# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting schemes and types

Setting schemes and deriving associated types, which delineate the target populations.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
schemes <- 
  read_schemes(lang = lang) %>% 
  filter(attribute_1 %in% c("GW", "SURF", "SOIL")) %>% 
  select(scheme, programme, attribute_1, attribute_2, attribute_3, tag_1, spatial_restriction)
```

Further restriction of schemes to the ones that are in scope for implementation in the first phase.

```{r}
core_schemes_1st_phase <- c(
  "GW_03.3",
  "SOIL_03.2",
  "SURF_03.4_lentic",
  "SURF_03.4_lotic"
)

schemes_1st_phase <- c(
  core_schemes_1st_phase,
  "GW_04.2",
  "GW_05.1_aq",
  "GW_05.1_terr", # don't take the whole target population here!! Only groups 2, 3, 5
  "GW_05.2",
  "GW_07.1",
  "SURF_064_lentic", # this one's existence is still to be confirmed
  "SURF_07.2"
)

schemes <- filter(schemes, scheme %in% schemes_1st_phase)
```

```{r restriction-1, eval=params$apply_restrictions, echo=params$apply_restrictions}
# ARTIFICIAL RESTRICTION 1
schemes <- 
  schemes %>% 
  filter(scheme %in% c(
    "GW_05.1_terr",
    "GW_07.1", 
    "SOIL_03.2", 
    "SURF_03.4_lentic", 
    "SURF_03.4_lotic"
  ))
```

Defining the target populations based on selected schemes.

```{r}
targetpops <- 
  read_scheme_types(lang = lang) %>% 
  select(scheme, type) %>% 
  semi_join(schemes, by = "scheme")
```

Further restriction of the target populations as decided for the first implementation phase.

<!-- This restricts target pops to 03.2/3/4, + adds 6430_mr for 05.1_terr (group 3 completion) & for 05.2 (to complete its target pop) -->

```{r}
targetpops_exclude_from_1st_phase <- 
  read_scheme_types(lang = lang) %>% 
  filter(
    scheme == "GW_05.1_terr",
    str_detect(typegroup, "GW_05.1_terr_group(1|4)")
  ) %>% 
  select(scheme, type)

targetpops <- 
  targetpops %>% 
  anti_join(targetpops_exclude_from_1st_phase, by = c("scheme", "type"))
```

```{r restriction-2, eval=params$apply_restrictions, echo=params$apply_restrictions}
# ARTIFICIAL RESTRICTION 2
type_shortlist <- c(
    "7220",
    "3260",
    "rbbah",
    "1330_hpr",
    "2190_a",
    "3130_aom",
    "3130_na",
    "2130_hd",
    "2310",
    "6430_bz",
    "7140_base",
    "9120",
    "7150"
  )
targetpops <- targetpops %>%
  filter(type %in% type_shortlist)
```

Extending `schemes` and `targetpops` with MHQ schemes.

```{r}
schemes_mhq <- 
  read_schemes(lang = lang) %>% 
  filter(programme == "MHQ") %>% 
  select(scheme, programme, attribute_1, attribute_2, attribute_3, tag_1, spatial_restriction)
schemes <- bind_rows(schemes, schemes_mhq)
targetpops <- 
  read_scheme_types(lang = lang) %>% 
  select(scheme, type) %>% 
  semi_join(schemes_mhq, by = "scheme") %>% 
  bind_rows(targetpops, .)
```

```{r restriction-3, eval=params$apply_restrictions, echo=params$apply_restrictions}
# RE-APPLY ARTIFICIAL RESTRICTION 2:
targetpops <- 
  targetpops %>%
  filter(type %in% type_shortlist)
schemes <- 
  schemes %>% 
  # dropping HQ-schemes not represented by type_shortlist:
  semi_join(targetpops, by = "scheme")
```

```{r}
targetpops
```

```{r}
schemes
```


## Setting modules and domains

### Modules

A module is like a project, that aims at realizing some well-defined part of MNE + MHQ.

- A module determines budget, cycle duration, end year and scheme-specific properties regarding sampling design, sample sizes and start year.
- In general, a module also determines which schemes and which part of the target populations are in scope.
  - We have defined `schemes` and `targetpops` globally, but the module will further determine whether the MHQ programme is included or not.
  - Another property of a module are the involved domains (see \@ref(domains-per-module)), and these geographical boundaries can in turn restrict `schemes` and `targetpops`.
- Also, we distinguish different components as different modules if it's reasonable to assume that funding or period may diverge in the future, even though they have the same characteristics below.

```{r}
modules <- tribble(
  ~module, ~cycle_duration_y, ~year_last, ~includes_mhq,
  "phase_1_flanders", 6, 2037, FALSE,
  "phase_1_sac_subset_5", 6, 2037, TRUE,
  "phase_1_be2100024", 3, 2030, TRUE
) %>% 
  mutate(
    module = fct(module),
    cycle_duration_y = as.integer(cycle_duration_y),
    year_last = as.integer(year_last)
  )
```


### Domains

Defining domains, including their geometry.

```{r}
sfg_flanders <- 
  read_admin_areas(dsn = "flanders") %>% 
  st_geometry() %>%
  .[[1]]
sac <- read_admin_areas(dsn = "sac")
sfg_sac_network <- st_combine(sac)[[1]]
sac_subset_5 <- c(
  "BE2100015", # Kalmthoutse Heide
  "BE2100024", # Vennen, heiden en moerassen rond Turnhout
  "BE2200028", # De Maten
  "BE2200035", # Mechelse heide en vallei van de Ziepbeek
  "BE2200039"  # Voerstreek
)
for (sac_i in sac_subset_5) {
  sac %>% 
  filter(sac_code == sac_i) %>% 
  st_combine() %>%
  .[[1]] %>% 
  assign(str_c("sfg_", tolower(sac_i)), ., envir = .GlobalEnv)
}
domains <- tribble(
  ~domain, ~domain_name_nl, ~domain_name_en, ~geometry,
  "Flanders", "Vlaams Gewest", "Flemish Region", sfg_flanders,
  "SAC_network", "SBZ-H-netwerk", "SAC network", sfg_sac_network,
  "BE2100015", "SBZ-H BE2100015", "SAC BE2100015", sfg_be2100015,
  "BE2100024", "SBZ-H BE2100024", "SAC BE2100024", sfg_be2100024,
  "BE2200028", "SBZ-H BE2200028", "SAC BE2200028", sfg_be2200028,
  "BE2200035", "SBZ-H BE2200035", "SAC BE2200035", sfg_be2200035,
  "BE2200039", "SBZ-H BE2200039", "SAC BE2200039", sfg_be2200039
) %>% 
  mutate(domain = fct(domain)) %>% 
  st_as_sf(crs = 31370)
```

### Domains per module

Each module is linked to one or more domains, hence the target populations (and consequently the schemes) will differ between modules.

At this level, we also define which domains get a specific sample size allocation for the target populations represented in the domain.

```{r}
module_domains <- tribble(
  ~module, ~domain, ~sample_size_predetermined,
  "phase_1_flanders", "Flanders", TRUE,
  "phase_1_flanders", "SAC_network", FALSE,
  "phase_1_sac_subset_5", "BE2100015", TRUE,
  "phase_1_sac_subset_5", "BE2100024", TRUE,
  "phase_1_sac_subset_5", "BE2200028", TRUE,
  "phase_1_sac_subset_5", "BE2200035", TRUE,
  "phase_1_sac_subset_5", "BE2200039", TRUE,
  "phase_1_be2100024", "BE2100024", TRUE
) %>% 
  mutate(
    module = factor(module, levels = levels(modules$module)),
    domain = factor(domain, levels = levels(domains$domain))
  )
```

**Where domains with predetermined sample size overlap and if the module(s) they belong to is/are active**, then the relevant sample parts can be stitched together with adjustment of sampling weights.
This means that overlapping domains give rise to:

- stratification -- if domain sizes differ -- of the larger domain into the domains from the other module + the remaining part;
- a split panel in case a different revisit design applies.

In order to not complicate terminology, we just denote all these geographical boundaries as 'domains' but we need to consider the stratification when generating samples and calculating sampling weights.


## Setting scheme properties

<!-- Only set the stable, module-agnostic scheme properties here. The module and domain specific scheme properties (for domains with predetermined sample size) can only be set completely once the spatial sampling frames are created for each scheme: it will also determine which types, hence schemes do apply in specific domains. -->

<!-- Revisit design and start year, depending on module. -->

**Missing functionality**: getting the primary and additional site characteristics and environmental variables for each scheme in MNE from existing datasets such as googlesheets.

Additional site characteristics / environmental variables might need spatial restrictions that are more stringent than the one at the scheme level.

For MHQ the variables will be summarized as one variable 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

For MNE we may want to define summarized sets of variables here, linked to a specific terrain method, and link _those_ to schemes.

<!-- Selectie van variabelen te voorzien. -->

## Setting data source versions



