# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting schemes and types

Setting schemes and deriving associated types, which delineate the target population.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
schemes <- 
  read_schemes(lang = lang) %>% 
  filter(attribute_1 %in% c("GW", "SURF", "SOIL")) %>% 
  select(scheme, programme, attribute_1, attribute_2, attribute_3, tag_1, spatial_restriction)
```

Further restriction of schemes to the ones that are in scope for implementation in the first phase.

```{r}
core_schemes_1st_phase <- c(
  "GW_03.3",
  "SOIL_03.2",
  "SURF_03.4_lentic",
  "SURF_03.4_lotic"
)

schemes_1st_phase <- c(
  core_schemes_1st_phase,
  "GW_04.2",
  "GW_05.1_aq",
  "GW_05.1_terr", # don't take the whole target population here!! Only groups 2, 3, 5
  "GW_05.2",
  "GW_07.1",
  "SURF_064_lentic", # this one's existence is still to be confirmed
  "SURF_07.2"
)

schemes <- filter(schemes, scheme %in% schemes_1st_phase)
```

```{r restriction-1}
# ARTIFICIAL RESTRICTION 1
schemes <- 
  schemes %>% 
  filter(scheme %in% c(
    "GW_05.1_terr",
    "GW_07.1", 
    "SOIL_03.2", 
    "SURF_03.4_lentic", 
    "SURF_03.4_lotic"
  ))
```

Defining the target population based on selected schemes.

```{r}
targetpop <- 
  read_scheme_types(lang = lang) %>% 
  select(scheme, type) %>% 
  semi_join(schemes, by = "scheme")
```

Further restriction of the target population as decided for the first implementation phase.

<!-- This restricts target pops to 03.2/3/4, + adds 6430_mr for 05.1_terr (group 3 completion) & for 05.2 (to complete its target pop) -->

```{r}
targetpop_exclude_from_1st_phase <- 
  read_scheme_types(lang = lang) %>% 
  filter(
    scheme == "GW_05.1_terr",
    str_detect(typegroup, "GW_05.1_terr_group(1|4)")
  ) %>% 
  select(scheme, type)

targetpop <- 
  targetpop %>% 
  anti_join(targetpop_exclude_from_1st_phase, by = c("scheme", "type"))
```

```{r restriction-2}
# ARTIFICIAL RESTRICTION 2
type_shortlist <- c(
    "7220",
    "3260",
    "rbbah",
    "1330_hpr",
    "2190_a",
    "3130_aom",
    "3130_na",
    "2130_hd",
    "2310",
    "6430_bz",
    "7140_base",
    "9120",
    "7150"
  )
targetpop <- targetpop %>%
  filter(type %in% type_shortlist)
```

Extending `schemes` and `targetpop` with MHQ schemes.

```{r}
schemes_mhq <- 
  read_schemes(lang = lang) %>% 
  filter(programme == "MHQ") %>% 
  select(scheme, programme, attribute_1, attribute_2, attribute_3, tag_1, spatial_restriction)
schemes <- bind_rows(schemes, schemes_mhq)
targetpop <- 
  read_scheme_types(lang = lang) %>% 
  select(scheme, type) %>% 
  semi_join(schemes_mhq, by = "scheme") %>% 
  bind_rows(targetpop, .)
```

```{r restriction-3}
# RE-APPLY ARTIFICIAL RESTRICTION 2:
targetpop <- 
  targetpop %>%
  filter(type %in% type_shortlist)
schemes <- 
  schemes %>% 
  # dropping HQ-schemes not represented by type_shortlist:
  semi_join(targetpop, by = "scheme")
```

```{r}
targetpop
```

```{r}
schemes
```


## Setting domains

Defining domains, including their geometry.

```{r}
sfg_flanders <- 
  read_admin_areas(dsn = "flanders") %>% 
  st_geometry() %>%
  .[[1]]
sac <- read_admin_areas(dsn = "sac")
sfg_sac_network <- st_combine(sac)[[1]]
sfg_sac_subset_5 <- 
  sac %>% 
  filter(sac_code %in% c(
    "BE2100015", # Kalmthoutse Heide
    "BE2100024", # Vennen, heiden en moerassen rond Turnhout
    "BE2200028", # De Maten
    "BE2200035", # Mechelse heide en vallei van de Ziepbeek
    "BE2200039"  # Voerstreek
  )) %>% 
  st_combine() %>%
  .[[1]]
sfg_be2100024 <- 
  sac %>% 
  filter(sac_code == "BE2100024") %>% 
  st_combine() %>%
  .[[1]]
domains <- tribble(
  ~domain, ~domain_name_nl, ~domain_name_en, ~geometry,
  "Flanders", "Vlaams Gewest", "Flemish Region", sfg_flanders,
  "SAC_network", "SBZ-H-netwerk", "SAC network", sfg_sac_network,
  "SAC_subset_5", "SBZ-H-subset van 5 PAS-maatwerkgebieden", "SAC subset of 5 protected areas", sfg_sac_subset_5,
  "BE2100024", "SBZ-H BE2100024", "SAC BE2100024", sfg_be2100024
) %>% 
  mutate(domain = factor(domain, levels = domain)) %>% 
  st_as_sf(crs = 31370)
```

The domains `SAC_subset_5` and `BE2100024` are more than just domains.
They also represent their own target populations for scheme implementations that aim at a higher precision than what would be obtained by subpopulation estimation from the Flemish schemes.
Hence, seen from the Flemish level, these represent a further level of stratification (crossed with type-based strata) as sample sizes are tailored to those domains.

In order to not complicate terminology, we just denote all these geographical boundaries as 'domains' but we need to consider the stratification when generating samples and calculating sampling weights.

<!-- Nog verder te verbeteren is domein SAC_subset_5, want het gaat om 5 domeinen, niet 1 domein. -->

## Setting scheme properties

**Missing functionality**: getting the primary and additional site characteristics and environmental variables for each scheme in MNE from existing datasets such as googlesheets.

Additional site characteristics / environmental variables might need spatial restrictions that are more stringent than the one at the scheme level.

For MHQ the variables will be summarized as one variable 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

For MNE we may want to define summarized sets of variables here, linked to a specific terrain method, and link _those_ to schemes.

<!-- Selectie van variabelen te voorzien. -->

## Setting data source versions



