# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting scheme sets and associated target populations (types)

Setting schemes and deriving associated types, which delineate the target populations.

The notion of a scheme being a 'core' scheme will be tied to modules (later).
In non-core schemes, only measuring the target variable is planned.
They are retained because of ± matching target population and variable with at least one core scheme in the same module.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
n2khab_schemes <- read_schemes(lang = lang)
pan_core_schemes_vec <- c(
  "GW_03.3",
  "SOIL_03.2",
  "SURF_03.4_lentic",
  "SURF_03.4_lotic"
) %>%
  factor(levels = levels(n2khab_schemes$scheme))
pan_transmon_schemes_vec <- c(
  as.character(pan_core_schemes_vec),
  "GW_04.2",
  "GW_05.1_aq",
  "GW_05.1_terr", # don't take the whole target population here!! Only groups 2 and 3
  "GW_05.2",
  "GW_07.1",
  "SURF_064_lentic", # this one's existence is still to be confirmed XXXXXXXXX
  "SURF_07.2"
)
pan_transmon_schemes <-
  n2khab_schemes %>%
  filter(scheme %in% pan_transmon_schemes_vec | programme == "MHQ") %>%
  mutate(
    is_core_scheme = scheme %in% pan_core_schemes_vec | 
      programme == "MHQ"
  ) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
# Since scheme SURF_07.2 has no habitat types, only RIBs, we should never list
# it as a member scheme for pan_effectmon modules:
pan_effectmon_schemes <- 
  pan_transmon_schemes %>% 
  filter(scheme != "SURF_07.2")
mbaa_mne_phase_1_schemes <-
  n2khab_schemes %>%
  filter(attribute_1 %in% c("GW", "SURF", "SOIL"), tag_1 == "focal") %>%
  mutate(is_core_scheme = TRUE) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
```

Defining a variant of `pan_effectmon_schemes` without the MHQ schemes.

```{r}
pan_effectmon_schemes_mne <-
  pan_effectmon_schemes %>%
  filter(programme == "MNE")
```

Preparing restriction of target populations for all 'pan' modules to that of schemes SOIL_03.2, GW_03.3 & SURF_03.4_*, including the addition of 6430_mr for GW_05.1_terr (typegroup 3 completion) & for GW_05.2 (to complete its target population).

```{r}
targetpops_exclude_from_pan <-
  read_scheme_types(lang = lang) %>%
  filter(
    scheme == "GW_05.1_terr",
    str_detect(typegroup, "GW_05.1_terr_group(1|4|5)")
  ) %>%
  select(scheme, type)
```

Note that 'completion' of target population takes into account that RIBs are not included in the target population of 'pan_effectmon' schemes by definition.

So we must also define these target populations in 'pan_effectmon' schemes as not considering RIBs.

Defining 'scheme + target population sets', i.e. different collections of schemes with associated target populations, that will be tied to modules.
A 'scheme + target population set' defines both the schemes and optional subsetting of types in the definition of associated target populations.
The 'core schemes' of a 'scheme + target population set' are the schemes with a custom sample size in a module.

```{r paged.print=FALSE}
stset_name_vec <- c(
  "pan_effectmon_schemes",
  "pan_effectmon_schemes_mne",
  "pan_transmon_schemes",
  "mbaa_mne_phase_1_schemes"
)
scheme_targetpop_sets <-
  tibble(
    stset_id = seq_along(stset_name_vec),
    stset_name = fct(stset_name_vec),
    scheme_set = map(stset_name_vec, ~ eval(as.name(.)))
  ) %>%
  mutate(
    targetpops = map2(scheme_set, stset_name, function(sts, stsn) {
      read_scheme_types(lang = lang) %>%
        select(scheme, type) %>%
        semi_join(sts, by = "scheme") %>%
        # restrict GW_05.1_terr target populations in 'pan' schemes:
        {
          if (str_detect(stsn, "pan_")) {
            anti_join(
              .,
              targetpops_exclude_from_pan,
              by = c("scheme", "type")
            )
          } else {
            .
          }
        } %>% 
        # restrict all 'pan_effectmon' schemes to habitat types:
        {
          if (str_detect(stsn, "pan_effectmon_")) {
            filter(., !str_detect(type, "^rbb"))
          } else {
            .
          }
        }
    })
  )
scheme_targetpop_sets
```

View of scheme_sets (for inspection):

```{r}
scheme_targetpop_sets %>%
  select(-targetpops) %>%
  unnest(cols = scheme_set)
```

Unioned set of schemes.

```{r}
schemes <-
  scheme_targetpop_sets$scheme_set %>%
  map(~ select(., -is_core_scheme)) %>%
  reduce(dplyr::union) %>%
  arrange(scheme)
```

Unioned set of target populations.

```{r}
targetpops <-
  scheme_targetpop_sets$targetpops %>%
  reduce(dplyr::union) %>%
  arrange(scheme, type)
targetpops
```


## Setting modules and domains

### Modules

A module is like a project, that aims at realizing some well-defined part of MNE + MHQ.

- A module determines budget, end year and scheme-specific properties regarding sampling design, sample sizes, cycle duration and start year.
- A module also determines which schemes and which part of the target populations are in scope.
  - Another property of a module are the involved domains (see \@ref(domains-per-module)), and these geographical boundaries can in turn restrict the applicable schemes and target populations in each domain.
  Since such restriction is not applied in `scheme_targetpop_sets`, this restriction is also not applied in the module definition below.
  The module-specific omission of schemes, based on domains, needs the scheme-specific sampling frames and will happen in chapter \@ref(scheme-objects).
- Also, we distinguish different components as different modules if it's reasonable to assume that funding or period may diverge in the future, even when they currently have the same characteristics.

```{r}
modules <- tribble(
  ~module, ~code_short, ~year_firstdefinition, ~year_last, ~stset_id,
  "pan_effectmon_flanders", "panfl", 2024, 2037, 2,
  "pan_effectmon_sac5_custommeas", "pan5", 2024, 2037, 1,
  "pan_transmon_be2100024", "pan2124", 2024, 2030, 3,
  "mbaa_mne_phase_1", "mbaa", 2024, 2031, 4,
  "map_mne", "map", 2024, NA_integer_, NA_integer_
) %>% 
  bind_cols(
    tibble(
      shortname_nl = c(
        "PAS-effectmonitoring in Vlaanderen", 
        "PAS-effectmonitoring in vijf maatwerk-SBZ-H’s",
        "Monitoring SBZ-H BE2100024 voor PAS-landbouwtransitie",
        "MBAG-MNM fase 1",
        "MAP-MNM"
      ),
      shortname_en = c(
        "PAN effect monitoring in Flanders",
        "PAN effect monitoring in five SACs with custom measures",
        "Monitoring SAC BE2100024 for PAN agricultural transition",
        "MBAA-MNE phase 1",
        "MAP-MNE"
      ),
      name_nl = c(
        "Programmatische Aanpak Stikstof (PAS) effectmonitoring in Vlaanderen", 
        "Programmatische Aanpak Stikstof (PAS) effectmonitoring in vijf Speciale Beschermingszones met maatwerk",
        "Monitoring van het natuurlijk milieu en de biotische kwaliteit in Speciale Beschermingszone BE2100024 (‘Vennen, heiden en moerassen rond Turnhout’) in het kader van landbouwtransitie voor de Programmatische Aanpak Stikstof (PAS)",
        "Meetnetten Natuurlijk Milieu in Vlaanderen (MNM) in het kader van het Meetnet Biodiversiteit Agrarisch Gebied, fase 1",
        "Meetnetten Natuurlijk Milieu in Vlaanderen (MNM) in het kader van het Mestactieplan (MAP)"
      ),
      name_en = c(
        "Programmatic Approach Nitrogen (PAN) effect monitoring in Flanders",
        "Programmatic Approach Nitrogen (PAN) effect monitoring in five Special Areas of Conservation with custom measures",
        "Monitoring of the natural environment and biotic quality in Special Area of Conservation BE2100024 (‘Vennen, heiden en moerassen rond Turnhout’) in the framework of agricultural transition for the Programmatic Approach Nitrogen (PAN)",
        "Monitoring Programme for the Natural Environment in Flanders (MNE) in the framework of Monitoring Biodiversity in Agricultural Areas (MBAA), phase 1",
        "Monitoring Programme for the Natural Environment in Flanders (MNE) in the framework of the Manure Action Plan (MAP)"
      )
    )
  ) %>%
  mutate(
    across(where(is.character), fct),
    across(where(is.numeric), as.integer)
  )
modules
```

**Missing functionality**: write this as a file, adding extra attributes like `date_last_updated`, `last_update_comment`, `date_cancelled`.

**Missing functionality (not for POC but for implementation)**: create a file to maintain module implementation data, e.g. with columns `module`, `module_definition_date` (i.e. referring to a `date_last_updated` above), `date_started`, `date_stopped`.

This workflow will restrict modules based on the parameter `active_modules`:

```{r}
params$active_modules
```

Applying the restriction by active modules.

```{r}
modules <-
  modules %>%
  filter(module %in% params$active_modules)
```

Flattening the schemes per module.
Note that below `module_schemes` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r}
module_schemes <- 
  modules %>% 
  select(module, stset_id) %>% 
  inner_join(
    scheme_targetpop_sets %>% 
      select(stset_id, scheme_set) %>% 
      unnest(scheme_set) %>% 
      select(stset_id, scheme, is_core_scheme),
    join_by(stset_id),
    unmatched = "drop",
    relationship = "many-to-many"
  ) %>% 
  select(-stset_id)
```

Flattening the target populations per module.
Note that below `module_targetpops` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r paged.print=FALSE}
module_targetpops <-
  modules %>%
  select(module, stset_id) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, targetpops),
    by = "stset_id"
  ) %>%
  select(-stset_id) %>%
  unnest(targetpops)
```


### Domains

Defining domains, including their geometry.

```{r}
sfg_flanders <-
  read_admin_areas(dsn = "flanders") %>%
  st_geometry() %>%
  .[[1]]
sac <- read_admin_areas(dsn = "sac")
sfg_sac_network <- st_combine(sac)[[1]]
sac_subset_5 <- c(
  "BE2100015", # Kalmthoutse Heide
  "BE2100024", # Vennen, heiden en moerassen rond Turnhout
  "BE2200028", # De Maten
  "BE2200035", # Mechelse heide en vallei van de Ziepbeek
  "BE2200039" # Voerstreek
)
for (sac_i in sac_subset_5) {
  sac %>%
    filter(sac_code == sac_i) %>%
    st_combine() %>%
    .[[1]] %>%
    assign(str_c("sfg_", tolower(sac_i)), ., envir = .GlobalEnv)
}
domains <- tribble(
  ~domain, ~domain_name_nl, ~domain_name_en, ~geometry,
  "Flanders", "Vlaams Gewest", "Flemish Region", sfg_flanders,
  "SAC_network", "SBZ-H-netwerk", "SAC network", sfg_sac_network,
  "BE2100015", "SBZ-H BE2100015", "SAC BE2100015", sfg_be2100015,
  "BE2100024", "SBZ-H BE2100024", "SAC BE2100024", sfg_be2100024,
  "BE2200028", "SBZ-H BE2200028", "SAC BE2200028", sfg_be2200028,
  "BE2200035", "SBZ-H BE2200035", "SAC BE2200035", sfg_be2200035,
  "BE2200039", "SBZ-H BE2200039", "SAC BE2200039", sfg_be2200039
) %>%
  mutate(across(where(is.character), fct)) %>%
  st_as_sf(crs = 31370)
domains
```

Defining spatial relations between domains.

```{r}
domain_overlaps <- tribble(
  ~domain, ~included_by,
  "BE2100015", "SAC_network",
  "BE2100024", "SAC_network",
  "BE2200028", "SAC_network",
  "BE2200035", "SAC_network",
  "BE2200039", "SAC_network",
  "BE2100015", "Flanders",
  "BE2100024", "Flanders",
  "BE2200028", "Flanders",
  "BE2200035", "Flanders",
  "BE2200039", "Flanders",
  "SAC_network", "Flanders"
) %>% 
  mutate(across(everything(), ~ factor(., levels = levels(domains$domain))))
```


### Domains per module

Each module is linked to one or more domains, hence the target populations (and consequently the schemes) will differ between modules.

At this level, we also define which domains get a specific sample size allocation for the target populations represented in the domain.

```{r}
module_domains <- tribble(
  ~module, ~domain, ~sample_size_predetermined,
  "pan_effectmon_flanders", "Flanders", TRUE,
  "pan_effectmon_flanders", "SAC_network", FALSE,
  "pan_effectmon_sac5_custommeas", "BE2100015", TRUE,
  "pan_effectmon_sac5_custommeas", "BE2100024", TRUE,
  "pan_effectmon_sac5_custommeas", "BE2200028", TRUE,
  "pan_effectmon_sac5_custommeas", "BE2200035", TRUE,
  "pan_effectmon_sac5_custommeas", "BE2200039", TRUE,
  "pan_transmon_be2100024", "BE2100024", TRUE,
  "mbaa_mne_phase_1", "Flanders", TRUE,
) %>%
  mutate(
    module = factor(module, levels = levels(modules$module)),
    domain = factor(domain, levels = levels(domains$domain))
  )
module_domains
```

Applying the restriction by active modules.

```{r}
module_domains <-
  module_domains %>%
  filter(module %in% params$active_modules)
```

**Where domains with predetermined sample size overlap and if the module(s) they belong to is/are active**, then the relevant sample parts can be stitched together with adjustment of sampling weights.
This means that overlapping domains give rise to:

- spatial poststratification -- if domain sizes differ -- of the larger domain into the domains from the other module + the remaining part;
- a split panel in case a different revisit design applies.

We need to consider the poststratification when generating spare units, in generating panels for revisit designs and in calculating sampling weights (chapter \@ref(sampling)).


## Setting the default proportion of spare units

A conservative approach is to assume that no local replacements are done but to always assume that the next (reverse) GRTS address is used.

Considering a conservative success rate of 1/4 (for many types it should be higher), we assume that 300% spare units is almost always sufficient:

```{r}
# considering a theoretical maximum of 20 consecutive replacements (each time because the previous replacement resulted in a non-match)
round(100 * (3/4)^(1:20))
sum(100 * (3/4)^(1:20)) %>% round()
```

```{r}
spare_coef_default <- 3L
```

We will however apply a much lower number for aquatic types, as these are mostly already known and monitored before.
When combining multiple revisit designs (multiple modules or panel splits), each needs its own spare units and this can cause large gaps in the GRTS address series.

```{r}
spare_coef_aq <- 0.2
```


## Setting cycle duration and implementation start

The remaining part of this chapter considers settings by module and scheme (active modules only), even while some may currently not differ between modules.

Note that schemes, hence scheme properties, can only be correctly linked to a module x domain combination (i.e. for domains with predetermined sample size) once the spatial sampling frames are created for each scheme: the sampling frames will also determine which types, hence schemes do apply in specific domains (because schemes with just a few rarer types will not apply to every domain).

Therefore settings by module and scheme will be filtered and rewritten based on `module_domain_schemes` after the latter has been created (chapter \@ref(scheme-objects)).

We set the variable `cycle_duration_y` per module x scheme, which indeed depends on the combination of module x scheme already.

```{r}
module_scheme_cycle_duration <-
  module_schemes %>%
  select(module, scheme) %>%
  mutate(cycle_duration_y = case_when(
    module == "pan_transmon_be2100024" ~ 3L,
    str_detect(scheme, "^SURF") ~ 3L,
    .default = 6L
  ))
```

Setting year and month of first implementation (this is further used in chapter \@ref(scheme-objects)).

```{r}
year_first_default <- 2025L
month_first_default <- 7L
```

## Defining activities and activity sequences

Activities are specific actions needed to collect data or to support the data collection.

Their definition and their planning, linked to the data collection needs, are essential in data collection plans.

Reading the list of activities and their attributes.
Note that only part of activities are field activities.

```{r read-gsheet-activities, eval = !file.exists(file.path(datapath, "text/raw/activities.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE, include=FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "activities", col_types = "ccllllc") %>%
  mutate(
    across(c(activity, activity_name), fct),
    protocol = factor(protocol)
  ) %>% 
  write_vc(
    file = "activities",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r read-activities}
activities <- 
  read_vc(
    file = "activities",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble()
```

An **activity sequence** is a set of activities that are tied together at location level for a specific scheme, with a specific chronological order.
Some activities in an activity sequence may have different panel designs, due to the nature of the activity or other requirements or restrictions, so the frequency of some activities may be higher but the overall chronology remains in place.

An **activity group** is a set of activities with the same panel design (panel count, period, group size and group repetition attributes) _and_ which take place at the same moment; for field activities this means during the same location visit.

The activity group to which an activity belongs depends on the applicable activity sequence.
This is due to the fact that different schemes require only _partly_ overlapping activities, hence a different activity sequence.

Also, different schemes may apply to the same location, hence different activity sequences may be applicable at one location.

Reading the list of activity sequences and the corresponding activities, chronology and activity groups.


```{r read-gsheet-activity-sequences, eval = !file.exists(file.path(datapath, "text/raw/activity_sequences.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE, include=FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "activity_sequences", col_types = "ciiclc") %>%
  mutate(
    activity_sequence = fct(activity_sequence),
    activity = factor(activity, levels = activities$activity),
    activity_group = fct_reorder(
      factor(activity_group),
      as.numeric(activity),
      first
    )
  ) %>%
  write_vc(
    file = "activity_sequences",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r read-activity-sequences}
activity_sequences <- 
  read_vc(
    file = "activity_sequences",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble()
```

## Setting environmental variables and associated (main) activities

### Environmental variables and main data collection methods

Reading and further amending the dataset that defines the environmental variables and their attributes, for each module and scheme.

Attributes include:

- variable set: a cluster of variables that are measured together (e.g. in the laboratory)
- main data collection method: it is an activity from above.

```{r read-gsheet-envvars, eval = !file.exists(file.path(datapath, "text/raw/vars.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE, include=FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "vars_per_scheme") %>%
  write_vc(
    file = "vars",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r construct-envvars}
envvars <- 
  read_vc(
    file = "vars",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble() %>%
  select(
    scheme = meetnet,
    target_site_factor = standplaatsfactor,
    environm_var = milieuvar,
    envvar_abbrev = afk_milieuvar,
    measurement_var = meetvar,
    target = primair,
    explanatory = verklarend,
    quality_control = kwaliteitscontrole,
    variable_set,
    main_datacollection_method,
    targetvar_temporal_targetpop:last_col()
  ) %>% 
  # warn about unknown schemes:
  {
    if (all(.$scheme %in% schemes$scheme)) {
      .
    } else {
      unknown <- unique(.$scheme[!(.$scheme %in% schemes$scheme)])
      warning(
        "Dropped unknown scheme(s): ", 
        paste(unknown, collapse = ", "),
        call. = FALSE
      )
      filter(., scheme %in% schemes$scheme)
    }
  } %>% 
  # check that data collection methods are known:
  {
    if (
      all(.$main_datacollection_method[!is.na(.$main_datacollection_method)] %in%
          activities$activity)
    ) {
      .
    } else {
      stop("Unknown data collection method(s) detected!", call. = FALSE)
    }
  } %>% 
  mutate(
    scheme = scheme %>% factor(levels = levels(n2khab_schemes$scheme)),
    main_datacollection_method = factor(
      main_datacollection_method, 
      levels = levels(activities$activity)
    ),
    across(target:quality_control, ~ !is.na(.)),
    across(where(is.character), factor),
    variable_set = factor(
      variable_set, 
      levels = c(levels(variable_set), "LSVI")
    )
  ) %>% 
  pivot_longer(
    cols = where(~ max(as.character(.), na.rm = TRUE) == "X"),
    names_to = "module",
    values_to = "keep"
  ) %>% 
  mutate(keep = !is.na(keep)) %>% 
  filter(keep) %>% 
  # warn about inactive or unknown modules:
  {
    if (all(.$module %in% modules$module)) {
      .
    } else {
      unknown <- unique(.$module[!(.$module %in% modules$module)])
      warning(
        "Dropped inactive or unknown module(s): ", 
        paste(unknown, collapse = ", "),
        call. = FALSE
      )
      filter(., module %in% modules$module)
    }
  } %>% 
  mutate(module = factor(module, levels = levels(modules$module))) %>% 
  relocate(module) %>% 
  select(-keep) %>% 
  # enforce consistency with the module_schemes object:
  semi_join(
    module_schemes,
    join_by(module, scheme)
  )
```

Note: explanatory environmental variables might need spatial restrictions that are more stringent than the one at the scheme level.

Checking which MNE schemes per module are not covered by `envvars` (this may yield cases where the applicable domain has no target population units available, since `module_schemes` will only take this into account in chapter \@ref(scheme-objects)).

```{r check-envvar-coverage}
module_schemes_mne <- 
  module_schemes %>% 
  filter(!str_detect(scheme, "^HQ"))
module_schemes_mne %>% 
  anti_join(
    envvars %>% 
      distinct(module, scheme),
    join_by(module, scheme)
  )
```

<!-- We currently see no unexpected cases. -->

Checking that no unplanned schemes are covered.

```{r check-envvar-coverage-2}
envvars %>% 
  distinct(module, scheme) %>% 
  anti_join(module_schemes_mne, join_by(module, scheme)) %>% 
  nrow() == 0
```

Some scheme x environmental variable combinations are module-specific:

```{r module-specific-envvars}
envvars %>% 
  distinct(
    module,
    scheme,
    environm_var,
    envvar_abbrev,
    variable_set,
    main_datacollection_method
  ) %>% 
  count(
    scheme,
    environm_var,
    envvar_abbrev,
    variable_set,
    main_datacollection_method,
    name = "n_modules"
  ) %>% 
  inner_join(
    envvars %>% 
      distinct(module, scheme) %>% 
      count(scheme, name = "n_modules_all"),
    join_by(scheme)
  ) %>% 
  filter(n_modules != n_modules_all)
```

Creating an object with _non-field_ data collection methods per module x scheme, which will add to later cost calculations per module.

```{r non-field-data-collection-methods}
mod_scheme_datacoll_nofield <- 
  envvars %>% 
  distinct(module, scheme, environm_var, envvar_abbrev, variable_set, main_datacollection_method) %>% 
  filter(
    !(main_datacollection_method %in% activities$activity[activities$is_field_activity]),
    !is.na(main_datacollection_method)
  ) %>% 
  arrange(module, scheme, variable_set, main_datacollection_method)
```

### Variable sets and associated main field methods

Defining variable sets and main field methods (i.e. main data collection methods restricted to field activities), per module x scheme.
We also add the variable set and field methods for the MHQ programme for modules where this applies.

```{r}
mod_scheme_variableset_mainfieldmethod <-
  # MNE schemes
  envvars %>%
  distinct(module, scheme, variable_set, main_datacollection_method) %>%
  # warn about missing data collection methods:
  {
    if (!any(is.na(.$main_datacollection_method))) {
      .
    } else {
      warning(
        "Dropped ",
        sum(is.na(.$main_datacollection_method)),
        " row(s) with missing data collection method.", 
        call. = FALSE
      )
      filter(., !is.na(main_datacollection_method))
    }
  } %>%
  # limit to field activities:
  filter(main_datacollection_method %in% activities$activity[activities$is_field_activity]) %>% 
  rename(main_field_method = main_datacollection_method) %>% 
  mutate(scheme = factor(scheme, levels = levels(n2khab_schemes$scheme))) %>%
  arrange(module, scheme, variable_set, main_field_method) %>% 
  # add variable set & field method for MHQ schemes:
  bind_rows(
    module_schemes %>% 
      filter(str_detect(scheme, "^HQ")) %>% 
      select(-is_core_scheme) %>% 
      inner_join(
        schemes %>%
          select(scheme, tag_1),
        join_by(scheme),
        unmatched = "drop",
        relationship = "many-to-one"
      ) %>%
      mutate(
        variable_set = "LSVI",
        main_field_method = ifelse(tag_1 == "aq", "LSVIAQ", "LSVITERR")
      ) %>% 
      select(-tag_1)
  ) %>%
  mutate(
    variable_set = factor(variable_set, levels = levels(envvars$variable_set)),
    main_field_method = factor(main_field_method, levels = levels(activities$activity))
  )
mod_scheme_variableset_mainfieldmethod
```

For MHQ the variables are summarized as one variable set 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

Checking that the set of module x schemes not covered by `mod_scheme_variableset_mainfieldmethod` equals the MNE-set not covered by `envvars`.

```{r check-mod-scheme-variableset-coverage}
stopifnot(
  module_schemes %>% 
    anti_join(
      mod_scheme_variableset_mainfieldmethod %>% 
        distinct(module, scheme),
      join_by(module, scheme)
    ) %>% 
    all.equal(
      module_schemes_mne %>% 
        anti_join(
          envvars %>% 
            distinct(module, scheme),
          join_by(module, scheme)
        )
    )
)
```

## Setting the panel design of field activities

### Determining the field activities of (spatial subsets of) schemes per module

'Field activity' can cover data collection, where it corresponds to 'main field method' from above, but it is broader and also comprises installation or maintenance of in-situ instruments.

The field activities are a subset of the activities:

```{r construct-field-activities}
field_activities <- 
  activities %>% 
  filter(is_field_activity) %>% 
  select(starts_with("activity")) %>% 
  rename_with(~ str_c("field_", .))
field_activities
```

Several field activities necessarily belong together, depending on the applicable sequence of activities, as expressed by the `activity_sequence` variable.

```{r construct-field-activity-sequences}
field_activity_sequences <- 
  activity_sequences %>% 
  semi_join(field_activities, join_by(activity == field_activity)) %>% 
  rename(
    field_activity = activity,
    field_activity_group = activity_group
  )
field_activity_sequences
```

Field activities can be linked to modules x schemes via the main field method by using the applicable activity sequence, and abstracting variable sets away.

Determining the main field methods per module x scheme, where useful split into spatial subsets (based on type properties in the target population) that will be linked to different activity sequences.
  
```{r}
mod_scheme_mainfieldmethod <- 
  mod_scheme_variableset_mainfieldmethod %>% 
  distinct(module, scheme, main_field_method) %>%
  # add the is_core_scheme variable, needed to filter activity_sequences later:
  inner_join(
    module_schemes,
    join_by(module, scheme),
    unmatched = c("error", "drop"),
    relationship = "many-to-one"
  ) %>% 
  # modify some main field methods for non-core schemes:
  mutate(
    main_field_method = case_when(
      scheme == "GW_05.1_aq" & 
        !is_core_scheme &  
        main_field_method == "SURFLEVREADDIVER" ~ "SURFLEVREADGAUGE",
      .default = main_field_method
    ) %>% 
      factor(levels = levels(activities$activity))
  )
mod_scheme_mainfieldmethod_subsets <- 
  mod_scheme_mainfieldmethod %>% 
  # Characterize all schemes with needed type features as needed to distinguish
  # between activity sequences, e.g. the distinction between aquatic and
  # terrestrial types. This gives rise to spatial subsets of module x scheme
  # based on type properties (if there's only one such subset, then the scheme
  # will not be split with regard to activity sequences).
  inner_join(
    module_targetpops %>% 
      inner_join(
        read_types() %>% 
          # duplicating types that have aquatic forms & terrestrial forms:
          filter(hydr_class == "HC23") %>% 
          uncount(2) %>% 
          mutate(in_aquatic_subset = rep(c(TRUE, FALSE), 2)) %>% 
          bind_rows(
            read_types() %>% 
              filter(hydr_class != "HC23") %>% 
              mutate(in_aquatic_subset = hydr_class == "HC3")
          ) %>% 
          select(type, in_aquatic_subset),
        join_by(type),
        unmatched = c("error", "drop"),
        relationship = "many-to-many"
      ) %>% 
      distinct(module, scheme, in_aquatic_subset) %>% 
      filter(
        !(scheme == "GW_05.1_terr" & in_aquatic_subset),
        !(scheme == "GW_05.1_aq" & !in_aquatic_subset),
        !(str_detect(scheme, "^SURF") & !in_aquatic_subset)
      ),
    join_by(module, scheme),
    unmatched = c("error", "drop"),
    relationship = "many-to-many"
  )
```

Expanding with activity sequences using the `field_activity_sequences` dataset and applying scheme, spatial subset, field method and `is_core_scheme` based filter rules.
  
```{r}
mod_scheme_mainfieldmethod_actseq <- 
  mod_scheme_mainfieldmethod_subsets %>% 
  inner_join(
    field_activity_sequences %>% 
      select(activity_sequence, field_activity),
    join_by(main_field_method == field_activity),
    unmatched = c("error", "drop"),
    relationship = "many-to-many"
  ) %>%
  apply_activity_sequence_filters()
```

Checking that each main field method per module, scheme and spatial subset was translated to exactly one sequence (this checks the `apply_activity_sequence_filters()` function).

```{r}
stopifnot(
  nrow(mod_scheme_mainfieldmethod_actseq) == 
    nrow(mod_scheme_mainfieldmethod_subsets)
)
```

Expanding the object with all field activities that apply for each combination of module, scheme and spatial subset.

```{r}
mod_scheme_field_activity <- 
  mod_scheme_mainfieldmethod_actseq %>% 
  # add missing field activities for each sequence
  nest(data = c(main_field_method)) %>% 
  mutate(data = map2(activity_sequence, data, function(s, df) {
    field_activity_sequences %>% 
      filter(activity_sequence == s) %>% 
      select(-activity_sequence)
  })) %>% 
  unnest(data) %>% 
  # don't keep incompatible combinations of spatial subsets and field
  # activities:
  filter(
    !(!in_aquatic_subset & field_activity == "LOCEVALAQ"), 
    !(in_aquatic_subset & field_activity == "LOCEVALTERR")
  ) %>% 
  # check that no incompatible situations remain:
  (
    function(df) {
      test <- df %>% 
        summarize(
          choice = sum(choose_one),
          .by = c(module, scheme, in_aquatic_subset, activity_sequence, rank)
        ) %>% 
        filter(choice > 1)
      if (nrow(test) > 0) {
        stop("Undecided activities detected; please choose one.")
      }
      df %>% select(-choose_one)
    }
  )
```

**Missing functionality**: some field activities will need more spatial subsets.
For example, within a scheme there may be differences by type for explanatory or quality-control variables: scheme sections are only defined for target environmental variables, not additional ones.
Another possible source are protocol variants that are scheme-specific _and_ have different costs.

### Field activity groups of (spatial subsets of) schemes per module

The 'field activity groups' (FAGs) are those activity groups that exist of field activities.
Hence, a FAG is a set of _field_ activities with the same panel design _and_ which take place during the same location visit.

We will construct FAGs per module, scheme and spatial subset by abstracting away the field activities and collapsing activity sequences, ranks and maximum ranks.
Also, the spatial subset is only given where relevant.

First, construct the FAGs while still retaining the activity sequence.

```{r}
mod_scheme_actseq_fag <- 
  mod_scheme_field_activity %>% 
  distinct(
    module,
    scheme,
    is_core_scheme,
    activity_sequence,
    field_activity_group,
    rank,
    max_rank,
    in_aquatic_subset
  ) %>% 
  # only consider spatial subsets where needed to distinguish between activity
  # groups:
  inner_join(
    mod_scheme_mainfieldmethod_subsets %>% 
      distinct(module, scheme, in_aquatic_subset) %>% 
      summarize(
        # is more than one subset covered by the module x scheme?
        multiple_subsets = n() > 1,
        .by = c(module, scheme)
      ),
    join_by(module, scheme),
    unmatched = "error",
    relationship = "many-to-one"
  ) %>% 
  mutate(
    # is more than one subset represented for the FAG?
    # This currently assumes that ALL subsets are hence represented for the FAG
    duplicate = n_distinct(in_aquatic_subset) > 1,
    in_aquatic_subset = ifelse(duplicate | !multiple_subsets, NA, in_aquatic_subset),
    .by = c(module, scheme, field_activity_group)
  ) %>% 
  # now spatial subsets are mere attributes of unique combination of
  # module, scheme, activity_sequence & field_activity_group:
  distinct(
    module,
    scheme,
    is_core_scheme,
    activity_sequence,
    field_activity_group,
    rank,
    max_rank,
    in_aquatic_subset
  )
```

Now we must verify that ranks will actually be meaningful after collapsing, by inspecting the cases that would be collapsed and filter for cases that have more than one rank number.
If this works when neglecting the module (useful for the revisit design), which we look at below, then of course it also holds when keeping the module.

```{r paged.print=FALSE}
scheme_fag_differentranks <- 
  mod_scheme_actseq_fag %>% 
  distinct(pick(-module)) %>% 
  mutate(
    n_actseqs = n_distinct(activity_sequence),
    has_different_ranks = n_distinct(rank) > 1,
    .by = c(scheme, in_aquatic_subset, field_activity_group)
  ) %>% 
  filter(n_actseqs > 1, has_different_ranks) %>% 
  arrange(scheme, in_aquatic_subset, field_activity_group, rank)
scheme_fag_differentranks
```

```{r}
stopifnot(nrow(scheme_fag_differentranks) == 0)
```

Next, construct the FAGs per module, scheme and spatial subset while collapsing activity sequences, ranks and maximum ranks.

```{r}
mod_scheme_fag <-
  mod_scheme_actseq_fag %>%
  summarize(
    rank = min(rank),
    max_rank = if (all(is.na(max_rank))) NA else max(max_rank, na.rm = TRUE),
    actseqs = str_c(unique(activity_sequence), collapse = ", ") %>% factor(),
    .by = c(
      module,
      scheme,
      is_core_scheme,
      field_activity_group,
      in_aquatic_subset
    )
  )
```

Overview of different existing combinations of `field_activity_group`, `scheme`, `is_core_scheme` and spatial subset:

```{r}
mod_scheme_fag %>%
  distinct(field_activity_group, scheme, is_core_scheme, in_aquatic_subset) %>%
  arrange(field_activity_group, scheme, is_core_scheme, in_aquatic_subset) 
```

We also construct an object with the FAG ranks at scheme x spatial subset level, hence after dropping module.
(But note we're actually not using it, it seems.)

```{r}
scheme_fag <- 
  mod_scheme_actseq_fag %>%
  summarize(
    rank = min(rank),
    max_rank = if (all(is.na(max_rank))) NA else max(max_rank, na.rm = TRUE),
    actseqs = str_c(activity_sequence, collapse = ", "),
    .by = c(
      scheme,
      field_activity_group,
      in_aquatic_subset
    )
  )
```

We already checked above that the ranks were constant at this level, so their interpretation remains straightforward.

#### Note on the role of activity sequences in a module x scheme

Note that quite some FAGs are present in more than one `activity_sequence`, within a module x scheme x spatial subset (below result), and this has been collapsed in `mod_scheme_fag`.

```{r}
mod_scheme_fag %>% 
  select(-c(actseqs, rank, max_rank)) %>% 
  inner_join(
    mod_scheme_actseq_fag %>% 
      mutate(
        n_actseq = n_distinct(activity_sequence),
        .by = c(module, scheme, is_core_scheme, field_activity_group, in_aquatic_subset)
      ) %>% 
      filter(n_actseq > 1),
    join_by(module, scheme, is_core_scheme, field_activity_group, in_aquatic_subset),
    relationship = "one-to-many",
    unmatched = c("drop", "error")
  ) %>% 
  select(-n_actseq)
```

The activity sequences serve two purposes:

- getting all field activities and FAGs for a given main field method of a module, scheme and spatial subset (see above in the creation of `mod_scheme_field_activity`);
- setting the required chronological order between field activities / FAGs.

By collapsing them in `mod_scheme_fag` we have made things more simple from now on, but note that we still need the requirements for the chronological order of FAGs, reflecting the (single or multiple) activity sequences they came from.

#### Note on duplication of field activities across schemes

Note that some field activities occur in multiple schemes.
Insofar their field activity groups get a different panel design (which may depend on module x scheme), care has to be taken to not unnecessarily duplicate some non-datacollection methods at different times.

Field activities and FAGs that occur more than once (considering all active modules together):

```{r}
mod_scheme_field_activity %>% 
  filter(is_core_scheme) %>% 
  count(field_activity, in_aquatic_subset) %>% 
  filter(n > 1)
```

```{r}
mod_scheme_fag %>% 
  count(field_activity_group, in_aquatic_subset) %>% 
  filter(n > 1)
```

#### Note on competing field activities

Something similar occurs when the same location gets two different but competing FAGs, one of which is superior to another, e.g. `GWINSTPIEZWELL` is superior to `GWINSTPIEZNODIVER`.
This will mostly be avoided by generating spatiotemporal samples for non-core schemes only for types (hence, locations) that are not covered by a competing core scheme (which would compete and win otherwise).

### Panel design of FAGs in (spatial subsets of) schemes per module

`mod_scheme_fag_paneldesign` is the panel design of field activity groups per module and scheme.
This will matter most for data collection plans.

```{r}
mod_scheme_fag_paneldesign <-
  mod_scheme_fag %>% 
  # Adding preliminary support for split panels; to be set as NA where
  # applicable. We call different splits of a panel design a 'panel_split' and
  # they are numbered within module x scheme x FAG x spatial subset (but I
  # actually expect only module x scheme to be relevant for the number of
  # distinguished panel_splits; later on we effectively consider this at scheme
  # level so that we get the same locations for all FAGs in a module x scheme)
  mutate(panel_split = 1L) %>% 
  inner_join(
    module_scheme_cycle_duration,
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  mutate(
    # revisit pattern of main periodicity
    revisit_pattern = case_when(
      str_detect(field_activity_group, "READDIVER|ADHOC") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    # main periodicity
    panel_period_main = case_when(
      # set some explicitly to the default of 1 year to avoid confusion:
      field_activity_group == "GWINSTPIEZNODIVER" ~ years(1),
      # SPATPOSITXXX: we will inflate the panel period with 3 months when making
      # the calendar! Not doing it here since that would disturb generating the
      # panel design.
      str_detect(scheme, "^SURF") & 
        field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE") ~ years(1),
      str_detect(
        field_activity_group, 
        "^GWINST|^SURFINST|SOIL.*SAMP|^SPATPOSITGAUGE"
      ) ~ months(3),
      str_detect(field_activity_group, "READDIVER") ~ months(6),
      # ADHOC FAGs are tricky: they CAN occur anytime, anywhere, but are 
      # rather rare!
      # One approach could be to artificially reduce the number of visits (e.g.
      # by sampling visits)
      str_detect(field_activity_group, "ADHOC") ~ months(1),
      field_activity_group == "SPATPOSITPIPE" & str_detect(scheme, "GW_05") ~ months(3),
      field_activity_group == "SPATPOSITPIPE" & is_core_scheme ~ months(3),
      .default = years(1)
    ),
    # overruling the cycle duration that was set by the module
    enforce_cycle_duration_y = case_when(
      str_detect(field_activity_group, "^GWINST") |
        field_activity_group %in% c("SURFINSTWELLDIVER", "SPATPOSITPIPE") ~ 6L,
      field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE") ~ 3L,
      str_detect(scheme, "^HQ") & 
        field_activity_group %in% c("LOCEVALAQ", "LSVIAQ") ~ 3L,
      .default = NA_integer_
    ),
    # effective cycle duration
    cycle_duration_y = ifelse(
      !is.na(enforce_cycle_duration_y),
      enforce_cycle_duration_y,
      cycle_duration_y
    ),
    # was cycle duration enforced?
    enforced_cycle_duration = !is.na(enforce_cycle_duration_y),
    # revisit pattern of subperiods, if applicable
    revisit_subpattern = case_when(
      str_detect(field_activity_group, "SURF.*DATACOLL") ~ "pure_panel",
      str_detect(scheme, "^SURF") & 
        field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE") ~ "pure_panel",
      str_detect(field_activity_group, "GW(SURF)?SHALL(CLEAN|SAMP)|LEVREAD") ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    # subperiodicity, if applicable
    panel_subperiod = case_when(
      str_detect(field_activity_group, "SURF.*DATACOLL") ~ months(1),
      str_detect(scheme, "^SURF") & 
        field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE") ~ months(1),
      str_detect(field_activity_group, "GW(SURF)?SHALL(CLEAN|SAMP)|LEVREAD") ~ months(3)
    ),
    # do measurements take place in an automated, continuous manner, which means
    # that data are collected more frequently at a location than the period
    # between two visits? This is relevant in the panel membership later on, as
    # these FAGs are not subject to the condition that consecutive panels must
    # be populated according to the GRTS series, allowing for more spatial
    # alignment with other FAG panel membership designs.
    auto_continuous_meas = str_detect(field_activity_group, "READDIVER"),
    # in certain FAGs, panel membership will not be aligned with that of other
    # revisit designs by following the GRTS series, but by aligning the
    # 'switching to a next panel' with the fastest switching revisit design with
    # which alignment is sought. When that fastest switching revisit design has
    # a longer cycle duration, this means that a single panel will be composed
    # by the locations of multiple, distinct panels from that revisit design.
    # This is the case for gauge installations in a groundwater scheme (aquatic
    # subset), where meaningful data collection will also depend on the
    # availability of piezometers or wells for the location, and the
    # installation of all these devices (SURFINSTGAUGE, GWINST) ideally is
    # aligned in the same date interval.
    memship_alignfastest = str_detect(scheme, "^GW") & 
        field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE"),
    # 'Enddate addition' does not affect the start date of the next panel, but 
    # only the end date of the current panel; this is used to make panels
    # overlap in time.
    panel_enddate_addition = case_when(
      str_detect(field_activity_group, "^SPATPOSIT") ~ months(3)
    ),
    # in certain FAGs, the whole panel design must be shifted by a fixed period
    # (negative shift = backward in time), relative to the default panel start
    # and end dates that would normally emerge
    panel_dateshift = case_when(
      # evaluate locations one year beforehand:
      str_detect(field_activity_group, "LOCEVAL") & 
        !str_detect(scheme, "^HQ") ~ years(-1),
      # postpone the start of groundwater sampling & diver reading (to align
      # location visits)
      str_detect(field_activity_group, "GW(SURF)?SHALL(CLEAN|SAMP)") ~ months(6),
      str_detect(field_activity_group, "READDIVER") ~ months(6)
    ),
    # optionally limit the number of subperiods in a main period; this causes
    # uneven coverage of the main period since only the first subperiod(s)
    # remain
    limit_n_subperiods = case_when(
      str_detect(scheme, "^SURF") & 
        field_activity_group %in% c("SURFINSTGAUGE", "SPATPOSITGAUGE") ~ 1L,
      .default = NA_integer_
    ),
    # optionally set a fixed number of subpanels in a main panel, which enforces
    # repetition of these panels in the main period in case of a serially
    # alternating subpattern
    limit_n_subpanels_per_panel = case_when(
      str_detect(field_activity_group, "GW(SURF)?SHALL(CLEAN|SAMP)") ~ 2L,
      .default = NA_integer_
    )
  ) %>% 
  select(-enforce_cycle_duration_y) %>% 
  relocate(cycle_duration_y, .before = enforced_cycle_duration) %>% 
  relocate(panel_split, .after = field_activity_group) %>% 
  arrange(field_activity_group)
```

**Still undecided**: should panel dateshift also apply to MHQ schemes?
(LOCEVAL; potentially LSVI.)
Only to be considered if this causes enough location overlap between panels in the same year.

**Missing functionality**: speed up of certain FAGs (such as INST) in just the first cycle if `enforce_cycle_duration_y` is `TRUE` (= FAG specific) while the cycle duration of this module is shorter.

Overview of different occurring panel designs per FAG across modules and schemes:

```{r}
mod_scheme_fag_paneldesign %>%
  arrange(scheme) %>% 
  summarize(
    scheme_first = first(scheme),
    scheme_last = last(scheme),
    n_schemes = n_distinct(scheme),
    .by = c(field_activity_group, panel_split, revisit_pattern:last_col())
  ) %>%
  relocate(contains("scheme")) %>% 
  arrange(field_activity_group)
```

## Panel design of measurements

`scheme_variableset_paneldesign` is the panel design of measurements.
It is created below to distinguish with `mod_scheme_fag_paneldesign` above, but yet unclear whether we'll need it to create data collection plans.
Its primary relevance is statistical (spatiotemporal pattern of available data), hence it underpins choices made above.

Also, it may be useful in cost calculations of data acquisition (next to field activity costs & other non-field datacollection costs – see before), since laboratory work is involved.

Notable differences with `mod_scheme_fag_paneldesign` are that:

- automated frequent measurements are reflected as 'frequent revisit' while fieldwork is much less frequent;
- currently no distinction between modules is supported;
- only core schemes are considered;
- no installation or maintenance is involved;
- the (assumed) temporal resolution of the target variable is added, which is an attempt to answer the question: which period does one value of the variable represent?

```{r}
scheme_variableset_paneldesign <-
  mod_scheme_variableset_mainfieldmethod %>%
  inner_join(
    module_schemes,
    join_by(module, scheme),
    unmatched = c("error", "drop"),
    relationship = "many-to-one"
  ) %>% 
  filter(is_core_scheme) %>% 
  distinct(scheme, variable_set, main_field_method) %>%
  mutate(
    targetvar_temporal_resolution = case_when(
      str_detect(scheme, "^HQ") ~ years(1),
      str_detect(scheme, "^GW_05") & str_detect(variable_set, "peil") ~ years(1),
      str_detect(variable_set, "peil.*automatisch") ~ days(1),
      str_detect(scheme, "^SURF_") ~ months(1),
      .default = months(3)
    ),
    revisit_pattern = case_when(
      str_detect(variable_set, "peil.*automatisch") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    panel_period_main = case_when(
      str_detect(variable_set, "peil.*automatisch") ~ days(1),
      str_detect(scheme, "^HQ|^SURF_|^GW_(03|04|07)") ~ years(1),
      .default = months(3)
    ),
    revisit_subpattern = case_when(
      str_detect(scheme, "^SURF_") ~ "pure_panel",
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    panel_subperiod = case_when(
      !is.na(revisit_subpattern) ~ targetvar_temporal_resolution
    ),
    limit_n_subpanels_per_panel = case_when(
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ 2L
    )
  )
scheme_variableset_paneldesign
```

## Setting data source versions for the construction of spatial sampling frames



## Overview of most relevant objects in this chapter {#overview-settings}

Objects subsetted with `[[1]]` mean that they (may) contain more, similarly structured elements.

Note that the following objects are not yet finished at this stage (see chapter \@ref(scheme-objects)):

- `scheme_targetpop_sets`
- `module_scheme_cycle_duration`
- `mod_scheme_datacoll_nofield`
- `mod_scheme_variableset_mainfieldmethod`
- `mod_scheme_mainfieldmethod`
- `mod_scheme_mainfieldmethod_subsets`
- `mod_scheme_mainfieldmethod_actseq`
- `mod_scheme_field_activity`
- `mod_scheme_actseq_fag`
- `mod_scheme_fag`
- `mod_scheme_fag_paneldesign`

```{r echo = FALSE}
for (name in c(
  "scheme_targetpop_sets",
  "scheme_targetpop_sets$scheme_set[[1]]",
  "scheme_targetpop_sets$targetpops[[1]]",
  "schemes",
  "targetpops",
  "modules",
  "domains",
  "module_domains",
  "module_schemes",
  "module_targetpops",
  "activities",
  "activity_sequences",
  "envvars",
  "field_activities",
  "field_activity_sequences",
  "module_scheme_cycle_duration",
  "mod_scheme_datacoll_nofield",
  "mod_scheme_variableset_mainfieldmethod",
  "mod_scheme_mainfieldmethod",
  "mod_scheme_mainfieldmethod_subsets",
  "mod_scheme_mainfieldmethod_actseq",
  "mod_scheme_field_activity",
  "mod_scheme_actseq_fag",
  "mod_scheme_fag",
  "mod_scheme_fag_paneldesign",
  "scheme_variableset_paneldesign"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
