# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting scheme sets and associated target populations (types)

Setting schemes and deriving associated types, which delineate the target populations.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
n2khab_schemes <- read_schemes(lang = lang)
pan_phase_1_core_schemes_vec <- c(
  "GW_03.3",
  "SOIL_03.2",
  "SURF_03.4_lentic",
  "SURF_03.4_lotic"
) %>%
  factor(levels = levels(n2khab_schemes$scheme))
pan_phase_1_schemes_vec <- c(
  as.character(pan_phase_1_core_schemes_vec),
  "GW_04.2",
  "GW_05.1_aq",
  "GW_05.1_terr", # don't take the whole target population here!! Only groups 2 and 3
  "GW_05.2",
  "GW_07.1",
  "SURF_064_lentic", # this one's existence is still to be confirmed XXXXXXXXX
  "SURF_07.2"
)
pan_phase_1_schemes <-
  n2khab_schemes %>%
  filter(scheme %in% pan_phase_1_schemes_vec | programme == "MHQ") %>%
  mutate(is_core_scheme = scheme %in% pan_phase_1_core_schemes_vec) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
mbaa_mne_phase_1_schemes <-
  n2khab_schemes %>%
  filter(attribute_1 %in% c("GW", "SURF", "SOIL"), tag_1 == "focal") %>%
  mutate(is_core_scheme = TRUE) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
```

Defining a variant of `pan_phase_1_schemes` without the MHQ schemes.

```{r}
pan_phase_1_schemes_mne <-
  pan_phase_1_schemes %>%
  filter(programme == "MNE")
```

Preparing restriction of target populations for pan_phase_1 to that of schemes SOIL_03.2, GW_03.3 & SURF_03.4_*, including the addition of 6430_mr for GW_05.1_terr (typegroup 3 completion) & for GW_05.2 (to complete its target population).

```{r}
targetpops_exclude_from_pan_phase_1 <-
  read_scheme_types(lang = lang) %>%
  filter(
    scheme == "GW_05.1_terr",
    str_detect(typegroup, "GW_05.1_terr_group(1|4|5)")
  ) %>%
  select(scheme, type)
```

Defining 'scheme + target population sets', i.e. different collections of schemes with associated target populations, that will be tied to modules.
A 'scheme + target population set' defines both the schemes and optional subsetting of types in the definition of associated target populations.
The 'core schemes' of a 'scheme + target population set' are the schemes with a custom sample size in a module.

```{r paged.print=FALSE}
stset_name_vec <- c(
  "pan_phase_1_schemes",
  "pan_phase_1_schemes_mne",
  "mbaa_mne_phase_1_schemes"
)
scheme_targetpop_sets <-
  tibble(
    stset_id = seq_along(stset_name_vec),
    stset_name = fct(stset_name_vec),
    scheme_set = map(stset_name_vec, ~ eval(as.name(.)))
  ) %>%
  mutate(
    targetpops = map2(scheme_set, stset_name, function(sts, stsn) {
      read_scheme_types(lang = lang) %>%
        select(scheme, type) %>%
        semi_join(sts, by = "scheme") %>%
        {
          if (!str_detect(stsn, "pan_phase_1")) {
            .
          } else {
            anti_join(
              .,
              targetpops_exclude_from_pan_phase_1,
              by = c("scheme", "type")
            )
          }
        }
    })
  )
scheme_targetpop_sets
```

View of scheme_sets (for inspection):

```{r}
scheme_targetpop_sets %>%
  select(-targetpops) %>%
  unnest(cols = scheme_set)
```

Unioned set of schemes.

```{r}
schemes <-
  scheme_targetpop_sets$scheme_set %>%
  map(~ select(., -is_core_scheme)) %>%
  reduce(dplyr::union) %>%
  arrange(scheme)
```

Unioned set of target populations.

```{r}
targetpops <-
  scheme_targetpop_sets$targetpops %>%
  reduce(dplyr::union) %>%
  arrange(scheme, type)
targetpops
```


## Setting modules and domains

### Modules

A module is like a project, that aims at realizing some well-defined part of MNE + MHQ.

- A module determines budget, end year and scheme-specific properties regarding sampling design, sample sizes, cycle duration and start year.
- A module also determines which schemes and which part of the target populations are in scope.
  - Another property of a module are the involved domains (see \@ref(domains-per-module)), and these geographical boundaries can in turn restrict the applicable schemes and target populations in each domain.
  Since such restriction is not applied in `scheme_targetpop_sets`, this restriction is also not applied in the module definition below.
  The module-specific omission of schemes, based on domains, needs the scheme-specific sampling frames and will happen in chapter \@ref(scheme-objects).
- Also, we distinguish different components as different modules if it's reasonable to assume that funding or period may diverge in the future, even when they currently have the same characteristics.

```{r}
modules <- tribble(
  ~module, ~year_last, ~stset_id,
  "pan_phase_1_flanders", 2037, 2,
  "pan_phase_1_sac_subset_5", 2037, 1,
  "pan_phase_1_be2100024", 2030, 1,
  "mbaa_mne_phase_1", 2031, 3
) %>%
  mutate(
    module = fct(module),
    across(where(is.numeric), as.integer)
  )
modules
```

This workflow can restrict modules based on the parameter `active_modules`:

```{r}
params$active_modules
```

Applying the restriction by active modules.

```{r}
modules <-
  modules %>%
  filter(module %in% params$active_modules)
```

Flattening the target populations per module.

```{r paged.print=FALSE}
module_targetpops <-
  modules %>%
  select(module, stset_id) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, targetpops),
    by = "stset_id"
  ) %>%
  select(-stset_id) %>%
  unnest(targetpops)
```


### Domains

Defining domains, including their geometry.

```{r}
sfg_flanders <-
  read_admin_areas(dsn = "flanders") %>%
  st_geometry() %>%
  .[[1]]
sac <- read_admin_areas(dsn = "sac")
sfg_sac_network <- st_combine(sac)[[1]]
sac_subset_5 <- c(
  "BE2100015", # Kalmthoutse Heide
  "BE2100024", # Vennen, heiden en moerassen rond Turnhout
  "BE2200028", # De Maten
  "BE2200035", # Mechelse heide en vallei van de Ziepbeek
  "BE2200039" # Voerstreek
)
for (sac_i in sac_subset_5) {
  sac %>%
    filter(sac_code == sac_i) %>%
    st_combine() %>%
    .[[1]] %>%
    assign(str_c("sfg_", tolower(sac_i)), ., envir = .GlobalEnv)
}
domains <- tribble(
  ~domain, ~domain_name_nl, ~domain_name_en, ~geometry,
  "Flanders", "Vlaams Gewest", "Flemish Region", sfg_flanders,
  "SAC_network", "SBZ-H-netwerk", "SAC network", sfg_sac_network,
  "BE2100015", "SBZ-H BE2100015", "SAC BE2100015", sfg_be2100015,
  "BE2100024", "SBZ-H BE2100024", "SAC BE2100024", sfg_be2100024,
  "BE2200028", "SBZ-H BE2200028", "SAC BE2200028", sfg_be2200028,
  "BE2200035", "SBZ-H BE2200035", "SAC BE2200035", sfg_be2200035,
  "BE2200039", "SBZ-H BE2200039", "SAC BE2200039", sfg_be2200039
) %>%
  mutate(domain = fct(domain)) %>%
  st_as_sf(crs = 31370)
domains
```

### Domains per module

Each module is linked to one or more domains, hence the target populations (and consequently the schemes) will differ between modules.

At this level, we also define which domains get a specific sample size allocation for the target populations represented in the domain.

```{r}
module_domains <- tribble(
  ~module, ~domain, ~sample_size_predetermined,
  "pan_phase_1_flanders", "Flanders", TRUE,
  "pan_phase_1_flanders", "SAC_network", FALSE,
  "pan_phase_1_sac_subset_5", "BE2100015", TRUE,
  "pan_phase_1_sac_subset_5", "BE2100024", TRUE,
  "pan_phase_1_sac_subset_5", "BE2200028", TRUE,
  "pan_phase_1_sac_subset_5", "BE2200035", TRUE,
  "pan_phase_1_sac_subset_5", "BE2200039", TRUE,
  "pan_phase_1_be2100024", "BE2100024", TRUE,
  "mbaa_mne_phase_1", "Flanders", TRUE,
) %>%
  mutate(
    module = factor(module, levels = levels(modules$module)),
    domain = factor(domain, levels = levels(domains$domain))
  )
module_domains
```

Applying the restriction by active modules.

```{r}
module_domains <-
  module_domains %>%
  filter(module %in% params$active_modules)
```

**Where domains with predetermined sample size overlap and if the module(s) they belong to is/are active**, then the relevant sample parts can be stitched together with adjustment of sampling weights.
This means that overlapping domains give rise to:

- stratification -- if domain sizes differ -- of the larger domain into the domains from the other module + the remaining part;
- a split panel in case a different revisit design applies.

In order to not complicate terminology, we just denote all these geographical boundaries as 'domains' but we need to consider the stratification when generating samples and calculating sampling weights.


## Setting scheme properties

**Missing functionality**: getting the primary and explanatory site factors and environmental variables for **remaining** schemes in MNE from existing datasets such as googlesheets.

Explanatory site factors / environmental variables might need spatial restrictions that are more stringent than the one at the scheme level.

Using what we have already:

<!-- currently the 'vars_phase_1_mbag' file is copied from the dcp_karen branch: -->

```{r}
envvars_mbaa <-
  read_vc(
    file = "vars_phase_1_mbag",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble() %>%
  select(
    scheme = meetnet,
    site_factor = standplaatsfactor,
    environm_var = milieuvar,
    envvar_abbrev = afk_milieuvar,
    measurement_var = meetvar,
    main = primair,
    explanatory = verklarend,
    quality_control = kwaliteitscontrole,
    variable_set,
    field_method
  ) %>%
  mutate(
    scheme = scheme %>% factor(levels = levels(n2khab_schemes$scheme)),
    main = !is.na(main),
    explanatory = !is.na(explanatory),
    quality_control = !is.na(quality_control)
  )
```

Defining variable sets and field methods per scheme.
This is module-agnostic.

```{r}
scheme_variableset_fieldmethod <-
  # MNE schemes phase 1
  envvars_mbaa %>%
  distinct(scheme, variable_set, field_method) %>%
  filter(
    !is.na(field_method),
    field_method != "te bepalen",
    variable_set != "te bepalen"
  ) %>%
  bind_rows(
    tribble(
      ~scheme, ~variable_set, ~field_method,
      "GW_04.2", "standaardpakket labo water anorganisch", "bemonsteren van ondiepe peilbuizen",
      "GW_05.2", "ondiep grondwaterpeil automatisch", "uitlezen diverloggers",
      "GW_07.1", "standaardpakket labo water anorganisch", "bemonsteren van ondiepe peilbuizen",
      "SURF_07.2", "standaardpakket labo water anorganisch", "bemonsteren waterkolom stilstaande wateren"
    )
  ) %>%
  mutate(scheme = factor(scheme, levels = levels(n2khab_schemes$scheme))) %>%
  # MHQ schemes
  bind_rows(
    schemes %>%
      filter(programme == "MHQ") %>%
      select(scheme) %>%
      mutate(
        variable_set = "LSVI",
        field_method = "LSVI-beoordeling"
      )
  ) %>%
  mutate(across(
    c(variable_set, field_method),
    ~ factor(.) %>% fct_relevel(~ .[!str_detect(., "LSVI")])
  ))
scheme_variableset_fieldmethod
```

For MHQ the variables are summarized as one variable 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

Checking that all schemes are covered by `scheme_variableset_fieldmethod`.

```{r}
schemes %>%
  anti_join(scheme_variableset_fieldmethod, by = "scheme") %>%
  nrow() == 0
```

'Field activity' can cover measurement, where it corresponds to 'field method' from above, but it is broader and also comprises installation or maintenance of in-situ instruments.

```{r}
field_activities <-
  tribble(
    ~field_activity, ~field_activity_name,
    "GWLEVINST", "installatie of onderhoud peilbuizen",
    "GWLEVREAD", "uitlezen diverloggers",
    "GWSHALLSAMP", "bemonsteren van ondiepe peilbuizen",
    "SOILUPPERSAMP", "bemonsteren bodem via oppervlaktemonsters",
    "SURFLEVINST", "installatie of onderhoud peillatten",
    "SURFLEVREAD", "aflezen peillat",
    "SURFLENTSAMP", "bemonsteren waterkolom stilstaande wateren",
    "SURFLOTSAMP", "bemonsteren waterkolom waterlopen",
    "SURFSECC", "bepaling Secchi-diepte",
    "SURFLENTFIELD", "veldmeting stilstaande wateren",
    "SURFLOTFIELD", "veldmeting waterlopen",
    "LSVI", "LSVI-beoordeling"
  ) %>%
  mutate(
    field_activity = fct(field_activity),
    field_activity_name = fct(field_activity_name)
  )
```

Setting other _**so-far**_ module-agnostic scheme design attributes.
'So far' because these are not guaranteed to apply to every future module.

- Field activities are scheme specific.

  **Missing functionality**: consider making field activities type-specific (i.e. join with the target population), since within a scheme there may be differences by type for explanatory or quality-control variables: scheme sections are only defined for main environmental variables, not additional ones.
  
  A special type-specific field activity is the evaluation of a location: does it belong to type X of the target population?
  Type specificity in this case will always invoke splitting of the field activity, since the type specificity concerns the panel design properties; especially the timing during the year depends on the type.
  The revisit design of these field activities will not be scheme specific, but given the dependency on the target population, different schemes will have different location evaluation field activities.

```{r}
scheme_field_activity <-
  scheme_variableset_fieldmethod %>%
  distinct(scheme, field_method) %>%
  rename(field_activity_name = field_method) %>%
  bind_rows(
    tibble(
      scheme = c(
        "GW_03.3",
        "GW_05.1_terr",
        "GW_05.1_aq",
        "GW_05.1_quarries",
        "GW_04.2",
        "GW_05.2",
        "GW_07.1"
      ),
      field_activity_name = "installatie of onderhoud peilbuizen"
    )
  ) %>%
  bind_rows(
    tibble(
      scheme = "GW_05.1_aq",
      field_activity_name = "installatie of onderhoud peillatten"
    )
  ) %>%
  mutate(
    scheme = factor(scheme, levels = levels(schemes$scheme)),
    field_activity_name = factor(
      field_activity_name,
      levels = levels(field_activities$field_activity_name)
    )
  ) %>%
  inner_join(
    field_activities,
    join_by(field_activity_name),
    unmatched = "error"
  ) %>%
  select(-field_activity_name) %>%
  arrange(scheme, field_activity)
```

- `field_activity_paneldesign` is the revisit design of field activities.
This will matter most for data collection plans.
Note that currently we don't handle this scheme specific, but that might change in future.

  **Missing functionality**: make the panel designs module and scheme specific.

```{r}
field_activity_paneldesign <-
  field_activities %>%
  mutate(
    revisit_pattern = case_when(
      str_detect(field_activity_name, "(af|uit)lezen") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    panel_period_main = case_when(
      str_detect(field_activity_name, "^LSVI|water|Secchi") ~ years(1),
      str_detect(field_activity_name, "bemonsteren.*peilbuizen") ~ years(1),
      str_detect(
        field_activity_name,
        "(af|uit)lezen"
      ) ~ months(6),
      .default = months(3)
    ),
    enforce_cycle_duration_y = case_when(
      str_detect(field_activity_name, "installatie.*onderhoud.*") ~ 6L
    ),
    revisit_subpattern = case_when(
      str_detect(field_activity_name, "water|Secchi") ~ "pure_panel",
      str_detect(
        field_activity_name,
        "bemonsteren.*peilbuizen|(af|uit)lezen"
      ) ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    panel_subperiod = case_when(
      str_detect(field_activity_name, "water|Secchi") ~ months(1),
      str_detect(
        field_activity_name,
        "bemonsteren.*peilbuizen|(af|uit)lezen"
      ) ~ months(3)
    ),
    limit_n_subpanels_per_panel = case_when(
      str_detect(field_activity_name, "bemonsteren.*peilbuizen") ~ 2L
    )
  ) %>%
  select(-field_activity_name)
field_activity_paneldesign
```

- `scheme_variableset_paneldesign` is the revisit design of measurements.
It is created below to distinguish with `field_activity_paneldesign` above, but yet unclear whether we'll need it to create data collection plans.
Its primary relevance is statistical (spatiotemporal pattern of available data), hence it underpins choices made above.
Notable differences with `field_activity_paneldesign` are that:
  - automated frequent measurements are reflected as 'frequent revisit' while fieldwork is much less frequent;
  - no installation or maintenance is involved;
  - the (assumed) temporal resolution of the target variable is added, which is an attempt to answer the question: which period does one value of the variable represent?

```{r}
scheme_variableset_paneldesign <-
  scheme_variableset_fieldmethod %>%
  distinct(scheme, variable_set) %>%
  mutate(
    targetvar_temporal_resolution = case_when(
      str_detect(scheme, "^HQ") | str_detect(variable_set, "peil") ~ years(1),
      str_detect(scheme, "^SURF_") ~ months(1),
      .default = months(3)
    ),
    revisit_pattern = case_when(
      str_detect(variable_set, "peil") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    panel_period_main = case_when(
      str_detect(variable_set, "peil") ~ days(1),
      str_detect(scheme, "^HQ|^SURF_|^GW_(03|04|07)") ~ years(1),
      .default = months(3)
    ),
    revisit_subpattern = case_when(
      str_detect(scheme, "^SURF_") ~ "pure_panel",
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    panel_subperiod = case_when(
      !is.na(revisit_subpattern) ~ targetvar_temporal_resolution
    ),
    limit_n_subpanels_per_panel = case_when(
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ 2L
    )
  )
scheme_variableset_paneldesign
```

We set the variable `cycle_duration_y` per module x scheme, which indeed depends on the combination of module x scheme already.

Note that schemes, hence scheme properties, can only be correctly linked to a module x domain combination (i.e. for domains with predetermined sample size) once the spatial sampling frames are created for each scheme: the sampling frames will also determine which types, hence schemes do apply in specific domains (because schemes with just a few rarer types will not apply to every domain).
So the below `module_scheme_cycle_duration` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r}
module_scheme_cycle_duration <-
  module_targetpops %>%
  distinct(module, scheme) %>%
  mutate(cycle_duration_y = case_when(
    module == "pan_phase_1_be2100024" ~ 3L,
    str_detect(scheme, "^SURF") ~ 3L,
    .default = 6L
  ))
```

Setting year and month of first implementation (this is further used in chapter \@ref(scheme-objects)).

```{r}
year_first_default <- 2025L
month_first_default <- 7L
```

## Setting data source versions



## Overview of most relevant objects in this chapter

Objects subsetted with `[[1]]` mean that they (may) contain more, similarly structured elements.

Note that the following objects are not yet finished at this stage (see chapter \@ref(scheme-objects)):

- `module_scheme_cycle_duration`
- `scheme_targetpop_sets`

```{r echo = FALSE}
for (name in c(
  "scheme_targetpop_sets",
  "scheme_targetpop_sets$scheme_set[[1]]",
  "scheme_targetpop_sets$targetpops[[1]]",
  "schemes",
  "targetpops",
  "modules",
  "domains",
  "module_domains",
  "scheme_variableset_fieldmethod",
  "scheme_field_activity",
  "field_activity_paneldesign",
  "scheme_variableset_paneldesign",
  "module_scheme_cycle_duration"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
