# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting scheme sets and associated target populations (types)

Setting schemes and deriving associated types, which delineate the target populations.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
n2khab_schemes <- read_schemes(lang = lang)
mne2024_pan_effectmon_core_schemes_vec <- c(
  "GW_03.3",
  "SOIL_03.2",
  "SURF_03.4_lentic",
  "SURF_03.4_lotic"
) %>%
  factor(levels = levels(n2khab_schemes$scheme))
mne2024_pan_schemes_vec <- c(
  as.character(mne2024_pan_effectmon_core_schemes_vec),
  "GW_04.2",
  "GW_05.1_aq",
  "GW_05.1_terr", # don't take the whole target population here!! Only groups 2 and 3
  "GW_05.2",
  "GW_07.1",
  "SURF_064_lentic", # this one's existence is still to be confirmed XXXXXXXXX
  "SURF_07.2"
)
mne2024_pan_schemes <-
  n2khab_schemes %>%
  filter(scheme %in% mne2024_pan_schemes_vec | programme == "MHQ") %>%
  mutate(is_core_scheme = scheme %in% mne2024_pan_effectmon_core_schemes_vec) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
mne2024_mbaa_mne_phase_1_schemes <-
  n2khab_schemes %>%
  filter(attribute_1 %in% c("GW", "SURF", "SOIL"), tag_1 == "focal") %>%
  mutate(is_core_scheme = TRUE) %>%
  select(
    scheme,
    is_core_scheme,
    programme,
    attribute_1,
    attribute_2,
    attribute_3,
    tag_1,
    spatial_restriction
  )
```

Defining a variant of `mne2024_pan_schemes` without the MHQ schemes.

```{r}
mne2024_pan_schemes_mne <-
  mne2024_pan_schemes %>%
  filter(programme == "MNE")
```

Preparing restriction of target populations for mne2024_pan to that of schemes SOIL_03.2, GW_03.3 & SURF_03.4_*, including the addition of 6430_mr for GW_05.1_terr (typegroup 3 completion) & for GW_05.2 (to complete its target population).

```{r}
targetpops_exclude_from_mne2024_pan <-
  read_scheme_types(lang = lang) %>%
  filter(
    scheme == "GW_05.1_terr",
    str_detect(typegroup, "GW_05.1_terr_group(1|4|5)")
  ) %>%
  select(scheme, type)
```

Defining 'scheme + target population sets', i.e. different collections of schemes with associated target populations, that will be tied to modules.
A 'scheme + target population set' defines both the schemes and optional subsetting of types in the definition of associated target populations.
The 'core schemes' of a 'scheme + target population set' are the schemes with a custom sample size in a module.

```{r paged.print=FALSE}
stset_name_vec <- c(
  "mne2024_pan_schemes",
  "mne2024_pan_schemes_mne",
  "mne2024_mbaa_mne_phase_1_schemes"
)
scheme_targetpop_sets <-
  tibble(
    stset_id = seq_along(stset_name_vec),
    stset_name = fct(stset_name_vec),
    scheme_set = map(stset_name_vec, ~ eval(as.name(.)))
  ) %>%
  mutate(
    targetpops = map2(scheme_set, stset_name, function(sts, stsn) {
      read_scheme_types(lang = lang) %>%
        select(scheme, type) %>%
        semi_join(sts, by = "scheme") %>%
        {
          if (!str_detect(stsn, "mne2024_pan")) {
            .
          } else {
            anti_join(
              .,
              targetpops_exclude_from_mne2024_pan,
              by = c("scheme", "type")
            )
          }
        }
    })
  )
scheme_targetpop_sets
```

View of scheme_sets (for inspection):

```{r}
scheme_targetpop_sets %>%
  select(-targetpops) %>%
  unnest(cols = scheme_set)
```

Unioned set of schemes.

```{r}
schemes <-
  scheme_targetpop_sets$scheme_set %>%
  map(~ select(., -is_core_scheme)) %>%
  reduce(dplyr::union) %>%
  arrange(scheme)
```

Unioned set of target populations.

```{r}
targetpops <-
  scheme_targetpop_sets$targetpops %>%
  reduce(dplyr::union) %>%
  arrange(scheme, type)
targetpops
```


## Setting modules and domains

### Modules

A module is like a project, that aims at realizing some well-defined part of MNE + MHQ.

- A module determines budget, end year and scheme-specific properties regarding sampling design, sample sizes, cycle duration and start year.
- A module also determines which schemes and which part of the target populations are in scope.
  - Another property of a module are the involved domains (see \@ref(domains-per-module)), and these geographical boundaries can in turn restrict the applicable schemes and target populations in each domain.
  Since such restriction is not applied in `scheme_targetpop_sets`, this restriction is also not applied in the module definition below.
  The module-specific omission of schemes, based on domains, needs the scheme-specific sampling frames and will happen in chapter \@ref(scheme-objects).
- Also, we distinguish different components as different modules if it's reasonable to assume that funding or period may diverge in the future, even when they currently have the same characteristics.

```{r}
modules <- tribble(
  ~module, ~year_firstdefinition, ~year_last, ~stset_id,
  "mne2024_pan_effectmon_flanders", 2024, 2037, 2,
  "mne2024_pan_effectmon_sac_subset_5", 2024, 2037, 1,
  "mne2024_pan_be2100024_restoration", 2024, 2030, 1,
  "mne2024_mbaa_mne_phase_1", 2024, 2031, 3,
  "mne2024_map", 2024, NA_integer_, NA_integer_
) %>% 
  bind_cols(
    tibble(
      module_name_nl = c(
        "PAS effectmonitoring in Vlaanderen", 
        "PAS effectmonitoring in maatwerkgebieden", 
        "Habitatmonitoring BE2100024 voor flankerend PAS-beleid",
        "MBAG-MNM fase 1",
        "MAP monitoring natuurlijk milieu"
      ),
      module_name_en = c(
        "PAN effect monitoring in Flanders",
        "PAN effect monitoring in areas with specific measures",
        "Habitat monitoring BE2100024 for PAN facilitation policy",
        "MBAA-MNE phase 1",
        "MAP natural environment monitoring"
      )
    )
  ) %>%
  mutate(
    across(where(is.character), fct),
    across(where(is.numeric), as.integer)
  )
modules
```

This workflow will restrict modules based on the parameter `active_modules`:

```{r}
params$active_modules
```

Applying the restriction by active modules.

```{r}
modules <-
  modules %>%
  filter(module %in% params$active_modules)
```

Flattening the schemes per module.
Note that below `module_schemes` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r}
module_schemes <- 
  modules %>% 
  select(module, stset_id) %>% 
  inner_join(
    scheme_targetpop_sets %>% 
      select(stset_id, scheme_set) %>% 
      unnest(scheme_set) %>% 
      select(stset_id, scheme, is_core_scheme),
    join_by(stset_id),
    unmatched = "drop",
    relationship = "many-to-many"
  ) %>% 
  select(-stset_id)
```

Flattening the target populations per module.
Note that below `module_targetpops` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r paged.print=FALSE}
module_targetpops <-
  modules %>%
  select(module, stset_id) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, targetpops),
    by = "stset_id"
  ) %>%
  select(-stset_id) %>%
  unnest(targetpops)
```


### Domains

Defining domains, including their geometry.

```{r}
sfg_flanders <-
  read_admin_areas(dsn = "flanders") %>%
  st_geometry() %>%
  .[[1]]
sac <- read_admin_areas(dsn = "sac")
sfg_sac_network <- st_combine(sac)[[1]]
sac_subset_5 <- c(
  "BE2100015", # Kalmthoutse Heide
  "BE2100024", # Vennen, heiden en moerassen rond Turnhout
  "BE2200028", # De Maten
  "BE2200035", # Mechelse heide en vallei van de Ziepbeek
  "BE2200039" # Voerstreek
)
for (sac_i in sac_subset_5) {
  sac %>%
    filter(sac_code == sac_i) %>%
    st_combine() %>%
    .[[1]] %>%
    assign(str_c("sfg_", tolower(sac_i)), ., envir = .GlobalEnv)
}
domains <- tribble(
  ~domain, ~domain_name_nl, ~domain_name_en, ~geometry,
  "Flanders", "Vlaams Gewest", "Flemish Region", sfg_flanders,
  "SAC_network", "SBZ-H-netwerk", "SAC network", sfg_sac_network,
  "BE2100015", "SBZ-H BE2100015", "SAC BE2100015", sfg_be2100015,
  "BE2100024", "SBZ-H BE2100024", "SAC BE2100024", sfg_be2100024,
  "BE2200028", "SBZ-H BE2200028", "SAC BE2200028", sfg_be2200028,
  "BE2200035", "SBZ-H BE2200035", "SAC BE2200035", sfg_be2200035,
  "BE2200039", "SBZ-H BE2200039", "SAC BE2200039", sfg_be2200039
) %>%
  mutate(across(where(is.character), fct)) %>%
  st_as_sf(crs = 31370)
domains
```

Defining spatial relations between domains.

```{r}
domain_overlaps <- tribble(
  ~domain, ~included_by,
  "BE2100015", "SAC_network",
  "BE2100024", "SAC_network",
  "BE2200028", "SAC_network",
  "BE2200035", "SAC_network",
  "BE2200039", "SAC_network",
  "BE2100015", "Flanders",
  "BE2100024", "Flanders",
  "BE2200028", "Flanders",
  "BE2200035", "Flanders",
  "BE2200039", "Flanders",
  "SAC_network", "Flanders"
) %>% 
  mutate(across(everything(), ~ factor(., levels = levels(domains$domain))))
```


### Domains per module

Each module is linked to one or more domains, hence the target populations (and consequently the schemes) will differ between modules.

At this level, we also define which domains get a specific sample size allocation for the target populations represented in the domain.

```{r}
module_domains <- tribble(
  ~module, ~domain, ~sample_size_predetermined,
  "mne2024_pan_effectmon_flanders", "Flanders", TRUE,
  "mne2024_pan_effectmon_flanders", "SAC_network", FALSE,
  "mne2024_pan_effectmon_sac_subset_5", "BE2100015", TRUE,
  "mne2024_pan_effectmon_sac_subset_5", "BE2100024", TRUE,
  "mne2024_pan_effectmon_sac_subset_5", "BE2200028", TRUE,
  "mne2024_pan_effectmon_sac_subset_5", "BE2200035", TRUE,
  "mne2024_pan_effectmon_sac_subset_5", "BE2200039", TRUE,
  "mne2024_pan_be2100024_restoration", "BE2100024", TRUE,
  "mne2024_mbaa_mne_phase_1", "Flanders", TRUE,
) %>%
  mutate(
    module = factor(module, levels = levels(modules$module)),
    domain = factor(domain, levels = levels(domains$domain))
  )
module_domains
```

Applying the restriction by active modules.

```{r}
module_domains <-
  module_domains %>%
  filter(module %in% params$active_modules)
```

**Where domains with predetermined sample size overlap and if the module(s) they belong to is/are active**, then the relevant sample parts can be stitched together with adjustment of sampling weights.
This means that overlapping domains give rise to:

- stratification -- if domain sizes differ -- of the larger domain into the domains from the other module + the remaining part;
- a split panel in case a different revisit design applies.

In order to not complicate terminology, we just denote all these geographical boundaries as 'domains' but we need to consider the stratification when generating samples and calculating sampling weights.


## Setting scheme properties

Note: explanatory environmental variables might need spatial restrictions that are more stringent than the one at the scheme level.

```{r read-gsheet-activities, eval = !file.exists(file.path(datapath, "text/raw/activities.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "activities", col_types = "ccllllc") %>%
  mutate(
    across(c(activity, activity_name), fct),
    protocol = factor(protocol)
  ) %>% 
  write_vc(
    file = "activities",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r}
activities <- 
  read_vc(
    file = "activities",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble()
```

```{r read-gsheet-activity-sequences, eval = !file.exists(file.path(datapath, "text/raw/activity_sequences.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "activity_sequences", col_types = "ciiclc") %>%
  mutate(
    activity_sequence = fct(activity_sequence),
    activity = factor(activity, levels = activities$activity),
    activity_group = fct_reorder(
      factor(activity_group),
      as.numeric(activity),
      first
    )
  ) %>%
  write_vc(
    file = "activity_sequences",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r read-gsheet-envvars, eval = !file.exists(file.path(datapath, "text/raw/vars.yml")) || params$update_raw_textdata, message = FALSE, warning = FALSE}
gs_id_envvars <- "1gJb2nY-Cs-SCNMpz0sGFslb0j69mzNx5BJu9tQXpHfc"
read_sheet(ss = gs_id_envvars, sheet = "vars_per_scheme") %>%
  write_vc(
    file = "vars",
    root = file.path(datapath, "text/raw"),
    optimize = FALSE,
    strict = FALSE
  )
```

```{r}
activity_sequences <- 
  read_vc(
    file = "activity_sequences",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble()
```

```{r}
envvars <- 
  read_vc(
    file = "vars",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble() %>%
  select(
    scheme = meetnet,
    target_site_factor = standplaatsfactor,
    environm_var = milieuvar,
    envvar_abbrev = afk_milieuvar,
    measurement_var = meetvar,
    target = primair,
    explanatory = verklarend,
    quality_control = kwaliteitscontrole,
    variable_set,
    main_datacollection_method,
    targetvar_temporal_targetpop:last_col()
  ) %>% 
  # warn about unknown schemes:
  {
    if (all(.$scheme %in% schemes$scheme)) {
      .
    } else {
      unknown <- unique(.$scheme[!(.$scheme %in% schemes$scheme)])
      warning(
        "Dropped unknown scheme(s): ", 
        paste(unknown, collapse = ", "),
        call. = FALSE
      )
      filter(., scheme %in% schemes$scheme)
    }
  } %>% 
  # check that data collection methods are known:
  {
    if (
      all(.$main_datacollection_method[!is.na(.$main_datacollection_method)] %in%
          activities$activity)
    ) {
      .
    } else {
      stop("Unknown data collection method(s) detected!", call. = FALSE)
    }
  } %>% 
  mutate(
    scheme = scheme %>% factor(levels = levels(n2khab_schemes$scheme)),
    main_datacollection_method = factor(
      main_datacollection_method, 
      levels = levels(activities$activity)
    ),
    across(c(target:quality_control, starts_with("mne")), ~ !is.na(.)),
    across(where(is.character), factor),
    variable_set = factor(
      variable_set, 
      levels = c(levels(variable_set), "LSVI")
    )
  ) %>% 
  pivot_longer(
    cols = starts_with("mne"),
    names_to = "module",
    values_to = "keep"
  ) %>% 
  filter(keep) %>% 
  # warn about inactive or unknown modules:
  {
    if (all(.$module %in% modules$module)) {
      .
    } else {
      unknown <- unique(.$module[!(.$module %in% modules$module)])
      warning(
        "Dropped inactive or unknown module(s): ", 
        paste(unknown, collapse = ", "),
        call. = FALSE
      )
      filter(., module %in% modules$module)
    }
  } %>% 
  mutate(module = factor(module, levels = levels(modules$module))) %>% 
  relocate(module) %>% 
  select(-keep)
```

Checking which MNE schemes per module are not covered by `envvars` (this may yield cases where the applicable domain has no target population units available, since `module_schemes` will only take this into account in chapter \@ref(scheme-objects)).

```{r}
module_schemes_mne <- 
  module_schemes %>% 
  filter(!str_detect(scheme, "^HQ"))
module_schemes_mne %>% 
  anti_join(
    envvars %>% 
      distinct(module, scheme),
    join_by(module, scheme)
  )
```

<!-- We currently see no unexpected cases. -->

Checking that no unplanned schemes are covered.

```{r}
envvars %>% 
  distinct(module, scheme) %>% 
  anti_join(module_schemes_mne, join_by(module, scheme)) %>% 
  nrow() == 0
```

<!-- Creating a subset of `envvars` that only involves data collection methods that are field activities. -->

<!-- ```{r} -->
<!-- envvars_field <-  -->
<!--   envvars %>%  -->
<!--   filter(data_collection_method %in% activities$activity[activities$is_field_activity]) -->
<!-- ``` -->

Some scheme x environmental variable combinations are module-specific:

```{r}
envvars %>% 
  distinct(module, scheme, environm_var, envvar_abbrev, variable_set, main_datacollection_method) %>% 
  count(scheme, environm_var, envvar_abbrev, variable_set, main_datacollection_method, name = "n_modules") %>% 
  inner_join(
    envvars %>% 
      distinct(module, scheme) %>% 
      count(scheme, name = "n_modules_all"),
    join_by(scheme)
  ) %>% 
  filter(n_modules != n_modules_all)
```

Creating an object with _non-field_ data collection methods per module x scheme, which will add to later cost calculations per module.

```{r}
mod_scheme_datacoll_nofield <- 
  envvars %>% 
  distinct(module, scheme, environm_var, envvar_abbrev, variable_set, main_datacollection_method) %>% 
  filter(
    !(main_datacollection_method %in% activities$activity[activities$is_field_activity]),
    !is.na(main_datacollection_method)
  ) %>% 
  arrange(module, scheme, variable_set, main_datacollection_method)
```

Defining variable sets and 'main data collection methods', restricted to field activities (hence: 'main field method'), per module x scheme (active modules only).
We also add the variable set and field methods for the MHQ programme for modules where this applies.

```{r}
mod_scheme_variableset_mainfieldmethod <-
  # MNE schemes
  envvars %>%
  distinct(module, scheme, variable_set, main_datacollection_method) %>%
  # warn about missing data collection methods:
  {
    if (!any(is.na(.$main_datacollection_method))) {
      .
    } else {
      warning(
        "Dropped ",
        sum(is.na(.$main_datacollection_method)),
        " row(s) with missing data collection method.", 
        call. = FALSE
      )
      filter(., !is.na(main_datacollection_method))
    }
  } %>%
  # limit to field activities:
  filter(main_datacollection_method %in% activities$activity[activities$is_field_activity]) %>% 
  rename(main_field_method = main_datacollection_method) %>% 
  mutate(scheme = factor(scheme, levels = levels(n2khab_schemes$scheme))) %>%
  arrange(module, scheme, variable_set, main_field_method) %>% 
  # add variable set & field method for MHQ schemes:
  bind_rows(
    module_schemes %>% 
      filter(str_detect(scheme, "^HQ")) %>% 
      select(-is_core_scheme) %>% 
      inner_join(
        schemes %>%
          select(scheme, tag_1),
        join_by(scheme),
        unmatched = "drop",
        relationship = "many-to-one"
      ) %>%
      mutate(
        variable_set = "LSVI",
        main_field_method = ifelse(tag_1 == "aq", "LSVIAQ", "LSVITERR")
      ) %>% 
      select(-tag_1)
  ) %>%
  mutate(
    variable_set = factor(variable_set, levels = levels(envvars$variable_set)),
    main_field_method = factor(main_field_method, levels = levels(activities$activity))
  )
mod_scheme_variableset_mainfieldmethod
```

For MHQ the variables are summarized as one variable set 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

Checking that all schemes are covered by `mod_scheme_variableset_mainfieldmethod`.

```{r}
schemes %>%
  anti_join(mod_scheme_variableset_mainfieldmethod, by = "scheme") %>%
  nrow() == 0
```

'Field activity' can cover data collection, where it corresponds to 'main field method' from above, but it is broader and also comprises installation or maintenance of in-situ instruments.

```{r}
field_activities <- 
  activities %>% 
  filter(is_field_activity) %>% 
  select(starts_with("activity")) %>% 
  rename_with(~ str_c("field_", .))
field_activities
```

Several field activities necessarily belong together, depending on the applicable sequence of activities, as expressed by the 'activity sequence' variable.

```{r}
field_activity_sequences <- 
  activity_sequences %>% 
  semi_join(field_activities, join_by(activity == field_activity)) %>% 
  rename(field_activity = activity)
field_activity_sequences
```

- Field activities are module x scheme specific.

  **Missing functionality**: consider making field activities type-specific (i.e. join with the target population), since within a scheme there may be differences by type for explanatory or quality-control variables: scheme sections are only defined for target environmental variables, not additional ones.
  
  A special type-specific field activity is the evaluation of a location: does it belong to type X of the target population?
  Type specificity in this case will always invoke splitting of the field activity, since the type specificity concerns the panel design properties; especially the timing during the year depends on the type.
  The revisit design of these field activities will not be scheme specific, but given the dependency on the target population, different schemes will have different location evaluation field activities.

  `mod_scheme_field_activity` is like `mod_scheme_variableset_mainfieldmethod`, but expanded with all field activities that are linked to the main field method through the applicable sequence (see intermediate result `mod_scheme_mainfieldmethod_sequence`).
  Also it abstracts variable sets away.
  
```{r}
mod_scheme_mainfieldmethod_sequence <- 
  mod_scheme_variableset_mainfieldmethod %>% 
  distinct(module, scheme, main_field_method) %>%
  inner_join(
    field_activity_sequences %>% 
      select(activity_sequence, field_activity),
    join_by(main_field_method == field_activity),
    unmatched = "drop",
    relationship = "many-to-many"
  ) %>%
  filter(
    # for GW_05.1_aq x GWLEVREAD, only include sequence gwsurf_lev
    !(
      scheme == "GW_05.1_aq" & 
        main_field_method == "GWLEVREAD" & 
        activity_sequence == "gw_lev"
    ),
    # for GW_03.3, 05.1_terr/quarries, 05.2 x GWLEVREAD, only include sequence gw_lev
    !(
      scheme %in% c("GW_03.3", "GW_05.1_terr", "GW_05.1_quarries", "GW_05.2") & 
        main_field_method == "GWLEVREAD" & 
        activity_sequence == "gwsurf_lev"
      )
  )
```

Check that one field method per module and scheme was translated to exactly one sequence:

```{r}
nrow(mod_scheme_mainfieldmethod_sequence) == 
  nrow(
    mod_scheme_variableset_mainfieldmethod %>% 
      distinct(module, scheme, main_field_method)
  )
```

```{r}
mod_scheme_field_activity <- 
  mod_scheme_mainfieldmethod_sequence %>% 
  rename(field_activity = main_field_method) %>% 
  # add missing field activities for each sequence
  nest(data = c(field_activity)) %>% 
  mutate(data = map2(activity_sequence, data, function(s, df) {
    levs <- field_activity_sequences %>% 
      filter(activity_sequence == s) %>% 
      pull(field_activity) %>% 
      as.character
    df %>% 
      mutate(field_activity = factor(field_activity, levels = levs)) %>% 
      expand(field_activity)
  })) %>% 
  unnest(data) %>% 
  left_join(
    field_activity_sequences,
    join_by(activity_sequence, field_activity),
    unmatched = "drop",
    relationship = "many-to-one"
  ) %>% 
  distinct(module, scheme, field_activity, choose_one)
```

We set the variable `cycle_duration_y` per module x scheme, which indeed depends on the combination of module x scheme already.

Note that schemes, hence scheme properties, can only be correctly linked to a module x domain combination (i.e. for domains with predetermined sample size) once the spatial sampling frames are created for each scheme: the sampling frames will also determine which types, hence schemes do apply in specific domains (because schemes with just a few rarer types will not apply to every domain).
So the below `module_scheme_cycle_duration` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

```{r}
module_scheme_cycle_duration <-
  module_schemes %>%
  select(module, scheme) %>%
  mutate(cycle_duration_y = case_when(
    module == "mne2024_pan_be2100024_restoration" ~ 3L,
    str_detect(scheme, "^SURF") ~ 3L,
    .default = 6L
  ))
```

Setting year and month of first implementation (this is further used in chapter \@ref(scheme-objects)).

```{r}
year_first_default <- 2025L
month_first_default <- 7L
```

A 'field activity group' (FAG) is a set of field activities (FA) with the same panel count, period, group size and group repetition attributes _and_ which typically take place during the same location visit.

```{r}
fa_fag <- 
  read_vc(
    file = "fa_fag",
    root = file.path(datapath, "text/raw")
  ) %>%
  as_tibble()
```

- `mod_scheme_fag_paneldesign` is the revisit design of field activity groups per module and scheme.
This will matter most for data collection plans.

First connect `mod_scheme_field_activity` to field activity groups.

```{r}
mod_scheme_fa_fag <- 
  mod_scheme_field_activity %>% 
  inner_join(
    fa_fag,
    join_by(field_activity),
    unmatched = "drop",
    relationship = "many-to-many"
  ) %>% 
  filter(
    # for groundwater quality, don't install surface water equipment (including aquatic types)
    !(str_detect(scheme, "03\\.3|04\\.2|05\\.2|07\\.1") & str_detect(field_activity_group, "^GWSURFLEV")),
    # for terrestrial groundwater, don't install surface water equipment
    !(str_detect(scheme, "_terr|_quarries") & str_detect(field_activity_group, "^GWSURFLEV")),
    # for aquatic groundwater, don't include the terrestrial-only fag
    !(scheme == "GW_05.1_aq" & str_detect(field_activity_group, "^GWLEV")),
    # for _lentic, don't include a lotic fag and vice-versa
    !(str_detect(scheme, "_lentic") & str_detect(field_activity_group, "LOT")),
    !(str_detect(scheme, "_lotic") & str_detect(field_activity_group, "LENT"))
  )
```

Check that the number of rows did not change:

```{r}
nrow(mod_scheme_fa_fag) == nrow(mod_scheme_field_activity)
```

Now collapse to field activity groups and define the panel designs.

Note that below `mod_scheme_fag_paneldesign` object will be filtered and rewritten based on `module_domain_schemes` once the latter has been created (chapter \@ref(scheme-objects)).

Note that some field activity groups are type-specific; this is indicated by the `choose_one` variable.

```{r}
# field_activity_paneldesign <-
mod_scheme_fag_paneldesign <-
  mod_scheme_fa_fag %>% 
  distinct(module, scheme, field_activity_group, choose_one) %>% 
  mutate(
    revisit_pattern = case_when(
      str_detect(field_activity_group, "LEVREAD") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    panel_period_main = case_when(
      str_detect(field_activity_group, "LEVINST$|SOIL.*SAMP") ~ months(3),
      str_detect(field_activity_group, "LEVREAD") ~ months(6),
      .default = years(1)
    ),
    enforce_cycle_duration_y = case_when(
      str_detect(field_activity_group, "LEVINST$") ~ 6L
    ),
    limit_n_cycles = case_when(
      str_detect(field_activity_group, "SAMPLPOINT") ~ 1L,
      .default = NA_integer_
    ),
    revisit_subpattern = case_when(
      str_detect(field_activity_group, "SURF.*DATACOLL") ~ "pure_panel",
      str_detect(field_activity_group, "GWSHALL(CLEAN|SAMP)|LEVREAD") ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    panel_subperiod = case_when(
      str_detect(field_activity_group, "SURF.*DATACOLL") ~ months(1),
      str_detect(field_activity_group, "GWSHALL(CLEAN|SAMP)|LEVREAD") ~ months(3)
    ),
    limit_n_subpanels_per_panel = case_when(
      str_detect(field_activity_group, "GWSHALL(CLEAN|SAMP)") ~ 2L
    )
  )
mod_scheme_fag_paneldesign
# mod_scheme_fag_paneldesign %>% distinct(pick(field_activity_group, revisit_pattern:last_col()))
```


- `scheme_variableset_paneldesign` is the revisit design of measurements.
It is created below to distinguish with `field_activity_paneldesign` above, but yet unclear whether we'll need it to create data collection plans.
Its primary relevance is statistical (spatiotemporal pattern of available data), hence it underpins choices made above.
Notable differences with `field_activity_paneldesign` are that:
  - automated frequent measurements are reflected as 'frequent revisit' while fieldwork is much less frequent;
  - no installation or maintenance is involved;
  - the (assumed) temporal resolution of the target variable is added, which is an attempt to answer the question: which period does one value of the variable represent?

```{r}
scheme_variableset_paneldesign <-
  scheme_variableset_fieldmethod %>%
  distinct(scheme, variable_set) %>%
  mutate(
    targetvar_temporal_resolution = case_when(
      str_detect(scheme, "^HQ") | str_detect(variable_set, "peil") ~ years(1),
      str_detect(scheme, "^SURF_") ~ months(1),
      .default = months(3)
    ),
    revisit_pattern = case_when(
      str_detect(variable_set, "peil") ~ "pure_panel",
      .default = "serially_alternating"
    ) %>%
      factor(),
    panel_period_main = case_when(
      str_detect(variable_set, "peil") ~ days(1),
      str_detect(scheme, "^HQ|^SURF_|^GW_(03|04|07)") ~ years(1),
      .default = months(3)
    ),
    revisit_subpattern = case_when(
      str_detect(scheme, "^SURF_") ~ "pure_panel",
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ "serially_alternating"
    ) %>%
      factor(levels = levels(revisit_pattern)),
    panel_subperiod = case_when(
      !is.na(revisit_subpattern) ~ targetvar_temporal_resolution
    ),
    limit_n_subpanels_per_panel = case_when(
      str_detect(scheme, "^GW_(03|04|07)") &
        revisit_pattern != "pure_panel" ~ 2L
    )
  )
scheme_variableset_paneldesign
```

## Setting data source versions



## Overview of most relevant objects in this chapter

Objects subsetted with `[[1]]` mean that they (may) contain more, similarly structured elements.

Note that the following objects are not yet finished at this stage (see chapter \@ref(scheme-objects)):

- `module_scheme_cycle_duration`
- `scheme_targetpop_sets`

```{r echo = FALSE}
for (name in c(
  "scheme_targetpop_sets",
  "scheme_targetpop_sets$scheme_set[[1]]",
  "scheme_targetpop_sets$targetpops[[1]]",
  "schemes",
  "targetpops",
  "modules",
  "domains",
  "module_domains",
  "scheme_variableset_fieldmethod",
  "scheme_field_activity",
  "field_activity_paneldesign",
  "scheme_variableset_paneldesign",
  "module_scheme_cycle_duration"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
