# Settings

<!-- For the schemes, also recognize that a subset may exist with extra variables. -->
<!-- Use dummy variables for now, but make sure various variable types exist (repetitions within one year, once per year, continuous measurements which need multiple visits per year but only one value per year, etc.) -->
<!-- Also include necessary aspects for resource allocation in the sampling design, especially time investments per variable. (this may be added later) -->

## Setting schemes, types and domains

Setting schemes and deriving associated types, which delineate the target population.

Also restricting columns to the necessary ones for workflows (unlike for 'displays').

```{r}
schemes <- 
  read_schemes(lang = lang) |> 
  filter(attribute_1 %in% c("GW", "SURF", "SOIL")) |> 
  select(scheme, programme, attribute_1, attribute_2, attribute_3, tag_1, spatial_restriction)
```


```{r restriction-1}
# ARTIFICIAL RESTRICTION 1
schemes <- 
  schemes |> 
  filter(scheme %in% c(
    "GW_05.1_terr", 
    "GW_07.1", 
    "SOIL_03.2", 
    "SURF_03.4_lentic", 
    "SURF_03.4_lentic", 
    "SURF_064"
  ))
```

```{r}
targetpop <- 
  read_scheme_types(lang = lang) |> 
  select(scheme, type) |> 
  semi_join(schemes, by = "scheme")
```

```{r restriction-2}
# ARTIFICIAL RESTRICTION 2
# targetpop <- targetpop |> 
#   filter(type %in% c(
#     
#   ))
```

Extending `schemes` with MHQ schemes for the selected types.

```{r}

```

Included types with some properties.

```{r}
types <- 
  read_types(lang = lang) |> 
  select(type, typelevel, main_type, typeclass, hydr_class, groundw_dep, flood_dep)
```



## Setting scheme properties

**Missing functionality**: getting the primary and additional site characteristics and environmental variables for each scheme in MNE from existing datasets such as googlesheets.

For MHQ the variables will be summarized as one variable 'LSVI', and the hierarchy of involved measurement variables for each scheme (type) is incorporated in the [LSVI](https://packages.inbo.be/LSVI/) package.

<!-- Selectie van variabelen te voorzien. -->

## Setting data source versions



