# Creating the base sampling frame

## Preparing elements of the base sampling frame

### Types to be distinguished and their properties

Summarizing the target populations of all schemes in MNE & MHQ (prefix `n2khab_`) as an overall list of types, including the expansion to main types or subtypes for each scheme for better spatial joining with `habitatmap_terr`, which is a 'best possible guess' of actually present terrestrial types (see its documentation).

```{r}
n2khab_targetpops <-
  read_scheme_types() %>%
  select(scheme, type)
n2khab_types <-
  n2khab_targetpops %>%
  distinct(type) %>%
  arrange(type)
n2khab_targetpops_expanded <-
  n2khab_targetpops %>%
  group_by(scheme) %>%
  expand_types(strict = FALSE) %>%
  ungroup()
n2khab_types_expanded <-
  n2khab_targetpops_expanded %>%
  distinct(type) %>%
  arrange(type)
```

The base sampling frame will include a layer for each of the resulting main types or subtypes.
More correctly, a layer is to be made for each stratum (see further), which can be further subdivisions of types.

Since the target population of each scheme is defined by non-overlapping type codes (never combining a main type with its subtypes), when deriving sampling frames for each scheme:

- locations of subtypes that were added by the code above will be merged in the sampling frame layer of their main type
- locations of an added (but unresolved) main type will be added (repeated) in each of its corresponding subtype sampling frame layers.
This only applies to subtypes that triggered the addition of the main type (see `?n2khab::expand_types(strict = FALSE)`).

This way, the sampling frames will present the exact set of type codes as in `targetpops`, while taking advantage of the type expansion since all codes added by expansion will be present in the base sampling frame.

In the ideal situation, the base sampling frame is 'resolved' by terrain visits by replacing main type codes by the correct subtype code, for main types that do have subtypes.

```{r eval=FALSE, include=FALSE}
# various main type codes are not added because habitatmap_terr doesn't have
# them OR they're not involved in MNE/MHQ anyway
read_types(lang = lang) %>%
  anti_join(n2khab_types_expanded, by = "type")
# following types were added by expansion:
n2khab_types_expanded %>%
  anti_join(n2khab_types, by = "type") %>%
  inner_join(read_types() %>% select(type, typelevel),
    by = "type"
  )
```

Gathering minimal needed properties of all types.

```{r}
n2khab_types_expanded_properties <-
  read_types() %>%
  semi_join(n2khab_types_expanded, by = "type") %>%
  select(type, typelevel, main_type, typeclass, hydr_class, groundw_dep, flood_dep) %>%
  left_join(
    read_vc(
      "030_preparations/010_explore_targetpop/type_approaches",
      root = gitroot
    ) %>%
      as_tibble() %>%
      mutate(
        grts_join_method = ifelse(approach_cellbased, "cell", "cell_center")
      ) %>%
      select(type, grts_join_method),
    by = "type"
  ) %>%
  mutate(
    grts_join_method = case_when(
      !is.na(grts_join_method) ~ grts_join_method,
      type == "7220" ~ "original_point",
      type == "3260" ~ "endpoint_of_segment",
      type == "8310" | hydr_class == "HC3" ~ "point_on_surface",
      .default = "cell_center"
    ) %>%
      factor(levels = c(
        "cell_center",
        "cell",
        "point_on_surface",
        "endpoint_of_segment",
        "original_point"
      )),
    sample_support = case_when(
      grts_join_method == "cell_center" ~ "area of type within 1024 m² cell (cell center must be type)",
      grts_join_method == "cell" ~ "area of type within 1024 m² cell",
      type == "3260" ~ "predefined watercourse segment (mostly 100 m)",
      type == "7220" ~ "population unit inferred from read_habitatsprings(units_7220 = TRUE)",
      type == "8310" ~ "population unit as defined by read_habitatquarries(filter_hab = TRUE)",
      .default = "watersurface"
    ) %>%
      factor(),
    sample_support_code = fct_recode(
      sample_support,
      cell_conditioned_on_center = "area of type within 1024 m² cell (cell center must be type)",
      cell = "area of type within 1024 m² cell",
      watercourse_segment = "predefined watercourse segment (mostly 100 m)",
      spring = "population unit inferred from read_habitatsprings(units_7220 = TRUE)",
      quarry = "population unit as defined by read_habitatquarries(filter_hab = TRUE)",
    )
  )
```

### Reading and filtering spatial data of the target populations

Reading and filtering spatial data of terrestrial types.

```{r}
hmt <- read_habitatmap_terr(keep_aq_types = FALSE, drop_7220 = TRUE)
typelev <- levels(hmt$habitatmap_terr_types$type)
hmt_occ <-
  hmt$habitatmap_terr_types %>%
  semi_join(n2khab_types_expanded,
    by = "type"
  ) %>% 
  # manually fixing the type of two polygons in habitatmap_terr_2023_v1:
  mutate(
    source = case_when(
      polygon_id == "184091_v2023" & type == "2190_overig" ~ "habitatmap_stdized + manual fix",
      polygon_id == "533826_v2023" & type == "2310" ~ "habitatmap_stdized + manual fix",
      .default = source
    ) %>% 
      factor(),
    type = case_when(
      polygon_id == "184091_v2023" & type == "2190_overig" ~ "9120",
      polygon_id == "533826_v2023" & type == "2310" ~ "2130",
      .default = type
    ) %>% 
      factor(levels = typelev)
  )
hmt_pol <-
  hmt$habitatmap_terr_polygons %>%
  semi_join(hmt_occ, by = "polygon_id") %>%
  select(polygon_id) %>%
  st_make_valid()
```

Reading and filtering spatial data of lentic aquatic types.

```{r}
wsh <- read_watersurfaces_hab(interpreted = TRUE)
wsh_occ <-
  wsh$watersurfaces_types %>%
  semi_join(n2khab_types_expanded,
    by = "type"
  )
wsh_pol <-
  wsh$watersurfaces_polygons %>%
  semi_join(wsh_occ, by = "polygon_id") %>%
  select(polygon_id)
```

Reading spatial data of type 3260.

<!-- maar beter ineens de afgeleide dataset opbouwen / inlezen met de segmenten: betrek daarbij het voorbereidend werk van Tom DD en vlgs googledoc steekproefkader is dit iets voor Toon W, incl koppeling met punten -->

```{r}
habstream <- read_habitatstreams()
```

Reading spatial data of type 7220.
Each row (point) already represents a population unit in this case.

```{r}
flanders_buffer <-
  read_admin_areas(dsn = "flanders") %>%
  st_buffer(40)
habspring_units <-
  read_habitatsprings(units_7220 = TRUE) %>%
  .[flanders_buffer, ]
```

Reading spatial data of type 8310.
Each row (polygon) does not necessarily represent a population unit, but the latter is indicated by the `unit_id` variable.
However in this case the filtering by 8310 already results in just one polygon per unit.

```{r}
habquarries <-
  read_habitatquarries(filter_hab = TRUE)
```

### Defining the strata used in sampling and inference

The strata are the types, in some cases subdivided according to area class of the population unit.
The base sampling frame layers will consist of the strata, not just the types.

```{r prepare-strata}
extract_interval_limits <- function(df) {
  mutate(
    df,
    lower = str_match(stratum, "^\\((\\d+),\\d+\\]$")[, 2] %>% as.integer(),
    upper = str_match(stratum, "^\\(\\d+,(\\d+)\\]$")[, 2] %>% as.integer()
  )
}
wsh_pol_area_class <-
  wsh_pol %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>%
  st_drop_geometry() %>%
  select(polygon_id, area) %>%
  mutate(
    stratum =
      area %>%
        cut(c(0, 1, 5, 50, 150)) %>%
        as.character()
  ) %>%
  inner_join(
    wsh_occ %>%
      select(polygon_id, type),
    .,
    by = "polygon_id"
  ) %>%
  rename(unit_id = polygon_id)
wsh_strata <-
  wsh_pol_area_class %>%
  extract_interval_limits() %>%
  distinct(type, lower, upper) %>%
  arrange(type, lower, upper)
habquarries_area_class <-
  habquarries %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>%
  st_drop_geometry() %>%
  select(unit_id, type, area) %>%
  mutate(
    stratum =
      area %>%
        cut(c(0, 2, 6, 40)) %>%
        as.character()
  ) %>%
  mutate(unit_id = as.character(unit_id))
habquarries_strata <-
  habquarries_area_class %>%
  extract_interval_limits() %>%
  distinct(type, lower, upper) %>%
  arrange(type, lower, upper)
```


```{r result-strata}
# n2khab_strata_expanded: defines the strata of the base sampling frame
n2khab_strata_expanded <-
  n2khab_types_expanded %>%
  left_join(
    bind_rows(wsh_strata, habquarries_strata),
    by = "type"
  ) %>%
  unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
  select(type, stratum) %>%
  mutate(stratum = factor(stratum))
# n2khab_strata: for sampling and inference from the spatial sampling frames,
# which it defines (except for further needed spatial restrictions)
n2khab_strata <-
  n2khab_types %>%
  left_join(
    bind_rows(wsh_strata, habquarries_strata),
    by = "type"
  ) %>%
  unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
  select(type, stratum) %>%
  mutate(
    stratum = factor(stratum, levels = levels(n2khab_strata_expanded$stratum))
  )
```

### Data sources with the population unit geometries

Reading the `GRTSmaster_habitats` data source that defines the population unit geometries of the terrestrial types.

```{r}
grts_mh <- read_GRTSmh()
```

For watersurface types, the unit geometries are defined by the `watersurfaces` data source (units are polygons).

For type 3260, the unit geometries are defined by `read_watercourse_100mseg(element = "lines")` (units are segments).

For type 7220, we already created the `habspring_units` object (undefined units referred to by points).

For type 8310, we already created the `habquarries` object (units are polygons).


### Defining the point - geometry relationship for specific types

These may encompass units not currently covered by the target populations.
This is especially true for watersurfaces and watersegments.
This allows for dynamic changes in the occurrence of types.

We don't need points for `grts_join_method == "cell"`.

We will use cell centers on the fly for `grts_join_method == "cell_center"`.

Defining points that represent the watersurface units (object `points_watsurf`).
This step requires the availability of data that are currently external to this repository and used packages.

```{r}
if (curl::has_internet()) {
  # Accessing a temporary data source for watersurfaces
  with_drive_quiet({
    path <- file.path(tempdir(), "watersurfaces_refpoints.tsv")
    drive_download(as_id("1XgtdZOebbqp-MJih60P_Vi-npz_5UVkY"), path = path)
    path2 <- file.path(tempdir(), "watersurfaces_refpoints.yml")
    drive_download(as_id("1Bb07utgZOyMK65TdR64-4VjGP7T6GqMa"), path = path2)
  })
  points_watsurf <- 
    read_vc("watersurfaces_refpoints", tempdir()) %>%
    as_tibble()
  invisible(file.remove(path, path2))
} else {
  # Clumsy approach; needs maintenance. In the longer term, the management of
  # all this needs more centralization and rigidity.
  mhq_ssfws_datapath <- file.path(
    dirname(gitroot),
    "n2khab-samplingframes/data/watersurfaces"
  )
  stopifnot(dir.exists(mhq_ssfws_datapath))
  points_watsurf <- 
    read_vc("watersurfaces_refpoints", root = mhq_ssfws_datapath) %>%
    as_tibble()
}
points_watsurf <- 
  points_watsurf %>%
  select(polygon_id, x, y) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370)
# adding all other watersurfaces
points_watsurf <-
  bind_rows(
    points_watsurf,
    read_watersurfaces() %>%
      filter(!(polygon_id %in% points_watsurf$polygon_id)) %>%
      select(polygon_id) %>%
      st_sf(agr = "identity") %>%
      st_point_on_surface() %>%
      rename(geometry = geom)
  )
```

Defining points that represent the watercourse segments (object `points_watcourse`).

```{r}
points_watcourse <- read_watercourse_100mseg(element = "points")
```

For type 7220, we can use the points in `habspring_units`.

Defining points that represent the type 8310 units (object `points_8310`).

```{r}
points_8310 <-
  habquarries %>%
  select(unit_id) %>%
  st_sf(agr = "identity") %>%
  st_point_on_surface()
```

Combining all points that refer to non-cell units into one object.

```{r}
points_non_cell <-
  bind_cols(
    tibble(sample_support_code = "watersurface"),
    points_watsurf %>% rename(unit_id = polygon_id)
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "watercourse_segment"),
      points_watcourse %>%
        unite(unit_id, vhag_code, rank) %>%
        rename(geometry = geom)
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "spring"),
      habspring_units %>%
        transmute(unit_id = as.character(unit_id))
    )
  ) %>%
  bind_rows(
    bind_cols(tibble(sample_support_code = "quarry"), points_8310) %>%
      mutate(unit_id = as.character(unit_id)) %>%
      rename(geometry = geom)
  ) %>%
  mutate(sample_support_code = factor(
    sample_support_code,
    levels = levels(n2khab_types_expanded_properties$sample_support_code)
  )) %>%
  st_sf(sf_column_name = "geometry")
```

The `sample_support_code` needs to be combined with `unit_id` to be globally unique!
Alternatively, they could be combined into a single column with unique IDs, but then it would loose connection with the corresponding columns of spatial data sources.

## Assigning a GRTS address to all population units of each stratum

The end result of this, `stratum_grts_n2khab`, is what constitutes the essence of the base sampling frame.

However in its full meaning the base sampling frame also comprises various linked attributes generated before.

### For the non-cell units

From `points_non_cell`, we need to filter those that fall within the target populations of MNE and MHQ.
This is already fine for types 7220 and 8310.

We determine the watercourse segments that have type 3260.

```{r}
# Quick-and-dirty approach to define the 3260 segments (i.e. it will miss a
# part and some may be false positives)
segm_3260 <- read_watercourse_100mseg(element = "lines")[habstream, ] %>%
  unite(unit_id, vhag_code, rank)
```

The watersurfaces with types are defined by `wsh_pol`; the segments with 3260 are defined by `segm_3260`.

```{r}
points_non_cell_n2khab <-
  points_non_cell %>%
  filter(
    sample_support_code %in% c("spring", "quarry") |
      (sample_support_code == "watersurface" & unit_id %in% wsh_pol$polygon_id) |
      (sample_support_code == "watercourse_segment" & unit_id %in% segm_3260$unit_id)
  )
```

Determining the GRTS address of the units represented by `points_non_cell_n2khab`.

```{r}
units_non_cell_n2khab_grts <-
  points_non_cell_n2khab %>%
  mutate(
    grts_address = extract(grts_mh, st_coordinates(points_non_cell_n2khab)) %>%
      pull(GRTSmaster_habitats)
  ) %>%
  st_drop_geometry()
```

<!-- Er zijn gevallen met hetzelfde GRTS-adres; een oplossing is hiervoor uit te werken: -->
<!-- > units_non_cell_n2khab_grts %>% count(sample_support_code, grts_address) %>% filter(n > 1) %>% count(sample_support_code, n) -->
<!-- Storing counts in `nn`, as `n` already present in input -->
<!-- ℹ Use `name = "new_name"` to pick a new name. -->
<!-- # A tibble: 7 × 3 -->
<!--   sample_support_code     n    nn -->
<!--   <fct>               <int> <int> -->
<!-- 1 watercourse_segment     2    21 -->
<!-- 2 watercourse_segment     3     1 -->
<!-- 3 watercourse_segment   122     1 -->
<!-- 4 watersurface            2    54 -->
<!-- 5 watersurface            3    13 -->
<!-- 6 watersurface            4     8 -->
<!-- 7 watersurface            5     2 -->

Listing the non-cell units per stratum (within the target populations of MNE and MHQ), needed in order to connect the GRTS addresses with strata.

```{r}
stratum_units_non_cell_n2khab <-
  bind_cols(
    tibble(sample_support_code = "watersurface"),
    wsh_pol_area_class %>%
      distinct() %>%
      extract_interval_limits() %>%
      unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
      select(stratum, unit_id)
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "watercourse_segment"),
      segm_3260 %>%
        st_drop_geometry() %>%
        mutate(stratum = "3260", .before = everything())
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "spring"),
      habspring_units %>%
        st_drop_geometry() %>%
        mutate(
          stratum = "7220",
          unit_id = as.character(unit_id),
          .keep = "none",
          .before = everything()
        )
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "quarry"),
      habquarries_area_class %>%
        extract_interval_limits() %>%
        unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
        select(stratum, unit_id)
    )
  ) %>%
  mutate(
    sample_support_code = factor(
      sample_support_code,
      levels = levels(n2khab_types_expanded_properties$sample_support_code)
    ),
    stratum = factor(stratum, levels = levels(n2khab_strata_expanded$stratum))
  ) %>%
  relocate(stratum)
```

Note that `stratum_units_non_cell_n2khab` has `r nrow(stratum_units_non_cell_n2khab)` rows which is larger than the `r nrow(units_non_cell_n2khab_grts)` rows of `units_non_cell_n2khab_grts`.
This is because some watersurface units are attached to more than 1 stratum (i.e. more than 1 type).

Assigning the GRTS addresses of the units to the corresponding strata in `n2khab_strata_expanded`.

```{r}
stratum_grts_non_cell_n2khab <-
  stratum_units_non_cell_n2khab %>%
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>%
  select(-sample_support_code, -unit_id) %>%
  # distinct() leads to less rows because of units with same GRTS address.
  distinct()
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_non_cell_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_non_cell_n2khab$grts_address)
```

### For the cell units

Filtering units of `GRTSmaster_habitats` that are part of the target populations of MNE and MHQ with sample support `cell_conditioned_on_center`, and storing their polygon ID and GRTS address.

```{r}
hmt_pol_cellcenter <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell_conditioned_on_center") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  semi_join(hmt_pol, ., by = "polygon_id") %>%
  vect()
```


```{r eval = params$regenerate_binary_files}
# takes < 1 minute:
hmt_pol_grts_cellcenter <-
  extract(grts_mh, hmt_pol_cellcenter, small = FALSE) %>%
  as_tibble() %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(hmt_pol_cellcenter)),
      polygon_id = hmt_pol_cellcenter$polygon_id
    ),
    .,
    by = "ID"
  ) %>%
  select(-ID, grts_address = GRTSmaster_habitats) %>%
  # filtering is needed since all polygons are listed by extract():
  filter(!is.na(grts_address))
saveRDS(
  hmt_pol_grts_cellcenter,
  file.path(datapath, "binary/intermediate/hmt_pol_grts_cellcenter.rds")
)
```

```{r}
hmt_pol_grts_cellcenter <-
  readRDS(file.path(
    datapath,
    "binary/intermediate/hmt_pol_grts_cellcenter.rds"
  ))
```

Filtering units of GRTSmaster_habitats that are part of the target populations of MNE and MHQ with sample support `cell`, and storing their polygon ID and GRTS address.

```{r}
hmt_pol_cell <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  semi_join(hmt_pol, ., by = "polygon_id") %>%
  vect()

hmt_pol_grts_cell <-
  # touches = TRUE allows about 2.5 times the number of cells
  extract(grts_mh, hmt_pol_cell, touches = TRUE) %>%
  as_tibble() %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(hmt_pol_cell)),
      polygon_id = hmt_pol_cell$polygon_id
    ),
    .,
    by = "ID"
  ) %>%
  select(-ID, grts_address = GRTSmaster_habitats)
```

Note that `hmt_pol_grts_cellcenter` and `hmt_pol_grts_cell` share a number of polygons so this should be used with care, i.e. not mixing the link of `polygon_id` with `grts_address` between cell-center vs. cell types.

```{r}
stratum_polygons_cell_all_n2khab <-
  n2khab_types_expanded_properties %>%
  filter(str_detect(sample_support_code, "^cell")) %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(stratum = type, polygon_id)
```

<!-- Evalueer of er daarom niet beter ook een object wordt gemaakt zoals `stratum_units_non_cell_n2khab`, maar dan een die de _polygonen_ oplijst per stratum (type) aangezien er geen unit_id's zijn; dus zou `stratum_polygons_cell_all_n2khab` heten. -->
<!-- Dit is inmiddels hierboven gedaan (wordt niet verder gebruikt); te bekijken of dit zinvol is. -->

Assigning the GRTS addresses of the units to the corresponding strata in `n2khab_strata_expanded`.

We can use `n2khab_types_expanded` since for terrestrial types the strata are the types without subdivision.

```{r}
hmt_pol_stratum_grts_cellcenter <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell_conditioned_on_center") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(polygon_id, stratum = type) %>%
  inner_join(
    hmt_pol_grts_cellcenter,
    by = "polygon_id",
    relationship = "many-to-many"
  )
stratum_grts_cellcenter_n2khab <-
  hmt_pol_stratum_grts_cellcenter %>%
  select(-polygon_id)
hmt_pol_stratum_grts_cell <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(polygon_id, stratum = type) %>%
  inner_join(
    hmt_pol_grts_cell,
    by = "polygon_id",
    relationship = "many-to-many"
  )
stratum_grts_cell_n2khab <-
  hmt_pol_stratum_grts_cell %>%
  select(-polygon_id) %>%
  # adjacent polygons can share the same type AND grts_address here, so:
  distinct()
stratum_grts_cell_all_n2khab <-
  bind_rows(stratum_grts_cellcenter_n2khab, stratum_grts_cell_n2khab)
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_cell_all_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_cell_all_n2khab$grts_address)
```

### Integrate the GRTS addresses for cell and non-cell units

```{r}
stratum_grts_n2khab <-
  stratum_grts_non_cell_n2khab %>%
  bind_rows(stratum_grts_cell_all_n2khab)
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_n2khab$grts_address)
```

```{r}
stratum_grts_n2khab %>% filter(is.na(grts_address))
```

<!-- REVISIT BELOW FIX! XXXXXXXXXXXXXXXXXX -->

Temporary fix:

```{r}
stratum_grts_n2khab <-
  stratum_grts_n2khab %>%
  filter(!is.na(grts_address))
```


## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "n2khab_types_expanded_properties",
  "n2khab_strata_expanded",
  "n2khab_strata",
  "n2khab_targetpops_expanded",
  "n2khab_targetpops",
  "points_non_cell",
  "points_non_cell_n2khab",
  "stratum_units_non_cell_n2khab",
  "units_non_cell_n2khab_grts",
  "stratum_grts_non_cell_n2khab",
  "stratum_grts_cellcenter_n2khab",
  "stratum_grts_cell_n2khab",
  "stratum_grts_cell_all_n2khab",
  "stratum_grts_n2khab"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```

<!-- ## Writing the base sampling frame -->


<!-- ## Reading the base sampling frame -->


