# Creating the base sampling frame and other essential starting points (datasets) to draw samples

## Creating elements of the base sampling frame

### Types to be distinguished and their properties

Summarizing the target populations of all schemes in MNE & MHQ (prefix `n2khab_`) as an overall list of types, including the expansion to main types or subtypes for each scheme for better spatial joining with `habitatmap_terr`, which is a 'best possible guess' of actually present terrestrial types (see its documentation).

```{r}
n2khab_targetpop <-
  read_scheme_types() %>%
  select(scheme, type)
n2khab_targetpop_expanded <-
  n2khab_targetpop %>%
  group_by(scheme) %>%
  expand_types() %>%
  ungroup()
n2khab_types_expanded <-
  n2khab_targetpop_expanded %>%
  distinct(type) %>%
  arrange(type)
```

The base sampling frame will include a layer for each of the resulting main types or subtypes.
More correctly, a layer is to be made for each stratum (see further), which can be further subdivisions.

Since the target populations of each scheme are defined by non-overlapping type codes (never combining a main type with its subtypes), when deriving sampling frames for each scheme, locations of subtypes that were added by the code above will be merged in the sampling frame layer of their main type, and the locations of an added (but unresolved) main type will be added (repeated) in each of its corresponding subtype sampling frame layers.

This way, the sampling frames will present the exact set of type codes as in `targetpop`, while taking advantage of the type expansion since all codes added by expansion will be present in the base sampling frame.

In the ideal situation, the base sampling frame is 'resolved' by terrain visits by replacing main type codes by the correct subtype code, for main types that do have subtypes.

```{r eval=FALSE, include=FALSE}
# various main type codes are not added because habitatmap_terr doesn't have
# them OR they're not involved in MNE/MHQ anyway
read_types(lang = lang) %>% 
  anti_join(n2khab_types_expanded, by = "type")
# following types were added by expansion:
n2khab_types_expanded %>%
    anti_join(n2khab_targetpop %>% distinct(type), by = "type") %>% 
    inner_join(read_types() %>% select(type, typelevel),
               by = "type")
```

Gathering minimal needed properties of all types.

```{r}
n2khab_types_expanded_properties <-
  read_types() %>% 
  semi_join(n2khab_types_expanded, by = "type") %>% 
  select(type, typelevel, main_type, typeclass, hydr_class, groundw_dep, flood_dep) %>% 
  left_join(
    read_vc(
      "030_preparations/010_explore_targetpop/type_approaches",
      root = find_root(is_git_root)
    ) %>%
      as_tibble() %>% 
      mutate(
        grts_join_method = ifelse(approach_cellbased, "cell", "centroid") %>% 
          factor(levels = c("centroid", "cell", "point"))
      ) %>% 
      select(type, grts_join_method),
    by = "type"
  ) %>% 
  mutate(
    grts_join_method = case_when(
        !is.na(grts_join_method) ~ grts_join_method,
        type %in% c("7220", "8310") | hydr_class == "HC3" ~ "point",
        .default = "centroid"
    ),
    sample_support = case_when(
      grts_join_method == "centroid" ~ "area of type within 1024 m² cell (centroid must be type)",
      grts_join_method == "cell" ~ "area of type within 1024 m² cell",
      type == "3260" ~ "predefined watercourse segment (mostly 100 m)",
      type == "7220" ~ "population unit as defined by read_habitatsprings(units_7220 = TRUE)",
      type == "8310" ~ "population unit as defined by read_habitatquarries(filter_hab = TRUE)",
      .default = "watersurface"
    ) %>% 
      factor()
  )
```

### Reading and filtering spatial data

Reading and filtering spatial data of terrestrial types.

```{r}
hmt <- read_habitatmap_terr(keep_aq_types = FALSE, drop_7220 = TRUE)
hmt_occ <-
    hmt$habitatmap_terr_types %>%
    semi_join(n2khab_types_expanded,
              by = "type")
hmt_pol <-
    hmt$habitatmap_terr_polygons %>%
    semi_join(hmt_occ, by = "polygon_id") %>% 
    select(polygon_id) %>% 
    st_make_valid
```

Reading and filtering spatial data of lentic aquatic types.

```{r}
wsh <- read_watersurfaces_hab(interpreted = TRUE)
wsh_occ <- 
    wsh$watersurfaces_types %>% 
    semi_join(n2khab_types_expanded,
              by = "type")
wsh_pol <-
    wsh$watersurfaces_polygons %>%
    semi_join(wsh_occ, by = "polygon_id") %>% 
    select(polygon_id)
```

Reading spatial data of type 3260.

<!-- maar beter ineens de afgeleide dataset opbouwen / inlezen met de segmenten: betrek daarbij het voorbereidend werk van Tom DD en vlgs googledoc steekproefkader is dit iets voor Toon W, incl koppeling met punten -->

```{r}
habstream <- read_habitatstreams()
```

Reading spatial data of type 7220.
Each row (point) already represents a population unit in this case.

```{r}
flanders_buffer <- 
    read_admin_areas(dsn = "flanders") %>% 
    st_buffer(40)
habspring_units <- 
  read_habitatsprings(units_7220 = TRUE) %>% 
  .[flanders_buffer, ]
```


```{r}
habquarries <- read_habitatquarries(filter_hab = TRUE) # is restricted to Flanders
```

<!-- Below chunks go back to targetpop, for exploratory purposes. Best move and refer them... -->

```{r exploration-1, eval = FALSE, include=FALSE}
# building blocks for the exploration which types to select from targetpop for the POC

pop_size_terr <- 
  read_vc(
    "030_preparations/010_explore_targetpop/pop_size_terr", 
    root = find_root(is_git_root)
  ) %>% 
  as_tibble() %>% 
  inner_join(
    read_vc(
      "030_preparations/010_explore_targetpop/type_approaches",
      root = find_root(is_git_root)
    ),
    by = "type") %>% 
  filter(count_method == "cellbased" & approach_cellbased |
           count_method == "centroidbased" & !approach_cellbased) %>% 
  select(-c(count_method, approach_cellbased)) %>% 
  mutate(stratum = NA_character_,
         population_size = round(population_size) %>% as.integer) %>% 
  select(type, stratum, population_size) %>% 
  arrange(type, stratum)
pop_size_7220 <-
  habspring_units %>% 
  st_drop_geometry %>% 
  count(system_type) %>% 
  pivot_wider(names_from = system_type,
              values_from = n) %>% 
  transmute(all = mire + rivulet + unknown,
            terrestrial = mire + unknown,
            aquatic = rivulet + unknown) %>% 
  mutate(type = "7220" %>% factor(levels = levels(pop_size_terr$type))) %>% 
  pivot_longer(cols = c("all", "terrestrial", "aquatic"),
               names_to = "stratum",
               values_to = "population_size")
pop_size_3260 <- 
  habstream %>% 
  mutate(length = st_length(.) %>% drop_units(),
         population_size = ceiling(length / 100) %>% as.integer) %>% 
  st_drop_geometry() %>% 
  summarise(type = first(type),
            stratum = first(NA_character_),
            population_size = sum(population_size))
pop_size_lentic <- 
  wsh_pol %>% 
  mutate(area = st_area(.) %>% set_units("ha")) %>% 
  st_drop_geometry %>% 
  select(polygon_id, area) %>% 
  mutate(stratum = 
           area %>% 
           cut(c(0, 1, 5, 50, 150)) %>% 
           as.character) %>% 
  inner_join(wsh_occ, by = "polygon_id", .) %>% 
  semi_join(targetpop, by = "type") %>% 
  count(type, stratum, name = "population_size")
pop_size_8310 <- 
  habquarries %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>% 
  st_drop_geometry %>% 
  select(type, area) %>% 
  mutate(stratum = 
           area %>% 
           cut(c(0, 2, 6, 40)) %>% 
           as.character) %>% 
  count(type, stratum, name = "population_size")
join_targetpop <- function(df) {
  df  %>% 
    inner_join(targetpop %>% 
                 select(scheme, 
                        type,
                        contains("typegroup")), 
               .,
               by = "type",
               relationship = "many-to-many") %>% 
    filter(type != "7220" |
             (type == "7220" & stratum == "all" & !str_detect(scheme, "aq|terr|SURF")) |
             (type == "7220" & stratum == "aquatic" & str_detect(scheme, "aq|SURF")) | 
             (type == "7220" & stratum == "terrestrial" & str_detect(scheme, "terr"))
    )
}
targetpop_stratumsize <- 
  bind_rows(pop_size_terr,
            pop_size_7220,
            pop_size_3260,
            pop_size_lentic,
            pop_size_8310) %>% 
  join_targetpop
```

```{r exploration-2, eval = FALSE, include=FALSE}
# targetpop_stratumsize comprises the whole targetpop!
targetpop %>% anti_join(targetpop_stratumsize, by = c("scheme", "type")) 
# targetpop_stratumsize has no empty population sizes!
targetpop_stratumsize %>% filter(is.na(population_size))
# targetpop_stratumsize has no type, stratum, population_size combinations that are different between schemes (it's obvious from the code above)
targetpop_stratumsize %>% 
  distinct(type, stratum, population_size) %>% 
  count(type, stratum) %>% 
  filter(n > 1)
# the type_stratum_props object created below is the object we use to explore and decide which types to select for prototyping
type_stratum_props <-
  targetpop_stratumsize %>% 
  distinct(type, stratum, population_size) %>% 
  arrange(type, stratum) %>%
  inner_join(
    n2khab_types_expanded_properties,
    by = "type"
  )
```


## Writing the base sampling frame

## Defining all inputs to draw samples

## Reading the base sampling frame (and other things), including phab-corrections


