# Creating the base sampling frame and other essential starting points (datasets) to draw samples

## Creating elements of the base sampling frame

Summarizing the target populations of all schemes in MNE & MHQ (prefix `n2khab_`) to an overall list of types, including the expansion to main types or subtypes for each scheme for better spatial joining with `habitatmap_terr`, which is a 'best possible guess' of actually present terrestrial types (see its documentation).

```{r}
n2khab_targetpop <-
  read_scheme_types() %>%
  select(scheme, type)
n2khab_targetpop_expanded <-
  n2khab_targetpop %>%
  group_by(scheme) %>%
  expand_types() %>%
  ungroup()
n2khab_types_expanded <-
  n2khab_targetpop_expanded %>%
  distinct(type) %>%
  arrange(type)
```

The base sampling frame will include a layer for each of the resulting main types or subtypes.

Since the target populations of each scheme are defined by non-overlapping type codes (never combining a main type with its subtypes), when deriving sampling frames for each scheme, subtypes that were added by the code above will be merged again into their main types and added (but unresolved) main types will be added (repeated) to each of its corresponding subtype sampling frame.

This way, the sampling frames will present the exact set of type codes as in `targetpop`, while taking advantage of the type expansion since all codes added by expansion will be present in the base sampling frame.

```{r eval=FALSE, include=FALSE}
# various main type codes are not added because habitatmap_terr doesn't have
# them OR they're not involved in MNE/MHQ anyway
read_types(lang = lang) %>% 
  anti_join(n2khab_types_expanded, by = "type")
# following types were added by expansion:
n2khab_types_expanded %>%
    anti_join(n2khab_targetpop %>% distinct(type), by = "type") %>% 
    inner_join(read_types() %>% select(type, typelevel),
               by = "type")
```

Gathering minimal needed properties of types.

```{r}
n2khab_types_expanded_properties <- 
  read_vc(
    "030_preparations/010_explore_targetpop/type_approaches",
    root = find_root(is_git_root)
  ) %>%
  as_tibble() %>% 
  mutate(
    grts_join_method = ifelse(approach_cellbased, "cell", "centroid") %>% 
      factor(levels = c("centroid", "cell", "point"))
  ) %>% 
  select(type, grts_join_method)%>%
  left_join(
    read_types() %>% 
      semi_join(n2khab_types_expanded, by = "type") %>% 
      select(type, typelevel, main_type, typeclass, hydr_class, groundw_dep, flood_dep),
    .,
    by = "type"
  )
```


<!-- We lezen de ruimtelijke data in van de terrestrische types (incl. 'semi-terrestrisch'), met uitsluiting van resterende aquatische types (binnen deze potentieel terrestrische polygonen). -->

```{r}
hmt <- read_habitatmap_terr(keep_aq_types = FALSE, drop_7220 = TRUE)
hmt_occ <-
    hmt$habitatmap_terr_types %>%
    semi_join(n2khab_types_expanded,
              by = "type")
hmt_pol <-
    hmt$habitatmap_terr_polygons %>%
    semi_join(hmt_occ, by = "polygon_id") %>% 
    select(polygon_id) %>% 
    st_make_valid
```

<!-- - We doen hetzelfde voor de stilstaande wateren (d.w.z. aquatische types uitgezonderd 3260 en 7220 partim beken): -->

```{r}
wsh <- read_watersurfaces_hab(interpreted = TRUE)
wsh_occ <- 
    wsh$watersurfaces_types %>% 
    semi_join(n2khab_types_expanded,
              by = "type")
wsh_pol <-
    wsh$watersurfaces_polygons %>%
    semi_join(wsh_occ, by = "polygon_id") %>% 
    select(polygon_id)
```

<!-- - We lezen de spatiale data ook in voor types 3260, 7220 en 8310. -->

```{r}
habstream <- read_habitatstreams()
```


```{r}
flanders_buffer <- 
    read_admin_areas(dsn = "flanders") %>% 
    st_buffer(40)
habspring_units <- 
  read_habitatsprings(units_7220 = TRUE) %>% 
  .[flanders_buffer, ]
```


```{r}
habquarries <- read_habitatquarries(filter_hab = TRUE) # is restricted to Flanders
```

<!-- Below chunks go back to targetpop, for exploratory purposes. Best move and refer them... -->

```{r exploration-1, eval = FALSE, include=FALSE}
# building blocks for the exploration which types to select from targetpop for the POC

pop_size_terr <- 
  read_vc(
    "030_preparations/010_explore_targetpop/pop_size_terr", 
    root = find_root(is_git_root)
  ) %>% 
  as_tibble() %>% 
  inner_join(
    read_vc(
      "030_preparations/010_explore_targetpop/type_approaches",
      root = find_root(is_git_root)
    ),
    by = "type") %>% 
  filter(count_method == "cellbased" & approach_cellbased |
           count_method == "centroidbased" & !approach_cellbased) %>% 
  select(-c(count_method, approach_cellbased)) %>% 
  mutate(stratum = NA_character_,
         population_size = round(population_size) %>% as.integer) %>% 
  select(type, stratum, population_size) %>% 
  arrange(type, stratum)
pop_size_7220 <-
  habspring_units %>% 
  st_drop_geometry %>% 
  count(system_type) %>% 
  pivot_wider(names_from = system_type,
              values_from = n) %>% 
  transmute(all = mire + rivulet + unknown,
            terrestrial = mire + unknown,
            aquatic = rivulet + unknown) %>% 
  mutate(type = "7220" %>% factor(levels = levels(pop_size_terr$type))) %>% 
  pivot_longer(cols = c("all", "terrestrial", "aquatic"),
               names_to = "stratum",
               values_to = "population_size")
pop_size_3260 <- 
  habstream %>% 
  mutate(length = st_length(.) %>% drop_units(),
         population_size = ceiling(length / 100) %>% as.integer) %>% 
  st_drop_geometry() %>% 
  summarise(type = first(type),
            stratum = first(NA_character_),
            population_size = sum(population_size))
pop_size_lentic <- 
  wsh_pol %>% 
  mutate(area = st_area(.) %>% set_units("ha")) %>% 
  st_drop_geometry %>% 
  select(polygon_id, area) %>% 
  mutate(stratum = 
           area %>% 
           cut(c(0, 1, 5, 50, 150)) %>% 
           as.character) %>% 
  inner_join(wsh_occ, by = "polygon_id", .) %>% 
  semi_join(targetpop, by = "type") %>% 
  count(type, stratum, name = "population_size")
pop_size_8310 <- 
  habquarries %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>% 
  st_drop_geometry %>% 
  select(type, area) %>% 
  mutate(stratum = 
           area %>% 
           cut(c(0, 2, 6, 40)) %>% 
           as.character) %>% 
  count(type, stratum, name = "population_size")
join_targetpop <- function(df) {
  df  %>% 
    inner_join(targetpop %>% 
                 select(scheme, 
                        type,
                        contains("typegroup")), 
               .,
               by = "type",
               relationship = "many-to-many") %>% 
    filter(type != "7220" |
             (type == "7220" & stratum == "all" & !str_detect(scheme, "aq|terr|SURF")) |
             (type == "7220" & stratum == "aquatic" & str_detect(scheme, "aq|SURF")) | 
             (type == "7220" & stratum == "terrestrial" & str_detect(scheme, "terr"))
    )
}
targetpop_stratumsize <- 
  bind_rows(pop_size_terr,
            pop_size_7220,
            pop_size_3260,
            pop_size_lentic,
            pop_size_8310) %>% 
  join_targetpop
```

```{r exploration-2, eval = FALSE, include=FALSE}
# targetpop_stratumsize comprises the whole targetpop!
targetpop %>% anti_join(targetpop_stratumsize, by = c("scheme", "type")) 
# targetpop_stratumsize has no empty population sizes!
targetpop_stratumsize %>% filter(is.na(population_size))
# targetpop_stratumsize has no type, stratum, population_size combinations that are different between schemes (it's obvious from the code above)
targetpop_stratumsize %>% 
  distinct(type, stratum, population_size) %>% 
  count(type, stratum) %>% 
  filter(n > 1)
# the type_stratum_props object created below is the object we use to explore and decide which types to select for prototyping
type_stratum_props <-
  targetpop_stratumsize %>% 
  distinct(type, stratum, population_size) %>% 
  arrange(type, stratum) %>%
  inner_join(
    n2khab_types_expanded_properties,
    by = "type"
  )
```


## Writing the base sampling frame

## Defining all inputs to draw samples

## Reading the base sampling frame (and other things), including phab-corrections


