# Creating the base sampling frame

## Preparing elements of the base sampling frame

### Types to be distinguished and their properties

Summarizing the target populations of all schemes in MNE & MHQ (prefix `n2khab_`) as an overall list of types, including the expansion to main types or subtypes for each scheme for better spatial joining with `habitatmap_terr`, which is a 'best possible guess' of actually present terrestrial types (see its documentation).

```{r}
n2khab_targetpops <-
  read_scheme_types() %>%
  select(scheme, type)
n2khab_types <-
  n2khab_targetpops %>%
  distinct(type) %>%
  arrange(type)
n2khab_targetpops_expanded <-
  n2khab_targetpops %>%
  group_by(scheme) %>%
  expand_types(strict = FALSE) %>%
  ungroup()
n2khab_types_expanded <-
  n2khab_targetpops_expanded %>%
  distinct(type) %>%
  arrange(type)
```

The base sampling frame will include a layer for each of the resulting main types or subtypes.
More correctly, a layer is to be made for each stratum (see further), which can be further subdivisions of types.

Since the target population of each scheme is defined by non-overlapping type codes (never combining a main type with its subtypes), when deriving sampling frames for each scheme:

- locations of subtypes that were added by the code above will be merged in the sampling frame layer of their main type
- locations of an added (but unresolved) main type will be added (repeated) in each of its corresponding subtype sampling frame layers.
This only applies to subtypes that triggered the addition of the main type (see `?n2khab::expand_types(strict = FALSE)`).

This way, the sampling frames will present the exact set of type codes as in `targetpops`, while taking advantage of the type expansion since all codes added by expansion will be present in the base sampling frame.

In the ideal situation, the base sampling frame is 'resolved' by terrain visits by replacing main type codes by the correct subtype code, for main types that do have subtypes.

```{r eval=FALSE, include=FALSE}
# various main type codes are not added because habitatmap_terr doesn't have
# them OR they're not involved in MNE/MHQ anyway
read_types(lang = lang) %>%
  anti_join(n2khab_types_expanded, by = "type")
# following types were added by expansion:
n2khab_types_expanded %>%
  anti_join(n2khab_types, by = "type") %>%
  inner_join(read_types() %>% select(type, typelevel),
    by = "type"
  )
```

Gathering minimal needed properties of all types.

```{r}
n2khab_types_expanded_properties <-
  read_types() %>%
  semi_join(n2khab_types_expanded, by = "type") %>%
  select(type, typelevel, main_type, typeclass, hydr_class, groundw_dep, flood_dep) %>%
  left_join(
    read_vc(
      "030_preparations/010_explore_targetpop/type_approaches",
      root = gitroot
    ) %>%
      as_tibble() %>%
      mutate(
        grts_join_method = ifelse(approach_cellbased, "cell", "cell_center")
      ) %>%
      select(type, grts_join_method),
    by = "type"
  ) %>%
  mutate(
    grts_join_method = case_when(
      !is.na(grts_join_method) ~ grts_join_method,
      type == "7220" ~ "original_point",
      type == "3260" ~ "endpoint_of_segment",
      type == "8310" | hydr_class == "HC3" ~ "point_on_surface",
      .default = "cell_center"
    ) %>%
      factor(levels = c(
        "cell_center",
        "cell",
        "point_on_surface",
        "endpoint_of_segment",
        "original_point"
      )),
    sample_support = case_when(
      grts_join_method == "cell_center" ~ "area of type within 1024 m² cell (cell center must be type)",
      grts_join_method == "cell" ~ "area of type within 1024 m² cell",
      type == "3260" ~ "predefined watercourse segment (mostly 100 m)",
      type == "7220" ~ "population unit inferred from read_habitatsprings(units_7220 = TRUE)",
      type == "8310" ~ "population unit as defined by read_habitatquarries(filter_hab = TRUE)",
      .default = "watersurface"
    ) %>%
      factor(),
    sample_support_code = fct_recode(
      sample_support,
      cell_conditioned_on_center = "area of type within 1024 m² cell (cell center must be type)",
      cell = "area of type within 1024 m² cell",
      watercourse_segment = "predefined watercourse segment (mostly 100 m)",
      spring = "population unit inferred from read_habitatsprings(units_7220 = TRUE)",
      quarry = "population unit as defined by read_habitatquarries(filter_hab = TRUE)",
    )
  )
```

### Reading and filtering spatial data of the target populations

Reading and filtering spatial data of terrestrial types.

```{r}
hmt <- read_habitatmap_terr(keep_aq_types = FALSE, drop_7220 = TRUE)
typelev <- levels(hmt$habitatmap_terr_types$type)
hmt_occ <-
  hmt$habitatmap_terr_types %>%
  semi_join(n2khab_types_expanded,
    by = "type"
  ) %>% 
  # manually fixing the type of two polygons in habitatmap_terr_2023_v1:
  mutate(
    source = case_when(
      polygon_id == "184091_v2023" & type == "2190_overig" ~ "habitatmap_stdized + manual fix",
      polygon_id == "533826_v2023" & type == "2310" ~ "habitatmap_stdized + manual fix",
      .default = source
    ) %>% 
      factor(),
    type = case_when(
      polygon_id == "184091_v2023" & type == "2190_overig" ~ "9120",
      polygon_id == "533826_v2023" & type == "2310" ~ "2130",
      .default = type
    ) %>% 
      factor(levels = typelev)
  )
hmt_pol <-
  hmt$habitatmap_terr_polygons %>%
  semi_join(hmt_occ, by = "polygon_id") %>%
  select(polygon_id) %>%
  st_make_valid()
```

Reading and filtering spatial data of lentic aquatic types.

```{r}
wsh <- read_watersurfaces_hab(interpreted = TRUE)
wsh_occ <-
  wsh$watersurfaces_types %>%
  semi_join(n2khab_types_expanded,
    by = "type"
  )
wsh_pol <-
  wsh$watersurfaces_polygons %>%
  semi_join(wsh_occ, by = "polygon_id") %>%
  select(polygon_id)
```

Reading spatial data of type 3260.

<!-- maar beter ineens de afgeleide dataset opbouwen / inlezen met de segmenten: betrek daarbij het voorbereidend werk van Tom DD en vlgs googledoc steekproefkader is dit iets voor Toon W, incl koppeling met punten -->

```{r}
habstream <- read_habitatstreams()
```

Reading spatial data of type 7220.
Each row (point) already represents a population unit in this case.

```{r}
flanders_buffer <-
  read_admin_areas(dsn = "flanders") %>%
  st_buffer(40)
habspring_units <-
  read_habitatsprings(units_7220 = TRUE) %>%
  .[flanders_buffer, ]
```

Reading spatial data of type 8310.
Each row (polygon) does not necessarily represent a population unit, but the latter is indicated by the `unit_id` variable.
However in this case the filtering by 8310 already results in just one polygon per unit.

```{r}
habquarries <-
  read_habitatquarries(filter_hab = TRUE)
```

### Defining the strata used in sampling and inference

The strata are the types, in some cases subdivided according to area class of the population unit.
The base sampling frame layers will consist of the strata, not just the types.

```{r prepare-strata}
extract_interval_limits <- function(df) {
  mutate(
    df,
    lower = str_match(stratum, "^\\((\\d+),\\d+\\]$")[, 2] %>% as.integer(),
    upper = str_match(stratum, "^\\(\\d+,(\\d+)\\]$")[, 2] %>% as.integer()
  )
}
wsh_pol_area_class <-
  wsh_pol %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>%
  st_drop_geometry() %>%
  select(polygon_id, area) %>%
  mutate(
    stratum =
      area %>%
        cut(c(0, 1, 5, 50, 150)) %>%
        as.character()
  ) %>%
  inner_join(
    wsh_occ %>%
      select(polygon_id, type),
    .,
    by = "polygon_id"
  ) %>%
  rename(unit_id = polygon_id)
wsh_strata <-
  wsh_pol_area_class %>%
  extract_interval_limits() %>%
  distinct(type, lower, upper) %>%
  arrange(type, lower, upper)
habquarries_area_class <-
  habquarries %>%
  mutate(area = st_area(.) %>% set_units("ha")) %>%
  st_drop_geometry() %>%
  select(unit_id, type, area) %>%
  mutate(
    stratum =
      area %>%
        cut(c(0, 2, 6, 40)) %>%
        as.character()
  ) %>%
  mutate(unit_id = as.character(unit_id))
habquarries_strata <-
  habquarries_area_class %>%
  extract_interval_limits() %>%
  distinct(type, lower, upper) %>%
  arrange(type, lower, upper)
```


```{r result-strata}
# n2khab_strata_expanded: defines the strata of the base sampling frame
n2khab_strata_expanded <-
  n2khab_types_expanded %>%
  left_join(
    bind_rows(wsh_strata, habquarries_strata),
    by = "type"
  ) %>%
  unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
  select(type, stratum) %>%
  mutate(stratum = factor(stratum))
# n2khab_strata: for sampling and inference from the spatial sampling frames,
# which it defines (except for further needed spatial restrictions)
n2khab_strata <-
  n2khab_types %>%
  left_join(
    bind_rows(wsh_strata, habquarries_strata),
    by = "type"
  ) %>%
  unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
  select(type, stratum) %>%
  mutate(
    stratum = factor(stratum, levels = levels(n2khab_strata_expanded$stratum))
  )
```

### Data sources with the population unit geometries

Reading the `GRTSmaster_habitats` data source that defines the population unit geometries of the terrestrial types.

```{r}
grts_mh <- read_GRTSmh()
```

For watersurface types, the unit geometries are defined by the `watersurfaces` data source (units are polygons).

For type 3260, the unit geometries are defined by `read_watercourse_100mseg(element = "lines")` (units are segments).

For type 7220, we already created the `habspring_units` object (undefined units referred to by points).

For type 8310, we already created the `habquarries` object (units are polygons).


### Defining the point - geometry relationship for specific types

These may encompass units not currently covered by the target populations.
This is especially true for watersurfaces and watersegments.
This allows for dynamic changes in the occurrence of types.

We don't need points for `grts_join_method == "cell"`.

We will use cell centers on the fly for `grts_join_method == "cell_center"`.

Defining points that represent the watersurface units (object `points_watsurf`).
This step requires the availability of data that are currently external to this repository and used packages.

```{r}
if (curl::has_internet()) {
  # Accessing a temporary data source for watersurfaces
  with_drive_quiet({
    path <- file.path(tempdir(), "watersurfaces_refpoints.tsv")
    drive_download(as_id("1XgtdZOebbqp-MJih60P_Vi-npz_5UVkY"), path = path)
    path2 <- file.path(tempdir(), "watersurfaces_refpoints.yml")
    drive_download(as_id("1Bb07utgZOyMK65TdR64-4VjGP7T6GqMa"), path = path2)
  })
  points_watsurf <- 
    read_vc("watersurfaces_refpoints", tempdir()) %>%
    as_tibble()
  invisible(file.remove(path, path2))
} else {
  # Clumsy approach; needs maintenance. In the longer term, the management of
  # all this needs more centralization and rigidity.
  mhq_ssfws_datapath <- file.path(
    dirname(gitroot),
    "n2khab-samplingframes/data/watersurfaces"
  )
  stopifnot(dir.exists(mhq_ssfws_datapath))
  points_watsurf <- 
    read_vc("watersurfaces_refpoints", root = mhq_ssfws_datapath) %>%
    as_tibble()
}
points_watsurf <- 
  points_watsurf %>%
  select(polygon_id, x, y) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370)
# adding all other watersurfaces from wsh_pol (note that these also have unit_id
# values not present in read_watersurfaces())
points_watsurf <-
  bind_rows(
    points_watsurf,
    wsh_pol %>%
      filter(!(polygon_id %in% points_watsurf$polygon_id)) %>%
      select(polygon_id) %>%
      st_sf(agr = "identity") %>%
      st_point_on_surface() %>%
      rename(geometry = geom)
  )
```

Defining points that represent the watercourse segments (object `points_watcourse`).

```{r}
points_watcourse <- read_watercourse_100mseg(element = "points")
```

For type 7220, we can use the points in `habspring_units`.

Defining points that represent the type 8310 units (object `points_8310`).

```{r}
points_8310 <-
  habquarries %>%
  select(unit_id) %>%
  st_sf(agr = "identity") %>%
  st_point_on_surface()
```

Combining all points that refer to non-cell units into one object.

```{r}
points_non_cell <-
  bind_cols(
    tibble(sample_support_code = "watersurface"),
    points_watsurf %>% rename(unit_id = polygon_id)
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "watercourse_segment"),
      points_watcourse %>%
        unite(unit_id, vhag_code, rank) %>%
        rename(geometry = geom)
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "spring"),
      habspring_units %>%
        transmute(unit_id = as.character(unit_id))
    )
  ) %>%
  bind_rows(
    bind_cols(tibble(sample_support_code = "quarry"), points_8310) %>%
      mutate(unit_id = as.character(unit_id)) %>%
      rename(geometry = geom)
  ) %>%
  mutate(sample_support_code = factor(
    sample_support_code,
    levels = levels(n2khab_types_expanded_properties$sample_support_code)
  )) %>%
  st_sf(sf_column_name = "geometry")
```

The `sample_support_code` needs to be combined with `unit_id` to be globally unique!
Alternatively, they could be combined into a single column with unique IDs, but then it would loose connection with the corresponding columns of spatial data sources.

## Assigning a GRTS address to all population units of each stratum

The end result of this, `stratum_grts_n2khab`, is what constitutes the essence of the base sampling frame.

Mind that we will make further updates to the base sampling frame in section \@ref(update-bsf).

However in its full meaning the base sampling frame also comprises various linked attributes generated before.

### For the non-cell units

From `points_non_cell`, we need to filter those that fall within the target populations of MNE and MHQ.
This is already fine for types 7220 and 8310.

We determine the watercourse segments that have type 3260.

```{r}
# Quick-and-dirty approach to define the 3260 segments (i.e. it will miss a
# part and some may be false positives)
segm_3260 <- read_watercourse_100mseg(element = "lines")[habstream, ] %>%
  unite(unit_id, vhag_code, rank)
```

The watersurfaces with types are defined by `wsh_pol`; the segments with 3260 are defined by `segm_3260`.

```{r}
points_non_cell_n2khab <-
  points_non_cell %>%
  filter(
    sample_support_code %in% c("spring", "quarry") |
      (sample_support_code == "watersurface" & unit_id %in% wsh_pol$polygon_id) |
      (sample_support_code == "watercourse_segment" & unit_id %in% segm_3260$unit_id)
  )
```

Determining the GRTS address of the units represented by `points_non_cell_n2khab`.

```{r}
units_non_cell_n2khab_grts <-
  points_non_cell_n2khab %>%
  mutate(
    grts_address = extract(grts_mh, st_coordinates(points_non_cell_n2khab)) %>%
      pull(GRTSmaster_habitats)
  ) %>%
  st_drop_geometry()
```

<!-- Er zijn gevallen met hetzelfde GRTS-adres; een oplossing is hiervoor uit te werken: -->
<!-- > units_non_cell_n2khab_grts %>% count(sample_support_code, grts_address) %>% filter(n > 1) %>% count(sample_support_code, n) -->
<!-- Storing counts in `nn`, as `n` already present in input -->
<!-- ℹ Use `name = "new_name"` to pick a new name. -->
<!-- # A tibble: 7 × 3 -->
<!--   sample_support_code     n    nn -->
<!--   <fct>               <int> <int> -->
<!-- 1 watercourse_segment     2    21 -->
<!-- 2 watercourse_segment     3     1 -->
<!-- 3 watercourse_segment   122     1 -->
<!-- 4 watersurface            2    54 -->
<!-- 5 watersurface            3    13 -->
<!-- 6 watersurface            4     8 -->
<!-- 7 watersurface            5     2 -->

Listing the non-cell units per stratum (within the target populations of MNE and MHQ), needed in order to connect the GRTS addresses with strata.

```{r}
stratum_units_non_cell_n2khab <-
  bind_cols(
    tibble(sample_support_code = "watersurface"),
    wsh_pol_area_class %>%
      distinct() %>%
      extract_interval_limits() %>%
      unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
      select(stratum, unit_id)
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "watercourse_segment"),
      segm_3260 %>%
        st_drop_geometry() %>%
        mutate(stratum = "3260", .before = everything())
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "spring"),
      habspring_units %>%
        st_drop_geometry() %>%
        mutate(
          stratum = "7220",
          unit_id = as.character(unit_id),
          .keep = "none",
          .before = everything()
        )
    )
  ) %>%
  bind_rows(
    bind_cols(
      tibble(sample_support_code = "quarry"),
      habquarries_area_class %>%
        extract_interval_limits() %>%
        unite(stratum, c(type, lower, upper), remove = FALSE, na.rm = TRUE) %>%
        select(stratum, unit_id)
    )
  ) %>%
  mutate(
    sample_support_code = factor(
      sample_support_code,
      levels = levels(n2khab_types_expanded_properties$sample_support_code)
    ),
    stratum = factor(stratum, levels = levels(n2khab_strata_expanded$stratum))
  ) %>%
  relocate(stratum)
```

Note that `stratum_units_non_cell_n2khab` has `r nrow(stratum_units_non_cell_n2khab)` rows which is larger than the `r nrow(units_non_cell_n2khab_grts)` rows of `units_non_cell_n2khab_grts`.
This is because some watersurface units are attached to more than 1 stratum (i.e. more than 1 type).

Assigning the GRTS addresses of the units to the corresponding strata in `n2khab_strata_expanded`.

```{r}
stratum_grts_non_cell_n2khab <-
  stratum_units_non_cell_n2khab %>%
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>%
  select(-sample_support_code, -unit_id) %>%
  # distinct() leads to less rows because of units with same GRTS address.
  distinct()
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_non_cell_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_non_cell_n2khab$grts_address)
```

### For the cell units

Filtering units of `GRTSmaster_habitats` that are part of the target populations of MNE and MHQ with sample support `cell_conditioned_on_center`, and storing their polygon ID and GRTS address.

```{r}
hmt_pol_cellcenter <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell_conditioned_on_center") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  semi_join(hmt_pol, ., by = "polygon_id") %>%
  vect()
```


```{r eval = params$regenerate_binary_files}
# takes < 1 minute:
hmt_pol_grts_cellcenter <-
  extract(grts_mh, hmt_pol_cellcenter, small = FALSE) %>%
  as_tibble() %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(hmt_pol_cellcenter)),
      polygon_id = hmt_pol_cellcenter$polygon_id
    ),
    .,
    by = "ID"
  ) %>%
  select(-ID, grts_address = GRTSmaster_habitats) %>%
  # filtering is needed since all polygons are listed by extract():
  filter(!is.na(grts_address))
saveRDS(
  hmt_pol_grts_cellcenter,
  file.path(datapath, "binary/intermediate/hmt_pol_grts_cellcenter.rds")
)
```

```{r}
hmt_pol_grts_cellcenter <-
  readRDS(file.path(
    datapath,
    "binary/intermediate/hmt_pol_grts_cellcenter.rds"
  ))
```

Filtering units of GRTSmaster_habitats that are part of the target populations of MNE and MHQ with sample support `cell`, and storing their polygon ID and GRTS address.

```{r}
hmt_pol_cell <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  semi_join(hmt_pol, ., by = "polygon_id") %>%
  vect()

hmt_pol_grts_cell <-
  # touches = TRUE allows about 2.5 times the number of cells
  extract(grts_mh, hmt_pol_cell, touches = TRUE) %>%
  as_tibble() %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(hmt_pol_cell)),
      polygon_id = hmt_pol_cell$polygon_id
    ),
    .,
    by = "ID"
  ) %>%
  select(-ID, grts_address = GRTSmaster_habitats)
```

Note that `hmt_pol_grts_cellcenter` and `hmt_pol_grts_cell` share a number of polygons so this should be used with care, i.e. not mixing the link of `polygon_id` with `grts_address` between cell-center vs. cell types.

```{r}
stratum_polygons_cell_all_n2khab <-
  n2khab_types_expanded_properties %>%
  filter(str_detect(sample_support_code, "^cell")) %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(stratum = type, polygon_id)
```

<!-- Evalueer of er daarom niet beter ook een object wordt gemaakt zoals `stratum_units_non_cell_n2khab`, maar dan een die de _polygonen_ oplijst per stratum (type) aangezien er geen unit_id's zijn; dus zou `stratum_polygons_cell_all_n2khab` heten. -->
<!-- Dit is inmiddels hierboven gedaan (wordt niet verder gebruikt); te bekijken of dit zinvol is. -->

Assigning the GRTS addresses of the units to the corresponding strata in `n2khab_strata_expanded`.

We can use `n2khab_types_expanded` since for terrestrial types the strata are the types without subdivision.

```{r}
hmt_pol_stratum_grts_cellcenter <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell_conditioned_on_center") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(polygon_id, stratum = type) %>%
  inner_join(
    hmt_pol_grts_cellcenter,
    by = "polygon_id",
    relationship = "many-to-many"
  )
stratum_grts_cellcenter_n2khab <-
  hmt_pol_stratum_grts_cellcenter %>%
  select(-polygon_id)
hmt_pol_stratum_grts_cell <-
  n2khab_types_expanded_properties %>%
  filter(sample_support_code == "cell") %>%
  semi_join(distinct(hmt_occ, polygon_id, type), ., by = "type") %>%
  select(polygon_id, stratum = type) %>%
  inner_join(
    hmt_pol_grts_cell,
    by = "polygon_id",
    relationship = "many-to-many"
  )
stratum_grts_cell_n2khab <-
  hmt_pol_stratum_grts_cell %>%
  select(-polygon_id) %>%
  # adjacent polygons can share the same type AND grts_address here, so:
  distinct()
stratum_grts_cell_all_n2khab <-
  bind_rows(stratum_grts_cellcenter_n2khab, stratum_grts_cell_n2khab)
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_cell_all_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_cell_all_n2khab$grts_address)
```

### Integrate the GRTS addresses for cell and non-cell units

```{r}
stratum_grts_n2khab <-
  stratum_grts_non_cell_n2khab %>%
  bind_rows(stratum_grts_cell_all_n2khab) %>% 
  mutate(
    typelevel_certain = NA,
    assessed_in_field = FALSE,
    assessment_date = NA_Date_,
    inaccessible = factor(NA, levels = c("short term", "long term")),
    replaced_by = NA_integer_,
    replacement_for = NA_integer_ 
  ) %>% 
  relocate(typelevel_certain, .after = stratum) %>% 
  relocate(assessed_in_field, .after = grts_address)
```

We also added some extra attributes that prepare for integrating information on assessments and local replacements.

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_n2khab))
)
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_n2khab$grts_address)
```

```{r paged.print = FALSE}
stratum_grts_n2khab %>% filter(is.na(grts_address))
```

<!-- REVISIT BELOW FIX! XXXXXXXXXXXXXXXXXX -->

Temporary fix:

```{r}
stratum_grts_n2khab <-
  stratum_grts_n2khab %>%
  filter(!is.na(grts_address))
```


## Updating the base sampling frame based on existing field assessments {#update-bsf}

The aim is to add existing field assessments to the base sampling frame.

These steps require the availability of data that are currently external to this repository and used packages.

```{r}
# A bit clumsy; needs maintenance. In the longer term, the management of all
# this needs more centralization and rigidity.
mhq_terr_datapath <- file.path(
  dirname(gitroot),
  "n2khab-sample-admin/data/mhq_terr/rapportage2025"
)
mhq_watersurfaces_datapath <- file.path(
  dirname(gitroot),
  "n2khab-sample-admin/data/mhq_watersurfaces/rapportage2025"
)
mhq_data_available <- 
  dir.exists(mhq_terr_datapath) && dir.exists(mhq_watersurfaces_datapath)
```

These data are `r if (!mhq_data_available) "_not_"` available.

```{r eval = mhq_data_available}
mhq_terr_popunits <-
  read_vc("mhq_terr_popunits", root = mhq_terr_datapath) %>%
  as_tibble() %>% 
  mutate(
    type = factor(type, levels = levels(n2khab_types$type)),
    source = factor(source),
    across(where(is.numeric), as.integer)
  ) %>% 
  select(-sac, -phab, -polygon_id) %>% 
  rename(
    grts_address = grts_ranking,
    grts_address_drawn = grts_ranking_draw
  )
mhq_terr_refpoints <-
  read_vc("mhq_terr_refpoints", root = mhq_terr_datapath) %>%
  as_tibble() %>% 
  distinct() %>% 
  select(
    point_code,
    grts_address = grts_ranking,
    is_cell_center = is_centroid
  )
mhq_terr_assessments <-
  read_vc("mhq_terr_assessments", root = mhq_terr_datapath) %>%
  as_tibble() %>% 
  filter(assessment_date == max(assessment_date), .by = point_code) %>%
  select(point_code, assessment_date, type, is_present, inaccessible) %>% 
  mutate(
    inaccessible = factor(inaccessible, levels = c("short term", "long term"))
  )
mhq_watersurfaces_assessments <-
  read_vc("mhq_watersurfaces_assessments", root = mhq_watersurfaces_datapath) %>%
  as_tibble()
mhq_watersurfaces_populationunits <-
  read_vc("mhq_watersurfaces_populationunits", root = mhq_watersurfaces_datapath) %>%
  as_tibble()
```

### Terrestrial field assessments

#### Valid terrestrial population units

Determining a list of valid terrestrial population units to update the base sampling frame.

```{r eval = mhq_data_available}
valid_extra_popunits_cell_all_n2khab <- 
  mhq_terr_popunits %>% 
  # only consider assessment data:
  filter(str_detect(source, "assessment")) %>% 
  inner_join(
    read_types() %>% 
      select(type, typelevel, hydr_class),
    join_by(type),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # expansion of main types to subtypes is needed for compatibility with MNE/MHQ
  # target populations; those main types are subsequently removed. We don't
  # expand subtypes to main types (that is only relevant in joining with some
  # geospatial layers)
  expand_types(mark = TRUE) %>% 
  filter(
    !(expanded & typelevel == "main_type"),
    is.na(expanded_to) | expanded_to != "main_type"
  ) %>% 
  select(-typelevel, -expanded, -source) %>% 
  # at the time of coding, 2 records from lhs needed to be dropped because
  # they are not cell-based types (but aquatic)
  filter(!(hydr_class %in% c("HC3"))) %>% 
  # adding grts_join_method (either cell or cell_center)
  inner_join(
    n2khab_types_expanded_properties %>% 
      select(type, grts_join_method) %>% 
      filter(str_detect(grts_join_method, "cell")),
    join_by(type),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # include whether a location is assessed at the cell center
  inner_join(
    mhq_terr_refpoints %>% 
      select(point_code, is_cell_center),
    join_by(point_code),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>%
  left_join(
    mhq_terr_assessments %>% 
      filter(is_present) %>% 
      select(-is_present),
    join_by(point_code, type),
    relationship = "one-to-one",
    unmatched = "drop"
  ) %>% 
  select(-point_code, -hydr_class) %>% 
  mutate(assessed_in_field = TRUE) %>% 
  filter(
    # don't include legacy sites except if grts_join_method is cell (also, note
    # that all 'bosinventarisatie' sites are considered legacy sites and are
    # never assessed at the cell center). At the time of coding such combination
    # did not exist BTW.
    !legacy_site | grts_join_method == "cell",
    # require that cell_center-joined types have an assessment at the cell
    # center
    grts_join_method == "cell_center" & is_cell_center |
      grts_join_method == "cell"
  ) %>% 
  # since the base sampling frame uses strata, not types, we must align
  rename(stratum = type) %>% 
  mutate(
    typelevel_certain = is.na(expanded_to),
    stratum = factor(stratum, levels = levels(n2khab_strata$stratum))
  ) %>% 
  select(-expanded_to, -grts_join_method, -legacy_site, -is_cell_center) %>% 
  relocate(grts_address_drawn) %>% 
  relocate(
    assessed_in_field,
    assessment_date,
    inaccessible,
    .after = last_col()
  )
```

#### Sanity checks and temporary fixes {#mhqterr-fixes}

Checking that non-replaced GRTS addresses for cell-center types are unique.

```{r eval = mhq_data_available}
stopifnot(
  valid_extra_popunits_cell_all_n2khab %>% 
    inner_join(
      n2khab_types_expanded_properties %>% 
        select(type, grts_join_method),
      join_by(stratum == type),
      relationship = "many-to-one",
      unmatched = c("error", "drop")
    ) %>% 
    filter(
      grts_address_drawn == grts_address,
      grts_join_method == "cell_center"
    ) %>% 
    count(grts_address_drawn) %>% 
    filter(n > 1) %>% 
    nrow() == 0
)
```

For each _drawn_ GRTS address x type, we expect only one occurrence.
Listing violations for further enquiry.

```{r paged.print = FALSE, eval = mhq_data_available}
grtsaddrdrawn_duplicates_in_stratum <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  count(grts_address_drawn, stratum) %>% 
  filter(n > 1) %>% 
  semi_join(
    valid_extra_popunits_cell_all_n2khab, 
    .,
    join_by(grts_address_drawn, stratum)
  )
grtsaddrdrawn_duplicates_in_stratum
```

Applying a temporary fix.

<!-- XXXXXX To be revisited! -->

```{r eval = mhq_data_available && nrow(grtsaddrdrawn_duplicates_in_stratum) > 0}
valid_extra_popunits_cell_all_n2khab <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  # from the duplicates, only keep the one with the highest GRTS address
  anti_join(
    grtsaddrdrawn_duplicates_in_stratum %>% 
      filter(
        grts_address != max(grts_address), 
        .by = c(grts_address_drawn, stratum)
      ),
    join_by(grts_address_drawn, grts_address, stratum)
  )
```

For each _real_ GRTS address x type, we expect only one occurrence.
Listing violations for further enquiry.

```{r paged.print = FALSE, eval = mhq_data_available}
grtsaddr_duplicates_in_stratum <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  count(grts_address, stratum) %>% 
  filter(n > 1) %>% 
  semi_join(
    valid_extra_popunits_cell_all_n2khab, 
    .,
    join_by(grts_address, stratum)
  ) %>% 
  arrange(grts_address)
grtsaddr_duplicates_in_stratum
```

Applying a temporary fix.

<!-- XXXXXX To be revisited! -->

```{r eval = mhq_data_available && nrow(grtsaddr_duplicates_in_stratum) > 0}
valid_extra_popunits_cell_all_n2khab <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  # don't include the duplicates with grts_address != grts_address_drawn; we
  # don't consider them as an actual local replacement
  anti_join(
    grtsaddr_duplicates_in_stratum %>% 
      filter(
        grts_address != grts_address_drawn, 
        .by = c(grts_address_drawn, stratum)
      ),
    join_by(grts_address_drawn, grts_address, stratum)
  )
```

Verifying that all strata can be found in `n2khab_strata_expanded`.

```{r eval = mhq_data_available}
stopifnot(
  valid_extra_popunits_cell_all_n2khab %>%
  distinct(stratum) %>%
  anti_join(n2khab_strata_expanded, join_by(stratum)) %>%
  nrow() == 0
)
```

Listing the strata that are not found in `n2khab_strata` and which will be collapsed (unexpanded) in section \@ref(collapse-bsf).
Best **verify** that `collapse_strata()` indeed covers these cases.

```{r paged.print = FALSE, eval = mhq_data_available}
valid_extra_popunits_cell_all_n2khab %>%
  distinct(stratum) %>%
  anti_join(n2khab_strata, join_by(stratum))
```

We expect that all GRTS addresses, both in `grts_address_drawn` and `grts_address`, exist in the `GRTSmaster_habitats` data source.

```{r eval = mhq_data_available}
stopifnot(
  valid_extra_popunits_cell_all_n2khab %>% 
    distinct(grts_address) %>% 
    pull(grts_address) %>% 
    all(. %in% values(grts_mh))
)
```

```{r eval = mhq_data_available}
valid_extra_popunits_cell_all_n2khab %>% 
  distinct(grts_address_drawn) %>% 
  pull(grts_address_drawn) %>% 
  all(. %in% values(grts_mh))
```

Listing violations for `grts_address_drawn`:

```{r eval = mhq_data_available}
grtsaddrdrawn_invalid <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  distinct(grts_address_drawn) %>% 
  filter(!grts_address_drawn %in% values(grts_mh))
grtsaddrdrawn_invalid
```

Applying a temporary fix.

<!-- XXXXXX To be revisited! -->

```{r eval = mhq_data_available && nrow(grtsaddrdrawn_invalid) > 0}
valid_extra_popunits_cell_all_n2khab <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  mutate(
    drawn_invalid = 
      grts_address_drawn %in% grtsaddrdrawn_invalid$grts_address_drawn,
    grts_address_drawn = ifelse(
      drawn_invalid,
      NA_integer_, 
      grts_address_drawn
    )
  ) %>% 
  select(-drawn_invalid)
```

Some `grts_address_drawn` values are very far away from their replacement (section \@ref(bsf-sanity)).

Applying a temporary fix for those cases: we set these values as missing.

<!-- XXXXXX To be revisited! -->

```{r}
existing_grts_addresses_suspected <- c(1, 22, 54, 73, 82, 86, 105)
valid_extra_popunits_cell_all_n2khab <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  mutate(grts_address_drawn = ifelse(
    grts_address_drawn %in% existing_grts_addresses_suspected,
    NA_integer_,
    grts_address_drawn
  ))
```

#### Tidy object to add valid units to the base sampling frame

Drawn GRTS addresses that have been locally replaced by another address will possibly be substituted in the sampling frame, since we need this to draw samples according to the GRTS address series.
The degree to which this happens depends on the outcome of phab-correction (section \@ref(#phab-correction)), as MHQ did not apply a phab-driven bernouilli sampling when drawing samples: only locations that are retained by phab-correction (while neglecting locally replaced drawn GRTS addresses) will lead to retaining the corresponding (replaced) drawn GRTS address.
Further, these drawn GRTS addresses should not be used to direct fieldwork, since they were indeed replaced by another GRTS address in the vicinity.

Creating a tidy object with a single GRTS address column, plus indication whether this replaces another address or is being replaced.
Depending on the purpose, one will need to filter the sampling frame one way or another, using this replacement information.

```{r eval = mhq_data_available}
valid_extra_popunits_cell_all_n2khab_tidy <- 
  valid_extra_popunits_cell_all_n2khab %>% 
  mutate(
    replaced_by = 
      case_when(
        is.na(grts_address_drawn) | is.na(grts_address) ~ NA_integer_,
        grts_address_drawn != grts_address ~ grts_address,
        .default = NA_integer_
        ),
    replacement_for = 
      case_when(
        is.na(grts_address_drawn) | is.na(grts_address) ~ NA_integer_,
        grts_address_drawn != grts_address ~ grts_address_drawn,
        .default = NA_integer_
      )
  ) %>% 
  (function(df) {
    bind_rows(
      # adding the GRTS addresses that were drawn
      df %>% 
        filter(
          !is.na(grts_address_drawn),
          grts_address_drawn != grts_address
        ) %>% 
        select(-grts_address) %>% 
        rename(grts_address = grts_address_drawn) %>% 
        relocate(stratum, typelevel_certain) %>% 
        mutate(replacement_for = NA_integer_),
      df %>% 
        select(-grts_address_drawn) %>% 
        relocate(stratum, typelevel_certain) %>% 
        mutate(replaced_by = NA_integer_)
    )
  }) %>% 
  arrange(stratum, grts_address)
```

Checking that combinations of GRTS address x type are unique.

```{r eval = mhq_data_available}
stopifnot(
  valid_extra_popunits_cell_all_n2khab_tidy %>% 
    distinct(stratum, grts_address) %>% 
    nrow() == nrow(valid_extra_popunits_cell_all_n2khab_tidy)
)
```


#### Assessed type absences

`mhq_terr_assessments` also contains information about types being absent from a location.
This information can be used to drop some population units from the base sampling frame.

As a precaution, we only consider the assessments that are not older than 'current year - 5'.

```{r eval = mhq_data_available}
absent_popunits_cell_all_n2khab <- 
  mhq_terr_assessments %>% 
  # we DON'T yet filter by `!is_present` because in some cases, the point_code
  # has multiple values for the same GRTS address, so this must be filtered and
  # summarized at the GRTS address level
  filter(
    !is.na(type),
    !is.na(is_present),
    year(assessment_date) >= year(today()) - 5
  ) %>% 
  select(point_code, type, is_present, assessment_date) %>% 
  inner_join(
    read_types() %>% 
      select(type, typelevel, hydr_class),
    join_by(type),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # expansion of main types to subtypes is needed for compatibility with MNE/MHQ
  # target populations; those main types are subsequently removed. With regard
  # to absences, the concept is simple: if a main type is assessed as absent, it
  # is safe to replace this by 'all its subtypes are absent'. We don't expand
  # subtypes to main types (that is only relevant in joining with some
  # geospatial layers)
  expand_types(mark = TRUE) %>% 
  # at the time of writing, the net operation is just the removal of main types,
  # since their subtypes were already present:
  filter(
    !(expanded & typelevel == "main_type"),
    is.na(expanded_to) | expanded_to != "main_type"
  ) %>% 
  select(-typelevel, -expanded, -expanded_to) %>% 
  # add grts_address & include whether a location is assessed at the cell center
  inner_join(
    mhq_terr_refpoints,
    join_by(point_code),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # some aquatic types seemed to be included:
  filter(hydr_class != "HC3") %>% 
  select(-hydr_class) %>% 
  inner_join(
    n2khab_types_expanded_properties %>% 
      select(type, grts_join_method) %>% 
      filter(str_detect(grts_join_method, "cell")),
    join_by(type),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  select(-point_code) %>% 
  # require that cell_center-joined types have an assessment at the cell
  # center in order to accept this as an absence
  filter(
    grts_join_method == "cell_center" & is_cell_center |
      grts_join_method == "cell"
  ) %>% 
  select(-grts_join_method, -is_cell_center) %>% 
  # since the base sampling frame uses strata, not types, we must align
  rename(stratum = type) %>% 
  # we can only honour stratum absences if these apply the same (main type /
  # subtype) level as in the base sampling frame. At the time of coding, and due
  # to the expansion operation above, this caused no further change.
  semi_join(
    n2khab_strata,
    join_by(stratum)
  ) %>% 
  # since multiple point_codes can occur within one type x GRTS address, we must
  # resolve this to a single outcome
  slice_max(assessment_date, by = c(stratum, grts_address)) %>% 
  select(-assessment_date) %>% 
  # to avoid potential multiple values of is_present at the same assessment date
  summarize(
    is_present = any(is_present),
    .by = c(stratum, grts_address)
  ) %>% 
  # only keep the absences
  filter(!is_present) %>% 
  select(-is_present) %>% 
  mutate(stratum = factor(stratum, levels = levels(n2khab_strata$stratum)))
```

We don't expect absences to be present in `valid_extra_popunits_cell_all_n2khab`.
Listing violations for further enquiry.

```{r eval = mhq_data_available}
present_absences_cell_all <- 
  absent_popunits_cell_all_n2khab %>% 
  semi_join(
    valid_extra_popunits_cell_all_n2khab,
    join_by(grts_address, stratum)
  ) %>% 
  arrange(stratum, grts_address)
present_absences_cell_all
```

As a safety precaution, we don't consider these locations as true absences.

Applying a temporary fix for those cases: dropping them from `absent_popunits_cell_all_n2khab`.

<!-- XXXXXX To be revisited! -->

```{r eval = mhq_data_available}
absent_popunits_cell_all_n2khab <- 
  absent_popunits_cell_all_n2khab %>% 
  anti_join(
    present_absences_cell_all,
    join_by(stratum, grts_address)
  )
```



### Substituting assessed population units in the base sampling frame

Updating the base sampling frame `stratum_grts_n2khab` by substituting with cells from `valid_extra_popunits_cell_all_n2khab_tidy`, and dropping those that have been assessed as absences in `absent_popunits_cell_all_n2khab`.

```{r eval = mhq_data_available}
stratum_grts_n2khab <- 
  bind_rows(
    valid_extra_popunits_cell_all_n2khab_tidy,
    stratum_grts_n2khab %>% 
      anti_join(
        valid_extra_popunits_cell_all_n2khab_tidy,
        join_by(stratum, grts_address)
      )
  ) %>% 
  anti_join(absent_popunits_cell_all_n2khab, join_by(stratum, grts_address))
```

Checking that the essential columns provide unique rows.

```{r}
stopifnot(
  stratum_grts_n2khab %>% 
    distinct(stratum, grts_address) %>% 
    nrow() == nrow(stratum_grts_n2khab)
)
```



<!-- ## Writing the base sampling frame -->


<!-- ## Reading the base sampling frame -->


## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "n2khab_types_expanded_properties",
  "n2khab_strata_expanded",
  "n2khab_strata",
  "n2khab_targetpops_expanded",
  "n2khab_targetpops",
  "points_non_cell",
  "points_non_cell_n2khab",
  "stratum_units_non_cell_n2khab",
  "units_non_cell_n2khab_grts",
  "stratum_grts_non_cell_n2khab",
  "stratum_grts_cellcenter_n2khab",
  "stratum_grts_cell_n2khab",
  "stratum_grts_cell_all_n2khab",
  "valid_extra_popunits_cell_all_n2khab_tidy",
  "absent_popunits_cell_all_n2khab",
  "stratum_grts_n2khab"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```

