# Processing the base sampling frame

Using the base sampling frame, objects are generated in this chapter that pertain to all schemes in MNE & MHQ.

These objects are needed in preparing scheme-specific sampling frames or samples.

## Raster layers that represent the base sampling frame

### Subsetting `GRTSmaster_habitats` according to the base sampling frame

Create SpatRaster layers that are masked with the 3 vector layers that represent the base sampling frame, and integrate them.

In spatial processing operations, this form (with much less integer values) has shown to be a bit faster to process than `grts_mh`.

```{r eval = params$regenerate_binary_files}
grts_mh_non_cell <- mask(grts_mh, points_non_cell_n2khab)
grts_mh_cellcenter <- mask(grts_mh, hmt_pol_cellcenter, touches = FALSE)
grts_mh_cell <- mask(grts_mh, hmt_pol_cell)
grts_mh_n2khab <- max(
  grts_mh_non_cell,
  grts_mh_cellcenter,
  grts_mh_cell,
  na.rm = TRUE
) %>%
  as.int()
writeRaster(
  grts_mh_n2khab,
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif"),
  overwrite = TRUE
)
```

```{r eval=FALSE, include=FALSE}
# compare to units_non_cell_n2khab_grts %>% distinct(grts_address) %>% nrow:
global(grts_mh_non_cell, "notNA")
# compare to nrow(hmt_pol_grts_cellcenter):
global(grts_mh_cellcenter, "notNA")
# compare to hmt_pol_grts_cell %>% distinct(grts_address) %>% nrow:
global(grts_mh_cell, "notNA")
# compare to stratum_grts_n2khab %>% distinct(grts_address) %>% nrow:
global(grts_mh_n2khab, "notNA")
```

```{r}
grts_mh_n2khab <- rast(
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif")
)
```

Creating a spatial index of the GRTS values.

```{r}
grts_mh_n2khab_index <- tibble(
  id = seq_len(ncell(grts_mh_n2khab)),
  grts_address = values(grts_mh_n2khab)[, 1]
) %>%
  filter(!is.na(grts_address))
```

## Determining the GRTS addresses per domain

Determining all GRTS addresses of the base sampling frame per domain.

Since this does not distinguish between strata, stratum collapsing (see further) is irrelevant.

```{r eval = params$regenerate_binary_files}
domain_grts_n2khab <-
  extract(grts_mh_n2khab, vect(domains)) %>%
  as_tibble() %>%
  filter(!is.na(GRTSmaster_habitats)) %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(domains)),
      domain = domains$domain
    ),
    .,
    by = "ID"
  ) %>%
  select(
    -ID,
    grts_address = GRTSmaster_habitats
  )
saveRDS(
  domain_grts_n2khab,
  file.path(datapath, "binary/intermediate/domain_grts_n2khab.rds")
)
```

```{r}
domain_grts_n2khab <-
  readRDS(file.path(
    datapath,
    "binary/intermediate/domain_grts_n2khab.rds"
  ))
```

Note that `grts_mh_n2khab` has a little more non-NA cells than if it is being restricted to domain 'Flanders'.
This needs further thought.

```{r paged.print=FALSE}
global(grts_mh_n2khab, "notNA")
domain_grts_n2khab %>%
  filter(domain == "Flanders") %>%
  nrow()
```

<!-- XXXXX ofwel voor Flanders touches=TRUE gebruiken, ofwel voor alle domeinen touches=TRUE gebruiken (maar dit geeft oververtegenwoordiging door eenheden), ofwel houden as-is. -->
<!-- Onderzoek nodig met welke eenheden dit verschil heeft te maken (beken op de grens?). -->

## Processing the base sampling frame in preparation of scheme-specific sampling frames

### Collapse the base sampling frame to match the strata in `n2khab_strata`

We 'unexpand' the base sampling frame in order to match it properly with the target populations in `scheme_targetpop_sets$targetpops`.
We don't include this step _inside_ the base sampling frame so that the latter better reflects the information from the spatial data sources on which it is based.

This collapsing comprises two aspects:

- replacing subtype codes by their main type code (where the target population is defined by the latter)
- adding units, associated with a main type, to each corresponding subtype layer that triggered the expansion to the main type.

**Missing functionality**: including the second aspect in an attribute that reflects uncertainty about the stated type.

In effect, 'collapsing' means less stratum levels, but _more_ rows.

Note that collapsing only affects cell-based units.

<!-- Note that below hasn't been coded in a generic way; it is specified based upon the observed expansion. -->
<!-- However it is clear that no other expansions occur in MNE + MHQ when comparing `n2khab_types` with `n2khab_types_expanded`. -->

```{r}
stratum_grts_n2khab_collapsed <-
  stratum_grts_n2khab %>%
  collapse_strata() %>%
  # subtypes and main types that are attached to the same GRTS address, i.e.
  # were attached to the same habitatmap polygon, have induced duplicates. We
  # need to solve that:
  distinct()
```

Checking that GRTS addresses occur only once in each stratum.

```{r}
stopifnot(
  !any(duplicated(stratum_grts_n2khab_collapsed))
)
```

Checking that no empty GRTS addresses occur.

```{r}
stopifnot(
  !anyNA(stratum_grts_n2khab_collapsed$grts_address)
)
```

Checking that no strata exist that don't occur in `n2khab_strata`.

```{r}
stopifnot(
  stratum_grts_n2khab_collapsed %>%
  distinct(stratum) %>%
  anti_join(n2khab_strata, by = "stratum") %>%
  nrow() == 0
)
```


### Apply bernouilli sampling to the collapsed base sampling frame, depending on phab value

`phab` values of `cell_conditioned_on_center` types in each polygon (no collapsing).

```{r}
hmt_pol_phab_cellcenter_n2khab <-
  hmt_occ %>%
  summarize(phab = sum(phab), .by = c(polygon_id, type)) %>%
  semi_join(
    n2khab_types_expanded_properties %>%
      filter(sample_support_code == "cell_conditioned_on_center"),
    by = "type"
  )
```

**Missing functionality**:
dissolving polygons _per type_ based on a similar value:

- aimed 21 interval levels are [0, 2.5], (2.5, 7.5], (7.5, 12.5], …, (97.5, 100].
(`cut(c(0, 100), c(0, seq(2.5, 97.5, 5), 100), include.lowest = TRUE)`).
- the resulting polygons are only needed for the replacement protocol.
Filtering in order to create spatial sampling frames can just be done at the cell level.

`phab` values of `cell_conditioned_on_center` types (strata) per GRTS address, including stratum collapsing.

```{r}
stratum_grts_phab_cellcenter_n2khab_collapsed <- 
  hmt_pol_grts_cellcenter %>%
  inner_join(
    hmt_pol_phab_cellcenter_n2khab,
    by = "polygon_id",
    relationship = "many-to-many",
    unmatched = c("error", "drop")
  ) %>%
  rename(stratum = type) %>%
  collapse_strata() %>% 
  select(-polygon_id) %>% 
  relocate(stratum) %>% 
  # due to the collapsing, we need to sum cases where types were translated into
  # the same code
  summarize(
    phab = sum(phab),
    .by = c(stratum, grts_address)
  )
```

Determining for each stratum which GRTS addresses are dropped by bernouilli sampling, based on the `raster_runif` data source.
Note that this step can be turned off with the `phab_correct` parameter.

```{r eval = params$phab_correct}
stratum_grts_n2khab_collapsed_phabexcluded <- 
  tibble(
    grts_address = values(grts_mh_n2khab, na.rm = TRUE)[, 1],
    # lower limit (minimum threshold) for phab:
    phab_ll = read_raster_runif() %>% 
      crop(grts_mh_n2khab, mask = TRUE) %>% 
      values(na.rm = TRUE) %>% 
      .[, 1] * 100
  ) %>% 
  inner_join(
    stratum_grts_phab_cellcenter_n2khab_collapsed,
    join_by(grts_address),
    relationship = "one-to-many",
    unmatched = c("drop", "error")
  ) %>% 
  filter(phab < phab_ll) %>% 
  select(grts_address, stratum)
```

Constructing the phab-corrected collapsed base sampling frame.

```{r}
stratum_grts_n2khab_phabcorrected <- 
  if (params$phab_correct) {
    stratum_grts_n2khab_collapsed %>% 
      anti_join(
        stratum_grts_n2khab_collapsed_phabexcluded,
        join_by(stratum, grts_address)
      )
  } else {
    # dummy step
    stratum_grts_n2khab_collapsed
  }
```


## Calculating the spatial population size of strata in each domain

Note: this is not useful for every scheme, since some schemes need further spatial restriction of the base sampling frame (see chapter \@ref(scheme-objects)).
Hence it will need scheme-specific recalculation anyway (chapter \@ref(scheme-objects)).

Determining the population size per domain for strata with non-cell-based units.

```{r}
domain_stratum_non_cell_nunits <-
  stratum_units_non_cell_n2khab %>%
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>%
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>%
  count(domain, stratum, name = "nunits")
```

Determining the population size per domain for strata with cell-center-based units.
Regardless whether phab-correction by bernouilli sampling has been carried out, we also determine population sizes by phab-based calculation from the collapsed base sampling frame, in order to compare both approaches.

<!-- Below simplification in domain_stratum_cellcenter_nunits_calculated (representing domains and polygons + their intersection as cells and then summing phab) can be slightly improved by actually intersecting polygons by domains, then calculating phab-corrected areas (see git history for such an approach but still lacking st_intersection()). -->

```{r}
domain_stratum_cellcenter_nunits_calculated <- 
  domain_grts_n2khab %>%
  inner_join(
    stratum_grts_phab_cellcenter_n2khab_collapsed,
    join_by(grts_address),
    relationship = "many-to-many",
    unmatched = c("drop", "drop")
  ) %>%
  summarize(
    nunits = sum(phab) / 100,
    .by = c(domain, stratum)
  ) %>%
  arrange(domain, stratum)

domain_stratum_cellcenter_nunits <- 
  if (params$phab_correct) {
    domain_grts_n2khab %>%
      inner_join(
        stratum_grts_n2khab_phabcorrected,
        join_by(grts_address),
        relationship = "many-to-many",
        unmatched = c("drop", "drop")
      ) %>% 
      semi_join(
        n2khab_types_expanded_properties %>%
          filter(sample_support_code == "cell_conditioned_on_center"),
        join_by(stratum == type)
      ) %>%
      count(domain, stratum, name = "nunits") %>%
      arrange(domain, stratum)
  } else {
    domain_stratum_cellcenter_nunits_calculated
  }
```

Listing largest relative and absolute differences between `domain_stratum_cellcenter_nunits` and `domain_stratum_cellcenter_nunits_calculated`:

```{r paged.print = FALSE}
domain_stratum_cellcenter_nunits_diffs <-
  domain_stratum_cellcenter_nunits_calculated %>%
  rename(nunits_calc = nunits) %>%
  left_join(
    domain_stratum_cellcenter_nunits,
    join_by(domain, stratum),
    relationship = "one-to-one",
    unmatched = "error"
  ) %>%
  mutate(
    nunits = replace_na(nunits, 0),
    absdiff = nunits - nunits_calc,
    reldiff = (nunits - nunits_calc) / nunits_calc
  )

domain_stratum_cellcenter_nunits_diffs %>%
  filter(abs(reldiff) > 0.25) %>% 
  arrange(desc(abs(reldiff)))

domain_stratum_cellcenter_nunits_diffs %>% 
  arrange(desc(abs(absdiff)))
```

Notes:

- The more accurate and fine-grained the base sampling frame (for the `cell_conditioned_on_center` strata) is, i.e. the more cells with phab = 100 occur, the less differences will emerge between both approaches.
- Large relative differences emerge only for (very) small populations.
- The bernouilli sampling uses a raster with numbers from a random uniform distribution, to apply a bernouilli sampling with cell-specific probability `phab / 100`.
This naturally results in deviations from the expected (calculated) population size, but at the same time there is no straightforward way to adjust this while maintaining these bernouilli-based inclusion probabilities.
Furthermore, post-hoc adjustments would not hold for every possible current or future domain at the same time.
That ultimately requires a perfect base sampling frame for the `cell_conditioned_on_center` strata, having phab = 100 in all cells.
- Spatially using a simple random approach within this bernouilli sampling indeed makes most sense: there is no reason to assume that cells to be selected (and that are assumed to represent phab = 100) should be spatially well spread or systematically distributed within a phab tier of the considered stratum.
- Consequently, we accept this inaccuracy when reducing the sampling frame for `cell_conditioned_on_center` strata based on phab.
- It is not possible to both apply the phab-based reduction and still use the calculated population size for finite population correction of sample sizes, because of the mismatch between the number of population units in the obtained base sampling frame (and which are available for sampling), and this calculated population size.
  - If the latter is larger than in the corrected base sampling frame, then for smaller populations the required sample size (as obtained from the calculated population size) may or will not available in the sampling frame.
  - If the calculated population size is smaller than in the corrected base sampling frame, then we could choose to neglect this difference and use the calculated population size for finite population correction, but unwanted side-effects (of more units available than expected) may then still occur, e.g. total spatial sample sizes would still arise in module combos or split panel designs (chapter \@ref(spatial-sampling)) that are larger than the calculated population size.

Determining the population size per domain for strata with cell-based units.

```{r}
domain_stratum_cell_nunits <-
  stratum_grts_cell_n2khab %>%
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>%
  count(domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
domain_stratum_nunits <-
  bind_rows(
    domain_stratum_non_cell_nunits,
    domain_stratum_cellcenter_nunits,
    domain_stratum_cell_nunits
  ) %>%
  arrange(pick(-nunits)) %>%
  mutate(nunits = round(nunits) %>% as.integer())
```

Also deriving the population size at type level.

```{r}
domain_type_nunits <-
  domain_stratum_nunits %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  summarize(nunits = sum(nunits), .by = c(domain, type))
```

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
cat("grts_mh_n2khab\n")
grts_mh_n2khab
cat("\n")

for (name in c(
  "grts_mh_n2khab_index",
  "domain_grts_n2khab",
  "stratum_grts_n2khab_collapsed",
  "hmt_pol_phab_cellcenter_n2khab",
  "stratum_grts_phab_cellcenter_n2khab_collapsed",
  "stratum_grts_n2khab_phabcorrected",
  "domain_stratum_nunits",
  "domain_type_nunits"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
