# Creating spatial sampling frames

## Collapse the base sampling frame to match the strata in `n2khab_strata`

We 'unexpand' the base sampling frame in order to match it properly with `targetpop`.
We don't include this step _inside_ the base sampling frame so that the latter better reflects the information from the spatial data sources on which it is based.

This collapsing comprises two aspects:

- replacing subtype codes by their main type code (where the target population is defined by the latter)
- adding units, associated with a main type, to each corresponding subtype layer that triggered the expansion to the main type.

**Missing functionality**: including the second aspect in an attribute that reflects uncertainty about the stated type.

In effect, 'collapsing' means less stratum levels, but _more_ rows.

Note that collapsing only affects cell-based units.

<!-- Note that below hasn't been coded in a generic way; it is specified based upon the observed expansion. -->
<!-- However it is clear that no other expansions occur in MNE + MHQ when comparing `n2khab_types` with `n2khab_types_expanded`. -->

```{r}
stratum_grts_n2khab_collapsed <- 
  stratum_grts_n2khab %>% 
  collapse_strata() %>% 
  # subtypes and main types that are attached to the same GRTS address, i.e.
  # were attached to the same habitatmap polygon, have induced duplicates. We
  # need to solve that:
  distinct()
```

Check that GRTS addresses occur only once in each stratum:

```{r}
!any(duplicated(stratum_grts_n2khab_collapsed))
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_n2khab_collapsed$grts_address)
```

Check that no strata exist that don't occur in `n2khab_strata`:

```{r}
stratum_grts_n2khab_collapsed %>% 
  distinct(stratum) %>% 
  anti_join(n2khab_strata, by = "stratum") %>% 
  nrow == 0
```


```{r}
glimpse(stratum_grts_n2khab_collapsed)
```

## Apply bernouilli sampling to the collapsed base sampling frame, depending on phab value

```{r}
# dummy step
stratum_grts_n2khab_phabcorrected <- stratum_grts_n2khab_collapsed
```


## Define target population as strata

This is needed to extract scheme- or stratum-specific sampling frames.

```{r}
targetpop_strata <- 
  targetpop %>% 
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>% 
  select(-type)
```

## Apply scheme-specific spatial restrictions to certain types

**Missing functionality**: apply spatial restrictions depending on scheme x stratum.







