# Processing the base sampling frame

Using the base sampling frame, objects are generated in this chapter that pertain to all schemes in MNE & MHQ.

These objects are needed in preparing scheme-specific sampling frames or samples.

## Raster layers that represent the base sampling frame

### Subsetting `GRTSmaster_habitats` according to the base sampling frame

Create SpatRaster layers that are masked with the 3 vector layers that represent the base sampling frame, and integrate them.

In spatial processing operations, this form (with much less integer values) has shown to be a bit faster to process than `grts_mh`.

```{r}
grts_mh_non_cell <- mask(grts_mh, points_non_cell_n2khab)
grts_mh_cellcenter <- mask(grts_mh, hmt_pol_cellcenter, touches = FALSE)
grts_mh_cell <- mask(grts_mh, hmt_pol_cell)
grts_mh_n2khab <- max(
   grts_mh_non_cell, 
   grts_mh_cellcenter, 
   grts_mh_cell,
   na.rm = TRUE
 ) %>% 
  as.int()
```


```{r eval=FALSE, include=FALSE}
# compare to units_non_cell_n2khab_grts %>% distinct(grts_address) %>% nrow:
global(grts_mh_non_cell, "notNA")
# compare to nrow(hmt_pol_grts_cellcenter):
global(grts_mh_cellcenter, "notNA")
# compare to hmt_pol_grts_cell %>% distinct(grts_address) %>% nrow:
global(grts_mh_cell, "notNA")
# compare to stratum_grts_n2khab %>% distinct(grts_address) %>% nrow:
global(grts_mh_n2khab, "notNA")
```

## Determining the GRTS addresses per domain

Determining all GRTS addresses of the base sampling frame per domain.

Since this does not distinguish between strata, stratum collapsing is irrelevant.

```{r eval = params$regenerate_binary_files}
domain_grts_n2khab <-
  extract(grts_mh_n2khab, vect(domains)) %>%
  as_tibble() %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  inner_join(
    tibble(
      ID = seq_len(nrow(domains)),
      domain = domains$domain
    ),
    .,
    by = "ID"
  ) %>% 
  select(
    -ID,
    grts_address = GRTSmaster_habitats
  )
saveRDS(
  domain_grts_n2khab, 
  file.path(datapath, "binary/intermediate/domain_grts_n2khab.rds")
)
```

```{r}
domain_grts_n2khab <-
  readRDS(file.path(
    datapath, 
    "binary/intermediate/domain_grts_n2khab.rds"
  ))
```

Note that `grts_mh_n2khab` has a little more non-NA cells than if it is being restricted to domain 'Flanders'.
This needs further thought.

```{r}
global(grts_mh_n2khab, "notNA")
domain_grts_n2khab %>% filter(domain == "Flanders") %>% nrow()
```

<!-- XXXXX ofwel voor Flanders touches=TRUE gebruiken, ofwel voor alle domeinen touches=TRUE gebruiken (maar dit geeft oververtegenwoordiging door eenheden), ofwel houden as-is. -->
<!-- Onderzoek nodig met welke eenheden dit verschil heeft te maken (beken op de grens?). -->

## Objects in preparation of scheme-specific sampling frames

### Collapse the base sampling frame to match the strata in `n2khab_strata`

We 'unexpand' the base sampling frame in order to match it properly with `targetpops`.
We don't include this step _inside_ the base sampling frame so that the latter better reflects the information from the spatial data sources on which it is based.

This collapsing comprises two aspects:

- replacing subtype codes by their main type code (where the target population is defined by the latter)
- adding units, associated with a main type, to each corresponding subtype layer that triggered the expansion to the main type.

**Missing functionality**: including the second aspect in an attribute that reflects uncertainty about the stated type.

In effect, 'collapsing' means less stratum levels, but _more_ rows.

Note that collapsing only affects cell-based units.

<!-- Note that below hasn't been coded in a generic way; it is specified based upon the observed expansion. -->
<!-- However it is clear that no other expansions occur in MNE + MHQ when comparing `n2khab_types` with `n2khab_types_expanded`. -->

```{r}
stratum_grts_n2khab_collapsed <- 
  stratum_grts_n2khab %>% 
  collapse_strata() %>% 
  # subtypes and main types that are attached to the same GRTS address, i.e.
  # were attached to the same habitatmap polygon, have induced duplicates. We
  # need to solve that:
  distinct()
```

Check that GRTS addresses occur only once in each stratum:

```{r}
!any(duplicated(stratum_grts_n2khab_collapsed))
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_n2khab_collapsed$grts_address)
```

Check that no strata exist that don't occur in `n2khab_strata`:

```{r}
stratum_grts_n2khab_collapsed %>% 
  distinct(stratum) %>% 
  anti_join(n2khab_strata, by = "stratum") %>% 
  nrow == 0
```


```{r}
glimpse(stratum_grts_n2khab_collapsed)
```

### Apply bernouilli sampling to the collapsed base sampling frame, depending on phab value

`phab` values of `cell_conditioned_on_center` types.

<!-- XXXXXXX I think we should further process below object to a form that extends the `stratum_grts_cellcenter` object with a phab column! -->

<!-- Also, since this has to be used for the collapsed sampling frame, the -->
<!-- collapsing should happen too. -->
<!-- Also prevent code duplication where such thing is already done (further down *) to -->
<!-- calculate number of population units per domain. -->
<!-- * See object `hmt_pol_stratum_cellcenter_nunits` -->

```{r}
hmt_pol_phab_cellcenter_n2khab <- 
  hmt_occ %>% 
  summarize(phab = sum(phab), .by = c(polygon_id, type)) %>% 
  semi_join(
    n2khab_types_expanded_properties %>% 
      filter(sample_support_code == "cell_conditioned_on_center"),
    by = "type"
  )
```

**Missing functionality**:
dissolving polygons _per type_ based on a similar value:

- aimed 21 interval levels are [0, 2.5], (2.5, 7.5], (7.5, 12.5], …, (97.5, 100].
(`cut(c(0, 100), c(0, seq(2.5, 97.5, 5), 100), include.lowest = TRUE)`).
- the resulting polygons are only needed for the replacement protocol.
Filtering in order to create spatial sampling frames can just be done at the cell level.

<!-- PHAB correction still to be done! -->

```{r}
# dummy step
stratum_grts_n2khab_phabcorrected <- stratum_grts_n2khab_collapsed
```


## Calculating the spatial population size of strata in each domain

### Approximately by using the phab-corrected data frame of stratum x grts addresses in MNE + MHQ

The phab-corrected (i.e. subsampled) data frame of stratum x grts addresses is one approach to approximate the population sizes; however beside inaccuracy due to phab-based subsampling, this is also compounded by the fact that some cells include multiple (non-cell) units that belong to the same stratum.

<!-- However we did NOT YET APPLY this phab correction!!
This results in an overestimation for the cell-centered types,
but it's still a good base to DROP strata from domains XXXXXX -->

```{r}
domain_stratum_ncells <- 
  stratum_grts_n2khab_phabcorrected %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "ncells")
```

### More correctly by combining different unit-level data sources

Determining the population size per domain for strata with non-cell-based units.

```{r}
domain_stratum_non_cell_nunits <- 
  stratum_units_non_cell_n2khab %>% 
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "nunits")
```

Determining the population size per domain for strata with cell-center-based units.

<!-- XXXXXX Below simplicifaction (taking domain-overlapping polygons 
completely into account) can be improved by actually intersecting by domains. -->

```{r}
# polygon_ids that intersect each domain:
domain_hmt_pol <- 
  domains %>% 
  st_drop_geometry() %>% 
  select(domain) %>% 
  mutate(id = unclass(st_intersects(domains, hmt_pol))) %>% 
  unnest(id) %>% 
  inner_join(
    tibble(
      id = seq_len(nrow(hmt_pol)),
      polygon_id = hmt_pol$polygon_id
    ),
    by = "id"
  ) %>% 
  select(-id)
# number of units per polygon_id x stratum (taking into account surface area & phab):
# (since we start from hmt, we also need to collapse the types to the n2khab_strata!)
hmt_pol_stratum_cellcenter_nunits <- 
  hmt_pol_phab_cellcenter_n2khab %>% 
  inner_join(
    hmt_pol %>% 
      mutate(area = st_area(.) %>% drop_units()) %>% 
      st_drop_geometry(),
    by = "polygon_id"
  ) %>% 
  mutate(nunits = area * phab / 100 / (32^2), .keep = "unused") %>% 
  rename(stratum = type) %>% 
  collapse_strata() %>% 
  summarize(
    nunits = sum(nunits),
    .by = c(polygon_id, stratum)
  )
# population size per domain x stratum:
domain_stratum_cellcenter_nunits <- 
  domain_hmt_pol %>% 
  inner_join(
    hmt_pol_stratum_cellcenter_nunits, 
    by = "polygon_id",
    relationship = "many-to-many"
  ) %>% 
  summarize(nunits = sum(nunits), .by = c(domain, stratum))
```

Determining the population size per domain for strata with cell-based units.

```{r}
domain_stratum_cell_nunits <- 
  stratum_grts_cell_n2khab %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
domain_stratum_nunits <- 
  bind_rows(
    domain_stratum_non_cell_nunits,
    domain_stratum_cellcenter_nunits,
    domain_stratum_cell_nunits
  )
```

Check that this coincides with the domain x stratum combinations of `domain_stratum_ncells`.

```{r}
inner_join(
  domain_stratum_ncells, 
  domain_stratum_nunits, 
  by = c("domain", "stratum")
) %>% 
  nrow() == nrow(domain_stratum_nunits)
```

<!-- XXXXX -->
<!-- FALSE ==> st_intersection() needed in crossing `hmt_pol` with `domains`, to achieve a better result! -->
<!-- It will normally still return 'something', but since no rbbmr _raster cells_ seem to involved, it should be much smaller than '1 unit'. -->
<!-- One could then further simplify by rounding the result and dropping zeros. -->

<!-- > domain_stratum_nunits %>% anti_join(domain_stratum_ncells, by = c("domain", "stratum")) -->
<!-- # A tibble: 1 × 3 -->
<!--   domain    stratum nunits -->
<!--   <fct>     <fct>    <dbl> -->
<!-- 1 BE2200039 rbbmr     1.02 -->

<!-- > n2khab_types_expanded_properties %>% filter(type == "rbbmr") %>% select(type, sample_support_code) -->
<!-- # A tibble: 1 × 2 -->
<!--   type  sample_support_code        -->
<!--   <fct> <fct>                      -->
<!-- 1 rbbmr cell_conditioned_on_center -->




