# Processing the base sampling frame

Using the base sampling frame, objects are generated in this chapter that pertain to all schemes in MNE & MHQ.

These objects are needed in preparing scheme-specific sampling frames or samples.

## Raster layers that represent the base sampling frame

### Subsetting `GRTSmaster_habitats` according to the base sampling frame

Create SpatRaster layers that are masked with the 3 vector layers that represent the base sampling frame, and integrate them.

In spatial processing operations, this form (with much less integer values) has shown to be a bit faster to process than `grts_mh`.

```{r eval = params$regenerate_binary_files}
grts_mh_non_cell <- mask(grts_mh, points_non_cell_n2khab)
grts_mh_cellcenter <- mask(grts_mh, hmt_pol_cellcenter, touches = FALSE)
grts_mh_cell <- mask(grts_mh, hmt_pol_cell)
grts_mh_n2khab <- max(
  grts_mh_non_cell,
  grts_mh_cellcenter,
  grts_mh_cell,
  na.rm = TRUE
) %>%
  as.int()
writeRaster(
  grts_mh_n2khab,
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif"),
  overwrite = TRUE
)
```

```{r eval=FALSE, include=FALSE}
# compare to units_non_cell_n2khab_grts %>% distinct(grts_address) %>% nrow:
global(grts_mh_non_cell, "notNA")
# compare to nrow(hmt_pol_grts_cellcenter):
global(grts_mh_cellcenter, "notNA")
# compare to hmt_pol_grts_cell %>% distinct(grts_address) %>% nrow:
global(grts_mh_cell, "notNA")
# compare to stratum_grts_n2khab %>% distinct(grts_address) %>% nrow:
global(grts_mh_n2khab, "notNA")
```

```{r}
grts_mh_n2khab <- rast(
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif")
)
```

Creating a spatial index of the GRTS values.

```{r}
grts_mh_n2khab_index <- tibble(
  id = seq_len(ncell(grts_mh_n2khab)),
  grts_address = values(grts_mh_n2khab)[, 1]
) %>%
  filter(!is.na(grts_address))
```

## Determining the GRTS addresses per domain

Determining all GRTS addresses of the base sampling frame per domain.

Since this does not distinguish between strata, stratum collapsing (see further) is irrelevant.

```{r eval = params$regenerate_binary_files}
domain_grts_n2khab <-
  extract(grts_mh_n2khab, vect(domains)) %>%
  as_tibble() %>%
  filter(!is.na(GRTSmaster_habitats)) %>%
  inner_join(
    tibble(
      ID = seq_len(nrow(domains)),
      domain = domains$domain
    ),
    .,
    by = "ID"
  ) %>%
  select(
    -ID,
    grts_address = GRTSmaster_habitats
  )
saveRDS(
  domain_grts_n2khab,
  file.path(datapath, "binary/intermediate/domain_grts_n2khab.rds")
)
```

```{r}
domain_grts_n2khab <-
  readRDS(file.path(
    datapath,
    "binary/intermediate/domain_grts_n2khab.rds"
  ))
```

Note that `grts_mh_n2khab` has a little more non-NA cells than if it is being restricted to domain 'Flanders'.
This needs further thought.

```{r paged.print=FALSE}
global(grts_mh_n2khab, "notNA")
domain_grts_n2khab %>%
  filter(domain == "Flanders") %>%
  nrow()
```

<!-- XXXXX ofwel voor Flanders touches=TRUE gebruiken, ofwel voor alle domeinen touches=TRUE gebruiken (maar dit geeft oververtegenwoordiging door eenheden), ofwel houden as-is. -->
<!-- Onderzoek nodig met welke eenheden dit verschil heeft te maken (beken op de grens?). -->

## Objects in preparation of scheme-specific sampling frames

### Collapse the base sampling frame to match the strata in `n2khab_strata`

We 'unexpand' the base sampling frame in order to match it properly with the target populations in `scheme_targetpop_sets$targetpops`.
We don't include this step _inside_ the base sampling frame so that the latter better reflects the information from the spatial data sources on which it is based.

This collapsing comprises two aspects:

- replacing subtype codes by their main type code (where the target population is defined by the latter)
- adding units, associated with a main type, to each corresponding subtype layer that triggered the expansion to the main type.

**Missing functionality**: including the second aspect in an attribute that reflects uncertainty about the stated type.

In effect, 'collapsing' means less stratum levels, but _more_ rows.

Note that collapsing only affects cell-based units.

<!-- Note that below hasn't been coded in a generic way; it is specified based upon the observed expansion. -->
<!-- However it is clear that no other expansions occur in MNE + MHQ when comparing `n2khab_types` with `n2khab_types_expanded`. -->

```{r}
stratum_grts_n2khab_collapsed <-
  stratum_grts_n2khab %>%
  collapse_strata() %>%
  # subtypes and main types that are attached to the same GRTS address, i.e.
  # were attached to the same habitatmap polygon, have induced duplicates. We
  # need to solve that:
  distinct()
```

Check that GRTS addresses occur only once in each stratum:

```{r}
!any(duplicated(stratum_grts_n2khab_collapsed))
```

Check that no empty GRTS addresses occur:

```{r}
!anyNA(stratum_grts_n2khab_collapsed$grts_address)
```

Check that no strata exist that don't occur in `n2khab_strata`:

```{r}
stratum_grts_n2khab_collapsed %>%
  distinct(stratum) %>%
  anti_join(n2khab_strata, by = "stratum") %>%
  nrow() == 0
```


```{r}
glimpse(stratum_grts_n2khab_collapsed)
```

### Apply bernouilli sampling to the collapsed base sampling frame, depending on phab value

`phab` values of `cell_conditioned_on_center` types.

<!-- XXXXXXX I think we should further process below object to a form that extends the `stratum_grts_cellcenter` object with a phab column! -->

<!-- Also, since this has to be used for the collapsed sampling frame, the -->
<!-- collapsing should happen too. -->
<!-- Also prevent code duplication where such thing is already done (further down *) to -->
<!-- calculate number of population units per domain. -->
<!-- * See object `hmt_pol_stratum_cellcenter_nunits` -->

```{r}
hmt_pol_phab_cellcenter_n2khab <-
  hmt_occ %>%
  summarize(phab = sum(phab), .by = c(polygon_id, type)) %>%
  semi_join(
    n2khab_types_expanded_properties %>%
      filter(sample_support_code == "cell_conditioned_on_center"),
    by = "type"
  )
```

**Missing functionality**:
dissolving polygons _per type_ based on a similar value:

- aimed 21 interval levels are [0, 2.5], (2.5, 7.5], (7.5, 12.5], …, (97.5, 100].
(`cut(c(0, 100), c(0, seq(2.5, 97.5, 5), 100), include.lowest = TRUE)`).
- the resulting polygons are only needed for the replacement protocol.
Filtering in order to create spatial sampling frames can just be done at the cell level.

<!-- PHAB correction still to be done! -->

```{r}
# dummy step
stratum_grts_n2khab_phabcorrected <- stratum_grts_n2khab_collapsed
phabcorrection_done <- FALSE
```


## Calculating the spatial population size of strata in each domain

Note: this is not useful for every scheme, since some schemes need further spatial restriction of the base sampling frame (see chapter \@ref(scheme-objects)).
Hence it will need scheme-specific recalculation anyway (chapter \@ref(scheme-objects)).

Determining the population size per domain for strata with non-cell-based units.

```{r}
domain_stratum_non_cell_nunits <-
  stratum_units_non_cell_n2khab %>%
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>%
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>%
  count(domain, stratum, name = "nunits")
```

Determining the population size per domain for strata with cell-center-based units.

<!-- Below simplification (representing domains and polygons + their intersection as cells and then using phab) can be slightly improved by actually intersecting polygons by domains, then calculating phab-corrected areas (see git history for such an approach but still lacking st_intersection()). -->

```{r}
domain_stratum_cellcenter_nunits <-
  domain_grts_n2khab %>%
  inner_join(hmt_pol_grts_cellcenter, by = "grts_address") %>%
  inner_join(
    hmt_pol_phab_cellcenter_n2khab,
    by = "polygon_id",
    relationship = "many-to-many"
  ) %>%
  rename(stratum = type) %>%
  collapse_strata() %>%
  summarize(
    nunits = sum(phab) / 100,
    .by = c(domain, stratum)
  )
```

Determining the population size per domain for strata with cell-based units.

```{r}
domain_stratum_cell_nunits <-
  stratum_grts_cell_n2khab %>%
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>%
  count(domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
domain_stratum_nunits <-
  bind_rows(
    domain_stratum_non_cell_nunits,
    domain_stratum_cellcenter_nunits,
    domain_stratum_cell_nunits
  ) %>%
  arrange(across(-nunits)) %>%
  mutate(nunits = round(nunits) %>% as.integer())
```

Also deriving the population size at type level.

```{r}
domain_type_nunits <-
  domain_stratum_nunits %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  summarize(nunits = sum(nunits), .by = c(domain, type))
```

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
cat("grts_mh_n2khab\n")
grts_mh_n2khab
cat("\n")

for (name in c(
  "grts_mh_n2khab_index",
  "domain_grts_n2khab",
  "stratum_grts_n2khab_collapsed",
  "hmt_pol_phab_cellcenter_n2khab",
  "stratum_grts_n2khab_phabcorrected",
  "domain_stratum_nunits",
  "domain_type_nunits"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
