# Generating scheme-specific objects in preparation of generating samples {#scheme-objects}

These data objects are restricted to the schemes and types defined by the target populations in `scheme_targetpop_sets$targetpops`.

## Defining the target populations using strata

This is needed to extract scheme- or stratum-specific sampling frames.

Add target populations using strata for each set of schemes + target populations, in `scheme_targetpop_sets`.

```{r}
scheme_targetpop_sets <-
  scheme_targetpop_sets %>%
  mutate(
    targetpops_strata = map(targetpops, function(tp) {
      inner_join(tp, n2khab_strata, by = "type", relationship = "many-to-many") %>%
        select(-type)
    })
  )
```

Flattening the target populations per module, using strata.

```{r paged.print=FALSE}
module_targetpops_strata <-
  modules %>%
  select(module, stset_id) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, targetpops_strata),
    by = "stset_id"
  ) %>%
  select(-stset_id) %>%
  unnest(targetpops_strata)
```

Creating unioned set of target populations using strata.

```{r}
targetpops_strata <-
  targetpops %>%
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>%
  select(-type)
```

Beware that the target populations in `targetpops_strata` may be too wide for application to a scheme in a specific module.

## Generating spatial sampling frames per scheme

### Applying scheme-specific spatial restrictions to certain types

By default, the spatial sampling frame of a scheme is the base sampling frame, subsetted by the scheme's target population.
This is a restriction that can easily be applied on the fly while defining the sample for each scheme.

However **some scheme x stratum combinations** apply a further spatial restriction of the spatial sampling frame.

To this end, we shrink the base sampling frame in different ways, in order to replace the base sampling frames for these schemes.

```{r paged.print=FALSE}
schemes %>%
  filter(!is.na(spatial_restriction)) %>%
  select(scheme, spatial_restriction)
```

```{r}
schemes %>%
  distinct(spatial_restriction) %>%
  filter(!is.na(spatial_restriction)) %>%
  pull(spatial_restriction)
```

- **Restriction 1**: _For terrestrial types: restrict to zones with shallow groundwater (in reach of vegetation). Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Masking `grts_mh_n2khab` according to the `shallowgroundwater` layer.
To be conservative, we don't require cell centers to intersect the polygons.

```{r}
sg <-
  locate_n2khab_data() %>%
  file.path("10_raw/shallowgroundwater/shallowgroundwater.gpkg") %>%
  vect()
sg_grts_n2khab <-
  mask(grts_mh_n2khab, sg) %>%
  values(na.rm = TRUE) %>%
  as.vector()
```

Preparing restricted result `stratum_grts_n2khab_phabc_1`.

```{r}
strata_for_restriction_1 <-
  n2khab_types_expanded_properties %>%
  filter(groundw_dep != "GD2", hydr_class != "HC3") %>%
  semi_join(n2khab_strata_expanded, ., by = "type") %>%
  pull(stratum)
strata_gd2 <-
  n2khab_types_expanded_properties %>%
  filter(groundw_dep == "GD2") %>%
  semi_join(n2khab_strata_expanded, ., by = "type") %>%
  pull(stratum)
stratum_grts_n2khab_phabc_1 <-
  stratum_grts_n2khab_phabcorrected %>%
  group_by(grts_address) %>%
  mutate(
    to_restrict = stratum %in% strata_for_restriction_1,
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>%
  ungroup() %>%
  filter(
    !to_restrict |
      gd2_in_cell |
      grts_address %in% sg_grts_n2khab
  ) %>%
  select(stratum, grts_address)
```

- **Restriction 2**: _Restrict to zones with shallow groundwater (in reach of vegetation). Exclude aquatic (i.e. rivulet) forms of type 7220. Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Preparing restricted result `stratum_grts_n2khab_phabc_2`.

```{r}
grts_7220_allowed_2 <-
  extract(
    grts_mh_n2khab,
    habspring_units %>%
      filter(system_type != "rivulet") %>%
      st_coordinates()
  ) %>%
  filter(!is.na(GRTSmaster_habitats)) %>%
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_2 <-
  stratum_grts_n2khab_phabcorrected %>%
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_2) %>%
  group_by(grts_address) %>%
  mutate(
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>%
  ungroup() %>%
  filter(gd2_in_cell | grts_address %in% sg_grts_n2khab) %>%
  select(stratum, grts_address)
```

- **Restriction 3**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable. Exclude locations where the surface water is not connected with (nor affected by) the phreatic aquifer_
- **Restriction 4**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable_

We combine both restrictions since the condition relating to surface water - groundwater interaction is to be assessed in situ.

Preparing restricted result `stratum_grts_n2khab_phabc_3`.

```{r}
grts_7220_allowed_3 <-
  extract(
    grts_mh_n2khab,
    habspring_units %>%
      filter(system_type != "mire") %>%
      st_coordinates()
  ) %>%
  filter(!is.na(GRTSmaster_habitats)) %>%
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_3 <-
  stratum_grts_n2khab_phabcorrected %>%
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_3)
```

- **Restriction 5**: _Restrict to 8310 marl quarries that are directly connected with (intersected by) a groundwater aquifer_

**Missing data**: new data source needed to apply this spatial restriction.

```{r}
# dummy steps
grts_8310_allowed_5 <-
  extract(
    grts_mh_n2khab,
    st_coordinates(points_8310)
  ) %>%
  filter(!is.na(GRTSmaster_habitats)) %>%
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_5 <-
  stratum_grts_n2khab_phabcorrected %>%
  filter(stratum != "8310" | grts_address %in% grts_8310_allowed_5)
```


### Defining spatial sampling frames

```{r}
sp_samplingframes <-
  tribble(
    ~sampling_frame_id, ~sampling_frame,
    0L, stratum_grts_n2khab_phabcorrected,
    1L, stratum_grts_n2khab_phabc_1,
    2L, stratum_grts_n2khab_phabc_2,
    3L, stratum_grts_n2khab_phabc_3,
    5L, stratum_grts_n2khab_phabc_5,
  )
```

Deriving the domain-wise sampling frames.

```{r}
sp_samplingframe_domain <-
  sp_samplingframes %>%
  mutate(
    sampling_frame_domain = map(
      sampling_frame,
      ~ inner_join(
        .,
        domain_grts_n2khab,
        by = "grts_address",
        relationship = "many-to-many"
      ) %>%
        relocate(domain)
    ),
    .keep = "unused"
  )
```


### Setting spatial sampling frames per scheme

Note that these spatial sampling frames generally still contain strata that don't belong to the scheme's target population, nor belong to the types that exist in a specific domain with preset sample size.

This means we must always join with the scheme's target population (in `scheme_targetpop_sets$targetpops_strata`) _and_ with a domain-specific set of units (`domain_grts_n2khab`).
Furthermore, module-specific restrictions of the MHQ target populations may be in place, as seen in paragraph \@ref(dom-scheme-popsize).  

```{r}
scheme_sampling_frame <-
  schemes %>%
  select(scheme) %>%
  mutate(
    sampling_frame_id = case_match(
      scheme,
      c("GW_03.3", "GW_04.2", "GW_05.2", "GW_07.1") ~ 1L,
      "GW_05.1_terr" ~ 2L,
      "GW_05.1_quarries" ~ 5L,
      c("GW_05.1_aq", "SURF_03.4_lotic") ~ 3L,
      .default = 0L
    )
  )
```


## Calculating scheme-specific population sizes per domain {#dom-scheme-popsize}

### Precalculations at the sampling frame level

Determining the sampling frame specific population size per domain for strata with non-cell-based units.

```{r}
ssf_domain_stratum_non_cell_nunits <-
  stratum_units_non_cell_n2khab %>%
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>%
  inner_join(
    sp_samplingframe_domain %>% unnest(sampling_frame_domain),
    by = c("stratum", "grts_address"),
    relationship = "many-to-many"
  ) %>%
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Determining the sampling frame specific population size per domain for strata with cell-center-based units.

<!-- Below simplifications (either counting cells in phab-corrected sampling frames, or representing domains and polygons + their intersection as cells and then summing phab) can be slightly improved by actually intersecting polygons by domains, then calculating phab-corrected areas (see git history for such an approach but still lacking st_intersection()). -->

```{r}
ssf_domain_stratum_cellcenter_nunits <-
  if (params$phab_correct) {
    sp_samplingframe_domain %>%
      unnest(sampling_frame_domain) %>%
      semi_join(
        n2khab_types_expanded_properties %>%
          filter(sample_support_code == "cell_conditioned_on_center"),
        by = c(stratum = "type")
      ) %>%
      count(sampling_frame_id, domain, stratum, name = "nunits")
  } else {
    stratum_grts_phab_cellcenter_n2khab_collapsed %>%
      inner_join(
        sp_samplingframe_domain %>% unnest(sampling_frame_domain),
        by = c("stratum", "grts_address")
      ) %>%
      summarize(nunits = sum(phab) / 100, .by = c(sampling_frame_id, domain, stratum))
  } %>%
  arrange(pick(-nunits))
```

Determining the sampling frame specific population size per domain for strata with cell-based units.

```{r}
ssf_domain_stratum_cell_nunits <-
  sp_samplingframe_domain %>%
  unnest(sampling_frame_domain) %>%
  semi_join(stratum_grts_cell_n2khab, by = c("stratum", "grts_address")) %>%
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
ssf_domain_stratum_nunits <-
  bind_rows(
    ssf_domain_stratum_non_cell_nunits,
    ssf_domain_stratum_cellcenter_nunits,
    ssf_domain_stratum_cell_nunits
  ) %>%
  arrange(pick(-nunits)) %>%
  mutate(nunits = round(nunits) %>% as.integer())
```

Some cases exist with less than 0.5 units, which comes down to 0 units:

```{r paged.print = FALSE}
ssf_domain_stratum_nunits %>% 
  filter(nunits == 0) %>% 
  arrange(stratum, domain)
```

We still keep them for reference, since future states of these strata may become larger.

### Applying result at the scheme level

```{r}
scheme_ssf_domain_stratum_nunits <-
  scheme_sampling_frame %>%
  inner_join(
    ssf_domain_stratum_nunits,
    by = "sampling_frame_id",
    relationship = "many-to-many"
  ) %>%
  semi_join(targetpops_strata, by = c("scheme", "stratum"))
```


### Filtering result per module

Each module has a restricted set of schemes, associated target populations, and domains.

```{r}
mod_dom_scheme_ssf_stratum_nunits <-
  scheme_ssf_domain_stratum_nunits %>%
  inner_join(
    module_domains %>% select(module, domain),
    .,
    by = "domain",
    relationship = "many-to-many"
  ) %>%
  semi_join(module_targetpops_strata, by = c("module", "scheme", "stratum")) %>%
  # adding the is_core_scheme variable, which is module x scheme specific
  inner_join(
    modules %>%
      select(module, stset_id) %>%
      inner_join(
        scheme_targetpop_sets %>%
          select(stset_id, scheme_set) %>%
          unnest(scheme_set) %>%
          select(stset_id, scheme, is_core_scheme),
        by = "stset_id",
        relationship = "many-to-many"
      ) %>%
      select(-stset_id),
    by = c("module", "scheme")
  ) %>%
  relocate(is_core_scheme, .after = scheme)
```

Note that we will update `mod_dom_scheme_ssf_stratum_nunits` once again after the restriction of MHQ schemes below.

### Restriction of MHQ schemes and associated types per domain based on required surface area

Module-specific rules can apply to exclude some types or schemes from the MHQ programme in a domain, based on the surface area in such a domain.

In fact, those excluded types (see objects below) are measured in other ways, but we don't consider them in the sampling and revisit design.
This choice is also made because for several of such types this rule had already been applied at the Flemish level in MHQ which makes them not even appear in `read_scheme_types()`, hence they were already absent from objects like `targetpops` and `scheme_targetpop_sets`.

Creating some helper objects that add types and categories to `mod_dom_scheme_ssf_stratum_nunits` partim MHQ.

```{r}
mhq_mod_dom_scheme_ssf_stratum_nunits_type <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  semi_join(
    module_domains %>% filter(sample_size_predetermined),
    join_by(module, domain)
  ) %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  filter(str_detect(scheme, "^HQ"))

mhq_mod_dom_scheme_ssf_stratum_nunits_type_category <-
  mhq_mod_dom_scheme_ssf_stratum_nunits_type %>%
  inner_join(
    read_types() %>%
      mutate(category = case_when(
        type %in% c("3260", "7220") ~ "lotic",
        str_detect(hydr_class, "3$") ~ "lentic",
        .default = "terrestrial"
      )) %>%
      select(type, category),
    by = "type"
  )

mhq_scheme_category <-
  mhq_mod_dom_scheme_ssf_stratum_nunits_type_category %>%
  distinct(scheme, category) %>%
  arrange(scheme, category)
```

The `mhq_sampled_types_per_domain` object below reflects which types are being sampled in the considered MHQ schemes according to a minimum surface area condition at the domain level.
Hence this is a subset of the target population of these schemes at the Flemish level.
The complement of this (non-sampled types in MHQ) is created in chunks that follow.
For these non-sampled types in MHQ (except RIBs) the whole target population is to be measured (a census).

```{r}
mhq_sampled_types_per_domain <-
  mhq_mod_dom_scheme_ssf_stratum_nunits_type_category %>%
  summarize(
    nunits = sum(nunits),
    .by = c(domain, scheme, type, category)
  ) %>%
  # don't sample terrestrial units when area is smaller than 10 ha
  filter(nunits >= 10e4 / 32^2 | category != "terrestrial") %>%
  select(-nunits)
```

Determining the types, resp strata, for each module x domain x **scheme**, that are _not_ measured with a sample in the considered MHQ schemes (i.e. in the modules that do have MHQ schemes).

```{r}
mhq_mod_dom_scheme_type_no_sample <-
  mhq_mod_dom_scheme_ssf_stratum_nunits_type %>%
  summarize(
    nunits = sum(nunits),
    .by = c(module, domain, scheme, type)
  ) %>%
  anti_join(mhq_sampled_types_per_domain, by = c("domain", "scheme", "type"))
mhq_mod_dom_scheme_stratum_no_sample <-
  mhq_mod_dom_scheme_ssf_stratum_nunits_type %>%
  anti_join(mhq_sampled_types_per_domain, by = c("domain", "scheme", "type")) %>%
  select(module, domain, scheme, stratum, nunits)
```

Both above objects are equal except for the column name and associated factor levels, since these types are not further divided as strata:

```{r}
all.equal(
  mhq_mod_dom_scheme_type_no_sample,
  mhq_mod_dom_scheme_stratum_no_sample %>%
    rename(type = stratum),
  check.attributes = FALSE
)
```

Determining the types, _for each module x domain_, that are _not_ measured with a sample in the considered MHQ schemes.
This also includes the types that are not defined as part of any MHQ scheme.
Note that, for clarity, we don't exclude RIBs in the below object.

```{r}
mhq_mod_dom_type_no_sample <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  filter(str_detect(scheme, "^HQ")) %>%
  distinct(module, domain) %>%
  inner_join(
    domain_type_nunits,
    by = "domain",
    relationship = "many-to-many"
  ) %>%
  anti_join(
    mhq_sampled_types_per_domain %>%
      distinct(domain, type),
    by = c("domain", "type")
  ) %>%
  left_join(
    targetpops %>%
      filter(str_detect(scheme, "^HQ")),
    by = "type"
  )
```

### Applying the MHQ restriction for module x domain x scheme x stratum to the complete set

Updating `mod_dom_scheme_ssf_stratum_nunits`.

```{r}
mod_dom_scheme_ssf_stratum_nunits <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  anti_join(
    mhq_mod_dom_scheme_stratum_no_sample,
    by = c("module", "domain", "scheme", "stratum")
  )
```

## Determining schemes and population sizes per domain for each module

We restrict all schemes based on the presence of strata in domains.
Some of the domains also play a role in scheme implementation (see \@ref(domains-per-module)).

Determining the schemes that apply to each domain in each module, according to the _current_ (!) presence of strata in domains.
This already includes the MHQ restriction of schemes and _sampled_ types that has been obtained above.
Hence from here on, we only consider sampling approaches.

```{r}
module_domain_schemes <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  distinct(module, domain, scheme, is_core_scheme)
```

Using this result to shrink earlier module x scheme objects.

```{r}
for (objectname in c(
  "module_scheme_cycle_duration",
  "module_schemes_mne",
  "mod_scheme_datacoll_nofield",
  "mod_scheme_variableset_mainfieldmethod",
  "mod_scheme_mainfieldmethod",
  "mod_scheme_mainfieldmethod_subsets",
  "mod_scheme_mainfieldmethod_actseq",
  "mod_scheme_field_activity",
  "mod_scheme_actseq_fag",
  "mod_scheme_fag",
  "mod_scheme_fag_paneldesign"
)) {
  result <- eval(as.name(objectname)) %>% 
    semi_join(module_domain_schemes, by = c("module", "scheme"))
  assign(objectname, result, envir = .GlobalEnv)
}
```

Calculating statistics of the population size per **stratum**, per module x domain x scheme.

```{r}
module_domain_scheme_stratumstats <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  summarize(
    stratum_count = n(),
    stratum_nunits_min = min(nunits),
    stratum_nunits_q1 = quantile(nunits, 0.25),
    stratum_nunits_q2 = quantile(nunits, 0.5),
    stratum_nunits_q3 = quantile(nunits, 0.75),
    stratum_nunits_max = max(nunits),
    .by = c(module, domain, scheme)
  ) %>%
  arrange(module, domain, scheme)
```

Calculating statistics of the population size per **type**, per module x domain x scheme.

```{r}
module_domain_scheme_typestats <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  summarize(nunits = sum(nunits), .by = c(module, domain, scheme, type)) %>%
  summarize(
    type_count = n(),
    type_nunits_min = min(nunits),
    type_nunits_q1 = quantile(nunits, 0.25),
    type_nunits_q2 = quantile(nunits, 0.5),
    type_nunits_q3 = quantile(nunits, 0.75),
    type_nunits_max = max(nunits),
    .by = c(module, domain, scheme)
  ) %>%
  arrange(module, domain, scheme)
```

<!-- XXXXX Note that quite some polygons have been dropped in domain_hmt_pol_grts after the join with hmt_pol_grts_cellcenter was added, leading to a clearly lower number of units; this may be because of the clipping effect of grts_mh (which is too small) and also because of dropping small polygons (small=FALSE), i.e. that didn't intersect any cell center. -->
<!-- This effect should be scrutinized further, but this effect may be want we want for now. -->
<!-- XXXXX Above result is to be confronted to the ncells result after PHAB correction has been applied -->


## Check the type overlap between core and non-core MNE schemes, by environmental compartment

Determining which types in a module x compartment (GW, SURF, SOIL) are covered by a non-core scheme and _not_ by a core scheme.

```{r}
non_core_types_per_module_and_compartment <-
  modules %>%
  select(
    -year_firstdefinition,
    -contains("name_"),
    -code_short,
    -contains("postpone_")
  ) %>% 
  semi_join(
    module_domains %>% filter(domain == "Flanders"),
    by = "module"
  ) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, scheme_set, targetpops_strata) %>%
      mutate(
        scheme_targetpop = map2(
          scheme_set, targetpops_strata, function(ss, tp) {
            inner_join(ss, tp, by = "scheme")
          }
        ),
        .keep = "unused"
      ) %>%
      unnest(cols = scheme_targetpop),
    by = "stset_id"
  ) %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  filter(programme == "MNE") %>%
  select(
    module,
    scheme,
    is_core_scheme,
    compartment = attribute_1,
    type
  ) %>%
  arrange(module, compartment, type) %>%
  mutate(
    compartment_has_core_scheme = any(is_core_scheme),
    .by = c(module, compartment, type)
  ) %>%
  mutate(
    all_compartments_have_core_scheme = all(compartment_has_core_scheme),
    .by = c(module, type)
  ) %>%
  filter(!all_compartments_have_core_scheme) %>%
  select(module, compartment, scheme, is_core_scheme, type)
non_core_types_per_module_and_compartment
```

The occurrence of types with non-predetermined sample size is very limited.

## Set and collect the module x domain specific design attributes of schemes x strata (x variable-sets)

The objects that will be created in this paragraph represent the design attributes of schemes x strata (and if needed variable sets) for each module x domain combination.
Together with `scheme_sampling_frame` and `sp_samplingframe_domain`, we will have the necessary and sufficient information to draw spatial samples and allocate locations to panels and date-time intervals (chapters \@ref(spatial-sampling) and \@ref(revisit)).

### Helper objects in distributing sample sizes

Preparing objects that will distribute sample sizes that have been set at a level higher than module x domain x scheme:

- distribution in module `pan_effectmon_sac5_custommeas` is over domains within schemes (MNE), or over domains x schemes within type categories terrestrial, lentic and lotic (MHQ)
- a specific distribution is done in module `pan_transmon_be2100024` over aquatic schemes (MHQ).

```{r}
mhq_proportion_aquatic_per_module <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  semi_join(
    module_domains %>% filter(sample_size_predetermined) %>% select(module, domain),
    join_by(module, domain)
  ) %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  semi_join(read_types() %>% filter(str_detect(hydr_class, "3$")), by = "type") %>%
  filter(str_detect(scheme, "^HQ")) %>%
  distinct(module, scheme, type) %>%
  mutate(
    prop_aq_in_module = 1 / n(),
    .by = module
  ) %>%
  summarize(
    prop_aq_in_module = sum(prop_aq_in_module),
    .by = c(module, scheme)
  ) %>%
  arrange(module, scheme)

mhq_proportions_sac_subset5 <-
  mhq_sampled_types_per_domain %>%
  semi_join(
    module_domains %>%
      filter(module == "pan_effectmon_sac5_custommeas", sample_size_predetermined),
    by = "domain"
  ) %>%
  # count the number of types that is sampled in each scheme
  mutate(
    type_count_mhq_sampling = n(),
    .by = c(domain, scheme)
  ) %>%
  mutate(
    prop_domaintype_per_cat = 1 / n(),
    .by = category
  ) %>%
  summarize(
    prop_mhq_domainscheme_per_cat = sum(prop_domaintype_per_cat),
    .by = c(domain, scheme, category, type_count_mhq_sampling)
  ) %>%
  arrange(domain, scheme)

mne_core_proportions_sac_subset5 <-
  mod_dom_scheme_ssf_stratum_nunits %>%
  semi_join(
    module_domain_schemes %>%
      filter(module == "pan_effectmon_sac5_custommeas", is_core_scheme),
    by = c("module", "domain", "scheme")
  ) %>%
  filter(!str_detect(scheme, "^HQ")) %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  distinct(domain, scheme, type) %>%
  mutate(
    prop_domaintype_per_mne_scheme = 1 / n(),
    .by = scheme
  ) %>%
  summarize(
    prop_domain_per_mne_scheme = sum(prop_domaintype_per_mne_scheme),
    .by = c(domain, scheme)
  ) %>%
  arrange(domain, scheme)
```

### Collecting and setting design attributes at the module x domain x scheme level

Joining module and scheme properties to the count statistics by module, domain and scheme, needing those attributes to generate further attributes, and repeating them here for now.
Also distributing the spatial sample sizes over schemes by module x domain.

Note that from here on, we only consider the domains with predetermined sample size, i.e. those that have their own design.

**Missing functionality**: divide sample sizes over panel splits.
Sampling for splits will be done for non-consecutive (in terms of reverse GRTS addresses) sets of sampling frame units, so the sample size must be determined beforehand.

```{r}
module_domain_scheme_designattr <-
  modules %>%
  select(-year_firstdefinition, -contains("name_"), -code_short) %>% 
  inner_join(module_domains, by = "module") %>%
  filter(sample_size_predetermined) %>%
  select(-sample_size_predetermined) %>%
  inner_join(
    module_domain_schemes,
    by = c("module", "domain"),
    relationship = "many-to-many"
  ) %>%
  # note that, for MHQ schemes, mod_dom_scheme_ssf_stratum_nunits was already
  # restricted to module x domain x type combinations that are approached
  # with a sample. See the 'mhq_mod_dom_type_no_sample' object to get the
  # remainder.
  # Below _typestats/_stratumstats objects are based on
  # mod_dom_scheme_ssf_stratum_nunits:
  inner_join(
    module_domain_scheme_typestats %>%
      select(module, domain, scheme, type_count),
    by = c("module", "domain", "scheme")
  ) %>%
  inner_join(
    module_domain_scheme_stratumstats %>%
      select(module, domain, scheme, stratum_count),
    by = c("module", "domain", "scheme"),
    relationship = "many-to-many"
  ) %>%
  inner_join(schemes %>% select(scheme, attribute_1, tag_1), by = "scheme") %>%
  inner_join(scheme_sampling_frame, by = "scheme") %>%
  inner_join(module_scheme_cycle_duration, by = c("module", "scheme")) %>%
  mutate(
    year_first = case_match(
      attribute_1,
      "GW" ~ year_first_default + yr_postpone_gw,
      "SURF" ~ year_first_default + yr_postpone_surf,
      "SOIL" ~ year_first_default + yr_postpone_soil,
      .default = year_first_default
    )
  ) %>%
  left_join(mhq_proportion_aquatic_per_module, by = c("module", "scheme")) %>%
  left_join(mhq_proportions_sac_subset5, by = c("domain", "scheme")) %>%
  left_join(mne_core_proportions_sac_subset5, by = c("domain", "scheme")) %>%
  mutate(
    sp_sample_size_all_panels = cycle_duration_y * case_when(
      # NA in MNE means adopting the locations of the core scheme in the same compartment:
      !is_core_scheme & !str_detect(scheme, "^HQ") ~ NA_real_,
      module == "pan_effectmon_flanders" & scheme == "GW_03.3" ~ 167,
      module == "pan_effectmon_flanders" & scheme == "SOIL_03.2" ~ 400,
      module == "pan_effectmon_flanders" & scheme == "SURF_03.4_lentic" ~ 64,
      module == "pan_effectmon_flanders" & scheme == "SURF_03.4_lotic" ~ 16,
      # pan_effectmon_sac5_custommeas partim MNE is taken as 'one sixth' of above
      module == "pan_effectmon_sac5_custommeas" & scheme == "GW_03.3" ~ 28 * prop_domain_per_mne_scheme,
      module == "pan_effectmon_sac5_custommeas" & scheme == "SOIL_03.2" ~ 67 * prop_domain_per_mne_scheme,
      module == "pan_effectmon_sac5_custommeas" & scheme == "SURF_03.4_lentic" ~ 11 * prop_domain_per_mne_scheme,
      module == "pan_effectmon_sac5_custommeas" & scheme == "SURF_03.4_lotic" ~ 3 * prop_domain_per_mne_scheme,
      # pan_effectmon_sac5_custommeas partim MHQ is set at category level
      module == "pan_effectmon_sac5_custommeas" & str_detect(scheme, "^HQ") & category == "terrestrial" ~ 338 * prop_mhq_domainscheme_per_cat,
      module == "pan_effectmon_sac5_custommeas" & str_detect(scheme, "^HQ") & category == "lentic" ~ 42 * prop_mhq_domainscheme_per_cat,
      module == "pan_effectmon_sac5_custommeas" & str_detect(scheme, "^HQ") & category == "lotic" ~ 17 * prop_mhq_domainscheme_per_cat,
      # pan_transmon_be2100024 is set at scheme level
      module == "pan_transmon_be2100024" & scheme == "GW_03.3" ~ 70,
      module == "pan_transmon_be2100024" & scheme == "SOIL_03.2" ~ 133,
      module == "pan_transmon_be2100024" & scheme == "SURF_03.4_lentic" ~ 14,
      module == "pan_transmon_be2100024" & scheme == "SURF_03.4_lotic" ~ 4,
      module == "pan_transmon_be2100024" & str_detect(scheme, "^HQ3") ~ 18 * prop_aq_in_module,
      module == "pan_transmon_be2100024" & str_detect(scheme, "^HQ(23|40|91(20|90))") ~ 28,
      module == "pan_transmon_be2100024" & str_detect(scheme, "^HQ6230") ~ 84,
      module == "pan_transmon_be2100024" & str_detect(scheme, "^HQ(7140|91E0)") ~ 56,
      # mbaa_mne_phase_1 is set at scheme level
      module == "mbaa_mne_phase_1" & scheme == "GW_03.3" ~ 1200 / 6,
      module == "mbaa_mne_phase_1" & scheme == "GW_05.1_terr" ~ 1200 / 6,
      module == "mbaa_mne_phase_1" & scheme == "GW_05.1_aq" ~ 400 / 6,
      module == "mbaa_mne_phase_1" & scheme == "GW_05.1_quarries" ~ 25 / 6,
      module == "mbaa_mne_phase_1" & attribute_1 == "SOIL" ~ 1400 / 6,
      module == "mbaa_mne_phase_1" & scheme == "SURF_03.4_lentic" ~ 106,
      module == "mbaa_mne_phase_1" & scheme == "SURF_03.4_lotic" ~ 27,
      # NA in MHQ means a census instead of a sample (but it's not present!):
      .default = NA_real_
    )
  ) %>%
  select(
    -attribute_1,
    -tag_1,
    -category,
    -starts_with("prop"),
    -type_count_mhq_sampling,
    -contains("postpone_")
  ) %>%
  mutate(
    sp_sample_size_all_panels_type = ifelse(
      is.na(sp_sample_size_all_panels),
      NA_real_,
      sp_sample_size_all_panels / type_count
    )
  ) %>%
  relocate(
    module,
    stset_id,
    domain,
    scheme,
    is_core_scheme,
    sampling_frame_id,
    type_count,
    stratum_count,
    cycle_duration_y,
    year_first,
    year_last
  )
```

Recalculating yearly spatial sample sizes at an aggregated level as checks.

```{r}
sample_size_check1 <-
  module_domain_scheme_designattr %>%
  mutate(yearly_sample_size = as.integer(round(sp_sample_size_all_panels / cycle_duration_y))) %>%
  summarize(
    yearly_sample_size = sum(yearly_sample_size, na.rm = TRUE),
    .by = c(module, scheme)
  ) %>%
  filter(!is.na(yearly_sample_size), yearly_sample_size > 0) %>%
  arrange(module, scheme)
sample_size_check1
```

```{r paged.print = FALSE}
aggregate_sample_size(
  module_domain_scheme_designattr,
  "sp_sample_size_all_panels",
  mhq_scheme_category
) %>%
  print(n = Inf)
```

### Generate non-corrected target sample sizes for strata in module x domain x scheme combinations

Expanding `module_domain_scheme_designattr` with strata and distributing spatial sample sizes over strata.
The 4th root of the number of population units per stratum (in module x domain x scheme x type) is taken to distribute the target sample size at type level.
This was found to be a good compromise for estimator precision at type level, between the more extreme scenarios of population size versus an area weighed distribution.

- Non-core MNE schemes do not adopt sample sizes from the core MNE schemes: only the locations that are common between the respective sampling frames will be adopted in the non-core schemes, hence this can be a subset and it happens at the spatial sampling stage.
However we do attribute extra sample sizes in non-core MNE schemes, for those types not covered by core schemes -- see object `non_core_types_per_module_and_compartment`.

```{r}
module_domain_scheme_stratum_target_sample_size <-
  module_domain_scheme_designattr %>%
  inner_join(
    # note that, for MHQ schemes, mod_dom_scheme_ssf_stratum_nunits was already
    # restricted to module x domain x type combinations that are approached
    # with a sample. See the 'mhq_mod_dom_type_no_sample' object to get the
    # remainder.
    mod_dom_scheme_ssf_stratum_nunits %>%
      select(-is_core_scheme, -sampling_frame_id),
    by = c("module", "domain", "scheme")
  ) %>%
  inner_join(n2khab_strata, by = "stratum") %>%
  left_join(
    schemes %>%
      filter(programme == "MNE") %>%
      select(
        scheme,
        compartment = attribute_1
      ),
    by = "scheme"
  ) %>%
  left_join(
    non_core_types_per_module_and_compartment %>%
      select(module, compartment, scheme, type) %>%
      mutate(add_to_sample = TRUE),
    by = c("module", "compartment", "scheme", "type")
  ) %>%
  mutate(
    sp_sample_size_all_panels_type = case_when(
      is.na(sp_sample_size_all_panels_type) & add_to_sample ~ 
        max(sp_sample_size_all_panels_type, na.rm = TRUE),
      nunits == 0 & !is.na(sp_sample_size_all_panels_type) ~ 0,
      .default = sp_sample_size_all_panels_type
    ),
    .by = c(module, compartment)
  ) %>%
  mutate(
    sp_sample_size_all_panels_stratum = ifelse(
      nunits == 0 & !is.na(sp_sample_size_all_panels_type),
      0,
      sp_sample_size_all_panels_type * nunits^0.25 / sum(nunits^0.25)
    ),
    .by = c(module, domain, scheme, type)
  ) %>%
  select(-compartment, -add_to_sample) %>%
  relocate(type, stratum, nunits, .before = sp_sample_size_all_panels_type)
```

Note that the above values for `sp_sample_size_all_panels_stratum` are pro-forma target sample sizes: they would apply if the populations were infinite.
Since they are finite _and_ since unequal population sizes apply, only the target sample sizes at an aggregated level matter, in verifying the finite-population sample sizes that we calculate below.

Rechecking the aggregated sample size.
Compared to the previous result, some non-core-scheme types that are not covered by core schemes cause a limited addition of sampling in non-core schemes.

```{r paged.print = FALSE}
aggregate_sample_size(
  module_domain_scheme_stratum_target_sample_size,
  "sp_sample_size_all_panels_stratum",
  mhq_scheme_category
) %>%
  print(n = Inf)
```

### Redistributing stratum-level target sample sizes based on finite-population correction

Overdimensioning spatial sample sizes to represent the infinite case and subsequently applying finite population correction in order to achieve sample sizes at a pooled level, aimed at the same size as the preset ones.

The infinite sample size is to be obtained by applying a fixed coefficient per module x scheme in MNE, or per module x category in MHQ -- we call such level a 'pool' -- to the target sample sizes per stratum (in module x domain x scheme).
This is needed to make sure the allocation for the _infinite_ population remains evenly spread over types (within a pool) as reflected in the above pro-forma stratum-level sample sizes, which we (for simplicity) assume to correspond to a similar estimator precision at the type level.

The reason to not set a separate coefficient for each domain (with 'predetermined sample size') of the same module x scheme-or-category (pool), hence include domain in the definition of a pool, is to enforce equal estimator precision across domains within such modules x scheme-or-category (pools): all type-level sample sizes for the infinite population will remain equal across domains in a pool.
A consequence of the finite population correction may be that the budgeted sample size at the domain level is somewhat shifted among domains, but the fact that this avoids differences in degree of overdimensioning across domains can be regarded as an advantage.
Hence the 'predetermined sample size' for domains in the same module refers specifically to:

- the fact that the sample size for _infinite_ populations is preset per domain x scheme in MNE (per domain x category in MHQ), in order to serve the domains proportionally (equal precision at domain x scheme-or-category x type level);
- the resulting FPC-corrected sample sizes per domain x scheme, i.e. the fact that a sample size is applied ('determined') at the domain level.

But as one can see, it is essentially the 'module x scheme' level (or 'module x category' level) sample size that is literally 'predetermined'.

First determining the infinite sample sizes and some intermediate results, by redistributing budgeted (target) sample sizes within pools.

```{r}
upper_optim_bound <- 100
module_domain_scheme_stratum_sample_size_infinite <-
  module_domain_scheme_stratum_target_sample_size %>%
  left_join(mhq_scheme_category, join_by(scheme)) %>%
  arrange(module, scheme) %>%
  mutate(
    pool_with_equal_type_precision = ifelse(
      str_detect(scheme, "^HQ"),
      str_c(module, "MHQ", category, sep = "_"),
      str_c(module, scheme, sep = "_")
    )
  ) %>%
  mutate(
    coef = optimize(
      get_fpc_sample_size_difference,
      c(1, upper_optim_bound),
      sp_sample_size_all_panels_stratum,
      nunits
    )$minimum,
    coef = ifelse(is.na(sp_sample_size_all_panels_stratum), NA_real_, coef),
    sp_sample_size_all_panels_stratum_inf =
      sp_sample_size_all_panels_stratum * coef,
    .by = pool_with_equal_type_precision
  )
```

Cases may exist for which the pooled budgeted sample size (target sample size of a pool) is larger than the corresponding population size.
These will need adjustment of target sample sizes for respective strata.
Check if such cases exist:

<!-- XXX In this case, type 3260 has not yet got a sufficiently reliable sampling frame, so this must be revisited. -->

```{r}
module_domain_scheme_stratum_sample_size_infinite %>%
  filter(coef > upper_optim_bound - 1) %>%
  summarize(
    pooled_pop_size = sum(nunits),
    budgeted_sp_sample_size_all_panels = sum(sp_sample_size_all_panels_stratum),
    .by = pool_with_equal_type_precision
  )
```

Overview of pools, pooled budgeted (target) sample sizes, and coefficients to set the infinite population sizes of the strata contained within each pool.

```{r}
module_domain_scheme_stratum_sample_size_infinite %>%
  filter(!is.na(coef)) %>%
  summarize(
    budgeted_sp_sample_size_all_panels = sum(sp_sample_size_all_panels_stratum),
    coef = first(coef),
    .by = pool_with_equal_type_precision
  )
```

Summary of coefficient values, excluding potential extremes (which relate to insufficient availability of population units).

```{r}
module_domain_scheme_stratum_sample_size_infinite %>%
  distinct(pool_with_equal_type_precision, coef) %>%
  filter(coef < upper_optim_bound - 1) %>%
  pull(coef) %>%
  summary()
```

Applying the finite population correction.

```{r}
module_domain_scheme_stratum_sample_size <-
  module_domain_scheme_stratum_sample_size_infinite %>%
  mutate(
    sp_sample_size_all_panels_stratum = apply_fpc(
      sp_sample_size_all_panels_stratum_inf,
      nunits
    ),
    sp_sample_size_all_panels_type = sum(sp_sample_size_all_panels_stratum),
    .by = c(module, domain, scheme, type)
  ) %>%
  mutate(
    sp_sample_size_all_panels = sum(sp_sample_size_all_panels_stratum, na.rm = TRUE),
    .by = c(module, domain, scheme)
  ) %>%
  select(-c(
    category,
    pool_with_equal_type_precision,
    coef,
    sp_sample_size_all_panels_stratum_inf
  )) %>%
  mutate(across(starts_with("sp_sample_size"), ~ as.integer(round(.)))) %>%
  arrange(module, domain, scheme, type, stratum)
```

Rechecking the aggregated sample size, i.e. the calculations at stratum and at scheme level.
Small differences with previous result arise because of rounding to integer values.

```{r paged.print = FALSE}
aggregate_sample_size(
  module_domain_scheme_stratum_sample_size,
  "sp_sample_size_all_panels_stratum",
  mhq_scheme_category
) %>%
  print(n = Inf)
```

```{r paged.print = FALSE}
module_domain_scheme_stratum_sample_size %>%
  distinct(
    module,
    domain,
    scheme,
    cycle_duration_y,
    type_count,
    sp_sample_size_all_panels
  ) %>%
  aggregate_sample_size(
    "sp_sample_size_all_panels",
    mhq_scheme_category
  ) %>%
  print(n = Inf)
```

### Adjustments depending on variable sets

Expanding with variables and field methods and adjusting sample sizes where needed.

<!-- XXXX Set sample size adjustment in 010_settings -->

```{r include=FALSE}
module_domain_scheme_designattr %>%
  inner_join(
    mod_scheme_variableset_mainfieldmethod,
    join_by(module, scheme),
    relationship = "many-to-many",
    unmatched = "error"
  )
```

## Overview of most relevant objects in this chapter

Objects subsetted with `[[1]]` mean that they (may) contain more, similarly structured elements.

Note that a whole series of `mod_scheme_xxxxx` and some other objects from chapter \@ref(settings) have been updated in this chapter; see the list in section \@ref(overview-settings).
Below, only `module_scheme_cycle_duration` and `scheme_targetpop_sets` are being shown again.

```{r echo = FALSE}
for (name in c(
  "scheme_targetpop_sets",
  "scheme_targetpop_sets$scheme_set[[1]]",
  "scheme_targetpop_sets$targetpops[[1]]",
  "scheme_targetpop_sets$targetpops_strata[[1]]",
  "module_targetpops_strata",
  "targetpops_strata",
  "sp_samplingframes",
  "sp_samplingframes$sampling_frame[[1]]",
  "sp_samplingframe_domain",
  "sp_samplingframe_domain$sampling_frame_domain[[1]]",
  "scheme_sampling_frame",
  "mhq_sampled_types_per_domain",
  "mhq_mod_dom_scheme_type_no_sample",
  "mhq_mod_dom_type_no_sample",
  "mod_dom_scheme_ssf_stratum_nunits",
  "module_domain_schemes",
  "module_scheme_cycle_duration",
  "module_domain_scheme_typestats",
  "module_domain_scheme_stratumstats",
  "module_domain_scheme_designattr",
  "module_domain_scheme_stratum_target_sample_size",
  "module_domain_scheme_stratum_sample_size"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```



