# Generating scheme-specific objects in preparation of generating samples {#scheme-objects}

These data objects are restricted to the schemes and types defined by the target populations in `scheme_targetpop_sets$targetpops`.

## Defining the target populations using strata

This is needed to extract scheme- or stratum-specific sampling frames.

Add target populations using strata for each set of schemes + target populations, in `scheme_targetpop_sets`.

```{r}
scheme_targetpop_sets <- 
  scheme_targetpop_sets %>% 
  mutate(
    targetpops_strata = map(targetpops, function(tp) {
      inner_join(tp, n2khab_strata, by = "type", relationship = "many-to-many") %>% 
        select(-type)
    })
  )
```

Flattening the target populations per module, using strata.

```{r paged.print=FALSE}
module_targetpops_strata <- 
  modules %>%
  select(module, stset_id) %>%
  inner_join(
    scheme_targetpop_sets %>%
      select(stset_id, targetpops_strata),
    by = "stset_id"
  ) %>% 
  select(-stset_id) %>% 
  unnest(targetpops_strata)
```

Creating unioned set of target populations using strata.

```{r}
targetpops_strata <- 
  targetpops %>% 
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>% 
  select(-type)
```

Beware that the target populations in `targetpops_strata` may be too wide for application to a scheme in a specific module.

## Generating spatial sampling frames per scheme

### Applying scheme-specific spatial restrictions to certain types

By default, the spatial sampling frame of a scheme is the base sampling frame, subsetted by the scheme's target population.
This is a restriction that can easily be applied on the fly while defining the sample for each scheme.

However **some scheme x stratum combinations** apply a further spatial restriction of the spatial sampling frame.

To this end, we shrink the base sampling frame in different ways, in order to replace the base sampling frames for these schemes.

```{r paged.print=FALSE}
schemes %>% 
  filter(!is.na(spatial_restriction)) %>% 
  select(scheme, spatial_restriction)
```

```{r}
schemes %>% 
  distinct(spatial_restriction) %>% 
  filter(!is.na(spatial_restriction)) %>% 
  pull(spatial_restriction)
```

- **Restriction 1**: _For terrestrial types: restrict to zones with shallow groundwater (in reach of vegetation). Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Masking `grts_mh_n2khab` according to the `shallowgroundwater` layer.
To be conservative, we don't require cell centers to intersect the polygons.

```{r}
sg <- 
  locate_n2khab_data() %>% 
  file.path("10_raw/shallowgroundwater/shallowgroundwater.gpkg") %>% 
  vect()
sg_grts_n2khab <- 
  mask(grts_mh_n2khab, sg) %>% 
  values(na.rm = TRUE) %>% 
  as.vector()
```

Preparing restricted result `stratum_grts_n2khab_phabc_1`.

```{r}
strata_for_restriction_1 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep != "GD2", hydr_class != "HC3") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
strata_gd2 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep == "GD2") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
stratum_grts_n2khab_phabc_1 <-
  stratum_grts_n2khab_phabcorrected %>% 
  group_by(grts_address) %>% 
  mutate(
    to_restrict = stratum %in% strata_for_restriction_1,
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(
    !to_restrict | 
      gd2_in_cell | 
      grts_address %in% sg_grts_n2khab
  ) %>% 
  select(stratum, grts_address)
```

- **Restriction 2**: _Restrict to zones with shallow groundwater (in reach of vegetation). Exclude aquatic (i.e. rivulet) forms of type 7220. Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Preparing restricted result `stratum_grts_n2khab_phabc_2`.

```{r}
grts_7220_allowed_2 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "rivulet") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_2 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_2) %>% 
  group_by(grts_address) %>% 
  mutate(
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(gd2_in_cell | grts_address %in% sg_grts_n2khab) %>% 
  select(stratum, grts_address)
```

- **Restriction 3**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable. Exclude locations where the surface water is not connected with (nor affected by) the phreatic aquifer_
- **Restriction 4**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable_

We combine both restrictions since the condition relating to surface water - groundwater interaction is to be assessed in situ.

Preparing restricted result `stratum_grts_n2khab_phabc_3`.

```{r}
grts_7220_allowed_3 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "mire") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_3 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_3)
```

- **Restriction 5**: _Restrict to 8310 marl quarries that are directly connected with (intersected by) a groundwater aquifer_

**Missing data**: new data source needed to apply this spatial restriction.

```{r}
# dummy steps
grts_8310_allowed_5 <- 
  extract(
    grts_mh_n2khab, 
    st_coordinates(points_8310)
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_5 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "8310" | grts_address %in% grts_8310_allowed_5)
```


### Defining spatial sampling frames

```{r}
sp_samplingframes <- 
  tribble(
    ~sampling_frame_id, ~sampling_frame,
    0L, stratum_grts_n2khab_phabcorrected,
    1L, stratum_grts_n2khab_phabc_1,
    2L, stratum_grts_n2khab_phabc_2,
    3L, stratum_grts_n2khab_phabc_3,
    5L, stratum_grts_n2khab_phabc_5,
  )
```

Deriving the domain-wise sampling frames.

```{r}
sp_samplingframe_domain <- 
  sp_samplingframes %>%
   mutate(
     sampling_frame_domain = map(
       sampling_frame,
       ~ inner_join(
         .,
         domain_grts_n2khab,
         by = "grts_address",
         relationship = "many-to-many"
       ) %>%
         relocate(domain)
     ),
     .keep = "unused"
   )
```


### Setting spatial sampling frames per scheme

Note that these spatial sampling frames generally still contain strata that don't belong to the scheme's target population, nor belong to the types that exist in a specific domain with preset sample size.

This means we must always join with the scheme's target population (in `scheme_targetpop_sets$targetpops_strata`) _and_ with a domain-specific set of units (`domain_grts_n2khab`).

```{r}
scheme_sampling_frame <- 
  schemes %>% 
  select(scheme) %>% 
  mutate(
    sampling_frame_id = case_match(
      scheme,
      c("GW_03.3", "GW_04.2", "GW_05.2", "GW_07.1") ~ 1L,
      "GW_05.1_terr" ~ 2L,
      "GW_05.1_quarries" ~ 5L,
      c("GW_05.1_aq", "SURF_03.4_lotic") ~ 3L,
      .default = 0L
    )
  )
```


## Calculating scheme-specific population sizes per domain

### Precalculations at the sampling frame level

Determining the sampling frame specific population size per domain for strata with non-cell-based units.

```{r}
ssf_domain_stratum_non_cell_nunits <- 
  stratum_units_non_cell_n2khab %>% 
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>% 
  inner_join(
    sp_samplingframe_domain %>% unnest(sampling_frame_domain), 
    by = c("stratum", "grts_address"),
    relationship = "many-to-many"
  ) %>% 
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Determining the sampling frame specific population size per domain for strata with cell-center-based units.

<!-- Below simplifications (either counting cells in phab-corrected sampling frames, or representing domains and polygons + their intersection as cells and then summing phab) can be slightly improved by actually intersecting polygons by domains, then calculating phab-corrected areas (see git history for such an approach but still lacking st_intersection()). -->

```{r}
ssf_domain_stratum_cellcenter_nunits <-
  if (phabcorrection_done) {
    sp_samplingframe_domain %>% 
      unnest(sampling_frame_domain) %>% 
      semi_join(
        n2khab_types_expanded_properties %>% 
          filter(sample_support_code == "cell_conditioned_on_center"),
        by = c(stratum = "type")
      ) %>% 
      count(sampling_frame_id, domain, stratum, name = "nunits")
  } else {
    hmt_pol_phab_cellcenter_n2khab %>% 
      rename(stratum = type) %>% 
      collapse_strata() %>% 
      summarize(phab = sum(phab), .by = c(polygon_id, stratum)) %>% 
      inner_join(
        hmt_pol_grts_cellcenter, 
        by = "polygon_id",
        relationship = "many-to-many"
      ) %>% 
      select(-polygon_id) %>% 
      inner_join(
        sp_samplingframe_domain %>% unnest(sampling_frame_domain),
        by = c("stratum", "grts_address")
      ) %>% 
      summarize(nunits = sum(phab) / 100, .by = c(sampling_frame_id, domain, stratum)) %>% 
      arrange(across(-nunits))
  }
```

Determining the sampling frame specific population size per domain for strata with cell-based units.

```{r}
ssf_domain_stratum_cell_nunits <- 
  sp_samplingframe_domain %>% 
  unnest(sampling_frame_domain) %>% 
  semi_join(stratum_grts_cell_n2khab, by = c("stratum", "grts_address")) %>% 
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
ssf_domain_stratum_nunits <- 
  bind_rows(
    ssf_domain_stratum_non_cell_nunits,
    ssf_domain_stratum_cellcenter_nunits,
    ssf_domain_stratum_cell_nunits
  ) %>% 
  arrange(across(-nunits)) %>% 
  mutate(nunits = round(nunits) %>% as.integer())
```


### Applying result at the scheme level

```{r}
scheme_ssf_domain_stratum_nunits <- 
  scheme_sampling_frame %>% 
  inner_join(
    ssf_domain_stratum_nunits, 
    by = "sampling_frame_id",
    relationship = "many-to-many"
    ) %>% 
  semi_join(targetpops_strata, by = c("scheme", "stratum"))
```


### Filtering result per module

Each module has a restricted set of schemes, associated target populations, and domains.

```{r}
mod_dom_scheme_ssf_stratum_nunits <- 
  scheme_ssf_domain_stratum_nunits %>% 
  inner_join(
    module_domains %>% select(module, domain), 
    ., 
    by = "domain",
    relationship = "many-to-many"
    ) %>% 
  semi_join(module_targetpops_strata, by = c("module", "scheme", "stratum"))
```


## Determining schemes per domain for each module

Some of the domains also play a role in scheme implementation (see \@ref(domains-per-module)).

Determining the schemes that apply to each domain in each module, according to the _current_ (!) presence of strata in domains.

```{r}
module_domain_schemes <- 
  mod_dom_scheme_ssf_stratum_nunits %>% 
  distinct(module, domain, scheme)
```

Calculating statistics of the population size per **stratum**, per module x domain x scheme.

```{r}
module_domain_scheme_stratumstats <-
  mod_dom_scheme_ssf_stratum_nunits %>% 
  summarize(
    stratum_count = n(),
    stratum_nunits_min = min(nunits),
    stratum_nunits_q1 = quantile(nunits, 0.25),
    stratum_nunits_q2 = quantile(nunits, 0.5),
    stratum_nunits_q3 = quantile(nunits, 0.75),
    stratum_nunits_max = max(nunits),
    .by = c(module, domain, scheme)
  ) %>% 
  arrange(module, domain, scheme)
```

Calculating statistics of the population size per **type**, per module x domain x scheme.

```{r}
module_domain_scheme_typestats <-
  mod_dom_scheme_ssf_stratum_nunits %>% 
  inner_join(n2khab_strata, by = "stratum") %>% 
  summarize(nunits = sum(nunits), .by = c(module, domain, scheme, type)) %>%
  summarize(
    type_count = n(),
    type_nunits_min = min(nunits),
    type_nunits_q1 = quantile(nunits, 0.25),
    type_nunits_q2 = quantile(nunits, 0.5),
    type_nunits_q3 = quantile(nunits, 0.75),
    type_nunits_max = max(nunits),
    .by = c(module, domain, scheme)
  ) %>% 
  arrange(module, domain, scheme)
```

<!-- XXXXX Note that quite some polygons have been dropped in domain_hmt_pol_grts after the join with hmt_pol_grts_cellcenter was added, leading to a clearly lower number of units; this may be because of the clipping effect of grts_mh (which is too small) and also because of dropping small polygons (small=FALSE), i.e. that didn't intersect any cell center. -->
<!-- This effect should be scrutinized further, but this effect may be want we want for now. -->
<!-- XXXXX Above result is to be confronted to the ncells result after PHAB correction has been applied -->


## Check the type overlap between core and non-core MNE schemes, by environmental compartment

Determining which types in a module x compartment (GW, SURF, SOIL) are covered by a non-core scheme and _not_ by a core scheme.

```{r}
non_core_types_per_module_and_compartment <- 
  modules %>% 
  semi_join(
    module_domains %>% filter(domain == "Flanders"),
    by = "module"
  ) %>% 
  inner_join(
    scheme_targetpop_sets %>% 
      select(stset_id, scheme_set, targetpops_strata) %>% 
      mutate(
        scheme_targetpop = map2(
          scheme_set, targetpops_strata, function(ss, tp) {
            inner_join(ss, tp, by = "scheme")
          }
        ),
        .keep = "unused"
      ) %>% 
      unnest(cols = scheme_targetpop),
    by = "stset_id"
  ) %>% 
  inner_join(n2khab_strata, by = "stratum") %>% 
  filter(programme == "MNE") %>% 
  select(
    module,
    scheme,
    is_core_scheme, 
    compartment = attribute_1,
    type
  ) %>%
  arrange(module, compartment, type) %>% 
  mutate(
    compartment_has_core_scheme = any(is_core_scheme), 
    .by = c(module, compartment, type)
  ) %>% 
  mutate(
    all_compartments_have_core_scheme = all(compartment_has_core_scheme),
    .by = c(module, type)
  ) %>% 
  filter(!all_compartments_have_core_scheme) %>% 
  select(module, compartment, scheme, is_core_scheme, type)
non_core_types_per_module_and_compartment
```

The occurrence of types with non-predetermined sample size is very limited.

## Set the module-specific properties of schemes x variable-sets

Some schemes can be dropped depending on the module.

```{r}
year_first_default <- 2025L
mhq_proportion_aquatic_per_module <- 
  scheme_ssf_domain_stratum_nunits %>% 
  inner_join(n2khab_strata, by = "stratum") %>% 
  semi_join(read_types() %>% filter(str_detect(hydr_class, "3$")), by = "type") %>% 
  filter(str_detect(scheme, "^HQ")) %>% 
  inner_join(
    module_domains %>% filter(sample_size_predetermined) %>% select(module, domain),
    by = "domain",
    relationship = "many-to-many"
  ) %>% 
  summarize(
    nunits = sum(nunits),
    .by = c(module, scheme)
  ) %>% 
  mutate(
    prop_in_module = nunits / sum(nunits),
    .by = module,
    .keep = "unused"
  ) %>% 
  arrange(across(-prop_in_module))
```

Joining module and scheme properties to `domain_scheme_stats`, needing those attributes to generate further attributes, keeping them duplicated here for now.

```{r}
modules %>% 
  inner_join(module_domains, by = "module") %>% 
  filter(sample_size_predetermined) %>% 
  select(-sample_size_predetermined) %>% 
  inner_join(
    domain_scheme_stats %>% select(domain, scheme, stratum_count),
    by = "domain",
    relationship = "many-to-many"
  ) %>% 
  # step to drop schemes depending on module:
  filter(includes_mhq | !str_detect(scheme, "^HQ")) %>% 
  select(-includes_mhq) %>% 
  inner_join(schemes %>% select(scheme, attribute_1, tag_1), by = "scheme") %>% 
  inner_join(scheme_design, by = "scheme") %>% 
  mutate(
    year_first = case_match(
      attribute_1,
      "SURF" ~ year_first_default + 1L,
      "SOIL" ~ year_first_default + 2L,
      .default = year_first_default
    ),
    panel_count = as.integer(cycle_duration_y * years(1) / panel_period)
  ) %>% 
  inner_join(
    scheme_variableset_fieldmethod,
    by = "scheme",
    relationship = "many-to-many"
  ) %>% 
  left_join(mhq_proportion_aquatic_per_module, by = c("module", "scheme")) %>% 
  mutate(
    sp_sample_size_all_panels = cycle_duration_y * case_when(
      tag_1 == "secondary" ~ NA_integer_,
      module == "phase_1_flanders" & attribute_1 == "GW" ~ 300,
      module == "phase_1_flanders" & attribute_1 == "SOIL" ~ 333,
      module == "phase_1_flanders" & scheme == "SURF_03.4_lentic" ~ 106,
      module == "phase_1_flanders" & scheme == "SURF_03.4_lotic" ~ 27,
      module == "phase_1_sac_subset_5" & attribute_1 == "GW" ~ 50,
      module == "phase_1_sac_subset_5" & attribute_1 == "SOIL" ~ 56,
      module == "phase_1_sac_subset_5" & scheme == "SURF_03.4_lentic" ~ 18,
      module == "phase_1_sac_subset_5" & scheme == "SURF_03.4_lotic" ~ 4,
      module == "phase_1_be2100024" & attribute_1 == "GW" ~ 70,
      # still need to split this one depending on variable set:
      module == "phase_1_be2100024" & attribute_1 == "SOIL" ~ 133,
      module == "phase_1_be2100024" & scheme == "SURF_03.4_lentic" ~ 14,
      module == "phase_1_be2100024" & scheme == "SURF_03.4_lotic" ~ 4,
      module == "phase_1_be2100024" & str_detect(scheme, "^HQ3") ~ 18 * prop_in_module,
      module == "phase_1_be2100024" & str_detect(scheme, "^HQ(23|40|91(20|60))") ~ 28,
      module == "phase_1_be2100024" & str_detect(scheme, "^HQ(6230|9190)") ~ 84,
      module == "phase_1_be2100024" & scheme == "HQ7140" ~ 84,
      # still missing here: MHQ schemes in phase_1_sac_subset_5
      # NA in MHQ means a census instead of a sample
      .default = NA_integer_
    ) %>% 
      as.integer()
  ) %>% 
  select(-attribute_1, -tag_1, -prop_in_module) %>% 
  glimpse
  # relocate(module, domain, scheme, ...)
```





