# Generating scheme-specific objects in preparation of generating samples

These data objects are restricted to the schemes and types defined by `targetpops`.

## Defining the target populations using strata

This is needed to extract scheme- or stratum-specific sampling frames.

```{r}
targetpops_strata <- 
  targetpops %>% 
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>% 
  select(-type)
```

## Determining schemes per domain

Some of the domains also play a role in scheme implementation (see \@ref(domains-per-module)).

The schemes that apply to each domain, according to the _current_ (!) presence of strata in domains:

```{r}
domain_scheme_stats <-
  targetpops_strata %>% 
  inner_join(domain_stratum_nunits, by = "stratum", relationship = "many-to-many") %>% 
  summarize(
    stratum_count = n(),
    stratum_nunits_min = min(nunits),
    stratum_nunits_q1 = quantile(nunits, 0.25),
    stratum_nunits_q3 = quantile(nunits, 0.75),
    stratum_nunits_max = max(nunits),
    .by = c(domain, scheme)
  ) %>% 
  arrange(domain, scheme)
```

## Generating spatial sampling frames per scheme

### Applying scheme-specific spatial restrictions to certain types

By default, the spatial sampling frame of a scheme is the base sampling frame, subsetted by the scheme's target population.
This is a restriction that can easily be applied on the fly while defining the sample for each scheme.

However **some scheme x stratum combinations** apply a further spatial restriction of the spatial sampling frame.

To this end, we shrink the base sampling frame in different ways, in order to replace the base sampling frames for these schemes.

```{r}
schemes %>% 
  filter(!is.na(spatial_restriction)) %>% 
  select(scheme, spatial_restriction)
```

```{r}
schemes %>% 
  distinct(spatial_restriction) %>% 
  filter(!is.na(spatial_restriction)) %>% 
  pull(spatial_restriction)
```

- **Restriction 1**: _For terrestrial types: restrict to zones with shallow groundwater (in reach of vegetation). Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Masking `grts_mh_n2khab` according to the `shallowgroundwater` layer.
To be conservative, we don't require cell centers to intersect the polygons.

```{r}
sg <- 
  locate_n2khab_data() %>% 
  file.path("10_raw/shallowgroundwater/shallowgroundwater.gpkg") %>% 
  vect()
sg_grts_n2khab <- 
  mask(grts_mh_n2khab, sg) %>% 
  values(na.rm = TRUE) %>% 
  as.vector()
```

Preparing restricted result `stratum_grts_n2khab_phabc_1`.

```{r}
strata_for_restriction_1 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep != "GD2", hydr_class != "HC3") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
strata_gd2 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep == "GD2") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
stratum_grts_n2khab_phabc_1 <-
  stratum_grts_n2khab_phabcorrected %>% 
  group_by(grts_address) %>% 
  mutate(
    to_restrict = stratum %in% strata_for_restriction_1,
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(
    !to_restrict | 
      gd2_in_cell | 
      grts_address %in% sg_grts_n2khab
  ) %>% 
  select(stratum, grts_address)
```

- **Restriction 2**: _Restrict to zones with shallow groundwater (in reach of vegetation). Exclude aquatic (i.e. rivulet) forms of type 7220. Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Preparing restricted result `stratum_grts_n2khab_phabc_2`.

```{r}
grts_7220_allowed_2 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "rivulet") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_2 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_2) %>% 
  group_by(grts_address) %>% 
  mutate(
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(gd2_in_cell | grts_address %in% sg_grts_n2khab) %>% 
  select(stratum, grts_address)
```

- **Restriction 3**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable. Exclude locations where the surface water is not connected with (nor affected by) the phreatic aquifer_
- **Restriction 4**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable_

We combine both restrictions since the condition relating to surface water - groundwater interaction is to be assessed in situ.

Preparing restricted result `stratum_grts_n2khab_phabc_3`.

```{r}
grts_7220_allowed_3 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "mire") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_3 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_3)
```




