# Generating scheme-specific objects in preparation of generating samples {#scheme-objects}

These data objects are restricted to the schemes and types defined by `targetpops`.

## Defining the target populations using strata

This is needed to extract scheme- or stratum-specific sampling frames.

```{r}
targetpops_strata <- 
  targetpops %>% 
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>% 
  select(-type)
```

## Generating spatial sampling frames per scheme

### Applying scheme-specific spatial restrictions to certain types

By default, the spatial sampling frame of a scheme is the base sampling frame, subsetted by the scheme's target population.
This is a restriction that can easily be applied on the fly while defining the sample for each scheme.

However **some scheme x stratum combinations** apply a further spatial restriction of the spatial sampling frame.

To this end, we shrink the base sampling frame in different ways, in order to replace the base sampling frames for these schemes.

```{r}
schemes %>% 
  filter(!is.na(spatial_restriction)) %>% 
  select(scheme, spatial_restriction)
```

```{r}
schemes %>% 
  distinct(spatial_restriction) %>% 
  filter(!is.na(spatial_restriction)) %>% 
  pull(spatial_restriction)
```

- **Restriction 1**: _For terrestrial types: restrict to zones with shallow groundwater (in reach of vegetation). Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Masking `grts_mh_n2khab` according to the `shallowgroundwater` layer.
To be conservative, we don't require cell centers to intersect the polygons.

```{r}
sg <- 
  locate_n2khab_data() %>% 
  file.path("10_raw/shallowgroundwater/shallowgroundwater.gpkg") %>% 
  vect()
sg_grts_n2khab <- 
  mask(grts_mh_n2khab, sg) %>% 
  values(na.rm = TRUE) %>% 
  as.vector()
```

Preparing restricted result `stratum_grts_n2khab_phabc_1`.

```{r}
strata_for_restriction_1 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep != "GD2", hydr_class != "HC3") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
strata_gd2 <- 
  n2khab_types_expanded_properties %>% 
  filter(groundw_dep == "GD2") %>% 
  semi_join(n2khab_strata_expanded, ., by = "type") %>% 
  pull(stratum)
stratum_grts_n2khab_phabc_1 <-
  stratum_grts_n2khab_phabcorrected %>% 
  group_by(grts_address) %>% 
  mutate(
    to_restrict = stratum %in% strata_for_restriction_1,
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(
    !to_restrict | 
      gd2_in_cell | 
      grts_address %in% sg_grts_n2khab
  ) %>% 
  select(stratum, grts_address)
```

- **Restriction 2**: _Restrict to zones with shallow groundwater (in reach of vegetation). Exclude aquatic (i.e. rivulet) forms of type 7220. Don't exclude locations with types that are '(almost) everywhere groundwater dependent' (GD2)_

Preparing restricted result `stratum_grts_n2khab_phabc_2`.

```{r}
grts_7220_allowed_2 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "rivulet") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_2 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_2) %>% 
  group_by(grts_address) %>% 
  mutate(
    gd2 = stratum %in% strata_gd2,
    gd2_in_cell = any(gd2)
  ) %>% 
  ungroup() %>% 
  filter(gd2_in_cell | grts_address %in% sg_grts_n2khab) %>% 
  select(stratum, grts_address)
```

- **Restriction 3**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable. Exclude locations where the surface water is not connected with (nor affected by) the phreatic aquifer_
- **Restriction 4**: _Exclude terrestrial (i.e. mire) forms of type 7220, if applicable_

We combine both restrictions since the condition relating to surface water - groundwater interaction is to be assessed in situ.

Preparing restricted result `stratum_grts_n2khab_phabc_3`.

```{r}
grts_7220_allowed_3 <- 
  extract(
    grts_mh_n2khab, 
    habspring_units %>% 
      filter(system_type != "mire") %>% 
      st_coordinates()
  ) %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  pull(GRTSmaster_habitats)
stratum_grts_n2khab_phabc_3 <-
  stratum_grts_n2khab_phabcorrected %>% 
  filter(stratum != "7220" | grts_address %in% grts_7220_allowed_3)
```


### Defining spatial sampling frames

```{r}
sp_samplingframes <- 
  tribble(
    ~sampling_frame_id, ~sampling_frame,
    0L, stratum_grts_n2khab_phabcorrected,
    1L, stratum_grts_n2khab_phabc_1,
    2L, stratum_grts_n2khab_phabc_2,
    3L, stratum_grts_n2khab_phabc_3,
  )
```

Deriving the domain-wise sampling frames.

```{r}
sp_samplingframe_domain <- 
  sp_samplingframes %>%
   mutate(
     sampling_frame_domain = map(
       sampling_frame,
       ~ inner_join(
         .,
         domain_grts_n2khab,
         by = "grts_address",
         relationship = "many-to-many"
       ) %>%
         relocate(domain)
     ),
     .keep = "unused"
   )
```


### Setting spatial sampling frames per scheme

Note that these spatial sampling frames generally still contain strata that don't belong to the scheme's target population, nor belong to the types that exist in a specific domain with preset sample size.

This means we must always join with the scheme's target population (in `targetpops_strata`) _and_ with a domain-specific set of units (`domain_grts_n2khab`).

```{r}
scheme_sampling_frame <- 
  schemes %>% 
  select(scheme) %>% 
  mutate(
    sampling_frame_id = case_match(
      scheme,
      c("GW_03.3", "GW_04.2", "GW_05.2", "GW_07.1") ~ 1L,
      "GW_05.1_terr" ~ 2L,
      c("GW_05.1_aq", "SURF_03.4_lotic") ~ 3L,
      .default = 0L
    )
  )
```


## Calculating scheme-specific population sizes per domain

### Precalculations at the sampling frame level

Determining the sampling frame specific population size per domain for strata with non-cell-based units.

```{r}
ssf_domain_stratum_non_cell_nunits <- 
  stratum_units_non_cell_n2khab %>% 
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>% 
  inner_join(
    sp_samplingframe_domain %>% unnest(sampling_frame_domain), 
    by = c("stratum", "grts_address"),
    relationship = "many-to-many"
  ) %>% 
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Determining the sampling frame specific population size per domain for strata with cell-center-based units.

<!-- XXXXXX Below simplification (taking domain-overlapping polygons 
completely into account) can be improved by actually intersecting by domains in
calculating surface area. -->

```{r}
# polygon_ids and associated grts addresses that intersect each domain:
domain_hmt_pol_grts <-
  domains %>% 
  st_drop_geometry() %>% 
  select(domain) %>% 
  mutate(id = unclass(st_intersects(domains, hmt_pol))) %>% 
  unnest(id) %>% 
  inner_join(
    tibble(
      id = seq_len(nrow(hmt_pol)),
      polygon_id = hmt_pol$polygon_id
    ),
    by = "id"
  ) %>% 
  select(-id) %>% 
  inner_join(
    hmt_pol_grts_cellcenter, 
    by = "polygon_id",
    relationship = "many-to-many"
  )
# number of units per polygon_id x stratum (taking into account surface area & phab):
# (since we start from hmt, we also need to collapse the types to the n2khab_strata!)
hmt_pol_stratum_cellcenter_nunits <- 
  hmt_pol_phab_cellcenter_n2khab %>% 
  inner_join(
    hmt_pol %>% 
      mutate(area = st_area(.) %>% drop_units()) %>% 
      st_drop_geometry(),
    by = "polygon_id"
  ) %>% 
  mutate(nunits = area * phab / 100 / (32^2), .keep = "unused") %>% 
  rename(stratum = type) %>% 
  collapse_strata() %>% 
  summarize(
    nunits = sum(nunits),
    .by = c(polygon_id, stratum)
  )
# population size per domain x stratum, for each sampling frame:
ssf_domain_stratum_cellcenter_nunits <-
  domain_hmt_pol_grts %>% 
  inner_join(
    hmt_pol_stratum_cellcenter_nunits, 
    by = "polygon_id",
    relationship = "many-to-many"
  ) %>% 
  inner_join(
    sp_samplingframe_domain %>% unnest(sampling_frame_domain),
    by = c("domain", "stratum", "grts_address")
  ) %>% 
  distinct(sampling_frame_id, domain, stratum, polygon_id, nunits) %>% 
  summarize(nunits = sum(nunits), .by = c(sampling_frame_id, domain, stratum)) %>% 
  arrange(across(-nunits))
```

Determining the sampling frame specific population size per domain for strata with cell-based units.

```{r}
ssf_domain_stratum_cell_nunits <- 
  stratum_grts_cell_n2khab %>% 
  inner_join(
    sp_samplingframe_domain %>% unnest(sampling_frame_domain),
    by = c("stratum", "grts_address")
  ) %>% 
  count(sampling_frame_id, domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
ssf_domain_stratum_nunits <- 
  bind_rows(
    ssf_domain_stratum_non_cell_nunits,
    ssf_domain_stratum_cellcenter_nunits,
    ssf_domain_stratum_cell_nunits
  ) %>% 
  arrange(across(-nunits))
```

### Applying result at the scheme level

```{r}
scheme_ssf_domain_stratum_nunits <- 
  scheme_sampling_frame %>% 
  inner_join(
    ssf_domain_stratum_nunits, 
    by = "sampling_frame_id",
    relationship = "many-to-many"
    ) %>% 
  semi_join(targetpops_strata, by = c("scheme", "stratum"))
```


## Determining schemes per domain

Some of the domains also play a role in scheme implementation (see \@ref(domains-per-module)).

The schemes that apply to each domain, according to the _current_ (!) presence of strata in domains:

```{r}
domain_scheme_stats <-
  scheme_ssf_domain_stratum_nunits %>% 
  summarize(
    stratum_count = n(),
    stratum_nunits_min = min(nunits),
    stratum_nunits_q1 = quantile(nunits, 0.25),
    stratum_nunits_q2 = quantile(nunits, 0.5),
    stratum_nunits_q3 = quantile(nunits, 0.75),
    stratum_nunits_max = max(nunits),
    .by = c(domain, scheme)
  ) %>% 
  arrange(domain, scheme)
```

<!-- XXXXX Note that quite some polygons have been dropped in domain_hmt_pol_grts after the join with hmt_pol_grts_cellcenter was added, leading to a clearly lower number of units; this may be because of the clipping effect of grts_mh (which is too small) and also because of dropping small polygons (small=FALSE), i.e. that didn't intersect any cell center. -->
<!-- This effect should be scrutinized further, but this effect may be want we want for now. -->
<!-- XXXXX Above result is to be confronted to the ncells result after PHAB correction has been applied -->



