# Generating scheme-specific objects in preparation of generating samples

These data objects are restricted to the schemes and types defined by `targetpop`.

## Defining the target population as strata

This is needed to extract scheme- or stratum-specific sampling frames.

```{r}
targetpop_strata <- 
  targetpop %>% 
  inner_join(n2khab_strata, by = "type", relationship = "many-to-many") %>% 
  select(-type)
```

## Determining schemes per domain

The schemes that apply to each domain, according to the _current_ (!) presence of strata in domains:

```{r}
domain_scheme_stats <-
  targetpop_strata %>% 
  inner_join(domain_stratum_ncells, by = "stratum", relationship = "many-to-many") %>% 
  summarize(
    stratum_count = n(),
    stratum_ncells_min = min(ncells),
    stratum_ncells_q1 = quantile(ncells, 0.25),
    stratum_ncells_q3 = quantile(ncells, 0.75),
    stratum_ncells_max = max(ncells),
    .by = c(domain, scheme)
  ) %>% 
  arrange(domain, scheme)
```

## Generating spatial sampling frames per scheme

### Applying scheme-specific spatial restrictions to certain types

**Missing functionality**: apply spatial restrictions depending on scheme x stratum.



