# Generating domain properties

## Subsetting the base sampling frame according to domain

Determining all GRTS addresses of the base sampling frame per domain.

Carried out for all schemes in MNE + MHQ.

Since this does not distinguish between strata, stratum collapsing is irrelevant.

```{r eval = params$regenerate_binary_files}
domain_grts_n2khab <-
  extract(grts_mh_n2khab, vect(domains)) %>%
  as_tibble() %>% 
  filter(!is.na(GRTSmaster_habitats)) %>% 
  inner_join(
    tibble(
      ID = seq_len(nrow(domains)),
      domain = domains$domain
    ),
    .,
    by = "ID"
  ) %>% 
  select(
    -ID,
    grts_address = GRTSmaster_habitats
  )
saveRDS(
  domain_grts_n2khab, 
  file.path(datapath, "binary/intermediate/domain_grts_n2khab.rds")
)
```

```{r}
domain_grts_n2khab <-
  readRDS(file.path(
    datapath, 
    "binary/intermediate/domain_grts_n2khab.rds"
  ))
```

## Calculating the spatial population size of strata in each domain

Carried out for all schemes in MNE + MHQ.

Needed in preparation of generating samples.

### Approximately by using the phab-corrected data frame of stratum x grts addresses in MNE + MHQ

The phab-corrected (i.e. subsampled) data frame of stratum x grts addresses is one approach to approximate the population sizes; however beside inaccuracy due to phab-based subsampling, this is also compounded by the fact that some cells include multiple (non-cell) units that belong to the same stratum.

<!-- However we did NOT YET APPLY this phab correction!!
This results in an overestimation for the cell-centered types,
but it's still a good base to DROP strata from domains XXXXXX -->

```{r}
domain_stratum_ncells <- 
  stratum_grts_n2khab_phabcorrected %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "ncells")
```

### More correctly by combining different unit-level data sources

Determining the population size per domain for strata with non-cell-based units.

```{r}
domain_stratum_non_cell_nunits <- 
  stratum_units_non_cell_n2khab %>% 
  inner_join(
    units_non_cell_n2khab_grts,
    by = c("sample_support_code", "unit_id")
  ) %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "nunits")
```

Determining the population size per domain for strata with cell-center-based units.

<!-- XXXXXX Below simplicifaction (taking domain-overlapping polygons 
completely into account) can be improved by actually intersecting by domains. -->

```{r}
# polygon_ids that intersect each domain:
domain_hmt_pol <- 
  domains %>% 
  st_drop_geometry() %>% 
  select(domain) %>% 
  mutate(id = unclass(st_intersects(domains, hmt_pol))) %>% 
  unnest(id) %>% 
  inner_join(
    tibble(
      id = seq_len(nrow(hmt_pol)),
      polygon_id = hmt_pol$polygon_id
    ),
    by = "id"
  ) %>% 
  select(-id)
# number of units per polygon_id x stratum (taking into account surface area & phab):
# (since we start from hmt, we also need to collapse the types to the n2khab_strata!)
hmt_pol_stratum_cellcenter_nunits <- 
  hmt_pol_phab_cellcenter_n2khab %>% 
  inner_join(
    hmt_pol %>% 
      mutate(area = st_area(.) %>% drop_units()) %>% 
      st_drop_geometry(),
    by = "polygon_id"
  ) %>% 
  mutate(nunits = area * phab / 100 / (32^2), .keep = "unused") %>% 
  rename(stratum = type) %>% 
  collapse_strata() %>% 
  summarize(
    nunits = sum(nunits),
    .by = c(polygon_id, stratum)
  )
# population size per domain x stratum:
domain_stratum_cellcenter_nunits <- 
  domain_hmt_pol %>% 
  inner_join(
    hmt_pol_stratum_cellcenter_nunits, 
    by = "polygon_id",
    relationship = "many-to-many"
  ) %>% 
  summarize(nunits = sum(nunits), .by = c(domain, stratum))
```

Determining the population size per domain for strata with cell-based units.

```{r}
domain_stratum_cell_nunits <- 
  stratum_grts_cell_n2khab %>% 
  inner_join(domain_grts_n2khab, by = "grts_address", relationship = "many-to-many") %>% 
  count(domain, stratum, name = "nunits")
```

Integrating these 3 different types of strata.

```{r}
domain_stratum_nunits <- 
  bind_rows(
    domain_stratum_non_cell_nunits,
    domain_stratum_cellcenter_nunits,
    domain_stratum_cell_nunits
  )
```

Check that this coincides with the domain x stratum combinations of `domain_stratum_ncells`.

```{r}
inner_join(
  domain_stratum_ncells, 
  domain_stratum_nunits, 
  by = c("domain", "stratum")
) %>% 
  nrow() == nrow(domain_stratum_nunits)
```

## Determining schemes per domain

Carried out for the schemes in the first phase.

The schemes that apply to each domain, according to the _current_ (!) presence of strata in domains:

```{r}
domain_scheme_stats <-
  targetpop_strata %>% 
  inner_join(domain_stratum_ncells, by = "stratum", relationship = "many-to-many") %>% 
  summarize(
    stratum_count = n(),
    stratum_ncells_min = min(ncells),
    stratum_ncells_q1 = quantile(ncells, 0.25),
    stratum_ncells_q3 = quantile(ncells, 0.75),
    stratum_ncells_max = max(ncells),
    .by = c(domain, scheme)
  ) %>% 
  arrange(domain, scheme)
```




