# Generating spatiotemporal samples {#sampling}

<!-- use hardcoded tibble from file 010 with design attributes per scheme -->

<!-- Take into account the revisit design, sample size, surplus sampling, domains -->

<!-- Sample sizes can be summed over active module-domains that refer to the same domain. -->
<!-- Where one module-domain overlaps (contains) another module-domain, those sample sizes cannot be added but an appropriate offset must be taken into account when sampling in the shared area (the smaller domain), in order to respect the GRTS ordering. -->
<!-- This also holds for MHQ Flanders (not in this POC) vs MHQ in SACs. -->

<!-- For this proof of concept, we restrict the allowed modules. -->

<!-- XXX Reconsider this: is module subselection actually needed? -->

<!-- Note that this is not needed! -->
<!-- Better just work with active modules, and drop the below. -->
<!-- Let's make this consideration after revisiting the revisit design code! -->

```{r eval=FALSE, include=FALSE}
module_selection <- if (all(
  c("mne2024_pan_effectmon_sac_subset_5", "mne2024_pan_effectmon_flanders") %in%
    params$active_modules
)) {
  c("mne2024_pan_effectmon_sac_subset_5", "mne2024_pan_effectmon_flanders")
} else {
  params$active_modules
}
module_domain_scheme_stratum_sample_size_2 <-
  module_domain_scheme_stratum_sample_size %>%
  filter(module %in% module_selection)
```

## Constructing the samples

The samples don't explicitly address the non-core schemes, except for its types that are not covered by core schemes.
There's no intention to do that, as for overlapping types the samples of the core schemes are the ones to be used for the non-core schemes.

Module labelling of sampling units: this is needed to apply different revisit designs to different modules, if applicable.
This should be tested first, because if module panel designs align, it's better for the membership design if all units are taken together.
Once units are taken together, it's difficult to pull them apart afterwards, since combining different modules with different revisit designs is like creating a split panel design: in the same period, you need two different sets of sampling units that receive another revisit design.
The conflict is at what levels you want to guarantee a single strict GRTS series: within the module, within the year (conflicts with previous, supposing that each module should cover the full cycle) ...
Since it's most reasonable to spread each module over the full cycle, this means that GRTS gaps will be inevitable either within a panel period (each module is a GRTS series split into periods, so for two or more modules this combines different GRTS subsets in the same period) or within a module (each panel is a GRTS series assigned to consecutive modules), notably between the different modules.

- Combining modules with the same revisit design: panels can be created for all units, afterwards each panel can be divided over both modules.
This optimizes sampling design & inference for the module combination.
  - Note that the panel _construction_ must be done for each domain separately, i.e. the overlapped domains and the non-overlapped (remainder) domain.
  Otherwise you would have that the higher GRTS addresses added in the included domains (~ other module) end up in the latest panels, hence not well spread in time.
  - Clearly, also the division of panels between modules only applies to the domains where both modules overlap.
- Combining modules with different revisit design: sampling units must be assigned to modules a priori, each module has its own GRTS series + spare addresses (hence mandatory gap needed).
Then panel designs are created for each module separately.

And now also think about the multiple revisit designs per scheme (multiple FAGs) and which one drives the foregoing choice.
Also, think about common field activity supergroups x sample size between module x schemes and how to proceed.

For now still without:

- GRTS address offset for MHQ at Flemish level = undefined module
- adding (and labelling) spare sampling units. These can be shared among modules when these have the same panel design, but their module membership is applied at the moment of replacement, according to module the replaced unit belonged to.
- collapsing nearby units for groundwater schemes
- gradual increase, during the first cycle, of the number of installed groundwater and surfacewater observation wells.
Also, the fact that the _installation_ of observation wells must be accelerated in the module's first cycle if the latter is shorter than the enforced cycle duration.
So it's currently assumed that all observation wells are in place already.

### Module combinations (module combos) per scheme based on common panel design of target variables

Determining module combos per scheme based on common panel design characteristics _of target variables_.
The target variables limit the main field methods of a module x scheme to a great extent, since the latter also comprise variables for explanation & quality control.

First create an object with the 'target' panel designs per module and scheme (target referring to the target variables restriction).

```{r}
mod_scheme_targetfieldmethod <- 
  mod_scheme_mainfieldmethod %>% 
  semi_join(
    envvars %>% 
      filter(target),
    join_by(module, scheme, main_field_method == main_datacollection_method)
  ) %>% 
  select(-is_core_scheme)
mod_scheme_targetfieldmethod_paneldesign <- 
  mod_scheme_targetfieldmethod %>% 
  # expand with FAG (note that some main field methods link to more than one
  # FAG, which has to do with the spatial subsets, which we can ignore here
  # since in the case of the MAIN field method these don't cause differences in
  # panel designs but only in activities.)
  inner_join(
    mod_scheme_field_activity %>% 
      distinct(module, scheme, field_activity, field_activity_group),
    join_by(module, scheme, main_field_method == field_activity),
    relationship = "one-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  # add the FAG panel design
  inner_join(
    mod_scheme_fag_paneldesign %>% 
      distinct(pick(module, scheme, field_activity_group, panel_split, revisit_pattern:last_col())),
    join_by(module, scheme, field_activity_group),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  distinct(pick(
    module,
    scheme,
    main_field_method,
    panel_split,
    revisit_pattern:last_col()
  ))
if (nrow(mod_scheme_targetfieldmethod_paneldesign) != nrow(mod_scheme_targetfieldmethod)) {
  stop("The number of rows changed unexpectedly.")
}
mod_scheme_target_paneldesign <- 
  mod_scheme_targetfieldmethod_paneldesign %>% 
  distinct(pick(-main_field_method, -enforced_cycle_duration))
```

Now determining the module combos per scheme.

```{r}
scheme_target_paneldesign <- 
  mod_scheme_target_paneldesign %>% 
  nest(module_combo = module) %>% 
  relocate(module_combo) %>% 
  arrange(scheme, panel_period_main, cycle_duration_y)
scheme_target_paneldesign
```

We end up with `r nrow(scheme_target_paneldesign)` different scheme x target panel design combinations, each of which applies to one or several modules.

Note that _different_ module combos within a scheme are handled as if they were different panel splits, i.e. they need a separate portion of the GRTS series to sample from; the starting address could be set based on sample size proportion but after that the starting address should be held fixed.
It should also be accepted that for rare types, these samples may by necessity overlap between module combos hence this should be simplified afterwards to avoid unnecessary duplication of fieldwork planning.

### Spatial samples

Determining the applicable domain relations.

```{r}
domain_relations_spsamples <- 
  domain_overlaps %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size %>% 
      distinct(module, domain),
    join_by(domain)
  ) %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size %>% 
      distinct(module, domain),
    join_by(included_by == domain)
  )
```

Drawing spatial samples.

```{r}
scheme_domain_stratum_spsamples <-
  module_domain_scheme_stratum_sample_size %>%
  select(module, domain, scheme, sampling_frame_id, stratum, sp_sample_size_all_panels_stratum) %>%
  # for MNE, non-core schemes (dropped below) have yet to be taken back (but I
  # think this step also drops things we always need to drop).
  # This is relevant for specifically added types for those non-core-schemes.
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>%
  # domains that are the same between different modules are integrated by
  # summing the sample sizes; the module is dropped:
  summarize(
    sp_sample_size_all_panels_stratum = sum(sp_sample_size_all_panels_stratum),
    .by = c(domain, scheme, sampling_frame_id, stratum)
  ) %>%
  nest(domain_stratum_spsamplesize = c(domain, stratum, sp_sample_size_all_panels_stratum)) %>%
  inner_join(sp_samplingframe_domain, by = join_by(sampling_frame_id)) %>%
  # drawing an independent sample per scheme, domain and stratum:
  mutate(
    spsample = map2(domain_stratum_spsamplesize, sampling_frame_domain, function(dss, sfd) {
      sfd %>%
        nest(grts_addresses = grts_address) %>%
        inner_join(dss, ., by = join_by(domain, stratum)) %>%
        mutate(
          sample =
            map2(sp_sample_size_all_panels_stratum, grts_addresses, function(sss, grts) {
              grts %>% slice_min(grts_address, n = sss)
            })
        ) %>% 
        nest(domain_spsample = c(
          domain, 
          sp_sample_size_all_panels_stratum, 
          grts_addresses,
          sample
        ))
    })
  ) %>%
  select(scheme, spsample) %>%
  unnest(spsample) %>% 
  # handling the overlap between domains: above, the samples have been drawn
  # independently so now this has to be corrected!
  mutate(
    domain_spsample = map(domain_spsample, function(ds) {
      ds <- left_join(
        ds,
        domain_relations_spsamples,
        join_by(domain),
        relationship = "one-to-one",
        unmatched = "drop"
      )
      # vector of including domains:
      including_domains <- unique(ds$included_by)
      # drop NAs:
      including_domains <- including_domains[!is.na(including_domains)]
      # only keep an including domain if it is effectively present as domain:
      including_domains <- including_domains[including_domains %in% ds$domain]
      if (length(including_domains) > 1) {
        stop("Multiple including domains in one stratum are not supported.")
      }
      ds %>% 
        mutate(
          including = domain %in% including_domains | length(including_domains) == 0,
          # put the sample of the including domain next to the included domains:
          grts_sampled = ifelse(
            including, 
            tibble(),
            sample[including]
          ),
          # determine the available unsampled units per domain:
          grts_addresses_2 = map2(grts_addresses, grts_sampled, function(ga, gs) {
            if (is.null(gs)) {
              return(NULL)
            }
            anti_join(ga, gs, join_by(grts_address))
          }),
          # redraw samples for included domains:
          sample_2 = pmap(
            list(including, sample, grts_addresses_2, sp_sample_size_all_panels_stratum),
            function(inc, sam, grts, sss){
              if (inc) {
                sam
              } else {
                grts %>% slice_min(grts_address, n = sss)
              }
            }
          ),
          # flag sampling units of included domains, as they should never be
          # considered apart from the including domain:
          extends_on_larger_domain = !including & length(including_domains) == 1
        ) %>% 
        select(domain, sample = sample_2, extends_on_larger_domain) %>% 
        # next step will also drop domains where no sampling units were
        # available (i.e. the whole population was sampled by the larger domain):
        unnest(sample)
    })
  ) %>% 
  unnest(domain_spsample)
```

Check that no sampling units overlap between domains within the combination of scheme and stratum:

```{r}
scheme_domain_stratum_spsamples %>% 
  count(scheme, stratum, grts_address) %>% 
  filter(n > 1) %>% 
  nrow() == 0
```


### Implementing the revisit design

**Missing functionality**: make the code below work for other field activities than those in module `mne2024_pan_effectmon_flanders` (e.g. wrt prefix extraction).

<!-- --------------------------- -->
Adding some ideas:

- organize the membership design for the combination of field activity supergroups, i.e. with the same panel design across FAGs, and sample size. Still within module x scheme since sample sizes may differ, but across schemes where they don't.
- also make membership design depend between such different supergroups of the same sample size, where these have a compatible number of panels, e.g. if one panel in one supergroup can be the sum of 4 panels of another supergroup.
- start with the supergroups that coincide with the panel designs of the _main_ field method of _target_ variables, as we plan to optimize the spatial samples in order to deal with the representation of those panel designs across modules, schemes and domains.
- apply non-core FAG panels only to locations not covered by a corresponding core FAG
- select a single winning FAG for competition in time (for the same FAG) or between interlinked FAGs (take the superior one and schedule it as the first one). However don't do this for the FAG with the main field method of each scheme: the data collection design must not be harmed!
- for MNE, the LOCEVAL activity must take place one year beforehand!
- SPATPOSITxxx: inflate the panel periods by 3 months at the end (so they will overlap)
- EMERGENCYxxx: reduce the number of occasions by sampling them
- in the first year of each compartment, start on 1 Jan and wind up to month_first_default
<!-- --------------------------- -->

#### Number of panels

Determining the _number_ of panels for each module x scheme x spatial subset x 'field activity group' (FAG).
We also add period, group size and group repetition attributes.
Grouping refers to the subgroup structure in nested serially alternated patterns.

```{r}
mod_scheme_dom_fag_panel_count <-
  module_domain_scheme_stratum_sample_size %>%
  # to be revisited filter step (see above); this is especially important since
  # field activities are done e.g. for GW_05.1_aq (SURFLEVINST, SURFLEVREAD)
  # that will also be dropped later on, as a consequence of this.
  # So even though for a specific module the non-core schemes don't have their
  # own sample size, they may have their own field activities that must be
  # taken into account

  # filter(!is.na(sp_sample_size_all_panels_stratum)) %>%
  distinct(module, scheme, domain, year_first, year_last) %>%
  inner_join(
    mod_scheme_fag_paneldesign, 
    join_by(module, scheme),
    relationship = "many-to-many",
    unmatched = "error"
  ) %>%
  mutate(
    panel_count = case_when(
      revisit_pattern == "pure_panel" ~ 1L,
      .default = as.integer(cycle_duration_y * years(1) / panel_period_main)
    ) *
      case_when(
        is.na(revisit_subpattern) ~ 1L,
        revisit_subpattern == "pure_panel" ~ 1L,
        !is.na(limit_n_subpanels_per_panel) ~ limit_n_subpanels_per_panel,
        .default = as.integer(panel_period_main / panel_subperiod)
      ),
    panel_period = case_when(
      is.na(panel_subperiod) ~ panel_period_main,
      .default = panel_subperiod
    ),
    panel_group_size = ifelse(
      !is.na(limit_n_subpanels_per_panel),
      limit_n_subpanels_per_panel,
      1L
    ),
    panel_group_repetitions = case_when(
      !is.na(limit_n_subpanels_per_panel) ~ as.integer(
        panel_period_main / panel_subperiod
      ) / limit_n_subpanels_per_panel,
      revisit_subpattern == "pure_panel" ~ min(
        as.integer(panel_period_main / panel_subperiod),
        limit_n_subcycles,
        na.rm = TRUE
      ),
      .default = 1L
    )
  ) %>%
  select(
    module,
    scheme,
    is_core_scheme,
    in_aquatic_subset,
    domain,
    field_activity_group,
    starts_with("year"),
    panel_period,
    panel_count,
    panel_group_size,
    panel_group_repetitions
  )
mod_scheme_dom_fag_panel_count %>%
  glimpse()
```

Different occurring patterns at the compartment level, with 'HQ' referring to the MHQ 'compartment', for information and thought:

```{r}
mod_scheme_dom_fag_panel_count %>%
  mutate(
    compartment = str_extract(scheme, "^[A-Z]+") %>% 
      factor() %>% 
      fct_relevel("HQ", after = Inf)
  ) %>%
  summarize(
    field_activity_groups = str_c(unique(field_activity_group), collapse = ", "),
    .by = c(compartment, starts_with("panel"))
  ) %>%
  arrange(compartment, panel_period) %>% 
  kable()
```

#### Panels and their properties

Defining panels.

```{r}
scheme_domain_fag_panels <-
  scheme_domain_fag_panel_count %>%
  uncount(panel_count) %>%
  mutate(
    panel = str_c(
      scheme,
      field_activity_group,
      str_c("PAN", str_pad(seq_len(n()), 2, pad = "0")),
      sep = "_"
    ),
    .by = c(scheme, field_activity_group)
  ) %>%
  mutate(panel = factor(panel))
```

Defining date intervals and assigning panels to these intervals.

```{r}
scheme_domain_fag_panel_calendar <-
  scheme_domain_fag_panels %>%
  nest(
    revisit = c(starts_with("year"), starts_with("panel"), -panel),
    panels = panel
  ) %>%
  mutate(
    revisit = map(revisit, function(df) {
      df %>%
        distinct() %>%
        mutate(
          # taking into account month_first_default:
          nr_periods =
            (years(year_last - year_first + 1) -
              months(13 - month_first_default)) / panel_period
        ) %>%
        uncount(nr_periods) %>%
        mutate(
          panel_periodmonths = as.period(panel_period, "months") %>%
            month() %>%
            as.integer(),
          panel_cum_months = cumsum(panel_periodmonths),
          panel_cum_months_lag = lag(panel_cum_months, default = 0L),
          # we only start in month_first_default of first year:
          date_start = make_date(year_first, month_first_default, 1) + months(panel_cum_months_lag),
          date_end = make_date(year_first, month_first_default, 1) + months(panel_cum_months) - days(1),
          date_interval = interval(date_start, date_end)
        ) %>%
        select(-c(panel_periodmonths, starts_with("panel_cum_months")))
    }),
    panel_date_assignment = map2(revisit, panels, function(revis, pan) {
      group_size <- revis$panel_group_size[1]
      group_repetitions <- revis$panel_group_repetitions[1]
      if (group_size == 1 && group_repetitions == 1) {
        pan2 <- pan
      } else {
        indexes_pan <- seq_len(nrow(pan))
        indexgrouping_pan <- rep(seq_len(nrow(pan) / group_size), each = group_size)
        indexes_grouped <- split(indexes_pan, indexgrouping_pan) %>%
          rep(each = group_repetitions) %>%
          unlist() %>%
          unname()
        pan2 <- pan[indexes_grouped, ]
      }
      indexes <- rep(seq_len(nrow(pan2)), nrow(revis) %/% nrow(pan2) + 1) %>%
        head(nrow(revis))
      bind_cols(
        revis %>% select(date_start, date_end, date_interval),
        pan2[indexes, ]
      )
    })
  ) %>%
  select(-panels, -revisit) %>%
  unnest(panel_date_assignment) %>%
  relocate(panel, .after = field_activity_group)
```

#### Panel membership of spatial sampling units

Panel membership is randomized by following the GRTS address sequence at the stratum level.

To determine the number of sampling units per panel, it must be taken into account that the stratum sample size is not equal to, or a multiple of, the number of panels, so panels usually cannot have an equal number of units.
Local pivotal sampling of panels is applied to get an even load of all panels while maintaining maximum temporal spread in the membership.
One wants to avoid a systematic effect that the first panels always have more units than those further in line, or the last panel.
This becomes even more important for strata with few sampling units; sometimes this number is smaller than the theoretical number of panels!

The interrelation between panels within a compartment is controlled:

- To get maximum synergy among schemes within a compartment, we consider the unioned sample per compartment, split that one to panels, after which the splits can again be divided over the different field activity groups.
- To get maximum synergy among different field activity groups within a compartment, different panel counts are handled consecutively: e.g. if both '2 panels' and '12 panels' occur in a compartment, then membership design is first done for 2 panels, and the members of each panel are then split into 6 groups to define the membership of the 12 panels.
This avoids repeating local pivotal sampling more than necessary, which would lead to unneeded mismatches.

**Missing functionality**: maximizing panel synergy between compartments (note that this would also need taking into account different start years between compartments), or conversely, avoid some synergies even within a compartment, to avoid the problem described now.
The challenge is the different scheme specific sample sizes (this may already hold for the current approach of integration within a compartment, but this was not a problem for the first module).
This would lead to the schemes with smaller sample sizes not being equally spread over the panels.
This might be an optimal fieldwork approach, but it is not statistically optimal.
So from a statistical viewpoint it will be safer to use the scheme level and only optimize between field activity groups within a scheme.
There's no such clash between schemes when their field activities don't overlap, but when the same field activity occurs this problem will be present for strata that are represented in more than one scheme, if they were assigned their own sample size.
This situation doesn't occur in the specific case of the `mne2024_pan_effectmon_flanders` module.

**Missing functionality**: expanding to all compartments + MHQ.

**Missing functionality**: defining and taking into account temporal dependencies between field activity groups.
An example is that a piezometer must be installed before a water sample can be taken from it.

**Missing functionality**: provide a gradual increase, during the first cycle, of the number of installed groundwater and surfacewater observation wells.
This has consequences for the READ and the SAMP activities in these compartments during the first cycle.

**Missing functionality**: apply more generic coding approach to prepare and create the panel memberships.
Currently this is customized to cater for a specific revisit approach in the `mne2024_pan_effectmon_flanders` module.

##### Membership of generic panels

Defining generic panel sets: each set corresponds to a different combination of compartment and number of panels.
This will be the basis to make different panel memberships.
The object `compartment_generic_panels` contains the generic panel set codes and defines the associated generic panels.

```{r paged.print = FALSE}
compartment_generic_panels <-
  scheme_domain_fag_panel_count %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)")) %>%
  distinct(compartment, panel_count) %>%
  arrange(compartment, panel_count) %>%
  uncount(panel_count, .remove = FALSE) %>%
  mutate(
    generic_panel_set = str_c(
      compartment,
      "_",
      panel_count[1],
      "PANELS"
    ),
    generic_panel = str_c(
      generic_panel_set,
      "_",
      str_pad(seq_len(panel_count[1]), 2, pad = "0")
    ),
    .by = c(compartment, panel_count)
  ) %>%
  mutate(across(where(is.character), fct))

compartment_generic_panels %>%
  distinct(generic_panel_set)
```

Defining the relation between panels of different generic panel sets, for compartments with more than one generic panel set.
This relation is about which panels of one set together constitute a panel of a different set.

```{r}
generic_panel_relations <-
  compartment_generic_panels %>%
  select(-panel_count) %>%
  nest(panels = generic_panel) %>%
  filter(n() > 1, .by = compartment) %>%
  pivot_wider(names_from = generic_panel_set, values_from = panels) %>%
  # XXXX generalize this code!!
  mutate(
    GW_2PANELS = map(GW_2PANELS, function(df) {
      tibble(generic_panel_2 = rep(df$generic_panel, 12))
    }),
    GW_12PANELS = map(GW_12PANELS, function(df) {
      # this code duplicates panels taking into account the serially alternating
      # pattern within a year, in order to properly match chronology of
      # GW_24PANELS. Hence each next line represents the next 3 months.
      df %>%
        mutate(yr = rep(seq_len(n() / 2), each = 2)) %>%
        bind_rows(., .) %>%
        arrange(yr) %>%
        select(-yr) %>%
        rename(generic_panel_12 = generic_panel)
    }),
    GW_24PANELS = map(GW_24PANELS, ~ rename(., generic_panel_24 = generic_panel))
  ) %>%
  unnest(starts_with("GW"))
```

Simplifying `scheme_domain_stratum_spsamples` by collapsing schemes per compartment, i.e. integrating their spatial samples.

```{r}
compartment_domain_stratum_spsamples <-
  scheme_domain_stratum_spsamples %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)") %>% factor()) %>%
  distinct(compartment, domain, stratum, grts_address) %>%
  arrange(compartment, domain, stratum, grts_address) %>%
  nest(spsample = grts_address)
```

Assigning spatial sampling units to generic panels, in several steps depending on the 'subdivision' relation between generic panel sets.

```{r}
set.seed(20240703)

simplest_panelsplit <-
  compartment_generic_panels %>%
  select(-panel_count) %>%
  nest(panels = generic_panel) %>%
  slice(1, .by = compartment)

simplest_spsample_membership <-
  compartment_domain_stratum_spsamples %>%
  inner_join(simplest_panelsplit, join_by(compartment)) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

additional_spsample_membership_1 <-
  simplest_spsample_membership %>%
  filter(compartment == "GW") %>%
  select(-panels, -spsample) %>%
  mutate(
    generic_panel_set = "GW_12PANELS" %>%
      factor(levels = levels(compartment_generic_panels$generic_panel_set))
  ) %>%
  unnest(spsample_panels) %>%
  nest(spsample = grts_address) %>%
  rename(generic_panel_2 = panel) %>%
  inner_join(
    generic_panel_relations %>%
      distinct(compartment, generic_panel_2, generic_panel = generic_panel_12) %>%
      nest(panels = generic_panel),
    join_by(compartment, generic_panel_2)
  ) %>%
  select(-generic_panel_2) %>%
  relocate(spsample, .after = stratum) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

additional_spsample_membership_2 <-
  additional_spsample_membership_1 %>%
  select(-panels, -spsample) %>%
  mutate(
    generic_panel_set = "GW_24PANELS" %>%
      factor(levels = levels(compartment_generic_panels$generic_panel_set))
  ) %>%
  unnest(spsample_panels) %>%
  nest(spsample = grts_address) %>%
  rename(generic_panel_12 = panel) %>%
  inner_join(
    generic_panel_relations %>%
      distinct(compartment, generic_panel_12, generic_panel = generic_panel_24) %>%
      nest(panels = generic_panel),
    join_by(compartment, generic_panel_12)
  ) %>%
  select(-generic_panel_12) %>%
  relocate(spsample, .after = stratum) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

generic_spsample_membership <-
  bind_rows(
    simplest_spsample_membership,
    additional_spsample_membership_1,
    additional_spsample_membership_2
  ) %>%
  select(-spsample, -panels) %>%
  unnest(spsample_panels) %>%
  rename(generic_panel = panel)
```

Defining spatial sampling unit groups based on `generic_panel_relations`, i.e. where different periodicities arise, to further pinpoint spatial relations between generic panels of different generic panel sets.

```{r}
generic_panel_spsunit_groups <-
  generic_panel_relations %>%
  pivot_longer(
    cols = c(generic_panel_2, generic_panel_12),
    names_to = "generic_panel_set",
    values_to = "generic_panel"
  ) %>%
  select(-generic_panel_set) %>%
  relocate(spsunit_group = generic_panel_24, .after = last_col()) %>%
  summarize(
    spsunit_groups = str_c(spsunit_group, collapse = ", "),
    .by = c(compartment, generic_panel)
  )
```

Adding the spatial sampling unit groups to `generic_spsample_membership` and extending them.

```{r}
generic_spsample_membership <-
  generic_spsample_membership %>%
  left_join(
    generic_panel_spsunit_groups,
    join_by(compartment, generic_panel),
    unmatched = "error",
    relationship = "many-to-one"
  ) %>%
  mutate(
    spsunit_groups = ifelse(
      is.na(spsunit_groups),
      as.character(generic_panel),
      spsunit_groups
    ) %>%
      factor()
  )
```

Inspecting number of spatial sampling units per generic panel.

```{r}
generic_spsample_membership %>%
  count(compartment, domain, generic_panel, spsunit_groups)
```

##### Panel membership

Deriving panel membership from generic panel membership.
A panel is defined for a scheme x field activity group (see `scheme_domain_fag_panels`).

```{r}
scheme_domain_fag_panel_stratum_spsamples <-
  scheme_domain_fag_panel_count %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)") %>% factor()) %>%
  select(compartment, scheme, domain, field_activity_group, panel_count) %>%
  inner_join(
    scheme_domain_fag_panels %>%
      select(-c(year_first:panel_group_repetitions)) %>%
      nest(panels = panel),
    join_by(scheme, domain, field_activity_group)
  ) %>%
  mutate(
    generic_panel_set = str_c(
      compartment,
      "_",
      panel_count,
      "PANELS"
    )
  ) %>%
  inner_join(
    compartment_generic_panels %>%
      select(-panel_count) %>%
      nest(generic_panels = generic_panel),
    join_by(compartment, generic_panel_set)
  ) %>%
  mutate(
    panels = map2(panels, generic_panels, ~ bind_cols(.x, .y))
  ) %>%
  select(-generic_panels) %>%
  unnest(panels) %>%
  # this step is scheme-agnostic:
  inner_join(
    generic_spsample_membership,
    join_by(compartment, domain, generic_panel_set, generic_panel),
    relationship = "many-to-many"
  ) %>%
  # aligning with scheme samples:
  semi_join(
    scheme_domain_stratum_spsamples,
    join_by(scheme, domain, stratum, grts_address)
  ) %>%
  select(
    scheme,
    domain,
    field_activity_group,
    panel,
    stratum,
    grts_address,
    spsunit_groups
  )
```

### Spatiotemporal samples

Joining the date intervals per panel with panel memberships to generate the spatiotemporal samples.

```{r}
scheme_domain_fag_stratum_spsamples_calendar <-
  scheme_domain_fag_panel_stratum_spsamples %>%
  inner_join(
    scheme_domain_fag_panel_calendar,
    join_by(scheme, domain, field_activity_group, panel),
    relationship = "many-to-many"
  )
```

### Field activities in space and time

Some field activity groups may overlap between schemes, i.e. when the same sampling unit is present in more than one scheme _and_ the same field activity group happens in these schemes at the same time for this sampling unit.

Further, the uncertainties in the sampling frame cause duplication of GRTS addresses in the (generated) sample because multiple strata can be connected to the same GRTS address.

While in below object `fag_grts_calendar` both sources of duplication have been collapsed, we also create an object `fag_stratum_grts_calendar` that better reflects the expected _number_ of field activities in space and time: the reduction by GRTS address will not happen in reality since the sample sizes must be maintained!
This means that new sampling units will be added from a list of spare sampling units, when a stratum on terrain is not the expected stratum, in order to maintain the sample size.
This specifically applies to the (many) types where the cell center in the field must be the type, so that only one 'cell_center' type can correspond to a GRTS address, contrary to the theoretical information from the sampling frame.

```{r}
fag_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address
  ) %>%
  arrange(pick(everything()))
```

```{r}
fag_stratum_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address,
    stratum
  ) %>%
  arrange(pick(everything()))
```

## Deriving datasets and layers for use and visualization

```{r}
scheme_domain_stratum_spsamples_sf <-
  scheme_domain_stratum_spsamples %>%
  add_point_coords_grts()
```

```{r}
samplinglocations_sf <-
  scheme_domain_stratum_spsamples %>%
  distinct(grts_address) %>%
  add_point_coords_grts()
```

## Optimizing resource allocation in time and space (within or between years)

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "scheme_domain_stratum_spsamples",
  "scheme_domain_stratum_spsamples_sf",
  "samplinglocations_sf",
  "scheme_fag_fa",
  "scheme_domain_fag_panel_count",
  "scheme_domain_fag_panels",
  "scheme_domain_fag_panel_calendar",
  "scheme_domain_fag_panel_stratum_spsamples",
  "scheme_domain_fag_stratum_spsamples_calendar",
  "fag_grts_calendar",
  "fag_stratum_grts_calendar"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
