# Generating spatiotemporal samples {#sampling}

<!-- use hardcoded tibble from file 010 with design attributes per scheme -->

<!-- Take into account the revisit design, sample size, surplus sampling, domains -->

<!-- Sample sizes can be summed over active module-domains that refer to the same domain. -->
<!-- Where one module-domain overlaps (contains) another module-domain, those sample sizes cannot be added but an appropriate offset must be taken into account when sampling in the shared area (the smaller domain), in order to respect the GRTS ordering. -->
<!-- This also holds for MHQ Flanders (not in this POC) vs MHQ in SACs. -->

For this proof of concept, we restrict the allowed modules.

<!-- XXX Reconsider this: is module subselection actually needed? -->

```{r}
module_selection <- if (all(
  c("pan_phase_1_sac_subset_5", "pan_phase_1_flanders") %in%
    params$active_modules
)) {
  c("pan_phase_1_sac_subset_5", "pan_phase_1_flanders")
} else {
  params$active_modules
}
module_domain_scheme_stratum_sample_size_2 <-
  module_domain_scheme_stratum_sample_size %>%
  filter(module %in% module_selection)
```

## Defining the samples

For now still without:

- non-core schemes (except for its types that are not covered by core schemes)
- GRTS address offset for MHQ at Flemish level = undefined module
<!-- - module labelling of sampling units: does this make sense? -->
- adding (and labelling) spare sampling units
- collapsing nearby units for groundwater schemes
- gradual increase, during the first cycle, of the number of installed groundwater and surfacewater observation wells.
So it's currently assumed that all observation wells are in place already.

### Spatial samples

Determining the applicable domain relations.

```{r}
domain_relations <- 
  domain_overlaps %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size_2 %>% 
      distinct(module, domain),
    join_by(domain)
  ) %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size_2 %>% 
      distinct(module, domain),
    join_by(included_by == domain)
  )
```

Drawing spatial samples.

```{r}
scheme_domain_stratum_spsamples <-
  module_domain_scheme_stratum_sample_size_2 %>%
  select(module, domain, scheme, sampling_frame_id, stratum, sp_sample_size_all_panels_stratum) %>%
  # for MNE, non-core schemes (dropped below) have yet to be taken back (but I
  # think this step also drops things we always need to drop):
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>%
  # domains that are the same between different modules are integrated by
  # summing the sample sizes; the module is dropped:
  summarize(
    sp_sample_size_all_panels_stratum = sum(sp_sample_size_all_panels_stratum),
    .by = c(domain, scheme, sampling_frame_id, stratum)
  ) %>%
  nest(domain_stratum_spsamplesize = c(domain, stratum, sp_sample_size_all_panels_stratum)) %>%
  inner_join(sp_samplingframe_domain, by = join_by(sampling_frame_id)) %>%
  # drawing an independent sample per scheme, domain and stratum:
  mutate(
    spsample = map2(domain_stratum_spsamplesize, sampling_frame_domain, function(dss, sfd) {
      sfd %>%
        nest(grts_addresses = grts_address) %>%
        inner_join(dss, ., by = join_by(domain, stratum)) %>%
        mutate(
          sample =
            map2(sp_sample_size_all_panels_stratum, grts_addresses, function(sss, grts) {
              grts %>% slice_min(grts_address, n = sss)
            })
        ) %>% 
        nest(domain_spsample = c(
          domain, 
          sp_sample_size_all_panels_stratum, 
          grts_addresses,
          sample
        ))
    })
  ) %>%
  select(scheme, spsample) %>%
  unnest(spsample) %>% 
  # handling the overlap between domains: above, the samples have been drawn
  # independently so now this has to be corrected!
  mutate(
    domain_spsample = map(domain_spsample, function(ds) {
      ds <- left_join(
        ds,
        domain_relations,
        join_by(domain),
        relationship = "one-to-one",
        unmatched = "drop"
      )
      # vector of including domains:
      including_domains <- unique(ds$included_by)
      # drop NAs:
      including_domains <- including_domains[!is.na(including_domains)]
      # only keep an including domain if it is effectively present as domain:
      including_domains <- including_domains[including_domains %in% ds$domain]
      if (length(including_domains) > 1) {
        stop("Multiple including domains in one stratum are not supported.")
      }
      ds %>% 
        mutate(
          including = domain %in% including_domains | length(including_domains) == 0,
          # unsampled units of the including domain:
          grts_available = ifelse(
            !including,
            tibble(),
            map2(grts_addresses, sample, function(ga, spl) {
              anti_join(ga, spl, join_by(grts_address))
            })
          ),
          # put the unsampled units next to the included domains:
          grts_restriction = ifelse(
            including, 
            tibble(),
            grts_available[including]
          ),
          # determine the available unsampled units per domain:
          grts_addresses_2 = map2(grts_addresses, grts_restriction, function(ga, gr) {
            if (is.null(gr)) {
              return(NULL)
            }
            semi_join(ga, gr, join_by(grts_address))
          }),
          # redraw samples for included domains:
          sample_2 = pmap(
            list(including, sample, grts_addresses_2, sp_sample_size_all_panels_stratum),
            function(inc, sam, grts, sss){
              if (inc) {
                sam
              } else {
                grts %>% slice_min(grts_address, n = sss)
              }
            }
          ),
          # flag sampling units of included domains, as they should never be
          # considered apart from the including domain:
          extends_on_larger_domain = !including & length(including_domains) == 1
        ) %>% 
        select(domain, sample = sample_2, extends_on_larger_domain) %>% 
        # next step will also drop domains where no sampling units were
        # available (i.e. the whole population was sampled by the larger domain):
        unnest(sample)
    })
  ) %>% 
  unnest(domain_spsample)
```

Check that no sampling units overlap between domains within the combination of scheme and stratum:

```{r}
scheme_domain_stratum_spsamples %>% 
  count(scheme, stratum, grts_address) %>% 
  filter(n > 1) %>% 
  nrow() == 0
```


### Panel design

Determining the number of panels for each scheme x 'field activity group' (FAG).
We also add period, group size and group repetition attributes.
Grouping refers to the subgroup structure in nested serially alternated patterns.

A 'field activity group' is a set of field activities with the same panel count, period, group size and group repetition attributes.

**Missing functionality**: make the code below work for other field activities than those in module `pan_phase_1_flanders` (e.g. wrt prefix extraction).

```{r}
scheme_domain_fa_panel_count <-
  module_domain_scheme_stratum_sample_size_2 %>%
  # to be revisited filter step (see above); this is especially important since
  # field activities are done e.g. for GW_05.1_aq (SURFLEVINST, SURFLEVREAD)
  # that will also be dropped later on, as a consequence of this.
  # So even though for a specific module the non-core schemes don't have their
  # own sample size, they may have their own field activities that must be
  # taken into account
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>%
  distinct(scheme, domain, cycle_duration_y, year_first, year_last) %>%
  inner_join(scheme_field_activity, join_by(scheme)) %>%
  inner_join(field_activity_paneldesign, join_by(field_activity)) %>%
  mutate(
    cycle_duration_y = ifelse(
      !is.na(enforce_cycle_duration_y),
      enforce_cycle_duration_y,
      cycle_duration_y
    ),
    panel_count = case_when(
      revisit_pattern == "pure_panel" ~ 1L,
      .default = as.integer(cycle_duration_y * years(1) / panel_period_main)
    ) *
      case_when(
        is.na(revisit_subpattern) ~ 1L,
        revisit_subpattern == "pure_panel" ~ 1L,
        !is.na(limit_n_subpanels_per_panel) ~ limit_n_subpanels_per_panel,
        .default = as.integer(panel_period_main / panel_subperiod)
      ),
    panel_period = case_when(
      is.na(panel_subperiod) ~ panel_period_main,
      .default = panel_subperiod
    ),
    panel_group_size = ifelse(
      !is.na(limit_n_subpanels_per_panel),
      limit_n_subpanels_per_panel,
      1L
    ),
    panel_group_repetitions = case_when(
      !is.na(limit_n_subpanels_per_panel) ~ as.integer(
        panel_period_main / panel_subperiod
      ) / limit_n_subpanels_per_panel,
      revisit_subpattern == "pure_panel" ~ as.integer(
        panel_period_main / panel_subperiod
      ),
      .default = 1L
    )
  ) %>%
  select(
    scheme,
    domain,
    field_activity,
    starts_with("year"),
    panel_period,
    panel_count,
    panel_group_size,
    panel_group_repetitions
  )

# relation between FA (field activity) and FAG (FA group); done scheme-specific
# so that joins still work one-to-one (because SURFSECC appears in two FAGs):
scheme_fag_fa <-
  scheme_domain_fa_panel_count %>%
  mutate(field_activity_prefix = str_extract(field_activity, "^(GW|SURF|SOIL)")) %>%
  mutate(
    field_activity_group = str_remove(field_activity, "^(GW|SURF|SOIL)") %>%
      str_flatten(collapse = "&") %>%
      str_c(first(field_activity_prefix), .),
    .by = -field_activity
  ) %>%
  select(-field_activity_prefix) %>%
  arrange(field_activity) %>%
  mutate(field_activity_group = fct(field_activity_group)) %>%
  arrange(scheme, field_activity) %>%
  distinct(scheme, field_activity_group, field_activity)

scheme_domain_fag_panel_count <-
  scheme_domain_fa_panel_count %>%
  inner_join(
    scheme_fag_fa,
    join_by(scheme, field_activity),
    unmatched = "error",
    relationship = "one-to-one"
  ) %>%
  distinct(pick(-field_activity)) %>%
  relocate(field_activity_group, .after = domain)

scheme_domain_fag_panel_count %>%
  kable()
```

Different occurring patterns at the compartment level, for information and thought:

```{r}
scheme_domain_fag_panel_count %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)")) %>%
  summarize(
    field_activity_groups = str_c(unique(field_activity_group), collapse = ", "),
    .by = c(compartment, starts_with("panel"))
  ) %>%
  kable()
```

Defining panels.

```{r}
scheme_domain_fag_panels <-
  scheme_domain_fag_panel_count %>%
  uncount(panel_count) %>%
  mutate(
    panel = str_c(
      scheme,
      field_activity_group,
      str_c("PAN", str_pad(seq_len(n()), 2, pad = "0")),
      sep = "_"
    ),
    .by = c(scheme, field_activity_group)
  ) %>%
  mutate(panel = factor(panel))
```

Defining date intervals and assigning panels to these intervals.

```{r}
scheme_domain_fag_panel_calendar <-
  scheme_domain_fag_panels %>%
  nest(
    revisit = c(starts_with("year"), starts_with("panel"), -panel),
    panels = panel
  ) %>%
  mutate(
    revisit = map(revisit, function(df) {
      df %>%
        distinct() %>%
        mutate(
          # taking into account month_first_default:
          nr_periods =
            (years(year_last - year_first + 1) -
              months(13 - month_first_default)) / panel_period
        ) %>%
        uncount(nr_periods) %>%
        mutate(
          panel_periodmonths = as.period(panel_period, "months") %>%
            month() %>%
            as.integer(),
          panel_cum_months = cumsum(panel_periodmonths),
          panel_cum_months_lag = lag(panel_cum_months, default = 0L),
          # we only start in month_first_default of first year:
          date_start = make_date(year_first, month_first_default, 1) + months(panel_cum_months_lag),
          date_end = make_date(year_first, month_first_default, 1) + months(panel_cum_months) - days(1),
          date_interval = interval(date_start, date_end)
        ) %>%
        select(-c(panel_periodmonths, starts_with("panel_cum_months")))
    }),
    panel_date_assignment = map2(revisit, panels, function(revis, pan) {
      group_size <- revis$panel_group_size[1]
      group_repetitions <- revis$panel_group_repetitions[1]
      if (group_size == 1 && group_repetitions == 1) {
        pan2 <- pan
      } else {
        indexes_pan <- seq_len(nrow(pan))
        indexgrouping_pan <- rep(seq_len(nrow(pan) / group_size), each = group_size)
        indexes_grouped <- split(indexes_pan, indexgrouping_pan) %>%
          rep(each = group_repetitions) %>%
          unlist() %>%
          unname()
        pan2 <- pan[indexes_grouped, ]
      }
      indexes <- rep(seq_len(nrow(pan2)), nrow(revis) %/% nrow(pan2) + 1) %>%
        head(nrow(revis))
      bind_cols(
        revis %>% select(date_start, date_end, date_interval),
        pan2[indexes, ]
      )
    })
  ) %>%
  select(-panels, -revisit) %>%
  unnest(panel_date_assignment) %>%
  relocate(panel, .after = field_activity_group)
```

### Panel membership of spatial sampling units

Panel membership is randomized by following the GRTS address sequence at the stratum level.

To determine the number of sampling units per panel, it must be taken into account that the stratum sample size is not equal to, or a multiple of, the number of panels, so panels usually cannot have an equal number of units.
Local pivotal sampling of panels is applied to get an even load of all panels while maintaining maximum temporal spread in the membership.
One wants to avoid a systematic effect that the first panels always have more units than those further in line, or the last panel.
This becomes even more important for strata with few sampling units; sometimes this number is smaller than the theoretical number of panels!

The interrelation between panels within a compartment is controlled:

- To get maximum synergy among schemes within a compartment, we consider the unioned sample per compartment, split that one to panels, after which the splits can again be divided over the different field activity groups.
- To get maximum synergy among different field activity groups within a compartment, different panel counts are handled consecutively: e.g. if both '2 panels' and '12 panels' occur in a compartment, then membership design is first done for 2 panels, and the members of each panel are then split into 6 groups to define the membership of the 12 panels.
This avoids repeating local pivotal sampling more than necessary, which would lead to unneeded mismatches.

**Missing functionality**: maximizing panel synergy between compartments (note that this would also need taking into account different start years between compartments), or conversely, avoid some synergies even within a compartment, to avoid the problem described now.
The challenge is the different scheme specific sample sizes (this may already hold for the current approach of integration within a compartment, but this was not a problem for the first module).
This would lead to the schemes with smaller sample sizes not being equally spread over the panels.
This might be an optimal fieldwork approach, but it is not statistically optimal.
So from a statistical viewpoint it will be safer to use the scheme level and only optimize between field activity groups within a scheme.
There's no such clash between schemes when their field activities don't overlap, but when the same field activity occurs this problem will be present for strata that are represented in more than one scheme, if they were assigned their own sample size.
This situation doesn't occur in the specific case of the `pan_phase_1_flanders` module.

**Missing functionality**: expanding to all compartments + MHQ.

**Missing functionality**: defining and taking into account temporal dependencies between field activity groups.
An example is that a piezometer must be installed before a water sample can be taken from it.

**Missing functionality**: provide a gradual increase, during the first cycle, of the number of installed groundwater and surfacewater observation wells.
This has consequences for the READ and the SAMP activities in these compartments during the first cycle.

**Missing functionality**: apply more generic coding approach to prepare and create the panel memberships.
Currently this is customized to cater for a specific revisit approach in the `pan_phase_1_flanders` module.

#### Membership of generic panels

Defining generic panel sets: each set corresponds to a different combination of compartment and number of panels.
This will be the basis to make different panel memberships.
The object `compartment_generic_panels` contains the generic panel set codes and defines the associated generic panels.

```{r paged.print = FALSE}
compartment_generic_panels <-
  scheme_domain_fag_panel_count %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)")) %>%
  distinct(compartment, panel_count) %>%
  arrange(compartment, panel_count) %>%
  uncount(panel_count, .remove = FALSE) %>%
  mutate(
    generic_panel_set = str_c(
      compartment,
      "_",
      panel_count[1],
      "PANELS"
    ),
    generic_panel = str_c(
      generic_panel_set,
      "_",
      str_pad(seq_len(panel_count[1]), 2, pad = "0")
    ),
    .by = c(compartment, panel_count)
  ) %>%
  mutate(across(where(is.character), fct))

compartment_generic_panels %>%
  distinct(generic_panel_set)
```

Defining the relation between panels of different generic panel sets, for compartments with more than one generic panel set.
This relation is about which panels of one set together constitute a panel of a different set.

```{r}
generic_panel_relations <-
  compartment_generic_panels %>%
  select(-panel_count) %>%
  nest(panels = generic_panel) %>%
  filter(n() > 1, .by = compartment) %>%
  pivot_wider(names_from = generic_panel_set, values_from = panels) %>%
  # XXXX generalize this code!!
  mutate(
    GW_2PANELS = map(GW_2PANELS, function(df) {
      tibble(generic_panel_2 = rep(df$generic_panel, 12))
    }),
    GW_12PANELS = map(GW_12PANELS, function(df) {
      # this code duplicates panels taking into account the serially alternating
      # pattern within a year, in order to properly match chronology of
      # GW_24PANELS. Hence each next line represents the next 3 months.
      df %>%
        mutate(yr = rep(seq_len(n() / 2), each = 2)) %>%
        bind_rows(., .) %>%
        arrange(yr) %>%
        select(-yr) %>%
        rename(generic_panel_12 = generic_panel)
    }),
    GW_24PANELS = map(GW_24PANELS, ~ rename(., generic_panel_24 = generic_panel))
  ) %>%
  unnest(starts_with("GW"))
```

Simplifying `scheme_domain_stratum_spsamples` by collapsing schemes per compartment, i.e. integrating their spatial samples.

```{r}
compartment_domain_stratum_spsamples <-
  scheme_domain_stratum_spsamples %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)") %>% factor()) %>%
  distinct(compartment, domain, stratum, grts_address) %>%
  arrange(compartment, domain, stratum, grts_address) %>%
  nest(spsample = grts_address)
```

Assigning spatial sampling units to generic panels, in several steps depending on the 'subdivision' relation between generic panel sets.

```{r}
set.seed(20240703)

simplest_panelsplit <-
  compartment_generic_panels %>%
  select(-panel_count) %>%
  nest(panels = generic_panel) %>%
  slice(1, .by = compartment)

simplest_spsample_membership <-
  compartment_domain_stratum_spsamples %>%
  inner_join(simplest_panelsplit, join_by(compartment)) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

additional_spsample_membership_1 <-
  simplest_spsample_membership %>%
  filter(compartment == "GW") %>%
  select(-panels, -spsample) %>%
  mutate(
    generic_panel_set = "GW_12PANELS" %>%
      factor(levels = levels(compartment_generic_panels$generic_panel_set))
  ) %>%
  unnest(spsample_panels) %>%
  nest(spsample = grts_address) %>%
  rename(generic_panel_2 = panel) %>%
  inner_join(
    generic_panel_relations %>%
      distinct(compartment, generic_panel_2, generic_panel = generic_panel_12) %>%
      nest(panels = generic_panel),
    join_by(compartment, generic_panel_2)
  ) %>%
  select(-generic_panel_2) %>%
  relocate(spsample, .after = stratum) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

additional_spsample_membership_2 <-
  additional_spsample_membership_1 %>%
  select(-panels, -spsample) %>%
  mutate(
    generic_panel_set = "GW_24PANELS" %>%
      factor(levels = levels(compartment_generic_panels$generic_panel_set))
  ) %>%
  unnest(spsample_panels) %>%
  nest(spsample = grts_address) %>%
  rename(generic_panel_12 = panel) %>%
  inner_join(
    generic_panel_relations %>%
      distinct(compartment, generic_panel_12, generic_panel = generic_panel_24) %>%
      nest(panels = generic_panel),
    join_by(compartment, generic_panel_12)
  ) %>%
  select(-generic_panel_12) %>%
  relocate(spsample, .after = stratum) %>%
  mutate(
    spsample_panels = map2(spsample, panels, distribute_sample_over_panels)
  )

generic_spsample_membership <-
  bind_rows(
    simplest_spsample_membership,
    additional_spsample_membership_1,
    additional_spsample_membership_2
  ) %>%
  select(-spsample, -panels) %>%
  unnest(spsample_panels) %>%
  rename(generic_panel = panel)
```

Defining spatial sampling unit groups based on `generic_panel_relations`, i.e. where different periodicities arise, to further pinpoint spatial relations between generic panels of different generic panel sets.

```{r}
generic_panel_spsunit_groups <-
  generic_panel_relations %>%
  pivot_longer(
    cols = c(generic_panel_2, generic_panel_12),
    names_to = "generic_panel_set",
    values_to = "generic_panel"
  ) %>%
  select(-generic_panel_set) %>%
  relocate(spsunit_group = generic_panel_24, .after = last_col()) %>%
  summarize(
    spsunit_groups = str_c(spsunit_group, collapse = ", "),
    .by = c(compartment, generic_panel)
  )
```

Adding the spatial sampling unit groups to `generic_spsample_membership` and extending them.

```{r}
generic_spsample_membership <-
  generic_spsample_membership %>%
  left_join(
    generic_panel_spsunit_groups,
    join_by(compartment, generic_panel),
    unmatched = "error",
    relationship = "many-to-one"
  ) %>%
  mutate(
    spsunit_groups = ifelse(
      is.na(spsunit_groups),
      as.character(generic_panel),
      spsunit_groups
    ) %>%
      factor()
  )
```

Inspecting number of spatial sampling units per generic panel.

```{r}
generic_spsample_membership %>%
  count(compartment, domain, generic_panel, spsunit_groups)
```

#### Panel membership

Deriving panel membership from generic panel membership.
A panel is defined for a scheme x field activity group (see `scheme_domain_fag_panels`).

```{r}
scheme_domain_fag_panel_stratum_spsamples <-
  scheme_domain_fag_panel_count %>%
  mutate(compartment = str_extract(scheme, "^(GW|SURF|SOIL)") %>% factor()) %>%
  select(compartment, scheme, domain, field_activity_group, panel_count) %>%
  inner_join(
    scheme_domain_fag_panels %>%
      select(-c(year_first:panel_group_repetitions)) %>%
      nest(panels = panel),
    join_by(scheme, domain, field_activity_group)
  ) %>%
  mutate(
    generic_panel_set = str_c(
      compartment,
      "_",
      panel_count,
      "PANELS"
    )
  ) %>%
  inner_join(
    compartment_generic_panels %>%
      select(-panel_count) %>%
      nest(generic_panels = generic_panel),
    join_by(compartment, generic_panel_set)
  ) %>%
  mutate(
    panels = map2(panels, generic_panels, ~ bind_cols(.x, .y))
  ) %>%
  select(-generic_panels) %>%
  unnest(panels) %>%
  # this step is scheme-agnostic:
  inner_join(
    generic_spsample_membership,
    join_by(compartment, domain, generic_panel_set, generic_panel),
    relationship = "many-to-many"
  ) %>%
  # aligning with scheme samples:
  semi_join(
    scheme_domain_stratum_spsamples,
    join_by(scheme, domain, stratum, grts_address)
  ) %>%
  select(
    scheme,
    domain,
    field_activity_group,
    panel,
    stratum,
    grts_address,
    spsunit_groups
  )
```

### Spatiotemporal samples

Joining the date intervals per panel with panel memberships to generate the spatiotemporal samples.

```{r}
scheme_domain_fag_stratum_spsamples_calendar <-
  scheme_domain_fag_panel_stratum_spsamples %>%
  inner_join(
    scheme_domain_fag_panel_calendar,
    join_by(scheme, domain, field_activity_group, panel),
    relationship = "many-to-many"
  )
```

### Field activities in space and time

Some field activity groups may overlap between schemes, i.e. when the same sampling unit is present in more than one scheme _and_ the same field activity group happens in these schemes at the same time for this sampling unit.

Further, the uncertainties in the sampling frame cause duplication of GRTS addresses in the (generated) sample because multiple strata can be connected to the same GRTS address.

While in below object `fag_grts_calendar` both sources of duplication have been collapsed, we also create an object `fag_stratum_grts_calendar` that better reflects the expected _number_ of field activities in space and time: the reduction by GRTS address will not happen in reality since the sample sizes must be maintained!
This means that new sampling units will be added from a list of spare sampling units, when a stratum on terrain is not the expected stratum, in order to maintain the sample size.
This specifically applies to the (many) types where the cell center in the field must be the type, so that only one 'cell_center' type can correspond to a GRTS address, contrary to the theoretical information from the sampling frame.

```{r}
fag_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address
  ) %>%
  arrange(pick(everything()))
```

```{r}
fag_stratum_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address,
    stratum
  ) %>%
  arrange(pick(everything()))
```

## Deriving datasets and layers for use and visualization

```{r}
scheme_domain_stratum_spsamples_sf <-
  scheme_domain_stratum_spsamples %>%
  add_point_coords_grts()
```

```{r}
samplinglocations_sf <-
  scheme_domain_stratum_spsamples %>%
  distinct(grts_address) %>%
  add_point_coords_grts()
```

## Optimizing resource allocation in time and space (within or between years)

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "scheme_domain_stratum_spsamples",
  "scheme_domain_stratum_spsamples_sf",
  "samplinglocations_sf",
  "scheme_fag_fa",
  "scheme_domain_fag_panel_count",
  "scheme_domain_fag_panels",
  "scheme_domain_fag_panel_calendar",
  "scheme_domain_fag_panel_stratum_spsamples",
  "scheme_domain_fag_stratum_spsamples_calendar",
  "fag_grts_calendar",
  "fag_stratum_grts_calendar"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
