# Generating spatiotemporal samples {#sampling}

## Considerations

The samples don't explicitly address the non-core schemes, except for their types that are not covered by core schemes.
There's no intention to do that, as for overlapping types the samples of the core schemes are the ones to be used for the non-core schemes.

Module labelling of sampling units: this is needed to apply different revisit designs to different modules, if applicable.
This should be tested first, because if module panel designs align, it's better for the membership design if all units are taken together.
Combining different modules with different revisit designs is like creating a split panel design: in the same period, you need two different sets of sampling units that receive another revisit design.
This is done to make these separate revisit designs form one strict GRTS series with their own spare units.
In this way, replacements don't cause spillovers of units from one revisit design to another.
The conflict is at what levels you want to guarantee a single strict GRTS series: within the module, within the year (conflicts with previous, supposing that each module should cover the full cycle) ...
Since it's most reasonable to spread each module over the full cycle, this means that GRTS gaps will be inevitable either within a panel period (each module is a GRTS series split into periods, so for two or more modules this combines different GRTS subsets in the same period) or within a module (each panel is a GRTS series assigned to consecutive modules), notably between the different modules.

- Combining modules with the same revisit design: panels can be created for all units, afterwards each panel can be divided over both modules for module labelling of units (if needed at all).
This approach optimizes sampling design & inference for the module combination.
  - Note that the panel _construction_ must be done for each domain separately, i.e. the overlapped domains and the non-overlapped (remainder) domain.
  Otherwise you would have that the higher GRTS addresses added in the included domains (~ other module) end up in the latest (common) panels, hence the units of an included domain would not be well spread over the panels when specifically looking at the domain.
  - The same holds for the generation of spare units: while module labelling is not needed within the same revisit pattern, a unit in a domain that is active because one of the applicable modules is, should be replaced _within the domain_ in order to maintain the sample size in the domain.
  Consequently the same holds for spare units in the non-overlapped (remainder) domain.
  We could also have applied this at the module x domain level, but then one needs to assign individual units to modules (and doing it in some randomized way, since we need interspersion of modules in time), even while the same revisit design applies, and it would lower the flexibility to maintain spatial balance when modules are stopped or added.
  - Clearly, also the division of panels between modules only applies to the domains where both modules overlap.
- Combining modules with different revisit design: sampling units must be assigned to modules a priori, each module has its own GRTS series + spare addresses (hence mandatory gap needed).
The challenge will be to maintain sufficient gap for spare units while also maintaining a fixed start point for a second module (or panel split) for the future that doesn't necessarily compound the rise of new modules.
Then panel designs are created for each module separately.
The same approach holds for split panel designs.

For now still without:

- GRTS address offset for MHQ at Flemish level = undefined module
- collapsing nearby units for groundwater schemes
- the fact that the _installation_ of observation wells must be accelerated in the module's _first_ cycle if the latter is shorter than the enforced cycle duration.

We will not take into account the fact that different modules can have _different end years_ (and maybe start years), so that their combination in a scheme and panel split would invoke the need for sample sizes and panel memberships on a yearly basis.
However, large shifts in the membership designs when a module in a module combo ends (invoking location redistribution over panels in order to restore an even spread) would effectively disrupt the revisit design of those locations.
Since the actual end years of modules are highly uncertain, we simply ignore this and assume that the maximum end year is most useful at the module combo level.
We do the same for the start years, using the minimum year per scheme x module combo.
If this situation is present, then consequences must be considered with care and some restrictions may need to be applied post-hoc to the spatiotemporal samples.
See also the _missing functionality_ in section \@ref(moco).

## Module combinations (module combos) per scheme based on common panel design of target variables {#moco}

Determining module combos per scheme based on common panel design characteristics _of target variables_.
The target variables limit the main field methods of a module x scheme to a great extent, since the latter also comprise variables for explanation & quality control.

First create an object with the 'target' panel designs per module and scheme (target referring to the target variables restriction).

```{r}
mod_scheme_targetfieldmethod <- 
  mod_scheme_mainfieldmethod %>% 
  left_join(
    envvars %>% 
      filter(target) %>% 
      distinct(module, scheme, main_datacollection_method) %>% 
      mutate(target_envvar = TRUE),
    join_by(module, scheme, main_field_method == main_datacollection_method),
    relationship = "one-to-one",
    unmatched = "drop"
  ) %>% 
  # the main field method of MHQ schemes also corresponds to a target variable;
  # in MHQ we only defined a 'variable set' LSVI...
  filter(target_envvar | str_detect(scheme, "^HQ")) %>% 
  select(-is_core_scheme, -target_envvar)
mod_scheme_targetfieldmethod_paneldesign <- 
  mod_scheme_targetfieldmethod %>% 
  # expand with FAG (note that some main field methods link to more than one
  # FAG, which has to do with the spatial subsets, which we can ignore here
  # since in the case of the MAIN field method these don't cause differences in
  # panel designs but only in activities.)
  inner_join(
    mod_scheme_field_activity %>% 
      distinct(module, scheme, field_activity, field_activity_group),
    join_by(module, scheme, main_field_method == field_activity),
    relationship = "one-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  # add the FAG panel design
  inner_join(
    mod_scheme_fag_paneldesign %>% 
      distinct(pick(module, scheme, field_activity_group, panel_split, revisit_pattern:last_col())),
    join_by(module, scheme, field_activity_group),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  distinct(pick(
    module,
    scheme,
    main_field_method,
    panel_split,
    revisit_pattern:last_col()
  ))
if (nrow(mod_scheme_targetfieldmethod_paneldesign) != nrow(mod_scheme_targetfieldmethod)) {
  stop("The number of rows changed unexpectedly.")
}
mod_scheme_target_paneldesign <- 
  mod_scheme_targetfieldmethod_paneldesign %>% 
  distinct(pick(-main_field_method, -enforced_cycle_duration))
```

Now determining the module combos per scheme and panel split.
Do note that the module combo code can only be interpreted as a unique revisit design in combination with scheme and panel split!
So a module combo has to be seen as an aggregation based on a common revisit design _between_ modules but _within_ scheme and panel split.

```{r}
scheme_target_paneldesign <- 
  mod_scheme_target_paneldesign %>% 
  inner_join(
    modules %>% select(module, code_short),
    join_by(module),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  summarize(
    module_combo_code = str_c(code_short, collapse = "_") %>% fct(),
    module_combo = list(tibble(module = module)),
    .by = -c(module, code_short)
  ) %>% 
  arrange(module_combo_code, scheme, panel_period_main, cycle_duration_y) %>% 
  relocate(module_combo_code, module_combo, .after = scheme)
scheme_target_paneldesign
```

We end up with `r nrow(scheme_target_paneldesign)` different scheme x target panel design combinations, each of which applies to one or several modules.

Several schemes have more than one revisit design, i.e. different module combinations x panel splits:

```{r}
scheme_target_paneldesign %>% 
  count(scheme, name = "nr_revisitpatterns_per_scheme") %>% 
  count(nr_revisitpatterns_per_scheme)
```

<!-- It can be seen that module 'pan_transmon_be2100024' (pan2124) is always present when a scheme has more than one revisit design. -->
<!-- So it is logical to handle it specially. -->

```{r eval=FALSE, include=FALSE}
# Not run since this chunk is too specific and depends on which modules are active.
scheme_target_paneldesign %>% 
  mutate(nr_revisitpatterns_per_scheme = n(), .by = scheme) %>% 
  filter(nr_revisitpatterns_per_scheme > 1) %>% 
  unnest(module_combo) %>% 
  summarize(
    has_pan2124 = any(module_combo_code == "pan2124"),
    .by = scheme
  ) %>% 
  pull(has_pan2124) %>% 
  all()
```

Creating an object that holds the relation between module, scheme, module combo, panel split and first & last years.

```{r}
mod_scheme_yrs_moco_ps <- 
  module_domain_scheme_stratum_sample_size %>%
  # next filter step is in alignment with earlier (see above); field
  # activities of non-core schemes should (by definition) only apply to the
  # explicitly (and only exceptionally) added sampling units for such non-core
  # schemes.
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>%
  distinct(module, scheme, year_first, year_last) %>%
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  arrange(scheme, module_combo_code, panel_split)

```

Checking which scheme x module combo x panel split rows have conflicting first or last years across modules.

- conflicting last years:

```{r paged.print=FALSE}
mod_scheme_yrs_moco_ps_yrsmoco <- 
  mod_scheme_yrs_moco_ps %>% 
  mutate(
    year_first_moco = min(year_first),
    year_last_moco = max(year_last),
    .by = c(scheme, panel_split, module_combo_code)
  )
mod_scheme_yrs_moco_ps_yrsmoco %>% 
  filter(year_last_moco != year_last)
```

- conflicting first years:

```{r paged.print=FALSE}
mod_scheme_ps_different_1styr <- 
  mod_scheme_yrs_moco_ps_yrsmoco %>% 
  filter(year_first_moco != year_first)
mod_scheme_ps_different_1styr
if (nrow(mod_scheme_ps_different_1styr) > 0L) {
  warning("Conflicting first years detected!")
} 
```

We ignore differing last years (assuming the maximum last year for the module combo; see section \@ref(considerations)), but we must pay attention to additional remedies in case of conflicting first years.

**Missing functionality**: we currently don't correct the spatial sampling unit calendar for '`year_first` mismatches' between modules in the same module combo, if those occur.
Handling this comes down to filtering sampling units in the first panel(s) of the first cycle where such mismatches occur, hence where not all modules have started yet.
This requires calculation of module-level spatial sample sizes in poststrata.
More generally, we should add support for temporal variation in which modules are active inside a module combo, and take (semi-automated) decisions to either filter sampling units (post-hoc, based on full panels) and accept the resulting gaps in the GRTS series, or rearrange the panel membership using a single sampled GRTS series (this seems more appropriate if reductions are for the longer term).
Expansion with new, unplanned modules will always lead to rearranging the panel membership.

From inspection of `scheme_target_paneldesign`, we decide to use the following order to prioritize module combos in the GRTS sampling.

```{r}
moco_order <- tribble(
  ~module_combo_code, ~moco_rank,
  "panfl_pan5_pan2124_mbaa", 1L,
  "panfl_pan5_pan2124", 1L,
  "panfl_pan5_mbaa", 1L,
  "panfl_pan5", 1L,
  "panfl", 1L,
  "pan5_pan2124", 1L,
  "pan5", 1L,
  "mbaa", 1L,
  "pan2124", 2L
) %>% 
  mutate(module_combo_code = factor(
    module_combo_code,
    levels = levels(scheme_target_paneldesign$module_combo_code)
  ))
```

Note that panel splits are always ranked in ascending order, assuming that panel split 1 represents the _largest_ number of sampling units.
This should be taken into account when defining the panel splits.

To maximize matching of spatial samples across schemes, the module combo codes with the same rank within a common panel split rank will be forced to use the same subset of a GRTS series (i.e. GRTS address partition) for a particular stratum and domain!

Check that no module combos exist that are missed by above object.

```{r}
stopifnot(all(
  scheme_target_paneldesign$module_combo_code %in% moco_order$module_combo_code
))
```

Note that _different_ module combos within a scheme are **currently** handled as if they were different panel splits, i.e. they need a separate portion of the GRTS series to sample from; the starting address could be set based on sample size proportion but after that the starting address should be held fixed.

Note that this may not hold in general.
This is currently invoked by the `pan_transmon_be2100024` module, which explicitly intends to apply the higher frequency monitoring to **new** (hence _extra_) monitoring locations, relative to the other modules.
However it can also happen that future modules increase frequencies at the locations of _existing_ modules.
Nevertheless, it should be warned that the latter approach can lead to challenging situations, once an original module stops: the extra module didn't foresee the needed budget to support the base frequency and the remaining visits (of the new module) may be pointless.

It should also be accepted that for rare types, these samples may by necessity overlap between module combos.
Hence, this should be simplified afterwards to avoid unnecessary duplication of fieldwork planning.

## Spatial samples

### Set GRTS address thresholds per domain and stratum for cross-scheme synergies

Determining fixed GRTS address thresholds per domain and stratum to assign to module combos and panel splits, applicable across schemes, in order to maintain maximum synergy across schemes.

The threshold is the GRTS address start where a series should start.
For this, also the relationship of module combos between schemes must be investigated and aligned.

Note that this should in general solve the following challenge:

- within one module (say module 1), strata of a non-core scheme (say scheme A) received no explicit sample size when they are present in another core scheme (say scheme B) of the same module where they got an explicit sample size;
- in another module (say module 2), that non-core scheme (A) may be present as a core scheme, where such strata do have an explicit sample size;
- consequently, for the scheme at hand (A), not all involved modules cover their part in the sample size of that stratum.

As long as the involved modules sit together in one module combo (!), in both involved schemes (A and B), we can expect these module combos to have the same rank between both schemes (insofar they don't just match), so that the GRTS partition thresholds for the involved strata will be determined by the core scheme.

**Missing functionality**: still, the number of explicit sample locations in scheme A for those strata will be solely determined by the module where A is a core-scheme; hence these locations will not take into account the coverage of sample locations by scheme B.
It actually depends on whether, within module 1, FAGs are still missing in _non-core_ scheme A and _core_ scheme B to fully cover the needs of scheme A as if it were a core module.
If these needs are indeed covered, then the missing functionality is about taking into account a GRTS offset in sampling locations for module 2 (or alternatively always be explicit about sample sizes, including non-core schemes).
**This is to be solved whenever two modules are effectively combined where this indeed happens**.
If the needs are not covered, then overlapping locations between module 1 and 2 for the same scheme makes sense; then we only need to deduplicate overlapping FAGs in the revisit design, something to be addressed anyway across schemes.

Spatial sampling frame 0 is the most common one and also the unrestricted one, so we use that one to choose GRTS thresholds.

```{r paged.print=FALSE}
module_domain_scheme_stratum_sample_size %>% 
  count(sampling_frame_id, sort = TRUE)
```

We can see that sampling frame 0 is not used in groundwater schemes though, where restrictions always apply:

```{r}
scheme_sampling_frame %>% 
  mutate(
    compartment = str_extract(scheme, "^[A-Z]+") %>% 
      factor() %>% 
      fct_relevel("HQ", after = Inf)
  ) %>% 
  count(compartment, sampling_frame_id)
```

However for synergy with schemes from other compartments, we will still use it in all situations, also taking into account that the sample sizes for schemes in the SOIL compartment are currently the largest, and based on sampling frame 0.
This choice is also because it's better to generate thresholds further apart than too close together for certain schemes.

Defining the GRTS partition number, based on the module combo **rank** and panel split, for each domain and stratum, hence across schemes.

```{r}
# below object is created separately for checking purposes (see further:
# reclassified spatial samples object (without spare units))
scheme_moco_ps_dom_stratum_sample_size <- 
  module_domain_scheme_stratum_sample_size %>%
  select(
    module,
    domain,
    scheme,
    sampling_frame_id,
    stratum,
    sp_sample_size_all_panels_stratum,
    nunits
  ) %>%
  # for MNE, non-core schemes are dropped, except for specifically added types
  # for those non-core-schemes.
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>% 
  # joining module combos
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # sum the module sample sizes for common panel designs, scheme, domain and
  # stratum, as these sizes are going to be drawn as one GRTS address series;
  # however limit this result to the maximum module-specific population size:
  summarize(
    total_sample_size_all_panels = min(
      sum(sp_sample_size_all_panels_stratum),
      max(nunits)
    ),
    .by = c(scheme, module_combo_code, panel_split, domain, stratum, nunits)
  )
moco_dom_stratum_grtspart <- 
  scheme_moco_ps_dom_stratum_sample_size %>% 
  # use maximum sample sizes across schemes for common panel designs, domain and
  # stratum, since these should drive GRTS series partitioning:
  summarize(
    max_sample_size_all_panels = max(total_sample_size_all_panels),
    .by = c(module_combo_code, panel_split, domain, stratum)
  ) %>% 
  # add module combo rank:
  inner_join(
    moco_order,
    join_by(module_combo_code),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # determine rank number of the GRTS partition
  mutate(
    grts_part = rank(moco_rank * 1000 + panel_split, ties.method = "min"),
    .by = c(domain, stratum)
  )
```

Now setting the GRTS threshold (variable `grts_address_start`) for each GRTS partition in a domain x stratum.

To do this, we use the maximum required spatial sample size for a GRTS partition in a domain x stratum, and extend with a number of spare units by applying the applicable coefficient to this sample size.
For smaller GRTS address series (smaller populations), it is tried to optimally distribute the partitions but overlaps between GRTS partitions that belong to different module combos may be unavoidable as spatial sample sizes are necessarily only limited by the total population size (in domain x stratum), not by the size of partitions thereof.

Also, note that the actual spatial sample size and number of spare units in a GRTS partition x domain x stratum **for a specific scheme** may be lower.
So the below GRTS-partition-level spatial sample sizes and number of spare units only help in setting the thresholds, but will not be used in drawing spatial samples.

```{r}
dom_stratum_grtspart_threshold <- 
  moco_dom_stratum_grtspart %>% 
  # determine max sample size for each GRTS partition rank per domain x stratum,
  # which will ultimately drive the GRTS partitioning for the same revisit
  # design and across module combos with the same rank (note that different
  # panel splits always trigger another rank)
  summarize(
    max_sample_size_all_panels = max(max_sample_size_all_panels),
    .by = c(domain, stratum, grts_part)
  ) %>% 
  # characterize strata with 'is_strictly_aquatic' to determine number of 
  # spare units
  add_col_is_strictly_aquatic(
    strata = n2khab_strata,
    type_properties = n2khab_types_expanded_properties
  ) %>% 
  arrange(domain, stratum, grts_part) %>% 
  nest(splits = -c(domain, stratum, is_strictly_aquatic)) %>% 
  inner_join(
    sp_samplingframe_domain %>% 
      filter(sampling_frame_id == 0L) %>% 
      pluck("sampling_frame_domain", 1),
    join_by(domain, stratum),
    relationship = "one-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  nest(grts_addresses = grts_address) %>% 
  # now determine threshold for each domain x stratum combination
  mutate(
    splits = pmap(
      list(is_strictly_aquatic, splits, grts_addresses), 
      function(aq, spl, grts) {
        spl %>% 
          mutate(
            # TEMPORARY HACK to fix two cases of 3260 (domains BE2100024 and
            # BE2200039) where more than one unit is linked to the same GRTS
            # address. TO BE SOLVED XXXXXXX
            # ################################
            max_sample_size_all_panels = ifelse(
              max_sample_size_all_panels > length(grts$grts_address),
              length(grts$grts_address),
              max_sample_size_all_panels
            ),
            ##################################
            n_spare_units_theoretical = round(max_sample_size_all_panels * 
              ifelse(aq, spare_coef_aq, spare_coef_default)),
            # first GRTS address of a GRTS partition
            grts_address_start = pick_grts_thresholds(
              max_sample_size_all_panels,
              n_spare_units_theoretical,
              grts$grts_address
            ),
            # index of the first GRTS address of a GRTS partition
            grts_index_start = which(
              sort(grts$grts_address) %in% grts_address_start
            ),
            # index of the last GRTS address of the max sample size per GRTS
            # partition. Note that this may exceed the index_start of the next
            # GRTS partitions if the partitions overlap.
            grts_index_end = grts_index_start + max_sample_size_all_panels - 1L,
            # index of the last GRTS address of the GRTS partition if we define
            # it using the index_start of the next partition minus 1. This
            # may be smaller than grts_index_end in overlap cases
            grts_gap_end = dplyr::lead(
              grts_index_start - 1L, 
              default = length(grts$grts_address)
            ),
            # GRTS address corresponding to grts_gap_end
            grts_address_gap_end = sort(grts$grts_address)[grts_gap_end],
            # maximum number of spare units that we will select in a GRTS
            # partition
            n_spare_units = pmin(
              n_spare_units_theoretical,
              grts_gap_end - grts_index_end
            ) %>% 
              as.integer()
          )
      })
  ) %>% 
  select(-grts_addresses, -is_strictly_aquatic) %>% 
  unnest(splits)
```

Some characteristics of `dom_stratum_grtspart_threshold`, especially with relation to number of spare units.

```{r}
nrow(dom_stratum_grtspart_threshold)
dom_stratum_grtspart_threshold %>% 
  filter(n_spare_units == 0) %>% 
  nrow()
dom_stratum_grtspart_threshold %>% 
  filter(n_spare_units == 0) %>% 
  pull(stratum) %>% 
  n_distinct()
dom_stratum_grtspart_threshold %>% 
  filter(n_spare_units < n_spare_units_theoretical) %>% 
  nrow()
```

Cases may exist where GRTS partitions overlap:

```{r paged.print = FALSE}
dom_stratum_grtspart_threshold %>% 
  filter(grts_index_end > grts_gap_end) %>% 
  count(domain, grts_part, name = "number_of_strata")
```

**Missing functionality**: write a file with these thresholds (GRTS partition definitions), so that it can be read in a static way from some point on!!


### Apply the GRTS address threshold to the module combos and panel splits

Expanding the GRTS-partition-level GRTS-threshold to the level of module combos and panel splits (per domain and stratum).

```{r}
moco_dom_stratum_grtspart_threshold <- 
  moco_dom_stratum_grtspart %>% 
  select(-max_sample_size_all_panels, -moco_rank) %>% 
  inner_join(
    dom_stratum_grtspart_threshold %>% 
      select(domain, stratum, grts_part, grts_address_start, grts_address_gap_end),
    join_by(grts_part, domain, stratum),
    relationship = "many-to-one",
    unmatched = "error"
  )
```

Joining the GRTS threshold with module, scheme and panel split.
<!-- This may be superfluous!! XXXXX -->

```{r eval=FALSE}
mod_dom_scheme_ps_stratum_grtspart <- 
  moco_dom_stratum_grtspart_threshold %>% 
  # expanding to modules and schemes:
  inner_join(
    scheme_target_paneldesign %>% 
      select(module_combo_code, module_combo, scheme, panel_split) %>% 
      unnest(module_combo),
    join_by(module_combo_code, panel_split),
    relationship = "many-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  # restrict to the target populations of the concerned mod x domain x scheme:
  semi_join(
    module_domain_scheme_stratum_sample_size,
    join_by(module, domain, scheme, stratum)
  ) %>% 
  select(-module_combo_code) %>% 
  relocate(module, domain, scheme, panel_split)
```

### Draw spatial samples

Determining the applicable domain relations.

```{r}
domain_relations_spsamples <- 
  domain_overlaps %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size %>% 
      distinct(module, domain),
    join_by(domain)
  ) %>% 
  semi_join(
    module_domain_scheme_stratum_sample_size %>% 
      distinct(module, domain),
    join_by(included_by == domain)
  )
including_domain <- unique(domain_relations_spsamples$included_by)
```

Drawing spatial samples per scheme and GRTS partition (hence for module combos x panel splits), for each stratum x domain.
This includes the step to redraw the sample for domains included by another domain (typically because of module combinations), to ensure that no units are drawn more than once and the GRTS series in the domain is one series.

Note that the term 'domain' causes some confusion when considered across modules:

- Each module targets one or multiple (geographically defined) domains with a predetermined sample size.
- While domains may overlap between modules, the domain name is kept as foreseen by a module, since the sample size must be maintained at that level for the module.
Sample sizes of two modules are summed only when domains are identical.
- Next, the specified number of sampling units is drawn for each domain.
- In a separate step, it is taken care of to have no common sampling units between modules where domains overlap.
In this step, the membership of each unit with respect to a smaller, included domain is taken into account, so that all units that belong to that domain, across modules, have contiguous GRTS addresses.
Nevertheless, their original (module-bound) domain attributeis still kept.
Sampling units from smaller domains that have a shifted GRTS address to avoid overlap with sampling units from a larger, including domain, have a value `TRUE` in the `extends_on_larger_domain` column of below created `scheme_moco_ps_stratum_dom_spsamples`.
- Consequently, which units belong to an overlapped domain across modules will not yet be apparent from the resulting object.
We will need this though in order to add spare units and apply the revisit design.
Preparing for this will be the topic see section \@ref(sppoststrata) (spatial poststrata).


```{r}
scheme_moco_ps_stratum_dom_spsamples_nested <-
  module_domain_scheme_stratum_sample_size %>%
  select(module, domain, scheme, sampling_frame_id, stratum, sp_sample_size_all_panels_stratum) %>%
  # For non-core-schemes, only specifically added types with explicit sample
  # size are kept, i.e. below filter step (the other sampling units are covered
  # already by the data collection of core types). This is the logical approach
  # for spatial sampling and data collection plans, however for inference we
  # also need to compile an expanded object with all applicable combinations of
  # sampling unit and scheme.
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>% 
  # joining module combos
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # Domains that are the same between different modules are integrated within
  # module combos by summing the sample sizes; no need to limit here by
  # population size since the sampling is done on the available units (however
  # note that population size already applies phab, while this may not (yet) be
  # the case for the spatial sampling frames). Modules are collapsed to their
  # combinations:
  summarize(
    sp_sample_size_all_panels_stratum = sum(sp_sample_size_all_panels_stratum),
    .by = c(scheme, module_combo_code, panel_split, sampling_frame_id, domain, stratum)
  ) %>% 
  # set first GRTS address of the GRTS partition (GRTS threshold)
  inner_join(
    moco_dom_stratum_grtspart_threshold %>% 
      select(-grts_part),
    join_by(module_combo_code, panel_split, domain, stratum),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  nest(moco_ps_domain_stratum_spsamplesize = c(
    module_combo_code, 
    panel_split,
    domain, 
    stratum, 
    sp_sample_size_all_panels_stratum,
    grts_address_start,
    grts_address_gap_end
  )) %>%
  inner_join(
    sp_samplingframe_domain, 
    join_by(sampling_frame_id),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>%
  # drawing an independent sample per scheme, domain, stratum and module combo & panel split:
  mutate(
    spsample = map2(moco_ps_domain_stratum_spsamplesize, sampling_frame_domain, function(dss, sfd) {
      dss %>% 
        inner_join(
          sfd %>% 
            nest(grts_addresses = grts_address),
          join_by(domain, stratum),
          relationship = "many-to-one",
          unmatched = c("error", "drop")
        ) %>% 
        # the grts_addresses_subset implements the GRTS threshold restriction
        mutate(
          grts_addresses_subset = map2(
            grts_address_start, 
            grts_addresses,
            function(grts_start, grts) {
              grts %>% filter(grts_address >= grts_start)
            }
          ),
          sample =
            map2(
              sp_sample_size_all_panels_stratum,
              grts_addresses_subset, 
              function(sss, grts_sub) {
                grts_sub %>% 
                  slice_min(grts_address, n = sss)
              }
            )
        ) %>% 
        select(-c(grts_addresses, grts_address_start, grts_address_gap_end)) %>% 
        nest(domain_spsample = c(
          domain, 
          sp_sample_size_all_panels_stratum, 
          grts_addresses_subset,
          sample
        ))
    })
  ) %>%
  select(scheme, spsample) %>%
  unnest(spsample) %>% 
  # handling the overlap between domains within module combo x panel split:
  # above, the samples have been drawn independently so now this has to be
  # corrected!
  mutate(
    domain_spsample = map(domain_spsample, function(ds) {
      ds <- left_join(
        ds,
        domain_relations_spsamples,
        join_by(domain),
        relationship = "one-to-one",
        unmatched = "drop"
      )
      # vector of including domains:
      including_domains <- unique(ds$included_by)
      # drop NAs:
      including_domains <- including_domains[!is.na(including_domains)]
      # only keep an including domain if it is effectively present as domain:
      including_domains <- including_domains[including_domains %in% ds$domain]
      if (length(including_domains) > 1) {
        stop(
          "Multiple including domains in one module combo x panel split",
          " x stratum are not supported."
        )
      }
      ds %>% 
        mutate(
          including = domain %in% including_domains | length(including_domains) == 0,
          # put the sample of the including domain next to the included domains:
          grts_sampled = ifelse(
            including, 
            tibble(),
            sample[including]
          ),
          # determine the available unsampled units per domain:
          grts_addresses_unsampled = map2(
            grts_addresses_subset, 
            grts_sampled, 
            function(ga, gs) {
              if (is.null(gs)) {
                return(NULL)
              }
              anti_join(ga, gs, join_by(grts_address))
            }),
          # redraw samples for included domains:
          sample_2 = pmap(
            list(including, sample, grts_addresses_unsampled, sp_sample_size_all_panels_stratum),
            function(inc, sam, grts, sss){
              if (inc) {
                sam
              } else {
                grts %>% slice_min(grts_address, n = sss)
              }
            }
          ),
          # flag sampling units of included domains, as they should never be
          # considered apart from the including domain:
          extends_on_larger_domain = !including & length(including_domains) == 1
        ) %>% 
        select(
          domain,
          including,
          grts_addresses_subset,
          sample = sample_2,
          extends_on_larger_domain
        )
    })
  )
      
scheme_moco_ps_stratum_dom_spsamples <- 
  scheme_moco_ps_stratum_dom_spsamples_nested %>% 
  mutate(
    domain_spsample = map(domain_spsample, function(ds) {
      ds %>% 
        select(domain, sample = sample, extends_on_larger_domain) %>% 
        # next step will also drop domains where no sampling units were
        # available (i.e. the whole population was sampled by the larger domain):
        unnest(sample)
    })
  ) %>% 
  unnest(domain_spsample)
```

Check that no sampling units overlap between domains within the combination of module combo, scheme, panel split and stratum:

```{r}
scheme_moco_ps_stratum_dom_spsamples %>% 
  count(scheme, module_combo_code, panel_split, stratum, grts_address) %>% 
  filter(n > 1) %>% 
  nrow() == 0
```

**Missing functionality**: generate object that reflects relationship between all sampling units and schemes, including non-core schemes, needed for the inference stage.
Currently, we only consider the needs for sampling and data collection, hence not the exact duplications between (parts of) core-schemes and (parts of) non-core schemes.
See also the above code comment about non-core schemes.
In essence this means duplicating existing sampling units using applicable rows from `module_domain_scheme_stratum_sample_size`, `mod_scheme_target_paneldesign` or `scheme_target_paneldesign` with relation to non-core schemes.
This object then also contains the spatial poststrata, needed for inference, and we also need to add the sampling weights at that level.


### Reclassify by spatial poststratum and add spare units {#sppoststrata}

The `scheme_moco_ps_stratum_dom_spsamples` object distinguishes domains which can spatially overlap.
This is caused by the domains envisaged by specific modules, and the sampling units attributed to the different domains are mutually exclusive.
It is also a consequence of the fact that sample sizes are typically set at the module x domain x scheme level, so it reflects this module-wise sampling approach.
Hence in this object it is clear which sampling units are added surplus in specific (included) domains in order to increase coverage of those domains.

However, given samples that belong to different and potentially overlapping domains, we must take into account several aspects:

- **Panel memberships** will be based on the GRTS series for each revisit design of a stratum in a scheme, hence for a module combo x panel split that applies to this scheme (this has already been prepared by implementing the GRTS partition thresholds).
This is done in order to maintain long-term spatial balance per revisit design and to minimize the possibility that sampling units jump from one revisit design to another as a consequence of changes in the spatial sampling frame.
Furthermore, the panel membership design is created separately for each included (or multiple non-overlapping) domain and for the remainder of the including domain.
Otherwise the units of an included domain would not be well spread over the panels when specifically looking at an included domain (for which a module is obviously present, hence an inference requested).
- We will refer to these mutually exclusive domains / domain remainders as _**spatial poststrata**_, as they will effectively be used for poststratification in the inference about the overall (including) domain.
- Consequently, a separate set of **spare units** is needed for each revisit design of a stratum in a scheme, hence they must be generated at the level of spatial poststrata in a revisit design.
The _number_ of spare units will be determined based on the stratum sample size of each spatial poststratum, for a given revisit design (~ module combo x panel split) in a scheme.

Now generating the spatial poststrata (which replace the domain level) and adding spare units.
(Note: for domain-specific inferences, re-introduction of domains is needed, where units in overlapped domains will be repeated for each domain but with different sampling weights.)

```{r}
scheme_moco_ps_stratum_sppost_spsamples_spares <- 
  scheme_moco_ps_stratum_dom_spsamples_nested %>% 
  # characterize strata with 'is_strictly_aquatic' to determine number of 
  # spare units
  add_col_is_strictly_aquatic(
    strata = n2khab_strata,
    type_properties = n2khab_types_expanded_properties
  ) %>% 
  # determine the coefficient to calculate the theoretical number of spare units
  mutate(
    spare_coef = ifelse(is_strictly_aquatic, spare_coef_aq, spare_coef_default)
  ) %>% 
  select(-is_strictly_aquatic) %>% 
  # add list columns for scheme x module combo x panel split x stratum:
  mutate(
    # 1. vector that defines relevant spatial poststrata
    sp_poststrata = map(domain_spsample, function(ds) {
      n_included_doms <- sum(!ds$including)
      ds %>% 
        mutate(
          sp_poststratum = ifelse(
            including & n_included_doms > 0,
            str_c(domain, "_remainder"),
            as.character(domain)
          )
        ) %>% 
        pull(sp_poststratum) %>% 
        unique()
    }),
    # 2. tibble with the GRTS addresses available for sampling (i.e. sampling
    # frame limited by applicable domains and by the GRTS partition),
    # integrated over all domains
    overall_grts_addresses_subset = map(domain_spsample, function(ds) {
      ds_poststrata <- 
      ds %>% 
        pull(grts_addresses_subset) %>% 
        reduce(dplyr::union)
    }),
    # 3. tibble with the sampled GRTS addresses, integrated over all domains
    overall_sample = map(domain_spsample, function(ds) {
      ds %>% 
        pull(sample) %>% 
        reduce(dplyr::union)
    }),
    # 4. tibble with: spatial poststrata, corresponding sampling units and
    # spare units
    sp_poststr_samples_spareunits = pmap(
      list(
        spare_coef,
        domain_spsample,
        sp_poststrata,
        overall_grts_addresses_subset,
        overall_sample
      ),
      function(coef, ds, poststr, grts, sam) {
        smaller_domains <- 
          ds %>% 
          # if the general including domain (Flanders) is not present, smaller
          # domains are marked as including
          filter(!including | !(domain %in% including_domain)) %>% 
          pull(domain)
        poststr_remainder <- poststr[!poststr %in% smaller_domains]
        if (length(poststr_remainder) > 1 && length(smaller_domains) > 0) {
          stop("Mismatch between spatial poststrata and domains!")
        }
        # create the relation between spatial poststrata and all available GRTS
        # addresses, and mark the sampling units
        sp_poststr_all <-
          grts %>%
          left_join(
            domain_grts_n2khab %>% 
              {
                if (length(smaller_domains) > 0) {
                  filter(., domain %in% smaller_domains)
                } else {
                  if (length(including_domain) != 1L) {
                    stop("The left_join approach needs to be revised.")
                  }
                  filter(., domain %in% including_domain)
                }
              },
            join_by(grts_address),
            relationship = "one-to-one",
            unmatched = "drop"
          ) %>%
          rename(sp_poststratum = domain) %>%
          relocate(sp_poststratum) %>%
          mutate(
            sp_poststratum = ifelse(
              is.na(sp_poststratum),
              poststr_remainder,
              as.character(sp_poststratum)
            ) %>%
              factor(),
            sample_status = ifelse(
              grts_address %in% sam$grts_address,
              "in_sample",
              NA_character_
            )
          ) %>%
          arrange(sp_poststratum, grts_address)
        # generate the spare units per spatial poststratum
        sp_poststr_spareunits <-
          sp_poststr_all %>%
          generate_spare_units(coef_spare = coef)
        # combine sample and spare units per spatial poststratum
        sp_poststr_all %>%
          filter(sample_status == "in_sample") %>%
          bind_rows(sp_poststr_spareunits) %>%
          mutate(sample_status = factor(sample_status)) %>% 
          arrange(sp_poststratum, grts_address)
      }
    )
  ) %>% 
  select(
    scheme,
    module_combo_code,
    panel_split,
    stratum,
    sp_poststr_samples_spareunits
  ) %>% 
  arrange(scheme, module_combo_code, panel_split, stratum)
```

Total size of `scheme_moco_ps_stratum_sppost_spsamples_spares` after unnesting.

```{r}
scheme_moco_ps_stratum_sppost_spsamples_spares %>% 
  unnest(sp_poststr_samples_spareunits) %>% 
  nrow()
```

Overall representation of spatial poststrata in the spatial samples (of all modules, schemes and strata):

```{r paged.print=FALSE}
scheme_moco_ps_stratum_sppost_spsamples_spares %>% 
  unnest(sp_poststr_samples_spareunits) %>% 
  mutate(sp_poststratum = factor(as.character(sp_poststratum))) %>%
  count(sp_poststratum)
```

In general, note that the above spatial poststrata are to be interpreted within a module combo x scheme x stratum; this holds especially for the 'remainder' poststratum which is defined relative to the smaller poststrata (domains).
This is stratum dependent since not each stratum is represented in the smaller poststrata (domains).

Overall distribution of sample status (across all modules, schemes and strata):

```{r paged.print=FALSE}
scheme_moco_ps_stratum_sppost_spsamples_spares %>% 
  unnest(sp_poststr_samples_spareunits) %>% 
  count(sample_status)
```


## Implementing the revisit design

**Missing functionality**: make the code below work for other field activities than those in module `pan_effectmon_flanders` (e.g. wrt prefix extraction).

<!-- --------------------------- -->
Adding some ideas:

- organize the membership design for the combination of field activity supergroups, i.e. with the same panel design across FAGs, and sample size. Still within module combo x scheme since sample sizes may otherwise differ, but across schemes where they don't.
- also make membership design linked between such different supergroups of the same sample size, where these have a compatible number of panels, e.g. if one panel in one supergroup can be the sum of 4 panels of another supergroup.
- start with the supergroups that coincide with the panel designs of the _main_ field method of _target_ variables, as we plan to optimize the spatial samples in order to deal with the representation of those panel designs across modules, schemes and domains.
- apply non-core FAG panels only to locations not covered by a corresponding core FAG
- select a single winning FAG for competition in time (for the same FAG) or between interlinked FAGs (take the superior one and schedule it as the first one). However don't do this for the FAG with the main field method of each scheme: the data collection design must not be harmed!
- for MNE, the LOCEVAL activity must take place one year beforehand!
- SPATPOSITxxx: inflate the panel periods by 3 months at the end (so they will overlap)
- ADHOCxxx: reduce the number of occasions by sampling them
- in the first year of each compartment, start on 1 Jan and wind up to month_first_default
<!-- --------------------------- -->

### Preparations

#### Isolating the spatial samples (without spare units)

`scheme_moco_ps_stratum_sppost_spsamples` is like `scheme_moco_ps_stratum_sppost_spsamples_spares` without the spare units.

```{r}
scheme_moco_ps_stratum_sppost_spsamples <- 
  scheme_moco_ps_stratum_sppost_spsamples_spares %>% 
  mutate(
    sp_poststr_samples_spareunits = map(sp_poststr_samples_spareunits, function(df) {
      filter(df, sample_status == "in_sample")
    })
  ) %>% 
  rename(sp_poststr_samples = sp_poststr_samples_spareunits)
```

Checking that all schemes x module combos x panel splits x strata are still in the running.
Listing the cases where the planned sample size is different than the one after GRTS series partitioning and reclassification (smaller: possible because of phab-corrections applied in the planned sample size but not (yet) in the samples; larger: possible because of the difference between domain 'Flanders' & spatial poststratum 'Flanders_remainder' (many; not shown), and sometimes also due to minor misfits in 3260 (popsize vs available GRTS addresses per domain)).

<!-- Perhaps look into this 3260 difference; not looked in to further now since it's very limited (2 SACs where the sample has 1 location less). Also, this may depend on the active modules. -->

```{r}
scheme_moco_ps_dom_stratum_sample_size_positive <- 
  scheme_moco_ps_dom_stratum_sample_size %>% 
  filter(total_sample_size_all_panels > 0)
scheme_moco_ps_dom_stratum_sample_size_plannedvspartitioned <- 
  scheme_moco_ps_dom_stratum_sample_size_positive %>% 
  rename(total_sample_size_planned = total_sample_size_all_panels) %>% 
  inner_join(
    scheme_moco_ps_stratum_sppost_spsamples %>% 
      mutate(
        ss_by_sppost = map(sp_poststr_samples, ~count(., sp_poststratum, name = "total_sample_size_after_partitioning")),
        .keep = "unused"
      ) %>% 
      unnest(ss_by_sppost),
    join_by(scheme, module_combo_code, panel_split, stratum),
    relationship = "many-to-many",
    unmatched = "error"
  ) %>% 
  filter(
    as.character(domain) == as.character(sp_poststratum) | 
      (sp_poststratum == "Flanders_remainder" & domain == "Flanders")
  )
stopifnot(
  nrow(scheme_moco_ps_dom_stratum_sample_size_positive) == 
    nrow(scheme_moco_ps_dom_stratum_sample_size_plannedvspartitioned)
)
scheme_moco_ps_dom_stratum_sample_size_plannedvspartitioned %>% 
  filter(
    total_sample_size_planned != total_sample_size_after_partitioning,
    sp_poststratum != "Flanders_remainder"
  ) %>% 
  relocate(stratum) %>% 
  arrange(stratum, scheme, module_combo_code, panel_split)
```

We can see that identical spatial samples ('identical' includes the sample size) across schemes, panel splits and module combos only occur for smaller strata:

```{r paged.print=FALSE}
scheme_moco_ps_stratum_sppost_spsamples %>% 
  count(stratum, sp_poststr_samples, name = "number_of_scheme_ps_moco") %>% 
  filter(number_of_scheme_ps_moco > 1) %>% 
  print(n = Inf)
```

Consequently we cannot generalize spatial samples above the level of scheme, panel split and module combo in the revisit design.

However several measures can be taken to maximize the synergy in panel memberships and revisit dates:

- Within a scheme we will combine multiple FAGs that have the same revisit design, called field activity supergroups (FASs).
- For the above strata where spatial samples do match across schemes, panel membership is to be aligned.
- Also, we will maximize the chance for cross-scheme and cross-compartment spatial alignment of sampling units by starting all panel designs on the same virtual date, and drop non-applicable (too early) parts afterwards.

#### Lifting mod_scheme_fag_paneldesign to the module combo & FA supergroup level

The module combos (of a scheme x panel split) reflect a common revisit design of the main field method of the target variable, across modules.
We use this entity to populate all the different FAG panel designs (grouped into field activity supergroups) with the same corresponding sampling units (from `scheme_moco_ps_stratum_sppost_spsamples`).
So this means that the spatial sample of a module combo in a scheme x panel split will be used for all field activity revisit designs of this scheme x panel split, including that of the main field method of the target variable.

A field activity supergroup (FAS) combines multiple FAGs with the same revisit design, in the same scheme x module combo x panel split x spatial subset.
Membership designs should be implemented at the FAS level.
FASs of different spatial subsets can share their revisit design, scheme, module combo and panel split, but they will normally contain a different (or only partially overlapping) set of FAGs; otherwise the spatial subset would be `NA` and the distinction between the FASs would not be made.

<!-- Check whether the combination of modules in a module combo where the same scheme is a core and a non-core scheme because of the different modules, causes a clash for the involved FAGs since all FAGs are applied to all sampling units of the module combo. -->

<!-- One solution may be to still lump those different incompatible FAGs of a FA supergroup across modules (as we currently do), and indeed apply the panel membership at this level, but afterwards filter the FAGs per sampling unit according to the module membership of each sampling unit, which assumes module labelling of sampling units! Module membership then may need to be an interspersion of the modules in the GRTS address series, since otherwise this conflicts with spread over panels for each module (but then we do sacrifice the spatial balance at module level!) -->

<!-- The other solution may be to work at the module level right away, hence respecting different sets of FAGs per module; this as well assumes module labelling of sampling units! -->

<!-- For now I prefer to ignore this on condition that no modules are effectively combined that pose this problem, hence we always need to verify the existence of clashes, for main field methods (which directly relate to target environmental vars) and for activity sequences (which trigger the FAGs). -->

<!-- Note that the following is wrong, since the spatial samples and spatial subsets already cover 'stratum': ~Since different modules can have different target populations for the same scheme, hence different applicable spatial subsets, we must do this at the stratum level.~ -->

Before we execute this plan, we must verify the degree to which:

- main field methods are repeated _in all modules_ in the same scheme x module combo x panel split.
Where this is not the case, it must be verified whether the combination of modules poses problems at the level of the interpretation of the involved (target) environmental variable, and it should be decided whether this combination of modules can be held.

  Main field methods not represented across all modules in the same scheme x module combo x panel split:

```{r}
moco_mod_scheme_ps_conflicting_mainfieldmethods <- 
  mod_scheme_mainfieldmethod %>% 
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  # determine number of different modules per scheme x module combo x panel
  # split:
  mutate(
    n_mod = n_distinct(module),
    .by = c(scheme, module_combo_code, panel_split)
  ) %>% 
  # differentiate number of different modules for each main field method
  mutate(
    n_mod_mainfieldmethod = n_distinct(module),
    .by = c(scheme, module_combo_code, panel_split, main_field_method)
  ) %>% 
  # we like both of these numbers to be the same, but which ones are different?
  filter(n_mod != n_mod_mainfieldmethod)
moco_mod_scheme_ps_conflicting_mainfieldmethods
```

<!-- A potential conflict is observed between the PAN effectmonitoring vs the MBAA modules, for reading surfacewater levels in GW_05.1_aq. -->
<!-- However the involved environmental variables are temporally aggregated variables, and these modules will normally not clash as they will rather apply upgrades to the field methods of each other's locations (but sample sizes will not always be equal) rather than add new locations on top of each other (this is not taken into account; their activity sequences have been chosen to represent the individual modules, not their combination). -->

  Target variables of the associated modules and schemes:

```{r}
envvars %>% 
  filter(target) %>% 
  semi_join(
    moco_mod_scheme_ps_conflicting_mainfieldmethods,
    join_by(module, scheme)
  ) %>% 
  distinct(pick(starts_with("env"), target, main_datacollection_method))
```

- activity sequences are repeated _in all modules_ in the same scheme x module combo x panel split.
Similarly we have to decide potential consequences (restrictions) in the combination of modules if this is not the case.

```{r}
mod_scheme_mainfieldmethod_actseq %>% 
  distinct(module, scheme, in_aquatic_subset, activity_sequence,) %>% 
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  # determine number of different modules per scheme x module combo x panel
  # split:
  mutate(
    n_mod = n_distinct(module),
    .by = c(scheme, module_combo_code, panel_split, in_aquatic_subset)
  ) %>% 
  # differentiate number of different modules for each activity sequence
  mutate(
    n_mod_actseq = n_distinct(module),
    .by = c(
      scheme, 
      module_combo_code, 
      panel_split, 
      in_aquatic_subset, 
      activity_sequence
    )
  ) %>% 
  # we like both of these numbers to be the same, but which ones are different?
  filter(n_mod != n_mod_actseq)
```

<!-- We only have a conflict between the PAN effectmonitoring vs the MBAA modules, for the activity sequence for groundwater (and surfacewater) levels. -->
<!-- We see no actual clashes for this combination, see reasoning above, provided that only one (the superior) activity sequence is actually applied to a single location. -->

We must bear in mind that mixing and applying such conflicting field method / activity sequence combinations to each module of the module combo (all sampling units) does not make sense, except when the  different field method / activity sequence is divided between different sampling units according to the module.
We currently don't handle a clash like this; for now we still expect the modules in a module combo to use only common methods.

**Missing functionality**: combinations of modules might in some cases (specific module combinations!) apply to the _same_ sampling units (or at least maximize the common sampling unit set), and only differ in the applicable field activities, so that a choice is to be made between the latter.
Our code currently does not handle this case; we just sum the module sample sizes in all cases and apply the overall set of FAGs to all involved sampling units.
When the above code chunks output a result, this will normally indicate such an unsupported module combination.

Creating preparatory objects to lift `mod_scheme_fag_paneldesign` to the module combo & FAS level.

```{r}
# object with revisit design for each scheme x moco x panel split x [FAS x
# spatial subset]
scheme_moco_ps_fas_paneldesign <-
  mod_scheme_fag_paneldesign %>% 
  select(-actseqs) %>% 
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme, panel_split),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  relocate(module_combo_code, .after = scheme) %>% 
  arrange(module, scheme, field_activity_group, in_aquatic_subset) %>% 
  # adding applicable activity sequences
  inner_join(
    mod_scheme_actseq_fag %>% 
      select(
        module, 
        scheme, 
        is_core_scheme, 
        field_activity_group, 
        in_aquatic_subset, 
        activity_sequence
      ),
    join_by(module, scheme, is_core_scheme, field_activity_group, in_aquatic_subset),
    relationship = "one-to-many",
    unmatched = "error"
  ) %>% 
  # prepare for collapsing
  mutate(
    actseqs = str_c(sort(unique(activity_sequence)), collapse = ", "),
    actseqs_vec = list(sort(unique(activity_sequence))),
    .by = !c(module, is_core_scheme, activity_sequence)
  ) %>% 
  summarize(
    fas_spsubset = list(
      tibble(
        field_activity_group = field_activity_group,
        in_aquatic_subset = in_aquatic_subset,
        rank = rank,
        max_rank = max_rank,
        actseqs = actseqs,
        actseqs_vec = actseqs_vec
      ) %>% 
        distinct()
    ),
    .by = !c(
      module, 
      is_core_scheme, 
      field_activity_group, 
      rank, 
      max_rank, 
      in_aquatic_subset, 
      activity_sequence,
      actseqs,
      actseqs_vec
    )
  ) %>% 
  relocate(fas_spsubset, .after = panel_split)

# object with FAGs and explicit spatial subsets (as in mod_scheme_field_activity)
scheme_moco_ps_fagset_spsubsetfull <-
  mod_scheme_field_activity %>%
  distinct(
    module, 
    scheme, 
    is_core_scheme, 
    field_activity_group, 
    in_aquatic_subset,
    activity_sequence, 
    rank,
    max_rank
  ) %>% 
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  relocate(module_combo_code, .after = scheme) %>% 
  arrange(module, scheme, field_activity_group, in_aquatic_subset) %>% 
  mutate(
    rank = min(rank),
    max_rank = if (all(is.na(max_rank))) NA else max(max_rank, na.rm = TRUE),
    actseqs = str_c(sort(unique(activity_sequence)), collapse = ", "),
    actseqs_vec = list(sort(unique(activity_sequence))),
    .by = c(
      scheme,
      module_combo_code,
      panel_split,
      field_activity_group,
      in_aquatic_subset
    )
  ) %>% 
  summarize(
    fag_spsubsetfull = list(
      tibble(
        field_activity_group = field_activity_group,
        in_aquatic_subset = in_aquatic_subset,
        rank = rank,
        max_rank = max_rank,
        actseqs = actseqs,
        actseqs_vec = actseqs_vec
      ) %>% 
        distinct()
    ),
    .by = !c(
      module, 
      is_core_scheme, 
      field_activity_group, 
      in_aquatic_subset,
      rank,
      max_rank,
      activity_sequence,
      actseqs,
      actseqs_vec
    )
  ) %>% 
  relocate(fag_spsubsetfull, .after = panel_split)
```

Creating final object.
It expands `scheme_moco_ps_fas_paneldesign` to explicit spatial subsets, but only where needed, i.e. we want each row of `scheme_moco_ps_fas_paneldesign` to have spatial subsets for all FAGs or for none.
Also, it unnests the spatial subset.

```{r}
scheme_moco_ps_spsubset_fas_paneldesign <-
  scheme_moco_ps_fas_paneldesign %>%
  arrange(scheme, module_combo_code) %>% 
  inner_join(
    scheme_moco_ps_fagset_spsubsetfull,
    join_by(scheme, module_combo_code, panel_split),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  mutate(
    fas_spsubset = map2(
      fas_spsubset, 
      fag_spsubsetfull, 
      function(fsp, f) {
        if (anyNA(fsp$in_aquatic_subset) && any(!is.na(fsp$in_aquatic_subset))) {
          fsp %>% 
            inner_join(
              f %>% 
                rename(
                  in_aquatic_subset_explicit = in_aquatic_subset,
                  rank_explicit = rank,
                  max_rank_explicit = max_rank,
                  actseqs_explicit = actseqs,
                  actseqs_vec_explicit = actseqs_vec
                ),
              join_by(field_activity_group),
              relationship = "one-to-many",
              unmatched = c("error", "drop")
            ) %>% 
            mutate(
              rank = ifelse(
                is.na(in_aquatic_subset),
                rank_explicit,
                rank
              ),
              max_rank = ifelse(
                is.na(in_aquatic_subset),
                max_rank_explicit,
                max_rank
              ),
              actseqs = ifelse(
                is.na(in_aquatic_subset),
                actseqs_explicit,
                actseqs
              ),
              actseqs_vec = ifelse(
                is.na(in_aquatic_subset),
                actseqs_vec_explicit,
                actseqs_vec
              ),
              in_aquatic_subset = ifelse(
                is.na(in_aquatic_subset),
                in_aquatic_subset_explicit,
                in_aquatic_subset
              )
            ) %>% 
            distinct(
              field_activity_group, 
              in_aquatic_subset, 
              rank, 
              max_rank, 
              actseqs,
              actseqs_vec
            )
        } else {
          fsp
        }
    })
  ) %>% 
  select(-fag_spsubsetfull) %>% 
  unnest(fas_spsubset) %>% 
  mutate(actseqs = factor(actseqs)) %>% 
  nest(fas = c(field_activity_group, rank, max_rank, actseqs, actseqs_vec)) %>% 
  relocate(fas, .after = in_aquatic_subset)
```

Let's verify that no FAGs have been duplicated because `rank` or `max_rank` were not unique.

```{r}
stopifnot(
  scheme_moco_ps_spsubset_fas_paneldesign %>% 
    mutate(
      fas_nrow = map_int(fas, ~ nrow(.)),
      fas_simple_nrow = map_int(fas, function(df) {
        df %>% 
          distinct(field_activity_group) %>% 
          nrow
      })
    ) %>% 
    filter(fas_nrow != fas_simple_nrow) %>% 
    nrow() == 0
)
```


### Number of panels {#nrpanels}

Determining the _number_ of panels for each scheme x module combo x panel split x spatial subset x 'field activity supergroup' (FAS).
We also generate (or keep) revisit design attributes needed to set date intervals in later steps.
Panel grouping as reflected by `panel_group_size` and `panel_group_repetitions` refers to the subgroup structure in nested serially alternated patterns.

Note, while below object does not distinguish between spatial poststrata, we will effectively populate the panels taking into account the spatial poststrata.
Also the spare units are considered per spatial poststratum, so that replacements with spare units cause a shift in the panel membership _within_ the spatial poststratum.

<!-- Still need to choose whether panels therefore must be defined separately between spatial poststrata, or not. -->
<!-- In fact they shouldn't need to be if the spatial poststratum is an attribute of the sampling (or spare) unit, although it is possible that this poses problems at some stage. -->

```{r}
scheme_moco_ps_spsubset_fas_panelcount_full <-
  mod_scheme_yrs_moco_ps %>% 
  # year_first has been set module and scheme specific, while year_last is just
  # module specific. However for year_first do note that we currently don't
  # correct the spatial sampling unit calendar for 'year_first mismatches'
  # between modules in the same module combo. See the earlier object
  # 'mod_scheme_ps_different_1styr' and associated missing functionality!
  summarize(
    year_first = min(year_first),
    year_last = max(year_last),
    .by = c(scheme, module_combo_code, panel_split)
  ) %>% 
  # Add revisit designs by FAS x spatial subset. The origins of
  # 'scheme_moco_ps_spsubset_fas_paneldesign' lie in objects
  # 'mod_scheme_variableset_mainfieldmethod' and 'activity_sequence', which
  # don't distinguish between schemes that got a sample size or not. So we
  # expect some cases to be dropped.
  inner_join(
    scheme_moco_ps_spsubset_fas_paneldesign,
    join_by(scheme, module_combo_code, panel_split),
    relationship = "one-to-many",
    unmatched = c("error", "drop")
  ) %>%
  # integrating main panel design and subpanel design to a single panel design
  mutate(
    panel_count = case_when(
      revisit_pattern == "pure_panel" ~ 1L,
      .default = as.integer(cycle_duration_y * years(1) / panel_period_main)
    ) *
      case_when(
        is.na(revisit_subpattern) ~ 1L,
        revisit_subpattern == "pure_panel" ~ 1L,
        !is.na(limit_n_subpanels_per_panel) ~ limit_n_subpanels_per_panel,
        .default = as.integer(panel_period_main / panel_subperiod)
      ),
    panel_period = case_when(
      is.na(panel_subperiod) ~ panel_period_main,
      .default = panel_subperiod
    ),
    # number of panels in a panel group, to be taken together in repetitions
    # (below)
    panel_group_size = ifelse(
      !is.na(limit_n_subpanels_per_panel),
      limit_n_subpanels_per_panel,
      1L
    ),
    # number of times a panel group must be repeated during a cycle
    panel_group_repetitions = case_when(
      !is.na(limit_n_subpanels_per_panel) ~ as.integer(
        (panel_period_main / panel_subperiod) / limit_n_subpanels_per_panel
      ),
      revisit_subpattern == "pure_panel" ~ pmin(
        as.integer(panel_period_main / panel_subperiod),
        limit_n_subperiods,
        na.rm = TRUE
      ),
      .default = 1L
    ),
    n_subperiods_per_main_period = ifelse(
      is.na(panel_subperiod),
      1L,
      as.integer(panel_period_main / panel_subperiod)
    ),
    panel_period_main_chr = map_chr(panel_period_main, simplify_period),
    panel_period_chr = map_chr(panel_period, simplify_period),
    panel_enddate_addition_chr = map_chr(panel_enddate_addition, simplify_period),
    panel_dateshift_chr = map_chr(panel_dateshift, simplify_period),
    revisit_pattern_abbrev = str_sub(revisit_pattern, 1, 3) %>% str_to_upper(),
    revisit_subpattern_abbrev = ifelse(
      !is.na(revisit_subpattern), 
      str_sub(revisit_subpattern, 1, 3), 
      ""
    ) %>% 
      str_to_upper(),
    # notation_pattern summarises the panel design
    notation_pattern = case_when(
      revisit_pattern_abbrev == "SER" & 
        revisit_subpattern_abbrev %in% c("SER", "PUR") ~
        str_c(
          revisit_pattern_abbrev, 
          panel_period_main_chr, 
          "-",
          revisit_subpattern_abbrev,
          ifelse(
            panel_group_size > 1L | panel_group_repetitions > 1L, 
            str_c("by", panel_group_size, "repeat", panel_group_repetitions),
            ""
          )
        ),
      revisit_pattern_abbrev == "PUR" & revisit_subpattern_abbrev == "SER" ~
        "SER",
      .default = str_c(
        revisit_pattern_abbrev, 
        ifelse(
          revisit_subpattern_abbrev != "",
          str_c("-", revisit_subpattern_abbrev),
          ""
        )
      )
    ),
    # notation_paneldesign combines notation_pattern & panel design attributes
    notation_paneldesign = str_c(
      panel_count,
      ifelse(panel_count == 1L, "panelof","panelsof"),
      ifelse(
        is.na(panel_enddate_addition),
        panel_period_chr,
        str_c(panel_period_chr,"[",  "+", panel_enddate_addition_chr, "]")
      ),
      ifelse(
        str_detect(notation_pattern, "-"),
        (str_c("for", cycle_duration_y, "y")),
        ""
      ),
      "(",
      ifelse(auto_continuous_meas, "auto|", ""),
      ifelse(memship_alignfastest, "fastalign|", ""),
      notation_pattern,
      ")",
      ifelse(
        !is.na(panel_dateshift), 
        str_c("[shift(", panel_dateshift_chr, ")]"),
        ""
      ),
      ifelse(
        !is.na(limit_n_subperiods), 
        str_c(limit_n_subperiods, "subperiod"),
        ""
      )
    ) %>% 
      fct()
    # PURE PANELS (main or sub) do not meet following check:
    # check = panel_count * panel_period * panel_group_repetitions ==
    #   cycle_duration_y * years(1)
  )
# scheme_moco_ps_spsubset_fas_panelcount_full %>% distinct(notation_paneldesign)
# scheme_moco_ps_spsubset_fas_panelcount_full %>% filter(!check)
scheme_moco_ps_spsubset_fas_panelcount <- 
  scheme_moco_ps_spsubset_fas_panelcount_full %>% 
  select(
    scheme,
    module_combo_code,
    panel_split,
    in_aquatic_subset,
    fas,
    starts_with("year"),
    panel_period,
    panel_enddate_addition,
    panel_dateshift,
    panel_count,
    n_subperiods_per_main_period,
    limit_n_subperiods,
    panel_group_size,
    panel_group_repetitions,
    auto_continuous_meas,
    memship_alignfastest,
    notation_paneldesign
  )
scheme_moco_ps_spsubset_fas_panelcount %>%
  glimpse()
```

Performing several checks on the result.

```{r}
# The notation_paneldesign column must effectively distinguish all occurring
# different revisit designs
stopifnot(
  scheme_moco_ps_spsubset_fas_panelcount_full %>% 
    # for pure panels the cycle_duration_y does not determine the number of
    # panels, hence is not part of the panel design notation
    mutate(
      cycle_duration_y = ifelse(
        revisit_pattern == "pure_panel", 
        NA, 
        cycle_duration_y)
    ) %>% 
    distinct(pick(revisit_pattern:last_col(), -enforced_cycle_duration)) %>% 
    nrow() == 
    n_distinct(scheme_moco_ps_spsubset_fas_panelcount$notation_paneldesign)
)
# The rows of scheme_moco_ps_spsubset_fas_panelcount are still unique if you
# drop fas and notation* columns
stopifnot(
  scheme_moco_ps_spsubset_fas_panelcount %>% 
    distinct(pick(-fas, -starts_with("notation"))) %>% 
    nrow() == nrow(scheme_moco_ps_spsubset_fas_panelcount)
)
# The uniqueness of the rows of scheme_moco_ps_spsubset_fas_panelcount is
# completely captured by scheme x module_combo_code x in_aquatic_subset x 
# notation_paneldesign
stopifnot(
  scheme_moco_ps_spsubset_fas_panelcount %>% 
    distinct(
      scheme,
      module_combo_code,
      in_aquatic_subset,
      notation_paneldesign
    ) %>% 
    nrow() == nrow(scheme_moco_ps_spsubset_fas_panelcount)
)
```

Note that the uniqueness of the rows of `scheme_moco_ps_spsubset_fas_panelcount` is completely captured by scheme x module combo x spatial subset x `notation_paneldesign`.

Different occurring patterns at the compartment level, with 'HQ' referring to the MHQ 'compartment', for information and thought:

```{r}
compartment_paneldesign_fags <- 
  scheme_moco_ps_spsubset_fas_panelcount %>%
  mutate(
    compartment = str_extract(scheme, "^[A-Z]+") %>% 
      factor() %>% 
      fct_relevel("SOIL", "HQ", after = Inf),
    fags = map(fas, ~.$field_activity_group),
    # actseqs_vec = map(fas, ~unique(unlist(.$actseqs_vec)))
  ) %>% 
  mutate(
    scheme_moco = str_c(
      scheme,
      ": ",
      str_c(sort(unique(module_combo_code)), collapse = ", ")
    ),
    .by = c(scheme, notation_paneldesign)
  ) %>% 
  summarize(
    fags = str_c(sort(unique(unlist(fags))), collapse = ", "),
    # actseqs = str_c(sort(unique(unlist(actseqs_vec))), collapse = ", "),
    scheme_moco = str_c(sort(unique(scheme_moco)), collapse = " | "),
    .by = c(compartment, notation_paneldesign)
  ) %>%
  arrange(compartment, notation_paneldesign)
compartment_paneldesign_fags %>%
  kable()
```

### Panels

Defining panels.

```{r}
scheme_moco_ps_spsubset_fas_panels <-
  scheme_moco_ps_spsubset_fas_panelcount %>%
  # spatial_subset_abbr is a temporary string used in the panel labels
  mutate(
    spatial_subset_abbr = if_else(in_aquatic_subset, "AQ", "TERR", ""),
    scheme_moco_has_multiple_panel_splits = n_distinct(panel_split) > 1,
    .by = c(scheme, module_combo_code)
  ) %>% 
  uncount(panel_count) %>%
  # 'panel' is essentially a label of an individual panel
  mutate(
    panel = str_c(
      if_else(
        spatial_subset_abbr != "", 
        str_c(scheme, "_", spatial_subset_abbr), 
        scheme
      ),
      ifelse(scheme_moco_has_multiple_panel_splits, 
             str_c(
               str_remove_all(module_combo_code, "_"), 
               "_", 
               "ps", 
               panel_split
             ), 
             str_remove_all(module_combo_code, "_")
      ),
      notation_paneldesign,
      str_c("PANEL", str_pad(seq_len(n()), 2, pad = "0")),
      sep = "_"
    ),
    .by = c(
      scheme, 
      module_combo_code, 
      panel_split, 
      in_aquatic_subset,
      notation_paneldesign
    )
  ) %>%
  mutate(panel = factor(panel)) %>% 
  select(-spatial_subset_abbr, -scheme_moco_has_multiple_panel_splits)
```

### Panel calendar

Defining date intervals and assigning panels to these intervals.

```{r}
scheme_moco_ps_spsubset_panel_calendar <-
  scheme_moco_ps_spsubset_fas_panels %>% 
  # year_first_overall will be used to build up the revisit design (incl the
  # membership design) in order to maximize chances for matching locations
  # between schemes x module combo
  mutate(year_first_overall = min(year_first)) %>% 
  nest(
    revisit = c(
      matches("^year|^limit_"), 
      n_subperiods_per_main_period,
      panel_period,
      panel_enddate_addition,
      panel_dateshift,
      panel_group_size, 
      panel_group_repetitions, 
      auto_continuous_meas,
      memship_alignfastest,
      notation_paneldesign
    ),
    panels = panel
  ) %>%
  # define date intervals from year_first_overall to year_last, by panel_period
  mutate(
    revisit = map(revisit, function(df) {
      df %>%
        distinct() %>%
        mutate(
          # we always define panels from the first (overall) year and the first
          # month; if a module or scheme starts later, then this part is dropped
          # from the first cycle
          nr_periods =
            years(year_last - year_first_overall + 1) / panel_period
        ) %>%
        uncount(nr_periods) %>%
        mutate(
          panel_periodmonths = as.period(panel_period, "months") %>%
            month() %>%
            as.integer(),
          panel_cum_months = cumsum(panel_periodmonths),
          panel_cum_months_lag = lag(panel_cum_months, default = 0L),
          date_start = make_date(year_first_overall, 1, 1) + 
            months(panel_cum_months_lag) +
            if_else(!is.na(panel_dateshift), panel_dateshift, months(0)),
          date_end = make_date(year_first_overall, 1, 1) + 
            months(panel_cum_months) +
            if_else(
              !is.na(panel_enddate_addition),
              panel_enddate_addition,
              months(0)
            ) +
            if_else(!is.na(panel_dateshift), panel_dateshift, months(0)) - 
            days(1),
          date_interval = interval(
            force_tz(date_start, "Europe/Brussels"), 
            force_tz(date_end, "Europe/Brussels")
          )
        ) %>%
        select(-c(panel_periodmonths, starts_with("panel_cum_months")))
    }),
    # assign panels to the date intervals, taking into account the revisit
    # design
    panel_date_assignment = map2(revisit, panels, function(revis, pan) {
      group_size <- revis$panel_group_size[1]
      group_repetitions <- revis$panel_group_repetitions[1]
      n_subperiods <- revis$n_subperiods_per_main_period[1]
      n_subperiods_limit <- revis$limit_n_subperiods[1]
      # set the panel order of one cycle according to group_size &
      # group_repetitions
      if (group_size == 1 && group_repetitions == 1) {
        pan2 <- pan
      } else {
        indexes_pan <- seq_len(nrow(pan))
        indexgrouping_pan <- rep(seq_len(nrow(pan) / group_size), each = group_size)
        indexes_grouped <- split(indexes_pan, indexgrouping_pan) %>%
          rep(each = group_repetitions) %>%
          unlist() %>%
          unname()
        pan2 <- pan[indexes_grouped, ]
      }
      # subset revis periodically to meet the n_subperiods_limit requirement
      if (is.na(n_subperiods_limit)) {
        revis2 <- revis
      } else {
        indexes_limitsubperiod <- 
          seq_len(nrow(revis)) %>% 
          # split the periods according to the number of subperiods in one main
          # period of the cycle
          split(rep(
            seq_len(ceiling(nrow(revis) / (n_subperiods))), 
            each = n_subperiods
          )) %>% 
          map_int(~ .[n_subperiods_limit]) %>% 
          unname()
        revis2 <- revis[indexes_limitsubperiod, ]
      }
      # determine indexes to recycle rows of pan2
      indexes <- rep(seq_len(nrow(pan2)), nrow(revis2) %/% nrow(pan2) + 1) %>%
        head(nrow(revis2))
      # combine the dates in revis2 with the panels in (recycled) pan2
      bind_cols(
        revis2 %>% select(date_start, date_end, date_interval),
        pan2[indexes, ]
      )
    })
  ) %>%
  select(-fas, -panels, -revisit) %>%
  unnest(panel_date_assignment) %>%
  relocate(panel, .after = in_aquatic_subset)
```

**Missing functionality**: ability to undo the dateshift of a backward shifted panel design in the first panel of the first cycle only, in combination with setting this panel's enddate earlier (i.e. making the panel duration shorter).
Not clear yet whether we'll need this, but it's relevant when there's a need to speed up implementation.


```{r eval=FALSE, echo=FALSE}
# For calendar inspection only
scheme_moco_ps_spsubset_panel_calendar %>% 
  inner_join(
    scheme_moco_ps_spsubset_fas_panels %>% 
      select(scheme:fas, panel),
    join_by(scheme, module_combo_code, panel_split, in_aquatic_subset, panel),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  mutate(
    fags = map_chr(fas, ~ str_c(.$field_activity_group, collapse = ", "))
  ) %>% 
  relocate(fags, .after = in_aquatic_subset) %>% 
  relocate(fas, .after = last_col()) %>% 
  View("calendar")
```

### Panel membership of spatial sampling units

Panel membership is randomized by following the GRTS address sequence at the stratum level.

To determine the number of sampling units per panel, it must be taken into account that the stratum sample size is not equal to, or a multiple of, the number of panels, so panels usually cannot have an equal number of units.
Local pivotal sampling of panels is applied to get an even load of all panels while maintaining maximum temporal spread in the membership.
One wants to avoid a systematic effect that the first panels always have more units than those further in line, or the last panel.
This becomes even more important for strata with few sampling units; sometimes this number is smaller than the theoretical number of panels!

First, we generate and populate **generic panels** for each scheme x module combo x panel split x stratum x spatial poststratum.
A generic panel only takes into account the number of panels of a revisit design, not the specific revisit design.
We use generic panel sets to allow for maximum synergy among different field activity groups with different revisit designs.
To achieve this, the panel memberships for panel designs with different panel counts are made interdependent: e.g. if both 3-panel-designs and 12-panel-designs occur, then membership design is first done for the 12-panel-designs, and the panel members are then lumped into 3 consecutive groups to define the membership for the 3-panel-designs.
This avoids repeating local pivotal sampling more than necessary, which would lead to unneeded mismatches.

Next, this is translated to the actual panels created before.

#### Membership of generic panels

Generating and populating generic panels.

```{r}
set.seed(20241206)
scheme_moco_ps_stratum_sppost_genericpanelmemship <- 
  scheme_moco_ps_stratum_sppost_spsamples %>% 
  unnest(sp_poststr_samples) %>% 
  select(-sample_status) %>% 
  nest(spsample = grts_address) %>% 
  inner_join(
    scheme_moco_ps_spsubset_fas_panelcount %>% 
      distinct(
        scheme, 
        module_combo_code, 
        panel_split, 
        panel_group_repetitions,
        auto_continuous_meas, 
        memship_alignfastest,
        panel_count
      ) %>% 
      # guessing the grouping pattern to use in the revisit panel design for
      # FAGs with automated measurements
      mutate(
        max_panel_group_repetitions = max(panel_group_repetitions),
        .by = c(scheme, module_combo_code, panel_split),
        .keep = "unused"
      ) %>% 
      distinct() %>% 
      # using auto_continuous_meas and memship_alignfastest as attributes next
      # to panel count, since they trigger a different panel membership approach
      nest(panel_counts = c(
        panel_count, 
        auto_continuous_meas,
        memship_alignfastest
      )),
    join_by(scheme, module_combo_code, panel_split),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  mutate(
    # generate data frame with the relation between generic panel numbers of
    # different panel setup (panel count & automated measurements) in a scheme x
    # moco x panel split. The rows of this data frame will be used for
    # partitioning spatial samples in GRTS order.
    generic_panels = map2(panel_counts, max_panel_group_repetitions, function(pc, mpgr) {
      pcs <- pc$panel_count
      max_panel_count <- max(pcs)
      # currently only panel counts are supported that are a divisor of the maximum
      divisors_cond <- all(pcs %in% numbers::divisors(max_panel_count))
      if (!divisors_cond) {
        stop("A number of panels occurs that is not a divisor of ", max_panel_count)
      }
      mat <- matrix(0L, nrow = max_panel_count, ncol = length(pcs))
      pclabels <- str_c(
        pc$panel_count,
        case_when(
          pc$auto_continuous_meas ~ "auto", 
          pc$memship_alignfastest ~ "fastalign",
          .default = ""
        )
      )
      colnames(mat) <- str_c("generic", pclabels, ifelse(pcs == 1, "panel", "panels"))
      df <- as_tibble(mat)
      for (i in seq_along(pcs)) {
        nr <- pcs[i]
        auto <- pc$auto_continuous_meas[i]
        fastalign <- pc$memship_alignfastest[i]
        df[, i] <- if (auto) {
          # automated measurements get a special pattern
          rep(seq_len(nr), each = mpgr) %>% 
            rep(max_panel_count / (nr * mpgr))
        } else if (fastalign) {
          # fast-aligned memberships
          rep(seq_len(nr), max_panel_count / nr)
        } else {
          # usual FAGs get a monotonically increasing panel number attribution
          rep(seq_len(nr), each = max_panel_count / nr)
        }
      }
      df %>% 
        mutate(id = row_number()) %>% 
        relocate(id)
    }),
    # distribute the spatial samples over the rows of foregoing data frame,
    # and rearrange the data so it will be possible to join with 
    # scheme_moco_ps_spsubset_fas_panels
    panel_membership = map2(spsample, generic_panels, function(sps, gen) {
      distribute_sample_over_panels(sps, gen) %>% 
        inner_join(
          gen,
          join_by(genericpanels_row == id),
          relationship = "many-to-one",
          # "drop" is relevant when nrow(sps) < nrow(gen)
          unmatched = c("error", "drop")
        ) %>% 
        select(-genericpanels_row) %>% 
        pivot_longer(
          cols = !grts_address,
          names_to = "panel_count",
          values_to = "panel_id",
          cols_vary = "slowest"
        ) %>% 
        mutate(
          auto_continuous_meas = str_detect(panel_count, "auto"),
          memship_alignfastest = str_detect(panel_count, "fastalign"),
          panel_count = str_extract(panel_count, "\\d+") %>% as.integer()
        ) %>% 
        relocate(auto_continuous_meas, memship_alignfastest, .after = panel_count)
    })
  ) %>% 
  select(
    scheme,
    module_combo_code,
    panel_split,
    stratum,
    sp_poststratum,
    panel_membership
  ) %>% 
  # we still need to handle 7220 better here!
  add_col_in_aquatic_subset(
    strata = n2khab_strata,
    type_properties = n2khab_types_expanded_properties
  ) %>% 
  relocate(in_aquatic_subset, .after = panel_split) %>% 
  arrange(scheme, module_combo_code, panel_split, in_aquatic_subset)
```

**Missing functionality**: split the 7220 locations in an aquatic and terrestrial spatial subset (which overlap).
This actually boils down to distinguishing rivulets and mires as two different strata within type 7220, for which independent sampling must take place.
Note that these 7220 strata also need special handling with respect to the target population of `GW_05.1_{aq,terr}` and of `SURF_*` schemes (much like the code to generate `mod_scheme_mainfieldmethod`).
Currently we just duplicate the samples for these two spatial subsets...

**Missing functionality**: for the rarer stratum x spatial poststratum combinations, exact matches of the spatial sample occur _across schemes_.
For these strata, we better share this distribution over the panels across schemes for efficiency reasons (matching field activities, reduced number of location visits).

#### Panel membership

Translating the generic panel memberships to panel memberships for each FAS (hence revisit design).

```{r}
scheme_moco_ps_spsubset_fas_stratum_sppost_panelmemship <- 
  scheme_moco_ps_spsubset_fas_panels %>% 
  select(
    scheme, 
    module_combo_code, 
    panel_split, 
    in_aquatic_subset, 
    fas, 
    auto_continuous_meas,
    memship_alignfastest,
    notation_paneldesign, 
    panel
  ) %>% 
  nest(panels = panel) %>% 
  mutate(
    panels = map(panels, ~ mutate(., panel_id = row_number())),
    panel_count = map_int(panels, nrow)
  ) %>% 
  # join the generic panel memberships to the panels based on panel count, on
  # whether the panel design is for automated continuous measurements, and on
  # whether the panel membership uses fast alignment:
  inner_join(
    scheme_moco_ps_stratum_sppost_genericpanelmemship %>% 
      rename(stratum_in_aquatic_subset = in_aquatic_subset) %>% 
      unnest(panel_membership) %>% 
      nest(panel_membership = c(grts_address, panel_id)),
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      panel_count,
      auto_continuous_meas,
      memship_alignfastest
    ),
    relationship = "many-to-many",
    unmatched = "error"
  ) %>% 
  # where a FAS applies to a spatial subset, the corresponding strata must be
  # limited as such:
  filter(
    in_aquatic_subset == stratum_in_aquatic_subset | is.na(in_aquatic_subset)
  ) %>% 
  select(
    -panel_count,
    -auto_continuous_meas,
    -memship_alignfastest,
    -stratum_in_aquatic_subset
  ) %>% 
  # XXXXXXXXXXXX
  # following distinct() only matters for 7220, as long as it isn't divided into
  # strata (the sample duplication will only remain for FASs that distinguish
  # spatial subsets):
  distinct() %>% 
  mutate(
    panel_membership = map2(panels, panel_membership, function(pan, mem) {
      pan %>% 
        inner_join(
          mem,
          join_by(panel_id),
          relationship = "one-to-many",
          unmatched = c("drop", "error")
        ) %>% 
        select(-panel_id)
    })
  ) %>% 
  select(-panels)
```

Note that (in a scheme x panel split x module combo) only the combination of stratum x spatial poststratum stands for a consecutive GRTS series, equally distributed over panels, and that replacement with spare units of the stratum must happen within spatial poststrata.

## Field activities & spatiotemporal samples

We will join panel memberships with the date intervals per panel to generate the spatiotemporal samples.
At the same time, each FAS is replaced by its constituting FAGs (field activity groups), which have a common revisit design.

### Calendar per FAG, stratum and spatial sampling unit

This calendar still distinguishes between scheme x module combo x panel split.

##### Initial spatial sampling unit calendar

Generating a sampling unit calendar, i.e. with timing per FAG, stratum and spatial sampling unit in a scheme x module combo x panel split x spatial poststratum.
Note that the spatial subset is a mere attribute, as is already expressed by stratum and also FAG (in case of 7220).

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar_initial <- 
  scheme_moco_ps_spsubset_fas_stratum_sppost_panelmemship %>% 
  unnest(panel_membership) %>% 
  unnest(fas) %>% 
  inner_join(
    scheme_moco_ps_spsubset_panel_calendar,
    join_by(scheme, module_combo_code, panel_split, in_aquatic_subset, panel),
    relationship = "many-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  relocate(
    sp_poststratum,
    panel,
    starts_with("date"),
    .after = notation_paneldesign
  ) %>% 
  arrange(
    scheme,
    module_combo_code,
    panel_split,
    in_aquatic_subset,
    actseqs,
    rank,
    field_activity_group,
    notation_paneldesign,
    sp_poststratum,
    date_start,
    panel,
    stratum,
    grts_address
  )
```

#### Checking the initial spatial sampling unit calendar

Checking that all calendar rows match a row of `scheme_moco_ps_spsubset_fas_paneldesign`.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar_initial %>% 
  distinct(pick(
    -c(panel, starts_with("date"), grts_address, stratum, sp_poststratum)
  )) %>% 
  inner_join(
    scheme_moco_ps_spsubset_fas_paneldesign %>% 
      unnest(fas),
    join_by(
      scheme, 
      module_combo_code, 
      panel_split, 
      in_aquatic_subset,
      field_activity_group,
      rank,
      max_rank,
      actseqs,
      actseqs_vec
    ),
    relationship = "one-to-one",
    # The origins of 'scheme_moco_ps_spsubset_fas_paneldesign' lie in objects
    # 'mod_scheme_variableset_mainfieldmethod' and 'activity_sequence', which
    # don't distinguish between schemes that got a sample size or not. So we
    # expect some cases to be dropped.
    unmatched = c("error", "drop")
  ) %>% 
  invisible()
```

Checking that all panel designs x FAGs x spatial subsets are present as expected (using `scheme_moco_ps_spsubset_fas_panelcount`), in each scheme x module combo x panel split.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar_initial %>% 
  distinct(pick(
    -c(panel, starts_with("date"), grts_address, stratum, sp_poststratum)
  )) %>% 
  inner_join(
    scheme_moco_ps_spsubset_fas_panelcount %>% unnest(fas),
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset,
      notation_paneldesign,
      field_activity_group,
      rank,
      max_rank,
      actseqs,
      actseqs_vec
    ),
    relationship = "one-to-one",
    unmatched = "error"
  ) %>% 
  invisible()
```

Note that it's irrelevant to verify the number of panels, since smaller sample sizes have led to dropping panels.

Since quite a few FAGs are applied at the same location, it is useful to derive the subset of FAGs ('target FAGs') that represent the main field method of the target variable(s) in each scheme, in order to use this for further selections.

```{r}
scheme_moco_ps_spsubset_targetfag <- 
  mod_scheme_targetfieldmethod %>% 
  # joining FAGs
  inner_join(
    mod_scheme_field_activity %>% 
      distinct(
        module,
        scheme,
        in_aquatic_subset,
        field_activity,
        field_activity_group
      ),
    join_by(module, scheme, main_field_method == field_activity),
    # in_aquatic_subset gives duplication:
    relationship = "one-to-many",
    unmatched = c("error", "drop")
  ) %>% 
  distinct(module, scheme, in_aquatic_subset, field_activity_group) %>% 
  rename(in_aq = in_aquatic_subset) %>% 
  # joining simplified values of 'in_aquatic_subset'
  inner_join(
    mod_scheme_fag,
    join_by(module, scheme, field_activity_group),
    relationship = "many-to-one",
    unmatched = c("error", "drop")
  ) %>% 
  # following filter appears to be superfluous:
  filter(in_aq == in_aquatic_subset | is.na(in_aquatic_subset)) %>% 
  select(module, scheme, in_aquatic_subset, field_activity_group) %>% 
  # join module combos:
  inner_join(
    scheme_target_paneldesign %>% 
      select(scheme, module_combo_code, module_combo, panel_split) %>% 
      unnest(module_combo),
    join_by(module, scheme),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  distinct(
    scheme, 
    module_combo_code, 
    panel_split, 
    in_aquatic_subset, 
    field_activity_group
  )
```

Checking the stratum spatial sample sizes for _target_ FAGs in the calendar against the spatial sample sizes from `scheme_moco_ps_stratum_sppost_spsamples`, for each scheme x module combo x panel split.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar_initial %>% 
  distinct(
    scheme,
    module_combo_code,
    panel_split,
    in_aquatic_subset,
    field_activity_group,
    stratum,
    grts_address
  ) %>% 
  semi_join(
    scheme_moco_ps_spsubset_targetfag,
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset, # this one doesn't matter actually
      field_activity_group
    )
  ) %>% 
  count(
    scheme,
    module_combo_code,
    panel_split,
    field_activity_group,
    stratum,
    name = "sample_size_fag"
  ) %>% 
  inner_join(
    scheme_moco_ps_stratum_sppost_spsamples %>% 
      mutate(
        total_sample_size = map_int(sp_poststr_samples, nrow),
        .keep = "unused"
      ),
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      stratum,
      sample_size_fag == total_sample_size
    ),
    # lhs has duplication due to > 1 target FAG per scheme in some cases
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  invisible()
```

<!-- check temporal spread of sample size -->

#### Deriving spatial sampling unit calendar {#cal-cleaning}

Cleaning is required with regard to the startup period:

- wind up panels outside the period considered to implement schemes in specific modules.
  - In this process, take care to actually _postpone_ certain panel visits X FAG combinations within a year if these represent a FAG needed to execute other FAGs happening in the same year within the considered period of implementation.
  This should also take into account (positive) timeshift properties of certain panel designs (i.e. needed delays).
    - This will typically be applied to panel designs where visits are limited to just one subperiod in the main period; the subperiod then needs to be postponed in the year in question.
    - We also apply this to INST & SPATPOSIT FAGs, even though this puts much more work into the period where this is moved into.
    In practice, it is expected that this work can be organized during a somewhat longer period than the one in the calendar.
    This may need reconsideration depending on evolving circumstances.
  - Also, for now neglect LOCEVAL FAGs as they will need more special consideration, e.g. they may need to be postponed a whole year, but only in the first year.
  - Note that we currently don't correct '`year_first` mismatches' between modules in the same module combo (see `mod_scheme_ps_different_1styr` object).
It requires calculation of module-level spatial sample sizes in poststrata.
See the _missing functionality_ in section \@ref(moco).
- SURFINSTGAUGE and SPATPOSITGAUGE activities in groundwater schemes (3-year periodicity) must require the presence of a corresponding GWINST activity (6-year periodicity) in the same or a preceding period in order to be relevant.
This will drop about half of the SURFINSTGAUGE activities in groundwater schemes during the first 3-year cycle.
- drop FAGs that miss their foregoing FAGs with a lower rank.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar <-
  scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar_initial %>% 
  # adding some properties of the revisit design
  inner_join(
    scheme_moco_ps_spsubset_fas_panelcount %>% 
      select(
        scheme, 
        module_combo_code, 
        panel_split, 
        in_aquatic_subset, 
        year_first,
        year_last,
        panel_enddate_addition,
        panel_dateshift,
        limit_n_subperiods,
        notation_paneldesign,
        memship_alignfastest
      ),
    join_by(scheme, module_combo_code, panel_split, in_aquatic_subset, notation_paneldesign),
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  # don't include inapplicable years
  filter(
    date_start >= make_date(year_first, 1, 1) + 
      if_else(!is.na(panel_dateshift), panel_dateshift, months(0)),
    date_end <= make_date(year_last + 1, 1, 1) + 
      if_else(
        !is.na(panel_enddate_addition),
        panel_enddate_addition,
        months(0)
      ) +
      if_else(!is.na(panel_dateshift), panel_dateshift, months(0)) - 
      days(1)
  ) %>% 
  # implement a delay for FAGs with a panel design with one subperiod and for
  # FAGs where permanent equipment is installed and spatially positioned, if
  # month(date_start) < month_first_default in the first year
  mutate(
    delay = if_else(
      (limit_n_subperiods == 1 | 
         str_detect(field_activity_group, "INST|SPATPOSIT")) & 
        year(date_start) == year_first & 
        month(date_start) < month_first_default,
      months(month_first_default - month(date_start)),
      months(0),
      # limit_n_subperiods can be NA, hence condition can be NA
      months(0)
    ),
    date_start = if_else(delay > months(0), date_start + delay, date_start),
    date_end = if_else(
      delay > months(0),
      date_end + days(1) + delay - days(1),
      date_end
    ),
    date_interval = if_else(
      delay > months(0),
      interval(
        force_tz(date_start, "Europe/Brussels"), 
        force_tz(date_end, "Europe/Brussels")
      ),
      date_interval
    )
  ) %>% 
  select(-delay) %>% 
  # drop inapplicable panels in the startup period based on month_first_default,
  # except for LOCEVAL
  filter(
    month(date_end) >= month_first_default | 
      year(date_end) > year_first |
      str_detect(field_activity_group, "LOCEVAL")
  ) %>% 
  # if a panel period overlaps month_first_default but starts earlier, then
  # wind up the start date, but not for LOCEVAL and HQ schemes
  mutate(
    delay = if_else(
      month(date_start) < month_first_default & 
        year(date_start) == year_first &
        !str_detect(field_activity_group, "LOCEVAL") &
        !str_detect(scheme, "^HQ"),
      months(month_first_default - month(date_start)),
      months(0)
    ),
    date_start = if_else(delay > months(0), date_start + delay, date_start),
    date_interval = if_else(
      delay > months(0),
      interval(
        force_tz(date_start, "Europe/Brussels"), 
        force_tz(date_end, "Europe/Brussels")
      ),
      date_interval
    )
  ) %>% 
  select(-delay) %>% 
  # drop SURFINSTGAUGE & SPATPOSITGAUGE activities in groundwater schemes that
  # take place earlier than the first corresponding GWINST FAG. Below code
  # assumes that the cases under consideration are represented by the
  # memship_alignfastest variable. Note that we can't use the FAG rank to
  # accomplish this, since SURFINSTGAUGE & GWINST share the same rank (i.e. they
  # are preferably carried out in the same date interval).
  mutate(
    first_date_gwinst = min(
      date_start[str_detect(field_activity_group, "GWINST")],
      ymd("21000101"),
      na.rm = TRUE
    ),
    .by = c(
      scheme,
      module_combo_code,
      panel_split,
      stratum,
      sp_poststratum,
      grts_address
    )
  ) %>%
  filter(
    (memship_alignfastest & date_start >= first_date_gwinst) |
      !memship_alignfastest
  ) %>% 
  select(!c(
    first_date_gwinst,
    memship_alignfastest,
    panel_enddate_addition,
    panel_dateshift,
    limit_n_subperiods
  ))
```

Generating a temporary, long variant of this intermediate sampling unit calendar result, with a column `activity_sequence`.

```{r}
spsamples_calendar_actseqs_long <- 
  scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar %>% 
  select(-actseqs) %>% 
  unnest_longer(actseqs_vec, values_to = "activity_sequence")
```

Updating the sampling unit calendar by dropping FAGs (and corresponding activity sequence) that miss their preceding FAGs on location-basis (and stratum-basis, since we still have a sampling frame with many coinciding strata on location-basis while this isn't the reality at least for terrestrial strata).
This condition is considered across schemes, module combos, panel splits and panels, but within activity sequence (but see code comment).

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar <- 
  spsamples_calendar_actseqs_long %>% 
  distinct(
    stratum,
    field_activity_group, 
    rank,
    max_rank, 
    activity_sequence,
    grts_address,
    pick(starts_with("date"))
  ) %>% 
  mutate(
    date_start_first_rank1 = if (length(date_start[rank == 1L]) == 0L) {
      ymd("21000101")
    } else {
      min(date_start[rank == 1L])
    },
    date_start_first_rank2 = if (length(date_start[rank == 2L]) == 0L) {
      ymd("21000101")
    } else {
      min(date_start[rank == 2L])
    },
    date_start_first_rank3 = if (length(date_start[rank == 3L]) == 0L) {
      ymd("21000101")
    } else {
      min(date_start[rank == 3L])
    },
    date_start_first_rank4 = if (length(date_start[rank == 4L]) == 0L) {
      ymd("21000101")
    } else {
      min(date_start[rank == 4L])
    },
    # A tricky thing below is that we again look at the activity_sequence level,
    # which would be logical, but we already lumped activity sequences per FAG
    # when making mod_scheme_fag (on which all the rest is based). This is
    # intensified by lumping in_aquatic_subset at the FAG level, not at the FAG
    # x activity_sequence level, when creating mod_scheme_actseq_fag, on which
    # mod_scheme_fag is based. This activity sequence agnostic approach works as
    # long as FAG ranks are the same across activity sequences, which we already
    # tested above. When this is no longer the case, we should not do the
    # lumping in mod_scheme_actseq_fag and mod_scheme_fag. A good example to
    # look at is FAG 'GWINSTPIEZWELL' in GW_03.3, which happens in 4 activity
    # sequences, 2 terrestrial and 2 aquatic (depending on which modules are
    # active).
    .by = c(stratum, activity_sequence, grts_address)
  ) %>% 
  filter(
    rank == 1L |
      (rank == 2L & 
         date_end > date_start_first_rank1 ) |
      (rank == 3L & 
         date_end > date_start_first_rank1 & 
         date_end > date_start_first_rank2) |
      (rank == 4L & 
         date_end > date_start_first_rank1 & 
         date_end > date_start_first_rank2 & 
         date_end > date_start_first_rank3) |
      (rank == 5L & 
         date_end > date_start_first_rank1 & 
         date_end > date_start_first_rank2 & 
         date_end > date_start_first_rank3 & 
         date_end > date_start_first_rank4),
    # similar comment about activity_sequence as above
    .by = c(stratum, activity_sequence, grts_address)
  ) %>% 
  semi_join(
    spsamples_calendar_actseqs_long,
    .,
    join_by(
      stratum,
      field_activity_group, 
      rank,
      max_rank,
      activity_sequence,
      grts_address,
      date_start,
      date_end
    )
  ) %>% 
  select(-year_first, -year_last) %>% 
  nest(actseqs = activity_sequence) %>% 
  # since the new FAG sets may be smaller than the original FA supergroups, due
  # to dropping FAGs while cleaning, we just call these sets 'fags' so that
  # 'fas' can keep its time-independent but revisit-design
  # (notation_paneldesign) dependent meaning
  nest(fags = c(field_activity_group, rank, max_rank, actseqs)) %>% 
  relocate(fags, .after = in_aquatic_subset) %>% 
  relocate(
    sp_poststratum,
    panel,
    starts_with("date"),
    .after = notation_paneldesign
  ) %>% 
  arrange(
    scheme,
    module_combo_code,
    panel_split,
    in_aquatic_subset,
    notation_paneldesign,
    sp_poststratum,
    date_start,
    panel,
    stratum,
    grts_address
  )
```

```{r include=FALSE}
rm(spsamples_calendar_actseqs_long)
gc()
```

#### Checking the spatial sampling unit calendar

<!-- - check consistency (join completeness) with earlier objects -->

```{r include=FALSE}
# to check if this runs without error
scheme_moco_ps_spsubset_fas_panels %>% 
  inner_join(
    scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar,
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset,
      notation_paneldesign,
      panel
    ),
    relationship = "one-to-many",
    # "drop" is allowed because some panels disappear due to the cleaning or
    # to small sample sizes
    unmatched = c("drop", "error")
  ) %>% 
  invisible()
```

Code to check which (FAS-specific) panel visits have been dropped (code not evaluated):

```{r eval=FALSE}
scheme_moco_ps_spsubset_fas_panels %>% 
  select(scheme:in_aquatic_subset, notation_paneldesign, panel) %>% 
  anti_join(
    scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar,
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset,
      notation_paneldesign,
      panel
    )
  )
```

The dropped panel visits arise:

- because of small sample sizes in all involved strata in the specific scheme x module_combo, leading to occasional panels with no sampling units at all;
- if `month_first_default > 1`: panel visits are dropped that start too early, and FAGs that depend on it but come later at an involved location are also dropped.

Note that 'dropping panel visits' happens because some module combos only last one cycle, otherwise the missing panel is present in the next cycle.
This means that within the first cycle actually more panel visits are dropped.
Also, because of the cleaning with respect to the order of FAGs, actual locations may be dropped from a panel in the first cycle while this is not true for the entire panel.

But also note that the current strategy is to avoid dropping panel visits and locations by postponing them (and this may be reconsidered; see \@ref(cal-cleaning)).

Checking that the same panel designs are present in each stratum and spatial poststratum, in the spatial sample calendar and in the panel membership object.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar %>% 
  # this leaves out the panels and the spatial sampling units:
  distinct(pick(scheme:sp_poststratum), stratum) %>% 
  inner_join(
    scheme_moco_ps_spsubset_fas_stratum_sppost_panelmemship %>% 
      select(-panel_membership),
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset,
      notation_paneldesign,
      stratum,
      sp_poststratum
    ),
    # in the calendar, we get panels of the same panel design and FAGs but with
    # different activity sequences because of the cleaning on FAG X actseq.
    # Hency we get "many-to-one"
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  invisible()
```

Checking the stratum spatial sample sizes for _target_ FAGs in the calendar against the spatial sample sizes from `scheme_moco_ps_stratum_sppost_spsamples`, for each scheme x module combo x panel split.
This is the same check as done for the initial calendar.

```{r}
scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar %>% 
  unnest(fags) %>% 
  distinct(
    scheme,
    module_combo_code,
    panel_split,
    in_aquatic_subset,
    field_activity_group,
    stratum,
    grts_address
  ) %>% 
  semi_join(
    scheme_moco_ps_spsubset_targetfag,
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      in_aquatic_subset, # this one doesn't matter actually
      field_activity_group
    )
  ) %>% 
  count(
    scheme,
    module_combo_code,
    panel_split,
    field_activity_group,
    stratum,
    name = "sample_size_fag"
  ) %>% 
  inner_join(
    scheme_moco_ps_stratum_sppost_spsamples %>% 
      mutate(
        total_sample_size = map_int(sp_poststr_samples, nrow),
        .keep = "unused"
      ),
    join_by(
      scheme,
      module_combo_code,
      panel_split,
      stratum,
      sample_size_fag == total_sample_size
    ),
    # lhs has duplication due to > 1 target FAG per scheme in some cases
    relationship = "many-to-one",
    unmatched = "error"
  ) %>% 
  invisible()
```

#### Displaying the (scheme-specific) spatial sampling unit panel calendar

We display some items in a googlesheet for visual exploration and checking.
For the code, see `helper_scripts/writing_to_googlesheet.R`.

- `compartment_paneldesign_fags` (from section \@ref(nrpanels)) shows the different panel designs (as notations) and which FAGs, schemes and module combos are attached to it.
- a 'panel_calendar' worksheet, generated from the spatial sampling unit calendar (excluding ADHOC FAGs), lists the panels and number of involved locations, for each combination of date interval (rows) and panel design (notation) (columns).
See previous dataset to find the connection with FAGs.

### Spatiotemporal samples

Subsetting the sampling unit calendar to the occasions with a target FAG.
This object comes closest to the notion of **spatiotemporal sample**.

```{r}
scheme_moco_ps_spsubset_targetfag_stratum_sppost_spsamples_calendar <- 
  scheme_moco_ps_spsubset_fag_stratum_sppost_spsamples_calendar %>% 
  semi_join(
    scheme_moco_ps_spsubset_targetfag,
    join_by(module_combo_code, scheme, in_aquatic_subset, field_activity_group)
  )
```

## Field activities in space and time

Some field activity groups may overlap between schemes, i.e. when the same sampling unit is present in more than one scheme _and_ the same field activity group happens in these schemes at the same time for this sampling unit.

Further, the uncertainties in the sampling frame cause duplication of GRTS addresses in the (generated) sample because multiple strata can be connected to the same GRTS address.

While in below object `fag_grts_calendar` both sources of duplication have been collapsed, we also create an object `fag_stratum_grts_calendar` that better reflects the expected _number_ of field activities in space and time: the reduction by GRTS address will not happen in reality since the sample sizes must be maintained!
This means that new sampling units will be added from a list of spare sampling units, when a stratum on terrain is not the expected stratum, in order to maintain the sample size.
This specifically applies to the (many) types where the cell center in the field must be the type, so that only one 'cell_center' type can correspond to a GRTS address, contrary to the theoretical information from the sampling frame.

```{r}
fag_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address
  ) %>%
  arrange(pick(everything()))
```

```{r}
fag_stratum_grts_calendar <-
  scheme_domain_fag_stratum_spsamples_calendar %>%
  distinct(
    field_activity_group,
    pick(starts_with("date")),
    grts_address,
    stratum
  ) %>%
  arrange(pick(everything()))
```

## Deriving datasets and layers for use and visualization

```{r}
scheme_stratum_spsamples_sf <-
  scheme_moco_ps_stratum_dom_spsamples %>%
  select(scheme, stratum, grts_address) %>% 
  add_point_coords_grts()
```

```{r}
samplinglocations_sf <-
  scheme_moco_ps_stratum_dom_spsamples %>%
  distinct(grts_address) %>%
  add_point_coords_grts()
```

## Optimizing resource allocation in time and space (within or between years)

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "scheme_domain_stratum_spsamples",
  "scheme_stratum_spsamples_sf",
  "samplinglocations_sf",
  "scheme_fag_fa",
  "scheme_domain_fag_panel_count",
  "scheme_domain_fag_panels",
  "scheme_domain_fag_panel_calendar",
  "scheme_domain_fag_panel_stratum_spsamples",
  "scheme_domain_fag_stratum_spsamples_calendar",
  "fag_grts_calendar",
  "fag_stratum_grts_calendar"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
