# Generating spatiotemporal samples {#sampling}

<!-- use hardcoded tibble from file 010 with design attributes per scheme -->

<!-- Take into account the revisit design, sample size, surplus sampling, domains -->

<!-- Sample sizes can be summed over active module-domains that refer to the same domain. -->
<!-- Where one module-domain overlaps (contains) another module-domain, those sample sizes cannot be added but an appropriate offset must be taken into account when sampling in the shared area (the smaller domain), in order to respect the GRTS ordering. -->
<!-- This also holds for MHQ Flanders (not in this POC) vs MHQ in SACs. -->

For this proof of concept, we restrict the allowed modules.

```{r}
module_domain_scheme_stratum_sample_size_2 <- 
  module_domain_scheme_stratum_sample_size %>% 
  filter(module %in% c("pan_phase_1_sac_subset_5", "mbaa_mne_phase_1"))
```

## Defining the samples

For now still without:

- non-core schemes (except for its types that are not covered by core schemes)
- GRTS address offset in the common area between different but overlapping domains (including MHQ at Flemish level = undefined module)
- panel membership (try to do this by field method (aggregating schemes) if this causes no clashes)
- module labelling of sampling units
- adding (and labelling) spare sampling units
- collapsing nearby units for groundwater schemes

```{r}
scheme_domain_stratum_spsamples <- 
  module_domain_scheme_stratum_sample_size_2 %>% 
  select(module, domain, scheme, sampling_frame_id, stratum, sp_sample_size_all_panels_stratum) %>% 
  # for MNE, non-core schemes (dropped below) have yet to be taken back (but I
  # think this step also drops things we always need to drop):
  filter(!is.na(sp_sample_size_all_panels_stratum)) %>% 
  summarize(
    sp_sample_size_all_panels_stratum = sum(sp_sample_size_all_panels_stratum),
    .by = c(domain, scheme, sampling_frame_id, stratum)
  ) %>% 
  nest(domain_stratum_spsamplesize = c(domain, stratum, sp_sample_size_all_panels_stratum)) %>% 
  inner_join(sp_samplingframe_domain, by = join_by(sampling_frame_id)) %>% 
  mutate(
    spsample = map2(domain_stratum_spsamplesize, sampling_frame_domain, function(dss, sfd) {
      sfd %>% 
        nest(grts_addresses = grts_address) %>% 
        inner_join(dss, ., by = join_by(domain, stratum)) %>% 
        mutate(sample = 
                 map2(sp_sample_size_all_panels_stratum, grts_addresses, function(sss, grts) {
                   grts %>% slice_min(grts_address, n = sss)
                 })
        ) %>% 
        select(domain, stratum, sample) %>% 
        unnest(sample)
    })
  ) %>% 
  select(scheme, spsample) %>% 
  unnest(spsample)
```

## Deriving datasets and layers for use and visualization

## Optimizing resource allocation in time and space (within or between years)

## Overview of most relevant objects in this chapter

```{r echo = FALSE}
for (name in c(
  "scheme_domain_stratum_spsamples"
)) {
  cat(name, "\n")
  glimpse(eval(str2lang(name)))
  cat("\n")
}
```
