# Introduction and data preparation

What we're going to explore here, is:

- Abundance of types
    - Overall
    - Terrestrial types: area proportion with PHAB < 10%
    - Divide between areas within and outside Special Areas of Conservation


- Relation with the shallowgroundwater layer


- Sample support
    - Terrestrial: homogeneity of the target population within discrete raster cells
        - Distribution
        - Surface estimation error


- GRTS reusal frequency of coarse raster cells when sampling fine raster cells, depending on type and coarse resolution


- Combining terrestrial, lentic and lotic types in making a sampling frame?


## Data preparation

### List of types for joining with habitat data sources

```{r}
scheme_types_mne <- 
    read_scheme_types(extended = TRUE) %>% 
    filter(programme == "MNE") %>% 
    distinct(scheme, type)
scheme_types_mne
```

Expanding types per scheme:

```{r}
scheme_types_mne_expanded <- 
    scheme_types_mne %>% 
    group_by(scheme) %>% 
    expand_types %>% 
    ungroup
scheme_types_mne_expanded
```

Unique types in the expanded list, with their hydrological class added:

```{r}
types_mne_expanded <- 
    scheme_types_mne_expanded %>% 
    distinct(type) %>% 
    inner_join(read_types() %>% select(type, hydroclass = tag_2_shortname),
               by = "type") %>% 
    arrange(type)
types_mne_expanded
```

The following types were added by expansion:

```{r}
scheme_types_mne %>% 
    distinct(type) %>% 
    anti_join(types_mne_expanded, .,
              by = "type")
```

Frequency of hydroclasses:

```{r}
types_mne_expanded %>% 
    count(hydroclass)
```


### Geospatial data sources of habitat types & RIB

- Terrestrial types (incl. 'semi-terrestrial'), and with exclusion of remaining aquatic-type patches (within potentially terrestrial polygons):

```{r}
hmt <- read_habitatmap_terr()
hmt_pat <- 
    hmt$habitatmap_terr_patches %>% 
    semi_join(types_mne_expanded %>% filter(!str_detect(hydroclass, "water")),
              by = "type")
hmt_pol <- hmt$habitatmap_terr_polygons
# Note that 
# hmt$habitatmap_terr_polygons %>% semi_join(hmt_pat, by = "polygon_id")
# yields the same number of polygons: non-terrestrial polygons were already excluded
```

```{r}
hmt_pat %>% 
    distinct(type) %>% 
    inner_join(types_mne_expanded, "type") %>% 
    count(hydroclass)
```

- Surface waters (excluding 3260):

```{r}
wsh <- read_watersurfaces_hab(interpreted = TRUE)
wsh_pol <- wsh$watersurfaces_polygons
wsh_pat <- wsh$watersurfaces_patches
```

```{r}
wsh_pat %>% 
    distinct(type) %>% 
    inner_join(types_mne_expanded %>% filter(str_detect(hydroclass, "water")),
              by = "type")
```


- Type 3260:

```{r}
habstream <- read_habitatstreams()
```

```{r}
types_mne_expanded %>% 
    filter(type == "3260")
```





