---
title: "Water Level Variograms"
format:
  html:
    toc: true
    html-math-method: katex
  pdf:
    fig-pos: 'H'
knitr:
  opts_chunk:
    echo: true
---

# Summary 
## TL;DR


## List of Relevant Figures

Analyzing water levels (mTAW):
- sample size: @fig-samplesize
- binned difference-distance plot: @fig-dw-dx-binned 
- variogram: @fig-variogram


# Introduction 

```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox


source("./watina_migration.R")

# option to work offline; requires previous data download
work_offline <- FALSE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

# options(browser = "librewolf") # a rather personal thing

# the user may choose to reload raw data
force_database_download <- FALSE
```



This notebook has similar purpose as the previous one, 
yet instead of working with water chemistry, I will query water levels 
to calculate their spatial interrelation.
This serves as

- a proxy for water chemistry, where data availability was sparse
- a measurement for coupling of sites used for mnw `GW_05.1`.


# Data
## Settings
Here some design choices on settings that will constrain the data set.

```{r data-settings}
distance_cap <- 1024 # (m) # DO NOT CHANGE without reloading.
n_bins <- 1024 # number of bins for binning of distances (best use 2^x)
plot_bins <- 128 # a little less bins for plotting
max_filterdepth <- 3 # (m) # choice to focus on shallow gw (ecologically relevant for GW_03.3 and GW_05.1)
taw_buffer <- 10 # allow mTAW observations to be below well depth (ReferentieNiveauTAW - PeilbuisLengte) 
maaiveld_margin <- 16 # differences in water level must be within a margin from differences in ReferentieNiveauMaaiveld
dtaw_cap <- 16 # (m) # difference in reference TAW must be below this value; rationale: not interesting in steep slope comparison
```


Other constraints:

- query `mTAW`
- maximal sampling interval: same day (join `DatumWID`)
- same `GebiedWID`
- same `soilclass` on both locations
- maximal filterdepth difference: 0.1m
- validation VLD, SDV, UNK



## Query building
This will require a similar query as with water chemistry.
The amount of available data exceeded sql capacity, yet the `dbplyr` strategy worked.


```{r check-existing-rawdata}
raw_data_export_file <- "data/wata_couplingdata_arrow.parquet"
reload <- force_database_download || !file.exists(raw_data_export_file)
```

```{r raw-data-query-dbplyr}


query_measurement <- function (){
  tbl(watina_dwh, "FactPeilMeting") %>%
    left_join(
      tbl(watina_dwh, "dimPeilpunt"),
      by = join_by(PeilpuntWID == PeilpuntWID),
      relationship = "many-to-many"
    ) %>%
    filter(PeilmetingStatusCode == "VLD",
      ReferentieNiveauTAW + PeilbuisLengte > mTAW - taw_buffer,
      !(PeilMetingWID %in% c(9078367)) # ZWAP outlier (see TODO)
    ) %>% 
    select(
      PeilMetingWID,
      PeilbuisLengte,
      ReferentieNiveauMaaiveld,
      ReferentieNiveauTAW,
      FilterLengte,
      PeilpuntStatusCode,
      PeilpuntCode,
      MeetpuntWID.x,
      DatumWID,
      PeilpuntXCoordinaat,
      PeilpuntYCoordinaat,
      mTAW
    ) %>%
    mutate(
      filterdepth = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
    ) %>%
    filter(filterdepth <= max_filterdepth) %>% 
    rename(
      idx = PeilMetingWID,
      pblen = PeilbuisLengte,
      ref_mv = ReferentieNiveauMaaiveld,
      ref_taw = ReferentieNiveauTAW,
      flen = FilterLengte,
      fdepth = filterdepth,
      pp_code = PeilpuntCode,
      status = PeilpuntStatusCode,
      mp_idx = MeetpuntWID.x,
      date = DatumWID,
      x = PeilpuntXCoordinaat,
      y = PeilpuntYCoordinaat,
      w = mTAW
    ) %>%
   filter_at(
    vars(flen, fdepth, pp_code, w),
    all_vars(!is.na(.))
  )

}

if (reload) {
  fpm1 <- query_measurement()
  fpm2 <- query_measurement()

  glimpse(fpm1)
  # fpm1 %>% count
  # unique(fpm1 %>% pull(status))
}

```


Downloading A LOT OF data:
```{r raw-data-join-dbplyr}

if (reload) {
  data_raw <- fpm1 %>%
    left_join(fpm2,
      by = join_by(
        idx > idx, # measurement index
        date == date # on the same day
      )
    ) %>% mutate(
      dr = sqrt((x.y-x.x)^2 + (y.y-y.x)^2),
      dtaw = abs(ref_taw.y - ref_taw.x),
      dw = abs(w.y - w.x),
    ) %>%
    filter(
      dr < distance_cap,
      dw - dtaw < maaiveld_margin,
      dtaw < dtaw_cap,
      mp_idx.x < mp_idx.y # unidirectional difference / avoid duplicates
      ) %>%
    collect

  write_parquet(data_raw, sink = raw_data_export_file)
} else {
  data_raw <- read_parquet(raw_data_export_file)
}

glimpse(data_raw)
data_raw %>% count
```

The data is saved to disc locally and restored





## data preparation
Renaming, selecting, mutating, re-arranging, summarizing.


```{r raw-data-processing}
data <- data_raw %>%
  rename(
    idx1 = idx.x,
    loc1 = pp_code.x, 
    mp1  = mp_idx.x, 
    x1   = x.x,
    y1   = y.x,
    z1   = fdepth.x, 
    w1   = w.x,
    ref1 = ref_taw.x,
    idx2 = idx.y,
    loc2 = pp_code.y, 
    mp2  = mp_idx.y, 
    x2   = x.y,
    y2   = y.y,
    z2   = fdepth.y, 
    w2   = w.y,
    ref2 = ref_taw.y,
  ) %>% select(
    idx1, loc1, mp1,
    x1, y1, z1, w1,
    idx2, loc2, mp2,
    x2, y2, z2, w2,
    ref1, ref2,
    date, dr
  ) %>%
  filter_at(
    vars(x1, x2, y1, y2, z1, z2, w1, w2, ref1, ref2),
    all_vars(!is.na(.))
  ) %>%
  mutate(
    region = substring(loc1, 1, 3),
    dx = sqrt((x2-x1)^2 + (y2-y1)^2),
    distance_bin = cut(dx, breaks = seq(-distance_cap/n_bins, distance_cap, length.out = n_bins+2)),
    dx_log = log10(dx + 1),
    dz = z2-z1,
    dref = ref2-ref1,
    dw = w2-w1
  ) %>% 
  arrange(idx1, idx2) %>% 
  mutate_at(
    vars(
      region, loc1, loc2, mp1, mp2, distance_bin
    ), as.factor
  ) %>%
  group_by(region, loc1, loc2, distance_bin) %>%
  summarize(
    dx = mean(abs(dx), na.rm = TRUE),
    dz = mean(dz, na.rm = TRUE),
    dw = mean(abs(dw), na.rm = TRUE),
    dref = mean(abs(dref), na.rm = TRUE),
    cnt = n()
  ) %>% ungroup
glimpse(data)
```



```{r fig-samplesize}
#| label: fig-samplesize
#| fig-cap: "Sample size (number of site pairs observed) per distance (${\\Delta x}$)."

data %>% 
  mutate(
    distance_bin = cut(
        dx,
        breaks = seq(-distance_cap/plot_bins, distance_cap, length.out = plot_bins+2)),
  ) %>% 
  group_by(distance_bin) %>% 
  summarize(cnt = sum(cnt)) %>% 
  ungroup %>% 
  arrange(distance_bin) %>% 
  mutate(distance = distance_cap/n_bins * as.numeric(distance_bin)) %>% 
  ggplot(aes(x = distance, y = cnt)) +
  geom_bar(stat = "identity", width = 1) +
  scale_x_continuous(breaks=seq(0,distance_cap,64)) +
  xlab("distance (m)") +
  ylab("number of observations (N)") +
  theme_bw()
```

NOTE: there are many co-located sites (\(dx \eq 0m\)).


# Soilclass

```{r check-existing-soilclass}
soilclass_export_file <- "data/soilclass.parquet"
reload_soil <- force_database_download || !file.exists(soilclass_export_file)
```


```{r load-soilclass}
if (reload_soil) {
  source("./zenodo_helper.R")
  soilmap_raw <- load_zenodo_data("soilmap")
  
  soilmap <- soilmap_raw %>%
    transmute(
      soilclass = case_when(
        stringr::str_detect(bsm_mo_tex, "V") ~ "peat",
        stringr::str_detect(bsm_mo_tex, "^U") ~ "heavy",
        stringr::str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
        stringr::str_detect(bsm_mo_tex, ".+") ~ "heavy",
        .default = "unknown"
      ) %>%
        factor()
    )

  soil_locations <- data_raw %>%
    select(pp_code.x, x.x, y.x) %>% 
    rename(loc = pp_code.x, x = x.x, y = y.x) %>% 
    distinct(.keep_all = TRUE)

  soil_locations <- data_raw %>%
    select(pp_code.y, x.y, y.y) %>% 
    rename(loc = pp_code.y, x = x.y, y = y.y) %>% 
    distinct(.keep_all = TRUE) %>%
    rbind(soil_locations) %>% 
    distinct(.keep_all = TRUE) %>% 
    sf::st_as_sf(coords = c("x", "y"), crs = 31370)

  lookup_soil <- sf::st_join(
    soil_locations, soilmap
  ) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown")) %>%
  sf::st_drop_geometry()
  

  write_parquet(lookup_soil, sink = soilclass_export_file)
} else {
  lookup_soil <- read_parquet(soilclass_export_file)
}

```


```{r join-soilclass}
data <- data %>%
  left_join(
    rename(lookup_soil, c("soilclass" = soilclass)),
    by = join_by(loc1 == loc),
    relationship = "many-to-one"
    ) %>%
  left_join(
    rename(lookup_soil, c("soilclass2" = soilclass)),
    by = join_by(loc2 == loc),
    relationship = "many-to-one"
    )

data <- data %>%
  dplyr::filter(soilclass == soilclass2) %>%
  select(-soilclass2)
```


```{r fig-samplesize-soilclass}
#| label: fig-samplesize-soilclass
#| fig-cap: "Sample size (number of site pairs observed) per distance (${\\Delta x}$), split per soil class."

data %>% 
  mutate(
    distance_bin = cut(
        dx,
        breaks = seq(-distance_cap/plot_bins, distance_cap, length.out = plot_bins+2)),
  ) %>% 
  group_by(distance_bin, soilclass) %>% 
  summarize(cnt = sum(cnt)) %>% 
  ungroup %>% 
  arrange(distance_bin) %>% 
  mutate(distance = distance_cap/n_bins * as.numeric(distance_bin)) %>% 
  ggplot(aes(x = distance, y = cnt, fill = soilclass)) +
  geom_bar(stat = "identity", width = 1) +
  scale_x_continuous(breaks=seq(0,distance_cap,64)) +
  xlab("distance (m)") +
  ylab("number of observations (N)") +
  theme_bw()
```

# Difference-Distance Plots

## unbinned

```{r plot-data-unbinned}
#| label: fig-dw-dx
#| fig-cap: "Distance-Difference-Plot of absolute difference in water level (mTAW) against distance (${\\Delta x}$)."

data %>%
  ggplot(aes(x = dx, y = abs(dw))) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  # xlim(0, 100) +
  # ylim(0,  25) +
  xlab("distance (m)") +
  ylab("absolute water level difference Δw (mTAW)") +
  theme_bw()
```


```{r plot-data-unbinned-dz}
#| label: fig-dw-dz
#| fig-cap: "Distance-Difference-Plot of absolute difference in water level (mTAW) against filter depth difference ($\\Delta z$)."

data %>%
  ggplot(aes(x = abs(dz), y = abs(dw))) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  # ylim(0, 100) +
  # xlim(0, 3) +
  xlab("filter depth difference dz (m)") +
  ylab("absolute water level difference Δw (mTAW)") +
  theme_bw()
```

NOTE: we do have some very deep drills (\(dz >> 50m\)), yet those were filtered above.


```{r plot-data-dref}
#| label: fig-dw-dref
#| fig-cap: "Influence of reference: absolute difference in water level (mTAW) against absolute difference in site reference TAW."

data %>%
  ggplot(aes(x = log10(abs(dref)), y = abs(dw))) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  # ylim(0, 100) +
  # xlim(0, 3) +
  xlab("log absolute difference in reference TAW") +
  ylab("absolute water level difference Δw (mTAW)") +
  theme_bw()
```


## binned data
*There must be a reason why 9 out of 10 variogram experts apply binning.*

https://library.soton.ac.uk/variance-standard-deviation-and-standard-error

```{r bin-data}
nanmean <- function (val) mean(val, na.rm = TRUE)
nanvar <- function (val) var(val, na.rm = TRUE)
nansd <- function (val) sd(val, na.rm = TRUE)
nansem <- function (val) sd(val, na.rm = TRUE)/sqrt(sum(!is.na(val)))
data_binned <- data %>%
  group_by(distance_bin, soilclass) %>% 
  summarize(
    vx = nanvar(dx), vz = nanvar(dz),
    vmw = nanvar(dw),
    sigmw = nansd(dw),
    semmw = nansem(dw),
    dx = nanmean(abs(dx)), dz = nanmean(abs(dz)), dw = nanmean(abs(dw)),
    cnt = sum(cnt)
  ) %>% ungroup %>% arrange(dx)
glimpse(data_binned)
```


```{r plot-data-binned}
#| label: fig-dw-dx-binned
#| fig-cap: "Distance-Difference-Plot of difference in water level (mTAW) against binned distance ($\\Delta x$)."

data_binned %>%
  ggplot(aes(x = dx, y = abs(dw))) +
  geom_point(alpha = 0.8, aes(color = soilclass)) +
  geom_smooth(color = "black") +
  xlab("site distance (m)") +
  ylab("mean absolute water level difference <|Δw|> (mTAW)") +
  theme_bw() 

```


# Variograms

```{r variogram}
#| label: fig-variogram
#| fig-cap: "variogram of difference in water level (mTAW) against binned distance ($\\Delta x$)."

data_binned %>%
  ggplot(aes(x = dx, y = vmw/2)) +
  geom_point(alpha = 0.8, aes(color = soilclass)) +
  geom_smooth(color = "black") +
  xlab("site distance (m)") +
  ylab("semivariance in water level difference var(Δw)/2 (mTAW)") +
  theme_bw()

```


TODO: 
- package variograms
- models



# Archive
```{r disconnect-watina, eval=FALSE}
#| include: false
if (!work_offline) {
  DBI::dbDisconnect(watina_dwh)
}
```



## query attempt

```{r raw-data-query-dbi, eval=FALSE}
#| include: false
#| eval: false

load_raw_data <- function () {
  DBI::dbGetQuery(watina_dwh,
    paste0("
        SELECT
          PEIL1.region AS region,
          PEIL1.PeilMetingWID AS id1,
          PEIL1.pp_code AS pp1,
          PEIL1.pp_status AS status1,
          PEIL1.filterdepth AS z1,
          PEIL1.MeetpuntWID AS mp1,
          PEIL1.DatumWID AS date1,
          PEIL1.mTAW AS mw1,
          PEIL2.PeilMetingWID AS id2,
          PEIL2.pp_code AS pp2,
          PEIL2.pp_status AS status2,
          PEIL2.filterdepth AS z2,
          PEIL2.MeetpuntWID AS mp2,
          PEIL2.DatumWID AS date2,
          PEIL2.mTAW AS mw2
        FROM (
          SELECT 
            FactPeilMeting.*,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
              AS filterdepth,
            dimPeilpunt.PeilpuntCode AS pp_code,
            dimPeilpunt.PeilpuntStatusCode AS pp_status
          FROM FactPeilMeting
          LEFT JOIN dimPeilpunt 
            ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
          ) PEIL1
        LEFT JOIN (
          SELECT 
            FactPeilMeting.*,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
              AS filterdepth,
            dimPeilpunt.PeilpuntCode AS pp_code,
            dimPeilpunt.PeilpuntStatusCode AS pp_status
          FROM FactPeilMeting
          LEFT JOIN dimPeilpunt
            ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
          ) PEIL2
          ON (PEIL1.DatumWID = PEIL2.DatumWID)
         AND (PEIL1.region = PEIL2.region)
        WHERE (PEIL1.PeilpuntWID < PEIL2.PeilpuntWID)
          AND ABS(PEIL2.filterdepth - PEIL1.filterdepth) <= 1.0
        ")
    ) 
}

```


## obswell locations
Again, for the query of spatial information, `get_locs` will be fine.

```{r raw-data-locs}
#| eval: false
#| include: false

area_code_selection <- data_raw %>%
  pull(pp_code.x, pp_code.y) %>%
  substring(first = 1, last = 3) %>% 
  unique


if (force_database_download || !file.exists("data/locs_backup.parquet")) {
  locs <- get_locs(
    watina_dwh,
    area_codes = area_code_selection,
    obswells = TRUE, 
    filterdepth_range = c(0, 1000),
    collect = TRUE
  ) 
  write_parquet(locs, sink = "data/locs_backup.parquet")
} else {
  locs <- read_parquet("data/locs_backup.parquet")

}
# 
# locs %>% filter(loc_code == "BUIP003") %>% arrange(loc_code)
```


## Database Exploration
List of tables, as a reminder...
```{r watina-table-list}
#| eval: false
table_list <- get_db_table_list()
table_list
```

Content of water level table:
```{r watina-peilmeting-1}
#| eval: false
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT TOP 100 *
    FROM dimPeilpunt
    ;")
  ))
```

```{r watina-peilmeting-2}
#| eval: false
waterlevels_preview <- tbl(watina_dwh, "FactPeilMeting")
glimpse(waterlevels_preview)
```

```{r watina-gebied-1}
#| eval: false
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT TOP 10 *
    FROM dimGebied
    ;")
  ))
```

```{r peilmeting-query-example}
#| eval: false
# heavy exec - ain't do! (HEAD)
# test <- DBI::dbGetQuery(watina_dwh,
#   paste0("
#     SELECT *
#     FROM FactPeilMeting
#   "))
# test
```


```{r peilmeting-query-dbplyr}
#| eval: false
test <- tbl(watina_dwh, "FactPeilMeting") %>%
    left_join(tbl(watina_dwh, "dimPeilpunt"),
              by = join_by(PeilpuntWID == PeilpuntID))
glimpse(test)
```


## Data Consistency
There are some observation bins wehre the variance goes crazy; worth investigating.

```{r find-outlier-bins}
#| eval: false

data_binned %>%
  arrange(-abs(dw))
  # arrange(-vmw)

```

- (761, 762]
- (952, 953]
- (954, 955]
- (1010, 1012]
- (811, 812]


what is going on in those?
```{r find-outliers-data}
#| eval: false


data %>%
  filter(dx > 811, dx <= 812)
  # filter(dx > 952, dx <= 953)
  # filter(dw >= 20)

```

```{r find-outliers-raw}
#| eval: false
outlier <- data_raw %>%
  filter(
    # pp_code.x == "ABEP115X",
    # pp_code.y == "ABEP111X"
    # pp_code.x == "ZWAP234X",
    # pp_code.y == "ZWAP212X"
    # pp_code.x == "BOBP218X",
    # pp_code.y == "BOBP201X"
    # pp_code.x == "GGVP012X",
    # pp_code.y == "GGVP024X"
    # pp_code.x == "RONP119X",
    # pp_code.y == "RONP124X"
    pp_code.x == "WALP066X",
    pp_code.y == "WALP043X"
  ) %>%
  mutate(dw = abs(w.y-w.x)) %>% 
  arrange(-dw)
glimpse(outlier)
```

```{r find-outliers-general}
#| eval: false
incosistent_depth <- fpm1 %>%
  filter(w - fdepth > 10) %>% 
  collect %>% 
  mutate(subdepth = w - fdepth) %>%
  arrange(-subdepth)
  
glimpse(incosistent_depth)
```

```{r find-outlier-query}
#| eval: false

outliers <- DBI::dbGetQuery(
  watina_dwh,
  paste0("
    SELECT 
      FactPeilMeting.*,
      (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
        AS filterdepth,
      dimPeilpunt.*,
      mTAW AS WTF
    FROM FactPeilMeting
    LEFT JOIN dimPeilpunt 
      ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
    WHERE PeilpuntCode IN ('ZWAP234X', 'ZWAP212X')
      AND DatumWID = 20011129
      -- WHERE (PeilbuisLengte - ReferentieNiveauMaaiveld) < 10
      -- AND (ReferentieNiveauTAW - PeilbuisLengte) < FactPeilMeting.mTAW
      -- AND FactPeilMeting.mTAW > 10
      -- AND mTAW > 100
    ORDER BY mTAW DESC
  ")
) %>% janitor::clean_names()
glimpse(outliers)
outliers %>% filter(peil_meting_wid == 4785841)
outliers %>%
  group_by(peilpunt_code) %>%
  summarize_all(mean)
    select(peilpunt_code, datum_wid, m_taw, referentie_niveau_taw, filterdepth, peilbuis_lengte) %>%
  head(10)
```

A query for the experts (✉20241127):
```{r}
#| eval: false
outliers <- DBI::dbGetQuery(
  watina_dwh,
  paste0("
    SELECT DISTINCT PeilpuntCode,
      PeilMetingStatusCode,
      AVG(PeilbuisLengte) AS PeilbuisLengte,
      AVG(ReferentieNiveauTAW) AS ReferentieNiveauTAW,
      AVG(mTAW) AS mTAW,
      AVG(Niveau) AS Niveau,
      COUNT(PeilMetingWID) AS n,
      MAX(mTAW - ReferentieNiveauTAW) AS max_verschil
    FROM FactPeilMeting
    LEFT JOIN dimPeilpunt 
      ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
    WHERE (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)) < 6
      AND ReferentieNiveauTAW + PeilbuisLengte + 10 < FactPeilMeting.mTAW
    GROUP BY PeilpuntCode, FactPeilMeting.PeilpuntWID, PeilMetingStatusCode
    ORDER BY max_verschil DESC
  ")
)
glimpse(outliers)
```


## Medewerker nVarchar?
I got an error that MedewerkerWID in dimPeilpunt is of dtype `nvarchar`.

```{r watina-medewerker-1}
#| eval: false
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT DISTINCT MedewerkerWID, COUNT(PeilpuntWID) AS cnt
    FROM dimPeilpunt
    GROUP BY MedewerkerWID
    ORDER BY MedewerkerWID ASC
    ;")
  ))
```

```{r watina-medewerker-2}
#| eval: false
query <- function(query_string) {
  rs <- DBI::dbGetQuery(watina_dwh, query_string)
  return(rs)
}

get_tablecolumns <- function(tbl_name) {
  my_columns <- query(paste0(
    "SELECT OBJECT_NAME(object_id), name, column_id, TYPE_NAME(system_type_id)
     FROM sys.columns
     WHERE (object_id = OBJECT_ID('",
    tbl_name, "')) ;")
  )
  return(my_columns)
}


table_list <- tbl(watina_dwh, "sysobjects") %>%
  filter(xtype == "U") %>%
  pull(name)


col_list <- get_tablecolumns("dimPeilpunt")
```

```{r watina-medewerker-3}
#| eval: false
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
SELECT
  OBJECT_NAME(object_id) AS tbl,
  name,
  column_id,
  TYPE_NAME(system_type_id) AS dtype
FROM sys.columns
WHERE (object_id = OBJECT_ID('dimPeilpunt'))
  AND name LIKE '%medew%'
;")
  ))

```


# TODO

## mTAW/refTAW inconsistency
Some colocated sites have inconsistent TAW references and/or mTAW measurements.

```{r zwap-inconsistency-query}
#| eval: false

outliers <- DBI::dbGetQuery(
  watina_dwh,
  paste0("
    SELECT 
      FactPeilMeting.*,
      (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
        AS filterdepth,
      dimPeilpunt.*,
      mTAW AS WTF
    FROM FactPeilMeting
    LEFT JOIN dimPeilpunt 
      ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntWID
    WHERE PeilpuntCode IN ('ZWAP234X', 'ZWAP212X')
      AND DatumWID = 20011129
    ORDER BY mTAW DESC
  ")
) %>% janitor::clean_names()
glimpse(outliers)
```

More examples:
```{r find-outliers-backup}
#| eval: false
outlier <- data_raw %>%
  filter(
    # pp_code.x == "ABEP115X",
    # pp_code.y == "ABEP111X"
    # pp_code.x == "ZWAP234X",
    # pp_code.y == "ZWAP212X"
    # pp_code.x == "BOBP218X",
    # pp_code.y == "BOBP201X"
    # pp_code.x == "GGVP012X",
    # pp_code.y == "GGVP024X"
    # pp_code.x == "RONP119X",
    # pp_code.y == "RONP124X"
    pp_code.x == "WALP066X",
    pp_code.y == "WALP043X"
  ) %>%
  mutate(dw = abs(w.y-w.x)) %>% 
  arrange(-dw)
glimpse(outlier)
```

