---
title: "Water Level Variograms"
format:
  html:
    toc: true
    html-math-method: katex
  pdf:
    fig-pos: 'H'
knitr:
  opts_chunk:
    echo: true
---


# Introduction 

```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox


source("./watina_migration.R")

if (!exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

# options(browser = "librewolf") # a rather personal thing

# the user may choose to reload raw data
choose_to_reload <- FALSE
```



This notebook has similar purpose as the previous one, 
yet instead of working with water chemistry, I will query water levels 
to calculate their spatial interrelation.


# Data
## Database Exploration
List of tables, as a reminder:
```{r watina-table-list}
table_list <- get_db_table_list()
table_list
```

Content of water level table:
```{r watina-peilmeting-1}
#| eval: false
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT *
    FROM FactPeilMeting
    LIMIT 1000
      ")
  ))
```

```{r watina-peilmeting-2}
waterlevels_preview <- tbl(watina_dwh, "FactPeilMeting")
glimpse(waterlevels_preview)
```


```{r peilmeting-query-example}
# test <- DBI::dbGetQuery(watina_dwh,
#   paste0("
#     SELECT *
#     FROM FactPeilMeting
#   "))
# test
# # long exec
```

```{r peilmeting-query-dbplyr}
test <- tbl(watina_dwh, "FactPeilMeting") %>%
    left_join(tbl(watina_dwh, "dimPeilpunt"),
              by = join_by(PeilpuntWID == PeilpuntID))
glimpse(test)
```

## Query building
This will require a similar query as with water chemistry.

Some constraints:

- query `mTAW`
- maximal sampling interval: 1d (join `DatumWID`)
- same `GebiedWID`
- maximal filterdepth difference: 0.1m
- validation VLD, SDV, UNK


```{r check-existing-rawdata}
raw_data_export_file <- "data/wata_couplingdata_arrow.parquet"
reload <- choose_to_reload || !file.exists(raw_data_export_file)
```

```{r raw-data-query-dbplyr}


query_measurement <- function (){
  tbl(watina_dwh, "FactPeilMeting") %>%
    left_join(
      tbl(watina_dwh, "dimPeilpunt"),
      by = join_by(PeilpuntWID == PeilpuntID),
      relationship = "many-to-many"
    ) %>%
    filter(PeilmetingStatusCode == "VLD") %>% 
    select(
      PeilMetingWID,
      PeilbuisLengte,
      ReferentieNiveauMaaiveld,
      FilterLengte,
      PeilpuntStatusCode,
      PeilpuntCode,
      MeetpuntWID.x,
      DatumWID,
      PeilpuntXCoordinaat,
      PeilpuntYCoordinaat,
      mTAW
    ) %>%
    mutate(
      filterdepth = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
    ) %>%
    rename(
      idx = PeilMetingWID,
      pblen = PeilbuisLengte,
      ref_mv = ReferentieNiveauMaaiveld,
      flen = FilterLengte,
      fdepth = filterdepth,
      pp_code = PeilpuntCode,
      status = PeilpuntStatusCode,
      mp_idx = MeetpuntWID.x,
      date = DatumWID,
      x = PeilpuntXCoordinaat,
      y = PeilpuntYCoordinaat,
      w = mTAW
    ) %>%
   filter_at(
    vars(flen, fdepth, pp_code, w),
    all_vars(!is.na(.))
  )

}

if (reload) {
  fpm1 <- query_measurement()
  fpm2 <- query_measurement()

  glimpse(fpm1)
  # fpm1 %>% count
  # unique(fpm1 %>% pull(status))
}

```

TODO attempting:
```{r raw-data-join-dbplyr}

if (reload) {
  data_raw <- fpm1 %>%
    left_join(fpm2,
      by = join_by(
        idx > idx,
        date == date
      )
    ) %>% mutate(
      dr = sqrt(abs(x.x-y.x) + abs(x.y-y.y))
    ) %>%
    filter(dr < 100) %>%
    collect

  write_parquet(data_raw, sink = raw_data_export_file)
} else {
  data_raw <- read_parquet(raw_data_export_file)
}

glimpse(data_raw)
```






```{r raw-data-query-function}
#| include: false
#| eval: false

load_raw_data <- function () {
  DBI::dbGetQuery(watina_dwh,
    paste0("
        SELECT
          PEIL1.region AS region,
          PEIL1.PeilMetingWID AS id1,
          PEIL1.pp_code AS pp1,
          PEIL1.pp_status AS status1,
          PEIL1.filterdepth AS z1,
          PEIL1.MeetpuntWID AS mp1,
          PEIL1.DatumWID AS date1,
          PEIL1.mTAW AS mw1,
          PEIL2.PeilMetingWID AS id2,
          PEIL2.pp_code AS pp2,
          PEIL2.pp_status AS status2,
          PEIL2.filterdepth AS z2,
          PEIL2.MeetpuntWID AS mp2,
          PEIL2.DatumWID AS date2,
          PEIL2.mTAW AS mw2
        FROM (
          SELECT 
            FactPeilMeting.*,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
              AS filterdepth,
            dimPeilpunt.PeilpuntCode AS pp_code,
            dimPeilpunt.PeilpuntStatusCode AS pp_status
          FROM FactPeilMeting
          LEFT JOIN dimPeilpunt 
            ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
          ) PEIL1
        LEFT JOIN (
          SELECT 
            FactPeilMeting.*,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            (PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2))
              AS filterdepth,
            dimPeilpunt.PeilpuntCode AS pp_code,
            dimPeilpunt.PeilpuntStatusCode AS pp_status
          FROM FactPeilMeting
          LEFT JOIN dimPeilpunt
            ON FactPeilMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
          ) PEIL2
          ON (PEIL1.DatumWID = PEIL2.DatumWID)
         AND (PEIL1.region = PEIL2.region)
        WHERE (PEIL1.PeilpuntWID < PEIL2.PeilpuntWID)
          AND ABS(PEIL2.filterdepth - PEIL1.filterdepth) <= 1.0
        ")
    ) 
}

```


```{r raw-data-loading}
raw_data_export_file <- "data/wata_couplingdata_arrow.parquet"
if (choose_to_reload || !file.exists(raw_data_export_file)) {
  data_raw <- load_raw_data()
  write_parquet(data_raw, sink = raw_data_export_file)
} else {
  data_raw <- read_parquet(raw_data_export_file)
}
glimpse(data_raw)
```

## obswell locations
Again, for the query of spatial information, `get_locs` will be fine.

```{r raw-data-locs}
area_code_selection <- data_raw %>%
  pull(region) %>%
  unique


if (choose_to_reload || !file.exists("data/locs_backup.parquet")) {
  locs <- get_locs(
    watina_dwh,
    obswells = TRUE, 
    filterdepth_range = c(0, 1000),
    collect = TRUE
  ) 
  write_parquet(locs, sink = "data/locs_backup.parquet")
} else {
  locs <- read_parquet("data/locs_backup.parquet")

}
# 
# locs %>% filter(loc_code == "BUIP003") %>% arrange(loc_code)
```

## double-joining locs
Comparing two locations, I will join the locs twice: once to every location in the pair.
Then, distances and differences are computed, and relevant entries filtered.
Finally, factors are generated from categorical variables.

```{r raw-data-join-locs}
data <- data_raw %>%
  left_join(
    select(locs, obswell_code, x, y),
    by = join_by(pp1 == obswell_code),
    relationship = "many-to-one"
  ) %>% rename(c("x1" = x, "y1" = y)) %>%
  left_join(
    select(locs, obswell_code, x, y),
    by = join_by(pp2 == obswell_code),
    relationship = "many-to-one"
  ) %>% rename(c("x2" = x, "y2" = y)) %>%
  filter_at(
    vars(x1, x2, y1, y2, z1, z2),
    all_vars(!is.na(.))
  ) %>%
  mutate(
    dx = sqrt((x2-x1)^2 + (y2-y1)^2),
    dx_log = log10(dx + 1),
    dt = sqrt((t2-t1)^2),
    dz = z2-z1,
    dmw = mw2 - mw1,
    # dmw_log = log(abs(dmw)+1.0),
    dmw_rel = (mw2 - mw1) / (abs(mw1) + abs(mw2))
  ) %>%
  filter(
    dx > 0,
    dx <= 1000,
    abs(dmw) <= 16
  ) %>%
  mutate_at(
    vars(
      region
    ), as.factor
  )

```


```{r brief-regionwise-count}
data %>%
  summarize(cnt = n(), .by = c(region))
```


# Difference-Distance Plots

## unbinned

```{r plot-data-unbinned}
#| label: fig-dmw-dx
#| fig-cap: "Distance-Difference-Plot of difference in water level (mTAW) against distance ($\\log{\\Delta x}$)."

data %>%
  ggplot(aes(x = dx_log, y = abs(dmw))) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  theme_bw()
```


```{r plot-data-unbinned-dz}
#| label: fig-dmw-dz
#| fig-cap: "Distance-Difference-Plot of difference in water level (mTAW) against filter depth difference ($\\Delta z$)."

data %>%
  ggplot(aes(x = abs(dz), y = abs(dmw))) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  theme_bw()
```

## binned variogram
*There must be a reason why 9 out of 10 variogram experts apply binning.*

```{r bin-data}
data_binned <- data %>%
  mutate(
    distance_bin = as.factor(ntile(dx, n = 200))
    )

data_binned <- data_binned %>%
  summarize_at(
    .funs = function (x) mean(x, na.rm = TRUE),
    .vars = vars(dx, dx_log, dt, dz, dmw, dmw_rel),
    .by = distance_bin
  ) %>%
  left_join(
    data_binned %>% summarize(count = n(), .by = distance_bin),
    by = "distance_bin",
    relationship = "one-to-one"
    )
glimpse(data_binned)
```


```{r plot-data-binned}
#| label: fig-dmw-dx-binned
#| fig-cap: "Distance-Difference-Plot of difference in water level (mTAW) against binned distance ($\\Delta x$)."

data_binned %>%
  ggplot(aes(x = dx, y = abs(dmw))) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  theme_bw() 

```


# Archive
```{r disconnect-watina}
#| include: false
DBI::dbDisconnect(watina_dwh)
```
