---
title: "Water Level Variograms and Models"
bibliography: references.bib
format:
  html:
    toc: true
    html-math-method: katex
  pdf:
    fig-pos: 'H'
knitr:
  opts_chunk:
    echo: true
---


# Intro
## goal
The purpose of this notebook is twofold.

- First, it can be used to query country-wide watina locations with all kind of relevant data.
- Second, this will be used for variogram modeling with `sdmTMB`.

The measurements of choice are "xG3", per hydrological year.


## assumptions
To get installation specificities per hydroyear joined onto the measurements (which are "per location"), some heuristic choices are made.
Namely, I take attributes of the newer installation in replacement years, except for filter depth, where the shallower value is used.


Furthermore, there are some settings below determining how sites are clustered (`cluster_dist`) 
and what range is used for elevation slope calculation (`margin <- 2.5`).


# Data

## prep
```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")

# option to work offline; requires previous data download
work_offline <- TRUE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

# options(browser = "librewolf") # a rather personal thing

# the user may choose to reload raw data
force_database_download <- FALSE

if ("" == getOption("browser")) options(browser = "librewolf")

```


```{r data-settings}

extent <- 1024 # units::set_units(1024, "m") # m
n_bins <- 128

# result of previous analysis
cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 16 # minimal cluster member count; fewer memeber clusters are removed


data_export_file <- paste0("data/xg3_data_cl", cluster_dist, ".parquet") # where data is stored
```


## query
```{r raw-data-query-xg3}
#| eval: true

query_xg3 <- function (){
  tbl(watina_dwh, "factBRPeilMetingJaar") %>%
    filter(
      IsHydroJaar == 1,
      ParameterSetWID == 1
    ) %>% 
    select(
      MeetpuntWID,
      Jaar,
      HGmTAWValue,
      LGmTAWValue,
      VGmTAWValue
    ) %>%
    filter_at(
      vars(HGmTAWValue, LGmTAWValue),
      all_vars(!is.na(.))
    ) %>% 
    left_join(
      tbl(watina_dwh, "dimMeetpunt"),
      by = join_by(MeetpuntWID == MeetpuntWID),
      relationship = "many-to-many"
    ) %>%
    filter(MeetpuntStatusCode == "VLD") %>%
    select(
      MeetpuntWID,
      # FilterLengte,
      # MeetpuntStatusCode,
      MeetpuntCode,
      Jaar,
      MeetpuntXCoordinaat,
      MeetpuntYCoordinaat,
      HGmTAWValue,
      LGmTAWValue,
    ) %>%
    rename(
      idx = MeetpuntWID,
      code = MeetpuntCode,
      hyear = Jaar,
      x = MeetpuntXCoordinaat,
      y = MeetpuntYCoordinaat,
      hw = HGmTAWValue,
      lw = LGmTAWValue
    ) %>%
   filter_at(
    vars(code, hyear, hw, lw),
    all_vars(!is.na(.))
  )

}

```


```{r query-data}
#| eval: true
reload <- force_database_download || !file.exists(data_export_file)

if ((!work_offline) && reload) {
  data_raw <- query_xg3() %>%
    select(idx, code, hyear, x, y, hw, lw) %>% 
    filter_at(
      vars(x, y, hw, lw),
      all_vars(!is.na(.))
    ) %>% 
    collect %>%
    mutate(
      region = substring(code, first = 1, last = 3)
    ) %>%
    filter(region != "VBS") # remove Luxemburg

  glimpse(data_raw)
}
```



## filter depth
```{r query-filter-depth}
#| eval: true
if ((!work_offline) && reload) {
  pp <- tbl(watina_dwh, "dimPeilpunt") %>%
    filter(
      OpenbaarheidWID == 4 # note: about O(150) locs are non-public
    ) %>%
    mutate(
      yr = sql("YEAR(CAST(PeilpuntPlaatsing AS date))"),
      month = sql("YEAR(CAST(PeilpuntPlaatsing AS date))"),
      start_hyear = yr + (if (month >= 10) 1 else 0),
      yr = sql("YEAR(CAST(PeilpuntStopzetting AS date))"),
      month = sql("YEAR(CAST(PeilpuntStopzetting AS date))"),
      stop_hyear = yr + (if (month >= 10) 1 else 0),
      ref = ReferentieNiveauTAW,
      z = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
    ) %>%
    rename(
      code = PeilpuntCode,
      idx = MeetpuntWID
    ) %>%
    select(idx, code, start_hyear, stop_hyear, z, ref) %>%
    collect %>%
    mutate(
      region = substring(code, first = 1, last = 3) 
    ) %>% 
    filter(region != "VBS") %>%  # remove Luxemburg
    group_by(idx, start_hyear, stop_hyear) %>%
    slice(which.min(z)) 
  
  # Note: 15964 -> 10389 rows??
  #       filter depth z duplicates removed by `slice`
  #       replacement "by date" is hard to cover in this analysis
  #       because we work on xG3 and they are not associated with a date
  #       decided to just take the shallower value per location/year
  
  glimpse(pp)
}
```


```{r join-peilpunt}
#| eval: true
if ((!work_offline) && reload) {

for (col in c("z", "ref")) {
  if (col %in% colnames(data_raw)) {
    data_raw <- data_raw %>%
      select(-one_of(col))
  }
}

if (FALSE) {
  # some locations are not in the dim

  pp %>%
    filter(region == "WES")
  data_raw %>%
    filter(region == "WES")

  pp %>%
    filter(!(idx %in% unique(data_raw$idx))) %>%
    arrange(idx)
  data_raw  %>%
    filter(!(idx %in% unique(pp$idx))) %>% 
    arrange(idx)
}

data_raw <- data_raw %>%
  inner_join(pp,
    by = join_by(
      idx == idx,
    ),
    # relationship = "many-to-many"
    # unmatched = c("drop", "drop"),
    relationship = "many-to-many",
    suffix = c("", ".duplicates")
  ) %>%
  select(-ends_with(".duplicates")) %>% 
  filter(
    hyear >= start_hyear, # the observation must be after loc startup
    (is.na(stop_hyear) | (hyear < stop_hyear)) # ... or loc is up and running
  ) %>%
  select(-start_hyear, -stop_hyear)
glimpse(data_raw )

# 28k -> 22k rows due to unmatched
}
```



## soilclass

```{r check-existing-soilclass}
#| eval: true
soilclass_export_file <- "data/soilclass_perlocation.parquet"
reload_soil <- force_database_download || !file.exists(soilclass_export_file)
```


```{r load-soilclass}
#| eval: true
if ((!work_offline) && reload) {
if (reload_soil) {
  source("./zenodo_helper.R")
  soilmap_raw <- load_zenodo_data("soilmap_simple")
  
  soilmap <- soilmap_raw %>%
    transmute(
      soilclass = case_when(
        stringr::str_detect(bsm_mo_tex, "V") ~ "peat",
        stringr::str_detect(bsm_mo_tex, "^U") ~ "heavy",
        stringr::str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
        stringr::str_detect(bsm_mo_tex, ".+") ~ "heavy",
        .default = "unknown"
      ) %>%
        factor()
    )

  # soilmap %>% 
  # mapview(
  #   map.types = c("OpenTopoMap"),
  #   zcol = "soilclass"
  # )

  soil_locations <- data_raw %>%
    select(code, x, y) %>% 
    rename(loc = code) %>% 
    distinct(.keep_all = TRUE) %>%
    sf::st_as_sf(coords = c("x", "y"), crs = 31370)

  lookup_soil <- sf::st_join(
    soil_locations, soilmap
  ) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown")) %>%
  sf::st_drop_geometry()
  
  # lookup_soil %>% arrange(desc(loc))

  write_parquet(lookup_soil, sink = soilclass_export_file)
} else {
  lookup_soil <- read_parquet(soilclass_export_file)
}


# remove cluster column if already present
if ("soilclass" %in% colnames(data_raw)) {
  data_raw <- data_raw %>%
    select(-soilclass)
}
data_raw <- data_raw %>%
  left_join(
    lookup_soil,
    by = join_by(code == loc),
    relationship = "many-to-one"
  )

# data_raw %>% arrange(desc(code))
}
```


## clusters
```{r cluster-function}
#| eval: true
cluster_locs <- function (
    input,
    xvar = "x",
    yvar = "y",
    max_dist = 1,
    output_var = "cluster_id"
  ) {
  coords <- input[, c(xvar, yvar)]
  cluster_data <- coords %>%
    dist %>%
    fastcluster::hclust(method = "complete") %>%
    stats::cutree(h = max_dist) %>%
    tibble::enframe(name = NULL, value = output_var) %>%
    bind_cols(input, .)
}
```


```{r perform-clustering}
#| eval: true
if ((!work_offline) && reload) {

locs <- data_raw %>% 
  distinct(x, y)

# basic clustering
locs_clustered <- cluster_locs(
  input = locs,
  max_dist = cluster_dist, # m 
  output_var = "cluster",
  xvar = "x",
  yvar = "y"
) 


# remove cluster column if already present
if ("cluster" %in% colnames(data_raw)) {
  data_raw <- data_raw %>%
    select(-cluster)
}

# join the cluster number to the data_raw
data_raw <- data_raw %>% 
  left_join(
    locs_clustered,
    by = join_by(x == x, y == y),
    relationship = "many-to-one"
  )
  
# plot number of cluster members
locs_clustered  %>%
  ggplot(aes(x = cluster)) +
  geom_histogram(bins = max(data_raw$cluster))


tiny_clusters <- data_raw %>%
  group_by(cluster) %>%
  summarize(count = n()) %>% 
  filter(count < min_cluster_size) %>%
  pull(cluster)

data_raw <- data_raw %>%  
  filter(!cluster %in% tiny_clusters)  

}
```


```{r cluster-wise-detrend}
#| eval: true

detrend_clusterwise <- function(cl) {
  sub_dc <- data_raw %>%
    filter(cluster == cl)
  sub_dc <- sub_dc %>%
    mutate(
      hwd2_cl = lm(hw ~ x + y, sub_dc)$residual,
      lwd2_cl = lm(lw ~ x + y, sub_dc)$residual
    )
  return(sub_dc)
}

if ((!work_offline) && reload) {
data_raw <- bind_rows(
  lapply(
    unique(data_raw$cluster),
    FUN = detrend_clusterwise
  )
)
}

```


## distance to water


```{r load-water-sources}
#| eval: true
source("./spatial_helpers.R")
source("./water_sources.R")

if ((!work_offline) && reload) {
  all_wata <- load_all_water_sources()
}
```


```{r test-water-distance}
#| eval: true
if ((!work_offline) && reload) {
  test_data <- data_raw %>%
    filter(region == "WES") %>%
    select(cluster, region, idx, code, x, y) %>% 
    distinct(.keep_all = TRUE) 
  test_sf <- st_as_sf(test_data, coords = c("x", "y"), crs = 31370)
  
  cr <- find_center_and_radius(sf::st_coordinates(test_sf))
  
  test_wata <- narrow_sources_radius(
    all_wata,
    cr$center,
    radius = cr$radius * 1.01 + 100 # add 100m safety
  )
  
  water_distances <- collect_combined_distances(test_sf, test_wata, point_index_col = "code")
  min_water_distances <- minimum_cross_distance(water_distances)
  
  
  test_sf <- test_sf %>% left_join(
    min_water_distances,
    by = join_by(code == idx),
    relationship = "one-to-one"
    )
  
  glimpse(test_sf)
}
```


```{r map-water-sources}
#| eval: false
st_geometry(test_wata$surfaces) <- "geometry"

m1 <- test_wata$surfaces %>%
    select(polygon_id) %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c()
  )

m2 <- test_wata$courses %>%
    select(polygon_id) %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c()
  )

m3 <- test_sf %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c("min_dist", "code", "cluster")
  )

m1 + m2 + m3

```


```{r apply-water-distance}

query_min_water_distances_clusterwise <- function (cluster_idx) {

  # subset of the data
  sub_data <- data_raw %>%
    filter(cluster == cluster_idx) %>%
    select(code, x, y) %>% 
    distinct(.keep_all = TRUE) 
  sub_data <- st_as_sf(sub_data, coords = c("x", "y"), crs = 31370)
  
  # center and radius
  cr <- find_center_and_radius(sf::st_coordinates(sub_data))

  # minimum radius
  cr$radius <- max(1000., cr$radius)
  
  # find adjacent water...
  sub_wata <- narrow_sources_radius(
    all_wata,
    cr$center,
    radius = cr$radius * 1.01 + 5000 # add safety margin
  )
  

  # ... or not. 
  if (sum(unlist(lapply(sub_wata, FUN = nrow))) <= 0) {
    # print(paste0("!!! NULL !!! ", cluster_idx))
    return(invisible(NULL))
  }
  
  # calculate water distances
  water_distances <- collect_combined_distances(
    sub_data, sub_wata, point_index_col = "code"
  )

  # get the minimum water distance
  min_water_distances <- minimum_cross_distance(water_distances)

  colnames(min_water_distances) <- c("code", "wata_min_dist", "wata_min_src", "wata_min_idx")
  return(min_water_distances)

}

waterdistance_export_file <- "data/waterdistance_perlocation.parquet"
reload_water <- force_database_download || !file.exists(waterdistance_export_file)

if ((!work_offline) && reload_water) {


  pb <- txtProgressBar(min = 0, max = max(data_raw$cluster), initial = 0, style = 1)
  waterdist_query_pb <- function (cluster_idx) {
    setTxtProgressBar(pb,cluster_idx)
    return(query_min_water_distances_clusterwise(cluster_idx))
  }

  lookup_waterdistance <- lapply(
    sort(unique(data_raw$cluster)),
    FUN = waterdist_query_pb
    )
  close(pb)

  lookup_waterdistance <- bind_rows(lookup_waterdistance) %>% 
    distinct(.keep_all = TRUE) 
  write_parquet(lookup_waterdistance, sink = waterdistance_export_file)
} else {
  lookup_waterdistance <- read_parquet(waterdistance_export_file)
}


if ((!work_offline) && reload) {
  for (col in c("wata_min_dist")) {
    if (col %in% colnames(data_raw)) {
      data_raw <- data_raw %>%
        select(-one_of(col))
    }
  }
  
  data_raw <- data_raw %>% 
    left_join(
      lookup_waterdistance %>%
        select(code, wata_min_dist),
      by = join_by(code == code),
      relationship = "many-to-one"
    )
}

# glimpse(data_raw %>% filter(is.na(wata_min_dist)))
# lookup_waterdistance  %>%
#   filter(code == "MSBP003")
```



## Height TAW 
(wcs query; should match `ref`)

```{r load-terra}
#| eval: true
library("terra")
source("./spatial_helpers.R")

```


```{r elevation-test-single}
#| eval: true

if ((!work_offline) && reload) {
  test_data <- data_raw %>%
    filter(region == "WES") %>%
    select(cluster, region, idx, code, x, y) %>% 
    distinct(.keep_all = TRUE) 

  # xy <- test_data
  # as.matrix(xy[c("x", "y")])
  
  m <- 12.5
  test_elevation <- inbospatial::get_coverage_wcs(
    wcs = "dhmv",
    bbox = find_bbox_for_points(xy = test_data[27,], margin = m),
    layername = "dhmvii_dtm_1m",
    version = "2.0.1",
    #version = "1.0.0",
    wcs_crs = "EPSG:31370",
    resolution = 1
    )
  # note: version 1.0.0 seems much faster
  
  plot(test_elevation, col = gray.colors(256))
  
  mat <- as.matrix(test_elevation, wide = TRUE)
  i <- as.integer(m+0.5)
  h <- mat[i,i] # elevation!
  z <- as.vector(mat)
  x <- rep(1:(2*m), each = 2*m)-m-0.5 # the geographical x -> columns
  y <- rep(1:(2*m), times = 2*m)-m-0.5 # the geographical y -> rows
  # print(cbind(x, y, z))
  
  ea <- prcomp(cbind(x, y, z))
  # attributes(ea)
  slope <- norm(ea$rotation[1:2,3], type = "2")
  
  print(paste(h, slope))
}
```

```{r elevation-test-region}
#| eval: true

normed <- function(vec) vec / norm(vec, type = "2")

get_elevation <- function (xy) {

  margin <- 2.5
  elevation <- inbospatial::get_coverage_wcs(
    wcs = "dhmv",
    bbox = sf::st_bbox(
      c(xmin = xy[1]-margin, xmax = xy[1]+margin,
        ymin = xy[2]-margin, ymax = xy[2]+margin
        ), crs = sf::st_crs(31370)),
    layername = "DHMVII_DTM_1m",
    version = "1.0.0",
    wcs_crs = "EPSG:31370",
    resolution = 1
    )

  mat <- as.matrix(elevation, wide = TRUE)
  i <- as.integer(margin+0.5)
  h <- mat[i,i] # elevation!

  z <- as.vector(mat)
  x <- rep(1:(2*margin), each = 2*margin)-margin-0.5 # the geographical x -> columns
  y <- rep(1:(2*margin), times = 2*margin)-margin-0.5 # the geographical y -> rows
  ea <- prcomp(cbind(x, y, z))
  # get the x, y component of the third eigenvector
  slope <- norm(normed(ea$rotation[,3])[1:2], type = "2")

  return(list("h" = h, "slope" = slope))
}

if ((!work_offline) && reload) {
  ele <- function (i) get_elevation(as.vector(as.matrix(test_data[i, c("x", "y")])))
  bind_rows(lapply(1:nrow(test_data), FUN=ele))
}
```


```{r query-and-join-elevation}

elevation_export_file <- "data/elevation_perlocation.parquet"
reload_elevation <- force_database_download || !file.exists(elevation_export_file)

if ((!work_offline) && reload_elevation) {

  distinct_locations <- data_raw %>%
    # filter(cluster == 16) %>%
    select(code, x, y) %>% 
    distinct(.keep_all = TRUE) 

  pb <- txtProgressBar(
    min = 0, max = nrow(distinct_locations),
    initial = 0, style = 1) 
  elevation_progress <- function(i) {
    setTxtProgressBar(pb,i)
    return(get_elevation(as.vector(as.matrix(distinct_locations[i, c("x", "y")]))))
  }
  all_elevations <- lapply(
    1:nrow(distinct_locations),
    FUN = elevation_progress
    )

  close(pb)

  lookup_elevations <- cbind(distinct_locations, bind_rows(all_elevations))
  write_parquet(lookup_elevations, sink = elevation_export_file)
} else {
  lookup_elevations <- read_parquet(elevation_export_file)
}


if ((!work_offline) && reload) {
  for (col in c("h", "slope")) {
    if (col %in% colnames(data_raw)) {
      data_raw <- data_raw %>%
        select(-one_of(col))
    }
  }
  
  data_raw <- data_raw %>% 
    left_join(
      lookup_elevations %>%
        select(code, h, slope),
      by = join_by(code == code),
      relationship = "many-to-one"
    )
}

```


```{r plot-elevation-vs-ref}
#| eval: true
if ((!work_offline) && reload) {
  data_raw %>%
    ggplot(aes(x = ref, y = h)) +
    geom_point(alpha = 0.1)
}
```


```{r map-elevation-vs-ref}
#| eval: false

ref_check <- data_raw %>%
  filter(abs(ref - h) > 4) %>% 
  select(region, idx, code, x, y, ref, h ) %>% 
  distinct(.keep_all = TRUE)  %>% 
  arrange(desc(abs(ref-h))) %>% 
  rename(ReferentieNiveauTAW = ref, dhmv_elevation = h, MeetpuntWID = idx) 

write.csv(ref_check,"data/elevation_reference_discrepancy.csv")

ref_check %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370) %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c("dhmv_elevation", "ReferentieNiveauTAW")
  )
```



```{r map-elevation}
#| eval: false
data_raw %>%
  # filter(region == "KAL") %>% 
  select(code, x, y, h, slope ) %>% 
  distinct(.keep_all = TRUE) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = 31370) %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c("h", "slope")
  )

```

## finalize data loading

```{r re-load-data}
if ((!work_offline) && reload) {
  data_raw <- data_raw %>% 
    mutate(
      region = substring(code, 1, 3)
    )
  write_parquet(data_raw, sink = data_export_file)
}
data <- read_parquet(data_export_file)

data <- data %>%
  mutate_at(
    vars(
      idx, code, region, hyear, soilclass, cluster
    ), as.factor) %>%
  mutate_at(
    vars(
      x, y, hw, hwd2_cl, lw, lwd2_cl, z, ref, h, slope, wata_min_dist
    ), function (col) units::set_units(col, NULL)) %>%
  mutate(
    filterdepth = z,
    elevation = h
  )

```


# Analysis (I): variogram

## difference data
```{r raw-data-join-dbplyr}
diffdata_export_file <- "data/xg3_full_diffdata.parquet"
reload <- TRUE || force_database_download || !file.exists(diffdata_export_file)

if (reload) {
  diffdata_raw <- data %>%
    left_join(data,
      by = join_by(
        idx > idx, # only different locations
        cluster == cluster, # from the same cluster
        hyear == hyear # in the same year
      ),
      suffix = c("1", "2")
    ) %>% mutate(
      dx = sqrt((x2-x1)^2 + (y2-y1)^2),
      dlg3 = abs(lw2 - lw1),
      dlg3_cl = abs(lwd2_cl2 - lwd2_cl1),
      dhg3 = abs(hw2 - hw1),
      dhg3_cl = abs(hwd2_cl2 - hwd2_cl1)
    ) %>%
    filter(
      dx < extent
      ) 

  write_parquet(diffdata_raw, sink = diffdata_export_file)
} else {
  diffdata_raw <- read_parquet(diffdata_export_file)
}

# glimpse(diffdata_raw)
diffdata_raw %>% count
```


```{r raw-data-processing}
diffdata <- diffdata_raw %>%
  select(
    idx1, code1, x1, y1, 
    lw1, lwd2_cl1, hw1, hwd2_cl1,
    idx2, code2, x2, y2, 
    lw2, lwd2_cl2, hw2, hwd2_cl2,
    hyear, dx
  ) %>%
  filter_at(
    vars(
    idx1, code1, x1, y1, 
    lw1, lwd2_cl1, hw1, hwd2_cl1,
    idx2, code2, x2, y2, 
    lw2, lwd2_cl2, hw2, hwd2_cl2,
    hyear, dx
    ),
    all_vars(!is.na(.))
  )

diffdata <- diffdata %>%
  mutate(
    region = substring(code1, 1, 3),
    dx = sqrt((x2-x1)^2 + (y2-y1)^2),
    distance_bin = cut(dx,
      breaks = seq(
        -extent/n_bins,
        extent,
        length.out = n_bins+2
      ) # cut to "meters" bin
    ),
    dx_log = log10(dx + 1),
    dlw = lw2-lw1,
    dhw = hw2-hw1,
    dlwd = lwd2_cl2 - lwd2_cl1, # cluster-wise de-trended difference
    dhwd = hwd2_cl2 - hwd2_cl1 # cluster-wise de-trended difference
  ) %>% 
  arrange(idx1, idx2) %>% 
  mutate_at(vars(region), as.factor) 
glimpse(diffdata)
```


- **TODO**: include differences in filter depth, elevation, water distance 


## sample size
 
```{r fig-samplesize}
#| label: fig-samplesize
#| fig-cap: "Sample size (number of site pairs observed) per distance (${\\Delta x}$)."
plot_bins <- 128
diffdata %>% 
  mutate(
    distance_bin = cut(dx,
      breaks = seq(
        -extent/plot_bins,
        extent,
        length.out = plot_bins+2)
    ), # more coarse binning than above (better vis)
  ) %>% 
  group_by(distance_bin) %>% 
  summarize(cnt = n()) %>% 
  ungroup %>% 
  arrange(distance_bin) %>% 
  mutate(distance = extent/n_bins * as.numeric(distance_bin)) %>% 
  ggplot(aes(x = distance, y = cnt)) +
  geom_bar(stat = "identity", width = 5) +
  scale_x_continuous(breaks = seq(0, extent, 64)) +
  xlab("distance (m)") +
  ylab("number of observations (N)") +
  theme_bw()
```


```{r plot-data-unbinned}
#| eval: true
#| label: fig-dw-dx
#| fig-cap: "Distance-Difference-Plot of cluster-detrended absolute difference in xG3 (m) against distance (${\\Delta x}$)."
#| layout-ncol: 2
#| fig-subcap: 
#|   - "LG3."
#|   - "HG3."

diffdata %>%
  ggplot(aes(x = dx, y = abs(dlwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

diffdata %>%
  ggplot(aes(x = dx, y = abs(dhwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

```


## binned data
*There must be a reason why 9 out of 10 variogram experts apply binning.*


```{r bin-data}
nanmean <- function (val) mean(val, na.rm = TRUE)
nanvar <- function (val) var(val, na.rm = TRUE)
# nansem <- function (val) sd(val, na.rm = TRUE)/sqrt(sum(!is.na(val)))
diffdata_binned <- diffdata %>%
  group_by(distance_bin) %>% 
  summarize(
    vlg3 = 0.5 * nanvar(dlw),
    vhg3 = 0.5 * nanvar(dhw),
    vlg3d = 0.5 * nanvar(dlwd),
    vhg3d = 0.5 * nanvar(dhwd),
    dx = nanmean(abs(dx)), 
    dlg3  = nanmean(dlw),
    dhg3  = nanmean(dhw),
    dlg3d = nanmean(dlwd),
    dhg3d = nanmean(dhwd),
    cnt = n()
  ) %>% ungroup %>% arrange(dx)
glimpse(diffdata_binned)
```

## self-made variogram
```{r variogram}
#| label: fig-variogram
#| fig-cap: "variogram of difference in xG3 against binned distance ($\\Delta x$)."
#| layout-ncol: 2
#| fig-subcap: 
#|   - "LG3."
#|   - "HG3."


trafo <- function(x) x
# trafo <- function(x) log(x)

diffdata_binned %>%
  ggplot(aes(x = dx, y = trafo(vlg3d))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended LG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none")


diffdata_binned %>%
  ggplot(aes(x = dx, y = trafo(vhg3d))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended HG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none")


```




# Analysis (II): Modeling with `sdmTMB`

## single cluster tryout

```{r subset-test-data}
  test_data <- data %>%
    filter(region == "WES")
```


```{r sdmTMB-mesh-test}
#| eval: true
test_mesh <- make_mesh(test_data, xy_cols = c("x", "y"), cutoff = 1)# as.integer(extent / n_bins))
print(test_mesh$mesh$n)
plot(test_mesh)

```


```{r sdmTMB-model-test}
#| eval: true
test_fit <- sdmTMB(
  hwd2_cl ~ filterdepth + elevation + soilclass + wata_min_dist,
  data = test_data,
  mesh = test_mesh,
  family = gaussian(link = "identity"),
  # time = "hyear",
  # spatiotemporal = "IID",
  spatial = "on"
)
test_fiti <- run_extra_optimization(test_fit)
sanity(test_fiti)
test_fiti
```


## full data set

### meshing Flanders

```{r sdmTMB-mesh-full}
#| eval: true
vlaanderen_mesh <- make_mesh(data, xy_cols = c("x", "y"), cutoff = as.integer(extent / n_bins))
print(vlaanderen_mesh$mesh$n)
plot(vlaanderen_mesh)
```


### LG3 {#sec-lg3}

`lwd2_cl` means:

- LG3
- cluster-wise de-trended

```{r sdmTMB-model-lg3}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd2_cl ~ filterdepth + elevation + soilclass + wata_min_dist,
  data = data,
  mesh = vlaanderen_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3i <- run_extra_optimization(fit_lg3)
sanity(fit_lg3i)
fit_lg3i
```

```{r extract-lg3-parameters}
#| eval: true
knitr::kable(tidy(fit_lg3i))
```

```{r extract-lg3-spatial}
#| eval: true
knitr::kable(tidy(fit_lg3i, "ran_pars"))
```


### HG3 {#sec-hg3}

`hwd2_cl` means:

- HG3
- cluster-wise de-trended


```{r sdmTMB-model-hg3}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd2_cl ~ filterdepth + elevation + soilclass + wata_min_dist,
  data = data,
  mesh = vlaanderen_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3i <- run_extra_optimization(fit_hg3)
sanity(fit_hg3i)
fit_hg3i
```

```{r extract-hg3-parameters}
#| eval: true
knitr::kable(tidy(fit_hg3i))
```

```{r extract-hg3-spatial}
#| eval: true
knitr::kable(tidy(fit_hg3i, "ran_pars"))
```

# Documentation

## Matérn parameterization

[(source)](https://pbs-assess.github.io/sdmTMB/articles/model-description.html#gaussian-random-fields)

The Matérn defines the covariance $\Phi\left( s_j, s_k\right)$ between spatial locations $s_j$ and $s_k$ as

$$\Phi\left( s_j, s_k\right) = \tau^2 / \Gamma\left(\nu\right) 2^{\nu -1} \left(\kappa d_{jk}\right)^\nu K_\nu\left(\kappa d_{jk}\right)$$,
<!-- Φ(sj,sk)=τ2/Γ(ν)2ν−1(κdjk)νKν(κdjk),-->

where $\tau_2$ controls the spatial variance, 
$\nu$ controls the smoothness, 
$\Gamma$ represents the Gamma function, 
$d_{jk}$ represents the distance between locations $s_j$ and $s_k$, 
$K_\nu$ represents the modified Bessel function of the second kind, 
and $\kappa$ represents the decorrelation rate.
The parameter $\nu$ is set to $1$ to take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GRF 
to greatly increase computational efficiency [@Lindgren2011].
Internally, the parameters $\kappa$ and $\tau$ are converted to range and marginal standard deviation $\sigma$ as 
$$range=\frac{8}{\kappa}$$
and 
$$\sigma=\left(4\pi e^{2log(\tau )+2log(\kappa)}\right)^{-\frac{1}{2}}$$
.


# References

::: {#refs}
:::



# Archive

```{r map-clusters}
#| eval: false
data %>%
  # filter(region == "KAL") %>% 
  select(x, y, cluster, z, ref, soilclass ) %>% 
  distinct(.keep_all = TRUE) %>%
  sf::st_as_sf(coords = c("x", "y"),
    crs = 31370
  ) %>% 
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = c("cluster", "z", "ref", "soilclass")
  )

```


```{r disconnect-watina, eval=FALSE}
#| include: true
if (!work_offline) {
  DBI::dbDisconnect(watina_dwh)
}
```


```{r watina-columnlist}
#| eval: false
query <- function(query_string) {
  rs <- DBI::dbGetQuery(watina_dwh, query_string)
  return(rs)
}

get_tablecolumns <- function(tbl_name) {
  my_columns <- query(paste0(
    "SELECT OBJECT_NAME(object_id), name, column_id, TYPE_NAME(system_type_id)
     FROM sys.columns
     WHERE (object_id = OBJECT_ID('",
    tbl_name, "')) ;")
  )
  return(my_columns)
}


col_list <- get_tablecolumns("dimPeilpunt")
col_list
```


```{r get-tablecols}
#| eval: false
get_tablecolumns("dimParameterSet")
head(tbl(watina_dwh, "dimParameterSet"))
get_tablecolumns("factBRPeilMetingJaar")
head(tbl(watina_dwh, "factBRPeilMetingJaar"))
factBRPeilMetingJaar
```


```{r find-missing-peilpunten}
#| eval: false
wesp_pp <- tbl(watina_dwh, "dimPeilpunt") %>%
  filter(PeilpuntCode %like% "WESP*")
wesp_pp
```

```{r match-meetpunten-peilpunten}
#| eval: false
pp_check <- tbl(watina_dwh, "dimPeilpunt") %>%
  pull("MeetpuntWID")
mp_check <- tbl(watina_dwh, "dimMeetpunt") %>%
  pull("MeetpuntWID")


DBI::dbGetQuery(watina_dwh,
  paste0("
SELECT * FROM (
    SELECT DISTINCT
        MP.MeetpuntWID AS mp_idx,
        PP.MeetpuntWID AS pp_idx
    FROM dimMeetpunt AS MP
    FULL OUTER JOIN (
        SELECT DISTINCT MeetpuntWID
        FROM dimPeilpunt 
    GROUP BY MeetpuntWID
    ) AS PP ON MP.MeetpuntWID = PP.MeetpuntWID
    GROUP BY MP.MeetpuntWID, PP.MeetpuntWID
) AS Locations
WHERE (mp_idx IS NULL) OR (pp_idx IS NULL)
ORDER BY mp_idx ASC
;")
)

```
