---
title: "Grand Data Aggregation"
format:
  html:
    toc: true
    html-math-method: katex
  pdf:
    fig-pos: 'H'
knitr:
  opts_chunk:
    echo: true
---


This notebook automates the procedure of loading data from various sources.

It starts at the `watina` locations, and then joins the following information:

- clusters (based on location)
- elevation (via `inbospatial`, from DHMV WCS)
- soilclass (from `n2khab::soilmap_simple`)
- distance from water (various `n2khab` data sources)



```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files

source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# connect watina 
watina_dwh <- connect_watina()

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

```


```{r user-settings}

cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 16 # minimal cluster member count; fewer memeber clusters are removed


# where data is stored
data_export_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet")

```


# Watina Locations

## raw locations

First, query the main data per location from `watina` database.

Here, we focus on `xG3` data.


This query is wrapped in a function for re-use here and portability elsewhere:
```{r raw-data-query-xg3}
#| eval: true

query_xg3 <- function (){
  tbl(watina_dwh, "factBRPeilMetingJaar") %>%
    filter(
      IsHydroJaar == 1,
      ParameterSetWID == 1
    ) %>% 
    select(
      MeetpuntWID,
      Jaar,
      HGmTAWValue,
      LGmTAWValue,
      VGmTAWValue
    ) %>%
    filter_at(
      vars(HGmTAWValue, LGmTAWValue),
      all_vars(!is.na(.))
    ) %>% 
    left_join(
      tbl(watina_dwh, "dimMeetpunt"),
      by = join_by(MeetpuntWID == MeetpuntWID),
      relationship = "many-to-many"
    ) %>%
    filter(MeetpuntStatusCode == "VLD") %>%
    select(
      MeetpuntWID,
      MeetpuntCode,
      Jaar,
      MeetpuntXCoordinaat,
      MeetpuntYCoordinaat,
      HGmTAWValue,
      LGmTAWValue,
    ) %>%
    rename(
      idx = MeetpuntWID,
      code = MeetpuntCode,
      hyear = Jaar,
      x = MeetpuntXCoordinaat,
      y = MeetpuntYCoordinaat,
      hw = HGmTAWValue,
      lw = LGmTAWValue
    ) %>%
   filter_at(
    vars(code, hyear, hw, lw),
    all_vars(!is.na(.))
  )

}

```


Raw data is collected by executing the function:

```{r query-data}
#| eval: true
data_raw <- query_xg3() %>%
  select(idx, code, hyear, x, y, hw, lw) %>% 
  filter_at(
    vars(x, y, hw, lw),
    all_vars(!is.na(.))
  ) %>% 
  collect %>%
  mutate(
    region = substring(code, first = 1, last = 3)
  ) %>%
  filter(region != "VBS") # remove Luxemburg

glimpse(data_raw)
```


## observation wells

Some information is stored in the `dimPeilpunt` table, e.g. filter depth.
It is queried and joined manually.

```{r query-filter-depth}
#| eval: true
pp <- tbl(watina_dwh, "dimPeilpunt") %>%
  filter(
    OpenbaarheidWID == 4 # note: about O(150) locs are non-public
  ) %>%
  mutate(
    yr = sql("YEAR(CAST(PeilpuntPlaatsing AS date))"),
    month = sql("YEAR(CAST(PeilpuntPlaatsing AS date))"),
    start_hyear = yr + (if (month >= 10) 1 else 0),
    yr = sql("YEAR(CAST(PeilpuntStopzetting AS date))"),
    month = sql("YEAR(CAST(PeilpuntStopzetting AS date))"),
    stop_hyear = yr + (if (month >= 10) 1 else 0),
    ref = ReferentieNiveauTAW,
    z = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
  ) %>%
  rename(
    code = PeilpuntCode,
    idx = MeetpuntWID
  ) %>%
  select(idx, code, start_hyear, stop_hyear, z, ref) %>%
  collect %>%
  mutate(
    region = substring(code, first = 1, last = 3) 
  ) %>% 
  filter(region != "VBS") %>%  # remove Luxemburg
  group_by(idx, start_hyear, stop_hyear) %>%
  slice(which.min(z)) 

# Note: 15964 -> 10389 rows??
#       filter depth z duplicates removed by `slice`
#       replacement "by date" is hard to cover in this analysis
#       because we work on xG3 and they are not associated with a date
#       decided to just take the shallower value per location/year

glimpse(pp)
```



```{r join-peilpunt}
#| eval: true

for (col in c("z", "ref")) {
  if (col %in% colnames(data_raw)) {
    data_raw <- data_raw %>%
      select(-one_of(col))
  }
}


data_raw <- data_raw %>%
  inner_join(pp,
    by = join_by(
      idx == idx,
    ),
    # relationship = "many-to-many"
    # unmatched = c("drop", "drop"),
    relationship = "many-to-many",
    suffix = c("", ".duplicates")
  ) %>%
  select(-ends_with(".duplicates")) %>% 
  filter(
    hyear >= start_hyear, # the observation must be after loc startup
    (is.na(stop_hyear) | (hyear < stop_hyear)) # ... or loc is up and running
  ) %>%
  select(-start_hyear, -stop_hyear)

# 28k -> 22k rows due to unmatched
glimpse(data_raw )
```


# Auxiliary Data

All the other required data can be appended easily, thanks to the auxiliary data query scripts.

```{r join-auxiliary-data}
data <- data_raw %>%
  join_clusters(characteristic_distance = cluster_dist) %>%
  remove_underpopulated_clusters(
    minimum_cluster_member_count = min_cluster_size) %>%
  join_elevation() %>%
  join_soilclass() %>%
  join_waterdistance(cluster_column = "idx")

write_parquet(data, sink = data_export_file)
dplyr::glimpse(data)
```


After this, data is stored in the file "`{r} data_export_file`" and can be loaded with: `data <- read_parquet("data/analysis_data_cl3400.parquet")`.

