---
title: "Water Level Variograms and Models"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs variogram analysis.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

```


Some parameters can be chosen by the user.

```{r data-settings}
#| eval: true

extent <- 1024 # m
n_bins <- 256

# result of previous analysis
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored
```

# Data Preparation
## load data

The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
A variogram is defined as a quantification of the semivariance of a difference measure depending on the (binned) spatial distance of points.
:::


We define the **semivariance** $\gamma$ [@Matheron1962, modified]:

$$ \gamma\left( h\right) = \frac{1}{2} \underset{h}{\mathrm{Var}} \left(z_j - z_i\right) \quad \forall \left\lbrace i, j\right\rbrace :\ \mid s_j-s_i \mid \approx h $$

Containing the following parameters:

- $N$ is the number of observation pairs $\{i, j\}$
- with spatial positions $s_{(\cdot)} = \left(x_{(\cdot)} , y_{(\cdot)}\right)$
- and measured values $z_{(\cdot)}$
- grouped (binned) to distance groups $h$; $h$ is also called the "lag vector magnitude", effectively the Euclidean distance
- $Var$ being the variance within bins of similar $h$
- $\gamma\left(h\right)$ is the semivariance (because $2\gamma$ is the variance)


Usually, the first step is to visualize the data in a difference-distance plot, further deriving a "semivariance-binned distance" plot.
To the latter, one can fit a Matérn-class function for parametrization and comparison.
Or one can apply more complex models.


Our raw data of choice are `xG3` values, which are `LG3` for a lower- and `HG3` for an upper reference of water levels [@ritzema2012].
<!-- see also /data/R/quickanalyses/HG3_timeseries.org -->

They are defined as follwos:

> Gemiddelde van de drie laagste|hoogste grondwaterstanden in een hydrologisch jaar (1 april t/m 31 maart) bij een meetfrequentie van tweemaal per maand (rond de 14e en 28e).

This is literally how they are calculated.


## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


## difference data
To efficiently calculate variogram, we first require difference data.
This means we have to combine a set of measurements with itself.

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- In contrast, a **prefix `d`** identifies the difference: e.g. `ds` for a difference in position, or `dlwd` for a difference in clusterwise de-trended `lg3`.
- There is the **prefix `v`**, indicating a variance calculation (e.g. `vhgd`, variance in clusterwise de-trended HG3).

:::


::: {.callout-warning}
I only join locations which are categorized under the same

- region and cluster
- soilclass
- hydroyear

:::


```{r raw-data-join-dbplyr}
#| eval: true
diffdata_raw <- data %>%
  left_join(data,
    by = join_by(
      idx > idx, # only different locations
      region == region, # from the same region
      cluster == cluster, # from the same cluster
      soilclass == soilclass, # from the same soilclass
      hyear == hyear # in the same year
    ),
    suffix = c("1", "2"),
    relationship = "many-to-many"
  ) %>% mutate(
    ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    dlw = abs(lw2 - lw1),
    dlwd = abs(lwd2 - lwd1),
    dhw = abs(hw2 - hw1),
    dhwd = abs(hwd2 - hwd1)
  ) %>%
  filter(
    ds < extent
    ) 

# glimpse(diffdata_raw)
diffdata_raw %>% count
```


Further processing is required:

- remove `NA`
- binning by distance 
- calculation of difference values
- optional log/logp1 transformation


```{r raw-data-processing}
diffdata <- diffdata_raw %>%
  filter_at(
    vars(
    idx1, code1, x1, y1, 
    lw1, lwd1, hw1, hwd1,
    idx2, code2, x2, y2, 
    lw2, lwd2, hw2, hwd2,
    hyear, ds
    ),
    all_vars(!is.na(.))
  )

diffdata <- diffdata %>%
  mutate(
    distance_bin = cut(ds,
      breaks = seq(
        -extent/n_bins,
        extent,
        length.out = n_bins+2
      ) # cut to "meters" bin
    ),
    dlw = lw2-lw1,
    dhw = hw2-hw1,
    dlwd = lwd2 - lwd1, # cluster-wise de-trended difference
    dhwd = hwd2 - hwd1, # cluster-wise de-trended difference
    dh = elevation_dhmv2 - elevation_dhmv1,
    dslope = slope_r2.5m2 - slope_r2.5m1,
    dw = wata_min_dist2 - wata_min_dist1,
    dz = z2 - z1,
    ds_log = log10(ds + 1)
  ) %>% 
  arrange(idx1, idx2) %>% 
  mutate_at(vars(
    region, code1, code2,
    wata_min_src1, wata_min_src2
  ), as.factor) 
# glimpse(diffdata)
```


## sample size

```{r plot-data-unbinned}
#| eval: true
#| label: fig-dw-ds
#| fig-cap: "Distance-Difference-Plot of cluster-detrended absolute difference in xG3 (m) against distance (${\\Delta s}$)."
#| layout-ncol: 2
#| fig-subcap: 
#|   - "LG3."
#|   - "HG3."

diffdata %>%
  ggplot(aes(x = ds, y = abs(dlwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

diffdata %>%
  ggplot(aes(x = ds, y = abs(dhwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

```

## binning and semivariance

Bins are defined above, but it makes sense to define how gins are compiled in terms of value fields.


::: {.callout-warning}
This is also where semivariance gets calculated.
:::


```{r bin-data}
nanmean <- function (val) mean(val, na.rm = TRUE)
nanvar <- function (val) var(val, na.rm = TRUE)

# a convenience function for calculated measures
summarize_groupeddata <- function(data) {
  data %>% 
    summarize(
      vlg = nanvar(dlw),
      vhg = nanvar(dhw),
      vlgd = nanvar(dlwd),
      vhgd = nanvar(dhwd),
      ds = nanmean(abs(ds)), 
      dh = nanmean(abs(dh)), 
      dslope = nanmean(abs(dslope)), 
      dw = nanmean(abs(dw)), 
      dz = nanmean(abs(dz)), 
      dlg  = nanmean(dlw),
      dhg  = nanmean(dhw),
      dlgd = nanmean(dlwd),
      dhgd = nanmean(dhwd),
      cnt = n()
    ) %>% ungroup %>% arrange(ds)
}


diffdata_binned <- diffdata %>%
  group_by(distance_bin) %>%
  summarize_groupeddata()

diffdata_binned_soilclass <- diffdata %>%
  group_by(soilclass, distance_bin) %>%
  summarize_groupeddata()

knitr::kable(head(diffdata_binned_soilclass, 5))
```


We can plot the sample size per bin:

```{r bin-samplesize}
#| label: fig-soilclass-samplesize
#| code-fold: true
#| fig-cap: "Sample size per soilclass and distance difference."

diffdata_binned_soilclass %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step() +
  theme_bw() 
  
```



Final detail: there will be transformations.

```{r variogram-transformations}
semi <- function(x) 0.5 * x
# trafo <- function(x) log(x)
```
  

# Variogram

## visualization

The first step is to visualize how data changes in bins.
With the data at hand, this is rather straight forward.

::: {.panel-tabset}
### LG3
```{r variogram-lg3}
#| label: fig-variogram-lg3
#| code-fold: true
#| fig-cap: "Variogram of difference in LG3 against binned distance ($\\Delta s$)."


diffdata_binned %>%
  ggplot(aes(x = ds, y = semi(vlgd))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended LG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none") 
```


### LG3
```{r variogram-hg3}
#| label: fig-variogram-hg3
#| code-fold: true
#| fig-cap: "Variogram of difference in LG3 against binned distance ($\\Delta s$)."
diffdata_binned %>%
  ggplot(aes(x = ds, y = semi(vhgd))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended HG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none")


```
:::

As can be seen, the `smooth` interpolation is not too accurate.
We should do better.


## Regression

### strategy
More specifically, we should regress with functions which are

- less sensitive to outliers
- optionally forced through zero (no distance, no difference)
- of different function class, e.g. Matérn or linear breakpoint


To understand the tools we have, start with linear.
Of course, I want a nice and [[https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/optim][slender optimizer, yet universal regression]] toolkit.


### implementation

There are some general convenience wrappers for classical regression.

- https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/optim


```{r regression-tools}

wrap_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  return(sum(differences^2))
}

create_prediction_function <- function(regressor, results) {
  fcn <- function (x) {
    regressor(x, results$par)
  }

  return(fcn)
}

print_regression_results <- function(orsl, label = "") {
  par <- paste(round(orsl$par, 2), collapse = ", ")
  conv <- orsl$convergence
  eps <- orsl$value
  print(sprintf("conv %i at (%s): eps %.1f %s", conv, par, eps, label))
}
```


To test this, behold the non-sensical linear regression!

```{r regression-test}

x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)

linear_function <- function(x, params) {
  return(params[1] + params[2]*x)
}

optimizer_results <- optim(
  par = c(0, 1),
  fn = function(params) {
    wrap_target_function(x, y, linear_function, params)
  }, method = "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "linear")

predictor_function <- create_prediction_function(linear_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```


### application

#### ramp function

A piecewise linear-to-constant function ("ramp") might be useful!


```{r regression-ramp}

x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)

ramp_function <- function(x, params) {
  pre <- x<=params[1]
  post <- !pre
  return(0 +
         params[2] * x * as.integer(pre) +
         params[2] * params[1] * as.integer(post))
}

optimizer_results <- optim(
  par = c(250, mean(y)),
  lower = c(0, 0),
  upper = c(max(x), +Inf),
  fn = function(params) {
    wrap_target_function(x, y, ramp_function, params)
  }, method = "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "ramp-plateau")

predictor_function <- create_prediction_function(ramp_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```

This did not converge well.


#### log-ramp
Yet we could use transformations:


```{r regression-logramp}

x <- diffdata_binned$ds
y <- log(1.+semi(diffdata_binned$vhgd))

logramp_function <- function(x, params) {
  pre <- x<=params[1]
  post <- !pre
  return(0 +
         params[2] * x * as.integer(pre) +
         params[2] * params[1] * as.integer(post))
}

optimizer_results <- optim(
  par = c(350, mean(y)),
  # lower = c(0, 0),
  # upper = c(max(x), +Inf),
  fn = function(params) {
    wrap_target_function(x, y, logramp_function, params)
  }, method = "Nelder-Mead" # "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "ramp-plateau")

predictor_function <- create_prediction_function(logramp_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```

Not *too* bad, but still bad.


#### Matérn-like function

Let's mimic what the pro's do!
What follows is an exact quote [(source)](https://pbs-assess.github.io/sdmTMB/articles/model-description.html#gaussian-random-fields).

The Matérn defines the covariance $\Phi\left( s_j, s_k\right)$ between spatial locations $s_j$ and $s_k$ as

$$\Phi\left( s_j, s_k\right) = \tau^2 / \Gamma\left(\nu\right) 2^{\nu -1} \left(\kappa d_{jk}\right)^\nu K_\nu\left(\kappa d_{jk}\right)$$,
<!-- Φ(sj,sk)=τ2/Γ(ν)2ν−1(κdjk)νKν(κdjk),-->

where $\tau^2$ controls the spatial variance, 
$\nu$ controls the smoothness, 
$\Gamma$ represents the Gamma function, 
$d_{jk}$ represents the distance between locations $s_j$ and $s_k$, 
$K_\nu$ represents the modified Bessel function of the second kind, 
and $\kappa$ represents the decorrelation rate.
The parameter $\nu$ is set to $1$ to take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GRF 
to greatly increase computational efficiency [@Lindgren2011].
Internally, the parameters $\kappa$ and $\tau$ are converted to range and marginal standard deviation $\sigma$ as 
$$range=\frac{8}{\kappa}$$
and 
$$\sigma=\left(4\pi e^{2log(\tau )+2log(\kappa)}\right)^{-\frac{1}{2}}$$
.

See also: https://stats.stackexchange.com/questions/611908/how-to-properly-implement-a-mat%C3%A9rn-kernel-function-in-r 


Should be fun to implement!


```{r regression-matérn}

x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)

# ## standardization helps function comprehension
# x <- x/max(x)
# y <- y / mean(y)

matern_function <- function(d, params) {
  phi <- params[1] # related to range; turning point
  sill <- params[2]^2
  nu <- 1. #params[3]
  d <- d
  z <- sqrt(2*nu)*d/phi
  K <- suppressWarnings(besselK(z, nu))
  result <- sill - sill * (z)^nu * K / (2^(nu-1)*gamma(nu))
  result[is.na(result)] <- 0
  return(result)
}


# for exploration:
# start_values <- c("tau" = mean(y), "kappa" = sqrt(8)/250)
# start_values <- c("phi" = 0.3, "sill" = 10) # , "nu" = 1. 
# params <- start_values
# d <- x
# plot(x, matern_function(x, start_values), type = "l") 

start_values <- c("phi" = mean(x), "sill" = mean(y)) # , "nu" = 1. 
optimizer_results <- optim(
  par = start_values,
  # lower = c(0.05, 0.01),
  # upper = c(max(x), max(y)),
  fn = function(params) {
    wrap_target_function(x, y, matern_function, params)
  }, method = "Nelder-Mead" # "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "Matérn")

predictor_function <- create_prediction_function(matern_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  labs(title = "Matérn regression") +
  theme_minimal()

```


The regression results are `{r} paste(round(optimizer_results$par, 4), collapse = ", ")`.

This can be converted to range and sill.




# References

::: {#refs}
:::
