---
title: "Water Level Variograms and Models"
bibliography: references.bib
format:
  html:
    toc: true
    html-math-method: katex
  pdf:
    fig-pos: 'H'
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs variogram analysis.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

```


Some parameters can be chosen by the user.

```{r data-settings}
#| eval: true

extent <- 1024 # m
n_bins <- 128

# result of previous analysis
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored
```

# Data Preparation
## load data

The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
A variogram is defined as a quantification of the semivariance of a difference measure depending on the (binned) spatial distance of points.
:::


We define the **semivariance** $\gamma$ [@Matheron1962]:

$\gamma\left(h\right) = \frac{1}{2N} \sum\limits_{N(h)} \left(z_j - z_i\right)^2 \quad \forall \left{i, j\right}: \mid s_j-s_i \mid = h$

Containing the following parameters:

- $N$ is the number of observation pairs $\{i, j\}$
- with spatial positions $s_\cdot = \left(x_\cdot , y_\cdot\right)$
- and measured values $z_\cdot$
- grouped (binned) to distance groups $h$; $h$ is also called the "lag vector magnitude", effectively the Euclidean distance
- $\gamma\left(h\right)$ is the semivariance (because $2\gamma$ is the variance)


Usually, the first step is to visualize the data in a difference-distance plot, further deriving a "semivariance-binned distance" plot.
To the latter, one can fit a Matérn-class function for parametrization and comparison.
Or one can apply more complex models.


Our raw data of choice are `xG3` values 
TODO: reference
<https://www.wur.nl/nl/onderzoek-resultaten/onderzoeksinstituten/environmental-research/faciliteiten-producten/software-en-modellen/grondwaterdynamiek.htm>


## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


## difference data
To efficiently calculate variogram, we first require difference data.
This means we have to combine a set of measurements with itself.

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- In contrast, a **prefix `d`** identifies the difference: e.g. `ds` for a difference in position, or `dlwd` for a difference in clusterwise de-trended `lg3`.



:::

```{r raw-data-join-dbplyr}
#| eval: true
diffdata_raw <- data %>%
  left_join(data,
    by = join_by(
      idx > idx, # only different locations
      cluster == cluster, # from the same cluster
      hyear == hyear # in the same year
    ),
    suffix = c("1", "2"),
    relationship = "many-to-many"
  ) %>% mutate(
    ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    dlw = abs(lw2 - lw1),
    dlwd = abs(lwd2 - lwd1),
    dhw = abs(hw2 - hw1),
    dhwd = abs(hwd2 - hwd1)
  ) %>%
  filter(
    dx < extent
    ) 

# glimpse(diffdata_raw)
diffdata_raw %>% count
```

# References

::: {#refs}
:::
