---
title: "Water Level Variograms and Regression"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs variogram analysis.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
# library("sdmTMB") # species distribution models Template Model Builder
# library("mapview") # viewing maps
# library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

```


Some parameters can be chosen by the user.

```{r data-settings}
#| eval: true

# extent <- 1024 # m
# n_bins <- 128
extent <- 1024*2 # m
n_bins <- 256*2

# result of previous analysis
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored
```

# Data Preparation
## load data

The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
A variogram is defined as a quantification of the semivariance of a difference measure depending on the (binned) spatial distance of points.
:::


We define the **semivariance** $\gamma$ [@Matheron1962, modified]:

$$ \gamma\left( h\right) = \frac{1}{2} \underset{N(h)}{\mathrm{Var}} \left(z_j - z_i\right) \quad \forall \left\lbrace i, j\right\rbrace :\ \mid s_j-s_i \mid \approx h $$

Containing the following parameters:

- $N$ is the number of observation pairs $\{i, j\}$ in a given distance bin
- grouped (binned) to distance groups $h$; $h$ is also called the "lag vector magnitude", effectively the Euclidean distance
- with spatial position vectors $s_{(\cdot)} = \left(x_{(\cdot)} , y_{(\cdot)}\right)$
- and measured values $z_{(\cdot)}$ at a given position $s_{(\cdot)}$
- $Var$ being the variance within bins of approximately similar $h$
- $\gamma\left(h\right)$ is the semivariance (because $2\gamma$ is the variance)


Usually, the first step is to visualize the data in a difference-distance plot, further deriving a "semivariance-binned distance" plot.
To the latter, one can fit a Matérn-class function for parametrization and comparison.
Or one can apply more complex models.


Our raw data of choice are `xG3` values, which are `LG3` for a lower- and `HG3` for an upper reference of water levels [@ritzema2012].
<!-- see also /data/R/quickanalyses/HG3_timeseries.org -->

They are defined as follwos:

> Gemiddelde van de drie laagste|hoogste grondwaterstanden in een hydrologisch jaar (1 april t/m 31 maart) bij een meetfrequentie van tweemaal per maand (rond de 14e en 28e).

This is literally how they are calculated.


## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


## difference data
To efficiently calculate variogram, we first require difference data.
This means we have to combine a set of measurements with itself.

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- In contrast, a **prefix `d`** identifies the difference: e.g. `ds` for a difference in position, or `dlwd` for a difference in clusterwise de-trended `lg3`.
- There is the **prefix `v`**, indicating a variance calculation (e.g. `vhgd`, variance in clusterwise de-trended HG3).

:::


::: {.callout-warning}
I only join locations which are categorized under the same

- region and cluster
- soilclass
- hydroyear ("water year", [see here](https://en.wikipedia.org/wiki/Water_year))

If the same pair of locations is measured in different hydrological years, this is treated as independent observations (simplifying assumption, implying that we are more certain about repeatedly observed co-variation).

:::



```{r raw-data-join-dbplyr}
#| eval: true
diffdata_raw <- data %>%
  left_join(data,
    by = join_by(
      idx > idx, # only different locations
      region == region, # from the same region
      cluster == cluster, # from the same cluster
      soilclass == soilclass, # from the same soilclass
      hyear == hyear # in the same year
    ),
    suffix = c("1", "2"),
    relationship = "many-to-many"
  ) %>% mutate(
    ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    dlw = abs(lw2 - lw1),
    dlwd = abs(lwd2 - lwd1),
    dhw = abs(hw2 - hw1),
    dhwd = abs(hwd2 - hwd1)
  ) %>%
  filter(
    ds < extent
    ) 

# glimpse(diffdata_raw)
diffdata_raw %>% count
```


Further processing is required:

- remove `NA`
- binning by distance 
- calculation of difference values
- optional log/logp1 transformation


```{r raw-data-processing}
diffdata <- diffdata_raw %>%
  filter_at(
    vars(
    idx1, code1, x1, y1, 
    lw1, lwd1, hw1, hwd1,
    idx2, code2, x2, y2, 
    lw2, lwd2, hw2, hwd2,
    hyear, ds
    ),
    all_vars(!is.na(.))
  )

diffdata <- diffdata %>%
  mutate(
    distance_bin = cut(ds,
      breaks = seq(
        -extent/n_bins,
        extent,
        length.out = n_bins+2
      ) # cut to "meters" bin
    ),
    dlw = lw2-lw1,
    dhw = hw2-hw1,
    dlwd = lwd2 - lwd1, # cluster-wise de-trended difference
    dhwd = hwd2 - hwd1, # cluster-wise de-trended difference
    dh = elevation_dhmv2 - elevation_dhmv1,
    dslope = slope_r2.5m2 - slope_r2.5m1,
    dw = wata_min_dist2 - wata_min_dist1,
    dz = z2 - z1,
    ds_log = log10(ds + 1)
  ) %>% 
  arrange(idx1, idx2) %>% 
  mutate_at(vars(
    region, code1, code2,
    wata_min_src1, wata_min_src2
  ), as.factor) 
# glimpse(diffdata)
```


## sample size

The database holds quite a number of locations, and if these are paired, an enormous data set emerges.
It is crucial to keep track of the sample size, as a function of distance.


```{r plot-data-unbinned}
#| eval: true
#| label: fig-dw-ds
#| fig-cap: "Distance-Difference-Plot of cluster-detrended absolute difference in xG3 (m) against distance (${\\Delta s}$)."
#| layout-ncol: 2
#| fig-subcap: 
#|   - "LG3."
#|   - "HG3."

diffdata %>%
  ggplot(aes(x = ds, y = abs(dlwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

diffdata %>%
  ggplot(aes(x = ds, y = abs(dhwd))) +
  geom_point(alpha = 0.05) +
  xlab("distance (m)") +
  ylab("de-trended hg3 difference (m)") +
  theme_bw()

```

This overview is concretized below, after binning.


## binning and semivariance

Bins are defined above (bins of `{r} extent / n_bins` meters), but it makes sense to define how bins are compiled in terms of value fields (averaging!).


::: {.callout-note}
This is also where semivariance gets calculated.
:::


```{r bin-data}
nanmean <- function (val) mean(val, na.rm = TRUE)
nanvar <- function (val) var(val, na.rm = TRUE)

# a convenience function for calculated measures
summarize_groupeddata <- function(data) {
  data %>% 
    summarize(
      vlg = nanvar(dlw),
      vhg = nanvar(dhw),
      vlgd = nanvar(dlwd),
      vhgd = nanvar(dhwd),
      ds = nanmean(abs(ds)), 
      dh = nanmean(abs(dh)), 
      dslope = nanmean(abs(dslope)), 
      dw = nanmean(abs(dw)), 
      dz = nanmean(abs(dz)), 
      dlg  = nanmean(dlw),
      dhg  = nanmean(dhw),
      dlgd = nanmean(dlwd),
      dhgd = nanmean(dhwd),
      cnt = n()
    ) %>% ungroup %>% arrange(ds)
}


diffdata_binned <- diffdata %>%
  group_by(distance_bin) %>%
  summarize_groupeddata()

diffdata_binned_soilclass <- diffdata %>%
  group_by(soilclass, distance_bin) %>%
  summarize_groupeddata()

knitr::kable(tail(diffdata_binned_soilclass, 10))
```


We can plot the sample size per bin.

::: { .callout-tip }
The relevance of binwise samplesize is not to be underestimated:

- sufficient bin size is essential for variance calculation (e.g. too small bins will have lower/too low variance)
- bin width affects the degree of outlier averaging (there are "outliers", but we currently have no logical reason to exclude them)
- the number of bins is the number of data points for regression (below); the more, the better

Note in particular how we have many observations from "heavy" soil at low distance (dense sampling of heavy soil regions?), whereas "light" soil data ramps up; both curves are zero-inflated.
On the other end, sample size diminishes above a kilometer distance, probably due to cluster size limit.

:::


```{r bin-samplesize}
#| label: fig-soilclass-samplesize
#| code-fold: true
#| fig-cap: "Sample size per soilclass and distance difference."

diffdata_binned_soilclass %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step() +
  theme_bw() 
  
```


Final detail: there will be transformations.
Yet I focus on the trivial step from "variance" to "semivariance".

```{r variogram-transformations}
semi <- function(x) 0.5 * x
# log_trafo <- function(x) log(x)
```
  

# Difference-Distance-Plot

## visualization

The first step is to visualize how data changes in bins.
With the data at hand, this is rather straight forward.

::: {.panel-tabset}
### LG3
```{r variogram-lg3}
#| label: fig-variogram-lg3
#| code-fold: true
#| fig-cap: "Variogram of difference in LG3 against binned distance ($\\Delta s$)."


diffdata_binned %>%
  ggplot(aes(x = ds, y = semi(vlgd))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended LG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none") 
```


### HG3
```{r variogram-hg3}
#| label: fig-variogram-hg3
#| code-fold: true
#| fig-cap: "Variogram of difference in LG3 against binned distance ($\\Delta s$)."
diffdata_binned %>%
  ggplot(aes(x = ds, y = semi(vhgd))) +
  geom_point(alpha = 0.2) +
  geom_smooth(color = 'darkblue') +
  xlab("site distance (m), binned") +
  ylab("semivariance in de-trended HG3: var(lg3)/2 (m²)") +
  theme_bw() +
  theme(legend.position = "none")


```
:::


::: { .callout-note }
Insights:

- Variation of variance is surprisingly large: it is hard to see clear patterns, despite the considerable sample size.
- Another take-home message: differences between `LG3` and `HG3` values seem subtle.
- As can be seen, the `smooth` interpolation is not too accurate. We should do better (regression).

:::


## Regression

### strategy
More specifically, we should regress with functions which are

- less sensitive to outliers
- optionally forced through zero (no distance, no difference)
- of different function class, e.g. Matérn or linear breakpoint


To understand the tools we have, I will start "linear", then increase complexity.


### implementation

There are some general convenience wrappers for classical regression.
However, [base-r `optim` does all we need](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/optim).

We choose between well-known "Nelder-Mead" optimization algorithm [@nelder1965], or the more versatile ["L-BFGS-B"](https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.minimize.html) [@zhu1997, @byrd1995].
They are interchangeable, for our practical purpose, except that the latter allows to define logical parameter boundaries to facilitate optimization convergence.


```{r regression-tools}

# wrap a regression function to generate residuals
# the result is the parameter to be minimized.
wrap_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  return(sum(sqrt(differences^2)))
}


# this can turn regression output into a usable function.
create_prediction_function <- function(regressor, results) {
  fcn <- function (x) {
    regressor(x, results$par)
  }

  return(fcn)
}

# a uniform way to print results
print_regression_results <- function(orsl, label = "") {
  par <- paste(round(orsl$par, 4), collapse = ", ")
  conv <- orsl$convergence
  eps <- orsl$value
  print(
    sprintf("%s: conv %i at (%s), mse %.1f", label, conv, par, eps)
  )
}

# note: the `epsilon` can be calculated manually with the formula:
#   sum((predictor_function(x) - y)^2)


# Finally, a quick histogram plot of residuals.
plot_residuals_histogram <- function (x, y, predictor_function, ...) {
  residuals <- predictor_function(x) - y
  ggplot(NULL, aes(x = residuals)) +
    geom_histogram(...) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    theme_bw()
}


```


To test this, behold the non-sensical linear regression!

```{r regression-test}

x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)

linear_function <- function(x, params) {
  return(params[1] + params[2]*x)
}

optimizer_results <- optim(
  par = c(0, 1),
  fn = function(params) {
    wrap_target_function(x, y, linear_function, params)
  }, method = "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "linear")

predictor_function <- create_prediction_function(linear_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```


Observations:

- The regression fits the data more or less well, quantified by the mean square error (`mse`).
- Optimizer did converge (`conv 0`, [see "convergence" here](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/optim.html)), which should not be overrated (the regression might still be irrelevant).
- Parameters can be measured, in this case intercept (`{r} round(optimizer_results$par[1], 2)`) and slope (`{r} round(optimizer_results$par[2], 4)`).

We can still do better.



### application

#### ramp function

A piecewise linear-to-constant function ("ramp") might be useful!


```{r regression-ramp}

x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

ramp_function <- function(x, params) {
  pre <- x <= params[1]
  post <- !pre
  return(0 +
         params[2] * x * as.integer(pre) +
         params[2] * params[1] * as.integer(post))
}

optimizer_results <- optim(
  par = c(250, mean(y)),
  lower = c(0, 0),
  upper = c(max(x), +Inf),
  fn = function(params) {
    wrap_target_function(x, y, ramp_function, params)
  }, method = "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "ramp-plateau")

predictor_function <- create_prediction_function(ramp_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```

This did converge, though values are practically zero.
Switch point is approximately `{r} round(optimizer_results$par[1], 0)`, and slope in the first section is `{r} round(optimizer_results$par[2], 4)`.

Note that, with other settings for `extent` and `n_bins`, this one was much more useful.
As it is now, with a local optimum at `(0, 0)`, the ramp function is a mere null reference for `mse`.


#### log-ramp
Yet we could use transformations:


```{r regression-logramp}

x <- diffdata_binned$ds
y <- log(1.+semi(diffdata_binned$vhgd))
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

logramp_function <- function(x, params) {
  pre <- x<=params[1]
  post <- !pre
  return(params[3] +
         params[2] * x * as.integer(pre) +
         params[2] * params[1] * as.integer(post))
}

optimizer_results <- optim(
  par = c(350, mean(y), 1),
  # lower = c(0, 0),
  # upper = c(max(x), +Inf),
  fn = function(params) {
    wrap_target_function(x, y, logramp_function, params)
  }, method = "Nelder-Mead" # "L-BFGS-B"
)

print_regression_results(optimizer_results, label = "log ramp-plateau")

predictor_function <- create_prediction_function(logramp_function, optimizer_results)

ggplot(NULL, aes(x = x, y = y)) +
  geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.4) +
  geom_line(aes(x = x, y = predictor_function(x))) +
  theme_minimal()

```

Not *too* bad, but still bad.
For example, the zero intercept made no sense, but introducing an intercept leaves the first section rather overmodeled with the limited number of bins.

Crucial limitation is that the data lacks a consistent pattern, even in log-space.


#### Matérn-like function

Let's mimic what the pro's do!

::: {.callout-tip}
There are logical reasons for choosing Matérn, namely the assumption of Gaussian interdependence.
:::


What follows is an exact quote [(source)](https://pbs-assess.github.io/sdmTMB/articles/model-description.html#gaussian-random-fields).

`<quote>`

The Matérn defines the covariance $\Phi\left( s_j, s_k\right)$ between spatial locations $s_j$ and $s_k$ as

$$\Phi\left( s_j, s_k\right) = \tau^2 / \Gamma\left(\nu\right) 2^{\nu -1} \left(\kappa d_{jk}\right)^\nu K_\nu\left(\kappa d_{jk}\right)$$
<!-- Φ(sj,sk)=τ2/Γ(ν)2ν−1(κdjk)νKν(κdjk),-->

where $\tau^2$ controls the spatial variance, 
$\nu$ controls the smoothness, 
$\Gamma$ represents the Gamma function, 
$d_{jk}$ represents the distance between locations $s_j$ and $s_k$, 
$K_\nu$ represents the modified Bessel function of the second kind, 
and $\kappa$ represents the decorrelation rate.
The parameter $\nu$ is set to $1$ to take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GRF 
to greatly increase computational efficiency [@Lindgren2011].
Internally, the parameters $\kappa$ and $\tau$ are converted to range and marginal standard deviation $\sigma$ as 
$$range=\frac{8}{\kappa}$$
and 
$$\sigma=\left(4\pi e^{2log(\tau )+2log(\kappa)}\right)^{-\frac{1}{2}}$$


`</quote>`

In code, this means:

```{r matern-definition}

matern_function <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}


```


(See also some experience [of another user](https://stats.stackexchange.com/questions/611908/how-to-properly-implement-a-mat%C3%A9rn-kernel-function-in-r).)


Should be fun to implement!
(First, for HG3, then for LG3.)

# Variogram Models

## Constraints

- We use the Matérn function for regression.
- Algorithm of choice is `L-BFGS-B`.
- The range parameter will be maximum-limited at the data extent (`{r} extent` m). It is *the* parameter of interest for spatial coupling.

## Helper Functions

```{r regression-matern-helpers}

# ## for exploration:
# start_values <- c("tau" = mean(y), "kappa" = sqrt(8)/250)
# start_values <- c("phi" = 0.3, "sill" = 10) # , "nu" = 1. 
# params <- start_values
# d <- x
# plot(x, matern_function(x, start_values), type = "l") 

fit_matern <- function(x, y, ...) {
  optimizer_results <- optim(
    fn = function(params) {
      wrap_target_function(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results) {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  sigma <- optimizer_results$par[2]
  range <- 2*sigma
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab("semivariance (m²)") +
    theme_minimal()

  return(g)
}

```

## HG3 

```{r regression-matérn-hg3}
x <- diffdata_binned$ds
y <- semi(diffdata_binned$vhgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

# standardization helps function comprehension
# x <- x/max(x)
# y <- y / mean(y)


matern_hg3 <- fit_matern(x, y,
  par = c(25, 250, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.001),
  upper = c(50, 500, 100, 64),
  # control = list("fnscale" = 1e-4),
  method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
)

print_regression_results(matern_hg3, label = "HG3")

```


Inspecting the residual pattern:

```{r regression-residuals-histogram-hg3}
plot_residuals_histogram(x, y, matern_hg3$predict,
  bins = 128, fill = "lightgray", color = "black")
```


```{r range-sill-conversion-hg3}
plot_matern(matern_hg3)
```

This computes a $2\sigma$ range of `{r} round(2*matern_hg3$par[2], 1)`.


## LG3 

Repeat for LG3:

```{r regression-matérn-lg3}
x <- diffdata_binned$ds
y <- semi(diffdata_binned$vlgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

# standardization helps function comprehension
# x <- x/max(x)
# y <- y / mean(y)


matern_lg3 <- fit_matern(x, y,
  par = c(6, 225, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.001),
  upper = c(50, 500, 100, 64),
  # control = list("fnscale" = 1e-4),
  method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
)

print_regression_results(matern_lg3, label = "LG3")

```


Inspecting the residual pattern:

```{r regression-residuals-histogram-lg3}
plot_residuals_histogram(x, y, matern_lg3$predict,
  bins = 128, fill = "lightgray", color = "black")
```


<!-- The regression results are `{r} paste(round(matern_lg3$par, 2), collapse = ", ")`. -->


```{r range-sill-conversion-lg3}
plot_matern(matern_lg3)
```

This computes a $2\sigma$ range of `{r} round(2*matern_lg3$par[2], 1)`.


## General Observations

::: { .callout-important }

- Both measurement variables indicate a $2\sigma$-range of about $510 m$.
- The plot of the residuals is cut on the right (due to variance being greater than zero; peak of residuals is above zero).
- Despite sufficient sample size, the data-situation seems not ideal (variability across bin variance, distribution of $\gamma$ values, number and influence of outliers).
- Another symptom: regression outcome strongly depends on binning parameters (`extent`, `n_bins`). 
- Regression range was limited to about a kilometer. Larger distances show decreasing variance due to lower sample size.

**It is crucial to repeat the exercise per soilclass.**

:::


# Soilclass

## recap

To recall: we initially joined difference data by soilclass, i.e. two measurement locations are only considered if they reside on the same type of soil.
However, for the regression, soil class mixed.

Furthermore, we saw contradicting bin size patterns in the sample size ([@fig-soilclass-samplesize]).


It might be instructive to repeat the regression separately.

```{r glimpse-soilclass-binned}
knitr::kable(
  diffdata_binned_soilclass %>%
  group_by(soilclass)  %>%
  summarize(count = n())
)
```


- I will focus on `LG3`, given the limited differences above.
- I nonchalantly filter outliers above a certain semivariance, just to see whether technically the regression benefits from it.


## soilclass-wise analysis

::: {.callout-warning}
Note the y-axis! (data filtered: $<=40$; plot scale capped on "heavy" soil)
:::

::: {.panel-tabset}
### light

```{r regression-matérn-light}
#| eval: true
#| code-fold: true

sc_selection <- "light"
sub_data <- diffdata_binned_soilclass %>%
  filter(
    vlgd < 40,
    soilclass == sc_selection
  )
x <- sub_data$ds
y <- semi(sub_data$vlgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]


matern_light_lg3 <- fit_matern(x, y,
  par = c(25, 225, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.01),
  upper = c(50, 500, 10, 64),
  # control = list("fnscale" = 1e-8),
  method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
)


print_regression_results(matern_light_lg3, label = "Matérn (LG3, light soil)")


```


```{r range-sill-conversion-light}
#| eval: true
#| code-fold: true

plot_matern(matern_light_lg3)
```


### heavy

```{r regression-matérn-heavy}
#| eval: true
#| code-fold: true

sc_selection <- "heavy"
sub_data <- diffdata_binned_soilclass %>%
  filter(
    vlgd < 40,
    soilclass == sc_selection
  )
x <- sub_data$ds
y <- semi(sub_data$vlgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

matern_heavy_lg3 <- fit_matern(x, y,
  par = c(5.01, 1., 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(10, 500, 10, 10),
  # control = list("fnscale" = 1e-8),
  method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
)


print_regression_results(matern_heavy_lg3, label = "Matérn (LG3, heavy soil)")


```


```{r range-sill-conversion-heavy}
#| eval: true
#| code-fold: true

g <- plot_matern(matern_heavy_lg3)
g + ylim(0, 1)

```



### peat

```{r regression-matérn-peat}
#| eval: true
#| code-fold: true

sc_selection <- "peat"
sub_data <- diffdata_binned_soilclass %>%
  filter(
    vlgd < 40,
    soilclass == sc_selection
  )
x <- sub_data$ds
y <- semi(sub_data$vlgd)
firsthalf <- x <= extent/2
y <- y[firsthalf]
x <- x[firsthalf]

matern_peat_lg3 <- fit_matern(x, y,
  par = c(0.01, 250, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(10, 500, 0.01, 1.01),
  # control = list("fnscale" = 1e-8),
  method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
)


print_regression_results(matern_peat_lg3, label = "Matérn (LG3, peat)")


```


```{r range-sill-conversion-peat}
#| eval: true
#| code-fold: true

plot_matern(matern_peat_lg3)

```

:::


# Summary

Soilclass Observations:

::: { .callout-important }

One major outcome:

- On heavy soil, variance is generally much lower, as is visible on the scale of the y-axis. ~~Heavy soil implies increased spatial covariance. (I got this wrong!)~~


I have concerns about the suitability of data:

- Regression is tricky: optimizer gets stuck in local minima (e.g. "light soil") or converges to irrelevant results (e.g. "heavy soil").
- ~~The $2\sigma$-range does not reduce, although the data sets should be more homogeneous.~~ It is futile to talk about homogeneity of this data set: it is heterogeneous.
- In fact, semivariance is not regularly distributed: there are many "low semivariance" bins and some high-$\gamma$ outliers (e.g. on peat).
- ~~Why are LG3 and HG3 almost identical?~~


In that regard, open questions remain:

- What does the low variance on heavy soil imply? 
- Should I have detrended per soil type?
- How does the variance relate to actual value difference? (e.g. for "1 cm" threshold)
- Does it make sense to put up models "per cluster", e.g. per RIB?
- Better to return to daily water levels, instead of xG3s?


I will temporally ignore some of these questions, 
(re-)turning to the more adequate `sdmTMB` models.

:::




# References

::: {#refs}
:::
