---
title: "Water Level Models"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs multivariate modeling with `sdmTMB`.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

# re-run models
rerun_models <- FALSE

```


Some parameters below are subject to selective variation.

```{r data-settings}
#| eval: true

# extent <- 1024 # m
# n_bins <- 128
extent <- 1024*2 # m
n_bins <- 256*2


mesh_cutoff <- as.integer(extent / n_bins)
```



# Data Preparation
## load data


The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored

data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- `filterdepth` was stored as `z`
- further parameters: 
  - `elevation`, the elevation from DHMV and `slope`, the local slope
  - `soilclass` ("unknown" category will be disregarded), 
  - `wata_min_dist`,
  - TODO: add `ecoregio` (watina-gebied | ecoregio | n2khab-types)
:::

```{r parameter-rename}
data <- data %>%
  rename(
    filterdepth = z,
    elevation = elevation_dhmv,
    slope = slope_r2.5m,
    water_distance = wata_min_dist
  ) #%>%
  # filter(soilclass != "unknown")

```

## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


# Meshing Flanders

The spatial models require a mesh of the country.

Even more, the outcome crucially depends on mesh characteristics.


```{r sdmTMB-mesh-full}
#| eval: true
flanders_mesh <- make_mesh(data, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(flanders_mesh$mesh$n)
plot(flanders_mesh)
```


# Modeling

::: {.panel-tabset}
## LG3 {#sec-lg3}

```{r sdmTMB-model-lg3}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + elevation + soilclass + water_distance, #+ slope
  data = data,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```

```{r extract-lg3-parameters}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```


## HG3 {#sec-hg3}


```{r sdmTMB-model-hg3}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + elevation + soilclass + water_distance, # + slope
  data = data,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```

```{r extract-hg3-parameters}
#| eval: true
knitr::kable(tidy(fit_hg3))
```

```{r extract-hg3-spatial}
#| eval: true
knitr::kable(tidy(fit_hg3, "ran_pars"))
```

:::

## residuals

There are residuals.
It is not entirely clear what they mean.

```{r residuals-hg3}

res <- residuals(fit_hg3, type = 'mle-mvn')

length(res) == nrow(data)

res <- res[!is.na(res)]
res <- res[res != Inf]
res <- res[res != -Inf]
sqrt(sum(res^2))

ggplot(NULL, aes(x = res)) +
  geom_histogram(bins = 256, color = "black", fill = "lightgray") +
  theme_bw()

```

Great, those are as one would expect.
Note that there are `NA` and `Inf` in there (minority) to be removed.



# generalize: many models

- http://www.cookbook-r.com/Formulas/Creating_a_formula_from_a_string/
- https://stackoverflow.com/questions/46310123/r-dplyr-operate-on-a-column-known-only-by-its-string-name
- https://cran.r-project.org/web/packages/sdmTMB/sdmTMB.pdf ?sanity

```{r sdmTMB-model-general}
#| eval: true
compute_model <- function(model_formula) {
  message("______________________________________________________________________________\n")  
  message(model_formula)
    
  # model_formula <- "hwd ~ filterdepth + elevation + soilclass + water_distance"
  fit <- sdmTMB(
    as.formula(model_formula),
    data = data,
    mesh = flanders_mesh,
    family = gaussian(link = "identity"),
    spatial = "on"
  ) %>% run_extra_optimization()
  san <- sanity(fit, silent = TRUE)
  res <- tidy(fit)
  spa <- tidy(fit, "ran_pars")
    
  outcome <- bind_rows(res, spa)

  for (i in 1:length(san)) {
    check <- names(san)[[i]]
    outcome[nrow(outcome)+1, "term"] <- check
    outcome[nrow(outcome), "estimate"] <- as.integer(san[[i]])
  }

  res <- residuals(fit, type = 'mle-mvn')
  res <- res[!is.na(res)]
  res <- res[res != Inf]
  res <- res[res != -Inf]
  outcome[nrow(outcome)+1, "term"] <- "residuals"
  outcome[nrow(outcome), "estimate"] <- sqrt(sum(res^2))

  outcome <- tidyr::pivot_longer(
    outcome, 
    cols = c("estimate", "std.error", "conf.low", "conf.high")
    ) %>% filter(!is.na(value))

  outcome$model_formula <- model_formula

  return(outcome)

  message("______________________________________________________________________________\n")  
}


# result <- compute_model("hwd ~ filterdepth + elevation + soilclass + water_distance")
```



```{r model-list}
#| eval: false

if (rerun_models){
  model_formulas <- c(
    "hwd ~ filterdepth + elevation + soilclass + water_distance",
    "hwd ~               elevation + soilclass + water_distance",
    "hwd ~ filterdepth             + soilclass + water_distance",
    "hwd ~ filterdepth + elevation             + water_distance",
    "hwd ~ filterdepth + elevation + soilclass                 ",
    "lwd ~ filterdepth + elevation + soilclass + water_distance",
    "lwd ~               elevation + soilclass + water_distance",
    "lwd ~ filterdepth             + soilclass + water_distance",
    "lwd ~ filterdepth + elevation             + water_distance",
    "lwd ~ filterdepth + elevation + soilclass                 ",
    "hw  ~ filterdepth + elevation + soilclass + water_distance",
    "hw  ~               elevation + soilclass + water_distance",
    "hw  ~ filterdepth             + soilclass + water_distance",
    "hw  ~ filterdepth + elevation             + water_distance",
    "hw  ~ filterdepth + elevation + soilclass                 ",
    "lw  ~ filterdepth + elevation + soilclass + water_distance",
    "lw  ~               elevation + soilclass + water_distance",
    "lw  ~ filterdepth             + soilclass + water_distance",
    "lw  ~ filterdepth + elevation             + water_distance",
    "lw  ~ filterdepth + elevation + soilclass                 ",
    "hwd ~ 1",
    "hw  ~ 1",
    "lwd ~ 1",
    "lw  ~ 1"
    )

  results <- lapply(model_formulas, FUN = compute_model)
  results <- bind_rows(results)
  
  write_parquet(results, sink = "data/sdmTMB_results.parquet")

}
```


```{r modelcomparison-residuals}

results <- read_parquet("data/sdmTMB_results.parquet") %>%
  mutate_at(vars(term, name, model_formula), as.factor)

knitr::kable(results %>%
  filter(term == "residuals", name == "estimate") %>%
  select(model_formula, value) %>%
  rename(rms_residual = value ) %>%
  arrange(substring(model_formula, 1, 4), rms_residual))

```

```{r modelcomparison-residuals}

extract <- function (df, selection) {
  df <- df %>%
    filter(name == selection) %>%
    select(model_formula, value)
  colnames(df) <- c("model_formula", selection) 
  return(df)
}

join_extract <- function(df, raw, selection) {
  extra <- df %>% 
    left_join(
      extract(raw, selection),
      by = join_by(model_formula),
      relationship = "one-to-one"
    )
  return(extra)
}

extract_term <- function(select_term){
  term_data <- results %>%
    filter(term == select_term) %>%
    select(model_formula, name, value)  

  term_data <- term_data %>% 
    extract("estimate" ) %>%
    join_extract(term_data, "conf.low") %>%
    join_extract(term_data, "conf.high")

  term_overview <- term_data %>% mutate(
    sig. = (sign(conf.low) * sign(conf.high)) > 0
  )

  return(term_overview)
}

print(levels(results$term))

extract_term("(Intercept)")
extract_term("filterdepth")
extract_term("elevation")
extract_term("soilclasslight")
extract_term("soilclasspeat")
extract_term("water_distance")

```


# TODO split by soilclass
