
---
title: "Water Level Models"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs multivariate modeling with `sdmTMB`.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

```


Some parameters below are subject to selective variation.

```{r data-settings}
#| eval: true

# extent <- 1024 # m
# n_bins <- 128
extent <- 1024*2 # m
n_bins <- 256*2


mesh_cutoff <- as.integer(extent / n_bins)
```



# Data Preparation
## load data


The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored

data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- `filterdepth` was stored as `z`
- further parameters: 
  - `elevation`, the elevation from DHMV
  - `slope`, the local slope
  - `soilclass`, 
  - `wata_min_dist`
:::

```{r parameter-rename}
data <- data %>%
  rename(
    filterdepth = z,
    elevation = elevation_dhmv,
    slope = slope_r2.5m,
    water_distance = wata_min_dist
)  

```

## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


# Meshing Flanders

The spatial models require a mesh of the country.

Even more, the outcome crucially depends on mesh characteristics.


```{r sdmTMB-mesh-full}
#| eval: true
vlaanderen_mesh <- make_mesh(data, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(vlaanderen_mesh$mesh$n)
plot(vlaanderen_mesh)
```


# Modeling

::: {.panel-tabset}
## LG3 {#sec-lg3}

```{r sdmTMB-model-lg3}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + elevation + soilclass + water_distance,
  data = data,
  mesh = vlaanderen_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```

```{r extract-lg3-parameters}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```


## HG3 {#sec-hg3}

```{r sdmTMB-model-hg3}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + elevation + soilclass + water_distance,
  data = data,
  mesh = vlaanderen_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```

```{r extract-hg3-parameters}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-hg3-spatial}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```

:::
