---
title: "Variograms Like a Pro"
author: "Falk Mielke"
date: "`r lubridate::today()`"
number-sections: true
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

# Introduction
I have been messing with variograms.
I should learn to do it correctly.
This is my learning trajectory, documented for others.

```{r setup}
#| eval: true

library("gstat")
library("sp")
library("dplyr") # for group-apply function

```


# Data

## settings
Some general settings:
```{r settings}
n_observations <- 2^12
extent <- 50
smooth_sigma <- 10
zrange <- pi

```

## raw random
Due to an [error on my side](https://github.com/inbo/n2khab-mne-design/commit/6f43f5991f9de56e8fd060b86a352fc175423fdf), 
I randomly paired locations instead of using acutal spatial distance as a coupling criterium.

> Shit happens.
But let's make something out of it.


The pseudo-random data showed a degree of spatial covariance.
Let's try with totally random data, with an amplitude of approximately 1.


```{r simulation}
gimme_points <- function(N = 100, scale = 1.0) (runif(N)*2.0-1.0) * scale
data <- data.frame(
  x = gimme_points(n_observations, extent),
  y = gimme_points(n_observations, extent),
  z_ = gimme_points(n_observations, zrange)
)
head(data, 5)
```

*Mental note: points towards the rim will tend to have fewer neighbors.*


## covariate

I also want to add in a covariate, say... `a`.

```{r covariate-definition}

a_slope <- pi/4

data$a <- 0.7 * data$x / extent + 0.3 * data$y / extent
data$z <- data$z_ + a_slope * data$a

```

*Mental note: `a` is correlated to `x` and `y`.*


Visualizing, with color:
```{r datavis}
rbpal <- colorRampPalette(c('black','gold'))

# https://stackoverflow.com/a/9946970
color <- rbpal(n_observations)[as.numeric(cut(data$z, breaks = n_observations))]

plot(data$x, data$y, col = color, pch = 20)
```



## smoothing

Just so that we can optionally incorporate some Tobler spirit: a smoother.

```{r cross-distance}
cross_diff <- function (vec) outer(X = vec, Y = vec, FUN = function(X, Y)  Y-X)
Euclid <- function (data) sqrt(cross_diff(data$x)^2 + cross_diff(data$y)^2 )
zdiff <- function (data) cross_diff(data$s)
zdiff_abs <- function (data) sqrt( cross_diff(data$s)^2 )


smooth <- function (data, sigma = NULL) {
  # u <- seq(0, 5, length.out = 201)
  # plot(u, dnorm(u, 0, 1))
  
  if (sigma <= 0 || is.null(sigma) || is.na(sigma)) {
    return(data$z)
  }

  dist <- Euclid(data)
  # sigma <- extent / 3
  weight <- dnorm(dist, 0, sigma) 
  weight <- weight / colSums(weight)
  # do.call("cbind", rep(list(data$z), length(data$z)))
  
  zmoothed <- weight %*% data$z
  return (zmoothed[,1])
}

data$s <- smooth(data, smooth_sigma)
color <- rbpal(n_observations)[as.numeric(cut(data$s, breaks = n_observations))]
plot(data$x, data$y, col = color, pch = 20)


```

Nice and smooth. 
Feel free to draw a sunrise.


## "De-Trending"

Just a desparate last try.
Pay good attention, please.

This will actually provide **the answer** [@sec-answer] below.

```{r detrending}
data$d <- lm(s ~ x + y, data)$residual
# color <- rbpal(n_observations)[as.numeric(cut(data$d, breaks = n_observations))]
# plot(data$x, data$y, col = color, pch = 20)

```


# Variograms
... should work out-of-the-box, right?
Then there remains the question: which box?


## distance-difference plots

It is good to keep track of the difference-distance plot.

```{r difference-distance-plot}

unwrap <- function (mat) mat[lower.tri(mat)]
detrend <- function(data) {
  fit <- lm(difference ~ distance, data)
  return(fit$residual)
  # return(fit$fitted.values)
}

# compute raw distance and difference
dist_diff <- data.frame(
  distance = unwrap(Euclid(data)),
  difference = unwrap(zdiff(data)),
  abs_difference = unwrap(zdiff_abs(data))
)

# limit range
dist_diff <- dist_diff %>% filter(distance < extent)

# "detrend"
dist_diff$detrend <- detrend(dist_diff)


# binning
dist_diff$bin <- as.factor(cut(dist_diff$distance, breaks = 65))
dist_diff_binned <- dist_diff %>% 
  group_by(bin) %>%
  summarize_all(
    c("mean" = mean,
      "var" = var,
      "sem" = function (vec) var(vec)/sqrt(length(vec)),
      "cressie" = function (h) (sum((abs(h))^(1/2))/length(h))^4 / (0.457 + 0.494/length(h)),
      "median" = function (h) median(abs(h)),
      "medsq" = function (h) (median((abs(h))^(1/2)))^2
    )
  ) # note: sem does not help
# cf. Cressie, 1993, p. 75 


# choose one parameter
dist_diff_binned$semivariance <- 0.5 * dist_diff_binned$difference_var
# dist_diff_binned$semivariance <- 0.5 * dist_diff_binned$detrend_mean # totally wrong.


# plot(distance, difference, pch = 21, color = 'black', alpha = 0.1)
head(dist_diff_binned)


# plot
par(mfrow = c(2,1))
plot(dist_diff_binned$distance_mean, dist_diff_binned$abs_difference_mean, type = "o", col = "darkblue")
plot(dist_diff_binned$distance_mean, dist_diff_binned$semivariance, type = "o", col = "darkgreen")

par(mfrow = c(1,1))
```

The lower plot is the variance, plotted against the distance.

Good to have.



## fields::vgram()

There is a variogram in the `fields` package.
It seems easy-to-use and intuitive, no large dependencies, and the documentation was fine.

```{r empirical-variogram}
#| eval: true

loc <- as.matrix(data)[,1:2]
y <- data$s
fields_vgram <- fields::vgram(loc = loc, y = y, N = 16, dmax = extent)

plot(fields_vgram, breaks = seq(0, extent, length.out = 65), type = "o", col = "darkgreen")

```


This is reassuring.
The outcome matches the semivariance-distance plot above. 

However, the smoothed data does not show a classical nugget-range-sill like plot.


Technically, I could work with this.
Might be a good reference.
Only downside: the performance is not great.



Next try.


## gstat::variogram() 
... has a [documentation](https://r-spatial.github.io/gstat/reference/variogram.html) which leaves me ridiculously clueless:

|   Calculate Sample or Residual Variogram or Variogram Cloud
|   Description:
|      Calculates the sample variogram from data, or in case of a linear
|      model is given, for the residuals, with options for directional,
|      robust, and pooled variogram, and for irregular distance
|      intervals.
|      In case spatio-temporal data is provided, the function
|      ‘variogramST’ is called with a different set of parameters.

Always careful with functions which doe multiple things.

Yet it's the `examples` that matter:
| Examples:
|    library(sp)
|    data(meuse)
|    # no trend:
|    coordinates(meuse) = ~x+y
|    variogram(log(zinc)~1, meuse)
|    # residual variogram w.r.t. a linear trend:
|    variogram(log(zinc)~x+y, meuse)
|    # directional variogram:
|    variogram(log(zinc)~x+y, meuse, alpha=c(0,45,90,135))
|    variogram(log(zinc)~1, meuse, width=90, cutoff=1300)
|
|    # GLS residual variogram:
|    v = variogram(log(zinc)~x+y, meuse)
|    v.fit = fit.variogram(v, vgm(1, "Sph", 700, 1))
|    v.fit
|    set = list(gls=1)
|    v
|    g = gstat(NULL, "log-zinc", log(zinc)~x+y, meuse, model=v.fit, set = set)
|    variogram(g)


Good, or not: I need to transform my data into an `sp` type with coordinates.
I then have to use a model forula, even if there is no model (yet).
I have a hundred options to include models, or create a "variogram cloud".

And, to my amazement, they reference the same book as the above package.

> Cressie, N.A.C., 1993, Statistics for Spatial Data, Wiley.

(Maybe I should read it?)



I will stick with the examples to finally get my variogram of the data.

```{r gstat-vgram-basic}

data_sp <- data
sp::coordinates(data_sp) = ~x+y
gstat_vgram <- gstat::variogram(s ~ 1, data_sp, cutoff = extent, width = extent / 64)

# compute a Matèrn model
#
m <- suppressWarnings({fit.variogram(gstat_vgram, vgm("Mat"), fit.kappa = TRUE)})

# plot(gstat_vgram, model = m, pch = 22, col = "darkblue", type = "o")
# grid(nx = NA, ny = NULL) # <- I would expect this to work, yet it doesn't

plot(gstat_vgram, model = m, pch = 22, col = "darkblue", type = "o")
```

Same as above. 
Nice and performant.
Still no nugget-sill story.


## linear trend, w.r.t.?!

Let us try another one of the examples!
I choose the
> residual variogram w.r.t. a linear trend:

Above, I clearly set `sp::coordinates(data_sp) = ~x+y`. 
It took me a while to figure that setting coordinates does not do the trick.
You seemingly have to add them to the variogram formula.


```{r gstat-vgram-formula}

# all as above, except the model formula
gstat_vgram_detrend <- gstat::variogram(s ~ x + y, data_sp, cutoff = extent, width = extent / 64)

m <- fit.variogram(gstat_vgram_detrend, vgm("Mat"), fit.kappa = TRUE)
plot(gstat_vgram_detrend, model = m, pch = 22, col = "darkblue", type = "o")

```

There we go!
This is the image I would have expected!


However, expecting is not enough.
Which one is right?
Certainly the "de-trended" one with `s ~ x + y`.
Yet, I do not understand why! 
The `~x+y` of the formula are not part of the `model` argument;
yet the y-values change drastically with this argument (residuals?).

The only hint that this is about "de-trending" is from the comment in the documentation.


There is more trouble.

Initially, [I saw no way](https://r-spatial.github.io/gstat/reference/plot.gstatVariogram.html) to set the `breaks`.

Yet they are needed for [gstat::variogram](https://r-spatial.github.io/gstat/reference/variogram.html)
> width: the width of subsequent distance intervals into which data point pairs are grouped for semivariance estimates


I also struggled getting a straight line into the plot, to see whether the bins are plotted on a slope.



## unsmoothed

Now, smoothing has a critical effect; I would like to see my variogram for the raw and edgy data.
I not yet remove the covariate.

```{r gstat-vgram-unsmoothed}

# all as above, except the outcome variable
gstat_vgram_unsmoothed <- gstat::variogram(z ~ x + y, data_sp, cutoff = extent, width = extent / 64)

m <- suppressWarnings({fit.variogram(gstat_vgram_unsmoothed, vgm("Mat"), fit.kappa = TRUE)})
plot(gstat_vgram_unsmoothed, model = m, pch = 22, col = "darkblue", type = "o")

```

At least, there are `warnings` (as supressed above), I guess it is for the model fit.
Variogram fitting is not trivial.


## plotting a horizontal line
Yet note that the empirical variogram line is rising!
Just that you cannot see it with the bare eye.
To show, I need to mess with that horrible `gstatVariogram` object.
It truly is a mess!
Just try to manually plot the variogram data... [see here for help]().

- The [documentation](https://r-spatial.github.io/gstat/reference/plot.gstatVariogram.html) is not helpful.
- I tried `?graphics::grid`; the vgram plot seems not to be a regular plot.
- I tried accessing the vgram data, but `attributes(vgram)` are not exactly obvious.
- The [source code](https://github.com/r-spatial/gstat/blob/main/R/plot.gstatVariogram.R) lacks relevant meaningful comments.


But, hey, I can live with wasting an hour just to not being able to plot straight lines.

Here is a workaround, using `vgram$gamma`.
Remember: `gamma` is the symbol of the variance.
I compare the unsmoothed, but covariate-trended data variogram to the very raw data.


```{r gstat-vgram-extractpoints}
v <- gstat_vgram_unsmoothed
# attributes(v)
breaks_ref <- seq(1, 64, length.out = 64)
dots_ref <- v$gamma[breaks_ref]

# remove the covariate
gstat_vgram_flat <- gstat::variogram(z_ ~ x + y, data_sp, cutoff = extent, width = extent / 64)
v <- gstat_vgram_flat
breaks_flat <- seq(1, 64, length.out = 64)
dots_flat <- v$gamma[breaks_flat]
plot(breaks_flat, dots_flat, ylim = c(3.0, 3.5), col = 'black', pch = 21, type = 'l')
points(breaks_ref, dots_ref, col = 'blue', pch = 19)
# plot(breaks_ref, dots_ref, col = 'blue', pch = 19)
grid(ny = NULL)
```

Good. 
That was non-instructive.
Unsmoothed data (but with the covariate) and "flat" data (without the covariate) yield exactly the same variogram.
I guess I should not have coupled `a` to `x, y`.


Maybe being able to extract variogram data helps someone.


On we go.


## reminder - what do we have

So far, so bad.
To wrap up: Intermediate Summary™!


- There are two kinds of people: those who understand what a variogram is (`fields`), and those who understand it differently (`gstat`).
- Above, I worked on random data which was smoothed with a kernel of about `{r} smooth_sigma`.
- Adding a covariate does not change the picture too drastically. 
- `gstat` seems to be most sophisticated, but also least understandable.


The first point is critical:
the "semivariance-distance" plot is not the end of the story. 
`gstat::variogram`, when used with the `~x+y` formula, performs some data transformation (linear detrending??) which brings the simulation data into a nice and smooth Matèrn shape.


# The Source (Code)
By now, I do not know why the `gstat` people understand variograms differently and how they bring their data to a space where fitting a Matèrn function goes so well.
My last hope is the source code.
And, oh boy, is it badly documented!

- <https://github.com/r-spatial/gstat/blob/main/R/variogram.default.R>


... 
...

...


Source code does not help.


Further leads:
- <https://stackoverflow.com/questions/29425101/why-does-my-detrending-not-result-in-a-different-semivariogram-in-r>

- <https://surferhelp.goldensoftware.com/vario/Variogram_Frequently_Asked_Questions.htm>


There is the formula:
\[2\gamma = \varepsilon \left[\left(z_j - z_i\right)^2 \right]\]

and the only things unclear to me are the epsilon \(\varepsilon\) and the brackets \(\left[\cdot\right]\).
In other places (Cressie, 1993, p.40, eq. 2.2.7) this is just taken as the mean difference (i.e. sum over all \(i, j\), and normalized by the count of lag vectors).


No success.




Hmmm... 

# The Answer {#sec-answer}
Watch This:

```{r vgram-o-matic}
vgram_o_matic <- function(formula) {
    vgram <- gstat::variogram(formula, data_sp, cutoff = extent, width = extent / 64)
}
```


My own variogram and that computed with `fields::variogram` has the following form, same as `gstat::variogram` with `s ~ 1`:

```{r vgram-s1}
plot(vgram_o_matic(s~1), pch = 22, col = "darkblue", type = "o")
```

However, the gstat "variogram" takes a different form

```{r vgram-sxy}
plot(vgram_o_matic(s ~ x + y), pch = 22, col = "darkblue", type = "o")

```



The obvious difference is the "formula": what *exactly* does the ` ~ x + y` part do?

```{r vgram-d}
plot(vgram_o_matic(d ~ 1), pch = 22, col = "darkblue", type = "o")
```


:mic-drop:



# Modeling Covariates

Time to recover the covariate introduced above.
No more postponing to read the `gstat` manual.

- https://www.gstat.org/gstat.pdf

... or better not. 
It is highly confusing (probably my bad).
And after all, there is a formula in the variogram function.


```{r gstat-vgram-covariate-1}

# work in progress
gstat_vgram_cov <- gstat::variogram(s ~ x+y+a, data_sp, cutoff = extent, width = extent / 64)

m <- fit.variogram(gstat_vgram_cov, vgm("Mat"), fit.kappa = TRUE)
plot(gstat_vgram_cov, model = m, pch = 22, col = "darkblue", type = "o")

```


To be continued!


# Notes

> Dat hoogteligging van het meetpunt gecorreleerd is met waterstand TAW, is te verwachten. Ook voor verschillen op grote afstand is dit te verwachten: grote verschillen in hoogteligging TAW gaan gepaard met grote verschillen in waterstand TAW.
< Karen Wuyts, 10:12 AM, Edited
Bedoel je dan dat er in het variogram best geplot wordt voor een aantal hoogteverschilklassen? Het lijkt me dan wel nuttig om te weten of de criteria voor hergebruik van bestaande piezo's/peilbuizen in de praktijk enkel via een desktopanalyse worden getest of op terrein. Om het op terrein toe te passen, moeten we niet afkomen met hoogteverschillen van een halve meter.

Kan je afstand en verschil in hoogte samen in het model steken? Ze zijn waarschijnlijk met elkaar gecorreleerd?

< Floris Vanderhaeghe, 10:15 AM
"het variogram best geplot wordt voor een aantal hoogteverschilklassen": dat was het idee dat ik had, ja.

< Floris Vanderhaeghe, 10:21 AM
Of je modelleert het variogram met hoogteverschil als covariaat, kan wellicht ook inderdaad (afhankelijk van wat de bestaande variogrammodellen toelaten; klassiek is alleen ifv afstandsverschil). Ik kan wel eens kijken wat er gaat. Dan kun je range (of een bepaald verschilcriterium) uitdrukken ifv afstandsverschil + hoogteverschil bv. Waarbij het variogram zelf nog steeds gedefinieerd is obv waterstand in m TAW, het gaat dus niet om het variogram van een residual.

Afstand tot waterloop was ook nog een criterium.


> Kan je afstand en verschil in hoogte samen in het model steken? Ze zijn waarschijnlijk met elkaar gecorreleerd?
Dat is inderdaad opletten geblazen.
Hoogteverschilklassen met aparte variogrammen gaan daar meer controle over geven lijkt me.

< Karen Wuyts, 10:28 AM
Hoe beslissen we welke grenzen we voor de klassen gebruiken? Bovenstaande figuur van log |delta W| ifv log |delta h| bevat natuurlijk ook het afstandseffect.

< Floris Vanderhaeghe, 10:49 AM
Ik zou die a priori kiezen obv intuïtie en kijken wat het geeft, nadien bijstellen indien nodig. Bv klassegrenzen 0, 0.1, 0.2, 0.3, 0.5, 0.75, 1.5.

< Floris Vanderhaeghe, 10:53 AM
Met gstat (dat uiteindelijk op kriging gericht is), kun je een variogram model fitten als volgt:

- eerst empirisch variogram maken met `variogram()`
- dan variogrammodel met initiële waarden definiëren met `vgm()` (zie `show.vgms()` voor beschikbare vormen)
- de modelparameters van dit object verfijnen obv de data met `fit.variogram()`

< Floris Vanderhaeghe, 10:56 AM
In kriging kun je extra covariaten meenemen, en in een variogram kun je anisotropieparameters meenemen en ook een model specifiëren om er de residuals van te kunnen gebruiken voor het variogram, maar een variogram dat 'variabel' is ifv covariaten heb ik niet onmiddellijk gezien (maar ik zou me veel meer moeten verdiepen in de materie daarrond, bv Cressie 1993 en de gstat manual).

