---
title: "Explore sdmTMB"
author: "Falk Mielke"
date: "`r lubridate::today()`"
number-sections: true
bibliography: references.bib
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

# Purpose

Testing the `sdmTMB` package on synthetic data.


```{r setup}
#| eval: true
silence <- suppressPackageStartupMessages
# library("sp") |> silence() # SPatial stuff
library("dplyr") |> silence() # for group-apply function
library("ggplot2") |> silence() # plotting
library("sdmTMB") |> silence() # modeling

```



# Data

## general settings

```{r settings}
n_observations <- 2^12
extent <- 50
smooth_sigma <- 8
zrange <- pi

# covariate effect magnitudes
a_slope <- pi/6
b_slope <- pi/8

```

## raw random

```{r simulation}
gimme_points <- function(N = 100, scale = 1.0) (runif(N)*2.0-1.0) * scale
data <- data.frame(
  x = gimme_points(n_observations, extent),
  y = gimme_points(n_observations, extent),
  z_ = gimme_points(n_observations, zrange)
)
```


## covariates

```{r covariate-definition}

data$a <- 0.7 * data$x / extent + 0.3 * data$y / extent
data$b <- as.integer(rbinom(n = n_observations, size = 1, prob = 0.4))

data$z <- data$z_ + a_slope * data$a + b_slope * data$b

```


## smoothing

```{r cross-distance}
cross_diff <- function (vec) outer(X = vec, Y = vec, FUN = function(X, Y)  Y-X)
Euclid <- function (data) sqrt(cross_diff(data$x)^2 + cross_diff(data$y)^2 )
zdiff <- function (data) cross_diff(data$s)
zdiff_abs <- function (data) sqrt( cross_diff(data$s)^2 )


smooth <- function (data, sigma = NULL) {

  if (sigma <= 0 || is.null(sigma) || is.na(sigma)) {
    return(data$z)
  }

  dist <- Euclid(data)
  weight <- dnorm(dist, 0, sigma)
  weight <- weight / colSums(weight)

  zmoothed <- weight %*% data$z
  return (zmoothed[,1])
}

data$s <- smooth(data, smooth_sigma)

```

## "De-Trending"

```{r detrending}
data$d <- lm(s ~ x + y, data)$residual

```


## glimpse

```{r glimpse-data}
head(data, 5)
```

# *Gadus macrocephalus*
([reference](https://pbs-assess.github.io/sdmTMB/index.html#basic-use))

> Our model contains a main effect of depth as a penalized smoother, a spatial random field, and Tweedie observation error. 

```{r pacific-cod}
head(pcod)
```
  

> We start by creating a mesh object that contains matrices to apply the SPDE approach.
> Here, cutoff defines the minimum allowed distance between points in the units of X and Y (km). Alternatively, we could have created any mesh via the fmesher or INLA packages and supplied it to make_mesh(). We can inspect our mesh object with the associated plotting method plot(mesh)
```{r cod-mesh}
mesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
plot(mesh)
```

> Fit a spatial model with a smoother for depth:

```{r cod-model}
fit <- sdmTMB(
  density ~ s(depth),
  data = pcod,
  mesh = mesh,
  family = tweedie(link = "log"),
  spatial = "on"
)
sanity(fit)
fit
```

- There are visualization and data extraction options.
- Different families have different kids.


([further reading](https://pbs-assess.github.io/sdmTMB/articles/basic-intro.html))


# Water Level Analysis (WLA™)

```{r basic-model}

```
