---
title: "Water Level Data Download"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---


# Preparation 

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("mapview")   |> void() # map visualization
library("doParallel")|> void() # parallel query of data


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# option to work offline; requires previous data download
work_offline <- FALSE

# verbosity
global_verbose <- FALSE
global_progress <- FALSE


# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./cache", showWarnings = FALSE)

# if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 2 # minimal cluster member count; fewer memeber clusters are removed
# metadata_export_file <- paste0("cache/waterlevel_metadata_cl", cluster_dist, ".parquet")
metadata_export_file <- here::here(, "_metadata.parquet")

# the user may choose to reload raw data
force_data_reassembly <- FALSE
reload <- force_data_reassembly || !file.exists(metadata_export_file)
reload_emd <- FALSE

```


# Clustered Peilpunt Data

## Strategy

- getting the peilpunt locations
- clustering
- pairing
- soilclass, distance, etc.


## Loading

```{r raw-data-query-dbplyr}

max_filterdepth <- 5.8 # 3 # (m) # choice to focus on shallow gw (ecologically relevant for GW_03.3 and GW_05.1)

peilpunten <- tbl(watina_dwh, "dimPeilpunt") %>%
  # filter(PeilpuntStatusCode == "VLD") %>% 
  filter(LandCode == "BE") %>%
    mutate(
      filterdepth = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
    ) %>% 
  filter(filterdepth <= max_filterdepth, filterdepth > 0) %>% 
  left_join(
    tbl(watina_dwh, "dimMeetpunt"),
    by = join_by(MeetpuntWID == MeetpuntWID),
    relationship = "many-to-many"
  ) %>%
  filter(!is.na(PeilpuntWID)) %>% 
  select(
    PeilpuntWID,
    PeilbuisLengte,
    ReferentieNiveauMaaiveld,
    ReferentieNiveauTAW,
    FilterLengte,
    filterdepth,
    PeilpuntStatusCode,
    PeilpuntCode,
    MeetpuntWID,
    PeilpuntXCoordinaat,
    PeilpuntYCoordinaat,
    PeilpuntStatusCode,
    MeetpuntStatusCode
  ) %>%
  rename(
    mp_idx = MeetpuntWID,
    pp_idx = PeilpuntWID,
    pblen = PeilbuisLengte,
    ref_mv = ReferentieNiveauMaaiveld,
    ref_taw = ReferentieNiveauTAW,
    flength = FilterLengte,
    fdepth = filterdepth,
    pp_code = PeilpuntCode,
    pp_vld = PeilpuntStatusCode,
    mp_vld = MeetpuntStatusCode,
    x = PeilpuntXCoordinaat,
    y = PeilpuntYCoordinaat,
  ) %>%
  collect %>% 
  mutate(
    region = substring(pp_code, first = 1, last = 3)
  ) %>%
  filter(region != "VBS") # remove Luxemburg

# print(colnames(peilpunten))
nrow(peilpunten)

peilpunten %>%
  mutate(vld = paste0(mp_vld, "_", pp_vld)) %>% 
  ggplot(aes(x = fdepth, fill = vld)) +
  geom_histogram(bins = 128, color = "black") +
  theme_bw()
```


## Clustering

Clustering the data with a cluster range of `{r} cluster_dist` m,
removing all clusters which contain only one location.

```{r cluster-join}

meta_data <- peilpunten %>%
  join_clusters(index_column = "pp_idx", characteristic_distance = cluster_dist) %>%
  remove_underpopulated_clusters(
    minimum_cluster_member_count = min_cluster_size
  ) 

knitr::kable(head(meta_data))
```


Cluster size:

```{r cluster-size}

meta_data %>%
  count(cluster) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = n)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()

# levels(meta_data$cluster)
```


```{r cluster-location}

st_clust <- meta_data %>%
  group_by(region, cluster) %>%
  summarize(x = mean(x), y = mean(y), count = n()) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_clust,
  map.types = c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"),
  zcol = "region",
  cex = "count",
  legend = FALSE
)

```


## local Belgian plate tectonics

Mental note: some locations have ambiguous co-ordinates

```{r location-duplicates}
duplicates <- meta_data[, c("cluster", "mp_idx", "pp_idx", "x", "y")]
duplicates <- duplicates %>% 
  filter(mp_idx %in% (duplicates %>% count(mp_idx) %>% filter(n>1) %>% pull(mp_idx))) %>%
  arrange(cluster, mp_idx)
knitr::kable(head(duplicates))
```

Some, but not all, might be due to historical coordinate rounding.


## soilclass

Soilclass is so fundamental that I consider it pert of the meta data.

```{r soilclass-join}
meta_data <- meta_data %>%
    join_soilclass(index_column = "pp_idx")


write_parquet(meta_data, sink = metadata_export_file)
```


```{r soilclass-map}
st_soil <- meta_data %>%
  group_by(soilclass, region, cluster) %>%
  summarize(x = mean(x), y = mean(y), count = n()) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_soil,
  map.types = c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"),
  zcol = "soilclass",
  cex = "count"
  # legend = FALSE
)

```

## elevation

... once we're at it.

```{r split-balanced}


backbone_data <- meta_data %>%
  select(pp_idx, cluster, x, y)
knitr::kable(head(backbone_data))


# backbone_data$cluster_group <- split_balanced(backbone_data, count_column = "cluster", n_bins = 17)

data_subgroups <- split(
  backbone_data,
  split_balanced(
    backbone_data,
    count_column = "cluster",
    n_bins = 8*17,
    show = TRUE
  )
)

# ggplot(NULL, aes(x = 1:length(data_subgroups), y = sapply(data_subgroups, FUN = nrow))) + geom_step()

```

```{r download-elevation}

label <- "elevation"
if (reload) {
  empty_cache_subfolder(paste0(label, collapse = ""))
}


parallel_query_to_cache(
  split_data = data_subgroups,
  label = label,
  query_function = query_elevation,
  index_column = "pp_idx",
  coordinate_columns = c("x", "y"),
  progress = FALSE
  )

# fi <- "cluster_grp00001"
# check <- read_parquet(here::here(local_cache_folder", "elevation", fi))
# fi <- "elevation.parquet"
# check <- read_parquet(here::here(local_cache_folder, fi))

```

```{r elevation-archive}

if (FALSE) {
  # meta_data <- meta_data %>%
  #   join_elevation(index_column = "pp_idx")
  
  ## "Oh No!", there were duplicates.
  # data <- meta_data
  # index_column = "mp_idx"
  # coordinate_columns <- c("x", "y")
  # elevation_lookup <- query_elevation(data, index_column = "mp_idx")
  # elevation_grouped <- elevation_lookup %>% 
  #   filter(mp_idx %in% (elevation_lookup %>% count(mp_idx) %>% filter(n>1) %>% pull(mp_idx)))
  # elevation_lookup %>% summarize(across(everything(), ~ mean(.x, na.rm = TRUE)), .by = !!index_column)
  # data <- data %>%
  #   dplyr::left_join(
  #     elevation_lookup,
  #     by = index_column,
  #     relationship = "many-to-one"
  #   )
}

```


```{r elevation-map}

fi <- "_elevation.parquet"
st_elevation <- read_parquet(here::here(local_cache_folder, fi)) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_elevation,
  map.types = c("OpenTopoMap", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap" ),
  zcol = "elevation_dhmv"
  # legend = FALSE
)

```

> 'I want to see mountains again, Gandalf, mountains, and then find somewhere where I can rest.'


## waterdistance

```{r download-waterdistance}

label <- "waterdistance"
if (reload) {
  empty_cache_subfolder(paste0(label, collapse = ""))
}


parallel_query_to_cache(
  split_data = data_subgroups,
  label = label,
  query_function = query_waterdistance,
  index_column = "pp_idx",
  coordinate_columns = c("x", "y"),
  progress = FALSE
  )


```


Okay, why not, let's map it.

```{r wata-map}

fi <- "_waterdistance.parquet"
st_wata <- read_parquet(here::here(local_cache_folder, fi)) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_wata,
  map.types = c("OpenTopoMap", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap" ),
  zcol = "wata_dist_class"
  # legend = FALSE
)

# TODO this is not correct yet; see coastal wells
```

# Quantiles

To correctly frame yearly water water fluctuations,
we would like to know various quantile levels per location.
Downstream, these will be used to determine "high level" and "low level" episodes,
kind of emulating "HG3" and "LG3", but for continuous water level measurement.


```{r download-quantiles}

label <- "quantiles"
if (reload) {
  empty_cache_subfolder(paste0(label, collapse = ""))
}


# query_waterlevel_quantiles(
#   data = backbone_data,
#   index_column = "pp_idx",
#   n_quantiles = 20,
#   db_conn = watina_dwh,
#   progress = TRUE
# )

# data <- data_subgroups[[1]]

parallel_query_to_cache(
  split_data = data_subgroups,
  label = label,
  query_function = query_waterlevel_quantiles,
  sequential = TRUE,
  verbose = global_verbose,
  index_column = "pp_idx",
  n_quantiles = 20,
  db_conn = watina_dwh,
  progress = global_progress
  )


```



# amplitude

## purpose
Some water levels oscillate a lot,
others are effectively flat.
We seem to miss a relevant co-variate which can explain the pronounced differences in amplitude.


Here, I extract the "momentary amplitude" of the water level measurements.
Then, I will measure their relation to water distance, elevation, *et cetera*. 


## EMD implementation

From <https://github.com/inbo/tutorials/tree/emd>.

```{r r-firstmode-function}
#' extracts from a signal (with an optional timeline vector)
#'
#' extra arguments will propagate to `findpeaks`
#'
#' @returns a list with
#'   - first empirical mode: mean of envelope
#'   - envelope: the upper and lower peak amplitude envelope
#'   - select: boolean vector of time subset
#'   - extrema: peaks and troughs
#'
#' 
#' usage:
#' > emd <- calculate_empirical_mode(y, t = x, threshold = 0.4, minpeakdistance = 20)
calculate_empirical_mode <- function (signal, t = NULL, ...) {

  # general purpose time vector
  if (is.null(t)) {
    t <- seq(0., 1., length.out = length(signal))
  }

  ## 1. find peaks
  peaks <- sort(pracma::findpeaks(signal, ...)[, 2])
  troughs <- sort(pracma::findpeaks(-signal, ...)[, 2])
  extrema <- sort(c(peaks, troughs))

  if ((length(peaks) < 4) || (length(troughs) < 4)) {
    # message("too few peaks or troughs.")
    return(NA)
  }
  
  # if (peaks[1] < troughs[1]) {
  #   peaks <- c(peaks, length(signal))
  #   troughs = c(1, troughs)
  # } else if (peaks[1] > troughs[1]) {
  #   peaks = c(1, peaks)
  #   troughs = c(troughs, length(signal))
  # }

  ## 2. interpolate -> envelope
  # selection <- t >= t[max(c(peaks[1], troughs[1]))] & t <= t[min(c(peaks[length(peaks)], troughs[length(troughs)]))]  
  selection <- function (pt) (t >= t[pt[1]]) & (t <= t[tail(pt, 1)])
  

  if (FALSE) {
    # debugging
    peaks <- sort(pracma::findpeaks(signal, zero = "-")[, 2])
    troughs <- sort(pracma::findpeaks(-signal, zero = "-")[, 2])
    ggplot(NULL, aes(x = t, y = signal)) +
      geom_line() +
      geom_point(aes(t[peaks], signal[peaks])) +
      geom_point(aes(t[troughs], signal[troughs]))
  }

  common_selection <- Reduce("&", list(selection(peaks), selection(troughs) ))

  # ti <- as.numeric(t[common_selection])
  # yi <- as.numeric(signal[common_selection])
  # py <- interpolators::evalInterpolator(interpolators::iprPCHIP(t[peaks], as.numeric(signal[peaks])), ti)
  # ty <- interpolators::evalInterpolator(interpolators::iprPCHIP(t[troughs], as.numeric(signal[troughs])), ti)
  py <- interpolators::evalInterpolator(interpolators::iprPCHIP(
    t[peaks],
    as.numeric(signal[peaks])),
    t[selection(peaks)]
  )
  if (peaks[1] > 1) py <- c(rep(py[1], peaks[1]), py[2:length(py)])
  if (tail(peaks, 1) < length(t)) py <- c(py, rep(tail(py, 1), length(t)-tail(peaks, 1)))
  py <- py[1:length(t)]

  ty <- interpolators::evalInterpolator(interpolators::iprPCHIP(
    t[troughs],
    as.numeric(signal[troughs])),
    t[selection(troughs)]
  )
  if (troughs[1] > 1) ty <- c(rep(ty[1], troughs[1]), ty[2:length(ty)])
  if (tail(troughs, 1) < length(t)) ty <- c(ty, rep(tail(ty, 1), length(t)-tail(troughs, 1)))
  # plot(t, ty)
  ty <- ty[1:length(t)]

  ## 3. empirical mode
  envelope <- cbind("upper" = py, "lower" = ty)
  rownames(envelope) <- t
  firstmode <- rowMeans(envelope)
  residual <- signal - firstmode
  amplitude <- abs(py-ty)/2

  return(list(
    "time" = t,
    "signal" = signal,
    "firstmode" = firstmode,
    # "envelope" = envelope,
    "envelope_upper" = py,
    "envelope_lower" = ty,
    "selection" = common_selection,
    "extrema" = extrema,
    "residual" = residual,
    "amplitude" = amplitude
  ))
}


plot_emd <- function(new_emd, raw_trace = NULL, show = TRUE, ...) {

  ti <- new_emd[["time"]]
  signal <- new_emd[["signal"]]
  firstmode <- new_emd[["firstmode"]]
  envelope_upper <- new_emd[["envelope_upper"]]
  envelope_lower <- new_emd[["envelope_lower"]]
  residual <- new_emd[["residual"]]
  amplitude <- new_emd[["amplitude"]]

  if (!is.null(raw_trace)) {
    x <- raw_trace[["t"]]
    y <- raw_trace[["y"]]
  } else {
    x <- ti
    y <- signal
  }

  h0 <- ggplot(NULL, aes(x = x, y = y)) +
    geom_point() +
    geom_step(alpha = 0.3, color = "darkgray") +
    geom_hline(yintercept = mean(y),
      alpha = 0.8, color = "black", linetype = "dashed") +
    geom_hline(yintercept = mean(y) - mean(amplitude),
      alpha = 0.8, color = "darkred", linetype = "dashed") +
    geom_hline(yintercept = mean(y) + mean(amplitude),
      alpha = 0.8, color = "darkred", linetype = "dashed") +
    theme_bw() +
    theme(legend.position = "none")

  h1 <- ggplot(NULL, aes(x = ti, y = firstmode)) +
    geom_line(color = "black") +
    # geom_point(aes(x = ti, y = signal)) +
    geom_line(aes(x = ti, y = signal), color = "darkgray", alpha = 0.4) +
    geom_line(aes(x = ti, y = envelope_upper), color = "orange", alpha = 0.4) +
    geom_line(aes(x = ti, y = envelope_lower), color = "orange", alpha = 0.4) +
    theme_bw() +
    labs(y = "first mode") +
    theme(legend.position = "none")
  
  h2 <- ggplot(NULL, aes(x = ti, y = residual)) +
    geom_line(color = "black") +
    geom_line(aes(x = ti, y = amplitude), color = "darkred", alpha = 0.8) +
    geom_line(aes(x = ti, y = -amplitude), color = "darkred", alpha = 0.8) +
    theme_bw() +
    labs(x = "time (arb. units)", y = "residual") +
    theme(legend.position = "none")

  if (!show) {
    return(list(h0, h1, h2))
  }

  h <- grid.arrange(h0, h1, h2, ncol = 1, ...) # gridExtra
  return(h)
}

```
  

## explore

```{r explore-emd}

get_trace <- function(pp_idx, show_plot = FALSE) {
  # pp_idx = 24

  peilmeting <- tbl(watina_dwh, "FactPeilMeting") %>%
    filter(
      PeilpuntWID == pp_idx,
      PeilmetingStatusCode == "VLD"
    ) %>%
    mutate(is_vld = PeilmetingStatusCode == "VLD") %>% 
    select(DatumWID, is_vld, mTAW) %>%
    arrange(DatumWID) %>%
    collect
  
  peilmeting <- peilmeting %>%
    mutate(date = as.Date(as.character(DatumWID), format = "%Y%m%d"))

  if (show_plot){
    h <- peilmeting %>%
      ggplot(aes(x = date, y = mTAW)) +
        geom_point() +
        geom_step(alpha = 0.3, color = "darkgray")
    print(h)
  }

  t <- peilmeting %>% pull(date) 
  t0 <- t[1]
  t <- as.numeric(difftime(t, t0, units = "days"))
  y <- peilmeting %>% pull(mTAW) 
  vld <- peilmeting %>% pull(is_vld) 
    
  return(list("t" = t[vld], "y" = y[vld], "t0" = t0, "vld" = vld, "tx" = t, "yx" = y))
}

get_emd_trace <- function(tr) calculate_empirical_mode(tr[["y"]], t = tr[["t"]], zero = "-")
count_trace <- function(tr) length(tr[["t"]])
median_sampling_interval <- function(tr) median(diff(trace[["t"]]))
try_emd_trace <- function(tr) {
  tryCatch({
     emd <- calculate_empirical_mode(tr[["y"]], t = tr[["t"]], zero = "-")
    }, error = function(cond) {
        return(NA)
    }, warning = function(cond) {
        return(NA)
    }
  )
  return(emd)
}
get_firstmode <- function(emd) list("t" = emd[["time"]], "y" = emd[["firstmode"]])
mean_amplitude <- function(emd) mean(emd[["amplitude"]])
get_info <- function(idx, info) paste0(meta_data %>% filter(pp_idx == idx) %>% pull(!!info), collapse = ", ")
get_locationcode <- function(idx) get_info(idx, "pp_code")

# example
idx <- 26 # 24
label <- get_locationcode(idx)
trace <- get_trace(idx, show_plot = FALSE)
emd <- get_emd_trace(trace)
# plot_emd(emd)
emd2 <- get_emd_trace(get_firstmode(emd))
# emd2[["signal"]] <- emd[["signal"]]
h <- plot_emd(emd2, raw_trace = trace, top = label) #,gp=gpar(fontsize=12,font=3)
ggsave("figures/test.pdf", h)
# print(h)
var(2*trace$y)
mean_amplitude(emd) + mean_amplitude(emd2)

# t <- emd2[["time"]]
# amp <- emd2[["amplitude"]] + emd[["amplitude"]][emd2[["select"]]]
# plot(t, amp, type = "l", col = "orange", ylim = c(0, max(amp)))
# lines(trace[["t"]], abs(trace[["y"]] - mean(trace[["y"]])))

median_sampling_interval(trace)


```

```{r test-splitting}
split_trace <- function(tr, msi = 30) {
  # mind the gap
  if (length(tr[["t"]]) < 9) return(list(tr))
  splits <- unique(as.numeric(c(1, 1+which(diff(tr[["t"]]) > 2*msi), length(tr[["t"]]))))
  sub_traces <- list()
  for (i in 1:(length(splits)-1)) {
    t1 <- splits[i]
    t2 <- splits[i+1]
    if(is.na(t2)) next
    # at least nine samples per EMD
    if (as.integer(t2-t1) < 9) {
      next
    }
    sub_traces <- append(sub_traces,
      list(list(
      "t0" = tr[["t0"]]+t1,
      "t" = tr[["t"]][t1:(t2-1)],
      "y" = tr[["y"]][t1:(t2-1)]
    )))
  }
  if (length(sub_traces) == 0) return(list(tr))
  return(sub_traces)
}

idx <- 26 # 24
label <- get_locationcode(idx)
trace <- get_trace(idx, show_plot = FALSE)

msi <- median_sampling_interval(trace)
tiny_traces <- split_trace(trace, msi = max(c(15, msi)))

for (j in 1:length(tiny_traces)) {
  tra <- tiny_traces[[j]]
  emd <- get_emd_trace(tra)
  if (all(is.na(emd))) {
    message("skip")
    next
  }

  hlist <- plot_emd(emd, show = FALSE)
  h <- grid.arrange(hlist[[1]], hlist[[2]], hlist[[3]],
                          ncol = 1) # gridExtra
  ggsave(paste0("figures/test_emd_", idx, "_", label, "_", as.character(j), ".png", collapse = ""), h, dpi = 300) |> suppressMessages()
}
```


```{r plot-multistep-emd}

trace[["t0"]]


join_sig <- function(multi, part, stitch_na = TRUE) {
  
  get_part <- function(mule) t(mule[[part]])
  if (stitch_na) get_part <- function(mule) c(NA, t(mule[[part]]), NA)
  if (all(is.na(multi))) return(NA)
  return(as.matrix(do.call(cbind, lapply(multi, FUN = get_part))))
}

smooth_sg <- function(tr, days = 15, rectify = TRUE, ...) {
  msi <- median_sampling_interval(tr)

  fl <- 2*floor(0.5*days/msi)+1
  fl <- as.integer(max(c(fl, 3)))
  if (is.na(fl)) return(tr[["y"]])
  if (((fl %% 2) == 0.)) {
    fl <- fl + 1
  }

  sg <- pracma::savgol(
    tr[["y"]],
    fl = fl,
    ...
    )
  if (rectify) {
    outs <- which(abs(sg - tr[["y"]]) > 3*var(tr[["y"]]))
    sg[outs] <- tr[["y"]][outs]
  }
  return(sg)
}

execute_multistep_emd <- function(trace, apply_savgol = TRUE) {

  # trace parts
  msi <- median_sampling_interval(trace)
  parts <- split_trace(trace, msi = max(c(64, msi)))

  # plot(join_sig(parts, "t"), join_sig(parts, "y"))
  
  collect_y <- list()
  collect_emd1 <- list()
  collect_emd2 <- list()

  for (j in 1:length(parts)) {
    tr <- parts[[j]]
    if (all(is.na(tr))) next
    
    ## SSGS (Subtle Savitzky-Golay Smoothing)
    # tr <- parts[[4]]
    # t <- tr[["t"]]
    # y <- tr[["y"]]
    # sg <- smooth_sg(tr, days = 15,
    #                 rectify = TRUE,
    #                 forder = 4,
    #                 dorder = 0)
    # ggplot(NULL) +
    #   geom_line(aes(t, y), color = "darkgrey", linewidth = 0.5) +
    #   geom_line(aes(t, sg), color = "black", linewidth = 1.0) +
    #   theme_bw()
    trs <- tr
    if (apply_savgol) trs[["y"]] <- smooth_sg(tr, days = 31, forder = 8)
    collect_y <- append(collect_y, list(trs))

    emd1 <- get_emd_trace(trs)
    if (all(is.na(emd1))) {
      next
    }
    collect_emd1 <- append(collect_emd1, list(emd1))

    emd2 <- get_emd_trace(get_firstmode(emd1))
    if (all(is.na(emd2))) {
      next
    }

    # gather emd's
    collect_emd2 <- append(collect_emd2, list(emd2))


  }

  if (length(collect_emd1) == 0) return(NA)

  # plot(join_sig(collect_emd1, "time"), join_sig(collect_emd1, "amplitude"), type = "l")
  return(list(
    "trace" = trace,
    "parts" = parts,
    "y_sg" = collect_y,
    "emd1" = collect_emd1,
    "emd2" = collect_emd2
    ))
}

plot_multistep_emd <- function(multi_emd, show = FALSE) {

  trace <- multi_emd[["trace"]]
  parts <- multi_emd[["parts"]]
  collect_y <- multi_emd[["y_sg"]]
  collect_emd1 <- multi_emd[["emd1"]]
  collect_emd2 <- multi_emd[["emd2"]]

  ## plot
  # (i) the signal
  t <- trace[["t0"]] + as.numeric(join_sig(parts, "t"))
  y <- as.numeric(join_sig(parts, "y"))
  tsg <- trace[["t0"]] + as.numeric(join_sig(collect_y, "t"))
  sg <- as.numeric(join_sig(collect_y, "y"))
  amp1 <- join_sig(collect_emd1, "amplitude")
  amp2 <- join_sig(collect_emd2, "amplitude")
  if(all(is.na(amp1))) amp1 <- 0
  if(all(is.na(amp2))) amp2 <- 0
  sel <- join_sig(collect_emd2, "selection")
  amplitude <- as.numeric(sqrt(amp1[sel]^2 + amp2[sel]^2))
  h0 <- ggplot(NULL, aes(x = t, y = y)) +
    geom_point(alpha = 0.6, size = 1) +
    geom_step(na.rm = FALSE, alpha = 0.8, color = "darkgray") +
    geom_path(aes(tsg, sg), na.rm = FALSE, alpha = 0.8, color = "black", linewidth = 0.5) +
    geom_hline(yintercept = mean(sg),
      alpha = 0.8, color = "black", linetype = "dashed") +
    geom_hline(yintercept = mean(sg) - mean(amplitude),
      alpha = 0.8, color = "darkred", linetype = "dashed") +
    geom_hline(yintercept = mean(sg) + mean(amplitude),
      alpha = 0.8, color = "darkred", linetype = "dashed") +
    labs(x = NULL, y = "raw signal") +
    theme_bw() +
    theme(legend.position = "none")

  # (i) emd1
  if (all(is.na(collect_emd1))) {
    h1 <- ggplot(NULL)
    h1r <- ggplot(NULL)
  } else {
    ti <- trace[["t0"]] + as.numeric(join_sig(collect_emd1, "time"))
    firstmode <- as.numeric(join_sig(collect_emd1, "firstmode"))
    residual <- as.numeric(join_sig(collect_emd1, "residual"))
    amplitude <- as.numeric(join_sig(collect_emd1, "amplitude"))
    envelope_upper <- as.numeric(join_sig(collect_emd1, "envelope_upper"))
    envelope_lower <- as.numeric(join_sig(collect_emd1, "envelope_lower"))
    h1 <- ggplot(NULL, aes(x = t, y = y)) +
      geom_path(na.rm = FALSE, color = "grey", alpha = 0.5, linewidth = 0.5) +
      geom_path(aes(x = ti, y = firstmode), na.rm = FALSE, color = "black") +
      geom_path(aes(x = ti, y = envelope_upper), na.rm = FALSE, color = "orange", alpha = 0.8) +
      geom_path(aes(x = ti, y = envelope_lower), na.rm = FALSE, color = "orange", alpha = 0.8) +
      theme_bw() +
      labs(x = NULL, y = "first mode") +
      theme(legend.position = "none")

    h1r <- ggplot(NULL, aes(x = ti, y = residual)) +
      geom_path(na.rm = FALSE, color = "black") +
      geom_path(aes(x = ti, y = amplitude), na.rm = FALSE, color = "darkred", alpha = 0.8) +
      geom_path(aes(x = ti, y = -amplitude), na.rm = FALSE, color = "darkred", alpha = 0.8) +
      theme_bw() +
      labs(x = NULL, y = "residual") +
      theme(legend.position = "none")
  }
  
  if (all(is.na(collect_emd2))) {
    h2 <- ggplot(NULL)
    h2r <- ggplot(NULL)
  } else {
    ti2 <- trace[["t0"]] + as.numeric(join_sig(collect_emd2, "time"))
    secondmode <- as.numeric(join_sig(collect_emd2, "firstmode"))
    residual2 <- as.numeric(join_sig(collect_emd2, "residual"))
    amplitude2 <- as.numeric(join_sig(collect_emd2, "amplitude"))
    envelope2_upper <- as.numeric(join_sig(collect_emd2, "envelope_upper"))
    envelope2_lower <- as.numeric(join_sig(collect_emd2, "envelope_lower"))
    h2 <- ggplot(NULL, aes(x = t, y = y)) +
      geom_path(na.rm = FALSE, color = "grey", alpha = 0.5, linewidth = 0.5) +
      geom_path(aes(x = ti2, y = secondmode), na.rm = FALSE, color = "black") +
      geom_path(aes(x = ti2, y = envelope2_upper), na.rm = FALSE, color = "orange", alpha = 0.8) +
      geom_path(aes(x = ti2, y = envelope2_lower), na.rm = FALSE, color = "orange", alpha = 0.8) +
      theme_bw() +
      labs(x = NULL, y = "second mode") +
      theme(legend.position = "none")

    h2r <- ggplot(NULL, aes(x = ti2, y = residual2)) +
      geom_path(na.rm = FALSE, color = "black") +
      geom_path(aes(x = ti2, y = amplitude2), na.rm = FALSE, color = "darkred", alpha = 0.8) +
      geom_path(aes(x = ti2, y = -amplitude2), na.rm = FALSE, color = "darkred", alpha = 0.8) +
      theme_bw() +
      labs(x = NULL, y = "residual") +
      theme(legend.position = "none")
  }
  

  if (!show) {
    return(list(h0, h1, h1r, h2, h2r))
  }

  h <- grid.arrange(h0, h1, h1r, h2, h2r,
    widths = c(2, 1),
    layout_matrix = rbind(c(1, NA),
                          c(2, 3),
                          c(4, 5)
                          )
   ) # gridExtra
  return(h)

  
}

multi_emd <- execute_multistep_emd(trace, apply_savgol = TRUE) 

plot_multistep_emd(multi_emd, show = TRUE)
```



TODO: is sampling frequency related? (implicit temporal averaging)
-> partly compensated by savgol `fl` 


```{r emd-loop-sites}

emd_multistep_function <- function (data, index_column = "idx", progress = FALSE) {  
  indices <- sort(unique(data[[index_column]]))

  if (progress) {
    pb <- txtProgressBar(
      min = 0, max = length(indices),
      initial = 0, style = 1
    )
  }
  emd_data <- data %>% select(!!enquo(index_column))
  colnames(emd_data) <- c("idx")
  emd_data$emd_amp1 <- as.numeric(NA)
  emd_data$emd_amp2 <- as.numeric(NA)
  emd_data$emd_n <- as.numeric(NA)
  emd_data$emd_residual_qdiff <- as.numeric(NA)
  emd_data$emd_qdiff <- as.numeric(NA)

  for (i_emd in 1:length(indices)) { 
  
    if (progress) setTxtProgressBar(pb, i_emd)
    idx <- indices[i_emd]
    locationcode <- get_locationcode(idx)
    trace <- get_trace(idx, show_plot = FALSE)

    label2 <- paste(
        as.character(idx),
        get_info(idx, "cluster"),
        get_info(idx, "pp_code"),
        collapse = " "
      )

    # print(label2)

    ### EMD
    multi_emd <- tryCatch({execute_multistep_emd(trace, apply_savgol = TRUE)},
      error = {next})
    
    if (all(is.na(multi_emd))) next

    parts <- multi_emd[["parts"]]
    collect_y <- multi_emd[["y_sg"]]
    collect_emd1 <- multi_emd[["emd1"]]
    collect_emd2 <- multi_emd[["emd2"]]

    sg <- as.numeric(join_sig(collect_y, "y"))
    amp1 <- as.numeric(join_sig(collect_emd1, "amplitude"))
    amp2 <- as.numeric(join_sig(collect_emd2, "amplitude"))
    firstmode <- as.numeric(join_sig(collect_emd1, "firstmode"))
    secondmode <- as.numeric(join_sig(collect_emd2, "firstmode"))

    # collect info
    emd_data[emd_data$idx == idx, "emd_n"] <- suppressWarnings(length(sg))
    emd_data[emd_data$idx == idx, "emd_amp1"] <- suppressWarnings(mean(amp1, na.rm = TRUE))
    emd_data[emd_data$idx == idx, "emd_amp2"] <- suppressWarnings(mean(amp2, na.rm = TRUE))

    # quantile difference
    qdiff <- function(mode) suppressWarnings(diff(as.numeric(quantile(mode, c(0.1, 0.9), na.rm = TRUE))))

    emd_data[emd_data$idx == idx, "emd_qdiff"] <- qdiff(sg)
    if (!all(is.na(amp2))) {
      emd_data[emd_data$idx == idx, "emd_residual_qdiff"] <- qdiff(secondmode)
    } else if (!all(is.na(amp1))) {
      emd_data[emd_data$idx == idx, "emd_residual_qdiff"] <- qdiff(firstmode)
    }
    

    hlist <- plot_multistep_emd(multi_emd, show = FALSE)

    h <- grid.arrange(hlist[[1]], hlist[[2]], hlist[[3]], hlist[[4]], hlist[[5]],
      widths = c(2, 1),
      layout_matrix = rbind(c(1, NA),
                            c(2, 3),
                            c(4, 5)
                            ),
      top = label2
    ) # gridExtra

    ggsave(paste0("figures/emd/", idx, "_", locationcode, ".png", collapse = ""), h,
           dpi = 300, height = 16, width = 24, units = "cm") |> suppressMessages()


  }
  
  if (progress) close(pb) # close the progress bar

  # combine amplitudes
  emd_data <- emd_data %>% mutate(emd_ampsum = sqrt(emd_amp1^2 + emd_amp2^2))

  return(emd_data %>% rename(c("pp_idx" = "idx")))
}


label <- "empiricalmodes"
if (reload) {
  empty_cache_subfolder(paste0(label, collapse = ""))
}

numCores <- parallel::detectCores() - 2
registerDoParallel(numCores)

query_step <- function(i_query) {
  system(paste0("mkdir -p '", here::here(local_cache_folder, label), "'"))
  cluster_group <- names(data_subgroups)[[i_query]]
  subdata <- data_subgroups[[i_query]]
  storage_path <- here::here(local_cache_folder, label, cluster_group)
  if (file.exists(storage_path)) return(FALSE)

  print(cluster_group)
  emd <- suppressWarnings(emd_multistep_function(
    subdata,
    index_column = "pp_idx",
    progress = global_progress
  ))

  write_parquet(emd, sink = storage_path)
  # if (verbose) print(storage_path)
}

if (reload_emd) {
  foreach(i = 1:length(data_subgroups), .combine=rbind) %do% {
    query_step(i)
  } -> count_nonexisting
}

combine_subfolder_data(label)

```



```{r amplitude-map}
#| eval: false


fi <- "_empiricalmodes.parquet"
st_amp <- read_parquet(here::here(local_cache_folder, fi)) %>%
  right_join(
    backbone_data,
    by = "pp_idx",
    relationship = "one-to-one"
  ) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)
  
  mapview(st_amp,
    map.types = c("CartoDB.Positron", "OpenTopoMap", "CartoDB.DarkMatter", "OpenStreetMap" ),
    zcol = "emd_ampsum"
    # legend = FALSE
  )
}

```



# differential data

## pairing

Retrieving all pairs of observation wells
which are below a certain distance,

```{r pairing}
extent <- 512 # m

# meta_data %>% count(cluster)

pair_sites <- function(cluster_idx) {
  sub_data <- meta_data %>%
    filter(cluster == cluster_idx) %>%
    select(mp_idx, pp_idx, cluster, x, y, soilclass)
  
  sub_data %>% 
    left_join(sub_data,
      by = join_by(
        mp_idx > mp_idx, # only different locations
        soilclass == soilclass, # from the same soilclass
        cluster == cluster, # from the same soilclass
      ),
      suffix = c("1", "2"),
      relationship = "many-to-many"
    ) %>% mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    ) %>%
    filter(ds >= 1.0/100, ds < extent)

}

pairs <- bind_rows(
  lapply(
    sort(unique(as.integer(meta_data$cluster))),
    FUN = pair_sites
  )
) %>% arrange(ds)

pairs %>%
  ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()
```

It is okay to keep the close range sites for now.


## water level query

query: pp1, pp2, date, mTAW1, mTAW2, validation statusses

```{r level-query}

query_measurement <- function (pp1, pp2){

  w1 <- tbl(watina_dwh, "FactPeilMeting") %>%
    filter(PeilpuntWID == pp1) %>%
    select(PeilpuntWID, DatumWID, mTAW, PeilmetingStatusCode) %>%
    rename(
      pp_idx1 = PeilpuntWID,
      date1 = DatumWID,
      w1 = mTAW,
      pm_vld1 = PeilmetingStatusCode
    )

  w2 <- tbl(watina_dwh, "FactPeilMeting") %>%
    filter(PeilpuntWID == pp2) %>%
    select(PeilpuntWID, DatumWID, mTAW, PeilmetingStatusCode) %>%
    rename(
      pp_idx2 = PeilpuntWID,
      date2 = DatumWID,
      w2 = mTAW,
      pm_vld2 = PeilmetingStatusCode
    )

  pair_diff <- w1 %>%
    inner_join(w2,
    join_by(date1 == date2),
    relationship = "many-to-many"
    ) %>%
    collect %>%
    mutate(
      across(date1, function(x) as.Date(as.character(x), format = "%Y%m%d")),
      dw = w2 - w1
    ) %>%
    rename(date = date1) %>%
    select(pp_idx1, pp_idx2, pm_vld1, pm_vld2, date, w1, w2, dw)

  return(pair_diff)
}


# pairs_test <- pairs
# pairs_test <- pairs_test %>%
#   head(100)

query_waterlevel_differences <- function (
    sub_pairs,
    progress = FALSE
  ) {
  n_rows <- nrow(sub_pairs)
  
  if (progress) {
    pb <- txtProgressBar(min = 0, max = n_rows,
                           initial = 0, style = 1)
  }
  rowwise_query <- function(i) {
    if (progress) setTxtProgressBar(pb, i)
    pp1 <- sub_pairs[[i, "pp_idx1"]]
    pp2 <- sub_pairs[[i, "pp_idx2"]]
    return(query_measurement(pp1, pp2))
  }
  
  
  diff_data <- bind_rows(
      lapply(
        1:n_rows,
        FUN = rowwise_query
      )
    )
  
  if (progress) close(pb)

  return(diff_data)
}

```


```{r waterlevel-split-apply-combine}


# knitr::kable(head(pairs))
pair_subgroups <- split(
  pairs,
  split_balanced(
    pairs,
    count_column = "cluster",
    n_bins = 2**10,
    show = TRUE
  )
)


label = "diffdata"
if (reload) {
  empty_cache_subfolder(paste0(label, collapse = ""))
}


numCores <- parallel::detectCores() - 2
registerDoParallel(numCores)

query_step <- function(i) {
  system(paste0("mkdir -p '", here::here(local_cache_folder, label), "'"))
  cluster_group <- names(pair_subgroups)[[i]]
  subdata <- pair_subgroups[[i]]
  storage_path <- here::here(local_cache_folder, label, cluster_group)
  if (file.exists(storage_path)) return(FALSE)

  print(cluster_group)
  waledi <- query_waterlevel_differences(
    subdata,
    progress = global_progress
  )

  write_parquet(waledi, sink = storage_path)
  # if (verbose) print(storage_path)
  return(TRUE)
}

foreach(i = 1:length(pair_subgroups), .combine=rbind) %do% {
  query_step(i)
} -> count_nonexisting

combine_subfolder_data(label)


```


# Reload All Data

All data has been stored to disk.
Here we will reload some of it, for demonstration and testing.


Metadata:

```{r reload-all-data}
source("./auxiliary_data_queries.R")
meta <- join_auxiliary_cache(label = "metadata")
# .data <- meta
meta %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") %>%
  glimpse

```

Quantiles:

```{r reload-quantiles}
meta %>%
  join_auxiliary_cache("quantiles", index_column = "pp_idx") %>%
  glimpse

```


Waterlevel differences:

```{r reload-diffdata}
diffdata <- join_auxiliary_cache(label = "diffdata")
meta %>%
  select(pp_idx, region, cluster, soilclass) %>%
  right_join(diffdata,
    by = join_by(pp_idx == pp_idx1),
      relationship = "many-to-many" 
  )

```
