---
title: "Water Level Models"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---


# Introduction 

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("mapview")   |> void() # map visualization
# library("gstat") # geospatial statistics
# library("sdmTMB") # species distribution models Template Model Builder


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# option to work offline; requires previous data download
work_offline <- FALSE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 2 # minimal cluster member count; fewer memeber clusters are removed
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")

# the user may choose to reload raw data
force_data_reassembly <- FALSE
reload <- force_data_reassembly || !file.exists(data_export_file)
```


# Clustered Peilpunt Data

## Strategy

- getting the peilpunt locations
- clustering
- pairing
- soilclass, distance, etc.


## Loading

```{r raw-data-query-dbplyr}

max_filterdepth <- 5.8 # 3 # (m) # choice to focus on shallow gw (ecologically relevant for GW_03.3 and GW_05.1)

peilpunten <- tbl(watina_dwh, "dimPeilpunt") %>%
  filter(PeilpuntStatusCode == "VLD") %>% 
  filter(LandCode == "BE") %>%
    mutate(
      filterdepth = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte/2)
    ) %>% 
  filter(filterdepth <= max_filterdepth, filterdepth > 0) %>% 
  left_join(
    tbl(watina_dwh, "dimMeetpunt"),
    by = join_by(MeetpuntWID == MeetpuntWID),
    relationship = "many-to-many"
  ) %>%
  filter(!is.na(PeilpuntWID)) %>% 
  select(
    PeilpuntWID,
    PeilbuisLengte,
    ReferentieNiveauMaaiveld,
    ReferentieNiveauTAW,
    FilterLengte,
    filterdepth,
    PeilpuntStatusCode,
    PeilpuntCode,
    MeetpuntWID,
    PeilpuntXCoordinaat,
    PeilpuntYCoordinaat,
  ) %>%
  rename(
    mp_idx = MeetpuntWID,
    pp_idx = PeilpuntWID,
    pblen = PeilbuisLengte,
    ref_mv = ReferentieNiveauMaaiveld,
    ref_taw = ReferentieNiveauTAW,
    flength = FilterLengte,
    fdepth = filterdepth,
    pp_code = PeilpuntCode,
    status = PeilpuntStatusCode,
    x = PeilpuntXCoordinaat,
    y = PeilpuntYCoordinaat,
  ) %>%
  collect %>% 
  mutate(
    region = substring(pp_code, first = 1, last = 3)
  ) %>%
    filter(region != "VBS") # remove Luxemburg

# print(colnames(peilpunten))
nrow(peilpunten)

peilpunten %>%
  ggplot(aes(x = fdepth)) +
  geom_histogram(bins = 128, fill = "lightgray", color = "black") +
  theme_bw()
```


## Clustering

Clustering the data with a cluster range of `{r} cluster_dist` m,
removing all clusters which contain only one location.

```{r cluster-join}

pp_data <- peilpunten %>%
  join_clusters(index_column = "pp_idx", characteristic_distance = cluster_dist) %>%
  remove_underpopulated_clusters(
    minimum_cluster_member_count = min_cluster_size
  ) 

knitr::kable(head(pp_data))
```


Cluster size:

```{r cluster-size}

pp_data %>%
  count(cluster) %>%
  arrange(desc(n)) %>%
  ggplot(aes(x = n)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()

# levels(pp_data$cluster)
```


```{r cluster-location}

st_clust <- pp_data %>%
  group_by(region, cluster) %>%
  summarize(x = mean(x), y = mean(y), count = n()) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_clust,
  map.types = c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"),
  zcol = "region",
  cex = "count",
  legend = FALSE
)

```

## duplicates

Mental note: some locations have ambiguous co-ordinates

```{r location-duplicates}
duplicates <- pp_data[, c("mp_idx", "pp_idx", "x", "y")]
duplicates <- duplicates %>% 
  filter(mp_idx %in% (duplicates %>% count(mp_idx) %>% filter(n>1) %>% pull(mp_idx))) %>%
  arrange(mp_idx)
knitr::kable(duplicates)
```

## soilclass

```{r soilclass-join}
source("./auxiliary_data_queries.R")

pp_data <- pp_data %>%
    join_soilclass(index_column = "pp_idx")

st_soil <- pp_data %>%
  group_by(soilclass, region, cluster) %>%
  summarize(x = mean(x), y = mean(y), count = n()) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_soil,
  map.types = c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"),
  zcol = "soilclass",
  cex = "count"
  # legend = FALSE
)

```

## elevation

... once we're at it.

```{r elevation-join}

if (reload) {
  pp_data <- pp_data %>%
    join_elevation(index_column = "pp_idx")
  
  ## "Oh No!", there were duplicates.
  # data <- pp_data
  # index_column = "mp_idx"
  # coordinate_columns <- c("x", "y")
  # elevation_lookup <- query_elevation(data, index_column = "mp_idx")
  # elevation_grouped <- elevation_lookup %>% 
  #   filter(mp_idx %in% (elevation_lookup %>% count(mp_idx) %>% filter(n>1) %>% pull(mp_idx)))
  # elevation_lookup %>% summarize(across(everything(), ~ mean(.x, na.rm = TRUE)), .by = !!index_column)
  # data <- data %>%
  #   dplyr::left_join(
  #     elevation_lookup,
  #     by = index_column,
  #     relationship = "many-to-one"
  #   )

  write_parquet(pp_data, sink = data_export_file)
}
pp_data <- read_parquet(data_export_file)
```


```{r elevation-map}
st_elevation <- pp_data %>%
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_elevation,
  map.types = c("OpenTopoMap", "CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap" ),
  zcol = "elevation_dhmv"
  # legend = FALSE
)

```


## waterdistance

```{r elevation-join}
if (reload) {
  pp_data <- pp_data %>%
    join_waterdistance(index_column = "pp_idx")

  ## today, nothing works as supposed.
  # source("./auxiliary_data_queries.R")
  # pp_test <- pp_data %>%
  #   head(100) %>% 
  #   join_waterdistance(index_column = "pp_idx")
  # wd_lookup <- query_waterdistance(pp_data, index_column = "pp_idx")

  # water_lookup <- wd_lookup
  # meters <- function(x) units::set_units(x, "m")
  # water_lookup <- water_lookup %>%
  #   mutate(
  #     wata_dist_class = factor(dplyr::case_when(
  #       wata_min_dist <= meters(50) ~ "mid",
  #       wata_min_dist <= meters(5) ~ "close",
  #       .default = "far"
  #   )))

  write_parquet(pp_data, sink = data_export_file)
}
pp_data <- read_parquet(data_export_file)
```

# amplitude

## purpose
Some water levels oscillate a lot,
others are effectively flat.
We seem to miss a relevant co-variate which can explain the pronounced differences in amplitude.


Here, I extract the "momentary amplitude" of the water level measurements.
Then, I will measure their relation to water distance, elevation, *et cetera*. 


## EMD implementation

From <https://github.com/inbo/tutorials/tree/emd>.

```{r r-firstmode-function}
#' extracts from a signal (with an optional timeline vector)
#'
#' extra arguments will propagate to `findpeaks`
#'
#' @returns a list with
#'   - first empirical mode: mean of envelope
#'   - envelope: the upper and lower peak amplitude envelope
#'   - select: boolean vector of time subset
#'   - extrema: peaks and troughs
#'
#' 
#' usage:
#' > emd <- extract_firstmode(y, t = x, threshold = 0.4, minpeakdistance = 20)
extract_firstmode <- function (signal, t = NULL, ...) {

  # general purpose time vector
  if (is.null(t)) {
    t <- seq(0., 1., length.out = length(signal))
  }

  ## 1. find peaks
  peaks <- sort(pracma::findpeaks(signal, ...)[, 2])
  troughs <- sort(pracma::findpeaks(-signal, ...)[, 2])
  extrema <- sort(c(peaks, troughs))
  
  # if (peaks[1] < troughs[1]) {
  #   peaks <- c(peaks, length(signal))
  #   troughs = c(1, troughs)
  # } else if (peaks[1] > troughs[1]) {
  #   peaks = c(1, peaks)
  #   troughs = c(troughs, length(signal))
  # }

  ## 2. interpolate -> envelope
  select <- t >= t[max(c(peaks[1], troughs[1]))] & t <= t[min(c(peaks[length(peaks)], troughs[length(troughs)]))]  
  select <- function (pt) t >= t[pt[1]] & t <= t[tail(pt, 1)]
  
  # ti <- as.numeric(t[select])
  # yi <- as.numeric(signal[select])
  # py <- interpolators::evalInterpolator(interpolators::iprPCHIP(t[peaks], as.numeric(signal[peaks])), ti)
  # ty <- interpolators::evalInterpolator(interpolators::iprPCHIP(t[troughs], as.numeric(signal[troughs])), ti)
  py <- interpolators::evalInterpolator(interpolators::iprPCHIP(
    t[peaks],
    as.numeric(signal[peaks])),
    t[select(peaks)]
  )
  if (peaks[1] > 1) py <- c(rep(py[1], peaks[1]), py[2:length(py)])
  if (tail(peaks, 1) < length(t)) py <- c(py, rep(tail(py, 1), length(t)-tail(peaks, 1)))
  # plot(t, py)

  ty <- interpolators::evalInterpolator(interpolators::iprPCHIP(
    t[troughs],
    as.numeric(signal[troughs])),
    t[select(troughs)]
  )
  if (troughs[1] > 1) ty <- c(rep(ty[1], troughs[1]), ty[2:length(ty)])
  if (tail(troughs, 1) < length(t)) ty <- c(ty, rep(tail(ty, 1), length(t)-tail(troughs, 1)))
  # plot(t, ty)

  ## 3. empirical mode
  envelope <- cbind("upper" = py, "lower" = ty)
  rownames(envelope) <- t
  firstmode <- rowMeans(envelope)
  residual <- signal - firstmode
  amplitude <- abs(py-ty)/2

  return(list(
    "time" = t,
    "signal" = signal,
    "firstmode" = firstmode,
    "envelope" = envelope,
    "select" = select,
    "extrema" = extrema,
    "residual" = residual,
    "amplitude" = amplitude
  ))
}


plot_emd <- function(emd, raw_trace = NULL) {


  ti <- emd[["time"]]
  signal <- emd[["signal"]]
  firstmode <- emd[["firstmode"]]
  envelope <- emd[["envelope"]]
  residual <- emd[["residual"]]
  amplitude <- emd[["amplitude"]]

  if (!is.null(raw_trace)) {
    x <- raw_trace[["t"]]
    y <- raw_trace[["y"]]
  } else {
    x <- ti
    y <- signal
  }

  h0 <- ggplot(NULL, aes(x = x, y = y)) +
    geom_point() +
    geom_step(alpha = 0.3, color = "darkgray")+
    theme_bw() +
    theme(legend.position = "none")

  h1 <- ggplot(NULL, aes(x = ti, y = firstmode)) +
    geom_line(color = "black") +
    geom_point(aes(x = ti, y = signal)) +
    geom_line(aes(x = ti, y = signal), color = "darkgray", alpha = 0.4) +
    geom_line(aes(x = ti, y = envelope[,1]), color = "orange", alpha = 0.4) +
    geom_line(aes(x = ti, y = envelope[,2]), color = "orange", alpha = 0.4) +
    theme_bw() +
    labs(y = "first mode") +
    theme(legend.position = "none")
  
  h2 <- ggplot(NULL, aes(x = ti, y = residual)) +
    geom_line(color = "black") +
    geom_line(aes(x = ti, y = amplitude), color = "darkred", alpha = 0.8) +
    geom_line(aes(x = ti, y = -amplitude), color = "darkred", alpha = 0.8) +
    theme_bw() +
    labs(x = "time (arb. units)", y = "residual") +
    theme(legend.position = "none")

  grid.arrange(h0, h1, h2, ncol = 1) # gridExtra
}

```


## explore

```{r explore-emd}

get_trace <- function(pp_idx, show_plot = FALSE) {
  # pp_idx = 24

  peilmeting <- tbl(watina_dwh, "FactPeilMeting") %>%
    filter(PeilpuntWID == pp_idx) %>%
    select(DatumWID, mTAW) %>%
    arrange(DatumWID) %>%
    collect
  
  peilmeting <- peilmeting %>%
    mutate(date = as.Date(as.character(DatumWID), format = "%Y%m%d"))

  if (show_plot){
    h <- peilmeting %>%
      ggplot(aes(x = date, y = mTAW)) +
        geom_point() +
        geom_step(alpha = 0.3, color = "darkgray")
    print(h)
  }

  t <- peilmeting %>% pull(date) 
  t <- difftime(t, t[1], units = "days")
  y <- peilmeting %>% pull(mTAW) 
    
  return(list("t" = as.numeric(t), "y" = y))
}

get_emd_trace <- function(tr) extract_firstmode(tr[["y"]], t = tr[["t"]])
get_firstmode <- function(emd) list("t" = emd[["time"]], "y" = emd[["firstmode"]])
mean_amplitude <- function(emd) mean(emd[["amplitude"]])

# example
trace <- get_trace(24, show_plot = FALSE)
emd <- get_emd_trace(trace)
# plot_emd(emd)
emd2 <- get_emd_trace(get_firstmode(emd))
plot_emd(emd2, raw_trace = trace)
var(2*trace$y)
mean_amplitude(emd) + mean_amplitude(emd2)

t <- emd2[["time"]]
amp <- emd2[["amplitude"]] + emd[["amplitude"]][emd2[["select"]]]
plot(t, amp, type = "l", col = "orange", ylim = c(0, max(amp)))
lines(trace[["t"]], abs(trace[["y"]] - mean(trace[["y"]])))

```


# pairing

Retrieving all pairs of observation wells
which are below a certain distance,

