---
title: "Water Level Variograms: Daily Data"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
knitr:
  opts_chunk:
    echo: true
---


# summary (tl;dr)


Differences compared to previous `xG3` analysis:

- daily water levels were used (watina `PeilMeting`)
- no de-trending was applied
- pooling/averaging ([@sec-pooling]):
  - observation pairs (as alternative to "random intercept", suggested by Floris)
  - distance bins
  - soilclass
- use of mean absolute difference $\langle dw \rangle$, instead of (semi)variance ($\gamma$, [@sec-parameters])
- variable bin width to achieve homogeneous sample size ([@sec-equalbins])


And in terms of the outcome ([@sec-variograms]):

- *light* soil is more variable, i.e. higher differences are expected to occur at closer distance (range: $\approx 30.6 m$)
- confirmative evidence that *heavy* soil shows longer range of spatial correlation (range: $\approx 62 m$)
- the $\langle dw\rangle$ (y-axis) of *heavy* soil ($\approx 1.3 m$) is smaller than on *light* soil ($\approx 1.8 m$), but overall value ranges are more similar (therefore plausible) than for $\gamma$
- *peat* data fits a rather long range ($> 100 m$) and differences are small ($\approx 0.5 m$), but there are outliers

::: {.callout-tip icon=false}
## Thresholds
According to these variogram models, the $1 cm$ threshold would be exceeded at distances of $1.2 m$ ("light" soil), $3 m$ ("heavy" soil), and $9 m$ ("peat"), respectively.
:::


::: {.callout-note icon=false}
## Lessons learned (FM):

- semivariance is calculated differently, i.e. not via R's `var` (see [@sec-parameters])
- $\langle dw \rangle$ is the more practical choice (semivariance has problems with pooling squared values)
- sample size per bin must be monitored, equal and large is better
- outliers, especially on short distance, have a major drag on the regression. 

In consequence, the results below ([@sec-variograms]) are the first convincing Matérn regressions I saw (good times!).

:::


# preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("ggcorrplot")|> void() # correlation, grid visualization
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges


if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")
diffdata_export_file <- paste0("data/waterlevel_diffdata_cl", cluster_dist, ".parquet")

```


The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.



```{r load-data}

pp_data <- read_parquet(data_export_file)

diff_data <- read_parquet(diffdata_export_file)

pp_join <- pp_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope = mean(c(slope_r2.5m2, slope_r2.5m1)) 
  )

```


# binning (preview)


Binning and joining the data is necessary, yet there is some procedural flexibility in how to apply it.

First and foremost, data is filtered to include spatial distance of up to `extent` ($512 m$).

```{r parameters}
#| code-fold: false
extent <- 512 # m
n_bins <- 32
```

Binning is optional, with `n_bins =` `{r} n_bins` giving the number of bins (equally-spaced at `{r} extent/n_bins` m); note however that the most relevant binning strategies might be two others:

1. Binning with fixed bin width of `1 m`.
2. Binning with variable bin width, to achieve **equal sample size per bin**.

More on that below.



```{r glimpse-data}

# diff_meta %>% glimpse
# diff_data %>% glimpse

diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    ),
    distance_bin = cut(ds,
      breaks = seq(
        -1,
        extent,
        length.out = extent+2
      )
    ), # cut to "meters" bin
    distance_widebin = cut(ds,
      breaks = seq(
        -extent/n_bins*2,
        extent,
        length.out = n_bins/2+2
      ) # cut to 8m bins
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(dw <= ds) # remove outliers

# glimpse(diff_all)


diff_widegroup <- diff_all %>%
  summarize(
    cnt = n(),
    #cnt = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*cnt)*sum(dw^2, na.rm = TRUE),
    dw = 1/(cnt)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
    .by = c(pair, distance_widebin)
  ) %>% ungroup %>% arrange(ds)
```


Briefly inspecting the wide bins, sample size (circle extent) is balanced; none outstandingly big.

```{r glimpse-data}
#| label: fig-bubbletea
#| fig-cap: "An irrelevant bubble tea plot. The size of the bubbles scales with the number of observations. Yes, this was only a humble first glimpse at the data; yes, I used `sqrt` as a 'cheat' to limit bubble size differences; yes, we will do better below. But the colors are nice."
diff_widegroup %>%
  ggplot(aes(x = ds, y = log(1.0+dw),
         color = soilclass)) +
  geom_point(aes(cex = sqrt(cnt)), alpha = 0.16) +
  theme_bw()
```



# ridge plots

... with the little "extra" to regular difference-distance plots.


```{r diff-ridges-dw}
#| label: fig-ridges
#| fig-cap: "Visualization of distance-dependent difference with ridges. With increasing distance (y-axis), the distribution of water level differences (x-axis, grey density distributions) increases (indicated by quantile lines)."
diff_widegroup %>%
  ggplot(aes(x = dw, y = distance_widebin)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 3,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi) +
  theme_bw()
```


```{r diff-ridges}
#| eval: false
#| include: false
diff_widegroup %>%
  ggplot(aes(x = semivar, y = distance_widebin)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 2,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi/2) +
  theme_bw()
```

This is a promising pattern. 
We better expect to re-capture it in the analysis below.


# grouping and binning

## pooled data: what to pool? {#sec-pooling}

Re-applying the groups with more fine-grained binning.

```{r bin-data}

diff_grouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
    .by = c(pair, distance_bin)
  ) %>% ungroup %>% arrange(ds)


combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)

diff_binned <- diff_grouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    dw_all = combine_grouped(dw, n_obs),
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_all = combine_grouped(semivar, n_obs),
    semivar_groups = mean(semivar, na.rm = TRUE),
    .by = c(distance_bin, soilclass)
  ) %>% ungroup %>% arrange(ds)


knitr::kable(rbind(head(diff_binned, 5), tail(diff_binned, 5)))
```


::: {.callout-note icon=false}
## Pooling

There are multiple variables to consider for pooling.

- pairs of locations (pseudo-replication)
- the distance bins
- soilclass

Even the grouping order matters, it better be in the listed order.
:::


## choice of difference parameter {#sec-parameters}

There are at least two possible distance parameters we can apply to this grouped data.
This has implications for grouping, with some practical trouble for semivariance (see [@sec-semivar-binning]).


First, we have **conventional semivariance** ["Method-of-Moments Estimator" according to @Cressie1993, p. 69, eqn. 2.4.2].
This originally goes back to @Matheron1962, and only holds under the stationarity assumption ("constant mean").

$$ \gamma = \frac{1}{2N} \sum\limits_{i,j} \left(z_j - z_i\right)^2 $$

```{r plot-semivariance-distance}
#| label: fig-semivariance-difference-plot
#| fig-cap: "Semivariance-distance plot: the increase with distance is not as pronounced on the semivariance measure. Log scale for better visualization."
diff_binned %>%
  ggplot(aes(x = ds, y = log(1+semivar_groups), color = soilclass)) +
  geom_point() 
```

This equation has the typical structure of "sum of square differences", the normalization with sample size $N$ effectively makes it an average. 
Yet the "square" is not inverted by a "square root", which has some pitfalls [@sec-semivar-binning].


Therefore, I suggest a simpler variant: **mean absolute difference** (MAD) in water level.

$$ \langle dw\rangle = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$


```{r plot-dw-distance}
#| label: fig-difference-distance-plot
#| fig-cap: "Difference-distance plot: rough top-down visualization to inspect if water level differences increases with distance. Apparently, it does. Log scale for better visualization."
diff_binned %>%
  ggplot(aes(x = ds, y = log(1+dw_groups), color = soilclass)) +
  geom_point() 
```


::: {.callout-warning}
Neither of these is the actual variance, as in `?var` in R, although with his limited maths talent the author cannot exclude that variance and "semivariance" are related in the stationary case.
:::


Let's go MAD!



## equally-sized bins {#sec-equalbins}

Re-visit the binning.

Sample size is rather rugged:

```{r semivar-regression-samplesize}
#| label: fig-sample-size
#| fig-cap: "Sample size per bin in standard grouping. Dark blue: per location pair, i.e. repeated observations from the same location pairs are averaged. Light blue: per observation (divided by 100 to fit the axis)."

diff_binned %>%
  ggplot(aes(x = ds, y = cnt)) +
  geom_step(color = "darkblue") +
  geom_step(aes(x = ds, y = n_obs/100), color = "lightblue", alpha = 0.6) +
  theme_bw()
```


There are some *rather empty* bins touching the axis; they probably account at least in part for the outliers below.
"Misery loves company:" some other bins are *spikes*, probably due to pseudo-replication (repeated sampling at high frequency).


How about using equal-sized bins, i.e. buckets of variable width which are all filled to about the same level?

Crucially, we'll have to group by location pair first because otherwise repeated measures fill single buckets.


```{r equibin-preparation}
#| code-fold: false

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(extent/2)
    ) # cut to equal sized bins
  ) %>%
  ungroup

    
diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    dw_all = combine_grouped(dw, n_obs),
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_all = combine_grouped(semivar, n_obs),
    semivar_groups = mean(semivar, na.rm = TRUE),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

```
    

```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Sample size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes."
diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step() +
  geom_hline(yintercept = 0, color = "darkgrey") +
  theme_bw()
```

Seems more balanced, let's see the outcome.


# regression

Here come the general toolkit for regression, based on `optim` ([docs](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/optim)).

Note that root-mean square difference (RMSD) is optionally distance-weighted, by dividing each observed `prediction-actual` difference by square root of the distance.


```{r regression-tools}

# wrap a regression function to generate residuals
# the result is the parameter to be minimized.
wrap_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  return(sqrt(mean(differences^2)))
}

distweighted_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  #differences <- log(1.0+y) - log(1.0+predictions)
  # differences <- 1000*differences * (1+1/x^2)
  # return(sqrt( sum(differences^2)/sum((1+1/x^2)) ))
  differences <- 1000*differences /sqrt(x)
  return(sqrt( mean(differences^2) ))
}

# this can turn regression output into a usable function.
create_prediction_function <- function(regressor, results) {
  fcn <- function (x) {
    regressor(x, results$par)
  }

  return(fcn)
}

calculate_limit <- function(orsl, threshold = 0.01) {
  par <- orsl$par
  par[3] <- 0 # no nugget
  # par[4] <- 1. # regular shape
  test_x <- seq(0., par[2], length.out = 1000)
  test_y <- matern_function(test_x, par)
  limit <- test_x[min(which(test_y > threshold))]
  return(limit)
}

# a uniform way to print results
print_regression_results <- function(orsl, label = "") {
  par <- paste(round(orsl$par, 4), collapse = ", ")
  conv <- orsl$convergence
  eps <- orsl$value

  threshold <- calculate_limit(orsl)

  print(
    sprintf("%s: conv %i at (%s), mse %.1f", label, conv, par, eps)
  )
  print(
    sprintf("==> Threshold of dw > 1cm reached at %.3f m distance.", threshold)
     
  )
}

# note: the `epsilon` can be calculated manually with the formula:
#   sum((predictor_function(x) - y)^2)


# Finally, a quick histogram plot of residuals.
plot_residuals_histogram <- function (x, y, predictor_function, ...) {
  residuals <- predictor_function(x) - y
  ggplot(NULL, aes(x = residuals)) +
    geom_histogram(...) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    theme_bw()
}


```


The Matérn parametrization below is related to the parametrs on semivariogram plots:

```
      ^
      | Range
      |<--->
Sill  |     :____,model
      |    /:
 VAR  |   / :
 IOG  |  /  :
 RAM  | /   :
      |/    :
Nugget| Cor.: Uncor.
     _|_____:___________>
      |              lag distance
```

We get:
$$ d\gamma = \gamma_{max} - \gamma_{min}$$
$$ \gamma = \gamma_{max} - d\gamma \cdot \left( \frac{\sqrt{2\nu}}{\sigma} x \right)^{\nu} \cdot K_{\nu}\left( \frac{\sqrt{2\nu}}{\sigma} x \right) \cdot \frac{1}{2^{\nu - 1} \cdot \Gamma\left(\nu\right)} $$

With sill $\gamma_{max}$, nugget $\gamma_{min}$, turning point $\sigma$ (related to range), and a shape parameter $\nu$ that controls the form of the sigmoid[^1]. 
(Compare [wikipedia](https://en.wikipedia.org/wiki/Mat%C3%A9rn_covariance_function).)


[^1]: In practical application, I soft-fix $\nu$ to a value of one ($\nu = 1.0$), and only give it moderate freedom in case of visual shape mismatches. 


In code:

```{r matern-definition}
#| code-fold: false

matern_function <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}

```

This particular parametrization of nugget, scale, and sill will facilitate optimization by ensuring that the nugget is always smaller than the sill.


Because we repeatedly apply the procedure, a couple of convenience functions below.

```{r regression-matern-helpers}

fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results, y_label = "semivariance (m²)") {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


(Finally, one more not-so-general helper function.)

```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x

regression_by_soilclass <- function(
    sc, reg_var, maxx = extent, maxy = NULL, use_equibin = TRUE,
    ...) {
  if (use_equibin) {
    diff_sc <- diff_equibinned %>% filter(soilclass == sc)
  } else {
    diff_sc <- diff_binned %>% filter(soilclass == sc)
  }
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))
  # firsthalf <- x <= extent/2
  # y <- y[firsthalf]
  # x <- x[firsthalf]

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]
  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
    # control = list("fnscale" = 1e-4),
  )
  
  print_regression_results(matern_fit, label = "regression:")
  ylabel <- "mean water level difference (pair-averaged)"
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"
  g <- plot_matern(matern_fit, y_label = ylabel)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, colour = "red", alpha = 0.2) 
  return(g)
}



```


# variograms {#sec-variograms}

## water level difference (MAD)

Feel free to adjust regression parameters; maybe most importantly, the outcome measure.

```{r dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
# # see also the "semivariance" options, below.
```

There are also axis upper limits, start values, parameter limits, RMSD weighting, and binning to choose.
Selected here are the most plausible settings (MAD, equal-sized bins, distance weighted, with sensible parameter starts and limits).


::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
#| label: fig-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass("heavy",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 6)

```


### light

```{r dw-regression-light}
#| label: fig-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass("light",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 6)
```


### peat

```{r dw-regression-peat}
#| label: fig-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass("peat",
  reg_var = column,
  maxx = 256, maxy = 2,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 6)
```

::: {.callout-warning}
Capped: $<= 2$ (see [@sec-peat-outliers])
:::


:::




# next

close-range variograms of further subsets:

- slope-classes
- classes of elevation (difference)?
- optional water distance filter


<hr>

# appendix {.appendix}


## why not semivariance? {#sec-semivar-binning}

Recall the two equations above:

$$ \langle dw\rangle = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$


$$ \gamma = \frac{1}{2N} \sum\limits_{i,j} \left(z_j - z_i\right)^2 $$


From these equations, semivariance ($\gamma$) involves subtle differences compared to the more intuitive MAD ($\langle dw\rangle$):

- $\gamma$ uses squared values
- $\gamma$ is divided by two, i.e. it calculates **semi**-variance (*triv.*)
- This is NOT the real variance! 

Therefore, semivariance should better be called: **half mean square difference** (HMSD).


Semivariance / HMSD be tried out with the code above by using either of the following:

```{r semivar-column}
#| code-fold: false

if (FALSE) {
  column <- "semivar_all"
  column <- "semivar_groups"
}
```

::: {.callout-warning}
The semivariance (i.e. HMSD) was not used here!
Instead, I went MAD.
:::

This has two justifications:

- observation of the outcome
- consideration about pooling data


Due to the square, pooling is non-trivial, which is best illustrated by a simple example. 

Assume we measure the following water levels, in meters.

```
day 1: z1 = 0.1
day 1: z2 = 0.5

day 2: z1 = 0.1
day 2: z2 = 0.3

day 3: z1 = 0.1
day 3: z3 = 0.4

```

Differences are:

```
d1 = 0.5-0.1 = 0.4
d2 = 0.2
d3 = 0.3

```

Plain average would be:
```
<d> = (0.4 + 0.2 + 0.3) / 3 = 0.3
```


The first two observations are pseudo-replications, because they come from the same sites (`z1` and `z2`).
We better average them:

```
<d> = ((0.4 + 0.2)/2 + 0.3) / 2 = 0.3
```

Of course, this is no co-incidence: I chose the values to yield identical results in grouped averaging.


Compare this to the *squared* average difference used in semivariance:

```
γ = (0.4^2 + 0.2^2 + 0.3^2) / (2*3) = 0.0483
γ = (((0.4 + 0.2)/2)^2 + 0.3^2) / (2*3) = 0.105
```

... but then we square the number of observations (positive integer, large numbers of observations become huge)... 
Or maybe average the square differences?

```
γ = ((0.4^2 + 0.2^2)/2 + 0.3^2) / (2*3) = 0.03167
```

... but then we square our generally small differences (measured in meters), making them even smaller.
Mr. Πυθαγόρας could have anticipated this.


::: {.callout-note}
This illustrates that, because water level differences are generally small, yet repeated sampling gives a lot of these small differences per location pair,
the square structure of the semivariance formula can complicate grouping of repeated observations.


The issue of squared values is consistent with the huge soilclass difference observed previously in *heavy versus light* soil.
:::



This is in part an *ex post* rationalization of the scattered values I see when attempting semivariance.
Primary reason was inspection of the outcome, and non-plausible Matérn regressions when using semivariance.
Another benefit is **unit retention**: differences in meters are easier to interpret and compare than the square meters produced by variance formulas.


Using $\langle dw\rangle$ just works better.


## peat outlier removal {#sec-peat-outliers}

For the regression of the `peat` subset, values above a given $\langle dw\rangle$ were excluded from regression.
This exclusion is not based on data-intrinsic reasons, but on the plausibility of the regression outcome.

For completeness and comparison, behold the regression without the outlier removal.


```{r dw-regression-peat-with-outliers}
#| label: fig-dw-regression-peat-with-outliers
#| fig-cap: "Matérn regression of water level distance (`dw (m)`) on 'peat', this time with the outliers included."
regression_by_soilclass("peat",
  reg_var = "dw_groups",
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 6)
```

The higher difference values attract the regression, leading to a much worse regression for the majority of points.
This issue is emphasized by the larger absolute values of the outliers, leading to larger RMSD; the regression method above is non-robust.


::: {.callout-note}
In conclusion, outlier removal is not ideal, but a necessary workaround to avoid well-known regression issues.
:::



## comparing co-variates

Just to show off, we have a number of useful co-variates at hand which might be insightful by themselves.


We can learn a lot about covariates which might influence water level differences.


Effect magnitude (log scale, which kills a number of zero differences):

```{r inspect-data-1}
#| label: fig-effect-dw
#| fig-cap: "Top-down, the water level difference follows and exponential distribution: large differences are increasingly unlikely."

diff_data %>% 
  ggplot(aes(x = log10(1+abs(dw)))) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  xlim(0, 2) +
  theme_bw()

```


Correlation among covariates:

```{r inspect-data-2}
#| label: fig-correlation-matrix
#| fig-cap: "Matrix visualization of correlations of parameters. The redder, the more correlated. Only the lower half of the matrix is shown."
p.mat <- pp_data %>%
  select(
       elevation_dhmv, slope_r2.5m,
       wata_min_dist,
       emd_n, emd_amp1, emd_amp2, emd_ampsum,
       emd_qdiff, emd_residual_qdiff
  ) %>% 
  cor_pmat()
ggcorrplot(p.mat, hc.order = TRUE, type = "lower",
  outline.col = "white",
  ggtheme = ggplot2::theme_bw,
  colors = c("#6D9EC1", "white", "#E46726"))
```


More detail about the pronounced `slope`/`amp` correlation:

```{r contour-two}
#| label: fig-contour-slopeamp
#| fig-cap: "Some parameters in the data set are correlated (here: slope and 2nd emd amplitude). However, the observations seem rather diffuse, even when separating soilclasses (rows) and water distances (columns)."
pp_data %>% 
  ggplot(aes(x = slope_r2.5m, y = emd_amp2)) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon") +
  scale_fill_continuous(low="lavenderblush", high="red") +
  geom_jitter(alpha=0.05, size = 0.5) +
  labs(subtitle = "") +
  facet_grid(soilclass ~ wata_dist_class) + 
  xlim(c(0, 0.16)) +
  ylim(c(0, 0.16)) +
  theme_bw()
```


Some of the more obvious co-variates (e.g. soilclass) were used in the analysis above.
The others might be interesting for other purposes, another day.


## references

::: {#refs}
:::
