---
title: "Water Level Variograms"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---


# Preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges
# library("gstat") # geospatial statistics
# library("sdmTMB") # species distribution models Template Model Builder


# source("./watina_migration.R")
# source("./auxiliary_data_queries.R")

# option to work offline; requires previous data download
work_offline <- TRUE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
# dir.create("./data", showWarnings = FALSE)

if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 2 # minimal cluster member count; fewer memeber clusters are removed
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")
diffdata_export_file <- paste0("data/waterlevel_diffdata_cl", cluster_dist, ".parquet")

# the user may choose to reload raw data
force_data_reassembly <- FALSE
reload <- force_data_reassembly || !file.exists(data_export_file)
```




```{r load-data}

pp_data <- read_parquet(data_export_file)

diff_data <- read_parquet(diffdata_export_file)

pp_join <- pp_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope = mean(c(slope_r2.5m2, slope_r2.5m1)) 
  )

```

# data inspection

We can learn a lot about covariates which might influence water level differences.


Effect magnitude (log scale):

```{r inspect-data-1}

diff_data %>% 
  ggplot(aes(x = log10(1+abs(dw)))) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  xlim(0, 2) +
  theme_bw()

```


Correlation among covariates:

```{r inspect-data-2}
library("ggcorrplot")
p.mat <- pp_data %>%
    select(
         elevation_dhmv, slope_r2.5m,
         wata_min_dist,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
           ) %>% 
    cor_pmat()
ggcorrplot(p.mat, hc.order = TRUE, type = "lower",
   outline.col = "white",
   ggtheme = ggplot2::theme_bw,
   colors = c("#6D9EC1", "white", "#E46726"))
```


More detail about the `slope`/`amp` correlation:

```{r contour-two}
pp_data %>% 
  ggplot(aes(x = slope_r2.5m, y = emd_amp2)) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon") +
  scale_fill_continuous(low="lavenderblush", high="red") +
  geom_jitter(alpha=0.05, size = 0.5) +
  labs(subtitle = "") +
  facet_grid(soilclass ~ wata_dist_class) + 
  xlim(c(0, 0.16)) +
  ylim(c(0, 0.16)) +
  theme_bw()
```


# binning

binning and joining the data:

```{r glimpse-data}

# diff_meta %>% glimpse
# diff_data %>% glimpse

extent <- 512 # m
n_bins <- 64

diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    ),
    distance_bin = cut(ds,
      breaks = seq(
        -1,
        extent,
        length.out = extent+2
      )
    ), # cut to "meters" bin
    distance_bin8 = cut(ds,
      breaks = seq(
        -extent/n_bins*2,
        extent,
        length.out = n_bins/2+2
      ) # cut to 8m bins
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(dw <= ds) # remove outliers

# glimpse(diff_all)

diff_grouped8 <- diff_all %>%
  group_by(pair, distance_bin8) %>%
  summarize(
    cnt = n(),
    #cnt = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*cnt)*sum(dw^2, na.rm = TRUE),
    dw = 1/(cnt)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
  ) %>% ungroup %>% arrange(ds)

diff_grouped8 %>%
  ggplot(aes(x = ds, y = log(1.0+dw),
         color = soilclass)) +
  geom_point(aes(cex = sqrt(cnt)), alpha = 0.16) +
  theme_bw()
```

Sample size (circle extent) is balanced; none too big.


# ridge plots

... as an extra to regular difference-distance plots.


```{r diff-ridges-dw}
diff_grouped8 %>%
  ggplot(aes(x = dw, y = distance_bin8)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 3,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi) +
  theme_bw()
```


```{r diff-ridges}
diff_grouped8 %>%
  ggplot(aes(x = semivar, y = distance_bin8)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 2,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi/2) +
  theme_bw()
```


# grouping

Re-applying the groups with more fine-grained binning

```{r bin-data}

diff_grouped <- diff_all %>%
  group_by(pair, distance_bin) %>%
  summarize(
    cnt = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*cnt)*sum(dw^2, na.rm = TRUE),
    dw = 1/(cnt)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
  ) %>% ungroup %>% arrange(ds)


combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)

# a convenience function for calculated measures
summarize_groupeddata <- function(data) {
  data %>% 
    summarize(
      ds = mean(ds), 
      dw_all = combine_grouped(dw, cnt),
      dw_groups = mean(dw, na.rm = TRUE), 
      semivar_all = combine_grouped(semivar, cnt),
      semivar_groups = mean(semivar, na.rm = TRUE),
      cnt_all = sum(cnt),
      cnt_groups = mean(cnt),
    ) %>% ungroup %>% arrange(ds)
}


diff_binned <- diff_grouped %>%
  group_by(distance_bin, soilclass) %>%
  summarize_groupeddata()

knitr::kable(rbind(head(diff_binned, 5), tail(diff_binned, 5)))
```


```{r plot_dw}
diff_binned %>%
  # filter(semivar < 1) %>% 
  ggplot(aes(x = ds, y = log(1+dw_groups), color = soilclass)) +
  geom_point() 
```

```{r}
diff_binned %>%
  # filter(semivar < 1) %>% 
  ggplot(aes(x = ds, y = log(1+semivar_all), color = soilclass)) +
  geom_point() 
```


# regression


```{r regression-tools}

# wrap a regression function to generate residuals
# the result is the parameter to be minimized.
wrap_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  return(sqrt(mean(differences^2)))
}

distweighted_target_function <- function(x, y, regressor, params) {
  predictions <- regressor(x, params)
  differences <- y - predictions
  #differences <- log(1.0+y) - log(1.0+predictions)
  # differences <- 1000*differences * (1+1/x^2)
  # return(sqrt( sum(differences^2)/sum((1+1/x^2)) ))
  differences <- 1000*differences /sqrt(x)
  return(sqrt( mean(differences^2) ))
}

# this can turn regression output into a usable function.
create_prediction_function <- function(regressor, results) {
  fcn <- function (x) {
    regressor(x, results$par)
  }

  return(fcn)
}

# a uniform way to print results
print_regression_results <- function(orsl, label = "") {
  par <- paste(round(orsl$par, 4), collapse = ", ")
  conv <- orsl$convergence
  eps <- orsl$value
  print(
    sprintf("%s: conv %i at (%s), mse %.1f", label, conv, par, eps)
  )
}

# note: the `epsilon` can be calculated manually with the formula:
#   sum((predictor_function(x) - y)^2)


# Finally, a quick histogram plot of residuals.
plot_residuals_histogram <- function (x, y, predictor_function, ...) {
  residuals <- predictor_function(x) - y
  ggplot(NULL, aes(x = residuals)) +
    geom_histogram(...) +
    geom_vline(xintercept = 0, color = "darkgrey") +
    theme_bw()
}


```


```{r matern-definition}

matern_function <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}


```


```{r regression-matern-helpers}

# ## for exploration:
# start_values <- c("tau" = mean(y), "kappa" = sqrt(8)/250)
# start_values <- c("phi" = 0.3, "sill" = 10) # , "nu" = 1. 
# params <- start_values
# d <- x
# plot(x, matern_function(x, start_values), type = "l") 

fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results, y_label = "semivariance (m²)") {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


Finally, one mor not-so-general helper function:

```{r regression-matérn-general}
# trafo <- function(x) log(1.+x)
trafo <- function(x) x

regression_by_soilclass <- function(sc, reg_var, maxx = extent, maxy = NULL, ...) {
  diff_sc <- diff_binned %>% filter(soilclass == sc)
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))
  # firsthalf <- x <= extent/2
  # y <- y[firsthalf]
  # x <- x[firsthalf]

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]
  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
    # control = list("fnscale" = 1e-4),
  )
  
  print_regression_results(matern_fit, label = "grp")
  ylabel <- "mean water level difference (pair-averaged)"
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"
  g <- plot_matern(matern_fit, y_label = ylabel)
  return(g)
}

```


# variograms


Sample size:

```{r semivar-regression-samplesize}
diff_binned %>%
  ggplot(aes(x = ds, y = cnt_groups)) +
  geom_step(color = "darkblue") +
  geom_step(aes(x = ds, y = cnt_all/10), color = "lightblue", alpha = 0.6) +
  theme_bw()
  #ggplot(aes(x = ds, stat = "count")) +
  #geom_histogram(bins = 256, fill = "lightgray", color = "black")
```

Nothing to see here.
There are some *rather empty* bins; they probably account for the outliers below.
However, the fine structure seems beneficial for regression.



## water level difference

Better called: **mean absolute difference** (MAD) in water level.

$$ \langle dw\rangle = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$



```{r dw-column}
column <- "dw_groups"
# column <- "dw_all"
```


::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
regression_by_soilclass("heavy",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 6)

```


### light

```{r dw-regression-light}
regression_by_soilclass("light",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 6)
```


### peat

```{r dw-regression-peat}
regression_by_soilclass("peat",
  reg_var = column,
  maxx = 256, maxy = 2,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 6)
```

::: {.callout-warning}
Capped: $<= 2$
:::


:::



## semivariance


Differences compared to MAD:

- use squared values
- divide by two, i.e. *semi*-variance
- NOT the real variance! 
- Better called: **half mean square difference.**

$$ \gamma = \frac{1}{2N} \sum\limits_{i,j} \left(z_j - z_i\right)^2 $$



```{r semivar-column}
column <- "semivar_all"
# column <- "semivar_groups"
```


::: {.panel-tabset group="soilclass"}

### heavy

```{r semivar-regression-heavy}
regression_by_soilclass("heavy",
  reg_var = column,
  maxx = 256, maxy = 6, 
  distweighted = TRUE,
  par = c(8., 100, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(16., 250, 0.08, 1.01)
) + ylim(0, 6)

```

::: {.callout-warning}
Capped: $<= 6$
:::


### light

```{r semivar-regression-light}
regression_by_soilclass("light",
  reg_var = column,
  maxx = 256, maxy = 6,
  distweighted = TRUE,
  par = c(8., 100, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(16., 250, 0.08, 1.01)
) + ylim(0, 6)
```

::: {.callout-warning}
Capped: $<= 6$
:::


### peat

```{r semivar-regression-peat}
regression_by_soilclass("peat",
  reg_var = column,
  maxx = 256, maxy = 1,
  distweighted = TRUE,
  par = c(8., 100, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(16., 250, 0.08, 1.01)
) + ylim(0, 6)
```

::: {.callout-warning}
Capped: $<= \mathbf{1}$
:::

:::




# Summary

::: {.callout-tip}

- semivariance is calculated differently, i.e. not via R's `var` (see formulas)
- mean absolute difference might be the more practical choice
- above are the first convincing Matérn regressions

:::

::: {.callout-tip}

compared to `xG3` analysis:

- no de-trending
- *light* soil is less variable (range: $\approx 35 m$)
- *heavy* soil is less flat, but probably still longer range than light (range: $\approx 70 m$)
- *peat* seems to have a very long range ($\approx 100 m$)


:::


# Next

close-range variograms of further subsets:

- slope-class
- elevation (difference)?
- optional water distance filter


