---
title: "Water Level Variograms: Daily Data"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---

::: {.callout-warning icon=false}
This notebook provides an introduction to a more complex analysis of the spatial coupling of groundwater level measurements.

According to "all-inclusive" variogram models derived below, a $1 cm$ difference threshold in the water level of two WATINA locations would be exceeded at distances of $1.5 m$ ("heavy" soil), $0.9 m$ ("light" soil), and $13.0 m$ ("peat"), respectively.
Although these calculations serve as a first reference, the remnant value of this notebook is explorative and educational.
Many more (reasonable) filter settings which affect the data (e.g. slope) remain to be explored in the following notebooks, and measurement uncertainty will be assessed by bootstrapping.


The core purpose of *this* notebook is to establish the choices for binning and difference parameters, showcase the general technique, and introduce vocabulary.
Fine-grained comparisons are found in the other notebooks.

:::



# summary (tl;dr)

In this notebook, I analyze difference data of **daily water levels** from the `watina` database.
Take two observation wells within a given distance, which are placed in the same type of soil ("soilclass"), and which have water level data (units: `mTAW`) available on the same day.
We group all possible such pairs of wells by distance ("binning"), use pair-average in case of repeated measurement, and analyze the mean absolute difference in water level.
To determine a range at which the average water level difference exceeds a given threshold (i.e. a range at which we would expect observation wells to yield more independent data), we use regression with a Matérn function.


This procedure is in line with the general algorithm of variograms ([see here](https://tutorials.inbo.be/tutorials/spatial_variograms)), but deviates in some design choices to what is conventionally understood as "variogram".
The most important differences are the choice of difference measure (mean absolute difference instead of semivariance), and adjustments to binning (homogeneous bin fill, but variable bin width).
In detail:

- Daily water levels were used (watina `PeilMeting`)
- Use **mean absolute difference** ($\langle dw \rangle$), instead of (semi)variance ($\gamma$, [@sec-parameters]).
- Categories for pooling/averaging ([@sec-pooling]), in the following order:
  - observation pairs 
  - distance bins
  - soilclass
- Variable bin width to achieve homogeneous sample size per bin ([@sec-equalbins]).
- No de-trending whatsoever was applied.
- No extraction of LG3 and HG3. (We get back to "seasonal" high and low values in another notebook via quantiles.)
- No "random intercept", as suggested by Floris: this is not applicable here because we disregarded the "modeling" approach (even de-trending).


Summary of the outcome ([@sec-variograms]):

- *light* soil is most variable, i.e. higher differences are expected to occur at closer distance (shortest range to $1 cm$ difference threshold: $\approx 0.9 m$)
- confirmative evidence that *heavy* soil shows slightly wider range of spatial correlation ($\approx 1.5 m$ distance to find $1 cm$ difference)
- *peat* data fits a comparatively long range ($13.0 m$ to $1 cm$ threshold) 
- in the limit, the value range of $\langle dw\rangle$ (y-axis range) of *heavy* and *light* soil are similar (sill $\approx 0.5 m$; nugget of $\approx 0.1$); we retrieve seemingly similar values for peat; in part this is due to using `mTAW` as the reference coordinate units.



::: {.callout-note icon=false}
## Lessons learned (technical notes FM):

The latest findings on an exciting journey.

- semivariance is calculated differently than what I thought, i.e. not via R's `var` (see [@sec-parameters])
- $\langle dw \rangle$ is the more practical choice (semivariance has problems with pooling squared values; $\langle dw \rangle$ is easier to interpret, e.g. because of units)
- Sample size per bin must be monitored, "equal" work better and "sufficiently large" is preferrable.
- Sample size per bin should be generous, sufficient to stabilize the vgram (i.e. minimize variance between bins, while retaining a sufficient number of bins for unambiguous regression).
- Outliers, especially on short distance, have a major drag on the regression (but few of these shall remain when bins are sufficiently large). 
- Data loading must be serialized. Even if the computer manages to load the data once, no guarantee it will work again.

:::

The outcome ([@sec-variograms]) are variograms with Matérn regressions, which provide evidence about spatial coupling of water levels.
To re-iterate: this notebook sets the stage for further detailed analyses to come.


# preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("ggcorrplot")|> void() # correlation, grid visualization
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges


if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")

# for reference
cluster_dist <- 3400 # m # parameter for clustering, used on data loading


# for a uniform look
soilclass_colors <- c(
  "heavy" = "sienna",
  "light" = "burlywood",
  "peat" = "darkseagreen",
  "unknown" = "slategray"
)
```


The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here; helper functions are collected in `auxiliary_data_queries.R`.



```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  ) 

```


# binning (preview)


Binning and joining the data is necessary, yet there is some procedural flexibility in how to apply it.

First and foremost, data is filtered to include spatial distance of up to `extent` ($500 m$).
Because that does not suffice to reach the inversion point of some curves, regression is allowed to reach longer distances (`maxx`).
We also set the `bin_target`, i.e. the bin fill aspiration, to $100$ observations per bin.


```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

n_bins <- 32 # number of bins in default binning
bin_target <- 100 # alternative: observations per bin
cap_dw <- 8 # m # maximum accepted water level difference
```

Binning is flexible, and we compare the differences below, with `n_bins =` `{r} n_bins` giving the number of equally-spaced bins (at `{r} extent/n_bins` m); note however that the most relevant binning strategies might be two others:

1. Binning with fixed bin width of `1 m`.
2. Binning with variable bin width, to achieve **equal sample size per bin**.

More on that below.



```{r glimpse-data}

# diff_meta %>% glimpse
# diff_data %>% glimpse

diff_meta <- diff_meta %>%
  filter(ds <= extent) %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    ),
    distance_bin = cut(ds,
      breaks = seq(
        -1,
        extent,
        length.out = extent+2
      )
    ), # cut to "meters" bin
    distance_widebin = cut(ds,
      breaks = seq(
        -extent/n_bins*2,
        extent,
        length.out = n_bins/2+2
      ) # cut to 8m bins
    )
  ) 

diff_joined <- diff_meta %>% 
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# glimpse(diff_joined)

```

:::{.callout-warning}
All single observations with a difference in water level of *more than 8m* were filtered out.

Given a maximum filter depth of $8 m$, extent (i.e. maximum pair distance) of $500 m$, and the flatness of Belgian topography, values exceeding that magnitude are not plausible.
Even $4 m$ requires ambitious justification, which is why on other scripts we occasionally use that as a stricter difference filter threshold.
:::


```{r widegroup-data}
diff_widegroup <- diff_joined %>%
  summarize(
    cnt = n(),
    #cnt = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*cnt)*sum(dw^2, na.rm = TRUE),
    dw = 1/(cnt)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
    .by = c(pair, distance_widebin)
  ) %>% arrange(ds)
```


Briefly inspecting the wide bins, sample size (circle extent) is balanced; none outstandingly big.

```{r glimpse-data}
#| label: fig-bubbletea
#| fig-cap: "An irrelevant ''bubble tea plot''™. The size of the bubbles scales with the number of observations. Yes, this was only a humble first glimpse at the data; yes, I used `sqrt` as a 'cheat' to limit bubble size differences; yes, we will do better below. But the colors are nice."
diff_widegroup %>%
  ggplot(aes(x = ds, y = log(1.0+dw),
         color = soilclass)) +
  geom_point(aes(cex = sqrt(cnt)), alpha = 0.16) +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```



# ridgeline plots

... with the little "extra" to regular difference-distance plots.


```{r diff-ridges-dw}
#| label: fig-ridges
#| fig-cap: "Visualization of distance-dependent difference with ridgelines. With increasing distance (x-axis), the distribution of water level differences (y-axis, grey density distributions) increases (indicated by quantile lines)."
diff_widegroup %>%
  ggplot(aes(x = dw, y = distance_widebin)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 3,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi) +
  theme_bw() + coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```


```{r diff-ridges}
#| eval: false
#| include: false
diff_widegroup %>%
  ggplot(aes(x = semivar, y = distance_widebin)) +
  geom_density_ridges(rel_min_height = 0.05,
    quantile_lines = TRUE, quantiles = 2,
    scale = 4.0, bandwidth = 0.05) +
  # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
  labs(x = "mean water level difference", y = "pair distance") +
  xlim(0, pi/2) +
  theme_bw()
```

This is a promising pattern: differences are lowest on short distances, and increase with distance. 
We better expect to re-capture it in the analysis below.


# grouping and binning

## pooled data: what to pool? {#sec-pooling}

Re-applying the groups with more fine-grained binning.

```{r bin-data}

diff_grouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    # dw = mean(abs(dw), na.rm = TRUE)
    .by = c(pair, distance_bin)
  ) %>% arrange(ds)


combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)

diff_binned <- diff_grouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    dw_all = combine_grouped(dw, n_obs),
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_all = combine_grouped(semivar, n_obs),
    semivar_groups = mean(semivar, na.rm = TRUE),
    .by = c(distance_bin, soilclass)
  ) %>% arrange(ds)


knitr::kable(rbind(head(diff_binned, 5), tail(diff_binned, 5)))
```


::: {.callout-note icon=false}
## Pooling

There are multiple variables to consider for pooling.

- pairs of locations (pseudo-replication)
- the distance bins
- soilclass

Even the grouping order matters, it better be in the listed order.
:::


## choice of difference parameter {#sec-parameters}

There are at least two possible distance parameters we can apply to this grouped data.
This has implications for grouping, with some practical trouble for semivariance (see [@sec-semivar-binning]).


First, we have **conventional semivariance** ["Method-of-Moments Estimator" according to @Cressie1993, p. 69, eqn. 2.4.2].
This originally goes back to @Matheron1962, and only holds under the stationarity assumption ("constant mean").

$$ \gamma = \frac{1}{2N} \sum\limits_{i,j} \left(z_j - z_i\right)^2 $$

```{r plot-semivariance-distance}
#| label: fig-semivariance-difference-plot
#| fig-cap: "Semivariance-distance plot: the increase with distance is not too pronounced on the semivariance measure, despite the log scale for better visualization."
diff_binned %>%
  ggplot(aes(x = ds, y = log(1+semivar_groups), color = soilclass)) +
  geom_point() +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```

This equation has the typical structure of "sum of square differences", the normalization with sample size $N$ effectively makes it an average. 
Yet the "square" is not inverted by a "square root", which has some pitfalls ([@sec-semivar-binning]).


Therefore, I suggest a simpler variant: **mean absolute difference** (MAD) in water level.

$$ \langle dw\rangle = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$


```{r plot-dw-distance}
#| label: fig-difference-distance-plot
#| fig-cap: "Difference-distance plot: rough top-down visualization to inspect if water level differences increases with distance. Apparently, it does. Log scale for better visualization."
diff_binned %>%
  ggplot(aes(x = ds, y = log(1+dw_groups), color = soilclass)) +
  geom_point() +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```


::: {.callout-warning}
Neither of these is the actual variance, as in `?var` in R, although with his limited maths talent the author cannot exclude that variance and "semivariance" are related in the stationary case.
:::


Let's go MAD!



## equally-sized bins {#sec-equalbins}

Re-visit the binning.

Sample size is rather rugged:

```{r semivar-regression-samplesize}
#| label: fig-sample-size
#| fig-cap: "Sample size per bin in standard grouping. Dark blue: per location pair, i.e. repeated observations from the same location pairs are averaged. Light blue: per observation (divided by 100 to fit the axis)."

diff_binned %>%
  ggplot(aes(x = ds, y = cnt)) +
  geom_step(color = "darkblue") +
  geom_step(aes(x = ds, y = n_obs/100), color = "lightblue", alpha = 0.6) +
  theme_bw()
```


There are some *rather empty* bins touching the axis; they probably account at least in part for potential outliers.
"Misery loves company:" some other bins are *spikes*, probably due to pseudo-replication (repeated sampling at high frequency).
This ruggedness is a problem, considering a bias potentially introduced by location clusters which were installed for an unknown research purpose and do not fit our "randomness" assumption (cf. notebooks `113a_cluster_distance_bias` and `113b_cluster_loo`).


How about using equal-sized bins, i.e. buckets of variable width which are all filled (by design) with approximately equal number of observations?

Crucially, we'll have to group by location pair first because otherwise repeated measures fill single buckets.

:::{.callout-warning}
After feedback by Thierry, I try to stabilize the variograms by widening bins,
aspiring to get approximately `{r} bin_target` observations per bin.
Even wider bins caused trouble by undersampling of the proximal areas.
The number of $100$ still seems to be a good allocation for most combinations of settings.
However, on some filter analyses in other notebooks the required bin fill (`bin_target`) can be reduced.
:::


```{r equibin-preparation}
#| code-fold: false

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup

    
diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    dw_all = combine_grouped(dw, n_obs),
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_all = combine_grouped(semivar, n_obs),
    semivar_groups = mean(semivar, na.rm = TRUE),
    .by = c(distance_equibin, soilclass)
  ) %>%
  arrange(ds)

```
    
    
    

```{r equibin-export}
#| eval: true
#| echo: false
#| code-fold: true

# plot export for presentation

require("ggdark")

h <- diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "grey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "grey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  dark_theme_minimal()

suppressWarnings(
  ggsave(paste0("figures/binsize_cumulated_dark.svg"), h, dpi = 300)
)
```


```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```

Seems more balanced, let's see what to do with these bins.


# regression {#sec-regression}

Here comes a blunt appearance of a general toolkit for regression, based on `optim` ([docs](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/optim)), which is thrown into the R script file `regression_tools.R` for convenient DRY re-use in this and the following notebooks.
It is derived in some more detail [here](https://tutorials.inbo.be/tutorials/spatial_variograms).

Note that root-mean square difference (RMSD) is optionally distance-weighted, by dividing each observed `prediction-actual` difference by square root of the distance.
This effectively gives extra relevance to the proximal bins (which would otherwise be outnumbered).


```{r regression-tools}

source("regression_tools.R") 

```


The Matérn parametrization below is related to the parameters on (semi-)variogram plots.
Some vocabulary:

```
      ^
      | Range
      |<--->
Sill  |     :____,model
      |    /:
 VAR  |   / :
 IOG  |  /  :
 RAM  | /   :
      |/    :
Nugget| Cor.: Uncorrelated
     _|_____:___________>
      |              lag distance
```

Most notably, $\gamma$ on the y-axis is called "semivariance", lending its name to this specific type of visualization.

We parametrize a bit differently, by using the difference in semivariance, $d\gamma$:
$$ d\gamma = \gamma_{max} - \gamma_{min}$$
$$ \gamma = \gamma_{max} - d\gamma \cdot \left( \frac{\sqrt{2\nu}}{\sigma} x \right)^{\nu} \cdot K_{\nu}\left( \frac{\sqrt{2\nu}}{\sigma} x \right) \cdot \frac{1}{2^{\nu - 1} \cdot \Gamma\left(\nu\right)} $$

With sill $\gamma_{max}$, nugget $\gamma_{min}$, turning point $\sigma$ (related to range), and a shape parameter $\nu$ that controls the form of the sigmoid[^1]. 
(Compare [wikipedia](https://en.wikipedia.org/wiki/Mat%C3%A9rn_covariance_function).)


[^1]: In practical application, I soft-fix $\nu$ to a value of one ($\nu = 1.0$), and only give it moderate freedom in case of visual shape mismatches. 


In code:

```{r matern-definition}
#| code-fold: false

matern_function <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}

```

This particular parametrization of nugget, scale, and sill will facilitate optimization by ensuring that the nugget is always smaller than the sill.


Because we repeatedly apply the procedure, a couple of convenience functions below.

```{r regression-matern-helpers}

fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results,
                        y_label = "mean absolute difference (m)",
                        color = "darkgrey") {

  # retrieve everything
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]

  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "grey") +
    geom_hline(yintercept = nugget, color = "grey") +
    geom_hline(yintercept = sill, color = "grey") +
    geom_point(size = 2.5, color = color, fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx)),
              color = color, linewidth = 1.2) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


(Finally, one more not-so-general helper function.)

```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x

regression_by_soilclass <- function(
    sc, reg_var, maxx = maxx, maxy = NULL, use_equibin = TRUE,
    ...) {
  if (use_equibin) {
    diff_sc <- diff_equibinned %>% filter(soilclass == sc)
  } else {
    diff_sc <- diff_binned %>% filter(soilclass == sc)
  }
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))
  # firsthalf <- x <= extent/2
  # y <- y[firsthalf]
  # x <- x[firsthalf]

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]
  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
    # control = list("fnscale" = 1e-4),
  )
  
  print_regression_results(matern_fit, label = "regression:")
  ylabel <- "mean water level difference (pair-averaged)"
  sc_color <- soilclass_colors[sc]
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"

  g <- plot_matern(matern_fit, y_label = ylabel, color = sc_color)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, color = "red", alpha = 0.2) 
  return(g)
}



```

These functions can also be found in the accompanying `regression_tools.R` script.


# variograms {#sec-variograms}

## water level difference (MAD)

Feel free to adjust regression parameters; maybe most importantly, the outcome measure ("target column").

```{r dw-column}
#| code-fold: false

target_column <- "dw_groups"
# target_column <- "dw_all"
# # see also the "semivariance" options, below.
```

There are also axis upper limits, start values, parameter limits, RMSD weighting, and binning to choose.
Selected here are the most plausible settings (MAD, equal-sized bins, distance weighted, with sensible parameter starts and limits).


Note that a crucial outcome determinant reported below is the **distance to $0.01 m$ threshold**.
We determin this by using the raw shape of the fit Matérn, deducing the nugget, and evaluating at what distance two hypothetical locations would be expected to differ by more than a centimeter in water level.
This distance is always lower than (but related to) the $\sigma$ range used for Matérn parametrization.


**Behold: The Variograms.**


::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
#| label: fig-dw-regression-heavy
#| fig-cap: "Matérn regression of water level difference $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass("heavy",
  reg_var = target_column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r dw-regression-light}
#| label: fig-dw-regression-light
#| fig-cap: "Matérn regression of water level difference $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass("light",
  reg_var = target_column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 100, 0.48, 1.01)
) + ylim(0, 1.6)
```


### peat

```{r dw-regression-peat}
#| label: fig-dw-regression-peat
#| fig-cap: "Matérn regression of water level difference $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass("peat",
  reg_var = target_column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  use_equibin = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```

:::


As mentioned, these are just an initial shot at the data. 
The remaining design choices are evaluated at length in the following notebooks.
Impatient readers are welcome to skip to the notebook `118_summary`.


<hr>

# appendix {.appendix}

## Questions

> Kunnen we de maximale afwijking van 1 cm in de HG3 en LG3 vertalen naar een maximale afwijking in dagelijkse waarden?

The HG3 and LG3 are non-linearly derived values that only indirectly measure highest or lowest water levels. 
Though they are deterministically calculated, I think their calculation method introduces extra jitter.
I do not see any additional value in using them.
To answer specific questions about high- and low water regimes, a more sophisticated analysis (e.g. using Empirical Mode Decomposition, EMD) might be in reach.


> Het is me niet duidelijk hoe de *observation pairs* zijn gedefinieerd: enkel pairs binnen een cluster (waardoor je niet moet detrenden noch met een random factor werken) op dezelfde dag gemeten? Zijn er voorwaarden obv filtertopdiepte?

Observation pairs must lie in the same cluster, because cluster distance (i.e. distance between the closest points of two clusters, chosen `{r} cluster_dist` m) is far longer than the maximum pair distance we allow in the data set (`{r} extent` m). 
We do not de-trend:
the chosen difference measure (MAD) is the absolute difference value, but because direction is not biased, de-trending would tend to reduce variability (increase spatial coupling).


> Waar staat *scale* in de Matérn-functie? Dat wordt gebruikt voor de berekening van de sill, maar ik zie niet waar ze juist in de vergelijking staat of er wordt van afgeleid.

The scale is simply the difference between *sill* and *nugget* (see ascii plot in [@sec-regression]).
The computation equation in the `matern_function` builds up step-wise.
Parametrization is chosen to avoid non-sensical optimizer outcomes (e.g. "nugget > sill" situations).


> Kan je ook de analyse doen voor de andere klassen die worden gecreëerd, i.e. water distance > 50 m en slope > 0.05? Je concludeert nl. dat de homogenisatie zorgt voor een toename in de range, maar geldt dit ook voor het andere deel van de dataset?

See the adjacent notebook files.


> Ook verwacht ik niet per sé homogenisatie wanneer een kleinere *treshold* voor afstand tot water wordt gekozen: ik zou denken dat we grotere verschillen in grondwaterstand kunnen verwachten op korte afstand wanneer je je dichter bij een waterloop bevindt. Kan een model lopen bij een *treshold* van 25 m?

I ~~will~~ did run that model and check. 
But there is other effects, such as bin size.
I do not understand the rationale for expecting that there should be larger differences at lower inter-well distances if we are closer to surface water bodies.




## why not semivariance? {#sec-semivar-binning}

Recall the two equations above:

$$ \langle dw\rangle = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$


$$ \gamma = \frac{1}{2N} \sum\limits_{i,j} \left(z_j - z_i\right)^2 $$


From these equations, semivariance ($\gamma$) involves subtle differences compared to the more intuitive MAD ($\langle dw\rangle$):

- $\gamma$ uses squared values
- $\gamma$ is divided by two, i.e. it calculates **semi**-variance (*triv.*)
- This is NOT the real variance! 

Therefore, semivariance should better be called: **half mean square difference** (HMSD).


Semivariance / HMSD be tried out with the code above by using either of the following:

```{r semivar-column}
#| code-fold: false

if (FALSE) {
  target_column <- "semivar_all"
  target_column <- "semivar_groups"
}
```

::: {.callout-warning}
The semivariance (i.e. HMSD) was not used here!
Instead, I went MAD.
:::

This has two justifications:

- observation of the outcome
- consideration about pooling data


Due to the square, pooling is non-trivial, which is best illustrated by a simple example. 

Assume we measure the following water levels, in meters.

```
day 1: z1 = 0.1
day 1: z2 = 0.5

day 2: z1 = 0.1
day 2: z2 = 0.3

day 3: z1 = 0.1
day 3: z3 = 0.4

```

Differences are:

```
d1 = 0.5-0.1 = 0.4
d2 = 0.2
d3 = 0.3

```

Plain average would be:
```
<d> = (0.4 + 0.2 + 0.3) / 3 = 0.3
```


The first two observations are pseudo-replications, because they come from the same sites (`z1` and `z2`).
We better average them:

```
<d> = ((0.4 + 0.2)/2 + 0.3) / 2 = 0.3
```

Of course, this is no co-incidence: I chose the values to yield identical results in grouped averaging.


Compare this to the *squared* average difference used in semivariance:

```
γ = (0.4^2 + 0.2^2 + 0.3^2) / (2*3) = 0.0483
γ = (((0.4 + 0.2)/2)^2 + 0.3^2) / (2*3) = 0.105
```

... but then we square the number of observations (positive integer, large numbers of observations become huge)... 
Or maybe average the square differences?

```
γ = ((0.4^2 + 0.2^2)/2 + 0.3^2) / (2*3) = 0.03167
```

... but then we square our generally small differences (measured in meters), making them even smaller.
Mr. Πυθαγόρας could have anticipated this.


::: {.callout-note}
This illustrates that, because water level differences are generally small, yet repeated sampling gives a lot of these small differences per location pair,
the square structure of the semivariance formula can complicate grouping of repeated observations.


The issue of squared values is consistent with the huge soilclass difference observed previously in *heavy versus light* soil.
:::



This is in part an *ex post* rationalization of the scattered values I see when attempting semivariance.
Primary reason was inspection of the outcome, and non-plausible Matérn regressions when using semivariance.
Another benefit is **unit retention**: differences in meters are easier to interpret and compare than the square meters produced by variance formulas.


Using $\langle dw\rangle$ just works better.


## comparing co-variates

Just to show off, we have a number of useful co-variates at hand which might be insightful by themselves.


We can learn a lot about covariates which might influence water level differences.


Effect magnitude (log scale, which kills a number of zero differences):

```{r inspect-data-1}
#| label: fig-effect-dw
#| fig-cap: "Top-down, the water level difference follows and exponential distribution: large differences are increasingly unlikely."

diff_data %>% 
  ggplot(aes(x = log10(1+abs(dw)))) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  xlim(0, 2) +
  theme_bw()

```


Correlation among covariates:

```{r inspect-data-2}
#| label: fig-correlation-matrix
#| fig-cap: "Matrix visualization of correlations of parameters. The redder, the more correlated. Only the lower half of the matrix is shown."
p.mat <- meta_data %>%
  select(
       elevation_dhmv, slope_r2.5m,
       fdepth, ref_taw, 
       wata_min_dist,
       emd_n, emd_amp1, emd_amp2, emd_ampsum,
       emd_qdiff, emd_residual_qdiff
  ) %>% 
  cor_pmat()
ggcorrplot(p.mat, hc.order = TRUE, type = "lower",
  outline.col = "white",
  ggtheme = ggplot2::theme_bw,
  colors = c("#6D9EC1", "white", "#E46726"))
```


More detail about the pronounced `slope`/`amp` correlation:

```{r contour-two}
#| label: fig-contour-slopeamp
#| fig-cap: "Some parameters in the data set are correlated (here: slope and 2nd emd amplitude). However, the observations seem rather diffuse, even when separating soilclasses (rows) and water distances (columns)."
meta_data %>% 
  ggplot(aes(x = slope_r2.5m, y = emd_amp1)) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon") +
  scale_fill_continuous(low="lavenderblush", high="red") +
  geom_jitter(alpha=0.05, size = 0.5) +
  labs(subtitle = "") +
  facet_grid(soilclass ~ wata_dist_class) + 
  xlim(c(0, 0.16)) +
  ylim(c(0, 0.16)) +
  theme_bw()
```


Some of the more obvious co-variates (e.g. soilclass) were used in the analysis above.
The others might be interesting for other purposes, another day.


## tracing outliers

If the following plots are empty, we effectively removed all outliers above.

```{r trace-outliers}
outlier_trace <- diff_equibinned %>%
  filter(
    # soilclass == "peat",
    dw_groups > 2
  ) %>%
  pull(distance_equibin)
outlier_trace


outlier_bins <- diff_equigrouped %>%
  filter(distance_equibin %in% outlier_trace)

outlier_bins %>%
  ggplot(aes(x = ds, y = dw)) +
  geom_point()


outlier_pairs <- outlier_bins %>%
  filter(dw > 5) %>%
  pull(pair)


diff_outliers <- diff_meta %>% 
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) 
  # filter(dw <= ds, dw <= 24) # remove outliers

outlier_observations <- diff_outliers %>%
  filter(pair %in% outlier_pairs)

outlier_observations %>%
  ggplot(aes(x = abs(dw))) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()

```


:::{.callout-warning}
Conclusion:

Removed observations with `dw > ` `{r} cap_dw`.
Even **if** there were good ones in there, 
they are a minority.
:::


## references

::: {#refs}
:::
