---
title: "Daily Water Level Variograms: Filtered Analysis"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

In this notebook, I filter the data for relevant subgroups in the following parameters:

- `wata_min_dist`: Water distance is measured on various water-related zenodo sources, and is known to suffer some inaccuracy (e.g. from water body dynamics).
- `mslope` (`dh/ds`): Slope is defined as the difference in elevation within a pair of wells, normalized by their distance.
- `dz` difference in filter depth: Difference in filter depth is what it says: the difference in filter depth.
- `pp_vld & mp_vld` validation status of the site/obswell


Then, the variogram procedure is repeated.


:::{.callout-warning}
**The effect of `slope` is conceptually relevant, so that a slope filter will be regarded as the default setting for other analysis components.**
:::


__Overview of distance to 0.01m difference threshold:__


| soilclass     | all in |      slope | water distance | water distance | filter depth | validation | slope and fdepth |
|:--------------|-------:|-----------:|---------------:|---------------:|-------------:|------------|------------------|
| `(threshold)` |   `NA` | ` < 0.01 ` |      ` < 50 m` |      ` < 25 m` |     ` < 1 m` | `TRUE`     | `<0.01 & <1m`    |
| (N sites)     |  34868 |      31389 |          14543 |      [☇]  6487 |        29574 | 18236      | 26724            |
| (N/1000)      |   6730 |       6017 |           2630 |      [☇]  1088 |         5906 | 4764       | 5287             |
| heavy         |  1.5 m |     10.0 m |          5.9 m |        (8.7 m) |        1.5 m | 1.7        | 10.8             |
| light         |  0.9 m |     12.7 m |      [☇] 2.3 m |        (5.9 m) |        1.1 m | 0.8        | 16.9             |
| peat          | 13.0 m |     16.7 m |         16.4 m |       (19.4 m) |       15.0 m | 15.5       | 16.7             |



# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("units")     |> void() # someone decided to soter water distance in meters...


# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
```


## parameters

```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

n_bins <- as.integer(extent/4) # the number of bins...
bin_target <- 100 # ... and the preferred alternative: observations per bin

cap_dw <- 8 # m # maximum accepted water level difference
```


## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff,
         pp_vld, mp_vld
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      validated1 = as.integer(pp_vld1 == "VLD") * as.integer(mp_vld1 == "VLD"),
      validated2 = as.integer(pp_vld2 == "VLD") * as.integer(mp_vld2 == "VLD"),
      validated = as.logical(validated1 * validated2),
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  ) %>% 
  filter(ds <= extent)

```



```{r count-all}
diff_joined <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

count_all <- nrow(diff_joined)
print(count_all)
```



## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 

```

*(... has been outsourced)*


# Slope 
## Filters


- ~~`slope_r2.5m` < 0.05~~
- `mslope` < 0.01 m/m

:::{.callout-warning}
We now interpret slope pairwise: 
difference in elevation, normalized by distance.
:::


```{r slope-distribution}
# slope_threshold <- Inf #0.05
slope_threshold <- 0.01/1 # 1 cm per meter

diff_meta %>%
  select(mslope) %>%
  mutate(filltered = mslope < slope_threshold) %>% 
  ggplot(aes(log10(mslope), fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r slope-filter}
somediff_meta <- diff_meta %>%
  filter(
    # slope_r2.5m1 < slope_threshold,
    # slope_r2.5m2 < slope_threshold
    mslope < slope_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r join-slope}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r groupwise-binning-slope}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r slope-equibin-samplesize}
#| label: fig-slope-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```


## variograms {#sec-slope}

using water level difference (MAD)


```{r dw-column-slope}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-slope-regression-heavy}
#| label: fig-slope-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = extent, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r dw-slope-regression-light}
#| label: fig-slope-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-slope-regression-peat}
#| label: fig-slope-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::




# Water Distance I: 50m
## Filters

Note: when filtering for location pairs "close" to water, only 298/18467 pairs would remain.
-> using "close" and "mid", i.e. distance <= 50m


```{r water-distribution}
water_threshold <- units::set_units(50, "m")

diff_meta %>%
  select(wata_min_dist1) %>%
  mutate(filltered = wata_min_dist1 < water_threshold) %>% 
  ggplot(aes(wata_min_dist1, fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r water-filter}
somediff_meta <- diff_meta %>%
  filter(
    wata_min_dist1 < water_threshold,
    wata_min_dist2 < water_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)

```


## Binning

Binning parameters; will use homogeneously filled bins:


```{r join-water}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r groupwise-binning}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
      # ds, as.integer(min(c(floor(n()/8), n_bins)))
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```




## variograms {#sec-waterdistance}

using water level difference (MAD)


```{r dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
#| label: fig-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```



### light

```{r dw-regression-light}
#| label: fig-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-regression-peat}
#| label: fig-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = 2,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::


# Water Distance II: 25m {#sec-wadi-threshold-25m}
## Filters

Again, with a narrower filter on water distance.


```{r water-distribution-25m}
water_threshold <- units::set_units(25, "m")

diff_meta %>%
  select(wata_min_dist1) %>%
  mutate(filltered = wata_min_dist1 < water_threshold) %>% 
  ggplot(aes(wata_min_dist1, fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r water-filter-25m}
somediff_meta <- diff_meta %>%
  filter(
    wata_min_dist1 < water_threshold,
    wata_min_dist2 < water_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning

(Binning parameters as above.)


```{r join-water-25m}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r groupwise-binning-25m}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r equibin-samplesize-25m}
#| label: fig-equibin-sample-size-25m
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```




## variograms {#sec-waterdistance-25m}

using water level difference (MAD)


```{r dw-column-25m}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy-25m}
#| label: fig-dw-regression-heavy-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```



### light

```{r dw-regression-light-25m}
#| label: fig-dw-regression-light-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-regression-peat-25m}
#| label: fig-dw-regression-peat-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::



# Filter Depth 
## Filters


- `dz < 1 m`


```{r dfdepth-distribution}
dfdepth_threshold <- 1 # Inf # m

diff_meta %>%
  select(dz) %>%
  mutate(filltered = abs(dz) < dfdepth_threshold) %>% 
  ggplot(aes(abs(dz), fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r dfdepth-filter}
somediff_meta <- diff_meta %>%
  filter(
    dz < dfdepth_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r join-dfdepth}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r groupwise-binning-dfdepth}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r dfdepth-equibin-samplesize}
#| label: fig-dfdepth-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```


## variograms {#sec-dfdepth}

using water level difference (MAD)


```{r dw-column-dfdepth}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-dfdepth-regression-heavy}
#| label: fig-dfdepth-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r dw-dfdepth-regression-light}
#| label: fig-dfdepth-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-dfdepth-regression-peat}
#| label: fig-dfdepth-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::



# Validation
## Filters

Validation, validation, validation.

In an effort to build on as broad a foundation as possible, I experimentally included unvalidated data.
"Unvalidated" does not mean "wrong" or "excluded", after all.

- `validated` == TRUE


```{r vali-distribution}
diff_meta %>%
  select(soilclass1, validated) %>%
  ggplot(aes(soilclass1, fill = validated, color = validated)) +
  geom_bar() +
  theme_bw()
    
```


```{r vali-filter}
somediff_meta <- diff_meta %>%
  filter(
    validated
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r join-vali}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r groupwise-binning-vali}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r vali-equibin-samplesize}
#| label: fig-vali-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```


## variograms {#sec-vali}

using water level difference (MAD)


```{r dw-column-vali}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-vali-regression-heavy}
#| label: fig-vali-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r dw-vali-regression-light}
#| label: fig-vali-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-vali-regression-peat}
#| label: fig-vali-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::

# Slope AND Filterdepth
## Filters


- ~~`slope_r2.5m` < 0.05~~
- `mslope` < 0.01 m/m

:::{.callout-warning}
We now interpret slope pairwise: 
difference in elevation, normalized by distance.
:::


```{r slope-and-fdepth-distribution}
# slope_threshold <- Inf #0.05
slope_threshold <- 0.01/1 # 1 cm per meter
dfdepth_threshold <- 1 # Inf # m

diff_meta %>%
  select(mslope, dz) %>%
  mutate(filltered = (mslope < slope_threshold)&(dz < dfdepth_threshold)) %>% 
  ggplot(aes(log10(mslope), fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r slope-and-fdepth-filter}
somediff_meta <- diff_meta %>%
  filter(
    # slope_r2.5m1 < slope_threshold,
    # slope_r2.5m2 < slope_threshold
    mslope < slope_threshold,
    dz < dfdepth_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r slope-and-fdepth-join}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r slope-and-fdepth-groupwise-binning}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r slope-and-fdepth-equibin-samplesize}
#| label: fig-slope-and-fdepth-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```


## variograms {#sec-slope-and-fdepth}

using water level difference (MAD)


```{r slope-and-fdepth-dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-slope-and-fdepth-regression-heavy}
#| label: fig-slope-and-fdepth-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = extent, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r dw-slope-and-fdepth-regression-light}
#| label: fig-slope-and-fdepth-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r dw-slope-and-fdepth-regression-peat}
#| label: fig-slope-and-fdepth-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::



# Archive
## Responses on earlier questions

> kan je ook de analyse doen voor de andere klassen die worden gecreëerd, i.e. water distance > 50 m en slope > 0.05? Je concludeert nl. dat de homogenisatie zorgt voor een toename in de range, maar geldt dit ook

For convenience, this is available in another notebook: please find `111b_waterlevel_complements.qmd`. 
Unfortunately, sample sizes of the complementary subsets are too small for definitive conclusions.



> Ook verwacht ik niet per sé homogenisatie wanneer een kleinere *treshold* voor afstand tot water wordt gekozen: ik zou denken dat we grotere verschillen in grondwaterstand kunnen verwachten op korte afstand wanneer je je dichter bij een waterloop bevindt. Kan een model lopen bij een *treshold* van 25 m?

*"Yes, we can!"* ([@sec-wadi-threshold-25m]).


--> still included above.


Note that the homogenization paradigma is just a general tendency: the low overall variability in "peat" and the detail difference in remaining datapoint provides a good counter-example. Regression landscape is generally flat, i.e. MSE differences are low.
Generally, my considerations on this are rather technical, to begin with: I can judge well the mathematical component of a change in threshold, and the effects on regression.
I am glad for expert input on water level influences.


--> to be reviewed; much change due to outlier exclusion.
