---
title: "Daily Water Level Variograms: Filtered Analysis"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

In this notebook, I filter the data for relevant subgroups in the following parameters:


Then, the variogram procedure is repeated.



**Overview of distance to 0.01m difference threshold:**


| soilclass     | all in |       Jan |       ... |
|:--------------|-------:|----------:|----------:|
| heavy         |  1.7 m |     3.1 m |     2.2 m |
| light         |  0.7 m |     2.3 m |     5.7 m |
| peat          | 10.2 m |    10.6 m |     7.8 m |



# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("ggcorrplot")|> void() # correlation, grid visualization
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges
library("units")     |> void() # someone decided to soter water distance in meters...


if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")
diffdata_export_file <- paste0("data/waterlevel_diffdata_cl", cluster_dist, ".parquet")

```


## parameters

```{r parameters}
#| code-fold: false
extent <- 512 # m
n_bins <- as.integer(extent/4)

bin_target <- 32 # alternative: observations per bin
cap_dw <- 8 # m # maximum accepted water level difference
```


## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}

pp_data <- read_parquet(data_export_file)

diff_data <- read_parquet(diffdata_export_file)

pp_join <- pp_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  )

# glimpse(diff_meta)
```


## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 


fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results,
                        g = NULL,
                        color = "darkgrey"
                        ) {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  if (is.null(g)) g <- ggplot(NULL)

  g <- g +
    geom_vline(xintercept = range, color = color) +
    geom_hline(yintercept = nugget, color = color) +
    geom_hline(yintercept = sill, color = color) +
    geom_point(aes(x = regx, y = regy), size = 2.5,
               colour = "black", fill = color, alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx)), color = color)

  return(g)
}

```


```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x

regression_filtered <- function(
    regression_data, 
    select_sc, reg_var,
    maxx = extent, maxy = NULL, 
    ...) {
  diff_sc <- regression_data %>%
    filter(soilclass == select_sc) 

  g <- ggplot(NULL)

  for (select_season in 1:12) {

    diff_season <- diff_sc %>% 
      filter(season == select_season)
    
    x <- diff_season$ds
    y <- trafo(diff_season %>% pull(!!reg_var))

    y <- y[x > 0]
    x <- x[x > 0]
    y <- y[x <= maxx]
    x <- x[x <= maxx]

    if (is.null(maxy)) maxy <- max(y)
    x <- x[y <= maxy]
    y <- y[y <= maxy]
    
    x <- x[!is.na(y)]
    y <- y[!is.na(y)]

    # standardization helps function comprehension
    # x <- x/max(x)
    # y <- y / mean(y)
    
    
    matern_fit <- fit_matern(x, y,
      method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
      ...
    )
    
    print_regression_results(matern_fit,
          label = paste0(select_sc, " month ", select_season, ":", collapse = ""))


    g <- plot_matern(matern_fit, g = g, color = "darkgrey")

    x_excl <- diff_sc$ds
    y_excl <- trafo(diff_sc %>% pull(!!reg_var))

    y_excl <- y_excl[x_excl <= maxx]
    x_excl <- x_excl[x_excl <= maxx]
    x_excl <- x_excl[y_excl > maxy]
    y_excl <- y_excl[y_excl > maxy]
    g <- g + geom_point(aes(x = x_excl, y = y_excl),
        size = 2.5, colour = "red", alpha = 0.2) 

  }


  g <- g +
    labs(title = "month-wise variograms") +
    xlab("distance (m)") + ylab("mean water level difference (pair-averaged)") +
    theme_minimal()


  return(g)
}


```




# Monthwise Analysis
## Filters

Note: when filtering for location pairs "close" to water, only 298/18467 pairs would remain.
-> using "close" and "mid", i.e. distance <= 50m


```{r water-distribution}
water_threshold <- units::set_units(50, "m")

diff_meta %>%
  select(wata_min_dist1) %>%
  mutate(filltered = wata_min_dist1 < water_threshold) %>% 
  ggplot(aes(wata_min_dist1, fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r water-filter}
somediff_meta <- diff_meta %>%
  filter(
    wata_dist_class1 %in% c("close", "mid"),
    wata_dist_class2 %in% c("close", "mid")
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning

Binning parameters; will use homogeneously filled bins:


```{r join-water}
diff_all <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw # remove outliers
  ) %>%
  mutate(
    season = lubridate::month(date)
  )

# knitr::kable(t(head(diff_all, 3)), decimals = 2)
```




```{r groupwise-binning}

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = c(pair, season)
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass, season) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
      # ds, as.integer(min(c(floor(n()/8), n_bins)))
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass, season)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = factor(season))) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  theme_bw()
```




## variograms {#sec-months}

using water level difference (MAD)


```{r dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
#| label: fig-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_filtered(
  diff_equibinned,
  select_sc = "heavy",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)

```

<!--
::: {.callout-warning}
Capped: $<= 2$ 
:::
-->


### light

```{r dw-regression-light}
#| label: fig-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_filtered(
  diff_equibinned,
  select_sc = "light",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```

<!--
::: {.callout-warning}
Capped: $<= 1.6$ 
:::
-->




### peat

```{r dw-regression-peat}
#| label: fig-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_filtered(
  diff_equibinned,
  select_sc = "peat",
  reg_var = column,
  maxx = 256, maxy = 2,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```

<!--
::: {.callout-warning}
Capped: $<= 2$ 
:::
-->


:::


