---
title: "Daily Water Level Variograms: Quantile Filters"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

In this derived workbook, I filter for paired observations of water levels which fall in certain quantile ranges of all water levels of that year.
Observations are only included if *both* sites were within their respecive quantile ranges.



| soilclass  | all in | low water | high water |
|:-----------|-------:|----------:|-----------:|
| `(filter)` |   `NA` |  `<= q30` |   `>= q70` |
| (N/1000)   |   7044 |      1648 |       1626 |
| heavy      |  1.4 m |     1.4 m |      1.2 m |
| light      |  1.0 m |     0.9 m |      0.8 m |
| peat       | 11.1 m |    11.4 m |      9.1 m |


# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox


# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")

```


## parameters

```{r parameters}
#| code-fold: false
extent <- 512 # m

n_bins <- as.integer(extent/4) # the number of bins...
bin_target <- 128 # ... and the preferred alternative: observations per bin

cap_dw <- 8 # m # maximum accepted water level difference


# quantile selection
lower_quantile <- "q030"
upper_quantile <- "q070"


```

## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}
meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

quantile_data <- join_auxiliary_cache(label = "quantiles") %>%
  select(pp_idx, yr, !!enquo(lower_quantile), !!enquo(upper_quantile)) %>%
  rename(lq = !!enquo(lower_quantile), hq = !!enquo(upper_quantile))

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  )

```

```{r join-quantiles}
diff_data <- diff_data %>%
  mutate(yr = lubridate::year(date)) %>%
  left_join(quantile_data,
    by = join_by(pp_idx1 == pp_idx, yr == yr),
    relationship = "many-to-one"
  ) %>%
  mutate(lq1 = w1 <= lq, hq1 = w1 >= hq) %>%
  select(-lq, -hq) %>%
  left_join(quantile_data,
    by = join_by(pp_idx2 == pp_idx, yr == yr),
    relationship = "many-to-one"
  ) %>%
  mutate(lq2 = w2 <= lq, hq2 = w2 >= hq) %>%
  select(-lq, -hq)

knitr::kable(sample_n(diff_data, 5))
```



## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 


fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results, y_label = "semivariance (m²)") {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x

regression_by_soilclass <- function(
    regression_data, 
    sc, reg_var,
    maxx = extent, maxy = NULL, 
    ...) {
  diff_sc <- regression_data %>% filter(soilclass == sc)
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
  )
  
  print_regression_results(matern_fit, label = "regression:")
  ylabel <- "mean water level difference (pair-averaged)"
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"
  g <- plot_matern(matern_fit, y_label = ylabel)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, colour = "red", alpha = 0.2) 
  return(g)
}


```


<hr>

# Low Waters
## Filters

We only take those simultaneous measurements which are in the lowest quantile for *both* observation points.

```{r lq-water-filter}
filter_columns <- c("lq1", "lq2")
some_diff_data <- diff_data %>%
  filter(if_all(filter_columns))

paste0(
  "selected ", nrow(some_diff_data), " / ",
  nrow(diff_data), " location pairs (",
  round(100*nrow(some_diff_data)/nrow(diff_data)), "%)."
)
    
```


## Binning

Binning parameters; will use homogeneously filled bins:


```{r lq-join-water}
diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(some_diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_all, 3)), decimals = 2)
```



```{r lq-groupwise-binning}

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
      # ds, as.integer(min(c(floor(n()/8), n_bins)))
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r lq-equibin-samplesize}
#| label: fig-lq-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  theme_bw()
```




## variograms {#sec-lowwater}

using water level difference (MAD)


```{r lq-dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r lq-dw-regression-heavy}
#| label: fig-lq-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  "heavy",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)

```


### light

```{r lq-dw-regression-light}
#| label: fig-lq-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```



### peat

```{r lq-dw-regression-peat}
#| label: fig-lq-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = 256, maxy = 2,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```

:::


<hr>

# High Waters
## Filters

We only take those simultaneous measurements which are in the lowest quantile for *both* observation points.

```{r hq-water-filter}
filter_columns <- c("hq1", "hq2")
some_diff_data <- diff_data %>%
  filter(if_all(filter_columns))

paste0(
  "selected ", nrow(some_diff_data), " / ",
  nrow(diff_data), " location pairs (",
  round(100*nrow(some_diff_data)/nrow(diff_data)), "%)."
)
    
```


## Binning

Binning parameters; will use homogeneously filled bins:


```{r hq-join-water}
diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(some_diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_all, 3)), decimals = 2)
```



```{r hq-groupwise-binning}

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
      # ds, n_bins
      # ds, as.integer(min(c(floor(n()/8), n_bins)))
    ) # cut to equal sized bins
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw_groups = mean(dw, na.rm = TRUE), 
    semivar_groups = mean(semivar, na.rm = TRUE),
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


Sample size:

```{r hq-equibin-samplesize}
#| label: fig-hq-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  theme_bw()
```




## variograms {#sec-highwater}

using water level difference (MAD)


```{r hq-dw-column}
#| code-fold: false

column <- "dw_groups"
# column <- "dw_all"
```



::: {.panel-tabset group="soilclass"}

### heavy

```{r hq-dw-regression-heavy}
#| label: fig-hq-dw-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  "heavy",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)

```


### light

```{r hq-dw-regression-light}
#| label: fig-hq-dw-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = 256, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```



### peat

```{r hq-dw-regression-peat}
#| label: fig-hq-dw-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = 256, maxy = 2,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, 2)
```

:::
