---
title: "Water Level Variograms: Resampling/Bootstrapping"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("ggplot2")   |> void() # for cut_number
library("arrow")     |> void() # writing parquet files
library("parallel")  |> void() # parallel processing

# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")

# regression has been done before
source("regression_tools.R") 


# for reference
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
# cluster_dist <- 3400 # m # parameter for clustering, used on data loading


soilclass_colors <- c(
  "heavy" = "sienna",
  "light" = "lightgoldenrod",
  "peat" = "darkseagreen",
  "unknown" = "slategray"
)

```


```{r parameters}
#| code-fold: false
extent <- 500 # m
maxx <- 1000 # for regression

bin_target <- 100 # alternative: observations per bin
cap_dw <- 4 # m # maximum accepted water level difference

slope_threshold    <- 0.01/1 # 1 cm per meter
wadi_threshold     <- units::set_units(Inf, "m")
dfdepth_threshold  <- Inf # m # Inf

n_bootstraps <- 2**16 # number of bootstrapping iterations
```



```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw, m1, m2, dm), ~ !is.na(.x)))

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  ) %>%
  filter(
    ds <= extent,
    wata_min_dist1 < wadi_threshold,
    wata_min_dist2 < wadi_threshold,
    mslope < slope_threshold,
    abs(dz) < dfdepth_threshold
  )



diff_meta %>% count(soilclass1)

```


:::{.callout-warning}
- water distance was filtered to be below `{r} sprintf("%s", wadi_threshold)`.
- Pairs with an elevation difference of more than `{r} slope_threshold` m per m distance were removed.
- Filterdepth difference was limited to observations below `{r} sprintf("%s", dfdepth_threshold)`.

After that, `{r} nrow(diff_meta)` location pairs remain, counting `{r} nrow(diff_data)` simultaneous water level measurements.
:::




```{r glimpse-data}

# diff_meta %>% glimpse
# diff_data %>% glimpse

diff_meta <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, # damp,
      # dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) 


diff_joined <- diff_meta %>% 
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw, abs(dm) <= cap_dw) # remove outliers

# glimpse(diff_joined)

# diff_meta %>% select(cluster1, pp_idx1, pp_idx2, ds, mslope) %>% arrange(-ds) %>% head
```



# grouping and binning

```{r equibin-preparation}
#| code-fold: false


# test <- diff_joined %>%
#   summarize(
#     n_obs = sum(as.integer( (!is.na(dw)) & (!is.na(dm)) )),
#     soilclass = first(soilclass1),
#     cluster = first(cluster1),
#     ds = mean(ds, na.rm = TRUE),
#     dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
#     dm = 1/(n_obs)*sum(abs(dm), na.rm = TRUE),
#     .by = pair 
#   ) %>% ungroup %>% arrange(-ds)

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer( (!is.na(dw)) & (!is.na(dm)) )),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    dm = 1/(n_obs)*sum(abs(dm), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = ggplot2::cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup

    
diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    dw = mean(dw, na.rm = TRUE), 
    dm = mean(dm, na.rm = TRUE), 
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

```


```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  scale_fill_manual(values = soilclass_colors) +
  scale_color_manual(values = soilclass_colors) +
  theme_bw()
```




## Regression Procedure Wrapper

```{r regression-wrapper}
#| code-fold: true

matern_function_3p <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- 1 # parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}


fit_matern <- function(x, y, ...) {
}
```



```{r regression-matérn-general}
trafo <- function(x) x

regression_wrapper <- function(
    regression_data, 
    regression_variable = "dw",
    ...) {
  
  x <- regression_data$ds
  y <- trafo(regression_data %>% pull(!!regression_variable))

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= extent]
  x <- x[x <= extent]

  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

 
  ## using L-BFGS-B to get the optimizer close
  pre_ <- optim(
    fn = function(params) {
      distweighted_target_function(x, y, matern_function_3p, params)
      # wrap_target_function(x, y, matern_function_3p, params)
    },
    # par = par,
    # upper = upper,
    # lower = lower,
    ...,
    method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
  )

  ## then performing a local Nelder-Mead adjustment
  optimizer_results <- pre_
  # optimizer_results <- optim(
  #   fn = function(params) {
  #     distweighted_target_function(x, y, matern_function_3p, params)
  #     # wrap_target_function(x, y, matern_function_3p, params)
  #   },
  #   par = pre_$par,
  #   method = "L-BFGS-B",
  #   lower = lower_limits, upper = upper_limits,
  #   control = list("fnscale" = 1000)
  # )
  # optimizer_results <- optim(
  #   fn = function(params) {
  #     distweighted_target_function(x, y, matern_function_3p, params)
  #     # wrap_target_function(x, y, matern_function_3p, params)
  #   },
  #   par = pre_$par,
  #   method = "Nelder-Mead" # "L-BFGS-B"
  # )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$pre_par <- pre_$par
  optimizer_results$fcn <- matern_function_3p
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function_3p,
    optimizer_results
  )

  # calculate 1cm threshold
  optimizer_results$threshold <- calculate_limit(
    optimizer_results,
    threshold = 0.01,
    fit_fcn = matern_function_3p,
    prep_fcn = shift_nugget_matern3p
  )

  # par <- optimizer_results$par
  # par[3] <- 0 # no nugget
  # threshold = 0.01

  return(optimizer_results)
}


```




# variograms {#sec-variograms}
... in repeated execution.





# bootstrapping
- <https://www.modernstatisticswithr.com/modchapter.html#a-general-approach>


## all in: getting starting values

```{r regression-reference}
lower_limits = c(0.0,  0.0, 0.0)# 0.1)
upper_limits = c(4.0, maxx, 4.0)#16.01)
reference <- regression_wrapper(
  regression_data = diff_equibinned,
  regression_variable = "dw",
  par = c(0.5, 10., 0.1), 
  lower = lower_limits,
  upper = upper_limits
)


g <- ggplot(NULL) %>% 
  add_regression_to_plot(reference, color = "darkgrey")
g + xlab("distance (m)") + ylab("mean absolute difference (m)") +
  ylim(0, 1.6) +
  theme_minimal()
```


```{r parallel-bootstrap}

# straps <- bind_rows(mclapply(
#   1:n_iterations,
#   datawrap_bootstrap("all", diff_equibinned),
#   mc.cores = detectCores()-1
#   ))
# 
# straps %>%
#   filter(conv == 0) %>% 
#   ggplot(aes(x = threshold)) +
#   geom_histogram(bins = 256, fill = "lightgray", color = "black") +
#   labs(x = "threshold", y = "bootstrap", title = sc) +
#   theme_bw()
    
```

# wrapping by soilclass

## bootstrap function
```{r bootstrap}


datawrap_bootstrap <- function(reference, pass_data) {

  bootstrap_function <- function(i){
    # Draw row numbers for the bootstrap sample:
    row_numbers <- sample(1:nrow(pass_data), as.integer(nrow(pass_data)),
                          replace = TRUE)
    
    # Obtain the bootstrap sample:
    boot_sample <- pass_data[row_numbers,]
    
    # jittering the observations
    # see https://influentialpoints.com/Training/properties-of-bootstrap_estimators.htm#howm
    boot_sample$dw <- jitter(boot_sample$dw, amount = 0.1*sd(boot_sample$dw))

    result <- regression_wrapper(
      regression_data = boot_sample,
      regression_variable = "dw",
      # par = reference$par,
      par = c(0.5, 10., 0.1),
      lower = lower_limits,
      upper = upper_limits
    )
    return(as.data.frame(list(
        "i" = i,
        "conv" = result$convergence,
        "rmse" = result$value,
        "scale" = result$par[1],
        "sigma" = result$par[2],
        "nugget" = result$par[3],
        # "nu" = result$par[4],
        "threshold" = result$threshold
      )))
  }
  return(bootstrap_function)
}


```


## subset

```{r subset-bootstrap}

ref_store <- list()

subset_bootstrap <- function(sc, sub_data, n_iterations = 999){
  reference <- regression_wrapper(
    regression_data = sub_data,
    par = c(0.5, 10, 0.1), # 1.0)
    lower = lower_limits,
    upper = upper_limits,
    control = list("fnscale" = 1000)
  )


  ref_store[[sc]] <<- as.data.frame(list(
      "i" = 0,
      "conv" = reference$convergence,
      "rmse" = reference$value,
      "scale" = reference$par[1],
      "sigma" = reference$par[2],
      "nugget" = reference$par[3],
      # "nu" = reference$par[4],
      "threshold" = reference$threshold,
      "soilclass" = sc
    ))

  straps <- bind_rows(mclapply(
    1:n_iterations,
    datawrap_bootstrap(reference, sub_data),
    mc.cores = detectCores()-2
    ))
  return (straps)
}

```

```{r soil-subset-bootstrap}

all_dolled_up <- list()
for (sc in c("heavy", "light", "peat")) {
  sub_data <- diff_equibinned %>% filter(soilclass == sc)
  straps <- subset_bootstrap(sc, sub_data, n_iterations = n_bootstraps)
  straps$soilclass <- sc
  all_dolled_up[[sc]] <- straps
}

all_dolled_up[["ref"]] <- bind_rows(ref_store)
all_dolled_up <- bind_rows(all_dolled_up)
all_dolled_up <- all_dolled_up %>%
  mutate(sill = nugget + scale)

# save bootstraps
arrow::write_parquet(all_dolled_up, sink = "./data/bootstrapping_dw_soilclass.parquet")

```


```{r count}

# count failed regressions
all_dolled_up %>% count(conv)

```




```{r one-parameter}

ref_boots <- all_dolled_up %>%
  filter(i == 0)
sub_boots <- all_dolled_up %>%
  filter(conv == 0, i > 0, if_all(everything(), ~ !is.na(.x)))

# param <- "threshold"
knitr::kable(t(ref_boots))

# knitr::kable(t(head(sub_boots)))
print(summary(sub_boots))
```



```{r plot-one-parameter}
plot_param <- function(param, limit_x = NA){
  h <- sub_boots %>%
    mutate(param = sub_boots[, param]) %>% 
    ggplot(aes(x = param, fill = soilclass)) +
    geom_histogram(bins = 256, width = 1.05, color = NA, alpha = 0.67, position = "identity") +
    geom_vline(xintercept = ref_boots[,param]) +
    xlab(param) +
    scale_fill_manual(values = soilclass_colors) +
    scale_color_manual(values = soilclass_colors) +
    theme_bw()

  # if (param == "threshold") h <- h + xlim(0, 50)
  # if (param == "scale") h <- h + xlim(lower_limits[1], upper_limits[1])
  # if (param == "sigma") h <- h + xlim(lower_limits[2], upper_limits[2])
  # if (param == "nugget") h <- h + xlim(lower_limits[3], upper_limits[3])

  # if (TRUE) h <- h + ylim(0, 100) # ARRRR! R blanks bars above the ylim 8(
  if (!is.na(limit_x)) h <- h + xlim(0, limit_x)

  ggsave(paste0("figures/bootstrap_dw_", param, ".png"), h, dpi = 300)
  return(h)
}

```


Quantiles for `threshold`:

```{r threshold-quantiles}
threshold_quantiles <- sub_boots %>%
  summarize(
    q02 = quantile(threshold, c(0.02)),
    median = quantile(threshold, c(0.50)),
    q98 = quantile(threshold, c(0.98)),
    .by = soilclass 
  ) 
knitr::kable(threshold_quantiles, digits = 1)
```


Plots:

```{r plot-all-parameters}
silence <- suppressWarnings
print(plot_param("threshold", limit_x = 35) |> silence())
print(plot_param("rmse") |> silence())
print(plot_param("scale", limit_x = 2) |> silence())
print(plot_param("sigma", limit_x = 500) |> silence())
print(plot_param("nugget") |> silence())
print(plot_param("sill", limit_x = 2) |> silence())

```



# Preliminary Status Quo (20250210)

After *a lot* of trial-and-error, I note the following:

- The default bootstrap regression (L-BFGS-N, with sensible limits) returns bimodal outcomes on the bootstraps if started with the reference outcome.
- Yet these depend on start values, bins, *et cetera*. With the default, some parameters come out at the limits; threshold (as a parameter mainly derived from sigma) is heavily quantized depending on secondary fit params.
- This is especially obvious for the peat subset.
- BFGS followed by Nelder-Mead seems stable; Nelder-Mead seems to find a result in a Normal range, yet values are relatively meaningless (sigma > limits, some nuggets < 0, ...).
- Filtered for water distance, slope, filter depth difference; `{r} round(100*4801/18741, 1)` \% of the observations pass all criteria.
- Binning plays a role, but is kept dynamic.
- Distance weighting of the regression RMSE might be influential; applied here together with the `fnscale` control parameter.
- Regressions are convergent because of the start values (low, then running into the lowest local minimum).

# Preliminary Status Quo (20250310)

After *even more* trial, little less error:

- start values are more robust
- ranges adjusted -> no more quantization
- filtering for `slope` is default
- binning improved
- distance weighting retained
- confirmation when using `mMaaiveld`
- still, several outcomes are bimodal (scales <- discussion Thierry)


seems (technically) robust now.
