---
title: "Closter LOO"
subtitle: "Leave-One-Out of the big, potentially influential clusters."
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

- Some of the big clusters contain locations of one or two soilclasses.
- Filtering them out affects those soilclass regressions to some extent (most obvious: `peat`).

Overall, the outcomes below seem plausible and expected.


As a bonus, and maybe the most valuable contribution of this partial analysis, some clusters are plotted on maps.


# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("units")     |> void() # someone decided to store water distance in meters...
# library("mapview")   |> void() # showing cluster locations on a map
library("sf")        |> void() # organizing spatial objects (map plotting)
library("terra")     |> void() # raster elevation map plotting
library("inbospatial") |> void() # query elevation data

# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
```


## parameters

```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

n_bins <- as.integer(extent/4) # the number of bins...
bin_target <- 100 # ... and the preferred alternative: observations per bin

cap_dw <- 8 # m # maximum accepted water level difference

# difference collumn used for variogram
diff_column <- "dw" # using water level difference (MAD)
```


## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx")
  # %>% join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         # emd_n, emd_amp1, emd_amp2, emd_ampsum,
         # emd_qdiff, emd_residual_qdiff,
         pp_vld, mp_vld
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      validated1 = as.integer(pp_vld1 == "VLD") * as.integer(mp_vld1 == "VLD"),
      validated2 = as.integer(pp_vld2 == "VLD") * as.integer(mp_vld2 == "VLD"),
      validated = as.logical(validated1 * validated2),
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      # damp = emd_ampsum2-emd_ampsum1,
      # dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  ) %>% 
  filter(ds <= extent)

```



```{r count-all}
diff_joined <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

count_all <- nrow(diff_joined)
print(count_all)
```


## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 

```

*(... has been outsourced)*


# Putatively Influential Clusters 

```{r sort-clusters}
cluster_count <- diff_meta %>% 
  count(cluster1) %>% 
  rename(c("cluster" = cluster1)) %>% 
  arrange(desc(n))
big_clusters <- cluster_count %>% filter(n > 500)
knitr::kable(big_clusters)
```

  
```{r show-clustersizes}
cluster_count %>% 
  ggplot(aes(x = 1:length(cluster), y = n)) +
  geom_vline(xintercept = nrow(big_clusters)) +
  geom_step()
```


# Procedure 

```{r filter}

i <- 1
cluster_idx <- as.integer(big_clusters[i, "cluster"][[1]])
```

```{r map-cluster}

map_cluster <- function(cluster_idx){

  message("________________________________________")
  message(cluster_idx)

  if (cluster_idx == 0){
    map_data <- diff_meta %>%
      sf::st_as_sf(coords = c("x1", "y1"), crs = 31370)
  } else {
    map_data1 <- diff_meta %>%
      filter(cluster1 == cluster_idx)  %>%
      mutate(x = x1, y = y1, idx = pp_idx1, code = pp_code1) %>%
      select(x, y, idx, code)
    map_data2 <- diff_meta %>%
      filter(cluster1 == cluster_idx, !(pp_idx2 %in% map_data1$idx) ) %>%
      mutate(x = x2, y = y2, idx = pp_idx2, code = pp_code2) %>%
      select(x, y, idx, code)

    map_data <- bind_rows(list(map_data1, map_data2)) %>% 
      mutate(
        region = substring(code, first = 1, last = 3)
      ) %>% 
      sf::st_as_sf(coords = c("x", "y"), crs = 31370)
  }

  regions <- paste(unique(map_data$region), collapse = ", ")

  # map_data %>% 
  #   mapview(cex = 3)
  coords <- sf::st_coordinates(map_data)

  # margin may not be too large, otherwise wcs query may fail
  # 200m seemed still okay
  bbox <- sf::st_bbox(
     c(xmin = min(coords[, 1]) - 0.4*extent,
       xmax = max(coords[, 1]) + 0.4*extent,
       ymin = min(coords[, 2]) - 0.4*extent,
       ymax = max(coords[, 2]) + 0.4*extent),
     crs = sf::st_crs(31370)
  )


  # cache raster
  filename <- here::here("cache", "maps",
                         paste0(
                           sprintf("map_cl%04.0f", cluster_idx),
                           ".raster"
                         ))
  if (file.exists(filename)){
    map_raster <- terra::readRDS(filename)
  } else {
    map_raster <- inbospatial::get_coverage_wcs(
      wcs = "dhmv",
      bbox = bbox,
      layername = "DHMVII_DTM_1m",
      version = "2.0.1",
      wcs_crs = "EPSG:31370",
      resolution = 1
    )
    terra::saveRDS(map_raster, filename)
  }


  # other locations in range
  other_data <- diff_meta %>%
    filter(cluster1 != cluster_idx) %>% 
    sf::st_as_sf(coords = c("x1", "y1"), crs = 31370) %>% 
    sf::st_intersection(sf::st_as_sfc(bbox))
  other_coords <- sf::st_coordinates(other_data)
  
  # plot map
  plot(map_raster, col = gray.colors(256),
       ylab = sprintf("cluster %i (%s)", cluster_idx, regions)
       )
  points(coords[, 1], coords[, 2], pch = 21, col = "black")
  points(coords[, 1], coords[, 2], pch = 16, col = "yellow")
  points(other_coords[, 1], other_coords[, 2], pch = 4, col = "darkorange")
  # title(sprintf("cluster %i", cluster_idx))

}

# map_cluster(cluster_idx)
# map_cluster(30) # one with critical size for wcs query
```


```{r map-cluster-loop}
for (i in 1:nrow(big_clusters)) suppressWarnings(
  map_cluster(as.integer(big_clusters[i, "cluster"][[1]]))
  )

```


## Binning


```{r join-diff-data}
join_diff_data_excluding <- function(exclude_cluster_idx) {

  somediff_meta <- diff_meta %>%
    filter(
      !(cluster1 == exclude_cluster_idx)
    )
  
  paste0(
    "selected ", nrow(somediff_meta), " / ",
    nrow(diff_meta), " location pairs (",
    round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
  )
  
  diff_joined <- somediff_meta %>%
    select(pp_idx1, pp_idx2, cluster1,
        x1, y1, x2, y2, 
        soilclass1, wata_dist_class1, wata_dist_class2,
        ds, dh, dz,
        # damp, dqdiff,
        dwadi, dslope,
        melevation, mslope
    ) %>%
    mutate(
      pair = factor(paste0(
        as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
      )
    ) %>%
    left_join(diff_data,
      by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
      relationship = "one-to-many"
    ) %>% 
    filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers
  
  # knitr::kable(t(head(diff_joined, 3)), decimals = 2)
  
  paste0(
    "selected ", nrow(diff_joined), " / ",
    count_all, " observations (",
    round(100*nrow(diff_joined)/(count_all)), "%)."
  )

  diff_equigrouped <- diff_joined %>%
    summarize(
      n_obs = sum(as.integer(!is.na(dw))),
      soilclass = first(soilclass1),
      cluster = first(cluster1),
      ds = mean(ds, na.rm = TRUE),
      semivar = 1/(2*n_obs)*sum(dw^2, na.rm = TRUE),
      dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
      n_pairs = 1,
      .by = pair 
    ) %>% ungroup %>% arrange(ds) %>%
    group_by(soilclass) %>% 
    mutate(
      distance_equibin = cut_number(
        ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
        # ds, n_bins
      ) # cut to equal sized bins
    ) %>%
    ungroup
  
  
  diff_equibinned <- diff_equigrouped %>%
    summarize(
      cnt = n(),
      ds = mean(ds, na.rm = TRUE), 
      dw = mean(dw, na.rm = TRUE), 
      n_obs = sum(n_obs),
      n_pairs = sum(n_pairs),
      .by = c(distance_equibin, soilclass)
    ) %>%
    ungroup %>% arrange(ds)
  
  return(diff_equibinned)
}

diff_equibinned <- join_diff_data_excluding(cluster_idx)
```




Sample size:

```{r equibin-samplesize}
#| label: fig-equibin-sample-size
#| fig-cap: "Cumulative size per bin for the different soilclass groups. Bin widths are variable to generate balanced sample sizes. The height (''Δy'') of each step is identical. Perfectly homogeneous bins would fall on the diagonal line; steep sloped ranges indicate narrow bins, whereas wider-than-average bins would manifest in shallower slopes."

plot_sample_size <- function(diff_equibinned) {
  diff_equibinned %>%
    ggplot(aes(x = ds, y = cnt, color = soilclass)) +
    geom_step(stat = "ecdf") +
    geom_hline(yintercept = 0, color = "darkgrey") +
    geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
    scale_fill_manual(values = soilclass_colors) +
    scale_color_manual(values = soilclass_colors) +
    theme_bw()
}

plot_sample_size(diff_equibinned)
```


```{r}
plot_regression <- function(diff_equibinned, soilclass) {
  regression_by_soilclass(
    diff_equibinned,
    sc = soilclass,
    reg_var = diff_column,
    maxx = extent, maxy = Inf,
    return_fit = FALSE,
    distweighted = TRUE,
    par = c(1., 10, 0., 1.),
    lower = c(0.0, 0.0, 0., 0.99),
    upper = c(4., 200, 0.5, 1.01)
  ) + ylim(0, 1.6)

}


perform_regression <- function(diff_equibinned, soilclass) {
  fit <- regression_by_soilclass(
    diff_equibinned,
    sc = soilclass,
    reg_var = diff_column,
    maxx = extent, maxy = Inf,
    return_fit = TRUE,
    distweighted = TRUE,
    par = c(1., 10, 0., 1.),
    lower = c(0.0, 0.0, 0., 0.99),
    upper = c(4., 200, 0.5, 1.01)
  ) 
  return(fit)
}
```


## variograms {#sec-example}


::: {.panel-tabset group="soilclass"}

### heavy

```{r dw-regression-heavy}
#| label: fig-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."

plot_regression(diff_equibinned, "heavy")

```


### light

```{r dw-regression-light}
#| label: fig-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."

plot_regression(diff_equibinned, "light")
```





### peat

```{r dw-regression-peat}
#| label: fig-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."

plot_regression(diff_equibinned, "peat")
```

:::


# Iterated Calculation

## Procedure

```{r quick_regression}

collect_regression <- function(i){
  collection <- list()
  if (i == 0) {
  cluster_idx <- 0
  } else {
  cluster_idx <- as.integer(big_clusters[i, "cluster"][[1]])
  }

  collection["cluster_idx"] <- cluster_idx
  # map_cluster(cluster_idx)

  diff_equibinned <- join_diff_data_excluding(cluster_idx)
  # plot_sample_size(diff_equibinned)
  collection["n"] <- sum(diff_equibinned$n_obs)
  collection["n_pairs"] <- sum(diff_equibinned$n_pairs)

  for (sc in c("heavy", "light", "peat")){
    fit <- perform_regression(diff_equibinned, sc)
    # print_regression_results(fit)
    # 
    diff_sc <- diff_equibinned %>% filter(soilclass == sc)
    # collection[paste0(sc, "_bins")] <- fit$regn
    collection[paste0(sc, "_pairs")] <- sum(diff_sc$n_pairs)
    # collection[paste0(sc, "_n")] <- sum(diff_sc$n_obs)
    collection[paste0(sc, "_threshold")] <- calculate_limit(fit, 0.01,
      matern_function,
      prep_fcn = shift_nugget_matern4p)
    collection[paste0(sc, "_scale")] <- fit$par[1]
    collection[paste0(sc, "_sigma")] <- fit$par[2]
    collection[paste0(sc, "_nugget")] <- fit$par[3]
  }
  return(as.data.frame(collection))
}

results <- bind_rows(lapply(
  0:nrow(big_clusters),
  FUN = collect_regression
))
```


## overview {#sec-overview}

```{r table-results}
# knitr::kable(results, digits = 2)
```


::: {.panel-tabset group="soilclass"}

### heavy

```{r table-heavy}

knitr::kable(results %>%
  select(cluster_idx, starts_with("heavy_")) %>%
  mutate(percent = round(100* heavy_pairs / results[1, "heavy_pairs"], 1))
  ,
  digits = 2)

```


### light

```{r table-regression-light}

knitr::kable(results %>%
  select(cluster_idx, starts_with("light_")) %>%
  mutate(percent = round(100* light_pairs / results[1, "light_pairs"], 1))
  ,
  digits = 2)
```


### peat

```{r table-peat}

knitr::kable(results %>%
  select(cluster_idx, starts_with("peat_")) %>%
  mutate(percent = round(100* peat_pairs / results[1, "peat_pairs"], 1))
  ,
  digits = 2)
```

:::


:::{.callout-note}
- fake cluster `0` is the reference result for "all clusters included" (the only one to have $100\%$ inclusion on all soilclasses)
- the other rows show results with one cluster filtered, respectively
- "*_pairs" column indicates the number of remaining location pairs included in calculation

:::
