---
title: "Water Level Variograms: Mean Absolute Differential Difference"
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Introduction

## purpose

Water levels of proximal observation wells can be offset,
e.g. due to differences in `mTAW` reference calibration,
or generally, slope.
When calculating a variogram based on "mean absolute difference", these differences will affect the variogram by exaggerating difference.


However, if two sites are coupled, the first temporal derivative of their water levels should also be coupled.
In other words: if water level on one site changes, then the simultaneous change on the other site is expected to This can be quantified by calculating `MADD`, **Mean Absolute Differential Difference**, from the `MAD` (Mean Absolute Difference).


:::{.callout-warinig}
Note: 
numeric differentials are horribly erratic. 
Treat results with some caution.

(Would be great to `savgol` the densely sampled timelines.)
:::


There are options for presentation of the output:

- **time-agnostic:** ignore time interval, just take absolute differences -> `ddw`
- **time-normalized** ("local tangent"): divide water level difference by time interval -> dtw
- **time-filtered**: "filter" time dimension, e.g. take `dt` as an extra variogram axis (-> *not shown*)


## Results

See [@sec-variograms-ddw]:

- only the time-agnostic analysis gives relatively acceptable outcomes
- noise is much higher, as expected
- **ranges are much shorter,** consistent with the fact that we use mTAW and the previously noted influence of slope
- this could be extended to analyze temporal coupling of water levels!



## libraries

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox


# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")

progress <- TRUE # print progress bar during reload
ddiff_file <- paste0("_waterlevels_differentials.parquet", collapse = "")
reload <- FALSE | !file.exists(here::here("cache", ddiff_file))
```

## settings 

```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

bin_target <- 100 # alternative: observations per bin
cap_ddw <- 4 # m # maximum accepted water level DIFFERENTIAL
cap_dt <- 32 # d # maximum accepted time difference in days
```


# Data 
## Loading


```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

diff_data <- diff_data %>%
  select(pp_idx1, pp_idx2, date, w1, w2) %>%
  arrange(pp_idx1, pp_idx2, date) 

pairs <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct()
```


## Differentials

The `lag` function should work fine, yet this is not the most efficient method.

Applying the most simple split/loop/aggregate/save strategy:

```{r differentials}

if (reload) {
  if (progress) pb <- txtProgressBar(min = 0, max = nrow(pairs),
    initial = 0, style = 1)
  
  
  pairwise_differential <- function(i){
  
    if (progress) setTxtProgressBar(pb,i)
    pp1 <- pairs[[i, "pp_idx1"]]
    pp2 <- pairs[[i, "pp_idx2"]]
  
    ddiff <- diff_data %>%
      filter(pp_idx1 == pp1, pp_idx2 == pp2) %>% 
      arrange(date) %>% 
      mutate(
        dt  = date - lag(date),
        dw1 = w1   - lag(w1),
        dw2 = w2   - lag(w2)
      ) %>%
      mutate(dt = as.integer(dt)/3600/24) %>% 
      select(-w1, -w2) %>%
      filter(!is.na(dt), dt > 0)
  
    return(ddiff)
  }
  
  ddiff <- dplyr::bind_rows(
    lapply(
      1:nrow(pairs),
      FUN = pairwise_differential
    )
  )
  
  if (progress) close(pb) # close the progress bar
  
  write_parquet(ddiff, sink = here::here("cache", ddiff_file))

}

ddiff <- read_parquet(here::here("cache", ddiff_file))
# ddiff <- ddiff %>%  mutate(dt = as.integer(dt)/3600/24)
# ddiff <- ddiff %>% filter(dt > 0)
ddiff <- ddiff %>%
  filter(abs(dw1) <= cap_ddw, abs(dw2) <= cap_ddw) %>% 
  mutate(ddw = abs(dw2 - dw1), dtw = ddw/as.integer(abs(dt))) %>%
  mutate(ddw = 1000*ddw, dtw = 1000*dtw)
knitr::kable(head(ddiff, 32))
```


## join metadata


```{r}

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- pairs %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
      mslope = abs(dh) / ds
  ) %>%
  filter(ds <= extent)

```



## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 


fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results, y_label = "semivariance (m²)") {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x

regression_by_soilclass <- function(
    regression_data, 
    sc, reg_var,
    maxx = maxx, maxy = NULL, 
    ...) {
  diff_sc <- regression_data %>% filter(soilclass == sc)
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
  )
  
  print_regression_results(matern_fit, label = "regression:")
  ylabel <- "mean water level difference (pair-averaged)"
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"
  g <- plot_matern(matern_fit, y_label = ylabel)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, colour = "red", alpha = 0.2) 
  return(g)
}


```


## Binning


[!] here comes the ddiff data


```{r join-water}
diff_joined <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(ddiff,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% filter(abs(dt) <= cap_dt)
# %>% filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(sample_n(diff_joined, 5)), decimals = 2)
```



```{r groupwise-binning}

diff_equigrouped <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(ddw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dt = mean(dt, na.rm = TRUE),
    ddw = 1/(n_obs)*sum(abs(ddw), na.rm = TRUE),
    dtw = 1/(n_obs)*sum(abs(dtw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds, dt) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ), # cut to equal sized bins
    time_bin = cut(dt,
      breaks = seq(0, cap_dt,
        length.out = as.integer(floor(cap_dt/7.)+1)
      ) # cut to 1d bins
    )
  ) %>%
  ungroup


diff_equibinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dt = mean(dt), 
    ddw = mean(ddw, na.rm = TRUE), 
    dtw = mean(dtw, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

diff_timebinned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dt = mean(dt), 
    ddw = mean(ddw, na.rm = TRUE), 
    dtw = mean(dtw, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, time_bin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)
# glimpse(diff_equibinned)
```


Sample size:

```{r equibin-samplesize-ds}
#| label: fig-equibin-sample-size-ds
#| fig-cap: ""

diff_equibinned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/extent, color = "darkgrey") +
  theme_bw()
```

```{r equibin-samplesize-dt}
#| label: fig-equibin-sample-size-dt
#| fig-cap: ""

diff_timebinned %>%
  ggplot(aes(x = dt, y = cnt, color = soilclass)) +
  geom_step(stat = "ecdf") +
  geom_hline(yintercept = 0, color = "darkgrey") +
  geom_abline(intercept = 0, slope = 1/cap_dt, color = "darkgrey") +
  theme_bw()
```



# Plain Vanilla Difference in Water Level Differentials {#sec-variograms-ddw}

Using **time-agnostic** water level differential difference (MADD),
i.e. the variant that is not time-interval-normalized.
Time intervals are filtered to be below `{r} cap_dt` days, but for the rest `dt` is not taken into account.


```{r dw-column}
#| code-fold: false

column <- "ddw"
```



::: {.panel-tabset group="soilclass"}

## heavy

```{r dw-regression-heavy}
#| label: fig-dw-regression-heavy-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  # diff_equibinned %>% filter(time_bin == "(24,32]"),
  # diff_equibinned %>% filter(time_bin == "(16,24]"),
  # diff_equibinned %>% filter(time_bin == "(8,16]"),
  #diff_equibinned %>% filter(time_bin == "(0,8]"),
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(10., 10, 10., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(400., 250, 100., 1.01),
  # upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, 1)

```



## light

```{r dw-regression-light}
#| label: fig-dw-regression-light-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(400., 250, 100., 1.01),
) 
```





## peat

```{r dw-regression-peat-25m}
#| label: fig-dw-regression-peat-25m
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(400., 250, 100., 1.01),
) 
```



:::





# Time-Normalized {#sec-variograms-dtw}

Using **time-normalized** water level differential difference (MADD),
i.e. true water level differentials, $\frac{dw_i}{dt}$.
Time intervals are filtered to be below `{r} cap_dt` days.
Water level differentials (i.e. temporal changes in water level) are normalized by the time interval over which they were observed.

(Because water levels only vary within a certain range, this might be over-normalization, leading to an exaggerated influence of the short intervals. 
The procedure will also amplify noise.)


```{r dw-column-dtw}
#| code-fold: false

column <- "dtw"
```



::: {.panel-tabset group="soilclass"}

## heavy

```{r dw-regression-heavy-dtw}
#| label: fig-dw-regression-heavy-dtw
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  # diff_equibinned %>% filter(time_bin == "(24,32]"),
  # diff_equibinned %>% filter(time_bin == "(16,24]"),
  # diff_equibinned %>% filter(time_bin == "(8,16]"),
  #diff_equibinned %>% filter(time_bin == "(0,8]"),
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(10., 10, 10., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(40., 250, 10., 1.01),
  # upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, 1)

```


## light

```{r dw-regression-light-dtw}
#| label: fig-dw-regression-light-dtw
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(40., 250, 10., 1.01),
) 
```



## peat

```{r dw-regression-peat-dtw}
#| label: fig-dw-regression-peat-dtw
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0., 0., 0., 0.99),
  upper = c(40., 250, 10., 1.01),
) 
```


:::



