---
title: "Sampling Frame Evaluations"
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
knitr:
  opts_chunk:
    echo: true
---

```{r libraries}
library("dplyr")
library("magrittr")
```

# scraping breadcrumbs

Using the example sampling frame drawn by @FV.

- load relevant data (stored via POC) 
- get spatial coordinates for `scheme == "GW_03.3"`
- combine with existing watina locations
- cross-distance
- plot cumulative number of location pairs at a given distance


see also

- <https://inbo.github.io/watina/reference/selectlocs_xg3.html>
- <https://inbo.github.io/watina/reference/selectlocs_chem.html>


Suggestions @FV

- find `spsamples_gw_sf_panflpan5`?
  - not in `objects_panflpan5.RData`, though that has a `samplinglocations_sf` (4514x2, grts_address and geometry)
  -> is a `gpkg` files: `spsamples_gw_sf_panflpan5.rds` and `spsamples_gw_sf_panflpan5.gpkg`
- filter `hydr_class != "HC3"`


```{r}
#| eval: false
grts_mh_n2khab <- rast(
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif")
)
grts_mh_n2khab_index <- tibble(
  id = seq_len(ncell(grts_mh_n2khab)),
  grts_address = values(grts_mh_n2khab)[, 1]
) %>%
  filter(!is.na(grts_address))
```


```{r}
#| eval: false
sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
domain_grts_n2khab <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
```

# load

```{r load-gpkg}
datapath <- "/home/falk/Downloads/steekproef"

# samples <- sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
samples <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
samples <- cbind(sf::st_drop_geometry(samples), sf::st_coordinates(samples))

samples %<>% dplyr::filter(hydr_class != "HC3")

knitr::kable(t(dplyr::sample_n(samples, 4)))
```


"only" 978 locations left.
