---
title: "Sampling Frame Evaluations"
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
knitr:
  opts_chunk:
    echo: true
---

```{r libraries}
library("dplyr")
library("magrittr")
library("ggplot2")

# data is loaded from cache
source("./auxiliary_data_queries.R")
source("regression_tools.R") 
```

# scraping breadcrumbs

Using the example sampling frame drawn by @FV.

- load relevant data (stored via POC) 
- get spatial coordinates for `scheme == "GW_03.3"`
- combine with existing watina locations
- cross-distance
- plot cumulative number of location pairs at a given distance


see also

- <https://inbo.github.io/watina/reference/selectlocs_xg3.html>
- <https://inbo.github.io/watina/reference/selectlocs_chem.html>


Suggestions @FV

- find `spsamples_gw_sf_panflpan5`?
  - not in `objects_panflpan5.RData`, though that has a `samplinglocations_sf` (4514x2, grts_address and geometry)
  -> is a `gpkg` files: `spsamples_gw_sf_panflpan5.rds` and `spsamples_gw_sf_panflpan5.gpkg`
- filter `hydr_class != "HC3"`


```{r}
#| eval: false
grts_mh_n2khab <- rast(
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif")
)
grts_mh_n2khab_index <- tibble(
  id = seq_len(ncell(grts_mh_n2khab)),
  grts_address = values(grts_mh_n2khab)[, 1]
) %>%
  filter(!is.na(grts_address))
```


```{r}
#| eval: false
sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
domain_grts_n2khab <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
```

# load

```{r load-gpkg}
datapath <- "/home/falk/Downloads/steekproef"

# samples <- sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
samples <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
samples <- cbind(sf::st_drop_geometry(samples), sf::st_coordinates(samples))

samples %<>% dplyr::filter(hydr_class != "HC3")

knitr::kable(t(dplyr::sample_n(samples, 4)))
```


"only" 978 locations left.


# samples by themselves


```{r cross-difference}
# compute the difference of all elements of one vector to each other
self_difference <- function (vec) outer(X = vec, Y = vec, FUN = function(X, Y)  Y-X)

# Calculate the Euclidean distance of the x and y columns in a data frame.
Euclid <- function (x, y) sqrt(self_difference(x)^2 + self_difference(y)^2 )

# return the lower triangle of a matrix, unpacking it into a vector of unique values
extract_lower_triangle <- function (mat) mat[lower.tri(mat)]

# retrieve elements of a vector which are below a given threshold
sub_threshold <- function (vec, thresh) vec[vec < thresh]

# cross distance of sampling frame locations
self_distances <- Euclid(samples$X, samples$Y)
# print(extract_lower_triangle(self_distances[1:6, 1:6]))
close_distances <- sub_threshold(extract_lower_triangle(self_distances), 256)
ggplot(NULL, aes(x = close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  theme_bw()

```

(For reference, the diagonal of a 32x32 grid cell is `{r} sqrt(32^2+32^2)` m.)


- As expected, there are a number of co-located samples. However, I would not expect so many.
- Cross distances are quantized, totally consistent with the GRTS grid cell arrangement of 32m resolution.


# existing sites

```{r load-watina}
watina_data <- join_auxiliary_cache(label = "metadata")

watina_cross_distances <- Euclid(watina_data$x, watina_data$y)
watina_close_distances <- sub_threshold(extract_lower_triangle(watina_cross_distances), 256)

ggplot(NULL, aes(x = watina_close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  scale_y_log10() +
  theme_bw()

```

Seems more evenly distributed.
Now the interesting part: how are the new samples positioned relative to the existing ones.


# samples and existing sites


```{r cross-distance}

cross_difference <- function (v1, v2) outer(X = v1, Y = v2, FUN = function(X, Y)  Y-X)
Euclid2 <- function (x1, y1, x2, y2) sqrt(cross_difference(x1, x2)^2 + cross_difference(y1, y2)^2 )

cross_distances <- Euclid2(
  x1 = watina_data$x,
  y1 = watina_data$y,
  x2 = samples$X,
  y2 = samples$Y
)

print(nrow(cross_distances))
print(ncol(cross_distances))
print(cross_distances[1:6, 1:6])

close_distances <- sub_threshold(cross_distances, 256)


ggplot(NULL, aes(x = close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  theme_bw()

```


For testing any hypothetical coupling rule, we can use the cumulated empirical distribution.

```{r cumulated-distances}
#| label: fig-cumulated-count
#| fig-cap: "Cumulated count of observed distances between existing watina locations and sampling frame draws."

count_threshold <- function (thresh){
  n <- length(sub_threshold(close_distances, thresh))
  return(tibble(distance = thresh, cumulated_count = n))
}

cumulative_crossdistcount <- bind_rows(
  lapply(
    seq(0, 256, length.out = 257),
    FUN = count_threshold)
)

cumulative_crossdistcount %>%
  ggplot(aes(x = distance, y = cumulated_count)) +
  geom_step() +
  theme_bw()
  
```


Via regression of a polynomial, we can approximate the empirical cumulative distribution of potentially re-used sites, at least for low ranges.


```{r fitting-definition}
# fitting_function <- function(d, parameters) {
#   A <- parameters[1]
#   b <- parameters[2]
#   k <- parameters[3]
#   s <- 1
#   result <- A * b^(k*d/s)-1.0
#   return(result)
# }

fitting_function <- function(d, parameters) {
  A <- parameters[1]
  k <- parameters[2]
  result <- A * d^k
  return(result)
}

fit_fitting <- function(x, y, ...) {
  target_fcn <- wrap_target_function
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, fitting_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- fitting_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    fitting_function,
    optimizer_results
  )


  return(optimizer_results)
}

print_regression_results <- function(orsl) {
  par <- paste(round(orsl$par, 4), collapse = ", ")
  conv <- orsl$convergence
  eps <- orsl$value

  print(
    sprintf("fit: conv %i at (%s), mse %.1f", conv, par, eps)
  )
}

x <- cumulative_crossdistcount$distance
y <- cumulative_crossdistcount$cumulated_count
cap <- 128.
y <- y[x<=cap]
x <- x[x<=cap]
fit_result <- fit_fitting(x, y,
  method = "Nelder-Mead", # "L-BFGS-B",
  par = c(0.25, 1.8),
  control = list("fnscale" = 1e-4)
)

print_regression_results(fit_result)

predict_n_sites <- function (val) round(fit_result$predict(val), 0)

y_fit <- fit_result$predict(x)

ggplot(NULL) +
  geom_step(aes(x = x, y = y), color = "black") +
  geom_step(aes(x = x, y = y_fit), color = "steelblue") +
  theme_bw()
```


With this, we can predict how many sites are potentially re-used:

```{r test-coupling}
#| code-fold: false
test <- 32 # m
n_sites <- predict_n_sites(test)
print(n_sites)
```

For example...

- If we set a coupling distance of `{r} 1` m, we can expect `{r} predict_n_sites(1)` locations within re-use range.
- If we set a coupling distance of `{r} 10` m, we can expect `{r} predict_n_sites(10)` locations within re-use range.
- If we set a coupling distance of `{r} test` m, we can expect `{r} predict_n_sites(test)` locations within re-use range.
