---
title: "Sampling Frame Evaluations"
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---

```{r libraries}
library("dplyr")
library("magrittr")
library("ggplot2")

# data is loaded from cache
source("./auxiliary_data_queries.R")
source("regression_tools.R") 
```

# scraping breadcrumbs

Using the example sampling drawn from the sampling frame by @FV.

- load relevant data (stored via POC) 
- get spatial coordinates for `scheme == "GW_03.3"`
- combine with existing watina locations
- cross-distance
- plot cumulative number of location pairs at a given distance


see also

- <https://inbo.github.io/watina/reference/selectlocs_xg3.html>
- <https://inbo.github.io/watina/reference/selectlocs_chem.html>


Suggestions @FV

- find `spsamples_gw_sf_panflpan5`?
  - not in `objects_panflpan5.RData`, though that has a `samplinglocations_sf` (4514x2, grts_address and geometry)
  -> is a `gpkg` files: `spsamples_gw_sf_panflpan5.rds` and `spsamples_gw_sf_panflpan5.gpkg`
- filter `hydr_class != "HC3"`


```{r referencecode-load-grts}
#| eval: false
grts_mh_n2khab <- rast(
  file.path(datapath, "binary/intermediate/grts_mh_n2khab.tif")
)
grts_mh_n2khab_index <- tibble(
  id = seq_len(ncell(grts_mh_n2khab)),
  grts_address = values(grts_mh_n2khab)[, 1]
) %>%
  filter(!is.na(grts_address))
```


```{r referencecode-load-panflpan5}
#| eval: false
sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
domain_grts_n2khab <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
```

# load

```{r load-gpkg}
datapath <- "/home/falk/Downloads/steekproef"

# samples <- sf::read_sf(file.path(datapath, "spsamples_gw_sf_panflpan5.gpkg"))
samples <- readRDS(file.path(datapath, "spsamples_gw_sf_panflpan5.rds"))
samples <- cbind(sf::st_drop_geometry(samples), sf::st_coordinates(samples))

samples %<>% dplyr::filter(hydr_class != "HC3")

knitr::kable(t(dplyr::sample_n(samples, 4)))
```


"only" 978 locations left.


# samples by themselves


```{r cross-difference}
# compute the difference of all elements of one vector to each other
self_difference <- function (vec) outer(X = vec, Y = vec, FUN = function(X, Y)  Y-X)

# Calculate the Euclidean distance of the x and y columns in a data frame.
Euclid <- function (x, y) sqrt(self_difference(x)^2 + self_difference(y)^2 )

# return the lower triangle of a matrix, unpacking it into a vector of unique values
extract_lower_triangle <- function (mat) mat[lower.tri(mat)]

# retrieve elements of a vector which are below a given threshold
sub_threshold <- function (vec, thresh) vec[vec < thresh]

# cross distance of sampling locations
self_distances <- Euclid(samples$X, samples$Y)
# print(extract_lower_triangle(self_distances[1:6, 1:6]))
close_distances <- sub_threshold(extract_lower_triangle(self_distances), 256)
ggplot(NULL, aes(x = close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  theme_bw()

```

(For reference, the diagonal of a 32x32 grid cell is `{r} round(sqrt(32^2+32^2), 1)` m.)


- Cross distances are quantized, totally consistent with the GRTS grid cell arrangement of 32m resolution.
- As expected, there are a number of co-located samples. These are the same raster addresses sampled for different habitat types or compartments, which might happen often on the priority locations with lower GRTS addresses, or due to monitoring network inter-dependencies.

This sample is the reference to be superimposed on the existing watina locations.


# existing sites

```{r load-watina}
watina_data <- join_auxiliary_cache(label = "metadata")

watina_cross_distances <- Euclid(watina_data$x, watina_data$y)
watina_close_distances <- sub_threshold(extract_lower_triangle(watina_cross_distances), 256)

ggplot(NULL, aes(x = watina_close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  scale_y_log10() +
  theme_bw()

```

The watina location cross-distances seem more evenly distributed.
Now the interesting part: how are the new samples positioned relative to the existing ones?


# samples and existing sites


```{r cross-distance}

cross_difference <- function (v1, v2) outer(X = v1, Y = v2, FUN = function(X, Y)  Y-X)
Euclid2 <- function (x1, y1, x2, y2) sqrt(cross_difference(x1, x2)^2 + cross_difference(y1, y2)^2 )

cross_distances <- Euclid2(
  x1 = watina_data$x,
  y1 = watina_data$y,
  x2 = samples$X,
  y2 = samples$Y
)

# print(nrow(cross_distances))
# print(ncol(cross_distances))
# print(cross_distances[1:6, 1:6])

close_distances <- sub_threshold(cross_distances, 256)
```


```{r cross-distance}
#| label: fig-cross-distance-count
#| fig-cap: "Observed distances between existing watina locations and sampling draws."

ggplot(NULL, aes(x = close_distances)) +
  geom_histogram(bins = 512, color = "black", fill = "lightgrey") +
  scale_x_continuous(name = "cross distance (m)", breaks = seq(0, 256, length.out = 17)) +
  theme_bw()

```


For testing any hypothetical coupling rule, we can use the cumulated empirical distribution.

```{r cumulated-distances}
#| label: fig-cumulated-count
#| fig-cap: "Cumulated count of observed distances between existing watina locations and sampling draws."

count_threshold <- function (thresh){
  n <- length(sub_threshold(close_distances, thresh))
  return(tibble(distance = thresh, cumulated_count = n))
}

cumulative_crossdistcount <- bind_rows(
  lapply(
    seq(0, 256, length.out = 257),
    FUN = count_threshold)
)

cumulative_crossdistcount %>%
  ggplot(aes(x = distance, y = cumulated_count)) +
  geom_step() +
  theme_bw()
  
```


# cross-distance regression

Via regression with a polynomial, we can approximate the empirical cumulative distribution of potentially re-used sites, at least for close distances.
A working regression model to map the number of re-used sites `r` to the distance threshold `d` is:

$$r = A\cdot d^{k}$$

... with scale `A` and exponent `k`.


```{r fitting-definition}
# fitting_function <- function(d, parameters) {
#   A <- parameters[1]
#   b <- parameters[2]
#   k <- parameters[3]
#   s <- 1
#   result <- A * b^(k*d/s)-1.0
#   return(result)
# }

fitting_function <- function(d, parameters) {
  A <- parameters[1]
  k <- parameters[2]
  return(A * d^k)
}

invert_fitting_function <- function(params){
  return(function(val) (val/params[1])^(1/params[2]))
}

fit_fitting <- function(x, y, ...) {
  target_fcn <- wrap_target_function
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, fitting_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- fitting_function
  optimizer_results$ifcn <- invert_fitting_function(optimizer_results$par)
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    fitting_function,
    optimizer_results
  )

  return(optimizer_results)
}

print_regression_results <- function(orsl) {
  par <- round(orsl$par, 4)
  conv <- orsl$convergence
  eps <- orsl$value

  print(
    sprintf("fit: r = %.2f d^%.2f (conv %s, mse %.1f)", par[1], par[2], conv==0, eps)
  )
}

x <- cumulative_crossdistcount$distance
y <- cumulative_crossdistcount$cumulated_count
cap <- 128.
y <- y[x<=cap]
x <- x[x<=cap]
fit_result <- fit_fitting(x, y,
  method = "Nelder-Mead", # "L-BFGS-B",
  par = c(0.25, 1.8),
  control = list("fnscale" = 1e-4)
)

print_regression_results(fit_result)

predict_n_sites <- function (val) round(fit_result$predict(val), 0)
inverse_predict_distance <- function(val) (val/fit_result$par[1])^(1/fit_result$par[2])

y_fit <- fit_result$predict(x) 

ggplot(NULL) +
  geom_step(aes(x = x, y = y, color = "simulated sample re-use")) +
  geom_line(aes(x = x, y = y_fit, color = "regression with $y=a x^k$"),
    linetype = "dashed", linewidth = 1.5) +
  labs(x = "distance `d` (m)", y = "cumulated re-use `r`") +
  scale_color_manual(name = "cumulative location re-use", values = c("simulated sample re-use" = "black", "regression with $y=a x^k$" = "steelblue")) +
  theme_bw()
```


With this, we can predict how many sites are potentially re-used:

```{r test-coupling}
#| code-fold: false
test <- 32 # m
n_sites <- predict_n_sites(test)
print(n_sites)
```

:::{.callout-note}
For example...

- If we set a coupling distance of `{r} 1` m, we can expect `{r} predict_n_sites(1)` locations within re-use range.
- If we set a coupling distance of `{r} 10` m, we can expect `{r} predict_n_sites(10)` locations within re-use range.
- If we set a coupling distance of `{r} test` m, we can expect `{r} predict_n_sites(test)` locations within re-use range.

:::

To put this into perspective: the total number of sites in the sampling is `{r} nrow(samples)`, so 10 sites would be about `{r} round(100*10/nrow(samples), 1)` % of the sample.


The regression equation can be inverted to calculate the distance for a given number of accepted site re-use.


$$d = \left(\frac{r}{A}\right)^{\frac{1}{k}}$$


For example, the re-use of `{r} 10` sites corresponds to a distance threshold of approximately `{r} round(inverse_predict_distance(10), 0)` m.



# Bayesian uncertainty

Just a quick glimpse at how sure we are about these regression parameters.

```{r load-rstan}
library("rstan")


```

```{r model-definition}

stan_model_definition <- paste0(
" data {
  int n;
  vector[n] r; // re-use site observations
  vector[n] d; // observed distances
  }",
" parameters {
  real <lower = 0> A; // scale
  real <lower = 0> k; // exponent
  real<lower = 0> residual;
  }",
" model {
  A ~ normal(0.25, 0.1);
  k ~ normal(1.8, 0.2);

  residual~ cauchy( 0.0, 1.0); // model residual

  // 'likelihood' / 'posterior'
  r ~ normal(A*(d^k), residual); 
  }",
""
)

model <- rstan::stan_model(
  model_code = stan_model_definition,
  model_name = "polynomial model",
)
```

```{r model-sampling}
#| eval: true
model_fit <- rstan::sampling(
  model,
  list(
    n = length(x),
    d = x,
    r = y
    ),
  iter = 2^14,
  chains = 8,
  cores = 8
  )

# model_fit
model_results <- as.data.frame(summary(model_fit)$summary)
saveRDS(model_results, here::here("cache", "regression", "steekproefkader_bayesian.rds"))
knitr::kable(t(model_results))
```


```{r stan-trace-plot}
#| label: fig-traceplot
#| fig-cap: "Traceplot of the MCMC sampling procedure, using `rstan`."
traceplot(model_fit)
```

```{r stan-forest-plot}
#| label: fig-forestplot
#| fig-cap: "Forest plot of model parameter posteriors."
plot(model_fit,
  show_density = TRUE, ci_level = 0.95,
  fill_color = "grey", fill_alpha = 0.6
  )
```


The regression for this particular sampling simulation is "quite a good fit", thus leaves little flexibility for the regression parameters.


However, this was just one particular sampling.


# bootstrapping

We have seen the word "sampling" herein as referring to a simulated sample from the MNE sampling frame.
We have seen sampling in the context of MCMC sampling for probabilistic inference.
To fulfill confusion, we can also sample from the sample and could than MCMC sample model parameters for the sample's sample model ("bootstrapping").


With bootstrapping, we can simulate a repeated sampling procedure and estimate the flexibility that the dataset leaves for the calculation of regression parameters.
(Culprit: this is just re-sampling of this particular sample; if it co-incidentally had few re-use locations, those will remain few.)


## technical wrap-up

```{r bootstrapping-extras}
library("parallel") 
```

```{r regression_wrapper}

regression_wrapper <- function(boot_distances) {

  boot_threshold <- function (thresh){
    n <- length(sub_threshold(boot_distances, thresh))
    return(data.frame(distance = thresh, cumulated_count = n))
  }
  
  boot_cumcrossdistcount <- bind_rows(
    lapply(
      seq(0, 128, length.out = 129),
      FUN = boot_threshold)
  )


  fit_result <- fit_fitting(
    x = boot_cumcrossdistcount$distance,
    y = boot_cumcrossdistcount$cumulated_count,
    method = "Nelder-Mead", # "L-BFGS-B",
    par = c(0.25, 1.8)
    # lower = c(0.01, 1.0),
    # upper = c(1.0, 10.0),
    # control = list("fnscale" = 1e-6)
  )
  return(fit_result)
} 

```



```{r regression-reference}
even_closer_distances <- close_distances[close_distances <= 128]
reference <- regression_wrapper(even_closer_distances)
reference["A"] <- reference$par[1]
reference["k"] <- reference$par[2]
print_regression_results(reference)
```


```{r bootstrap}
bootstrap_function <- function(i){

  boot_distances <- sample(
    even_closer_distances,
    as.integer(length(even_closer_distances)),
    replace = TRUE)

  result <- regression_wrapper(boot_distances)
  return(as.data.frame(list(
      "i" = i,
      "conv" = result$convergence == 0,
      "rmse" = result$value,
      "A" = result$par[1],
      "k" = result$par[2]
    )))
}

n_iterations <- 2^12
straps <- bind_rows(mclapply(
  1:n_iterations,
  bootstrap_function,
  mc.cores = detectCores()-2
  ))

```


```{r plot-distribution}

plot_param <- function(param){
  h <- straps %>%
    mutate(param = straps[, param]) %>% 
    ggplot(aes(x = param)) +
    geom_histogram(bins = 256, width = 1.05,
                   color = NA, fill = "darkgray",
                   alpha = 1., position = "identity") +
    geom_vline(xintercept = reference[[param]]) +
    xlab(param) +
    theme_bw()

  ggsave(paste0("figures/spk_bootstrap_", param, ".png"), h, dpi = 300)
  return(h)
}
```

## parameter `A` (scale)

```{r plot-A-distribution}
plot_param("A")
knitr::kable(t(round(quantile(straps$A, probs = c(0.05, 0.5, 0.95)), 2)))
```

## parameter `k` (exponent)
```{r plot-k-distribution}
plot_param("k")
knitr::kable(t(round(quantile(straps$k, probs = c(0.05, 0.5, 0.95)), 2)))
```


## confidence interval

Via bootstrapping, we can add confidence ranges to the plot above:


```{r confidence-lines}
test_d <- seq(0, 128, length.out = 129)
calculate_rs <- function (i) {
  test_r <- fitting_function(test_d, c(straps[i, "A"], straps[i, "k"]))
  return(test_r)
}
test_calculations <- sapply(
  1:nrow(straps),
  FUN = calculate_rs
)
get_line <- function(perc){
  as.numeric(sapply(1:nrow(test_calculations),
    FUN = function (i) {
      quantile(test_calculations[i,], probs = c(perc))
      }
  ))
}
# get_line(0.05)
```



```{r plot-confidence}
#| label: fig-confidence
#| fig-cap: "Distance-rule-dependent site re-use. The higher we set the distance threshold for inclusion of existing sites (x), the more existing watina locations would potentially be re-used for the MNM sampling. Black line: actual sampling; blue dashed line: regression; grey: bootstrapping interval."

h <- ggplot(NULL) +
  geom_ribbon(aes(x = test_d, ymin = get_line(0.02), ymax = get_line(0.98)),
              color = "gray", alpha = 0.3) +
  geom_step(aes(x = x, y = y), color = "black") +
  geom_line(aes(x = x, y = y_fit), color = "steelblue",
    linetype = "dashed", linewidth = 1.5) +
  labs(x = "inclusion distance threshold (m)", y = "cumulative site re-use") +
  xlim(0, 32) +
  ylim(0, 1.05*fit_result$predict(32)) +
  theme_bw()
  
suppressWarnings(
  ggsave(paste0("figures/steekproefkader_sites_per_distance.png"), h, dpi = 300)
)

suppressWarnings(print(h))
```


Translated to specific examples:


```{r check1}
check_dist <- 5 # m
check_r <- round(quantile(test_calculations[check_dist,], probs = c(0.05, 0.5, 0.95)), 1)
knitr::kable(t(check_r))
```
At a distance of `{r} check_dist` m, we would expect between `{r} round(check_r[[1]])` and `{r} round(check_r[[3]])` sites to qualify for re-use.


```{r check2}
check_dist <- 8 # m
check_r <- round(quantile(test_calculations[check_dist,], probs = c(0.05, 0.5, 0.95)), 1)
knitr::kable(t(check_r))
```
At a distance of `{r} check_dist` m, we would expect between `{r} round(check_r[[1]])` and `{r} round(check_r[[3]])` sites to qualify for re-use.
