---
title: "Daily Water Level Variograms: Slope Filter Refinement"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

:::{.callout-warning}
**The effect of `slope` is conceptually relevant, so that a slope filter will be regarded as the default setting for other analysis components.**
:::




Settings refinned:

- fixed `extent` (data limit) to $500 m$
- fixed `maxx` (regression upper limit for $\sigma$) to $1000 m$
- fixed `bin_target` to $100$ observations per bin (can be less in case of the data-draining filters)



`mMaaiveld` repetition:

- results seem consistent (slope filter necessary), but *sill* is different
- `peat`:
  - looses sill almost completely; suddenly matches the others!
  - limited slope is probably inherent
  - volatile regression due to influence of first few bins (insufficient close range sample size)


Bootstrapping:

- bootstrap ranges are a few meters wide (-> reporting accuracy)
- thresholds are always bimodal for `mMaaiveld`


- TO RETHINK is slope-filter a form of soft de-trending? (side-effect of smoothing the data)


# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("units")     |> void() # someone decided to store water distance in meters...


# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
source("regression_tools.R") 
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
```


## parameters

```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

bin_target <- 100 # ... and the preferred alternative: observations per bin

column <- "dw"
cap_dw <- 4 # m # maximum accepted water level difference within a pair
```


## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx") %>%
  join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff,
         pp_vld, mp_vld
         )


diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
    validated1 = as.integer(pp_vld1 == "VLD") * as.integer(mp_vld1 == "VLD"),
    validated2 = as.integer(pp_vld2 == "VLD") * as.integer(mp_vld2 == "VLD"),
    validated = as.logical(validated1 * validated2),
    ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    dh = elevation_dhmv2-elevation_dhmv1,
    dz = fdepth2-fdepth1,
    damp = emd_ampsum2-emd_ampsum1,
    dqdiff = emd_qdiff2-emd_qdiff1,
    dwadi = wata_min_dist2-wata_min_dist1,
    dslope = slope_r2.5m2-slope_r2.5m1,
    melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
    mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
    mslope = abs(dh) / ds
  ) %>%
  filter(ds <= extent)

```



```{r count-all}
diff_joined <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

count_all <- nrow(diff_joined)
print(count_all)
```


<hr>

# All Data

## Filter

```{r all-no-filter}
somediff_meta <- diff_meta 

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r all-join-diffdate}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r all-groupwise-binning}

diff_equibinned <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw = mean(dw, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```



## variograms {#sec-vgrams}

using water level difference (MAD)


::: {.panel-tabset group="soilclass"}

### heavy

```{r all-regression-heavy}
#| label: fig-all-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "allin",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r all-regression-light}
#| label: fig-all-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "allin",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



### peat

```{r all-regression-peat}
#| label: fig-all-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "allin",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::



# Slope Filtered

- `mslope` < 0.01 m/m

:::{.callout-warning}
We now interpret slope pairwise: 
difference in elevation, normalized by distance.
:::


## Filter

```{r slope-distribution}
# slope_threshold <- Inf #0.05
slope_threshold <- 0.01/1 # 1 cm per meter

diff_meta %>%
  select(mslope) %>%
  mutate(filltered = mslope < slope_threshold) %>% 
  ggplot(aes(log10(mslope), fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r slope-filter}
somediff_meta <- diff_meta %>%
  filter(
    mslope < slope_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r slope-join-diffdata}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r slope-groupwise-binning}

diff_equibinned <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw = mean(dw, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


## variograms {#sec-slope}

using water level difference (MAD)


::: {.panel-tabset group="soilclass"}

### heavy

```{r slope-regression-heavy}
#| label: fig-slope-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "slopefilter",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r slope-regression-light}
#| label: fig-slope-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "slopefilter",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```





### peat

```{r slope-regression-peat}
#| label: fig-slope-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dw\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "slopefilter",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```


:::


# Relative Ground Level with `mMaaiveld`


In all previous analyses, we used `mTAW` ("Tweede Algemene Waterpassing", 1947-1968).
For comparison, the relative water level (below field) can be used by querying `mMaaiveld`.


## Filter

This information is stored in the column `dm`.
Slope will be filtered as above, or not; both is kind of interesting.

```{r filter-maaiveld-data}
column <- "dm"

slope_threshold <- 0.01/1 # 1 cm per meter # Inf #
    
somediff_meta <- diff_meta %>%
  filter(
    mslope < slope_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```



## Binning


```{r mM-join-diffdate}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(if_all(c(m1, m2, dm), ~ !is.na(.x))) %>% 
  filter(abs(dm) <= ds, abs(dm) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r mM-groupwise-binning}

diff_equibinned <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dm))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dm = 1/(n_obs)*sum(abs(dm), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dm = mean(dm, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```



## variograms {#sec-maaiveld}

using water level difference (MAD) of `mMaaiveld`


::: {.panel-tabset group="soilclass"}

### heavy

```{r mM-regression-heavy}
#| label: fig-mM-regression-heavy
#| fig-cap: "Matérn regression of water level distance $\\langle dm\\rangle$ (m) on 'heavy' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "heavy",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "maaiveld",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### light

```{r mM-regression-light}
#| label: fig-mM-regression-light
#| fig-cap: "Matérn regression of water level distance $\\langle dm\\rangle$ (m) on 'light' soil."
regression_by_soilclass(
  diff_equibinned,
  sc = "light",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "maaiveld",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



### peat

```{r mM-regression-peat}
#| label: fig-mM-regression-peat
#| fig-cap: "Matérn regression of water level distance $\\langle dm\\rangle$ (m) on 'peat'. Red points indicate measurements which were excluded from the regression."
regression_by_soilclass(
  diff_equibinned,
  sc = "peat",
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = "maaiveld",
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)
```



:::

