---
title: "Daily Water Level Variograms: Ecoregion"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Summary

With slope filter set, all measurements in place...

:::{.callout-warning}
this notebook will attempt to further filter the resulting variograms by ecoregion.
:::


Main observations:

- results are largely **consistent with previous data, confirming the 10m rule**.
  - slope > 1% was filtered
  - soilclasses were lumped
- `mMaaiveld` vs `mTAW`: 
  - some water-independent habitats (beide huivelzones, Kempen, MV overgang) **show slope-dependence** (ramp in mTAW, but not in mMaaiveld)
  - others have opposite indication, i.e. slightly better variograms with mTAW (kustduinen, riviervalleien), putatively due to **waterbody-dependence**


# Preparation

## libraries
... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("units")     |> void() # someone decided to store water distance in meters...


# if (getOption("browser") == "") options(browser = "librewolf") 

# data is loaded from cache
source("./auxiliary_data_queries.R")
source("regression_tools.R") 
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
#

# https://r-charts.com/colors/
ecoregion_colors <- c(
  "zuidoostelijke heuvelzone" = "darkorchid",
  "Kempen" = "sienna",
  "polders en de getijdenschelde" = "slategray",
  "kustduinen" = "burlywood",
  "Pleistocene riviervalleien" = "deepskyblue",
  "zuidwestelijke heuvelzone" = "coral",
  "Midden-Vlaamse overgangsgebieden" = "gold",
  "krijt-leemgebieden" = "honeydew",
  "cuesta's" = "palevioletred",
  "grindrivieren" = "darkseagreen",
  "westelijke interfluvia" = "hotpink"
)
```


## parameters

```{r parameters}
#| code-fold: false
extent <- 500 # m # data included
maxx <- 1000 # 768 # upper limit for regression

bin_target <- 100 # ... and the preferred alternative: observations per bin

column <- "dw"
cap_dw <- 4 # m # maximum accepted water level difference within a pair
```


## data

The script `100_download_and_plot_waterlevel_for_vgram.qmd` should be used to assemble the data set loaded here.


```{r load-data}

meta_data <- join_auxiliary_cache(label = "metadata") %>%
  join_auxiliary_cache(label = "elevation", index_column = "pp_idx") %>%
  join_auxiliary_cache("waterdistance", index_column = "pp_idx")
  # %>% join_auxiliary_cache("empiricalmodes", index_column = "pp_idx") 

meta_join <- meta_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         # emd_n, emd_amp1, emd_amp2, emd_ampsum,
         # emd_qdiff, emd_residual_qdiff,
         pp_vld, mp_vld
         )


diff_data <- join_auxiliary_cache(label = "diffdata") %>% 
  filter(if_all(c(w1, w2, dw), ~ !is.na(.x)))

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    meta_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    meta_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
    validated1 = as.integer(pp_vld1 == "VLD") * as.integer(mp_vld1 == "VLD"),
    validated2 = as.integer(pp_vld2 == "VLD") * as.integer(mp_vld2 == "VLD"),
    validated = as.logical(validated1 * validated2),
    ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    dh = elevation_dhmv2-elevation_dhmv1,
    dz = fdepth2-fdepth1,
    # damp = emd_ampsum2-emd_ampsum1,
    # dqdiff = emd_qdiff2-emd_qdiff1,
    dwadi = wata_min_dist2-wata_min_dist1,
    dslope = slope_r2.5m2-slope_r2.5m1,
    melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
    mslope_r2.5m = mean(c(slope_r2.5m2, slope_r2.5m1)),
    mslope = abs(dh) / ds
  ) %>%
  filter(ds <= extent)

```


## ecoregion

```{r prepare-ecoregion}
ecoregions <- n2khab::read_ecoregions() %>% 
  select(region_name) %>% 
  rename(ecoregion = region_name) %>% 
  mutate(ecoregion = stringr::str_replace(ecoregion, "Ecoregio van de ", ""))

# n2khab::read_ecoregions() %>% sf::st_drop_geometry() %>% distinct(region_name)
ecoregion_lookup <- bind_rows(
  diff_meta %>% select(pp_idx1, x1, y1) %>% rename(pp_idx = pp_idx1, x = x1, y = y1),
  diff_meta %>% select(pp_idx2, x2, y2) %>% rename(pp_idx = pp_idx2, x = x2, y = y2)
  ) %>%
  distinct() %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370) %>% 
  sf::st_join(ecoregions)

ecoregion_lookup %>% 
  count(ecoregion) %>% arrange(desc(n)) %>% knitr::kable()

missing_points <- ecoregion_lookup %>% filter(is.na(ecoregion))
# mapview::mapview(ecoregions) +
#   mapview::mapview(missing_points)

```

... guess we can drop these two.


```{r join-ecoregion}

diff_meta <- diff_meta %>%
  left_join(
    ecoregion_lookup,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one"
  )

diff_meta <- diff_meta %>%
  filter(ecoregion != "NA")
  
diff_meta %>% 
  count(ecoregion) %>% arrange(desc(n)) %>% knitr::kable()

```


## diff data

```{r count-all}
diff_joined <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      ecoregion,
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

count_all <- nrow(diff_joined)
print(count_all)
```


## regression by ecoregion

a useful function for the tasks below.

```{r}

# regression_data <- diff_equibinned
# ecoreg <- "Kempen"
regression_by_ecoregion <- function(
    regression_data,
    ecoreg, reg_var,
    maxx = extent, maxy = NULL,
    label = "",
    return_fit = FALSE,
    ...) {


  stopifnot("arrow" = require("arrow"),
            "dplyr" = require("dplyr"))

  diff_sc <- regression_data %>% filter(ecoregion == ecoreg)

  diff_sc <- diff_sc %>%
    filter(ds > 0, ds < maxx) %>%
    select(ds, cnt, !!reg_var)

  n_sc <- sum(diff_sc %>% pull(cnt))

  sink_path <- here::here("cache", "regression")
  sink_file <- glue::glue("{label}_{reg_var}_{ecoreg}.parquet")
  write_parquet(diff_sc, sink = here::here(sink_path, sink_file))

  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= maxx]
  x <- x[x <= maxx]

  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]

  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)


  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
  )

  if (return_fit) return(matern_fit)

  sink_file <- glue::glue("{label}_{reg_var}_{ecoreg}.rds")
  saveRDS(matern_fit, file = here::here(sink_path, sink_file))

  print_regression_results(
    matern_fit,
    label = glue::glue("regression {ecoreg}, N={n_sc}:")
  )

  ylabel <- glue::glue("mean water level difference ({ecoreg}, N={n_sc})")
  sc_color <- ecoregion_colors[ecoreg]
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"

  g <- plot_matern(matern_fit, y_label = ylabel, color = sc_color)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, colour = "red", alpha = 0.2)
  return(g)
}

```

<hr>

# Slope-Filtered Data

## Filter

```{r slope-distribution}
# slope_threshold <- Inf #0.05
slope_threshold <- 0.01/1 # 1 cm per meter

diff_meta %>%
  select(mslope) %>%
  mutate(filltered = mslope < slope_threshold) %>% 
  ggplot(aes(log10(mslope), fill = filltered, color = filltered)) +
  geom_histogram(bins = 256) +
  theme_bw()
    
```


```{r all-no-filter}
# somediff_meta <- diff_meta 
somediff_meta <- diff_meta %>%
  filter(
    mslope < slope_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```


## Binning


```{r all-join-diffdate}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      ecoregion,
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(abs(dw) <= ds, abs(dw) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r all-groupwise-binning}

diff_equibinned <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dw))),
    soilclass = first(soilclass1),
    ecoregion = first(ecoregion),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dw = 1/(n_obs)*sum(abs(dw), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(ecoregion) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dw = mean(dw, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, ecoregion)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)
```


## variograms `mTAW` {#sec-vgrams}

using water level difference (MAD)


::: {.panel-tabset group="ecoregion"}

### zuidoostelijke heuvelzone

```{r}
select_ecoregion <- "zuidoostelijke heuvelzone"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Kempen

```{r}
select_ecoregion <- "Kempen"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Polders en de Getijdenschelde

```{r}
select_ecoregion <- "polders en de getijdenschelde"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Kustduinen

```{r}
select_ecoregion <- "kustduinen"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Pleistocene riviervalleien

```{r}
select_ecoregion <- "Pleistocene riviervalleien"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### zuidwestelijke heuvelzone

```{r}
select_ecoregion <- "zuidwestelijke heuvelzone"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Midden-Vlaamse overgangsgebieden

```{r}
select_ecoregion <- "Midden-Vlaamse overgangsgebieden"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


:::



# Relative Ground Level with `mMaaiveld`


In all previous analyses, we used `mTAW` ("Tweede Algemene Waterpassing", 1947-1968).
For comparison, the relative water level (below field) can be used by querying `mMaaiveld`.


## Filter

This information is stored in the column `dm`.
Slope will be filtered as above, or not; both is kind of interesting.

```{r filter-maaiveld-data}
column <- "dm"

somediff_meta <- diff_meta %>%
  filter(
    mslope < slope_threshold
  )

paste0(
  "selected ", nrow(somediff_meta), " / ",
  nrow(diff_meta), " location pairs (",
  round(100*nrow(somediff_meta)/nrow(diff_meta)), "%)."
)
```



## Binning

```{r mM-join-diffdate}
diff_joined <- somediff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      ecoregion,
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>% 
  filter(if_all(c(m1, m2, dm), ~ !is.na(.x))) %>% 
  filter(abs(dm) <= ds, abs(dm) <= cap_dw) # remove outliers

# knitr::kable(t(head(diff_joined, 3)), decimals = 2)

paste0(
  "selected ", nrow(diff_joined), " / ",
  count_all, " observations (",
  round(100*nrow(diff_joined)/(count_all)), "%)."
)
```



```{r mM-groupwise-binning}

diff_equibinned <- diff_joined %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dm))),
    soilclass = first(soilclass1),
    ecoregion = first(ecoregion),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dm = 1/(n_obs)*sum(abs(dm), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(ecoregion) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  ungroup %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    dm = mean(dm, na.rm = TRUE), 
    n_obs = sum(n_obs),
    .by = c(distance_equibin, ecoregion)
  ) %>%
  ungroup %>% arrange(ds)

# glimpse(diff_equibinned)

```



## variograms `mMaaiveld` {#sec-maaiveld}


::: {.panel-tabset group="ecoregion"}

### zuidoostelijke heuvelzone

```{r}
select_ecoregion <- "zuidoostelijke heuvelzone"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Kempen

```{r}
select_ecoregion <- "Kempen"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Polders en de Getijdenschelde

```{r}
select_ecoregion <- "polders en de getijdenschelde"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Kustduinen

```{r}
select_ecoregion <- "kustduinen"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Pleistocene riviervalleien

```{r}
select_ecoregion <- "Pleistocene riviervalleien"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### zuidwestelijke heuvelzone

```{r}
select_ecoregion <- "zuidwestelijke heuvelzone"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```


### Midden-Vlaamse overgangsgebieden

```{r}
select_ecoregion <- "Midden-Vlaamse overgangsgebieden"
```

```{r}
regression_by_ecoregion(
  diff_equibinned,
  ecoreg = select_ecoregion,
  reg_var = column,
  maxx = maxx, maxy = Inf,
  label = select_ecoregion,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 480, 0.48, 1.01)
) + ylim(0, 1.6)

```



:::


ecoregions which are not shown had too few observations for valid variograms:

- krijt-leemgebieden
- cuesta’s
- grindrivieren
- westelijke interfluvia
