---
title: "Water Chemistry Data Download"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---


# Preparation 


ChemVarWID -> holds the ID of compounds; 

- `6` = $PO_4$, 
- `7` = $NO_3$,
- `8` = $NO_2$ (not used),
- `9` = $NH_4$



```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("mapview")   |> void() # map visualization
library("scatterplot3d") |> void() # simple 3d vis


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# option to work offline; requires previous data download
work_offline <- FALSE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
min_cluster_size <- 2 # minimal cluster member count; fewer memeber clusters are removed
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")
diffdata_export_file <- paste0("data/waterchemistry_diffdata_cl", cluster_dist, ".parquet")

# # the user may choose to reload raw data
# force_data_reassembly <- FALSE
# reload <- force_data_reassembly || !file.exists(data_export_file)
```


Oh, look! 
Location Pairs stay the same!
We can re-use the location data stored earlier! (`100_download_and_plot_waterlevel_for_vgram.qmd`)


```{r amplitude-map}
pp_data <- read_parquet(data_export_file)
```


# WAIT!

Actually, we will not replace any site based on a single chemical alone.
**We want to measure coupling of water chemistry in general!**
Or, at least some chemicals.
```{r chem-variables}
chemvar_selection <- c(6, 7, 9)

chemvars <- tbl(watina_dwh, "dimChemVar") %>%
  select(ChemVarWID, ChemVarCode) %>%
  filter(
    ChemVarWID %in% chemvar_selection
  ) %>%
  mutate(ChemVarCode = as.character(lower(ChemVarCode)))
knitr::kable(chemvars %>% collect)
```

Without further ado, let us try good old **PCA** (Principal Component Analysis).

```{r glimpse-chemicals}
#| eval: false
peek <- tbl(watina_dwh, "FactChemischeMeting")
glimpse(peek)
```


```{r pca-data}
all_chem <- tbl(watina_dwh, "FactChemischeMeting") %>%
  filter(
    MetingStatusCode == "VLD",
    ChemVarWID %in% c(6, 7, 9)
  ) %>%
  left_join(
    chemvars,
    by = join_by(ChemVarWID),
    relationship = "many-to-many",
    unmatched = "drop"
  ) %>% 
  select(PeilpuntWID, ChemVarCode, DatumWID, Meetwaarde) %>%
  rename(
    pp_idx = PeilpuntWID,
    chem = ChemVarCode,
    date = DatumWID,
    m = Meetwaarde
  ) %>%
  tidyr::pivot_wider(
    names_from = chem,
    values_from = m
  ) %>%
  rename_with(~stringr::str_remove(., '-') ) %>% 
  collect

```

```{r pca-data-colors}
sel <- colnames(all_chem)[3:5]

legend_colors <- c(
  "nnh4" = "darkblue",
  "nno3" = "darkred",
  "ppo4" = "darkorange"
  )
```


```{r pca-data-distributions}
#| label: fig-pca-data-distributions
#| fig-cap: "Distributions of log concentrations of groundwater chemicals."
ggplot(NULL) +
  geom_density(
    aes(x = log10(all_chem %>% pull(!!sel[1])),
    color = sel[1])
  ) +
  geom_density(
    aes(x = log10(all_chem %>% pull(!!sel[2])),
    color = sel[2])
  ) +
  geom_density(
    aes(x = log10(all_chem %>% pull(!!sel[3])),
    color = sel[3])
  ) +
  labs(x = "log10 concentration") +
  scale_color_manual(values = legend_colors) + 
  theme_bw() + theme(legend.position = "right")
```


```{r pca-data-3d}
#| label: fig-pca-data-scatter
#| fig-cap: "Log concentrations of groundwater chemicals, data space visualization."
scatterplot3d(log10(all_chem[, sel]),
  main = "log10 concentrations",
  xlab = paste0("log(c(", sel[1], "))", collapse = ""),
  ylab = paste0("log(c(", sel[2], "))", collapse = ""),
  zlab = paste0("log(c(", sel[3], "))", collapse = ""),
  pch = 16, color = "steelblue", angle = 30
  )

```

correlation

```{r pca-data-correlation}

# log, filter na/inf
std_chem <- all_chem %>%
  mutate(across(any_of(sel), function(vec) log10(vec))) %>%
  filter(if_all(everything(), ~!is.na(.))) %>% 
  filter(if_all(everything(), ~is.finite(.)))

# center
means <- colMeans(std_chem[, sel], na.rm = TRUE)
std_chem <- std_chem %>%
  mutate(across(any_of(sel), function (vec) vec - mean(vec, na.rm = TRUE)))

# scale
vars <- colMeans(abs(std_chem[, sel]), na.rm = TRUE)
std_chem <- std_chem %>%
  mutate(across(any_of(sel), function (vec) vec / mean(abs(vec), na.rm = TRUE)))

chemcov <- cov(std_chem[, sel], use = "complete.obs")
knitr::kable(chemcov, digits = 2)
```

```{r pca-data-3d}
#| label: fig-pca-data-std
#| fig-cap: "Standardized log concentrations of groundwater chemicals. These are data input to a PCA."
scatterplot3d(std_chem[, sel],
  main = "log10 concentrations",
  xlab = paste0("standardized log(c(", sel[1], "))", collapse = ""),
  ylab = paste0("standardized log(c(", sel[2], "))", collapse = ""),
  zlab = paste0("standardized log(c(", sel[3], "))", collapse = ""),
  pch = 16, color = "steelblue", angle = 30
  )

```


eigenanalysis
```{r}
s <- svd(chemcov, nu = 3)
pca.val <- s$d
pca.weight <- pca.val / sum(pca.val)
pca.weight_ <- paste0(round(100*pca.weight, 1), "%")
pca.mat <- s$u

datamat <- as.matrix(std_chem[, sel], wide = TRUE) 
pcs <- paste0("PC", c(1, 2, 3))
std_chem[, pcs] <- datamat %*% t(pca.mat)
```


```{r pca-data-3d}
#| label: fig-pca-data-pca
#| fig-cap: "PCA space of groundwater chemicals. The first axis will be used for differential data analysis (variograms)."
scatterplot3d(std_chem[, pcs],
  main = "log10 concentrations",
  xlab = paste0(pcs[1], " (", pca.weight_[1], ")"),
  ylab = paste0(pcs[2], " (", pca.weight_[2], ")"),
  zlab = paste0(pcs[3], " (", pca.weight_[3], ")"),
  pch = 16, color = "purple", angle = 20
  )

```


There is not all that much variability concentrated on PC1 (`{r} pca.weight_[1]`).

But, nevertheless, it is the most general "chemistry" measure we have.


# differential data

## pairing

Retrieving all pairs of observation wells
which are below a certain distance,

```{r pairing}
extent <- 512 # m

# pp_data %>% count(cluster)

pair_sites <- function(cluster_idx) {
  sub_data <- pp_data %>%
    filter(cluster == cluster_idx) %>%
    select(mp_idx, pp_idx, x, y, soilclass)
  
  sub_data %>% 
    left_join(sub_data,
      by = join_by(
        mp_idx > mp_idx, # only different locations
        soilclass == soilclass, # from the same soilclass
      ),
      suffix = c("1", "2"),
      relationship = "many-to-many"
    ) %>% mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    ) %>%
    filter(ds >= 1.0/100, ds < extent)

}

pairs <- bind_rows(
  lapply(
    sort(unique(as.integer(pp_data$cluster))),
    FUN = pair_sites
  )
) %>% arrange(ds)
```



```{r pair-distance-distribution}
#| label: fig-pairs-ds
#| fig-cap: "Distribution of pair distances."
pairs %>%
  ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()
```

It is okay to keep the close range sites for now.




## water chem query


query: pp1, pp2, date1, date2, Meetwaarde1, Meetwaarde2

```{r chemistry-query}

query_measurement <- function (pp1, pp2){

  w1 <- std_chem %>%
    filter(pp_idx == pp1,
      ) %>%
    select(pp_idx, date, PC1) %>%
    rename(
      pp_idx1 = pp_idx,
      date1 = date,
      m1 = PC1
    )

  w2 <- std_chem %>%
    filter(pp_idx == pp2,
      ) %>%
    select(pp_idx, date, PC1) %>%
    rename(
      pp_idx2 = pp_idx,
      date2 = date,
      m2 = PC1
    )

  pair_diff <- w1 %>%
    cross_join(w2) %>%
    mutate(
      across(c(date1, date2), function(x) as.Date(as.character(x), format = "%Y%m%d")),
      dt = date2 - date1,
      dc = m2 - m1
    ) %>%
    filter(abs(dt) < 30) %>% 
    select(pp_idx1, pp_idx2, date1, date2, dc)

  return(pair_diff)
}

pairs_test <- pairs
# pairs_test <- pairs_test %>%
#   head(100)
n_rows <- nrow(pairs_test)

pb <- txtProgressBar(min = 0, max = n_rows,
                       initial = 0, style = 1)
rowwise_query <- function(i) {
  setTxtProgressBar(pb, i)
  pp1 <- pairs[[i, "pp_idx1"]]
  pp2 <- pairs[[i, "pp_idx2"]]
  return(query_measurement(pp1, pp2))
}


diff_data <- bind_rows(
    lapply(
      1:n_rows,
      FUN = rowwise_query
    )
  )

close(pb)
write_parquet(diff_data, sink = diffdata_export_file)

```
