---
title: "Water Chemistry Data Download"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# Preparation 


ChemVarWID -> holds the ID of compounds; 

- `6` = $PO_4$, 
- `7` = $NO_3$,
- `8` = $NO_2$ (not used),
- `9` = $NH_4$


Concentrations will be transformed.

```{r logtrafo}
# log_trafo <- \(x) log10(1e-6 + x)
# # 20260106: only five values are at lower limit: one for NO3, four for PO4

log_trafo <- \(x) log10(x)

```




```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("mapview")   |> void() # map visualization
library("scatterplot3d") |> void() # simple 3d vis


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# option to work offline; requires previous data download
work_offline <- FALSE

# connect watina 
if (!work_offline && !exists("watina_dwh")) {
  watina_dwh <- connect_watina()
}

# create a data folder (if not present)
dir.create("./data", showWarnings = FALSE)

if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
t_interval <- "alltimes"
min_cluster_size <- 2 # minimal cluster member count; fewer memeber clusters are removed
data_export_file <- glue::glue("data/waterlevel_data_cl{cluster_dist}.parquet")
allchem_export_file <- glue::glue("data/waterchemistry_alldata.parquet")
chem_export_file <- glue::glue("data/waterchemistry_data_cl{cluster_dist}_{t_interval}d.parquet")
diffdata_export_file <- glue::glue("data/waterchemistry_diffdata_cl{cluster_dist}_{t_interval}d.parquet")

# # the user may choose to reload raw data
# force_data_reassembly <- FALSE
# reload <- force_data_reassembly || !file.exists(data_export_file)
```


Oh, look! 
Location Pairs stay the same!
We can re-use the location data stored earlier! (`100_download_and_plot_waterlevel_for_vgram.qmd`)


```{r amplitude-map}
pp_data <- read_parquet(data_export_file)
```


# Chemical Variables: Concentrations

Actually, we will not replace any site based on a single chemical alone.
**We want to measure coupling of water chemistry in general!**
Or, at least some chemicals.

```{r chem-variables}
chemvar_selection <- c(6, 7, 9)

chemvars <- tbl(watina_dwh, "dimChemVar") %>%
  select(ChemVarWID, ChemVarCode) %>%
  filter(
    ChemVarWID %in% chemvar_selection
  ) %>%
  mutate(ChemVarCode = as.character(lower(ChemVarCode)))

knitr::kable(chemvars %>% collect)
```


```{r data-colors}

legend_colors <- c(
  "nh4" = "darkblue",
  "no3" = "darkred",
  "po4" = "darkorange"
  )

color_of <- \(chem) legend_colors[[chem]]

```



```{r glimpse-chemicals}
#| eval: false
peek <- tbl(watina_dwh, "FactChemischeMeting")
glimpse(peek)
```

 StaalWID <int>, [...]
 Meetwaarde <dbl>,
 MeetwaardeMEQ <dbl>,
 MeetWaardeMMOL_L <dbl>,
 MetingStatusCode <chr>,
 MetingStatus <chr>,
 IsBelowLOQ <lgl>,
 IsInferred <lgl>,
 IsDeleted <lgl>,
 ChangeReason <chr>,
 InsertAuditKey <int>,
 UpdateAuditKey <int>



```{r pca-data}
all_chem <- tbl(watina_dwh, "FactChemischeMeting") %>%
  filter(
    # MetingStatusCode == "VLD",
    !IsDeleted,
    ChemVarWID %in% c(6, 7, 9)
  ) %>%
  left_join(
    chemvars,
    by = join_by(ChemVarWID),
    relationship = "many-to-many",
    unmatched = "drop"
  ) %>% 
  select(
    PeilpuntWID,
    StaalWID,
    ChemVarCode,
    DatumWID,
    MeetwaardeMEQ,
    MeetWaardeMMOL_L,
    MetingStatusCode,
    IsBelowLOQ,
    IsInferred
  ) %>%
  rename(
    pp_idx = PeilpuntWID,
    sample_idx = StaalWID,
    chem = ChemVarCode,
    date = DatumWID,
    m = MeetwaardeMEQ,
    mmol = MeetWaardeMMOL_L,
    vld = MetingStatusCode,
    loq = IsBelowLOQ,
    infer = IsInferred
  ) %>%
  collect

all_chem %>% 
  write_parquet(sink = allchem_export_file)

```

```{r}
nrow(all_chem)
max(all_chem$date)

```

with MetingStatusCode == "VLD",

 -  N = 8361
 -  max(date) = 20121203

taking ALL:

 -  N = 16630
 -  max(date) = 20251027
 
... numbers now differ because widening was postponed.

*a propos:*
 
# "Widening"! Are we still on track?

 
This is how data was previously "widened":

```{r}
#| eval: false
null_chem <- all_chem %>% 
  tidyr::pivot_wider(
    names_from = chem,
    values_from = m
  ) %>%
  rename_with(~stringr::str_remove(., '-') ) 
```

Rationale: chemical measurements are in theory captured on the same samples, at the same lab cycles.
The database has the exact info about which of those co-incide, by "StaalID" = `sample_idx`.


First of all, turns out there are duplicate measurements, differing in validation status and/or LOQ.
I will just discard these.

```{r}

double_chemss <- all_chem %>%
  count(pp_idx, sample_idx, date, chem) %>%
  filter(n > 1) %>%
  arrange(desc(n))

all_chem %>%
  semi_join(
    double_chemss,
    by = join_by(pp_idx, sample_idx, date, chem)
  ) %>%
  arrange(sample_idx, date, chem) %>% 
  knitr::kable()

uq_chem <- all_chem %>%
  anti_join(
    double_chemss,
    by = join_by(pp_idx, sample_idx, date, chem)
  ) 

```



```{r}

backbone <- uq_chem %>%
  distinct(pp_idx, sample_idx, date) %>%
  arrange(date, sample_idx)

chem_data <- backbone %>%
  left_join(
    uq_chem %>%
      filter(chem == "n-nh4") %>%
      select(-chem, -vld, -infer) %>% 
      rename(meq_nh4 = m, mmol_nh4 = mmol, loq_nh4 = loq),
    by = join_by(pp_idx, sample_idx, date)
  ) %>%
  left_join(
    uq_chem %>%
      filter(chem == "n-no3") %>%
      select(-chem, -vld, -infer) %>% 
      rename(meq_no3 = m, mmol_no3 = mmol, loq_no3 = loq),
    by = join_by(pp_idx, sample_idx, date)
  ) %>%
  left_join(
    uq_chem %>%
      filter(chem == "p-po4") %>%
      select(-chem, -vld, -infer) %>% 
      rename(meq_po4 = m, mmol_po4 = mmol, loq_po4 = loq),
    by = join_by(pp_idx, sample_idx, date)
  )

glimpse(chem_data)

```

TODO CONTINUE



```{r log-data-distributions}
#| label: fig-log-data-distributions
#| fig-cap: "Distributions of log concentrations of groundwater chemicals."
measure <- "meq" # "meq" # "mmol"
ggplot(NULL) +
  geom_density(
    aes(x = log_trafo(chem_data %>% pull(!!glue::glue("{measure}_nh4"))),
    color = "nh4")
  ) +
  geom_density(
    aes(x = log_trafo(chem_data %>% pull(!!glue::glue("{measure}_no3"))),
    color = "no3")
  ) +
  geom_density(
    aes(x = log_trafo(chem_data %>% pull(!!glue::glue("{measure}_po4"))),
    color = "po4")
  ) +
  labs(x = glue::glue("log trafo(concentration|{measure})")) +
  scale_color_manual(values = legend_colors) + 
  theme_bw() + theme(legend.position = "right")
```

 ... Without further ado, let us try good old **PCA** (Principal Component Analysis).

```{r pca-data-3d}
#| label: fig-pca-data-scatter
#| fig-cap: "Log concentrations of groundwater chemicals, data space visualization."

# log, filter na/inf
value_columns <- c(
  paste("meq", names(legend_colors), sep = "_"),
  paste("mmol", names(legend_colors), sep = "_")
)

std_chem <- bind_cols(
  chem_data %>%
    mutate(across(any_of(value_columns), log_trafo)) %>% 
    rename_with(~stringr::str_c(., "_log"), !!value_columns),
  chem_data %>%
    select(!!value_columns) %>% 
    rename_with(~stringr::str_c(., "_raw"), everything())
  )

log_columns <- paste(value_columns, "log", sep = "_")
pca_chem <- std_chem %>%
  filter(if_all(log_columns, ~!is.na(.))) %>% 
  filter(if_all(log_columns, ~is.finite(.)))

sel <- paste(measure, names(legend_colors), "log", sep = "_")

scatterplot3d(
  pca_chem %>%
    select(!!sel),
  main = glue::glue("log trafo(concentrations|{measure})"),
  xlab = glue::glue("log trafo(c(NH₄)|meq)"),
  ylab = glue::glue("log trafo(c(NO₃)|meq)"),
  zlab = glue::glue("log trafo(c(PO₄)|meq)"),
  pch = 16, color = "steelblue", angle = 30
  )

```

correlation

```{r pca-data-correlation}
do_standardize <- FALSE

# center
means <- colMeans(std_chem[, sel], na.rm = TRUE)
if(do_standardize){
  std_chem <- std_chem %>%
    mutate(across(any_of(sel), function (vec) vec - mean(vec, na.rm = TRUE)))
}

# scale
vars <- colMeans(abs(std_chem[, sel]), na.rm = TRUE)
if(do_standardize){
  std_chem <- std_chem %>%
    mutate(across(any_of(sel), function (vec) vec / var(abs(vec), na.rm = TRUE)))
}

chemcov <- cov(pca_chem[, sel], use = "complete.obs")
knitr::kable(chemcov, digits = 2)
```

```{r pca-data-3d}
#| label: fig-pca-data-std
#| fig-cap: "Standardized log concentrations of groundwater chemicals. These are data input to a PCA."
scatterplot3d(std_chem[, sel],
  main = "log trafo(concentration)",
  xlab = paste0("log(c(", sel[1], "))", collapse = ""),
  ylab = paste0("log(c(", sel[2], "))", collapse = ""),
  zlab = paste0("log(c(", sel[3], "))", collapse = ""),
  pch = 16, color = "steelblue", angle = 30
  )

# count concentrations ≤ 0.0001
std_chem[, sel] %>% filter(meq_nh4_log <= -5)
std_chem[, sel] %>% filter(meq_no3_log <= -5)
std_chem[, sel] %>% filter(meq_po4_log <= -5)
```


```{r standardized-data-distributions}
#| label: fig-standardized-data-distributions
#| fig-cap: "Distributions of log concentrations of groundwater chemicals. Should be the same as above."

lograw <- "log"

ggplot(NULL) +
  geom_density(
    aes(x = std_chem %>% pull(!!glue::glue("{measure}_nh4_{lograw}")),
    color = "nh4")
  ) +
  geom_density(
    aes(x = std_chem %>% pull(!!glue::glue("{measure}_no3_{lograw}")),
    color = "no3")
  ) +
  geom_density(
    aes(x = std_chem %>% pull(!!glue::glue("{measure}_po4_{lograw}")),
    color = "po4")
  ) +
  labs(x = glue::glue("log trafo(concentration|{measure})")) +
  scale_color_manual(values = legend_colors) + 
  # xlim(0, 0.05) +  # raw
  theme_bw() + theme(legend.position = "right")

```



eigenanalysis:

```{r}
s <- svd(chemcov, nu = 3)
pca.val <- s$d
pca.weight <- pca.val / sum(pca.val)
pca.weight_ <- paste0(round(100*pca.weight, 1), "%")
pca.mat <- s$u

datamat <- as.matrix(std_chem[, sel], wide = TRUE) 
pcs <- paste0("PC", c(1, 2, 3))
std_chem[, pcs] <- datamat %*% t(pca.mat)
```


```{r pca-data-3d}
#| label: fig-pca-data-pca
#| fig-cap: "PCA space of groundwater chemicals. The first axis will be used for differential data analysis (variograms)."
scatterplot3d(std_chem[, pcs],
  main = "PCA of log trafo(concentrations)",
  xlab = paste0(pcs[1], " (", pca.weight_[1], ")"),
  ylab = paste0(pcs[2], " (", pca.weight_[2], ")"),
  zlab = paste0(pcs[3], " (", pca.weight_[3], ")"),
  pch = 16, color = "purple", angle = 20
  )

```


There is not all that much variability concentrated on PC1 (`{r} pca.weight_[1]`).

But, nevertheless, it is the most general "chemistry" measure we have.


```{r re-store-data}
std_chem %>% 
  write_parquet(sink = chem_export_file)


std_chem <- read_parquet(chem_export_file)
```

# differential data

## pairing

Retrieving all pairs of observation wells
which are below a certain distance,

```{r pairing}
extent <- 512 # m

# pp_data %>% count(cluster)

pair_sites <- function(cluster_idx) {
  sub_data <- pp_data %>%
    filter(cluster == cluster_idx) %>%
    select(mp_idx, pp_idx, x, y, soilclass)
  
  pair_data <- sub_data %>% 
    left_join(sub_data,
      by = join_by(
        mp_idx <= mp_idx, # only different locations
        soilclass == soilclass, # from the same soilclass
      ),
      suffix = c("1", "2"),
      relationship = "many-to-many"
    ) %>% mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
    ) %>%
    # filter((pp_idx1 == 8064) | (pp_idx1 == 8070)) %>% 
    # filter((pp_idx2 == 8064) | (pp_idx2 == 8070)) %>% 
    filter(ds >= 1.0/100, ds < extent) 

  return(pair_data)
}

pairs <- bind_rows(
  lapply(
    sort(unique(as.integer(pp_data$cluster))),
    FUN = pair_sites
  )
) %>% arrange(ds)
```



```{r pair-distance-distribution}
#| label: fig-pairs-ds
#| fig-cap: "Distribution of pair distances."
pairs %>%
  ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()
```

It is okay to keep the close range sites for now.




## water chem query


query: pp1, pp2, date1, date2, Meetwaarde1, Meetwaarde2

```{r chemistry-query}
# std_chem %>% count(pp_idx) %>% arrange(desc(n))

query_columns <- c(
  "pp_idx", "date",
  paste(value_columns, "log", sep = "_"),
  paste(value_columns, "raw", sep = "_"),
  "PC1", "PC2", "PC3"
)

query_measurement <- function (pp1, pp2){

  # pp1 <- 8064
  # pp2 <- 8070

  w1 <- std_chem %>%
    filter(pp_idx == pp1) %>%
    select(!!!rlang::syms(query_columns))
    # %>% arrange(date) %>% head(4)

  w2 <- std_chem %>%
    filter(pp_idx == pp2) %>%
    select(!!!rlang::syms(query_columns))
    
  pair_diff <- w1 %>%
    cross_join(
      w2,
      suffix = c("1", "2")
    ) %>%
    mutate(
      across(
        c(date1, date2),
        function(x) as.Date(as.character(x), format = "%Y%m%d")
      ),
      dt = as.numeric(date2 - date1, "days")) %>% 
    filter(dt >= 0) %>% #, dt < max_measurement_interval) %>%
    mutate(
      dnh4 = meq_nh4_log2 - meq_nh4_log1,
      dno3 = meq_no3_log2 - meq_no3_log1,
      dpo4 = meq_po4_log2 - meq_po4_log1,
      dnh4r = meq_nh4_raw2 - meq_nh4_raw1,
      dno3r = meq_no3_raw2 - meq_no3_raw1,
      dpo4r = meq_po4_raw2 - meq_po4_raw1,
      dPC1  = PC12  - PC11,
      dPC2  = PC22  - PC21,
      dPC3  = PC32  - PC31,
      is_same_location = pp_idx1 == pp_idx2
    ) %>%
    select(
      pp_idx1, date1, pp_idx2, date2, dt,
      date1, date2,
      meq_nh4_log1, meq_nh4_log2,
      meq_no3_log1, meq_no3_log2,
      meq_po4_log1, meq_po4_log2,
      dnh4,  dno3,  dpo4,
      dnh4r, dno3r, dpo4r,
      dPC1, dPC2, dPC3,
      is_same_location
    )

  # pair_diff %>% distinct(date1, date2)

  # find the one with least dt...
  get_first_diff_per_chem <- function(dchem) {
    pair_diff %>% 
      filter(!is.na(!!!rlang::syms(dchem))) %>% 
      # select(!!dchem)
      arrange(dt) %>% 
      group_by(date1) %>% 
      reframe(head(pick(everything()), 1)) %>% 
      arrange(dt) %>% 
      group_by(date2) %>% 
      reframe(head(pick(everything()), 1)) %>% 
      select(pp_idx1, date1, pp_idx2, date2) %>%
      mutate(is_first = TRUE) %>%
      return()
  }
  # get_first_diff_per_chem("dnh4")
  # get_first_diff_per_chem(dchem = "dno3")
  # knitr::kable(pair_diff %>% select(pp_idx1, date1, dt, date2, pp_idx2))

  # ... and flag the first difference
  pair_diff <- pair_diff %>%
    mutate(is_first = FALSE) %>% 
    left_join(
      get_first_diff_per_chem("dnh4"),
      by = join_by(pp_idx1, date1, pp_idx2, date2),
      suffix = c("", "_dnh4")
    ) %>%
    mutate(is_first = FALSE) %>% 
    left_join(
      get_first_diff_per_chem("dno3"),
      by = join_by(pp_idx1, date1, pp_idx2, date2),
      suffix = c("", "_dno3")
    ) %>%
    mutate(is_first = FALSE) %>% 
    left_join(
      get_first_diff_per_chem("dpo4"),
      by = join_by(pp_idx1, date1, pp_idx2, date2),
      suffix = c("", "_dpo4")
    ) %>%
    select(-is_first) %>% 
    mutate_at(
      vars(is_first_dnh4, is_first_dno3, is_first_dpo4),
      ~tidyr::replace_na(., FALSE)
    )
  # pair_diff %>%
  #   distinct(is_first_dnh4, is_first_dno3, is_first_dpo4)
  
  # pair_diff %>%
  #   filter(is_first_dno3) %>%
  #   knitr::kable()

  return(pair_diff)
}

# pp_data %>% filter(pp_idx == 8064) %>%
#   select(mp_idx, pp_idx, pp_code, cluster, soilclass)

pairs_test <- pairs
# pairs_test <- pairs_test %>% head(100)
# pairs %>%
#   filter(pp_idx1 >= 8064, pp_idx1 <= 8070) %>% 
#   filter(pp_idx2 >= 8064, pp_idx2 <= 8070)  %>%
#   arrange(pp_idx2)
n_rows <- nrow(pairs_test)

pb <- txtProgressBar(min = 0, max = n_rows,
                       initial = 0, style = 1)
rowwise_query <- function(i) {
  setTxtProgressBar(pb, i)
  pp1 <- pairs[[i, "pp_idx1"]]
  pp2 <- pairs[[i, "pp_idx2"]]
  return(query_measurement(pp1, pp2))
}


diff_data <- bind_rows(
    lapply(
      1:n_rows,
      FUN = rowwise_query
    )
  )

close(pb)
write_parquet(diff_data, sink = diffdata_export_file)

```

# distribution of time differences

```{r time-differences}
diff_data %>%
  ggplot2::ggplot(aes(x = log2(1+as.integer(dt)))) +
  geom_histogram(bins = 257)
```

# Important Changes! (20260112)

- units: `c("meq_", "mmol_")` (practically identical)
- pca of MEQ values
- did not remove `NA` values -> also measurements with only a single chemical
- not only "first differences" are stored, also intervals
  -> use `is_first_d{chem}`
- also comparison of a site to itself (for temporal analysis)
  -> use `!is_same_location` to remove those
