---
title: "Water Level Variograms: Detection Limit"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---

# summary (tl;dr)

In this notebook, I briefly inspect the chemical data which is available in `WATINA`.

Some observations:

- Detection Limit is the lowest concentration which our facilities can measure. That changed over time. Rumors have it that, in some periods, chemical analysis outcomes below limit were fixed to *half the detection limit.*
- Since 2015, a database flag becomes active to filter these (`IsBelowLOQ`). However, the fact that they are below limit is a valid observation, which should not be discarded thoughtlessly.
- Validation status was ignored below.
- Units of ["Milli-Equivalents"](https://en.wikipedia.org/wiki/Equivalent_concentration) were arbitrarily chosen; database values seem to be consistent.
- Empirically, 1997, 2008, and 2015 seem to be the years in which our lab inventory changed, thought there is some indication for more frequent discontinuities.
- For **NO₃**, there are rather few measurements at the lower limit since 2008.
- In contrast, for **PO₄**, more than 50% of all measurements since 2008 scratch the lower detection limit (seemed less, before).
- For **NH₄**, critical measurements seem to have been ramping up since the nineties.
- Detection limits before 1997 were not investigated, because there were much fewer measurements in that period.



# preparation

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("tidyr")     |> void() # some pivoting of tables
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox


if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- glue::glue("data/waterlevel_data_cl{cluster_dist}.parquet")
allchem_export_file <- glue::glue("data/waterchemistry_alldata.parquet")

```


The script `120_waterchemistry_reload.qmd` must be run to assemble the data set loaded here.


```{r reload-data}

chem_data <- read_parquet(allchem_export_file) %>%
  rename(meq = m) %>%
  mutate(
    year = floor(date/10000),
    yearmonth = floor(date/100)
  ) %>%
  arrange(date) %>%
  left_join(
    read_parquet(data_export_file) %>%
      select(pp_idx, soilclass),
    by = join_by(pp_idx)
  )



```

# Data Exploration

    ... better late than never


## Inference Tag


None are IsInferred <lgl>

```{r}
chem_data %>% count(infer)
```

## Below LOQ

Some observations are flagged IsBelowLOQ <lgl>

```{r}
chem_data %>% count(loq)

chem_data %>%
  count(year, loq) %>%
  pivot_wider(
    names_from = loq,
    values_from = n
  ) %>%
  knitr::kable()
```


something changed in 2014.


## Validation, *quo vadis*?

```{r}
chem_data %>%
  count(year, vld) %>%
  pivot_wider(
    names_from = vld,
    values_from = n
  ) %>%
  knitr::kable()
  
```

The codes changed in 2013.

## Units

MEQ and millimol are practically identical.

```{r}
chem_data %>%
  mutate(check_units = meq == mmol) %>%
  count(year, check_units) %>%
  pivot_wider(
    names_from = check_units,
    values_from = n
  ) %>%
  knitr::kable()

```

## Negatives

There are even six negative MEQs.


```{r}

chem_data %>%
  count(meq < 0) %>% 
  knitr::kable()

```


# counting

```{r}

t <- chem_data$date
years <- chem_data %>%
  distinct(year) %>% arrange()

year_breaks <- seq(min(years) - 0.5, max(years) + 0.5, length.out = max(years)-min(years)+1)
years_with_chemical_measurements <- chem_data %>% pull(year)
hist(years_with_chemical_measurements, breaks = year_breaks)

```


# peak finding

```{r find-peak-function}

c_trafo <- \(x) log10(x)
c_retrafo <- \(x) 10^x


find_concentration_peaks <- function(log_concentration, n_peaks = 10, label = "") {
  step <- 0.0001
  breaks <- seq(
    floor(min(log_concentration)) - step/2,
    ceiling(max(log_concentration)) + step/2,
    step
  )# 
  
  h <- hist(log_concentration, breaks = breaks, plot = FALSE)
  # names(h)
  mids <- h$mids
  counts <- h$counts
  stepped_counts <- c(h$counts, h$counts[length(h$counts)])
  
  
  # find the peaks
  peaks <- order(counts, decreasing = TRUE)[1:n_peaks]
  x_peak <- mids[peaks]
  y_peak <- counts[peaks]
  
  g <- ggplot2::ggplot(NULL) +
    ggplot2::geom_step(aes(x = c_retrafo(breaks), y = stepped_counts)) +
    ggplot2::geom_point(aes(x = c_retrafo(x_peak), y = y_peak), color = "red") +
    ggplot2::xlim(0, 0.02) +
    ggplot2::geom_text(
      aes(x = c_retrafo(x_peak), y = y_peak, label = round(c_retrafo(x_peak), 4)),
      hjust = 1, vjust = 0
    ) +
    ggplot2::labs(x = glue::glue("observed {label} concentrations"), y = "observation count")

  ggsave(glue::glue("figures/peaks_{label}.png"), g, width = 16, height = 9, dpi = 300)
  print(g)
  
  return(c_retrafo(x_peak))
}

```


```{r}

one_var <- \(var) chem_data %>%
  # filter(is.na(loq) | !loq) %>% # optionally filter those under detection limit
  filter(chem == var) %>%
  select(pp_idx, year, yearmonth, date, loq, meq, mmol, soilclass) %>%
  filter(is.finite(meq), is.finite(mmol))

nonna <- \(x) x[!is.na(x)]
get_finite <- \(x) x[is.finite(x)]

pull_var <- \(var) one_var(var) %>%
  filter(is.na(loq) | !loq) %>% # optionally filter those under detection limit
  pull(meq) 

log_c_nh4 <- get_finite(nonna(c_trafo(pull_var("n-nh4"))))
log_c_no3 <- get_finite(nonna(c_trafo(pull_var("n-no3"))))
log_c_po4 <- get_finite(nonna(c_trafo(pull_var("p-po4"))))

nh4_peaks <- find_concentration_peaks(log_c_nh4, label = "NH₄")
no3_peaks <- find_concentration_peaks(log_c_no3, label = "NO₃")
po4_peaks <- find_concentration_peaks(log_c_po4, label = "PO₄")

nh4_peaks
no3_peaks
po4_peaks
```

NH4
including `loq`s:  [1] 0.0014276 0.0071384 0.0028550 0.0027836 0.0003569 0.0035686 0.0142758 ...
excluding `loq`s:  [1] 0.0071384 0.0028550 0.0027836 0.0003569 0.0035686 0.0142758 ...
-> 0.0014276 is the latest detection limit

NO3
including `loq`s:  [1] 0.00356862 0.00071384 0.00178443 0.00007138 0.00035686 0.00142758 ...
excluding `loq`s:  [1] 0.00356862 0.00178443 0.00007138 0.00035686 0.00142758 ...
-> 0.00071384 is the latest detection limit

PO4
including `loq`s:  [1] 0.0009687 0.0006458 0.0010332 0.0003229 0.0012915 0.0032292 0.001937  
excluding `loq`s:  [1] 0.0006458 0.0003229 0.0010332 0.0009687 0.0012915 0.0032292 0.0019373
-> 0.0009687 is the latest detection limit


# yearly changes

follow peaks over the years

```{r}

rate_between <- \(vals, lo, hi) sum((vals >= lo) & (vals <= hi)) / length(vals)
# vals <- c_retrafo(log_c_nh4)
# lo <- 0.00140
# hi <- 0.00145
# rate_between(c_retrafo(log_c_nh4), 0.00135, 0.00145)

data_nh4 <- one_var("n-nh4") %>%
  filter(meq > 0)
data_no3 <- one_var("n-no3") %>%
  filter(meq > 0)
data_po4 <- one_var("p-po4") %>%
  filter(meq > 0)

```


# NH₄

checking NH4 rates of (0.0014276), 0.0071384, (0.0028550 0.0027836)

```{r}
nh4_peaks

result_rates <- data_nh4 %>%
  mutate_at(vars(year, yearmonth), as.character) %>% 
  # filter(is.na(loq) | !loq) %>% 
  group_by(year) %>% 
  group_modify(~ {
    .x %>%
      summarize(
        r_latest = rate_between(meq, 0.00140, 0.00145),
        r_peak1 = rate_between(meq, 0.00690, 0.0074),
        r_peak2 = rate_between(meq, 0.00275, 0.0029)
      )
    }
  ) %>%
  arrange(year)

result_rates <- bind_rows(
  result_rates,
  result_rates[nrow(result_rates),]
)

g <- result_rates %>% 
  ggplot2::ggplot(aes(x = as.integer(year))) +
  ggplot2::geom_step(aes(y = r_latest), color = "darkred") +
  ggplot2::annotate("text", x = 2020, y = 0.5, label = "N[c == 0.0014]/N", color = "darkred") +
  ggplot2::geom_step(aes(y = r_peak1), color = "skyblue") +
  ggplot2::annotate("text", x = 2004, y = 0.32, label = "N[c == 0.007]/N", color = "skyblue") +
  ggplot2::geom_step(aes(y = r_peak2), color = "orange") +
  ggplot2::annotate("text", x = 2012, y = 0.45, label = "N[c == 0.0028]/N", color = "orange") +
  ggplot2::geom_vline(aes(xintercept = 1997.5)) +
  ggplot2::geom_vline(aes(xintercept = 2008.5)) +
  ggplot2::geom_vline(aes(xintercept = 2015.5)) +
  ggplot2::labs(x = "year/month", y = "observation probability of a given NH₄ concentration")

ggsave("figures/rates_nh4.png", g, width = 16, height = 9, dpi = 300)
print(g)
```


# NO₃

checking NO3 rates of 0.00356862 (0.00071384) 0.00178443 

```{r}
no3_peaks

result_rates <- data_no3 %>%
  mutate_at(vars(year, yearmonth), as.character) %>% 
  # filter(is.na(loq) | !loq) %>% 
  group_by(year) %>% 
  group_modify(~ {
    .x %>%
      summarize(
        r_latest = rate_between(meq, 0.00070, 0.00073),
        r_peak1 = rate_between(meq, 0.0035, 0.0036),
        r_peak2 = rate_between(meq, 0.00177, 0.00179)
      )
    }
  ) %>%
  arrange(year)

result_rates <- bind_rows(
  result_rates,
  result_rates[nrow(result_rates),]
)

g <- result_rates %>% 
  ggplot2::ggplot(aes(x = as.integer(year))) +
  ggplot2::geom_step(aes(y = r_latest), color = "darkred") +
  ggplot2::annotate("text", x = 2020, y = 0.2, label = "N[c == 0.00071]/N", color = "darkred") +
  ggplot2::geom_step(aes(y = r_peak1), color = "skyblue") +
  ggplot2::annotate("text", x = 2004, y = 0.2, label = "N[c == 0.0035]/N", color = "skyblue") +
  ggplot2::geom_step(aes(y = r_peak2), color = "orange") +
  ggplot2::annotate("text", x = 1994, y = 0.4, label = "N[c == 0.00178]/N", color = "orange") +
  ggplot2::geom_vline(aes(xintercept = 1997.5)) +
  ggplot2::geom_vline(aes(xintercept = 2008.5)) +
  ggplot2::geom_vline(aes(xintercept = 2015.5)) +
  ggplot2::labs(x = "year/month", y = "observation probability of a given NO₃ concentration")

ggsave("figures/rates_no3.png", g, width = 16, height = 9, dpi = 300)
print(g)
```


# PO₄

checking PO4 rates of (0.0009687) 0.0006458 0.0010332 

```{r}
po4_peaks

result_rates <- data_po4 %>%
  mutate_at(vars(year, yearmonth), as.character) %>% 
  filter(year > 1982) %>% 
  # filter(is.na(loq) | !loq) %>% 
  group_by(year) %>% 
  group_modify(~ {
    .x %>%
      summarize(
        r_latest = rate_between(meq, 0.0009, 0.0010),
        r_peak1 = rate_between(meq, 0.0010, 0.00110),
        r_peak2 = rate_between(meq, 0.00060, 0.00069)
      )
    }
  ) %>%
  arrange(year)

result_rates <- bind_rows(
  result_rates,
  result_rates[nrow(result_rates),]
)

g <- result_rates %>% 
  ggplot2::ggplot(aes(x = as.integer(year))) +
  ggplot2::geom_step(aes(y = r_latest), color = "darkred") +
  ggplot2::annotate("text", x = 2020, y = 0.9, label = "N[c == 0.0014]/N", color = "darkred") +
  ggplot2::geom_step(aes(y = r_peak1), color = "skyblue") +
  ggplot2::annotate("text", x = 2012, y = 0.78, label = "N[c == 0.0010]/N", color = "skyblue") +
  ggplot2::geom_step(aes(y = r_peak2), color = "orange") +
  ggplot2::annotate("text", x = 1998, y = 0.45, label = "N[c == 0.00065]/N", color = "orange") +
  ggplot2::geom_vline(aes(xintercept = 1990.5)) +
  ggplot2::geom_vline(aes(xintercept = 2008.5)) +
  ggplot2::geom_vline(aes(xintercept = 2015.5)) +
  ggplot2::labs(x = "year/month", y = "observation probability of a given PO₄ concentration")

ggsave("figures/rates_po4.png", g, width = 16, height = 9, dpi = 300)
print(g)
```




# PO₄ - heavy soil

Checking PO4 rates of "heavy" soil only: hypothesis that there might be relevant variation in detectionlimit rates across soilclasses.

```{r}
po4_peaks

result_rates <- data_po4 %>%
  filter(soilclass == "heavy") %>% 
  mutate_at(vars(year, yearmonth), as.character) %>% 
  filter(year > 1982) %>% 
  # filter(is.na(loq) | !loq) %>% 
  group_by(year) %>% 
  group_modify(~ {
    .x %>%
      summarize(
        r_latest = rate_between(meq, 0.0009, 0.0010),
        r_peak1 = rate_between(meq, 0.0010, 0.00110),
        r_peak2 = rate_between(meq, 0.00060, 0.00069)
      )
    }
  ) %>%
  arrange(year)

result_rates <- bind_rows(
  result_rates,
  result_rates[nrow(result_rates),]
)

g <- result_rates %>% 
  ggplot2::ggplot(aes(x = as.integer(year))) +
  ggplot2::geom_step(aes(y = r_latest), color = "darkred") +
  ggplot2::annotate("text", x = 2020, y = 0.9, label = "N[c == 0.0014]/N", color = "darkred") +
  ggplot2::geom_step(aes(y = r_peak1), color = "skyblue") +
  ggplot2::annotate("text", x = 2012, y = 0.78, label = "N[c == 0.0010]/N", color = "skyblue") +
  ggplot2::geom_step(aes(y = r_peak2), color = "orange") +
  ggplot2::annotate("text", x = 1998, y = 0.45, label = "N[c == 0.00065]/N", color = "orange") +
  ggplot2::geom_vline(aes(xintercept = 1990.5)) +
  ggplot2::geom_vline(aes(xintercept = 2008.5)) +
  ggplot2::geom_vline(aes(xintercept = 2015.5)) +
  ggplot2::labs(x = "year/month", y = "observation probability of a given PO₄ concentration")

ggsave("figures/rates_po4_heavy.png", g, width = 16, height = 9, dpi = 300)
print(g)
```


