---
title: "Water Level Variograms: Daily Data"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


# summary (tl;dr)

Variogram analysis of groundwater chemistry indicates close range spatial co-variation.
The analysis was carried out per soil type.
Visualization ([@sec-variograms]) and an overview table ([@sec-overview]) are available at the end of this document.


- Only same-day chemical measurements in the same soilclass were included (18211 observations).
- Once more, dynamic-width bins with homogeneous filling were defined.
- The difference measure is mean absolute difference of *standardized* log(concentration) or PC values. (standardized `= (value-mean)/sd` )
- Many simultaneous measurements happen at the same site, therefore observation counts are zero-inflated. Those well group measurements show unexpectedly large variation, complicating Matérn regression. Those were excluded by filtering for `dx > 2m` (approximatly one bin).
- raw concentrations were appended, too, for transparency.


# preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("ggcorrplot")|> void() # correlation, grid visualization
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges


if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- paste0("data/waterlevel_data_cl", cluster_dist, ".parquet")
chem_export_file <- paste0("data/waterchemistry_data_cl", cluster_dist, ".parquet")
diffdata_export_file <- paste0("data/waterchemistry_diffdata_cl", cluster_dist, ".parquet")

# parameters
extent <- 256 # m
n_bins <- 50
```


The script `120_waterchemistry_reload.qmd` should be used to assemble the data set loaded here.



```{r load-data}

pp_data <- read_parquet(data_export_file)

std_chem <- read_parquet(file = chem_export_file) %>%
  mutate(date = as.Date(as.character(date), format = "%Y%m%d"))
pp_chem <- std_chem %>%
  select(-date) %>% 
  summarize(
    across(everything(), ~mean(., na.rm = TRUE)),
    .by = pp_idx
  )
```


There are maps available (here: example for $NH_4$), yet measurement pairs are relatively scattered.

```{r map-data}

st_data <- pp_data %>%
  left_join(
    pp_chem,
    by = join_by(pp_idx),
    relationship = "one-to-one"
  ) %>% 
  filter(!is.na(nnh4_raw)) %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370)

mapview(st_data,
  map.types = c("CartoDB.Positron", "CartoDB.DarkMatter", "OpenStreetMap", "OpenTopoMap"),
  zcol = "nnh4_raw",
  cex = 3,
  legend = TRUE
)

```


```{r load-diffdata}

diff_data <- read_parquet(diffdata_export_file)

pp_join <- pp_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  select(pp_idx1, pp_idx2) %>%
  distinct() %>%
  left_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      damp = emd_ampsum2-emd_ampsum1,
      dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope = mean(c(slope_r2.5m2, slope_r2.5m1)) 
  )

```


```{r glimpse-data}

# diff_meta %>% glimpse
# diff_data %>% glimpse

diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, wata_dist_class1, wata_dist_class2,
      ds, dh, dz, damp,
      dqdiff, dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(paste0(
      as.character(pp_idx1), "_", as.character(pp_idx2), sep = "")
    )
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>%
  filter(
    # ds >= extent/n_bins
    date2 - date1 == 0
  ) %>%
  mutate(date = date1) %>% select(-date1, -date2)
  # mutate(dt = abs(difftime(date1, date2, units = "days")))
    

diff_all %>% ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()

```

# binning


After some good prior experience, I will once more apply binning with variable bin width, to achieve **equal sample size per bin**.




```{r equibin-preparation}
#| code-fold: false

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!is.na(dPC1))),
    soilclass = first(soilclass1),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    # dt = mean(abs(date2-date1), na.rm = TRUE),
    # semivar = 1/(2*n_obs)*sum(dPC1^2, na.rm = TRUE),
    dnnh4 = 1/(n_obs)*sum(abs(dnnh4), na.rm = TRUE),
    dnno3 = 1/(n_obs)*sum(abs(dnno3), na.rm = TRUE),
    dppo4 = 1/(n_obs)*sum(abs(dppo4), na.rm = TRUE),
    dnnh4r = 1/(n_obs)*sum(abs(dnnh4r), na.rm = TRUE),
    dnno3r = 1/(n_obs)*sum(abs(dnno3r), na.rm = TRUE),
    dppo4r = 1/(n_obs)*sum(abs(dppo4r), na.rm = TRUE),
    dPC1 = 1/(n_obs)*sum(abs(dPC1), na.rm = TRUE),
    dPC2 = 1/(n_obs)*sum(abs(dPC2), na.rm = TRUE),
    dPC3 = 1/(n_obs)*sum(abs(dPC3), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    # distance_equibin = cut(ds,
    #   breaks = seq(
    #     -0, extent,
    #     length.out = as.integer(extent/2)+1
    #   )
    # ) # cut to equal sized bins
    distance_equibin = cut_number(
      ds, as.integer(n_bins)
    ) # cut to equal sized bins
  ) %>%
  ungroup

    
combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)

diff_binned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    # dt = mean(dt), 
    n_obs = sum(n_obs),
    # semivariance_groupmean = mean(semivar, na.rm = TRUE),
    # semivariance_all = combine_grouped(semivar, cnt),
    dPC1_all = combine_grouped(dPC1, n_obs),
    dnnh4 = mean(dnnh4, na.rm = TRUE), 
    dnno3 = mean(dnno3, na.rm = TRUE), 
    dppo4 = mean(dppo4, na.rm = TRUE), 
    dnnh4r = mean(dnnh4r, na.rm = TRUE), 
    dnno3r = mean(dnno3r, na.rm = TRUE), 
    dppo4r = mean(dppo4r, na.rm = TRUE), 
    dPC1 = mean(dPC1, na.rm = TRUE), 
    dPC2 = mean(dPC2, na.rm = TRUE), 
    dPC3 = mean(dPC3, na.rm = TRUE), 
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>% arrange(ds)

```



```{r equibin-samplesize}
#| label: fig-equibin-sample-size
diff_binned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step() +
  geom_hline(yintercept = 0, color = "darkgrey") +
  theme_bw()
```


## Regression Procedure Wrapper

```{r regression-wrapper}

source("regression_tools.R") 


fit_matern <- function(x, y, distweighted = FALSE, ...) {
  target_fcn <- wrap_target_function
  if (distweighted) {
    target_fcn <- distweighted_target_function
  }
  optimizer_results <- optim(
    fn = function(params) {
      target_fcn(x, y, matern_function, params)
    },
    ...
  )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$fcn <- matern_function
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function,
    optimizer_results
  )


  return(optimizer_results)
}


plot_matern <- function(optimizer_results, y_label = "semivariance (m²)") {

  # retrieve everything 
  fcn <- optimizer_results$fcn
  regx <- optimizer_results$regx
  regy <- optimizer_results$regy
  predict <- optimizer_results$predict

  # extract parameters
  scale <- optimizer_results$par[1]
  range <- optimizer_results$par[2]
  nugget <- optimizer_results$par[3]
  sill <- scale + nugget


  # plotting
  plotx <- regx # seq(0, extent, length.out = 2*extent + 1)
  plotx <- plotx[plotx>0]
  
  g <- ggplot(NULL, aes(x = regx, y = regy)) +
    geom_vline(xintercept = range, color = "darkgrey") +
    geom_hline(yintercept = nugget, color = "darkgrey") +
    geom_hline(yintercept = sill, color = "darkgrey") +
    geom_point(size = 2.5, colour = "black", fill = "white", alpha = 0.2) +
    geom_line(aes(x = plotx, y = predict(plotx))) +
    labs(title = paste0("Matérn regression: ",
      paste(round(optimizer_results$par, 4), collapse = ", "))) +
    xlab("distance (m)") + ylab(y_label) +
    theme_minimal()

  return(g)
}

```


```{r regression-matérn-general}
# trafo <- function(x) log(1.+x) # feel free to try other transformations.
trafo <- function(x) x


list_of_variables <- c(
    "dnnh4" , "dnno3" , "dppo4",
    "dnnh4r" , "dnno3r" , "dppo4r",
    "dPC1" , "dPC2" , "dPC3"
)
list_of_soilclasses <- c("heavy", "light", "peat")
regression_storage <- matrix(rep(NA,
  length(list_of_variables) * length(list_of_soilclasses)
  ), length(list_of_variables), byrow = TRUE)
regression_storage <- as.data.frame(regression_storage)

colnames(regression_storage) <- list_of_soilclasses
rownames(regression_storage) <- list_of_variables


regression_by_soilclass <- function(
    regression_data, 
    sc, reg_var,
    maxx = extent, maxy = NULL, 
    ...) {
  diff_sc <- regression_data %>% filter(soilclass == sc)
  
  x <- diff_sc$ds
  y <- trafo(diff_sc %>% pull(!!reg_var))

  # DO NOT exclude the first two bins
  minx <- 1.0*extent/n_bins
  lowy <- y[x <= minx]
  lowx <- x[x <= minx]
  y <- y[x > minx]
  x <- x[x > minx]
  y <- y[x <= maxx]
  x <- x[x <= maxx]



  if (is.null(maxy)) maxy <- max(y)
  x <- x[y <= maxy]
  y <- y[y <= maxy]
  
  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

  # standardization helps function comprehension
  # x <- x/max(x)
  # y <- y / mean(y)
  
  
  matern_fit <- fit_matern(x, y,
    method = "L-BFGS-B", # "Nelder-Mead" # "L-BFGS-B"
    ...
  )

  # https://stackoverflow.com/a/71621059 
  regression_storage[reg_var, sc] <<- matern_fit$par[2]
  
  print_regression_results(
    matern_fit,
    label = "regression",
    indicate_threshold = FALSE
    )
  ylabel <- "mean difference (pair-averaged, then binned)"
  # if ("semivar" reg_var) ylabel <- "semivariance (non-pair-averaged)"
  g <- plot_matern(matern_fit, y_label = ylabel)

  x_excl <- diff_sc$ds
  y_excl <- trafo(diff_sc %>% pull(!!reg_var))

  y_excl <- y_excl[x_excl <= maxx]
  x_excl <- x_excl[x_excl <= maxx]
  x_excl <- x_excl[y_excl > maxy]
  y_excl <- y_excl[y_excl > maxy]
  g <- g + geom_point(aes(x = x_excl, y = y_excl),
      size = 2.5, colour = "red", alpha = 0.2) 
  g <- g + geom_point(aes(x = lowx, y = lowy),
      size = 2.5, colour = "red", alpha = 0.2) 
  return(g)
}


```


There are several options,
using mean-absolute difference 
in (standardized) chemical concentration 
or principal components.


I will cycle through those.

```{r dchem-column-overview}
#| code-fold: false

column <- "dnnh4"
column <- "dnno3"
column <- "dppo4"
# column <- "dPC1"
# column <- "dPC2"
# column <- "dPC3"
# 
# column <- "dPC1"


common_ylim <- 2
common_maxx <- 256
```


# variograms {#sec-variograms}

## PC1

```{r dchem-column-pc1}
#| code-fold: false
column <- "dPC1"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-pc1}
#| label: fig-dchem-regression-heavy-pc1
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-pc1}
#| label: fig-dchem-regression-light-pc1
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-pc1}
#| label: fig-dchem-regression-peat-pc1
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::

## PC2

```{r dchem-column-pc2}
#| code-fold: false
column <- "dPC2"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-pc2}
#| label: fig-dchem-regression-heavy-pc2
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-pc2}
#| label: fig-dchem-regression-light-pc2
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-pc2}
#| label: fig-dchem-regression-peat-pc2
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::

## PC3

```{r dchem-column-pc3}
#| code-fold: false
column <- "dPC3"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-pc3}
#| label: fig-dchem-regression-heavy-pc3
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-pc3}
#| label: fig-dchem-regression-light-pc3
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-pc3}
#| label: fig-dchem-regression-peat-pc3
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::

## $NH_4$

```{r dchem-column-nh4}
#| code-fold: false
column <- "dnnh4"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-nh4}
#| label: fig-dchem-regression-heavy-nh4
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-nh4}
#| label: fig-dchem-regression-light-nh4
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-nh4}
#| label: fig-dchem-regression-peat-nh4
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::

## $NO_3$

```{r dchem-column-no3}
#| code-fold: false
column <- "dnno3"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-no3}
#| label: fig-dchem-regression-heavy-no3
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-no3}
#| label: fig-dchem-regression-light-no3
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-no3}
#| label: fig-dchem-regression-peat-no3
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::

## $PO_4$

```{r dchem-column-po4}
#| code-fold: false
column <- "dppo4"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-po4}
#| label: fig-dchem-regression-heavy-po4
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-po4}
#| label: fig-dchem-regression-light-po4
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-po4}
#| label: fig-dchem-regression-peat-po4
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) + ylim(0, common_ylim)
```


:::



## $NH_4$ Raw

```{r dchem-column-nh4r}
#| code-fold: false
column <- "dnnh4r"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-nh4r}
#| label: fig-dchem-regression-heavy-nh4r
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-nh4r}
#| label: fig-dchem-regression-light-nh4r
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-nh4r}
#| label: fig-dchem-regression-peat-nh4r
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


:::

## $NO_3$ Raw

```{r dchem-column-no3r}
#| code-fold: false
column <- "dnno3r"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-no3r}
#| label: fig-dchem-regression-heavy-no3r
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-no3r}
#| label: fig-dchem-regression-light-no3r
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-no3r}
#| label: fig-dchem-regression-peat-no3r
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


:::

## $PO_4$ Raw

```{r dchem-column-po4r}
#| code-fold: false
column <- "dppo4r"
```

::: {.panel-tabset group="soilclass"}

### heavy

```{r dchem-regression-heavy-po4r}
#| label: fig-dchem-regression-heavy-po4r
regression_by_soilclass(
  diff_binned,
  "heavy",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)

```


### light

```{r dchem-regression-light-po4r}
#| label: fig-dchem-regression-light-po4r
regression_by_soilclass(
  diff_binned,
  sc = "light",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


### peat

```{r dchem-regression-peat-po4r}
#| label: fig-dchem-regression-peat-po4r
regression_by_soilclass(
  diff_binned,
  sc = "peat",
  reg_var = column,
  maxx = common_maxx, maxy = Inf,
  distweighted = TRUE,
  par = c(1., 10, 0., 1.),
  lower = c(0.0, 0.0, 0., 0.99),
  upper = c(4., 200, 0.48, 1.01)
) # + ylim(0, common_ylim)
```


:::


# overview {#sec-overview}

```{r}
knitr::kable(regression_storage, digits = 2)
```
