---
title: "Water Level Variograms: Resampling/Bootstrapping"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---


::: {.callout-warning}
In contrast to the previous notebook (which worked on `mTAW`), this one bootstraps water levels with the local coordinate reference (`mMaaiveld`).
:::


# preparation

... the usual.

```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("ggplot2")   |> void() # for cut_number
library("arrow")     |> void() # writing parquet files
library("parallel")  |> void() # parallel processing

# if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- glue::glue("data/waterlevel_data_cl{cluster_dist}.parquet")
chem_export_file <- glue::glue("data/waterchemistry_data_cl{cluster_dist}_alltimesd.parquet")
diffdata_export_file <- glue::glue("data/waterchemistry_diffdata_cl{cluster_dist}_alltimesd.parquet")


# for reference
# metadata_export_file <- here::here(local_cache_folder, "_metadata.parquet")
# cluster_dist <- 3400 # m # parameter for clustering, used on data loading


soilclass_colors <- c(
  "heavy" = "sienna",
  "light" = "lightgoldenrod",
  "peat" = "darkseagreen",
  "unknown" = "slategray"
)

```


```{r parameters}
#| code-fold: false

# parameters
extent <- 512 # m
# n_bins <- 50
min_pair_distance <- 1 # m
bin_target <- 100

analysis_label <- "waterchemistry_filterdepthfiltered"
regression_storage_path <- here::here("cache", "regression")

n_bootstraps <- 2**16 # number of bootstrapping iterations
```



```{r load-data}

pp_data <- read_parquet(data_export_file)

std_chem <- read_parquet(file = chem_export_file) %>%
  mutate(date = as.Date(as.character(date), format = "%Y%m%d"))
pp_chem_nh4 <- std_chem %>%
  select(pp_idx, sample_idx, date, meq_nh4_log, meq_nh4_raw) %>% 
  summarize(
    across(everything(), ~mean(., na.rm = TRUE)),
    .by = pp_idx
  )

# std_chem %>%
#   filter(pp_idx %in% c(4178, 2586)) %>%
#   arrange(desc(date))

```


loading the differentials and inspecting time intervals:

```{r load-diffdata}

diff_data_raw <- read_parquet(diffdata_export_file) %>%
  filter(!is_same_location) # only different locations

dt <- diff_data_raw %>%
  filter(is_first_dnh4 | is_first_dno3 | is_first_dpo4) %>% 
  # mutate(dt = as.numeric(date2 - date1, "days")) %>% already stored
  pull(dt)

breaks <- c(0., 0.4, 1.1, 6., 8., 13., 15., 20., 22., 25., 35., 50., 60., 80., 120., 400.)
hist(dt[dt <= max(breaks)], breaks)

message(sum(dt == 0) / length(dt))

```



```{r timefilter-diff-data}

join_a_chemical <- function(.data, chem) {
  # .data <- diff_backbone
  # chem <- "nh4"

  cols <- c(
    "pp_idx1", "date1", "pp_idx2", "date2", "dt",
    glue::glue("meq_{chem}_log1"),
    glue::glue("meq_{chem}_log2"),
    glue::glue("d{chem}"),
    glue::glue("d{chem}r"),
    glue::glue("is_first_d{chem}")
  )
  
  .data %>% 
    left_join(
      diff_data_raw %>% 
        select(!!!rlang::syms(cols)) %>%
        filter(!!as.symbol(glue::glue("is_first_d{chem}"))) %>% 
        filter(dt < 1)
      ,
      by = join_by(pp_idx1, date1, pp_idx2),
      suffix = c("", glue::glue("_{chem}")),
      relationship = "many-to-many"
    ) %>%
    return()
}

diff_backbone <- diff_data_raw %>%
  distinct(pp_idx1, date1, pp_idx2) %>%
  mutate(date2 = NA, dt = NA)

diff_data_mixed <- diff_backbone %>%
  join_a_chemical("nh4") %>%
  join_a_chemical("no3") %>%
  join_a_chemical("po4") %>%
  select(-date2, -dt)

diff_data <- diff_data_mixed  %>%
  filter(!(is.na(dnh4)&is.na(dno3)&is.na(dpo4)))

diff_data %>%
  mutate(
    nan_nh4 = is.na(dnh4),
    nan_no3 = is.na(dno3),
    nan_po4 = is.na(dpo4)
  ) %>% 
  count(nan_nh4, nan_no3, nan_po4) %>%
  knitr::kable()

```


```{r join-diffdata}

pp_join <- pp_data %>%
  filter(fdepth <= 5) %>% 
  select(pp_idx, fdepth, flength, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )
# pp_join %>% nrow()
# pp_join %>% distinct() %>% nrow()
# yes, they are unique

diff_meta <- diff_data %>%
  distinct(pp_idx1, pp_idx2) %>%
  inner_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  inner_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      # damp = emd_ampsum2-emd_ampsum1,
      # dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope = mean(c(slope_r2.5m2, slope_r2.5m1)) 
  )

diff_meta %>%
  mutate(
    short_filters = (flength1 <= 1.0) & (flength2 <= 1.0),
    equiprofund = abs(dz) <= 1.0
  ) %>% 
  filter(short_filters, equiprofund) %>% 
  nrow()

```

`diff_meta` are the observed, distinct pairs. There are <19k.



```{r glimpse-data}

# diff_meta %>% glimpse

diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      flength1, flength2,
      x1, y1, x2, y2, 
      soilclass1, # ecoregion,
      wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(glue::glue("{pp_idx1}_{pp_idx2}"))
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>%
  filter(ds >= min_pair_distance) %>% 
  # filter(
  #   # ds >= extent/n_bins
  #   date2 - date1 == 0 # already filtered
  # ) %>%
  mutate(date = date1) %>% select(-date1)
  # mutate(dt = abs(difftime(date1, date2, units = "days")))


g <- diff_all %>% ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  labs(x = "afstand (m)", y = "aantal paren") +
  theme_bw()

print(g)


diff_all %>% glimpse()
# diff_all %>% filter(is_first_dnh4) %>% glimpse()
# diff_all %>% filter(pair == "4178_2586") %>% glimpse()
# diff_all %>% filter(pair == "4178_2586") %>% select(ds, date, dnh4)

```

`diff_all` contains all the observations of water chemistry at the pairs (~53k), but only those which are "firsts" for differences of either chemical.

# pruning

Here, the filter criteria are applied.

```{r}
diff_all_raw <- diff_all
diff_all <- diff_all %>%
  mutate(
    short_filters = (flength1 <= 1.0) & (flength2 <= 1.0),
    equiprofund = abs(dz) <= 1.0
  )

```

```{r}

diff_all %>% ggplot(aes(
    x = ds,
    group = as.factor(short_filters),
    fill = as.factor(short_filters)
    )) +
  geom_histogram(bins = 64, color = "black") +
  theme_bw() +
  theme(legend.position = c(.9,.9))

```

```{r}

diff_all %>% ggplot(aes(
    x = ds,
    group = as.factor(equiprofund),
    fill = as.factor(equiprofund)
    )) +
  geom_histogram(bins = 64, color = "black") +
  theme_bw() +
  theme(legend.position = c(.9,.9))

```

```{r}
diff_all %>%
  count(short_filters, equiprofund)
```

```{r}
diff_all <- diff_all %>%
  filter(short_filters, equiprofund)
```



# grouping and binning

```{r}
parameter_of_choice <- "dnh4"
```


After some good prior experience, I will once more apply binning with variable bin width, to achieve **equal sample size per bin**.
Many location pairs are repeatedly measured, i.e. both of the pair are simultaneously measured on different days.



```{r equibin-preparation}
#| code-fold: false

# this is the full length data, just with a group column.
diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!(
      is.na(dnh4[is_first_dnh4]) |
      is.na(dno3[is_first_dno3]) |
      is.na(dpo4[is_first_dpo4])
    ))),
    n_nh4 = sum(as.integer(!is.na(dnh4[is_first_dnh4]))),
    n_no3 = sum(as.integer(!is.na(dno3[is_first_dno3]))),
    n_po4 = sum(as.integer(!is.na(dpo4[is_first_dpo4]))),
    soilclass = first(soilclass1),
    # ecoregion = first(ecoregion),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    # dt = mean(abs(date2-date1), na.rm = TRUE),
    # semivar = 1/(2*n_obs)*sum(dPC1^2, na.rm = TRUE),
    dnh4 = 1/(n_nh4)*sum(abs(dnh4[is_first_dnh4]), na.rm = TRUE),
    dno3 = 1/(n_no3)*sum(abs(dno3[is_first_dno3]), na.rm = TRUE),
    dpo4 = 1/(n_po4)*sum(abs(dpo4[is_first_dpo4]), na.rm = TRUE),
    dnh4r = 1/(n_nh4)*sum(abs(dnh4r[is_first_dnh4]), na.rm = TRUE),
    dno3r = 1/(n_no3)*sum(abs(dno3r[is_first_dno3]), na.rm = TRUE),
    dpo4r = 1/(n_po4)*sum(abs(dpo4r[is_first_dpo4]), na.rm = TRUE),
    # dPC1 = 1/(n_obs)*sum(abs(dPC1), na.rm = TRUE),
    # dPC2 = 1/(n_obs)*sum(abs(dPC2), na.rm = TRUE),
    # dPC3 = 1/(n_obs)*sum(abs(dPC3), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(soilclass) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  relocate(distance_equibin) %>% 
  ungroup

# diff_equigrouped %>% glimpse()
# diff_equigrouped %>% head(9) %>% tail(1) %>% knitr::kable()
# 4178_2586
    
combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)
mean_finite <- function (val) mean(val[is.finite(val)], na.rm = TRUE)

# Here, grouped observations get thrown into bins.
diff_binned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    # dt = mean(dt), 
    n_obs = sum(n_obs),
    n_nh4 = sum(n_nh4),
    n_no3 = sum(n_no3),
    n_po4 = sum(n_po4),
    # semivariance_groupmean = mean(semivar, na.rm = TRUE),
    # semivariance_all = combine_grouped(semivar, cnt),
    # dPC1_all = combine_grouped(dPC1, n_obs),
    dnh4  = mean_finite(dnh4),
    dno3  = mean_finite(dno3),
    dpo4  = mean_finite(dpo4),
    dnh4r = mean_finite(dnh4r),
    dno3r = mean_finite(dno3r),
    dpo4r = mean_finite(dpo4r),
    # dPC1 = mean(dPC1, na.rm = TRUE), 
    # dPC2 = mean(dPC2, na.rm = TRUE), 
    # dPC3 = mean(dPC3, na.rm = TRUE), 
    .by = c(distance_equibin, soilclass)
  ) %>%
  ungroup %>%
  arrange(ds) %>%
  filter(ds <= extent)

print(diff_binned %>% pull(cnt) %>% sum())
print(diff_binned %>% filter(ds < 100) %>% pull(cnt) %>% sum())
print(sum(diff_binned$n_obs))
print(sum(diff_binned$n_nh4))
print(sum(diff_binned$n_no3))
print(sum(diff_binned$n_po4))
```

There are 18345 pairs, which are assigned to a distance bin. 
Repeated observations (simultaneous measurement on different occasions) are averaged.



```{r equibin-samplesize}
#| label: fig-equibin-sample-size
diff_binned %>%
  ggplot(aes(x = ds, y = cnt, color = soilclass)) +
  geom_step() +
  geom_hline(yintercept = 0, color = "darkgrey") +
  theme_bw()
```



## Regression Procedure Wrapper

```{r regression-wrapper}
#| code-fold: true

matern_function_3p <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- 1 # parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}


fit_matern <- function(x, y, ...) {
}
```



```{r regression-matérn-general}
trafo <- function(x) x

regression_wrapper <- function(
    regression_data, 
    regression_variable = parameter_of_choice,
    ...) {
  
  x <- regression_data$ds
  y <- trafo(regression_data %>% pull(!!regression_variable))

  y <- y[x > 0]
  x <- x[x > 0]
  y <- y[x <= extent]
  x <- x[x <= extent]

  x <- x[!is.na(y)]
  y <- y[!is.na(y)]

 
  ## using L-BFGS-B to get the optimizer close
  pre_ <- optim(
    fn = function(params) {
      distweighted_target_function(x, y, matern_function_3p, params)
      # wrap_target_function(x, y, matern_function_3p, params)
    },
    # par = par,
    # upper = upper,
    # lower = lower,
    ...,
    method = "L-BFGS-B" # "Nelder-Mead" # "L-BFGS-B"
  )

  ## then performing a local Nelder-Mead adjustment
  optimizer_results <- pre_
  # optimizer_results <- optim(
  #   fn = function(params) {
  #     distweighted_target_function(x, y, matern_function_3p, params)
  #     # wrap_target_function(x, y, matern_function_3p, params)
  #   },
  #   par = pre_$par,
  #   method = "L-BFGS-B",
  #   lower = lower_limits, upper = upper_limits,
  #   control = list("fnscale" = 1000)
  # )
  # optimizer_results <- optim(
  #   fn = function(params) {
  #     distweighted_target_function(x, y, matern_function_3p, params)
  #     # wrap_target_function(x, y, matern_function_3p, params)
  #   },
  #   par = pre_$par,
  #   method = "Nelder-Mead" # "L-BFGS-B"
  # )
  
  # store everything in one list;
  # do I sense a smidgen of OOP here? No, not really.
  optimizer_results$pre_par <- pre_$par
  optimizer_results$fcn <- matern_function_3p
  optimizer_results$regx <- x
  optimizer_results$regy <- y

  # wrap the function with the optimized parameters
  optimizer_results$predict <- create_prediction_function(
    matern_function_3p,
    optimizer_results
  )

  # calculate 1cm threshold
  scale <- optimizer_results$par[1]
  optimizer_results$thresholds <- lapply(
      c(0.01, 0.02, 0.03, 0.05, 0.10, 0.05 * scale, 0.10 * scale, 0.15 * scale),
      FUN = \(thresh) calculate_limit(
        optimizer_results,
        threshold = thresh,
        fit_fcn = matern_function_3p,
        prep_fcn = shift_nugget_matern3p
      )
    )

  # par <- optimizer_results$par
  # par[3] <- 0 # no nugget
  # threshold = 0.01

  return(optimizer_results)
}


```


```{r}
diff_equibinned <- diff_binned
```


# bootstrapping
- <https://www.modernstatisticswithr.com/modchapter.html#a-general-approach>


## all in: getting starting values

```{r regression-reference}
lower_limits = c(0.0,  0.0, 0.0)# 0.1)
upper_limits = c(4.0, extent, 4.0)#16.01)
reference <- regression_wrapper(
  regression_data = diff_equibinned,
  regression_variable = parameter_of_choice,
  par = c(0.5, 10., 0.1), 
  lower = lower_limits,
  upper = upper_limits
)


g <- ggplot(NULL) %>% 
  add_regression_to_plot(reference, color = "darkgrey")
g + xlab("distance (m)") + ylab("mean absolute difference (m)") +
  ylim(0, 1.6) +
  theme_minimal()
```


```{r parallel-bootstrap}

# straps <- bind_rows(mclapply(
#   1:n_iterations,
#   datawrap_bootstrap("all", diff_equibinned),
#   mc.cores = detectCores()-1
#   ))
# 
# straps %>%
#   filter(conv == 0) %>% 
#   ggplot(aes(x = threshold)) +
#   geom_histogram(bins = 256, fill = "lightgray", color = "black") +
#   labs(x = "threshold", y = "bootstrap", title = sc) +
#   theme_bw()
    
```

# wrapping by soilclass

## bootstrap function
```{r bootstrap}


datawrap_bootstrap <- function(reference, pass_data) {

  bootstrap_function <- function(i){
    # Draw row numbers for the bootstrap sample:
    row_numbers <- sample(1:nrow(pass_data), as.integer(nrow(pass_data)),
                          replace = TRUE)
    
    # Obtain the bootstrap sample:
    boot_sample <- pass_data[row_numbers,]
    
    # jittering the observations
    # see https://influentialpoints.com/Training/properties-of-bootstrap_estimators.htm#howm
    # boot_sample$dm <- jitter(boot_sample$dm, amount = 0.1*sd(boot_sample$dm))
    boot_sample$dnh4 <- jitter(boot_sample$dnh4, amount = 0.1*sd(boot_sample$dnh4))
    boot_sample$dno3 <- jitter(boot_sample$dno3, amount = 0.1*sd(boot_sample$dno3))
    boot_sample$dpo4 <- jitter(boot_sample$dpo4, amount = 0.1*sd(boot_sample$dpo4))

    result <- regression_wrapper(
      regression_data = boot_sample,
      regression_variable = parameter_of_choice,
      # par = reference$par,
      par = c(0.5, 10., 0.1),
      lower = lower_limits,
      upper = upper_limits
    )
    return(as.data.frame(list(
        "i" = i,
        "conv" = result$convergence,
        "rmse" = result$value,
        "scale" = result$par[1],
        "sigma" = result$par[2],
        "nugget" = result$par[3],
        # "nu" = result$par[4],
        "threshold"     = result$thresholds[[1]],
        "threshold001"  = result$thresholds[[1]],
        "threshold002"  = result$thresholds[[2]],
        "threshold003"  = result$thresholds[[3]],
        "threshold005"  = result$thresholds[[4]],
        "threshold010"  = result$thresholds[[5]],
        "threshold005p" = result$thresholds[[6]],
        "threshold010p" = result$thresholds[[7]],
        "threshold015p" = result$thresholds[[8]]
      )))
  }
  return(bootstrap_function)
}


```


## subset

```{r subset-bootstrap}

ref_store <- list()

subset_bootstrap <- function(sc, sub_data, n_iterations = 999){
  reference <- regression_wrapper(
    regression_data = sub_data,
    par = c(0.5, 10, 0.1), # 1.0)
    lower = lower_limits,
    upper = upper_limits,
    control = list("fnscale" = 1000)
  )


  ref_store[[sc]] <<- as.data.frame(list(
      "i" = 0,
      "conv" = reference$convergence,
      "rmse" = reference$value,
      "scale" = reference$par[1],
      "sigma" = reference$par[2],
      "nugget" = reference$par[3],
      # "nu" = reference$par[4],
        "threshold" =     reference$thresholds[[1]],
        "threshold001" =  reference$thresholds[[1]],
        "threshold002" =  reference$thresholds[[2]],
        "threshold003" =  reference$thresholds[[3]],
        "threshold005" =  reference$thresholds[[4]],
        "threshold010" =  reference$thresholds[[5]],
        "threshold005p" = reference$thresholds[[6]],
        "threshold010p" = reference$thresholds[[7]],
        "threshold015p" = reference$thresholds[[8]],
      "soilclass" = sc
    ))

  straps <- bind_rows(mclapply(
    1:n_iterations,
    datawrap_bootstrap(reference, sub_data),
    mc.cores = detectCores()-2
    ))
  return (straps)
}

```

```{r soil-subset-bootstrap}

all_dolled_up <- list()
for (sc in c("heavy", "light", "peat")) {
  sub_data <- diff_equibinned %>% filter(soilclass == sc)
  straps <- subset_bootstrap(sc, sub_data, n_iterations = n_bootstraps)
  straps$soilclass <- sc
  all_dolled_up[[sc]] <- straps
}

all_dolled_up[["ref"]] <- bind_rows(ref_store)
all_dolled_up <- bind_rows(all_dolled_up)
all_dolled_up <- all_dolled_up %>%
  mutate(sill = nugget + scale)

# save bootstraps
arrow::write_parquet(all_dolled_up, sink = "./data/bootstrapping_dm_soilclass.parquet")

```


```{r count}

# count failed regressions
all_dolled_up %>% count(conv)

```




```{r one-parameter}

ref_boots <- all_dolled_up %>%
  filter(i == 0)
sub_boots <- all_dolled_up %>%
  filter(conv == 0, i > 0, if_all(everything(), ~ !is.na(.x)))

# param <- "threshold"
knitr::kable(t(ref_boots))

# knitr::kable(t(head(sub_boots)))
print(summary(sub_boots))
```



```{r plot-one-parameter}
plot_param <- function(param, limit_x = NA){
  h <- sub_boots %>%
    mutate(param = sub_boots[, param]) %>% 
    ggplot(aes(x = param, fill = soilclass)) +
    geom_histogram(bins = 256, width = 1.05, color = NA, alpha = 0.67, position = "identity") +
    geom_vline(xintercept = ref_boots[,param]) +
    xlab(param) +
    scale_fill_manual(values = soilclass_colors) +
    scale_color_manual(values = soilclass_colors) +
    theme_bw()

  # if (param == "threshold") h <- h + xlim(0, 50)
  # if (param == "scale") h <- h + xlim(lower_limits[1], upper_limits[1])
  # if (param == "sigma") h <- h + xlim(lower_limits[2], upper_limits[2])
  # if (param == "nugget") h <- h + xlim(lower_limits[3], upper_limits[3])

  # if (TRUE) h <- h + ylim(0, 100) # ARRRR! R blanks bars above the ylim 8(
  if (!is.na(limit_x)) h <- h + xlim(0, limit_x)

  ggsave(paste0("figures/bootstrap_dm_", param, ".png"), h, dpi = 300)
  ggsave(paste0("figures/bootstrap_dm_", param, ".pdf"), h, dpi = 300)
  return(h)
}

```


Quantiles for `threshold`:

```{r threshold-quantiles}
threshold_quantiles <- sub_boots %>%
  summarize(
    q02 = quantile(threshold, c(0.02)),
    median = quantile(threshold, c(0.50)),
    q98 = quantile(threshold, c(0.98)),
    .by = soilclass 
  ) 
knitr::kable(threshold_quantiles, digits = 1)
```


Plots:

```{r plot-all-parameters}
silence <- suppressWarnings
print(plot_param("threshold", limit_x = 32) |> silence())
print(plot_param("threshold001") |> silence())
print(plot_param("threshold002") |> silence())
print(plot_param("threshold003") |> silence())
print(plot_param("threshold005") |> silence())
print(plot_param("threshold010") |> silence())
print(plot_param("threshold005p") |> silence())
print(plot_param("threshold010p") |> silence())
print(plot_param("threshold015p") |> silence())
print(plot_param("rmse") |> silence())
print(plot_param("scale", limit_x = 0.5) |> silence())
print(plot_param("sigma", limit_x = 200) |> silence())
print(plot_param("nugget") |> silence())
print(plot_param("sill", limit_x = 0.6) |> silence())

```

