---
title: "Spatial Coupling: Water Levels"
subtitle: "Retrieval of variogram regression parameters form spatial coupling analysis of water levels."
bibliography: references_csl.json
date: "2026-01-13"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: false
knitr:
  opts_chunk:
    echo: true
---

# Theory - Reminder


As a difference parameter $y$ on the y-axis, we use **Mean Absolute Difference (MAD)**.

$$ y = \frac{1}{N} \sum\limits_{i,j} \left| z_j - z_i\right| $$


We parametrize the Matérn a bit differently than usual, by using the difference in MAD, $dy$:
$$ dy = y_{max} - y_{min}$$
$$ y = y_{max} - dy \cdot \left( \frac{\sqrt{2\nu}}{\sigma} x \right)^{\nu} \cdot K_{\nu}\left( \frac{\sqrt{2\nu}}{\sigma} x \right) \cdot \frac{1}{2^{\nu - 1} \cdot y\left(\nu\right)} $$

With sill $y_{max}$, nugget $y_{min}$, turning point $\sigma$ (related to range), and a shape parameter $\nu$ that controls the form of the sigmoid[^1]. 
(Compare [wikipedia](https://en.wikipedia.org/wiki/Mat%C3%A9rn_covariance_function).)


[^1]: In practical application, I soft-fix $\nu$ to a value of one ($\nu = 1.0$), and only give it moderate freedom in case of visual shape mismatches. 


Bins are of dynamic width: we group a variable number of paired difference-opservations, ensuring that all bins are homogeneously filled with **approximately 100 observations each**.


This methodology is derived in some more detail in the original analysis, and summarized  [in a tutorial here](https://tutorials.inbo.be/tutorials/spatial_variograms).


# Preparation

```{r libraries}
#| code-fold: true
#| eval: true
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("rprojroot") |> void() # folder dynamics

# uniform regression tools and plot helpers
source("./regression_tools.R")

base_path <- rprojroot::find_root(rprojroot::is_rstudio_project)
regression_storage_folder <- here::here(base_path, "cache", "regression")

data_version <- "20260122"  
```



```{r aesthetics-and-style}
soilclasses <- c("heavy", "light", "peat")

```


The following functions are convenience tools to retrieve the previously stored regression data.

```{r data-paths}
#| code-fold: false
#| eval: true

stitch_filepath <- function(label, sc, reg_var, extension = "rds", path_location = ".") {
  # examples: 
  #   label = "waterlevels", "allin", "slopefilter", "maaiveld"
  #   sc = "heavy", "light", "peat"
  #   reg_var = "dw", "dm"
  #
  here::here(
      path_location,
      glue::glue("{label}_{reg_var}_{sc}.{extension}") 
    ) %>% 
    return()
}


load_fitdata <- function(
    label = c("waterlevels"), #, "allin", "maaiveld"),
    soilclass = c("heavy", "light", "peat"),
    regression_variable = c("dw", "dm")
  ) {
  stitch_filepath(
      label,
      soilclass,
      regression_variable,
      extension = "rds",
      path_location = regression_storage_folder
    ) %>%
    readRDS() %>% 
    return()
}

```


```{r}

print_results <- function(regressions) {
  knitr::kable(
    t(sapply(names(regressions),
      FUN = \(sc) calculate_limit(
          regressions[[sc]],
          threshold = 0.01,
          fit_fcn = matern_function,
          prep_fcn = shift_nugget_matern4p
        )
    )),
    digits = 1
  )
}
```

```{r plot-function}

plot_matern_regression <- function(regressions, soilclass) {
  chosen_regression <- regressions[[soilclass]]
  matern_parameters <- chosen_regression$par
  
  scale  <- matern_parameters[1] # related to semivariance
  sigma  <- matern_parameters[2] # related to actual range; turning point
  nugget <- matern_parameters[3] # variance at zero distance
  nu     <- matern_parameters[4] # shape parameter; here: soft-fixed to ~1
  sill   <- scale + nugget
  
  distance_m <- seq(0., 500, length.out = 1001)[2:1001]
  mean_abs_difference_m <- matern_function(distance_m, matern_parameters)
  
  plot(
    distance_m,
    mean_abs_difference_m,
    type = "l",
    ylim = c(nugget - 0.1, sill + 0.1)
  )
  points(chosen_regression$regx, chosen_regression$regy, type = "p")
  abline(h = nugget, lty = 2)
  abline(h = sill, lty = 2)
  abline(v = sigma, lty = 2)
  title(soilclass)
}

```



```{r print-data-tables}

print_input_data_table <- function(regressions, soilclass) {
  chosen_regression <- regressions[[soilclass]]

  data.frame(
    "mean bin distance (m)" = chosen_regression$regx,
    "mean abs difference (m)" = chosen_regression$regy
  ) %>% 
  return()
}

print_data_table_matern_regression <- function(regressions, soilclass) {
  chosen_regression <- regressions[[soilclass]]
  matern_parameters <- chosen_regression$par
  
  scale  <- matern_parameters[1] # related to semivariance
  sigma  <- matern_parameters[2] # related to actual range; turning point
  nugget <- matern_parameters[3] # variance at zero distance
  nu     <- matern_parameters[4] # shape parameter; here: soft-fixed to ~1
  sill   <- scale + nugget
  
  distance_m <- seq(0., 100., length.out = 101)
  distance_m <- distance_m[distance_m > 0]
  mean_abs_difference_m <- matern_function(distance_m, matern_parameters)
  difference_nuggetless_m <- mean_abs_difference_m - nugget
  

  data.frame(
    "distance (m)" = distance_m,
    "mean abs difference (m)" = mean_abs_difference_m,
    "mean abs difference *ex nugget* (m)" = difference_nuggetless_m
  ) %>%
  return()
}

```

# Results

## mTAW

```{r load-regressions-dw}
regressions_dw <- list(
  "heavy" = load_fitdata("waterlevels", "heavy", "dw"),
  "light" = load_fitdata("waterlevels", "light", "dw"),
  "peat" = load_fitdata("waterlevels", "peat", "dw")
)

```


```{r results-table-dw}
print_results(regressions_dw)
```


::: {.panel-tabset group="soilclass"}
### heavy
```{r}
soilclass <- "heavy"

print_regression_results(regressions_dw[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dw, soilclass)
print_input_data_table(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
```

### light
```{r}
soilclass <- "light"
print_regression_results(regressions_dw[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dw, soilclass)
print_input_data_table(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
```

### peat
```{r}
soilclass <- "peat"
print_regression_results(regressions_dw[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dw, soilclass)
print_input_data_table(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dw, soilclass) %>% knitr::kable(digits = 3)
```
:::


```{r}
analysis_label <- "waterlevels"
regression_storage_path <- here::here("cache", "regression")
reg_var <- "mTAW"

for (sc in soilclasses) {
  matern_fit <- regressions_dw[[sc]]

  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_regression.rds")
  saveRDS(matern_fit, file = here::here(regression_storage_path, sink_file))
  
  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_binned_observations.csv")
  regdat <- extract_input_data_table(matern_fit, units = "m|TAW")
  write.csv2(regdat, file = here::here(regression_storage_path, sink_file))

  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_regression_values.csv")
  regresult <- extract_data_table_matern_regression(matern_fit, units = "m|TAW")
  write.csv2(regresult, file = here::here(regression_storage_path, sink_file))
}

```

## mMaaiveld

```{r load-regressions-dm}
regressions_dm <- list(
  "heavy" = load_fitdata("maaiveld", "heavy", "dm"),
  "light" = load_fitdata("maaiveld", "light", "dm"),
  "peat" = load_fitdata("maaiveld", "peat", "dm")
)

```


```{r results-table-dm}
print_results(regressions_dm)
```


::: {.panel-tabset group="soilclass"}
### heavy
```{r}
soilclass <- "heavy"
print_regression_results(regressions_dm[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dm, soilclass)
print_input_data_table(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
```

### light
```{r}
soilclass <- "light"
print_regression_results(regressions_dm[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dm, soilclass)
print_input_data_table(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
```


### peat
```{r}
soilclass <- "peat"
print_regression_results(regressions_dm[[soilclass]], label = "regression:")
plot_matern_regression(regressions_dm, soilclass)
print_input_data_table(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
print_data_table_matern_regression(regressions_dm, soilclass) %>% knitr::kable(digits = 3)
```
:::


```{r}
analysis_label <- "waterlevels"
regression_storage_path <- here::here("cache", "regression")
reg_var <- "mMaaiveld"

for (sc in soilclasses) {
  matern_fit <- regressions_dm[[sc]]
  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_regression.rds")
  saveRDS(matern_fit, file = here::here(regression_storage_path, sink_file))
  
  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_binned_observations.csv")
  regdat <- extract_input_data_table(matern_fit, units = "m|Maaiveld")
  write.csv2(regdat, file = here::here(regression_storage_path, sink_file))

  sink_file <- glue::glue("{analysis_label}_{reg_var}_{sc}_regression_values.csv")
  regresult <- extract_data_table_matern_regression(matern_fit, units = "m|Maaiveld")
  write.csv2(regresult, file = here::here(regression_storage_path, sink_file))
}

```
