---
title: "Water Level Variograms: Daily Data"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
    code-fold: true
    embed-resources: true
knitr:
  opts_chunk:
    echo: true
---



```{r setup}
#| include: false
void <- suppressPackageStartupMessages
library("dplyr")     |> void() # our favorite data wrangling toolbox
library("dbplyr")    |> void() # dplyr + databases
library("arrow")     |> void() # writing parquet files
library("ggplot2")   |> void() # a visualization toolbox
library("gridExtra") |> void() # panels with ggplot
library("ggcorrplot")|> void() # correlation, grid visualization
library("mapview")   |> void() # map visualization
library("ggridges")  |> void() # density ridges


if (getOption("browser") == "") options(browser = "librewolf") 

# user-set parameters
cluster_dist <- 3400 # m # parameter for clustering
data_export_file <- glue::glue("data/waterlevel_data_cl{cluster_dist}.parquet")
chem_export_file <- glue::glue("data/waterchemistry_data_cl{cluster_dist}_alltimesd.parquet")
diffdata_export_file <- glue::glue("data/waterchemistry_diffdata_cl{cluster_dist}_alltimesd.parquet")

# parameters
extent <- 512 # m
# n_bins <- 50
bin_target <- 128 # pairs per bin
min_pair_distance <- 1.0 # m


ecoregion_colors <- c(
  "zuidoostelijke heuvelzone" = "darkorchid",
  "Kempen" = "sienna",
  "polders en de getijdenschelde" = "slategray",
  "kustduinen" = "burlywood",
  "Pleistocene riviervalleien" = "deepskyblue",
  "zuidwestelijke heuvelzone" = "coral",
  "Midden-Vlaamse overgangsgebieden" = "gold",
  "krijt-leemgebieden" = "honeydew3",
  "cuesta's" = "palevioletred",
  "grindrivieren" = "darkseagreen",
  "westelijke interfluvia" = "hotpink"
)
```


The script `120_waterchemistry_reload.qmd` should be used to assemble the data set loaded here.



```{r load-data}

pp_data <- read_parquet(data_export_file)

std_chem <- read_parquet(file = chem_export_file) %>%
  mutate(date = as.Date(as.character(date), format = "%Y%m%d"))
pp_chem_nh4 <- std_chem %>%
  select(pp_idx, sample_idx, date, meq_nh4_log, meq_nh4_raw) %>% 
  summarize(
    across(everything(), ~mean(., na.rm = TRUE)),
    .by = pp_idx
  )

```



```{r load-diffdata}

diff_data_raw <- read_parquet(diffdata_export_file) %>%
  filter(!is_same_location) # only different locations


join_a_chemical <- function(.data, chem) {
  # .data <- diff_backbone
  # chem <- "nh4"

  cols <- c(
    "pp_idx1", "date1", "pp_idx2", "date2", "dt",
    glue::glue("meq_{chem}_log1"),
    glue::glue("meq_{chem}_log2"),
    glue::glue("d{chem}"),
    glue::glue("d{chem}r"),
    glue::glue("is_first_d{chem}")
  )
  
  .data %>% 
    left_join(
      diff_data_raw %>% 
        select(!!!rlang::syms(cols)) %>%
        filter(!!as.symbol(glue::glue("is_first_d{chem}"))) %>% 
        filter(dt < 1)
      ,
      by = join_by(pp_idx1, date1, pp_idx2),
      suffix = c("", glue::glue("_{chem}")),
      relationship = "many-to-many"
    ) %>%
    return()
}

diff_backbone <- diff_data_raw %>%
  distinct(pp_idx1, date1, pp_idx2) %>%
  mutate(date2 = NA, dt = NA)

diff_data_mixed <- diff_backbone %>%
  join_a_chemical("nh4") %>%
  join_a_chemical("no3") %>%
  join_a_chemical("po4") %>%
  select(-date2, -dt)

diff_data <- diff_data_mixed  %>%
  filter(!(is.na(dnh4)&is.na(dno3)&is.na(dpo4)))

diff_data %>%
  mutate(
    nan_nh4 = is.na(dnh4),
    nan_no3 = is.na(dno3),
    nan_po4 = is.na(dpo4)
  ) %>% 
  count(nan_nh4, nan_no3, nan_po4) %>%
  knitr::kable()


```


```{r join-location-data}


pp_join <- pp_data %>%
  select(pp_idx, fdepth, pp_code, mp_idx, cluster,
         x, y, soilclass, elevation_dhmv, slope_r2.5m,
         wata_min_dist, wata_dist_class,
         emd_n, emd_amp1, emd_amp2, emd_ampsum,
         emd_qdiff, emd_residual_qdiff
         )

diff_meta <- diff_data %>%
  distinct(pp_idx1, pp_idx2) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one",
    suffix = c("", "1")
  ) %>%
  left_join(
    pp_join,
    by = join_by(pp_idx2 == pp_idx),
    relationship = "many-to-one",
    suffix = c("1", "2")
  ) %>%
  mutate(
      ds = sqrt((x2-x1)^2 + (y2-y1)^2),
      dh = elevation_dhmv2-elevation_dhmv1,
      dz = fdepth2-fdepth1,
      # damp = emd_ampsum2-emd_ampsum1,
      # dqdiff = emd_qdiff2-emd_qdiff1,
      dwadi = wata_min_dist2-wata_min_dist1,
      dslope = slope_r2.5m2-slope_r2.5m1,
      melevation = mean(c(elevation_dhmv2, elevation_dhmv1)),
      mslope = mean(c(slope_r2.5m2, slope_r2.5m1)) 
  )


```

Ecoregions:

```{r prepare-ecoregion}
ecoregions <- n2khab::read_ecoregions() %>% 
  select(region_name) %>% 
  rename(ecoregion = region_name) %>% 
  mutate(ecoregion = stringr::str_replace(ecoregion, "Ecoregio van de ", ""))

# n2khab::read_ecoregions() %>% sf::st_drop_geometry() %>% distinct(region_name)
ecoregion_lookup <- bind_rows(
  diff_meta %>% select(pp_idx1, x1, y1) %>% rename(pp_idx = pp_idx1, x = x1, y = y1),
  diff_meta %>% select(pp_idx2, x2, y2) %>% rename(pp_idx = pp_idx2, x = x2, y = y2)
  ) %>%
  distinct() %>% 
  sf::st_as_sf(coords = c("x", "y"), crs = 31370) %>% 
  sf::st_join(ecoregions)

ecoregion_lookup %>% 
  sf::st_drop_geometry() %>% 
  count(ecoregion) %>% arrange(desc(n)) %>%
  knitr::kable()

missing_points <- ecoregion_lookup %>% filter(is.na(ecoregion))
# mapview::mapview(ecoregions) +
#   mapview::mapview(missing_points)

```



```{r join-ecoregion}

diff_meta <- diff_meta %>%
  left_join(
    ecoregion_lookup,
    by = join_by(pp_idx1 == pp_idx),
    relationship = "many-to-one"
  )

diff_meta <- diff_meta %>%
  filter(
    ecoregion != "NA",
    !is.na(ecoregion)
  )
  
# diff_meta %>% 
#   count(ecoregion) %>% arrange(desc(n)) %>% knitr::kable()

```



```{r glimpse-data}

diff_all <- diff_meta %>%
  select(pp_idx1, pp_idx2, cluster1,
      x1, y1, x2, y2, 
      soilclass1, ecoregion,
      wata_dist_class1, wata_dist_class2,
      ds, dh, dz,
      # damp, dqdiff,
      dwadi, dslope,
      melevation, mslope
  ) %>%
  mutate(
    pair = factor(glue::glue("{pp_idx1}_{pp_idx2}"))
  ) %>%
  left_join(diff_data,
    by = join_by(pp_idx1 == pp_idx1, pp_idx2 == pp_idx2),
    relationship = "one-to-many"
  ) %>%
  filter(ds >= min_pair_distance) %>% 
  mutate(date = date1) %>% select(-date1)
    

diff_all %>% ggplot(aes(x = ds)) +
  geom_histogram(bins = 256, fill = "lightgray", color = "black") +
  theme_bw()

```


# binning


After some good prior experience, I will once more apply binning with variable bin width, to achieve **equal sample size per bin**.


```{r equibin-preparation}
#| code-fold: false

diff_equigrouped <- diff_all %>%
  summarize(
    n_obs = sum(as.integer(!(
      is.na(dnh4[is_first_dnh4]) |
      is.na(dno3[is_first_dno3]) |
      is.na(dpo4[is_first_dpo4])
    ))),
    n_nh4 = sum(as.integer(!is.na(dnh4[is_first_dnh4]))),
    n_no3 = sum(as.integer(!is.na(dno3[is_first_dno3]))),
    n_po4 = sum(as.integer(!is.na(dpo4[is_first_dpo4]))),
    soilclass = first(soilclass1),
    ecoregion = first(ecoregion),
    cluster = first(cluster1),
    ds = mean(ds, na.rm = TRUE),
    dnh4 = 1/(n_nh4)*sum(abs(dnh4[is_first_dnh4]), na.rm = TRUE),
    dno3 = 1/(n_no3)*sum(abs(dno3[is_first_dno3]), na.rm = TRUE),
    dpo4 = 1/(n_po4)*sum(abs(dpo4[is_first_dpo4]), na.rm = TRUE),
    dnh4r = 1/(n_nh4)*sum(abs(dnh4r[is_first_dnh4]), na.rm = TRUE),
    dno3r = 1/(n_no3)*sum(abs(dno3r[is_first_dno3]), na.rm = TRUE),
    dpo4r = 1/(n_po4)*sum(abs(dpo4r[is_first_dpo4]), na.rm = TRUE),
    .by = pair 
  ) %>% ungroup %>% arrange(ds) %>%
  group_by(ecoregion) %>% 
  mutate(
    distance_equibin = cut_number(
      ds, as.integer(floor(n()/min(c(n()/8, bin_target))))
    ) # cut to equal sized bins
  ) %>%
  relocate(distance_equibin) %>% 
  ungroup


    
combine_grouped <- function (val, cnt) sum(val*cnt, na.rm = TRUE)/sum(cnt, na.rm = TRUE)
mean_finite <- function (val) mean(val[is.finite(val)], na.rm = TRUE)

diff_binned <- diff_equigrouped %>%
  summarize(
    cnt = n(),
    ds = mean(ds), 
    n_obs = sum(n_obs),
    n_nh4 = sum(n_nh4),
    n_no3 = sum(n_no3),
    n_po4 = sum(n_po4),
    dnh4  = mean_finite(dnh4),
    dno3  = mean_finite(dno3),
    dpo4  = mean_finite(dpo4),
    dnh4r = mean_finite(dnh4r),
    dno3r = mean_finite(dno3r),
    dpo4r = mean_finite(dpo4r),
    .by = c(distance_equibin, ecoregion)
  ) %>%
  ungroup %>%
  arrange(ds) %>%
  filter(ds <= extent)

```



```{r equibin-samplesize}
#| label: fig-equibin-sample-size
diff_binned %>%
  ggplot(aes(x = ds, y = cnt, color = ecoregion)) +
  geom_step() +
  geom_hline(yintercept = 0, color = "darkgrey") +
  scale_color_manual(values = ecoregion_colors) +
  theme_minimal()
```


```{r}
ridgeplot_by_ecoregion <- function(er, reg_var) {
  # er = "Kempen"
  # reg_var = "dpo4"

  diff_equigrouped %>%
    filter(ecoregion == er, ds < 100, ds > 0) %>% 
    filter(is.finite(!!rlang::sym(reg_var))) %>% 
    ggplot(aes(x = !!rlang::sym(reg_var), y = distance_equibin)) +
    geom_density_ridges(rel_min_height = 0.08,
      quantile_lines = TRUE, quantiles = 4,
      scale = pi, bandwidth = 0.05,
      fill = ecoregion_colors[[er]],
      alpha = 0.8
    ) +
    # scale_fill_viridis_c(option = "plasma", name = "") +  xlim(0, 4) +
    labs(
      x = glue::glue("mean log(concentration) difference {reg_var}"),
      y = "well pair distance"
    ) +
    xlim(0, 1.0) +
    theme_bw() + coord_flip() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

}

```

# equibin ridges {#sec-ridgeplots}

## $NH_4$

```{r dchem-column-nh4}
#| code-fold: false
column <- "dnh4"
```

::: {.panel-tabset group="ecoregion"}

### zuidoostelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidoostelijke heuvelzone",
  reg_var = column
)

```


### Kempen

```{r}
ridgeplot_by_ecoregion(
  "Kempen",
  reg_var = column
)

```


### polders en de getijdenschelde

```{r}
ridgeplot_by_ecoregion(
  "polders en de getijdenschelde",
  reg_var = column
)

```


### kustduinen

```{r}
ridgeplot_by_ecoregion(
  "kustduinen",
  reg_var = column
)

```


### Pleistocene riviervalleien

```{r}
ridgeplot_by_ecoregion(
  "Pleistocene riviervalleien",
  reg_var = column
)

```


### zuidwestelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidwestelijke heuvelzone",
  reg_var = column
)

```


### Midden-Vlaamse overgangsgebieden

```{r}
ridgeplot_by_ecoregion(
  "Midden-Vlaamse overgangsgebieden",
  reg_var = column
)

```

### krijt-leemgebieden

```{r}
ridgeplot_by_ecoregion(
  "krijt-leemgebieden",
  reg_var = column
)

```


### cuesta's

```{r}
ridgeplot_by_ecoregion(
  er = "cuesta's",
  reg_var = column
)
```


### grindrivieren

```{r}
ridgeplot_by_ecoregion(
  er = "grindrivieren",
  reg_var = column
)
```


### westelijke interfluvia

```{r}
ridgeplot_by_ecoregion(
  er = "westelijke interfluvia",
  reg_var = column
)
```


:::

## $NO_3$

```{r dchem-column-no3}
#| code-fold: false
column <- "dno3"
```

::: {.panel-tabset group="ecoregion"}

### zuidoostelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidoostelijke heuvelzone",
  reg_var = column
)

```


### Kempen

```{r}
ridgeplot_by_ecoregion(
  "Kempen",
  reg_var = column
)

```


### polders en de getijdenschelde

```{r}
ridgeplot_by_ecoregion(
  "polders en de getijdenschelde",
  reg_var = column
)

```


### kustduinen

```{r}
ridgeplot_by_ecoregion(
  "kustduinen",
  reg_var = column
)

```


### Pleistocene riviervalleien

```{r}
ridgeplot_by_ecoregion(
  "Pleistocene riviervalleien",
  reg_var = column
)

```


### zuidwestelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidwestelijke heuvelzone",
  reg_var = column
)

```


### Midden-Vlaamse overgangsgebieden

```{r}
ridgeplot_by_ecoregion(
  "Midden-Vlaamse overgangsgebieden",
  reg_var = column
)

```

### krijt-leemgebieden

```{r}
ridgeplot_by_ecoregion(
  "krijt-leemgebieden",
  reg_var = column
)

```


### cuesta's

```{r}
ridgeplot_by_ecoregion(
  er = "cuesta's",
  reg_var = column
)
```


### grindrivieren

```{r}
ridgeplot_by_ecoregion(
  er = "grindrivieren",
  reg_var = column
)
```


### westelijke interfluvia

```{r}
ridgeplot_by_ecoregion(
  er = "westelijke interfluvia",
  reg_var = column
)
```


:::

## $PO_4$

```{r dchem-column-po4}
#| code-fold: false
column <- "dpo4"
```

::: {.panel-tabset group="ecoregion"}

### zuidoostelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidoostelijke heuvelzone",
  reg_var = column
)

```


### Kempen

```{r}
ridgeplot_by_ecoregion(
  "Kempen",
  reg_var = column
)

```


### polders en de getijdenschelde

```{r}
ridgeplot_by_ecoregion(
  "polders en de getijdenschelde",
  reg_var = column
)

```


### kustduinen

```{r}
ridgeplot_by_ecoregion(
  "kustduinen",
  reg_var = column
)

```


### Pleistocene riviervalleien

```{r}
ridgeplot_by_ecoregion(
  "Pleistocene riviervalleien",
  reg_var = column
)

```


### zuidwestelijke heuvelzone

```{r}
ridgeplot_by_ecoregion(
  "zuidwestelijke heuvelzone",
  reg_var = column
)

```


### Midden-Vlaamse overgangsgebieden

```{r}
ridgeplot_by_ecoregion(
  "Midden-Vlaamse overgangsgebieden",
  reg_var = column
)

```

### krijt-leemgebieden

```{r}
ridgeplot_by_ecoregion(
  "krijt-leemgebieden",
  reg_var = column
)

```


### cuesta's

```{r}
ridgeplot_by_ecoregion(
  er = "cuesta's",
  reg_var = column
)
```


### grindrivieren

```{r}
ridgeplot_by_ecoregion(
  er = "grindrivieren",
  reg_var = column
)
```


### westelijke interfluvia

```{r}
ridgeplot_by_ecoregion(
  er = "westelijke interfluvia",
  reg_var = column
)
```


:::



