---
title: "Spatial Coupling of Observation Wells"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("watina")
library("n2khab")
library("sf")
library("stringr")
library("ggplot2")
library("mapview")

source("./watina_migration.R")

watina_dwh <- connect_watina()

```

# Summary
In this document, I will assemble and test components and functions to extract information which is potentially relevant to infer spatial coupling of groundwater observation wells.

Those information are:

- spatial distance
- distance to water body / stream
- height TAW difference
- watina region, eco-region, n2khab-type
- soil class

As a bonus, this will work on the _new_ watina data warehouse.
Another bonus is the extraction of unit tests for `watina` usage.

```{r}
get_db_table_list(watina_dwh)
```

# Spatial Distance
A use case to query spatial distance is as follows:

- take a given location (observation well candidate)
- get existing location from `watina` 
- find locations in the area

step 1: migrate `watina::get_locs()`
```{r}
source("./watina_migration.R")

obswells_db <- get_locs(
  watina_dwh,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  area_codes = c("HOP"),
  obswells = TRUE
) %>% mutate(line_nr = row_number())

glimpse(obswells_db)
# obswells_db %>% count
```

TODO: unit test - query of a test area will not be empty


NEXT: (i) bbox; 
(ii) circular mask
```{r}
# @param bbox Optional geospatial fiter (rectangle).
# A bounding box (class \code{bbox}), or a vector of four named elements
# \code{xmin}, \code{xmax}, \code{ymin}, \code{ymax} defining the
# boundary coordinates of a bounding box.
# If provided, only locations within this rectangular area will be returned.
# The CRS must be Belgian Lambert 72 (EPSG-code
# \href{https://epsg.io/31370}{31370}).

# @param mask An optional geospatial filter of class \code{sf}.
# If provided, only locations that intersect with \code{mask} will be returned,
# with the value of \code{buffer} taken into account.
# The CRS must be Belgian Lambert 72 (EPSG-code
# \href{https://epsg.io/31370}{31370}).
```
