---
title: "Spatial Coupling of Observation Wells: Variograms"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
# library("dbplot")
library("watina")
library("n2khab")
# library("terra")
# library("sf")
library("stringr")
library("ggplot2")
# library("mapview")
library("DBI")
library("arrow")

source("./watina_migration.R")
# source("./spatial_helpers.R")

# watina_dwh <- connect_watina()
watina_dwh <- connect_watina()


# n2khab_data_path <- locate_n2khab_data()
# options(n2khab_data_path = "C:\\R\\n2khab_data")

# options(browser = "librewolf") # a rather personal thing
```


With this notebook, I plan to download chemical data from watina,
cross compute differences,
and quantify variogram parameters
to estimate spatial similarity range.

# Chemical Data
## rationale/demarcation
The measurement intervals and data structure of hydrogeographical and chemical data in `watinA` are fundamentally different, and justify separate investigation.
In particular:

- Temporal sampling: water levels are monitored quasi-continuously (ideally), whereas chemical measurements happen twice a year at best.
- Chemistry: water levels are one simple measure, whereas chemistry comes as a set of multiple compoud classess.
- Methodology: there are detection limits on the chemistry side, and known pressure/level interdependencies (e.g. well/seepage areas) on the physical side.
- Recurrency: both measurement types show recurrent structure, which might be linked, but not directly.


Herein, I focus on chemistry, and more specifically on the following compound classes:

- \(NO_2\)
- \(NO_3\)
- \(NH_4\)
- \(PO_4\)


I will query pairs of (approximately) simultaneously samples sites and quantify differences of measurement outcome, depending on site distance. 
And then we'll see.


## glimpse at the database
To create the query below, knowledge of the database structure is beneficial.

The `dimDatum` table:
```{r}
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT *
    FROM dimDatum
      ")
  ))
```

I will use `Jaardag_Nummer/365.0 + Jaar` for restricting time difference.



The `dimPeilpunt` table:
```{r}
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT *,
        SUBSTRING(PeilpuntCode, 1, 4) AS region
    FROM dimPeilpunt
      ")
  ))
```

I will use `SUBSTRING(PeilpuntCode, 1, 3) = GebiedCode` equality for restricting spatial distance.



The `dimGebied` table:
```{r}
glimpse(DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT *
    FROM dimGebied
      ")
  ))
```

-> `GebiedCode` can be used to join locs afterwards.


## data query
There are existing `watina` functions for chemical data.
However, in my experience, SQL is efficient at cross-joining large tables.

Therefore, I will use a direct query to get the difference-distance data.


```{r, eval=FALSE}
data_raw <- DBI::dbGetQuery(watina_dwh,
  paste0("
    SELECT * FROM (
      SELECT
        CHM1.ChemMetingWID AS id1,
        CHM1.pp_code AS pp1,
        CHM1.MetingStatusCode AS status1,
        CHM1.DatumWID AS date1,
        CHM1.time AS t1,
        CHM1.Meetwaarde AS mw1,
        CHM2.ChemMetingWID AS id2,
        CHM2.pp_code AS pp2,
        CHM2.MetingStatusCode AS status2,
        CHM2.DatumWID AS date2,
        CHM2.Meetwaarde AS mw2,
        CHM1.ChemVarWID AS chemvar_id,
        CHM1.region AS region
      FROM (
        SELECT
            FactChemischeMeting.*,
            Jaardag_Nummer/365.0 + Jaar AS time,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            dimPeilpunt.PeilpuntCode AS pp_code
        FROM FactChemischeMeting
        LEFT JOIN dimDatum ON FactChemischeMeting.DatumWID = dimDatum.DatumWID
        LEFT JOIN dimPeilpunt ON FactChemischeMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
        ) CHM1
      LEFT JOIN (
        SELECT
            FactChemischeMeting.*,
            Jaardag_Nummer/365.0 + Jaar AS time,
            SUBSTRING(PeilpuntCode, 1, 3) AS region,
            dimPeilpunt.PeilpuntCode AS pp_code
        FROM FactChemischeMeting
        LEFT JOIN dimDatum ON FactChemischeMeting.DatumWID = dimDatum.DatumWID
        LEFT JOIN dimPeilpunt ON FactChemischeMeting.PeilpuntWID = dimPeilpunt.PeilpuntID
        ) CHM2
        ON (CHM1.ChemVarWID = CHM2.ChemVarWID)
       AND (CHM1.region = CHM2.region)
      WHERE CHM1.ChemVarWID IN (6, 7, 9)
        AND CHM2.MetingStatusCode = CHM1.MetingStatusCode
        AND CHM2.DatumWID >= CHM1.DatumWID
        AND (CHM2.time - CHM1.time) < (30.0/365)
        AND (CHM2.ChemMetingWID != CHM1.ChemMetingWID)
        AND (CHM1.PeilpuntWID < CHM2.PeilpuntWID)
        AND CHM1.MetingStatusCode IN ('VLD', 'SDV', 'UNK')
      ) AS ChemRepeated
    LEFT JOIN (
      SELECT
        ChemVarWID,
        ChemVarCode AS compound,
        ChemVarEenheid AS unit
      FROM DimChemVar
      ) AS DCV
      ON DCV.ChemVarWID = ChemRepeated.chemvar_id
      ")
  )
write_parquet(data_raw, sink = "chem_couplingdata_arrow.parquet")
```


```{r}
data_raw <- read_parquet("chem_couplingdata_arrow.parquet")
glimpse(data_raw)
```


## obswell locations
On the other hand, for the query of spatial information, `get_locs` will be fine.

```{r}
area_code_selection <- data_raw %>%
  pull(region) %>%
  unique

locs <- get_locs(
  watina_dwh,
  obswells = TRUE, 
  collect = TRUE
) 
locs %>% count
```


```{r}
test_locs <- locs %>%
  filter(area_code == "BUI") %>%
  select(obswell_code, loc_code, x, y) %>%
  arrange(obswell_code)
test_data <- data_raw %>%
  filter(region == "BUI") %>%
  arrange(pp1)


```


## double-joining locs
Comparing two locations, I will join the locs twice: 

```{r}
data <- data_raw %>%
  left_join(
    select(locs, obswell_code, x, y),
    by = join_by(pp1 == obswell_code),
    relationship = "many-to-one"
  ) %>% rename(c("x1" = x, "y1" = y)) %>%
  left_join(
    select(locs, obswell_code, x, y),
    by = join_by(pp2 == obswell_code),
    relationship = "many-to-one"
  ) %>% rename(c("x2" = x, "y2" = y))

glimpse(data)
```



```{r}
locs %>%
  filter(obswell_code == 6100)
```
