---
title: "Spatial Coupling of Observation Wells"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("watina")
library("n2khab")
library("sf")
library("stringr")
library("ggplot2")
library("mapview")

source("./watina_migration.R")

watina_dwh <- connect_watina()

```

# Summary
In this document, I will assemble and test components and functions to extract information which is potentially relevant to infer spatial coupling of groundwater observation wells.

Those information are:

- spatial distance
- distance to water body / stream
- height TAW difference
- watina region, eco-region, n2khab-type
- soil class
- quality of existing well

As a bonus, this will work on the _new_ watina data warehouse.
Another bonus is the extraction of unit tests for `watina` usage.

```{r}
get_db_table_list(watina_dwh)
```

# Spatial Distance
## data
A use case to query spatial distance is as follows:

- take a given location (observation well candidate)
- get existing location from `watina` 
- find locations in the area

step 1: migrate `watina::get_locs()`
```{r}
source("./watina_migration.R")

obswells_db <- get_locs(
  watina_dwh,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  area_codes = c("HOP"),
  obswells = TRUE
) %>% mutate(line_nr = row_number())

glimpse(obswells_db)
# obswells_db %>% count
```
TODO: unit test - query of a test area will not be empty


## orientation

```{r}

source("./watina_migration.R")
soilmap_raw <-
  read_soilmap()

soilmap <- soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  )

joined_spatial <- st_join(
  watina::as_points(
    obswells_db %>%
      collect()
  ), soilmap
) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown"))


joined_spatial %>%
  mapview(
    map.types = c("OpenStreetMap", "OpenTopoMap"),
    zcol = "soilclass"
  )


```


## spatial relation
NEXT: (i) bbox; 
      (ii) circular bounding

### bbox
Query locations within a bounding box 

```{r}

source("./watina_migration.R")

x_test <- 148600
y_test <- 208900
radius <- 1000 # (+/- m)
hopo_boxed <- st_bbox(c(
  xmin = x_test - radius,
  xmax = x_test + radius,
  ymin = y_test - radius,
  ymax = y_test + radius
  ),
  crs = st_crs(31370))

obswells_db <- get_locs(
  watina_dwh,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  # area_codes = c("HOP"),
  bbox = hopo_boxed,
  obswells = TRUE
) %>% mutate(line_nr = row_number())

glimpse(obswells_db)

```

### circular

- [related, on stack overflow](https://stackoverflow.com/questions/48650274/spatial-efficient-way-of-finding-all-points-within-x-meters-of-a-point)

```{r}

center <- c(x_test, y_test)
hopo_points <- watina::as_points(obswells_db %>% collect())

focus_sites <- points_within_radius(center, 100, hopo_points )
```

## Usage

```{r}
source("./watina_migration.R")

get_locs_in_radius <- function(conn, point, radius = 10, ...) {

  # create bounding box from point and radius
  bbox <- st_bbox(c(
    xmin = point[[1]] - radius,
    xmax = point[[1]] + radius,
    ymin = point[[2]] - radius,
    ymax = point[[2]] + radius
    ), crs = st_crs(31370))

  # query with bbox
  close_sites <- get_locs(
    conn,
    bbox = bbox,
    obswells = FALSE,
    collect = TRUE,
    ...
  )

  # convert to points
  close_sites <- watina::as_points(close_sites)

  # refine: radius
  closer_sites <- points_within_radius(point, radius, close_sites )

  return(closer_sites)
}

x_test <- 148593
y_test <- 208862
rad_test <- 1000 # (+/- m)

hopo_points <- get_locs_in_radius(conn = watina_dwh,
    point = c(x_test, y_test),
    radius = rad_test,
    filterdepth_range = c(0, 5),
    filterdepth_guess = FALSE,
    loc_type = "P",
    loc_validity = c("VLD", "ENT")
    )

glimpse(hopo_points)

```


# distance to water body / stream
## data source?

- https://www.vmm.be/water/projecten/grachtenvalidatie
- https://www.vlaanderen.be/DataCatalogRecord/408b8ef8-b731-4358-a5b0-9e028e074180
- https://zenodo.org/records/7440931 n2khab data!

```{r}
watersurf_raw <-
  read_watersurfaces()
```


Looking at the example region:
```{r}

x_test <- 148593
y_test <- 208862

hopo_watersurfaces <- points_within_radius(c(x_test, y_test),
                                       radius = 1200,
                                       watersurf_raw)

m1 = mapview(hopo_watersurfaces,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "usage"
  )
m2 = mapview(hopo_points,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "filterdepth"
  )

m1+m2
```

## distance

... narrowing the focus region a bit

```{r}
x_test <- 148593
y_test <- 208862

hopo_watersurfaces <- points_within_radius(c(x_test, y_test),
                                       radius = 100,
                                       watersurf_raw)


hopo_points <- get_locs_in_radius(conn = watina_dwh,
    point = c(x_test, y_test),
    radius = 60,
    filterdepth_range = c(0, 5),
    filterdepth_guess = FALSE,
    loc_type = "P",
    loc_validity = c("VLD", "ENT")
    )


cross_distances <- st_distance(hopo_points, hopo_watersurfaces)
knitr::kable(cross_distances)
```


# height TAW difference

## Digital Elevantion Models

- https://r.geocompx.org/raster-vector
- https://cran.r-project.org/web/packages/rgugik/vignettes/DEM.html



## Digitaal Hoogtemodel Vlaanderen 
- https://www.vlaanderen.be/digitaal-vlaanderen/onze-diensten-en-platformen/earth-observation-data-science-eodas/het-digitaal-hoogtemodel
- raster https://download.vlaanderen.be/product/937-digitaal-hoogtemodel-vlaanderen-ii-dsm-raster-1-m
- raster https://download.vlaanderen.be/product/939-digitaal-hoogtemodel-vlaanderen-ii-dtm-raster-1-mper
- punten https://download.vlaanderen.be/product/56/configureer
- https://assets.vlaanderen.be/image/upload/v1676620086/Infosessie_BVKOpenData_LiDAR_21022017_1_p34z1s_uiihcm.pdf

others:

- Grootschaalig Referentiebestand of Basiskaart (GRB): https://www.vlaanderen.be/digitaal-vlaanderen/onze-diensten-en-platformen/basiskaart-vlaanderen-grb 
- https://pydov.readthedocs.io/en/stable/tutorials.html#tutorials
- topografie  https://nl-be.topographic-map.com/map-lqq9m/Vlaanderen/



# watina region, eco-region, n2khab-type
[reference](https://github.com/inbo/n2khab/blob/main/vignettes/v025_geospatial_hab.Rmd)

- **`habitatmap`** ([Zenodo-link](https://doi.org/10.5281/zenodo.3354381)): geospatial polygons of BWK and Natura 2000 habitat types in the Flemish Region, originally published by the Research Institute for Nature and Forest (INBO) [@de_saeger_biologische_2020] and distributed by 'Informatie Vlaanderen'. 
- **`habitatmap_stdized`** ([Zenodo-link](https://doi.org/10.5281/zenodo.3355192)):
derived from `habitatmap`.
This datasource only contains polygons with habitat (sub)type or RIB codes.
All codes conform to the `types` reference list (see `read_types()`).
The datasource is a GeoPackage with essentially two objects: 1) the geospatial polygons; 2) a long-formatted table with the associated types (multiple records can occur per `polygon_id`).
For more details see `read_habitatmap_stdized()`. 

# soil class

# quality of existing well


# misc
The `n2khab_data` for `soilmap` is needed.
```{r, eval=FALSE}
n2khab_data_path <- locate_n2khab_data()
soilmap_path <- file.path(n2khab_data_path, "10_raw/soilmap")
dir.create(soilmap_path)
download_zenodo(
  doi = "10.5281/zenodo.3387008",
  path = soilmap_path,
  parallel = TRUE
)
```
