---
title: "Spatial Coupling of Observation Wells"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("watina")
library("n2khab")
library("terra")
library("sf")
library("stringr")
library("ggplot2")
library("mapview")

source("./watina_migration.R")
source("./spatial_helpers.R")

watina_dwh <- connect_watina()


n2khab_data_path <- locate_n2khab_data()

# options(browser = "librewolf") # a rather personal thing
```

# Summary
In this document, I will assemble and test components and functions to extract information which is potentially relevant to infer spatial coupling of groundwater observation wells.
```{r}
# (i) existing watina
# (ii) co-located sampling
```


Those information are:

- spatial distance
- distance to water body / stream
- height TAW difference
- watina region, eco-region, n2khab-type
- soil class
- quality of existing well

As a bonus, this will work on the _new_ watina data warehouse.
Another bonus is the extraction of unit tests for `watina` usage.

```{r}
get_db_table_list(watina_dwh)
```

# Spatial Distance
## data
A use case to query spatial distance is as follows:

- take a given location (observation well candidate)
- get existing location from `watina` 
- find locations in the area

step 1: migrate `watina::get_locs()`
```{r}
source("./watina_migration.R")

obswells_db <- get_locs(
  watina_dwh,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  area_codes = c("HOP"),
  obswells = TRUE
) %>% mutate(line_nr = row_number())

glimpse(obswells_db)
# obswells_db %>% count
```
TODO: unit test or vignette - query of a test area will not be empty


## orientation

```{r}

source("./watina_migration.R")
soilmap_raw <-
  read_soilmap()

soilmap <- soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  )

joined_spatial <- st_join(
  watina::as_points(
    obswells_db %>%
      collect()
  ), soilmap
) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown"))


joined_spatial %>%
  mapview(
    map.types = c("OpenStreetMap", "OpenTopoMap"),
    zcol = "soilclass"
  )


```


## spatial relation
NEXT: (i) bbox; 
      (ii) circular bounding

### bbox
Query locations within a bounding box 

```{r}

source("./watina_migration.R")

x_test <- 148600
y_test <- 208900
radius <- 1000 # (+/- m)
hopo_boxed <- st_bbox(c(
  xmin = x_test - radius,
  xmax = x_test + radius,
  ymin = y_test - radius,
  ymax = y_test + radius
  ),
  crs = st_crs(31370))

obswells_db <- get_locs(
  watina_dwh,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  # area_codes = c("HOP"),
  bbox = hopo_boxed,
  obswells = TRUE
) %>% mutate(line_nr = row_number())

# glimpse(obswells_db)

```

### circular

- [related, on stack overflow](https://stackoverflow.com/questions/48650274/spatial-efficient-way-of-finding-all-points-within-x-meters-of-a-point)

```{r}

center <- c(x_test, y_test)
hopo_points <- watina::as_points(obswells_db %>% collect())

focus_sites <- geometry_within_radius(center, 100, hopo_points )
```

## Usage

```{r}
source("./watina_migration.R")

get_locs_in_radius <- function(conn, point, radius = 10, ...) {

  # create bounding box from point and radius
  bbox <- st_bbox(c(
    xmin = point[[1]] - radius,
    xmax = point[[1]] + radius,
    ymin = point[[2]] - radius,
    ymax = point[[2]] + radius
    ), crs = st_crs(31370))

  # query with bbox
  close_sites <- get_locs(
    conn,
    bbox = bbox,
    obswells = FALSE,
    collect = TRUE,
    ...
  )

  # convert to points
  close_sites <- watina::as_points(close_sites)

  # refine: radius
  closer_sites <- geometry_within_radius(point, radius, close_sites )

  return(closer_sites)
}

x_test <- 148593
y_test <- 208862
rad_test <- 1000 # (+/- m)

hopo_points <- get_locs_in_radius(conn = watina_dwh,
    point = c(x_test, y_test),
    radius = rad_test,
    filterdepth_range = c(0, 5),
    filterdepth_guess = FALSE,
    loc_type = "P",
    loc_validity = c("VLD", "ENT")
    )

# glimpse(hopo_points)

```


# TODO distance to water body / stream
## data source?

- [nvt.] https://inbo.github.io/n2khab/reference/read_habitatstreams.html
- [?] vmm grachtenkaart https://www.vmm.be/water/projecten/grachtenvalidatie
- https://www.vlaanderen.be/DataCatalogRecord/408b8ef8-b731-4358-a5b0-9e028e074180
- https://zenodo.org/records/7440931 n2khab data!

- digitale atlas waterlopen https://vha.waterinfo.be/arcgis/rest/services/digitale_atlas/MapServer
- GRB https://geo.api.vlaanderen.be/GRB/wms?request=getcapabilities
  - baangracht, gracht, GRB_WGR

- [!] Vlaamse Hydrografische Atlas - Waterlopen https://zenodo.org/records/5584530



There are multiple valid sources for close-by water.
```{r}
source("./water_sources.R")
watersurf_raw <- read_watersurfaces()
waterstreams_raw <- read_habitatstreams()
watercourses_raw <- read_watercourses()
watercourses_100m <- read_watercourse_100mseg(element = "lines")

```


Looking at the example region:
```{r}

source("./spatial_helpers.R")
source("./water_sources.R")

x_test <- 148593
y_test <- 208862
# x_test <- 149111.968
# y_test <- 208712.173

hopo_watersurfaces <- geometry_within_radius(c(x_test, y_test),
                                       radius = 1200,
                                       watersurf_raw)

hopo_morewater <- geometry_within_radius(c(x_test, y_test),
                                       radius = 1200,
                                       watercourses_raw)
hopo_evenmorecourses <- geometry_within_radius(c(x_test, y_test),
                                       radius = 1200,
                                       watercourses_100m)

# wells
m1 = mapview(hopo_points,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "filterdepth"
  )

# water surfaces
m2 = mapview(hopo_watersurfaces,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "usage"
  )
# Schelde is not in there! -> waterlopen

# water courses
m3 = mapview(hopo_morewater,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "rank"
  )
m4 = mapview(hopo_morewater,
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  zcol = "rank"
  )

# show all on a map
m1+m2+m3+m4
 
```

## distance


```{r}
source("./water_sources.R")
x_test <- 148593
y_test <- 208862
# x_test <- 149111.968
# y_test <- 208712.173


hopo_points <- get_locs_in_radius(conn = watina_dwh,
    point = c(x_test, y_test),
    radius = 100,
    filterdepth_range = c(0, 5),
    filterdepth_guess = FALSE,
    loc_type = "P",
    loc_validity = c("VLD", "ENT")
    )
```


Loading all water data sources takes some time... to be reviewed when implementing the whole procedure.
```{r}
all_wata <- load_all_water_sources()
```


... narrowing the focus region a bit
```{r}
hopo_waterdata <- narrow_sources_radius(
  all_wata,
  c(x_test, y_test),
  radius = 1000
)

```


compute cross_distances
```{r}
cross_distances <- collect_combined_distances(hopo_points, hopo_waterdata)

closest_distances <- minimum_cross_distance(cross_distances)

knitr::kable(closest_distances)
```


# Height TAW Difference

## general info

- https://r.geocompx.org/raster-vector
- https://r.geocompx.org/read-write#geographic-web-services
- https://cran.r-project.org/web/packages/rgugik/vignettes/DEM.html

_Edit 09/10/2024:_ Better use web services! (suggestion Thierry)

- https://inbo.github.io/tutorials/tutorials/spatial_wfs_services
- https://github.com/inbo/inbospatial/pull/12


## Digitaal Hoogtemodel Vlaanderen 

[source](https://www.vlaanderen.be/digitaal-vlaanderen/onze-diensten-en-platformen/earth-observation-data-science-eodas/het-digitaal-hoogtemodel)



- documentation:  https://assets.vlaanderen.be/image/upload/v1676620086/Infosessie_BVKOpenData_LiDAR_21022017_1_p34z1s_uiihcm.pdf


**alternatives:**
- Grootschaalig Referentiebestand of Basiskaart (GRB): https://www.vlaanderen.be/digitaal-vlaanderen/onze-diensten-en-platformen/basiskaart-vlaanderen-grb 
- https://pydov.readthedocs.io/en/stable/tutorials.html#tutorials
- topografie  https://nl-be.topographic-map.com/map-lqq9m/Vlaanderen/


## Elevation from Web Services
... using the DHMV WCS for DTM (1m resolution)

```{r}

# functions are in another dedicated file
source("./spatial_helpers.R")


# re-use the bounding box from above:
x_test <- c(148600)
y_test <- c(208900)
radius <- 10 # (+/- m)

hopo_sf <- st_as_sf(
  data.frame(x_test, y_test),
  coords = c("x_test", "y_test"),
  crs = 31370
)
# point_sf <- hopo_sf
# margin <- radius
# wcs = "DHMV"
# resolution = 1
# layername <- "DHMVII_DTM_1m"

elevation <- query_elevation_wcs(
  hopo_sf,
  margin = radius
)

elevation

# debug_once??
```

Does this work for a collection of points?
```{r}

hopo_wells_sf <- st_as_sf(hopo_points, coords = c("x", "y"), crs = 31370)
# point_sf <- hopo_wells_sf
elevation_wells <- query_elevation_wcs(
  hopo_wells_sf,
  margin = radius
) %>%
  select(-id) %>%
  st_join(hopo_wells_sf)

elevation_wells
```

It does. Good times.



## OBSOLETE: Elevation from Points
An "offline" approach, downloading all Flanders elevation point data (LiDAR raw data, including canopy).

- point data https://download.vlaanderen.be/product/56/configureer


There is a linux shell script to collect all data in an sqlite database: `unzip_dhmv_points.sh` (>40GB).
This sqlite file is stored as `dhmv_points.db` in the current working directory.



```{r, eval=FALSE}
require("sf")
require("DBI")

dhmv_conn <- dbConnect(RSQLite::SQLite(), file.path("./dhmv_points.db"))

source("./watina_migration.R")
source("./spatial_helpers.R")

# re-use the bounding box from above:
x_test <- c(148600)
y_test <- c(208900)
radius <- 10 # (+/- m)

hopo_sf <- st_as_sf(data.frame(x_test, y_test), coords = c("x_test", "y_test"), crs = 31370)

elevation <- query_elevation_sqlite(
  hopo_sf,
  conn = dhmv_conn,
  frame = radius
)


```


```{r, eval=FALSE}

hopo_wells_sf <- st_as_sf(hopo_points, coords = c("x", "y"), crs = 31370)
# point_sf <- hopo_wells_sf
well_elevations <- query_elevation_sqlite(
  hopo_wells_sf,
  conn = dhmv_conn,
  frame = radius
)

```

Great, we can get elevations!

```{r, eval=FALSE}
dbDisconnect(dhmv_conn)
``` 



# TODO watina region, eco-region, n2khab-type

## evaluate sources

(via `n2khab/read_adminareas`)
(via `n2khab/read_ecoregions`)

ecoregions, eco-district (hierarchisch)
◀ @FV
 * inlezen <https://github.com/inbo/n2khab-mne-design/blob/5f5b0ae0bbd33f630ab4f478d1d4b4c70bb4c9cb/110_design_groundwater/020_design_elaborate/210_data_targetpop.Rmd#L409-L416>
 * aan puntlocaties hangen en een probleempje oplossen <https://github.com/inbo/n2khab-mne-design/blob/5f5b0ae0bbd33f630ab4f478d1d4b4c70bb4c9cb/110_design_groundwater/020_design_elaborate/220_data_env.Rmd#L390-L403>
 * html-downloadlink van dit bookdownproject: data-exploratie <https://drive.google.com/uc?export=download&id=15ZNdeZEQ5Ul8yKP0XA62gIovZ2pUUcYC> ter
   voorbereiding van poweranalysen (2020-2021)


**NOT:**
- **`habitatmap`** ([Zenodo-link](https://doi.org/10.5281/zenodo.3354381)): geospatial polygons of BWK and Natura 2000 habitat types in the Flemish Region, originally published by the Research Institute for Nature and Forest (INBO) [@de_saeger_biologische_2020] and distributed by 'Informatie Vlaanderen'. 
- **`habitatmap_stdized`** ([Zenodo-link](https://doi.org/10.5281/zenodo.3355192)):
derived from `habitatmap`.
This datasource only contains polygons with habitat (sub)type or RIB codes.
All codes conform to the `types` reference list (see `read_types()`).
The datasource is a GeoPackage with essentially two objects: 1) the geospatial polygons; 2) a long-formatted table with the associated types (multiple records can occur per `polygon_id`).
For more details see `read_habitatmap_stdized()`. 

[reference](https://github.com/inbo/n2khab/blob/main/vignettes/v025_geospatial_hab.Rmd)



# TODO soil class
(heavy, light, peat, unknown; as before)


# TODO quality of existing well

- matching filter depth
- validation status
- amount of data
- ...


# TODO stitch this all into a single object

e.g. `well_candidate <- CandidatePoint$new(xy)`

- contains a set of locations
- `sf.Points`; indifferent: existing watina or novel location
- can query all the information above


# disconnect databases

```{r}
# dbDisconnect(watina_dwh)
# dbDisconnect(dhmv_conn)
``` 


# misc
Some `n2khab_data` sources are required:

- `soilmap`:
```{r, eval=FALSE}
n2khab_data_path <- locate_n2khab_data()
soilmap_path <- file.path(n2khab_data_path, "10_raw/soilmap")
dir.create(soilmap_path)
download_zenodo(
  doi = "10.5281/zenodo.3387008",
  path = soilmap_path,
  parallel = TRUE
)
```

 
- ... and watercourses:
```{r, eval=FALSE}
watercourse_simple_path <- file.path(n2khab_data_path, "20_processed/watercourse_100mseg")
dir.create(watercourse_simple_path, recursive = TRUE)
download_zenodo(
  doi = "10.5281/zenodo.4452577",
  path = watercourse_simple_path
)
```

- ... and watercourses:
```{r, eval=FALSE}
n2khab_data_path <- locate_n2khab_data()
watercourse_raw_path <- file.path(n2khab_data_path, "10_raw/watercourses")
dir.create(watercourse_raw_path, recursive = TRUE)
download_zenodo(
  doi = "10.5281/zenodo.4420905",
  path = watercourse_raw_path
)
```
