---
title: "Effects of phab-correction on the relation between MNE base sampling frame + samples and MHQ samples"
date: today
published-title: Date
params:
  lang: nl
engine: knitr
format:
  html:
    embed-resources: true
    include-in-header: ../../../header_qmd.html
    theme: cosmo
    keep-md: true
    number-sections: true
    code-fold: true
    fig-width: 10
    fig-height: 5
    df-print: paged
    toc: true
    toc-location: left
    toc-expand: true
---

```{r setup, message = FALSE}
options(scipen = 999)
conflictRules("n2khab", exclude = c("read_schemes", "read_scheme_types"))
conflictRules("git2rdata", exclude = c("pull", "meta"))
conflictRules("tidyr", exclude = "extract")
conflictRules("knitr", exclude = "spin")
conflictRules("kableExtra", exclude = "group_rows")
lang <- params$lang
library(rprojroot)
projroot <- find_root(is_rstudio_project)
gitroot <- find_root(is_git_root)
source(file.path(projroot, "R/functions.R"))
datapath <- file.path(projroot, "data")
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(purrr)
library(sf)
library(units)
library(lubridate)
library(terra)
library(git2rdata)
library(n2khab)
library(n2khabmon)
opts_chunk$set(
  echo = TRUE,
  dpi = 300
)
our_kbl <- function(df) {
  df %>% 
    kbl(table.attr = 'data-quarto-disable-processing="true"') %>% 
    kable_styling(full_width = FALSE, bootstrap_options = "striped")
}
```


## Data setup

The two involved RData files have been created by running following shell commands in the `010_prototyping_sampling_and_revisit_design` directory at commit `ee267a53`.

```bash
Rscript -e 'bookdown::render_book("index.Rmd", "bookdown::html_document2", params = list(regenerate_binary_files = FALSE, active_modules = c("pan_effectmon_flanders", "pan_effectmon_sac5_custommeas"), update_spatiotemp_gsheet = FALSE, save_rdata = TRUE))'

Rscript -e 'bookdown::render_book("index.Rmd", "bookdown::html_document2", params = list(regenerate_binary_files = FALSE, active_modules = c("pan_effectmon_flanders", "pan_effectmon_sac5_custommeas"), update_spatiotemp_gsheet = FALSE, save_rdata = TRUE, phab_correct = FALSE))'
```

Lazy loading the objects of these RData files in separate environments.

```{r}
prepare_lazy_get(
  file.path(datapath, "binary/results/objects_panflpan5.RData"),
  new_envir_name = "panflpan5"
)
prepare_lazy_get(
  file.path(datapath, "binary/results/objects_panflpan5_nophabcorrection.RData"),
  new_envir_name = "panflpan5_nophabcorrection"
)
```

Loading specific objects in the global environment.

```{r}
# ls(envir = panflpan5)
modules <- get("modules", envir = panflpan5)
mhq_samples <- get("mhq_samples", envir = panflpan5)
bsf_corrected <- get("stratum_grts_n2khab_phabcorrected", envir = panflpan5)
bsf_uncorrected <- get("stratum_grts_n2khab_phabcorrected", envir = panflpan5_nophabcorrection)
n2khab_types_expanded_properties <- get("n2khab_types_expanded_properties", envir = panflpan5)
n2khab_targetpops <- get("n2khab_targetpops", envir = panflpan5)
n2khab_types <-
  n2khab_targetpops %>%
  distinct(type) %>%
  arrange(type)
n2khab_strata <- get("n2khab_strata", envir = panflpan5)
active_strata <- get("module_targetpops_strata", envir = panflpan5) %>% distinct(stratum)
```

Reading a separate source of considered terrestrial population units in MHQ (`mhq_terr_popunits`).

```{r}
mhq_terr_datapath <- file.path(
  dirname(gitroot),
  "n2khab-sample-admin/data/mhq_terr/rapportage2025"
)
mhq_terr_popunits <-
  read_vc("mhq_terr_popunits", root = mhq_terr_datapath) %>%
  as_tibble() %>% 
  mutate(
    type = factor(type, levels = levels(n2khab_types$type)),
    source = factor(source),
    across(where(is.numeric), as.integer)
  ) %>% 
  select(-sac, -phab, -polygon_id) %>% 
  rename(
    grts_address = grts_ranking,
    grts_address_drawn = grts_ranking_draw
  )
```


## Introduction

In this notebook we compare the effect

We specifically consider this for the samples generated when following modules are active:

```{r}
#| echo: FALSE
modules %>% 
  select(module, shortname_nl) %>% 
  our_kbl()
```


## Comparing the base sampling frame with the Flemish MHQ samples

Current MHQ samples, restricted to those sampled by a GRTS (alike) strategy, were acquired and copied under `data/text/raw/mhq_samples`.
The essential information has been integrated into a single data frame:

```{r}
glimpse(mhq_samples)
```

Both columns are also represented in the phab-corrected and uncorrected base sampling frame.

We only consider terrestrial types to make comparisons, so we restrict all objects to these.
We also restrict strata to the ones covered by the active modules (this essentially removes RIBs from the base sampling frame).

```{r}
restrict_to_active_terrestrial_strata <- function(df) {
  df %>% 
    semi_join(active_strata, join_by(stratum)) %>% 
    inner_join(
      n2khab_strata,
      join_by(stratum),
      relationship = "many-to-one",
      unmatched = c("error", "drop")
    ) %>% 
    semi_join(
      n2khab_types_expanded_properties %>% 
        filter(hydr_class != "HC3"),
      join_by(type)
    ) %>% 
    select(-type)
}
mhq_samples_terr <- restrict_to_active_terrestrial_strata(mhq_samples)
bsf_corrected_terr <- restrict_to_active_terrestrial_strata(bsf_corrected)
bsf_uncorrected_terr <- restrict_to_active_terrestrial_strata(bsf_uncorrected)
```

### Matching of types (regardless of phab correction)

The (remaining) types covered by the Flemish MHQ samples are:

```{r}
mhq_samples_terr %>% 
  count(stratum, sort = TRUE) %>% 
  our_kbl()
```

A short note on terminology.
In MHQ, small clusters of a few strata, or a single stratum, corresponds to a single scheme.
This is why we talk about 'samples' (plural): one sample corresponds to a single scheme.
Moreover, one stratum can belong to only one scheme in the case of MHQ, unlike MNE.
The term 'stratum' comes from the MNE schemes, where each scheme mostly consists of many strata, for which a stratified design is made.
Since the same entities (types or subdivisions thereof) are used in both MHQ and MNE, we just use the term stratum in this case.
In the case of terrestrial types, the stratum level equals the type level, so that both terms can be used interchangeably.

Several (remaining) types in the sampling frame are not present in the considered MHQ samples:

```{r}
bsf_uncorrected_terr %>% 
  count(stratum, sort = TRUE) %>% 
  anti_join(mhq_samples_terr %>% distinct(stratum), join_by(stratum)) %>% 
  our_kbl()
```

Are all types of the MHQ samples are present in the sampling frame?

```{r}
bsf_uncorrected_terr %>% 
  distinct(stratum) %>% 
  anti_join(mhq_samples_terr %>% distinct(stratum), ., join_by(stratum)) %>% 
  nrow() == 0
```

To make comparisons, we only consider types that are present in both data sources.

### Matching of sampling units

The distribution of all (unique) stratum x GRTS address combinations from the MHQ samples is calculated, regarding their membership in each of the base samplingframe (bsf) variants.

```{r}
bsf_match_counts <- 
  mhq_samples_terr %>% 
  left_join(
    bsf_uncorrected_terr %>% 
      select(stratum, grts_address) %>% 
      mutate(in_bsf_uncorrected = TRUE),
    join_by(stratum, grts_address)
  ) %>% 
  left_join(
    bsf_corrected_terr %>% 
      select(stratum, grts_address) %>% 
      mutate(in_bsf_corrected = TRUE),
    join_by(stratum, grts_address)
  ) %>% 
  mutate(across(contains("in_bsf"), \(x) if_else(x, TRUE, FALSE, FALSE))) %>% 
  count(pick(contains("in_bsf")))

bsf_match_counts %>% 
  mutate(pct = round(n / sum(n) * 100, 1)) %>% 
  our_kbl()
```

- About 80% has a match in both the uncorrected and the corrected base sampling frame.
- About 5% is dropped in the corrected base sampling frame.
- The remaining 15% are sampling units missed by each base sampling frame.

We find (below) that quite a lot of MHQ sampling units is not present in the separate MHQ source `mhq_terr_popunits`.
This is a probable cause of sampling units missing in both variants of the (MNE) base sampling frame.
That is because it is the `mhq_terr_popunits` data source, notably its assessed data (constrained to cell-center assessments for most terrestrial types), that has been used to augment the base sampling frame in MNE with MHQ data.
The mismatch has been [reported as an issue](https://github.com/inbo/n2khab-sample-admin/issues/49).

Number of rows in the MHQ samples missing from `mhq_terr_popunits`:

```{r}
mhq_samples_terr %>% 
  anti_join(
    mhq_terr_popunits,
    # we used the drawn GRTS address in the MHQ samples to identify the units,
    # so we need to join accordingly
    join_by(stratum == type, grts_address == grts_address_drawn)
  ) %>% 
  nrow()
```

When comparing both variants of the base sampling frame, it makes more sense to ignore these missing locations.

```{r}
bsf_match_counts %>% 
  filter(in_bsf_uncorrected | in_bsf_corrected) %>% 
  mutate(pct = round(n / sum(n) * 100, 1)) %>% 
  our_kbl()
```

We conclude that the phab-correction leads to a loss of only about 6% of MHQ sampling locations in the sampling frame.
Hence, most that were still available for sampling, remain available for sampling after phab-correction.
This is due to the fact that non-replaced GRTS addresses have been set to phab = 100 at the level of the specific unit.
This introduces no bias in the sampling frame, since these are genuine matches at those locations, for which the local setting of phab = 100 is correct.







