---
title: "Revising sample sizes with regard to disturbance of rare types"
date: today
published-title: Date
csl: "`r if (curl::has_internet()) 'https://raw.githubusercontent.com/citation-style-language/styles/master/research-institute-for-nature-and-forest.csl' else system.file('research-institute-for-nature-and-forest.csl', package = 'INBOmd')`"
bibliography: ../../../refs_mne_design.yaml
params:
  lang: nl
engine: knitr
format:
  html:
    embed-resources: true
    include-in-header: ../../../header_qmd.html
    theme: cosmo
    keep-md: true
    number-sections: true
    code-fold: true
    fig-width: 10
    fig-height: 5
    df-print: paged
    toc: true
    toc-location: left
    toc-depth: 4
    toc-expand: true
---

```{r setup, message = FALSE}
options(scipen = 999)
conflictRules("git2rdata", exclude = c("pull", "meta"))
conflictRules("tidyr", exclude = "extract")
conflictRules("knitr", exclude = "spin")
conflictRules("kableExtra", exclude = "group_rows")
lang <- params$lang
library(rprojroot)
projroot <- find_root(is_rstudio_project)
gitroot <- find_root(is_git_root)
datapath <- file.path(projroot, "data")
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(lubridate)
library(ggplot2)
library(sf)
source(file.path(projroot, "R/functions.R"))
opts_chunk$set(
  echo = TRUE,
  dpi = 300,
  fig.align = "center"
)
our_kbl <- function(df) {
  df %>% 
    kbl(table.attr = 'data-quarto-disable-processing="true"') %>% 
    kable_styling(full_width = FALSE, bootstrap_options = "striped")
}
our_scroll_kbl <- function(df, width = NULL, height = NULL) {
  our_kbl(df) %>% 
    scroll_box(width = width, height = height)
}
scenario_name <- rmarkdown::yaml_front_matter(file.path(projroot, 'index.Rmd'))$params$sample_adjustment_scenario
path_ref <- file.path(datapath, "binary/results/objects_panflpan5.RData")
path_new <- file.path(
  datapath,
  str_glue("binary/results/objects_panflpan5_{scenario_name}.RData")
)
path_gpkg <- file.path(datapath, "binary/intermediate/spatial_sampling.gpkg")
```

**SCENARIO: `r scenario_name`**

## Data setup {-}

The two involved RData files have been created by running following shell command in the `010_prototyping_sampling_and_revisit_design` directory, at tag `poc_0.13.0` and at the specific branch with a sample scenario, respectively.

```bash
Rscript -e 'bookdown::render_book("index.Rmd", "bookdown::html_document2", params = list(regenerate_binary_files = FALSE, active_modules = c("pan_effectmon_flanders", "pan_effectmon_sac5_custommeas"), update_spatiotemp_gsheet = FALSE, save_rdata = TRUE))'
```

Lazy loading the objects of these RData files in separate environments.

```{r}
prepare_lazy_get(path_ref, new_envir_name = "ref")
prepare_lazy_get(path_new, new_envir_name = scenario_name)
```

Loading specific objects in the global environment.

```{r}
# ls(envir = ref)
sps_ref <- get("scheme_moco_ps_stratum_sppost_spsamples_sf", envir = ref)
sps_new <- get(
  "scheme_moco_ps_stratum_sppost_spsamples_sf",
  envir = eval(str2lang(scenario_name))
)
ssizes_ref <- get("module_domain_scheme_ps_stratum_sample_size", envir = ref)
ssizes_new <- get(
  "module_domain_scheme_ps_stratum_sample_size",
  envir = eval(str2lang(scenario_name))
)
cal_ref <- get("fag_stratum_grts_calendar", envir = ref)
cal_new <- get(
  "fag_stratum_grts_calendar",
  envir = eval(str2lang(scenario_name))
)


moco_ssizes_new <- get(
  "scheme_moco_ps_dom_stratum_sample_size",
  envir = eval(str2lang(scenario_name))
)
targetsizes_ref <- get("module_domain_scheme_ps_stratum_target_sample_size", envir = ref)
targetsizes_new <- get(
  "module_domain_scheme_ps_stratum_target_sample_size",
  envir = eval(str2lang(scenario_name))
)

type_properties <- get("n2khab_types_expanded_properties", envir = ref)
mhq_scheme_category <- get("mhq_scheme_category", envir = ref)
n2khab_strata <- get("n2khab_strata", envir = ref)
dom_scheme_stratum_nunits <-
  get("submod_dom_scheme_ssf_stratum_nunits", envir = ref) %>%
  distinct(domain, scheme, stratum, nunits)
```


Defining cell types & non-cell types.

```{r}
cell_types <-
  type_properties %>%
  filter(
    str_detect(grts_join_method, "cell"),
    type != "7140_mrd"
  )
non_cell_types <-
  type_properties %>%
  filter(
    !str_detect(grts_join_method, "cell") | type == "7140_mrd"
  )
```

Loading more functions.

```{r}
source(file.path(projroot, "R/ssize_scenario_functions.R"))
```


## Comparing sample sizes

Grand totals per scheme: reference.

```{r}
get_total_sample_sizes(sps_ref)
```

Grand totals per scheme: new.

```{r}
get_total_sample_sizes(sps_new)
```

Same, but limited to non-cell types. Reference.

```{r}
get_total_sample_sizes(sps_ref, non_cell_types)
```

Same, but limited to non-cell types. New.

```{r}
get_total_sample_sizes(sps_new, non_cell_types)
```

Yearly totals, distinguishing modules. Reference.

```{r}
aggregate_sample_size(
  ssizes_ref,
  "sp_sample_size_all_panels_stratum",
  mhq_scheme_category
) %>%
  print(n = Inf)
```

Yearly totals, distinguishing modules. New.

```{r}
aggregate_sample_size(
  ssizes_new,
  "sp_sample_size_all_panels_stratum",
  mhq_scheme_category
) %>%
  print(n = Inf)
```

Table of sample size differences per module, domain, scheme and stratum (only for cell types).

```{r}
compare_ssizes_per_stratum(ssizes_new) %>% 
  our_scroll_kbl(width = "100%", height = "500px")
```

<br/>

Percentage of changed sample sizes in above table.

```{r}
#| paged.print: FALSE
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(ssize_differs = ssize_stratum_altered != ssize_stratum) %>%
  count(ssize_differs)
```

Percentage of changed sample sizes in above table, depending on whether sample was truncated using a maximum population proportion, or not.

```{r}
#| paged.print: FALSE
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(ssize_differs = ssize_stratum_altered != ssize_stratum) %>%
  count(spss_stratum_truncated, ssize_differs)
```

Quantiles of absolute and relative sample size differences.

```{r}
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(ssize_diff = ssize_stratum_altered - ssize_stratum) %>%
  pull(ssize_diff) %>%
  quantile(seq(0, 1, 0.1))
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(
    ssize_diff_rel = round(
      (ssize_stratum_altered - ssize_stratum) / ssize_stratum,
      2
    )) %>%
  pull(ssize_diff_rel) %>%
  quantile(seq(0, 1, 0.1))
```


```{r}
#| echo: FALSE
#| eval: FALSE
# investigate unplanned but obtained sample size differences (they are always
# lower): this is the consequence of the lower target sample size ranges in the
# new (smaller) precision pools (i.e. after excluding strata with sample size
# limitation), leading to lower coef for infinite sample size and consequently
# lower FPC-corrected sample size. The difference is limited though.
if (FALSE) {
  compare_ssizes_per_stratum(ssizes_new) %>%
    mutate(ssize_differs = ssize_stratum_altered != ssize_stratum) %>%
    filter(!spss_stratum_truncated, ssize_differs)
}
```

```{r}
#| echo: FALSE
#| eval: FALSE
# comparing and checking unplanned differences in target sample size: apart from
# left-out-types from the densification submodules, differences have to do with
# the different weights that the domains get in each scheme, within the
# densification submodule in pan5; hence this is restricted to the types that
# are densified.
if (FALSE) {
  targetsizes_new %>%
    select(
      module,
      domain,
      scheme,
      panel_set,
      stratum,
      nunits,
      targsize_stratum_altered = sp_sample_size_all_panels_stratum,
      spss_stratum_truncated
    ) %>%
    inner_join(
      targetsizes_ref %>%
        select(
          module,
          domain,
          scheme,
          panel_set,
          stratum,
          targsize_stratum = sp_sample_size_all_panels_stratum
        ),
      join_by(module, domain, scheme, panel_set, stratum),
      relationship = "one-to-one",
      unmatched = "error"
    ) %>%
    filter(!spss_stratum_truncated) %>%
    mutate(
      targsize_differs = targsize_stratum_altered != targsize_stratum,
      targsize_diff_rel = (targsize_stratum_altered - targsize_stratum) / targsize_stratum
    ) %>%
    filter(targsize_differs) %>%
    summarize(
      targsize_diff_rel_min = min(targsize_diff_rel),
      targsize_diff_rel_max = max(targsize_diff_rel),
      .by = c(scheme, domain, stratum)
    ) %>%
    arrange(scheme, stratum, domain) %>%
    print(n = Inf)
  
  get(
    "submodule_domain_scheme_ps_designattr",
    envir = eval(str2lang(scenario_name))
  ) %>%
    select(
      module,
      submodule,
      domain,
      scheme,
      type_count_2 = type_count,
      panel_set,
      ssize_all_2 = sp_sample_size_all_panels,
      ssize_type_2 = sp_sample_size_all_panels_type
    ) %>%
    inner_join(
      ref$submodule_domain_scheme_ps_designattr %>%
        select(
          module,
          submodule,
          domain,
          scheme,
          type_count = type_count,
          panel_set,
          ssize_all = sp_sample_size_all_panels,
          ssize_type = sp_sample_size_all_panels_type
        ),
      join_by(module, submodule, domain, scheme, panel_set),
      relationship = "one-to-one",
      unmatched = "error"
    ) %>%
    filter(ssize_type_2 != ssize_type)
}
```

Relative sample size decrease per stratum.

```{r}
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(ssize_diff_rel = (ssize_stratum_altered - ssize_stratum) / ssize_stratum) %>%
  ggplot(aes(x = ssize_diff_rel)) +
  geom_density()
```

Relative sample size decrease per stratum, distinguishing truncated vs not-truncated sample size.

```{r}
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(ssize_diff_rel = (ssize_stratum_altered - ssize_stratum) / ssize_stratum) %>%
  ggplot(aes(x = ssize_diff_rel, colour = spss_stratum_truncated)) +
  geom_density()
```

Plotting relative sample size decrease per stratum in the **groundwater** compartment, limited to sample size truncations plus sample size changes larger than 5%.

```{r}
#| fig.height: 12
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(
    ssize_diff_rel = (ssize_stratum_altered - ssize_stratum) / ssize_stratum,
    compartment = str_match(scheme, "^(\\w+)_")[, 2] %>% factor(),
    stratum_domain = str_c(stratum, "_", domain) %>% fct_rev()
  ) %>%
  select(-scheme) %>%
  filter(!is.na(compartment)) %>%
  filter(compartment == "GW") %>%
  filter(spss_stratum_truncated | abs(ssize_diff_rel) > 0.05) %>%
  distinct() %>%
  ggplot(aes(x = stratum_domain, y = ssize_diff_rel, fill = spss_stratum_truncated)) +
  geom_col() +
  facet_wrap(~compartment) +
  coord_flip()
```

Plotting relative sample size decrease per stratum in the **soil** compartment, limited to sample size truncations plus sample size changes larger than 5%.

```{r}
#| fig.height: 14
compare_ssizes_per_stratum(ssizes_new) %>%
  mutate(
    ssize_diff_rel = (ssize_stratum_altered - ssize_stratum) / ssize_stratum,
    compartment = str_match(scheme, "^(\\w+)_")[, 2] %>% factor(),
    stratum_domain = str_c(stratum, "_", domain) %>% fct_rev()
  ) %>%
  select(-scheme) %>%
  filter(!is.na(compartment)) %>%
  filter(compartment == "SOIL") %>%
  filter(spss_stratum_truncated | abs(ssize_diff_rel) > 0.05) %>%
  distinct() %>%
  ggplot(aes(x = stratum_domain, y = ssize_diff_rel, fill = spss_stratum_truncated)) +
  geom_col() +
  facet_wrap(~compartment) +
  coord_flip()
```

Plotting all absolute sample sizes in both scenarios and compartments, in domain Flanders.

```{r}
#| fig.height: 10
#| fig.width: 12
plot_abs_sample_sizes(ssizes_new, flanders = TRUE)
```

Plotting all absolute sample sizes in both scenarios and compartments, in the other domains.

```{r}
#| fig.height: 10
#| fig.width: 12
plot_abs_sample_sizes(ssizes_new, flanders = FALSE)
```

Up to here, domains were considered independently (samples are indeed drawn independently).
But domains can overlap.
Below, the integrated sample proportions (sample size / population size) per domain are calculated for each stratum, and filtered according to a proportion threshold that is exceeded (i.e. the compartment-specific proportion threshold used by the scenario).
This is only done for the new scenario.

Some 'overshooting' of the sample proportion limit is possible due to:

- 'independency' of densification in included domains
- (few cases:) increasing zero- or one-sized strata to two units at Flemish level

In groundwater:

```{r}
gw_thresh <- as.numeric(str_match(scenario_name, "gw([\\d\\.]+)")[,2])
get_integrated_ssizes_per_domain(sps_new, gw_thresh, "^GW") %>% 
  our_kbl()
```

In soil:

```{r}
soil_thresh <- as.numeric(str_match(scenario_name, "soil([\\d\\.]+)")[,2])
get_integrated_ssizes_per_domain(sps_new, soil_thresh, "^SOIL") %>% 
  our_kbl()
```


## Comparing FAG calendars

### LOCEVAL FAGs that were currently scheduled in 2025 for groundwater

LOCEVAL FAGs for cell types _currently_ scheduled in 2025, that serve **groundwater** schemes.

```{r}
loceval_2025_ref <-
  cal_ref %>%
  filter(
    str_detect(field_activity_group, "LOCEVAL"),
    year(date_start) < 2026
  ) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  unnest(scheme_moco_ps) %>% 
  filter(str_detect(scheme, "^GW")) %>% 
  distinct(stratum, grts_address, grts_address_final)
loceval_2025_ref
```

LOCEVAL FAGs for cell types currently scheduled in 2025 but nowhere in the new
FAG calendar (i.e. all compartments, including biotic).

```{r}
cal_new %>%
  filter(str_detect(field_activity_group, "LOCEVAL")) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  distinct(stratum, grts_address) %>%
  anti_join(loceval_2025_ref, ., join_by(stratum, grts_address)) %>% 
  print(n = Inf)
```

LOCEVAL FAGs for cell types currently scheduled in 2025 but nowhere in the new
FAG calendar, partim **groundwater**.

```{r}
loceval_2025_ref_missing_gw <-
  cal_new %>%
  filter(str_detect(field_activity_group, "^GW")) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  distinct(stratum, grts_address) %>%
  anti_join(loceval_2025_ref, ., join_by(stratum, grts_address))
loceval_2025_ref_missing_gw
```

Distribution of the above result according to strata:

```{r}
#| paged.print: FALSE
loceval_2025_ref_missing_gw %>%
  count(stratum) %>% 
  print(n = Inf)
```

LOCEVAL FAGs for cell types _scheduled in 2025 in the new FAG calendar for **groundwater** but absent from current 2025 schedule_.

```{r}
cal_new %>%
  filter(
    str_detect(field_activity_group, "LOCEVAL"),
    year(date_start) < 2026
  ) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  unnest(scheme_moco_ps) %>% 
  filter(str_detect(scheme, "^GW")) %>% 
  distinct(stratum, grts_address, grts_address_final) %>%
  anti_join(loceval_2025_ref, join_by(stratum, grts_address))
```

LOCEVAL FAGs for cell types currently scheduled in 2025 for groundwater AND present in the new
FAG calendar: counting the corresponding LOCEVALs per year in the new FAG
calendar.

```{r}
#| paged.print: FALSE
cal_new %>%
  filter(str_detect(field_activity_group, "LOCEVAL")) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  semi_join(loceval_2025_ref, join_by(stratum, grts_address)) %>%
  mutate(year_start = year(date_start)) %>%
  count(year_start)
```


### GW sampling FAGs

GW sampling FAGs for cell types _currently_ scheduled in 2025 or 2026.

```{r}
gwsamp_2025_ref <-
  cal_ref %>%
  filter(
    str_detect(field_activity_group, "SHALLSAMP"),
    year(date_start) < 2027
  ) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  summarize(
    first_year = year(min(date_start)),
    .by = c(stratum, grts_address, grts_address_final)
  )
gwsamp_2025_ref
```

Temporal redistribution of these FAGs from old to new calendar.

```{r}
cal_new %>%
  filter(str_detect(field_activity_group, "SHALLSAMP")) %>%
  semi_join(cell_types, join_by(stratum == type)) %>%
  summarize(
    new_first_date = min(date_start),
    .by = c(stratum, grts_address, grts_address_final)
  ) %>% 
  right_join(
    gwsamp_2025_ref,
    join_by(stratum, grts_address, grts_address_final),
    relationship = "one-to-one"
  ) %>% 
  count(first_year, new_first_date) %>% 
  our_kbl()
```


## Comparing MHQ overlap

Reference:

```{r}
#| paged.print: FALSE
show_mhq_overlap(sps_ref)
```

New:

```{r}
#| paged.print: FALSE
show_mhq_overlap(sps_new)
```



## Writing to GeoPackage layers

```{r}
# Run the outcommented ones only once (hence, interactively):
# write_to_gpkg_layer(sps_ref, "GW_ref")
# write_to_gpkg_layer(sps_ref, "GW_NONCELL_ref", types = non_cell_types)
# write_to_gpkg_layer(sps_ref, "SOIL_ref", "^SOIL")
write_to_gpkg_layer(sps_new, str_c("GW_", scenario_name))
write_to_gpkg_layer(
  sps_new,
  str_c("GW_NONCELL_", scenario_name),
  types = non_cell_types
)
write_to_gpkg_layer(sps_new, str_c("SOIL_", scenario_name), "^SOIL")
```




