---
title: "Water Level Models"
bibliography: references_csl.json
format:
  html:
    toc: true
    html-math-method: katex
knitr:
  opts_chunk:
    echo: true
---

This script loads previously stored data and performs multivariate modeling with `sdmTMB`.


```{r setup}
#| include: false
library("dplyr") # our favorite data wrangling toolbox
library("dbplyr") # dplyr + databases
library("arrow") # writing parquet files
library("ggplot2") # a visualization toolbox
library("gstat") # geospatial statistics
library("sdmTMB") # species distribution models Template Model Builder
library("mapview") # viewing maps
library("inbospatial") # elevation WCS


source("./watina_migration.R")
source("./auxiliary_data_queries.R")

# browser choice, a rather personal thing
if (!is.function(getOption("browser"))) options(browser = "librewolf")

# re-run models
rerun_models <- FALSE

```


Some parameters below are subject to selective variation.

```{r data-settings}
#| eval: true

# extent <- 1024 # m
# n_bins <- 128
extent <- 1024*2 # m
n_bins <- 256*2


mesh_cutoff <- as.integer(extent / n_bins)
```



# Data Preparation
## load data


The data was stored in a parquet file by the script `050_load_all_data.qmd`.

```{r load-data}
#| eval: true
cluster_dist <- 3400 # m # parameter for clustering
data_storage_file <- paste0("data/analysis_data_cl", cluster_dist, ".parquet") # where data is stored

data <- read_parquet(data_storage_file)
knitr::kable(head(data, 5))
```


## recipe

::: {.callout-note}
For notation, I heavily abbreviate, but try to keep the abbreviations consistent.

- Measurements `lw` and `hw` (`l` for "low", `h` for "high", `w` for "water") cover the `lg3` and `xg3` values queried from the watina database.
- Clusterwise de-trending is indicated by a suffix `d`: we get `lwd` and `hwd`
- `filterdepth` was stored as `z`
- further parameters: 
  - `elevation`, the elevation from DHMV and `slope`, the local slope
  - `soilclass` ("unknown" category will be disregarded), 
  - `wata_min_dist`,
  - TODO: add `ecoregio` (watina-gebied | ecoregio | n2khab-types)
:::

```{r parameter-rename}
data <- data %>%
  rename(
    filterdepth = z,
    elevation = elevation_dhmv,
    slope = slope_r2.5m,
    water_distance = wata_min_dist
  ) #%>%
  # filter(soilclass != "unknown")

```

## de-trending
The semivariance is anisotropically biased by any linear trends in the data, which is why we optionally "de-trend" the raw measurement.
"De-trending" is just a fancy word for working on the residuals of a linear regression against spatial co-ordinates, e.g. $z \sim x + y$.


In preparing the data, we have identified clusters by a distance of $`{r} cluster_dist` m$.
This distance is larger than the variogram range found below.
Hence, it is opportune to perform the de-trending somewhat more locally, i.e. in a clusterwise iteration.


```{r clusterwise-detrend}
#| eval: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data <- bind_rows(
  lapply(
    unique(data$cluster),
    FUN = detrend_clusterwise
  )
)

```


# Meshing Flanders

The spatial models require a mesh of the country.

Even more, the outcome crucially depends on mesh characteristics.


```{r sdmTMB-mesh-full}
#| eval: true
flanders_mesh <- make_mesh(data, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(flanders_mesh$mesh$n)
plot(flanders_mesh)
```


# Modeling

::: {.panel-tabset}
## LG3 {#sec-lg3}

```{r sdmTMB-model-lg3}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + elevation + soilclass + water_distance, #+ slope
  data = data,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```

```{r extract-lg3-parameters}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```


## HG3 {#sec-hg3}


```{r sdmTMB-model-hg3}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + elevation + soilclass + water_distance, # + slope
  data = data,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```

```{r extract-hg3-parameters}
#| eval: true
knitr::kable(tidy(fit_hg3))
```

```{r extract-hg3-spatial}
#| eval: true
knitr::kable(tidy(fit_hg3, "ran_pars"))
```

:::

## residuals

There are residuals.
It is not entirely clear what they mean.

```{r residuals-hg3}

res <- residuals(fit_hg3, type = 'mle-mvn')

length(res) == nrow(data)

res <- res[!is.na(res)]
res <- res[res != Inf]
res <- res[res != -Inf]
sqrt(sum(res^2))

ggplot(NULL, aes(x = res)) +
  geom_histogram(bins = 256, color = "black", fill = "lightgray") +
  theme_bw()

```

Great, those are as one would expect.
Note that there are `NA` and `Inf` in there (minority) to be removed.



# generalize: many models

- http://www.cookbook-r.com/Formulas/Creating_a_formula_from_a_string/
- https://stackoverflow.com/questions/46310123/r-dplyr-operate-on-a-column-known-only-by-its-string-name
- https://cran.r-project.org/web/packages/sdmTMB/sdmTMB.pdf ?sanity

```{r sdmTMB-model-general}
#| eval: true
compute_model <- function(model_formula) {
  message("______________________________________________________________________________\n")  
  message(model_formula)
    
  # model_formula <- "hwd ~ filterdepth + elevation + soilclass + water_distance"
  fit <- sdmTMB(
    as.formula(model_formula),
    data = data,
    mesh = flanders_mesh,
    family = gaussian(link = "identity"),
    spatial = "on"
  ) %>% run_extra_optimization()
  san <- sanity(fit, silent = TRUE)
  res <- tidy(fit)
  spa <- tidy(fit, "ran_pars")
    
  outcome <- bind_rows(res, spa)

  for (i in 1:length(san)) {
    check <- names(san)[[i]]
    outcome[nrow(outcome)+1, "term"] <- check
    outcome[nrow(outcome), "estimate"] <- as.integer(san[[i]])
  }

  res <- residuals(fit, type = 'mle-mvn')
  res <- res[!is.na(res)]
  res <- res[res != Inf]
  res <- res[res != -Inf]
  outcome[nrow(outcome)+1, "term"] <- "residuals"
  outcome[nrow(outcome), "estimate"] <- sqrt(sum(res^2))

  outcome <- tidyr::pivot_longer(
    outcome, 
    cols = c("estimate", "std.error", "conf.low", "conf.high")
    ) %>% filter(!is.na(value))

  outcome$model_formula <- model_formula

  return(outcome)

  message("______________________________________________________________________________\n")  
}


# result <- compute_model("hwd ~ filterdepth + elevation + soilclass + water_distance")
```



```{r model-list}
#| eval: false

if (rerun_models){
  model_formulas <- c(
    "hwd ~ filterdepth + elevation + soilclass + water_distance",
    "hwd ~               elevation + soilclass + water_distance",
    "hwd ~ filterdepth             + soilclass + water_distance",
    "hwd ~ filterdepth + elevation             + water_distance",
    "hwd ~ filterdepth + elevation + soilclass                 ",
    "lwd ~ filterdepth + elevation + soilclass + water_distance",
    "lwd ~               elevation + soilclass + water_distance",
    "lwd ~ filterdepth             + soilclass + water_distance",
    "lwd ~ filterdepth + elevation             + water_distance",
    "lwd ~ filterdepth + elevation + soilclass                 ",
    "hw  ~ filterdepth + elevation + soilclass + water_distance",
    "hw  ~               elevation + soilclass + water_distance",
    "hw  ~ filterdepth             + soilclass + water_distance",
    "hw  ~ filterdepth + elevation             + water_distance",
    "hw  ~ filterdepth + elevation + soilclass                 ",
    "lw  ~ filterdepth + elevation + soilclass + water_distance",
    "lw  ~               elevation + soilclass + water_distance",
    "lw  ~ filterdepth             + soilclass + water_distance",
    "lw  ~ filterdepth + elevation             + water_distance",
    "lw  ~ filterdepth + elevation + soilclass                 ",
    "hwd ~ 1",
    "hw  ~ 1",
    "lwd ~ 1",
    "lw  ~ 1"
    )

  results <- lapply(model_formulas, FUN = compute_model)
  results <- bind_rows(results)
  
  write_parquet(results, sink = "data/sdmTMB_results.parquet")

}
```


```{r modelcomparison-residuals}

results <- read_parquet("data/sdmTMB_results.parquet") %>%
  mutate_at(vars(term, name, model_formula), as.factor)

knitr::kable(results %>%
  filter(term == "residuals", name == "estimate") %>%
  select(model_formula, value) %>%
  rename(rms_residual = value ) %>%
  arrange(substring(model_formula, 1, 4), rms_residual))

```

```{r modelcomparison-effect-extract}

extract <- function (df, selection) {
  df <- df %>%
    filter(name == selection) %>%
    select(model_formula, value)
  colnames(df) <- c("model_formula", selection) 
  return(df)
}

join_extract <- function(df, raw, selection) {
  extra <- df %>% 
    left_join(
      extract(raw, selection),
      by = join_by(model_formula),
      relationship = "one-to-one"
    )
  return(extra)
}

extract_term <- function(select_term){
  term_data <- results %>%
    filter(term == select_term) %>%
    select(model_formula, name, value)  

  term_data <- term_data %>% 
    extract("estimate" ) %>%
    join_extract(term_data, "conf.low") %>%
    join_extract(term_data, "conf.high")

  term_overview <- term_data %>% mutate(
    sig. = (sign(conf.low) * sign(conf.high)) > 0
  )

  return(term_overview)
}

print(levels(results$term))

extract_term("(Intercept)")
extract_term("filterdepth")
extract_term("elevation")
extract_term("soilclasslight")
extract_term("soilclasspeat")
extract_term("water_distance")

```


::: { .callout-note }

- Residuals / RMSE do not favor any model; likely caused by substantial residual variation (as with the variograms). 
- When including `soilclass` in the model, the **`intercept`** receives a completely different meaning: it is then the average `xG3` on **heavy** soil.
- In **`soilclass == light`**, compared to `heavy`, `xG3`s are about $0.37 m$ higher (both `LG3` and `HG3`). In contrast, `xG3` on **`soilclass == peat`** are indifferent from `heavy`.
- **`filterdepth`** does influence `xG3`s in every possible model (confidence excludes zero). The effect is about $-0.013 m$ per meter filterdepth after de-trending, or $-0.014 m/m$ if not de-trending.
- An effect of **`elevation`** of approximately $1.0 m$ per meter elevation (1:1 coupling) is only detected when not detrending. Spatial detrending removes it entirely.
- There are no effects of **`water_distance`**: magnitudes are generally small and include zero, which might be a result of limitations in the data source accuracy.
- Interaction terms (e.g. `(filterdepth|soilclass)`) were not regarded, but see below for soilclass splitting.
- TODO Convergence checks were recorded, but not inspected, except for the "all in" model (all lights "green").

:::



# Split by Soilclass

Re-do the analysis, per soil class.


## Heavy
### data
```{r data-filter-heavy}
sc_selection <- "heavy"

data_sc <- data %>%
  filter(soilclass == sc_selection)

```

```{r clusterwise-detrend-heavy}
#| eval: true
#| code-fold: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data_sc %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data_sc <- bind_rows(
  lapply(
    unique(data_sc$cluster),
    FUN = detrend_clusterwise
  )
)

```


```{r sdmTMB-mesh-heavy}
#| eval: true
flanders_mesh <- make_mesh(data_sc, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(flanders_mesh$mesh$n)
# plot(flanders_mesh)
```


### Models 

```{r sdmTMB-model-lg3-heavy}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + water_distance, 
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```



```{r sdmTMB-model-hg3-heavy}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + water_distance, # + slope,
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```

## Results: Heavy Soil

::: {.panel-tabset}
### LG3
```{r extract-lg3-parameters-heavy}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial-heavy}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```

### HG3
```{r extract-hg3-parameters-heavy}
#| eval: true
knitr::kable(tidy(fit_hg3))
```

```{r extract-hg3-spatial-heavy}
#| eval: true
knitr::kable(tidy(fit_hg3, "ran_pars"))
```

:::

## Light
### data
```{r data-filter-light}
sc_selection <- "light"

data_sc <- data %>%
  filter(soilclass == sc_selection)

```

```{r clusterwise-detrend-light}
#| eval: true
#| code-fold: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data_sc %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data_sc <- bind_rows(
  lapply(
    unique(data_sc$cluster),
    FUN = detrend_clusterwise
  )
)

```


```{r sdmTMB-mesh-light}
#| eval: true
flanders_mesh <- make_mesh(data_sc, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(flanders_mesh$mesh$n)
# plot(flanders_mesh)
```



### Models 

```{r sdmTMB-model-lg3-light}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + water_distance, 
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```


```{r sdmTMB-model-hg3-light}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + water_distance, # + slope,
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```



## Results: Light Soil

::: {.panel-tabset}
### LG3
```{r extract-lg3-parameters-light}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial-light}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```


### HG3
```{r extract-hg3-parameters-light}
#| eval: true
knitr::kable(tidy(fit_hg3))
```

```{r extract-hg3-spatial-light}
#| eval: true
knitr::kable(tidy(fit_hg3, "ran_pars"))
```

:::


## Peat
### data
```{r data-filter-peat}
sc_selection <- "peat"

data_sc <- data %>%
  filter(soilclass == sc_selection)

```

```{r clusterwise-detrend-peat}
#| eval: true
#| code-fold: true

detrend_clusterwise <- function(cluster_idx) {
  sub_data <- data_sc %>%
    filter(cluster == cluster_idx)
  sub_data <- sub_data %>%
    mutate(
      hwd = lm(hw ~ x + y, sub_data)$residual,
      lwd = lm(lw ~ x + y, sub_data)$residual
    )
  return(sub_data)
}

data_sc <- bind_rows(
  lapply(
    unique(data_sc$cluster),
    FUN = detrend_clusterwise
  )
)

data_sc <- data_sc %>% 
  filter_at(
    vars(lwd, hwd, filterdepth, water_distance),
    all_vars(!is.na(.))
  )  

```


```{r sdmTMB-mesh-peat}
#| eval: true
flanders_mesh <- make_mesh(data_sc, xy_cols = c("x", "y"), cutoff = mesh_cutoff)
print(flanders_mesh$mesh$n)
# plot(flanders_mesh)
```


### Models 

```{r sdmTMB-model-lg3-peat}
#| eval: true
fit_lg3 <- sdmTMB(
  lwd ~ filterdepth + water_distance, 
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_lg3 <- run_extra_optimization(fit_lg3)
sanity(fit_lg3)
fit_lg3
```



```{r sdmTMB-model-hg3-peat}
#| eval: true
fit_hg3 <- sdmTMB(
  hwd ~ filterdepth + water_distance, # + slope,
  data = data_sc,
  mesh = flanders_mesh,
  family = gaussian(link = "identity"),
  spatial = "on"
)
fit_hg3 <- run_extra_optimization(fit_hg3)
sanity(fit_hg3)
fit_hg3
```

## Results: Peat

::: {.panel-tabset}
### LG3
```{r extract-lg3-parameters-peat}
#| eval: true
knitr::kable(tidy(fit_lg3))
```

```{r extract-lg3-spatial-peat}
#| eval: true
knitr::kable(tidy(fit_lg3, "ran_pars"))
```

### HG3
```{r extract-hg3-parameters-peat}
#| eval: true
knitr::kable(tidy(fit_hg3))
```

```{r extract-hg3-spatial-peat}
#| eval: true
knitr::kable(tidy(fit_hg3, "ran_pars"))
```

:::




## Summary: Soilclass

Note: elevation showed no effect on de-trended data, and was removed.

::: { .callout-note }

When performing the analysis only on **heavy soil**:

- Water distance has an effect $O(0.002 m/m)$.
- The effect of `filterdepth` disappears.
- Matérn range plummets ($0.11 m$). 

:::


::: { .callout-note }

In contrast, on **light soil**:

- Matérn range remains high ($> 200 m$). 
- Water distance has no effect.
- The effect of `filterdepth` is quantified to about $-0.011 m / m$.

:::


::: { .callout-note }

Finally **peat**:

- Modeling issues with the variogram component.
- Matérn range not determinable. 
- tendency: like "light" (filter depth yes; water distance no)


:::

# quick Matérn inspection

finding the threshold of $1 cm$ on a non-arbitrary Matérn.

```{r matern-estimate}

matern_function <- function(d, parameters) {
  scale <- parameters[1] # related to semivariance
  sigma <- parameters[2] # related to actual range; turning point
  nugget <- parameters[3]
  nu <- parameters[4]

  sill <- scale + nugget
  z <- sqrt(2 * nu) * d / sigma
  K <- suppressWarnings(besselK(z, nu))
  matern <- (z)^nu * K / (2^(nu - 1) * gamma(nu))
  result <- sill - scale * matern
  
  result[is.na(result)] <- 0
  return(result)
}

parameters <- c(1., 230., 0., 1.)
x <- seq(0., 1*230., length.out=1*2300+1)
y <- matern_function(x, parameters)
threshold <- x[min(which(y > 0.01))]
print(threshold)

plot(x, y)
abline(h = 0.01)
abline(v = threshold)
```


# Overall Summary / Recommendation

Distance:

- Matérn range of $230m$ implies a threshold of $13 m$ to stay below $1 cm$ difference.
- Rule: **recommendation: $d \le 10m$**


Soil class:

- Should be the same (rough classes). @FV how about sub-classes?
- probably covered by distance.
- Rule: **same soil class**.


Filter depth:

- difficult to implement rules: should *all* be equal relative to reduction line
- relevant effect ($-1 cm/m$) only on "light" soil -> influence of surface seepage?
- Rule: **target depth $\pm 1 m$**


Water distance:

- Water distance is unclear $\Rightarrow$ limit $5 m$ range (from slope on heavy soil).
- May be looser on light soil.
- Rule: **water distance $\pm 5m$**


Elevation:

- A threshold of $\pm 1cm$ would be below measurement accuracy.
- There is indication that water level "tov. maaiveld" is relatively constant (see detrending and elevation effect).
- Rule: **Stick with $\Delta h \le 0.3m$** (expert suggestion Toon)


# Questions Floris (20250108)
## across

> `mutate_at()`, `vars()`: 'superseded' volgens hun documentatie, zie hoger

I find `across()` very confusing and the syntax non-intuitive; in particular [fliter(if_*())](https://dplyr.tidyverse.org/articles/colwise.html#filter) reads horribly does not cover what I need.

-> I will attempt to adjust in future applications.

## fairly large mesh

> *Warning: The x or y column values are fairly large*: genegeerd, maar kan reëel effect betreffen?

As I understand it ("the spatial range is dependent on the scale of coordinates") this is only a problem if values "from zero to fairly large" are included.

- I could center, yet there should be no difference (range stays the same).
- Scaling to kilometers is no option: effects in the order of millimeters are already problem for regression.

-> should be fine.


## cluster-based random intercept

> spatial detrending ofwel alternatief: cluster-based random intercept ?

Random intercepts "steal" the variability of a model, i.e. they remove effects and sample size from other estimators. 
Should be used with care.
Detrending here has a different function; and I use it optionally to compare (e.g. elevation).

## correlation soilclass/elevation 

> "In soilclass == light, compared to heavy, xG3s are about 0.37m higher": 

> vermelden dat dit specifiek voor de detrended data is. 

Correct, thanks!


> Komt mogelijks omdat de soilclasses gecorreleerd kunnen zijn met elevation (die niet meer verklarend is in deze modellen)

Yes, if I remember correctly, a majority of light soils are close to the coast.


## explanation filterdepth effect

> "filterdepth does influence xG3s in every possible model": kan dit verklaard worden?

Yes!
Water level depends on filter depth: the deeper we drill, the lower the water level ($\approx -1 \frac{cm}{m}$).
Note that further down, this can be associated with soilclass "light".

I would speculate that this has to do with the balance of (i) capilarity and chemical water/soil interactions and (ii) hydrostatic pressure.
Vague working hypothesis: in light soil, where i presume water is more mobile, interactions with the soil matrix partially counteract hydrostatics; in the tube (absence of soil), full water weight presses down and can even push water outside (the image of a "net capillary pull" seems to attract my thoughts).

The deeper, the more hydrostatics inside; hydrostatics outside increases slower because capillarity adds up.
In heavy soil, water is unable to penetrate filter surrounding.


Again: just guessing.

## collinearity of water distance
> "no effects of water_distance": misschien is er sprake van collineariteit?

Indeed!

I would not expect filterdepth and soilclass to be waterdistance-related.

However, waterdistance could be related to elevation in some clusters (therefore killed by de-trending). 

But then a rule on elevation should cover it as well.

-> do we have to care about co-linearity at all, using an `or`-combination of exclusion rules?


## Split by Soilclass:

> heavy: inderdaad binnen verwachting dat er minder spatiale relatie is

My earlier statement on this was wrong!

It appeared from the updated variograms that the spatial relation is not all that different in heavy soil (range of >200m is still plausible, if the data would show any pattern). 
The difference is the magnitude of variance: there seems to be much less variance overall.

To be discussed.

## Modeling best practices
> algemeen ivm modellering: normaal doen we in functie van modellering ook:

1. data-exploratie
  - outliers covariaten & respons?
  - collineariteit covariaten?
  - datahoeveelheid vlgs categorieën & interacties?
  - zie verder bv/Zuur A.F., Ieno E.N. & Elphick C.S. (2010). A protocol for data exploration to avoid common statistical problems: Data exploration. Methods in Ecology and Evolution 1 (1): 3–14. https://doi.org/10.1111/j.2041-210X.2009.00001.x./

2. modeldiagnose (checking violations against model assumptions ~ reliability of estimates & p-values)
3. modelevaluatie (predictive accuracy measures, k-fold cross validation ...)

*Ad 1.*
Yes, there are outliers. Yes, there is collinearity. Yes, some categories might be small (there is a lot of data; this is less concern). 
And generally I absolutely agree, these things need to be in order.
Data exploration happens, to a certain extent, in script `060_Variogram`.

Assume that all these data problems apply.
What would be the consequence for our present analysis?

In the present case, I would personally tend to just accept the data as we have it.


Some options to give us more confidence:

- return to daily water levels (larger -> more robust data set)
- return to water chemistry (independent, but smaller data set; might not work)


*Ad 2. & 3.*
I appreciate leads and guidance in how to implement that, given that you think the outcome would be consequential (i.e. given that we have a chance to react, even if "reaction" is just knowing that a givenmodel is useless).
Note that I fear that I would be limited to the capabilities of the `sdmTMB` toolbox.
