---
title: "Explore sdmTMB"
author: "Falk Mielke"
date: "`r lubridate::today()`"
number-sections: true
bibliography: references.bib
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

# Purpose

Testing the `sdmTMB` package on synthetic data.


```{r setup}
#| eval: true
silence <- suppressPackageStartupMessages
# library("sp") |> silence() # SPatial stuff
library("dplyr") |> silence() # for group-apply function
library("ggplot2") |> silence() # plotting
library("sdmTMB") |> silence() # modeling

```



# Data

## general settings

```{r data-settings}
n_observations <- 2^12 # 4096
extent <- 100 
smooth_sigma <- 2*pi # 6.283
zrange <- pi # 3.1416
intercept <- pi/2 # 1.57

# covariate effect magnitudes
a_slope <- pi/6 # 0.5236
b_slope <- pi/5 # 0.6283

```

## raw random

```{r data-raw}
gimme_points <- function(N = 100, scale = 1.0) (runif(N)*2.0-1.0) * scale
data <- data.frame(
  x = gimme_points(n_observations, extent),
  y = gimme_points(n_observations, extent),
  z_ = gimme_points(n_observations, zrange)
)
data$z_ <- data$z_ + intercept
```

## smoothing

```{r data-cross-distance}
cross_diff <- function (vec) outer(X = vec, Y = vec, FUN = function(X, Y)  Y-X)
Euclid <- function (data) sqrt(cross_diff(data$x)^2 + cross_diff(data$y)^2 )


smooth_pre <- function (data, sigma = NULL) {

  if (sigma <= 0 || is.null(sigma) || is.na(sigma)) {
    return(data$z_)
  }

  dist <- Euclid(data)
  weight <- dnorm(dist, 0, sigma)
  weight <- weight / colSums(weight)

  zmoothed <- weight %*% data$z_
  return (zmoothed[,1])
}

data$s_ <- smooth_pre(data, smooth_sigma)

```

## covariates

```{r data-covariate-definition}

data$a <- 0.7 * data$x / extent + 0.3 * data$y / extent
data$b <- as.integer(rbinom(n = n_observations, size = 1, prob = 0.4))

data$s <- data$s_ + a_slope * data$a + b_slope * data$b

```



## "De-Trending"

```{r data-detrending}
data$d <- lm(s ~ x + y, data)$residual

```



## glimpse

```{r data-glimpse}
head(data, 5)
```

```{r data-visualization}
rbpal <- colorRampPalette(c('black','gold'))
color <- rbpal(n_observations)[as.numeric(cut(data$s, breaks = n_observations))]

plot(data$x, data$y, col = color, pch = as.integer(18 + 2*data$b))

```

# Distance-Difference-Calculation
To get an overview of whether I got the data right.

```{r difference-distance-plot}

zdiff <- function (data) cross_diff(data$s)
ddiff <- function (data) cross_diff(data$d)
unwrap <- function (mat) mat[lower.tri(mat)]

# compute raw distance and difference
dist_diff <- data.frame(
  distance = unwrap(Euclid(data)),
  difference = unwrap(zdiff(data)),
  det_diff = unwrap(ddiff(data))
)

# limit range
dist_diff <- dist_diff %>% filter(distance < extent)

# binning
dist_diff$bin <- as.factor(cut(dist_diff$distance, breaks = 65))
dist_diff_binned <- dist_diff %>% 
  group_by(bin) %>%
  summarize_all(
    c("mean" = mean,
      "var" = var
    )
  ) 

# choose one parameter
dist_diff_binned$semivariance <- 0.5 * dist_diff_binned$det_diff_var


# plot(distance, difference, pch = 21, color = 'black', alpha = 0.1)
head(dist_diff_binned)


# plot
plot(dist_diff_binned$distance_mean, dist_diff_binned$semivariance,
   type = "o", col = "darkgreen",
   ylab = "semivariance"
   )

```


# Water Level Analysis (WLA™)

```{r mesh}
mesh <- make_mesh(data, xy_cols = c("x", "y"), cutoff = 8)
print(mesh$mesh$n)
plot(mesh)
```
Fascinating!

```{r model}
fit <- sdmTMB(
  s ~ 1 + b,
  # d ~ 1 + b,
  spatial_varying = ~a,
  data = data,
  mesh = mesh,
  family = gaussian(link = "identity"), 
  spatial = "on" 
)

run_extra_optimization(fit) # to give our model the little extra fit!

# https://pbs-assess.github.io/sdmTMB/articles/model-description.html#observation-model-families
# https://pbs-assess.github.io/sdmTMB/articles/model-description.html#spatial-random-fields
sanity(fit)

print(paste0(
  "intercept: ", round(intercept, 4),
  "a: ", round(a_slope, 4),
  ", b: ", round(b_slope, 4)
))
fit
```

Fun fact: they spell "Gamma" with capital "G", but "gaussian" with small letter.


# Observations

Great general tool, does exactly what we intend!

- execution order for data generation is crucial: smooth first, then covariates 
- `b` (random categorical, unsmoothed) is well recovered
- ~~"a" (xy-correlated slope) is not easily recovered~~ use `spatial_varying` where possible -> `a` is also recovered 
- Matérn range does ~~not~~ reflect "sigma" (as a proxy), if spatial varying parameters are well incorporated




# Reference: *Gadus macrocephalus*
([reference](https://pbs-assess.github.io/sdmTMB/index.html#basic-use))

*Below I just copy-pasted the `sdmTMB` vignette to understand general working.*


> Our model contains a main effect of depth as a penalized smoother, a spatial random field, and Tweedie observation error. 

```{r pacific-cod}
head(pcod)
```
  

> We start by creating a mesh object that contains matrices to apply the SPDE approach.
> Here, cutoff defines the minimum allowed distance between points in the units of X and Y (km). Alternatively, we could have created any mesh via the fmesher or INLA packages and supplied it to make_mesh(). We can inspect our mesh object with the associated plotting method plot(mesh)
```{r cod-mesh}
codmesh <- make_mesh(pcod, xy_cols = c("X", "Y"), cutoff = 10)
plot(codmesh)
```

> Fit a spatial model with a smoother for depth:

```{r cod-model}
codfit <- sdmTMB(
  density ~ s(depth),
  data = pcod,
  mesh = codmesh,
  family = tweedie(link = "log"),
  spatial = "on"
)
sanity(codfit)
codfit
```

- There are visualization and data extraction options.
- Different families have different kids.


([further reading](https://pbs-assess.github.io/sdmTMB/articles/basic-intro.html))

