---
title: "Sensibility of a shallow top of a piezometer filter"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("watina")
library("n2khab")
library("sf")
library("stringr")
library("ggplot2")
library("ggside")
library("ggridges")
library("mapview")


watina_conn <- connect_watina()
save_figures <- TRUE # whether/not to save plots in utils::mkdir("./figures")
```

---
# installation of the cmdstanr package is optional:
# 
# - `remotes::install_github("stan-dev/cmdstanr")`
# - `cmdstanr::check_cmdstan_toolchain(fix = TRUE)`
# - `cmdstanr::install_cmdstan()`
---


# Purpose
In this workbook, I will attempt to redo and extend some of the previous analyses (`110_shallow_filtertop_sensibility/explore_filtertops.Rmd`).
My goal is to stick with the `watina` capabilities this time.

Tasks:

- Get `obswells` data and `xg3` values for 
  - a scatter of filter (top) depth *versus* HG3 
  - bonus: proportion of HG3 above ground per filter depth 
  - bonus: per soil class

- measurement clusters with `watina::cluster_locs()`: compare proximal wells of different depth

# Load Data
## concept
We have to distinguish several levels of spatiotemporal organization of observation well.

One level is **location**, or site. 
Imagine, per analogy a water well in grandma's garden. 
Someone drilled a hole in the ground 100 years ago for groundwater retrieval.
This well is one location.

Yet over time, the actual well installation (**observation well**) was altered. 
In the beginning, there were stone walls and a bucket.
At some point, someone installed a mechanical pump to facilitate water retrieval - yet they used exactly the same spot (/location) in granny's garden.
The latest observation well at that location is an electric pump which is programmed to water granny's potatoes at regular intervals.


Transporting this analogy to actual groundwater measurement with `watina`:

- We can query either on *location* level, which might be agnostic to temporary technical exchange of the well hardware; this usually links to installation parameters for the latest installment (requires *aggregation*).
- Alternatively, we can query the *observation well* level. Consecutive installations will produce multiple lines in the output, which all share the location.


In the `watina::get_locs()` function, the `obswells` argument provides choice of level, and `obswell_aggr` provides control of the aggregation strategy.


## well locations
First, I will load the well data. 
By setting `obswells = FALSE`, they will be on location level (which is different from the strategy in the previous overview).

```{r}
obswells_db <- get_locs(
  watina_conn,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  obswells = FALSE
) %>% mutate(filtertop_depth = filterdepth - filterlength / 2)
# (remember: filterdepth is positive downwards;
#    thus filtertop_depth must be smaller)

obswells_db %>% count()
```


Always worth comparing with previous excerpts:

```{r}
obswells_reference <- readRDS("geospatial_example_subset.rds") %>%
  mutate(filtertop_depth = filterdepth - filterlength / 2)

obswells_reference %>%
  group_by(loc_code) %>%
  count() %>%
  arrange(desc(n))
location_selection <- "ZWIP017"
```

(I have tried `ZWIP034` and `ZWIP006`, yet their data availability is less instructive than `ZWIP017`.)


```{r}
glimpse(obswells_reference %>% filter(loc_code == location_selection))

obswells_location <- obswells_db %>%
  collect() %>%
  filter(grepl("zwin", area_name, ignore.case = TRUE))
glimpse(obswells_location %>% filter(loc_code == location_selection))
```

Note that the well on that location was replaced in May 2010 (and again recently).
This implies a discontinuity: filter top depth changed by about 2m towards the surface.


This is brief confirming evidence that we work "per location" now, aggregation picks the latest values.
(Will check back below whether `HG3` also restricts to the latest installment.)


## xg3 values
Then, we require the xg3 data for all locations:

```{r}
xg3s_db <- obswells_db %>%
  get_xg3(
    watina_conn,
    startyear = 1960,
    truncated = FALSE,
    with_estimated = FALSE,
    vert_crs = "local",
    collect = FALSE
  )
glimpse(xg3s_db %>% arrange(desc(hydroyear), loc_code))
xg3s_db %>% count()
```

Obviously, XG3 values are collected per location, per year, so we expect *location ⨉ year* table entries with sparse yearly coverage.

How is it with our sample location?

```{r}
xg3_samplelocation <- xg3s_db %>%
  filter(loc_code == location_selection) %>%
  arrange(hydroyear) %>%
  collect()
xg3_samplelocation %>%
  ggplot(aes(x = hydroyear, y = vg3_lcl)) +
  geom_line() +
  geom_line(aes(y = hg3_lcl), color = "darkred", linetype = "dashed") +
  geom_line(aes(y = lg3_lcl), color = "blue", linetype = "dashed") +
  geom_vline(xintercept = c(2010.5)) +
  geom_hline(yintercept = c(0)) +
  ggtitle(paste0("measured water levels of ", location_selection, " location"))
```

There seems to have been a change in 2010 which affects VG3, yet that might be co-incidence.

## soilmap
We might categorize by soil class, and the relevant geospatial data are loaded here.

```{r}
soilmap_raw <-
  read_soilmap()

soilmap <- soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  )

glimpse(soilmap)
```


## actual water levels
Now, wouldn't it be fun if we could query actual water levels for this location in 2010?

Sit back, and behold.

```{r}
water_query <- tbl(watina_conn, "FactPeilMeting") %>%
  filter(.data$MeetpuntWID %in% 6017) %>%
  select(PeilmetingCategorie, TijdWID, Niveau) %>%
  mutate(hydroyear = floor((TijdWID - 400) / 10000)) %>%
  arrange(TijdWID)
tijd <- tbl(watina_conn, "DimTijd") %>%
  rename(TijdWID = DatumWID) %>%
  select(TijdWID, HydroJaar)
water_query <- water_query %>%
  left_join(tijd, join_by(TijdWID)) %>%
  collect() %>%
  mutate(date = lubridate::ymd(TijdWID))
glimpse(water_query)
```

```{r}
legend_colors <- c(
  "niveau" = "black",
  "lg3" = "darkblue",
  "hg3" = "darkred",
  "vg3" = "darkgray"
)

water_query %>%
  left_join(xg3_samplelocation, join_by(hydroyear)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = Niveau, color = "niveau")) +
  geom_line(aes(y = vg3_lcl, color = "vg3"), linetype = "dashed") +
  geom_line(aes(y = hg3_lcl, color = "hg3"), linetype = "dashed") +
  geom_line(aes(y = lg3_lcl, color = "lg3"), linetype = "dashed") +
  geom_vline(xintercept = as.Date("2010-05-21")) +
  geom_hline(yintercept = c(0)) +
  labs(colour = "quantity") +
  xlim(as.Date("2004-01-01"), as.Date("2022-12-31")) +
  scale_color_manual(values = legend_colors) +
  theme(text = element_text(size = 20))
# ggtitle(
#   paste0(
#     "measured water levels of ",
#     location_selection,
#     " location (depth change in 2010)"
#   )
# )


if (save_figures) {
  ggsave("figures/hg3_zwip017.svg",
    dpi = 300,
    width = 16, height = 8, units = "in"
  )
}
```

Interesting. To be followed up.

# Population Plots

## hg3 ~ filtertop

I am interested in the distribution of HG3 values as a function of filter top depth.
This requires a many-to-one join of `xg3` onto `obswells`.

(At this point, data needs to be collected for spatial join.)


TODO: there are hidden duplicates in the database.
```{r, eval=FALSE}
rs <- DBI::dbGetQuery(watina_conn,
  "
   SELECT mp1.*, cnt
   FROM vwDimMeetpunt AS mp1
   LEFT JOIN (
   SELECT DISTINCT
      MeetpuntXCoordinaat, MeetpuntYCoordinaat, COUNT(*) AS cnt
   FROM vwDimMeetpunt
   GROUP BY MeetpuntXCoordinaat, MeetpuntYCoordinaat
   HAVING COUNT(*) > 1
   ) AS mp2 ON mp1.MeetpuntXCoordinaat = mp2.MeetpuntXCoordinaat AND mp1.MeetpuntYCoordinaat = mp2.MeetpuntYCoordinaat
   WHERE cnt IS NOT NULL
   ORDER BY mp1.MeetpuntXCoordinaat ASC
     ;
  "
  )
rs  %>% head()
```

```{r}
# first, join spatial information
joined_spatial <- st_join(
  watina::as_points(
    obswells_db %>%
      collect()
  ), soilmap
) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown"))

# next, append water levels
xg3_obs <- joined_spatial %>%
  select(loc_code, filtertop_depth, filterdepth, soilclass) %>%
  left_join(xg3s_db %>% collect(), join_by(loc_code))
glimpse(xg3_obs)


soilclass_counts <- xg3_obs %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
soilclass_colors <- c("#284971", "#8eafd7", "#f0d70f", "#cccccc")
names(soilclass_colors) <- names(soilclass_counts)
```


The filtertop depth has a soilclass-dependent effect on HG3:

```{r}
p1 <- xg3_obs %>%
  filter(!is.na(hg3_lcl)) %>%
  # filter(hydroyear > 2018) %>%
  ggplot(aes(x = filtertop_depth, y = hg3_lcl, color = soilclass)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  theme_bw()

p1 +
  geom_vline(xintercept = c(0, 0.3)) +
  geom_hline(yintercept = c(0)) +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  scale_x_reverse() +
  coord_cartesian(xlim = c(4, 0), expand = FALSE) +
  # ggtitle("soil class determines how much HG3 drops with filtertop depth ") +
  xlab("filtertop depth (m)") + ylab("local HG3 (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.8, 0.15),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )

if (save_figures) {
  ggsave("figures/hg3_filtertopdepth.png",
    dpi = 300,
    width = 16, height = 8, units = "in"
  )
}
```


## stats
### conceptual
This cries for a Bayesian model of the form:

<center>
$$hg3 \sim 1 + soilclass + (filtertop_depth : soilclass) + \varepsilon\cdot a$$
</center>

With a Student's T posterior (robust regression)


```{r}
library("brms")
```


### "fixed" slope everything
... although everything *is* random.

*Thanks to input by Hans!*

```{r}
model_file <- "models/fixedslope.rds"
fixedslope_model <- brm(
  formula = hg3_lcl ~ 0 + soilclass + (filtertop_depth:soilclass),
  data = xg3_obs,
  family = student,
  warmup = 2^10,
  iter = 2^11,
  chains = 4,
  cores = 4,
  backend = "cmdstanr",
  file = model_file,
  file_refit = "on_change",
  seed = 123
)
```

```{r}
mcmc_plot(fixedslope_model) # forest plot
plot(fixedslope_model) # traceplot
summary(fixedslope_model)
```

### results summary | summary results

- sigma 0.22
- nu 2.23
- heavy:    0.02 - 0.05 ftd 
- light:    0.08 - 0.17 ftd 
- peat:     0.05 - 0.03 ftd 
- unknown:  0.07 - 0.18 ftd 


### technical notes

- `brms` is disappointingly slow (yet I did not give it specific priors)
- ... and limited (e.g. could not use double-hierarchical).
- but it *looks* and *feels* nice!

*(mental note: learn STAN)*


### python reference
Intermediately struggling with `brms`, I felt I needed a refresher.

```{r}
readr::write_delim(
  xg3_obs %>% select(-geometry),
  "xg3obs_data.csv",
  delim = ";"
)
```

Results match those in R.
```{r, out.width = "80%", eval=FALSE}
py_figure <- "figures/python_model_fit_hg3_lcl.png"
if (exists(py_figure)) {
  knitr::include_graphics(py_figure)
}
```

If you would like to try the python code below, start with:
```{r eval=FALSE}
reticulate::install_python(version = "3.12.6") # note: "latest" will not work
reticulate::py_install(c(
  "scipy",
  "numpy",
  "pandas",
  "seaborn",
  "pymc",
  "matplotlib"
))
```

Code reference:
```{python eval=FALSE}
# import numpy as NP
import scipy.stats as STATS
import numpy as NP
import pandas as PD
import matplotlib as MP
import matplotlib.pyplot as PLT
import seaborn as SB
import pymc as PM
import pytensor as PP
import pytensor.tensor as PT

data = PD.read_csv("xg3obs_data.csv", sep = ";")
y_var = 'hg3_lcl'

# data cleansing
data = data.loc[:, ['loc_code', 'hydroyear', 'filtertop_depth', 'soilclass', y_var]]
data.dropna(inplace = True)
print(data.loc[data[y_var].values > 10, :])
data = data.loc[NP.logical_and(data[y_var].values < 2, data[y_var].values > -5), :]

# soilclass category
data['soilclass'] = PD.Categorical(data['soilclass'])
categories = data['soilclass'].cat.codes.values
n_categories = len(NP.unique(data['soilclass'].values))
cat_order = data['soilclass'].cat.categories
print(cat_order)

# data dimensions
n_observations = data.shape[0]
population_mean = data[y_var].mean()
population_sd = data[y_var].std()

# values
x = data['filtertop_depth'].values.ravel()#.reshape((-1,1))
y = data[y_var].values.ravel()#.reshape((-1, 1))


# model construction
with PM.Model() as model:
    a = PM.Normal('intercept', mu = population_mean, sigma = 2*population_sd, shape = n_categories)

    b = PM.Normal(f'class_slope', mu = 0, sigma = population_sd, shape = n_categories)
    d = PM.Data(f'xdata', x)
    estimator = a[categories] + d * b[categories]


    residual = PM.HalfNormal('residual')
    nu = PM.Normal('dof')

    posterior = PM.StudentT('y', mu = estimator, sigma = residual, nu = nu, observed = y)


# sampling
with model:
   trace = PM.sample(draws = 2**10 \
                     , tune = 2**10 \
                     , chains = 4 \
                     , progressbar = True \
                   )
   result = PM.summary(trace)
   print(PD.DataFrame(result))
   PM.plot_trace(trace, combined = True)
   PLT.show()

# data- and regression plot
dpi = 300
cm = 1/2.54
PLT.rcParams.update({"text.usetex": True})
fig = PLT.figure(figsize = (16*cm, 8*cm), dpi = dpi)
fig.subplots_adjust(top = 0.98 , right = 0.98, bottom = 0.16, left = 0.10 )

ax = fig.add_subplot(1, 1, 1)
x = NP.array([  NP.min(data['filtertop_depth'].values) \
              , NP.max(data['filtertop_depth'].values) \
             ])
colors = {'heavy': "#284971", 'light': "#8eafd7", 'peat': "#f0d70f"}

for nr, cat in enumerate(cat_order):
    # print(nr, cat)
    subdat = data.loc[data['soilclass'] == cat, :]
    ax.scatter(subdat['filtertop_depth'].values, subdat[y_var].values \
               , s = 1, alpha = 0.2, color = colors.get(cat, "#cccccc") \
               , zorder = 10)

    m = result.loc[f"class_slope[{nr}]", "mean"]
    n = result.loc[f"intercept[{nr}]", "mean"]
    y = m*x+n
    ax.plot(x, y, color = colors.get(cat, "#cccccc"), label = f"{cat}: \\(y={n:.3f} {m:+.3f}x\\) " \
            , zorder = 20)

ax.set_xlim(ax.get_xlim()[::-1])
ax.spines[['top', 'right']].set_visible(False)
ax.set_xlabel("filtertop depth (m)")
ax.set_ylabel("local HG3 (m)")
PLT.legend(loc = 'best', fontsize = 8)
fig.savefig(f"figures/python_model_fit_{y_var}.png", dpi = 300, transparent = False)
PLT.tight_layout()
PLT.show()
```


# Location, Location, Location
## concept
The next idea is to compare location clusters.

We search for locations where there are **clusters** ( := groups of wells, some shallow top, some deep), in order to 

- (ia) visualize/compare the local HG3 value difference 
- (ib) quantify the local HG3 value correlation
- (ii) comparatively explore the actual water level measurement time series


An extended analysis might go to temporal differences: find exchanged wells which change more or less drastically in filter (top) depth, again comparing

- (i) HG3
- (ii) level time series

## find clusters
... exploring `watina::cluster_locs()`
```{r}
obswell_clusters <- cluster_locs(
  joined_spatial,
  max_dist = 2,
  output_var = "cluster_id",
  # xvar = x, yvar = y
)
count_clusters <- obswell_clusters %>%
  group_by(cluster_id) %>%
  count(name = "n_clustermembers") %>%
  filter(n_clustermembers > 1) %>%
  st_drop_geometry() 

obswell_clusters <- obswell_clusters %>%
  left_join(count_clusters, join_by(cluster_id)) %>%
  filter(!is.na(n_clustermembers)) %>%
  mutate_at("cluster_id", as.factor)

obswell_clusters %>% count()
```

Quick notes: 

- (with max_dist=1) clusters never larger than 2, about 600 in total
- marginally more with max_dist = 2 (n=650), but several with >2 members


Here they are:
```{r}
obswell_clusters %>%
  mapview(
    map.types = c("OpenTopoMap"),
    zcol = "soilclass"
  )
```

## cluster depth differences

Just for fun, quantify the difference of filterdepth within clusters:
```{r}
cluster_spans <- obswell_clusters %>%
  group_by(cluster_id) %>%
  summarize(
    n = n(),
    min_depth = min(filterdepth),
    max_depth = max(filterdepth),
    span_filtertop_depth = abs(max(filterdepth) - min(filterdepth)),
  ) %>%
  filter(
    max_depth != min_depth,
    span_filtertop_depth > 0.1
  ) %>%
  arrange(desc(span_filtertop_depth))

cluster_spans %>%
  ggplot(aes(x = span_filtertop_depth)) +
  geom_histogram()
```


Then, identify the ones within the clusters which have the largest filterdepth:
```{r}
deep_members <- obswell_clusters %>%
  st_drop_geometry() %>%
  left_join(cluster_spans, join_by(cluster_id)) %>%
  filter(!is.na(max_depth), filterdepth == max_depth) %>%
  select(cluster_id, loc_wid) %>%
  group_by(cluster_id) %>%
  summarize(first_deep = min(loc_wid)) %>%
  select(cluster_id, first_deep) %>%
  unique() %>%
  rename(deep_ref = first_deep)

obswell_clusters <-
  obswell_clusters %>%
  left_join(deep_members, join_by(cluster_id),
            relationship = "many-to-one")
```


## bring in water levels

Now, xg3's join in.
```{r}
xg3_clustered <- obswell_clusters %>%
  left_join(xg3s_db %>% collect(), join_by(loc_code))
glimpse(xg3_clustered)
``` 


Now we can split the dataset into deepest and others:
```{r}
deepest <- xg3_clustered %>%
  filter(loc_wid == deep_ref)

undeep <- xg3_clustered %>%
  filter(loc_wid != deep_ref)
```

YOLO!


Quick check: are there sites with multiple undeep per `loc_wid`?
yes!
```{r}
undeep %>%
  st_drop_geometry() %>%
  count(loc_wid) %>%
  arrange(desc(n))
```

## within-cluster variance

Get the relevant parameters and join again.

NOTE: focusing on HG3, for now.

```{r}
dp <- deepest %>%
  st_drop_geometry() %>%
  filter(!is.na(hg3_lcl)) %>%
  select(loc_wid, loc_code, hydroyear, filterdepth, filtertop_depth, hg3_lcl)
names(dp) <- paste0("deep.", names(dp))

ud <- undeep %>%
  filter(!is.na(hg3_lcl)) %>%
  select(
    loc_wid, loc_code, soilclass,
    hydroyear, filterdepth, filtertop_depth,
    hg3_lcl, deep_ref
  )

hg3_diffs <- ud %>%
  left_join(dp,
    by = join_by(deep_ref == deep.loc_wid, hydroyear == deep.hydroyear),
    relationship = "many-to-many"
  ) %>%
  filter(!is.na(deep.hg3_lcl)) %>%
  mutate(
    diff_hg3 = deep.hg3_lcl - hg3_lcl,
    diff_filterdepth = deep.filterdepth - filterdepth,
    diff_filtertop_depth = deep.filtertop_depth - filtertop_depth
  ) %>%
  group_by(loc_code, soilclass, diff_filterdepth, diff_filtertop_depth) %>%
  summarize(diff_hg3 = mean(diff_hg3))
```


```{r}
soilclass_counts <- hg3_diffs %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
soilclass_colors <- c("#284971", "#8eafd7", "#f0d70f", "#cccccc")
names(soilclass_colors) <- names(soilclass_counts)

p <- hg3_diffs %>%
  ggplot(aes(x = diff_filtertop_depth, y = diff_hg3)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  theme_bw() +
  geom_hline(yintercept = c(0)) +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  ggtitle(
    "obswell clusters: filtertop depth has limited influence on local HG3"
  ) +
  xlab("difference in filtertop depth (m)") +
  ylab("site-averaged difference in local HG3 (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.9, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )


p2 <- ggExtra::ggMarginal(p, type = "histogram", binwidth = 0.05)
p2

if (save_figures) {
  ggsave("figures/diff_hg3_soilclass.pdf", p2,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```


At close inspection, there seems to be a trend in heavy soils. 
```{r}
p <- hg3_diffs %>%
  filter(soilclass != "unknown") %>%
  ggplot(aes(x = diff_filtertop_depth, y = diff_hg3)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  geom_hline(yintercept = c(0)) +
  facet_wrap(~soilclass, ncol = 2) +
  theme_bw() +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  ggtitle(
    "obswell clusters: on heavy soil, filtertop depth influences HG3"
  ) +
  xlab("difference in filtertop depth (m)") +
  ylab("site-averaged difference in local HG3 (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.9, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )

if (save_figures) {
  ggsave("figures/diff_hg3_soilclass_specific.svg", p,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```



## within-cluster variance, VG3!

Repeat for VG3, where we have more data.

```{r}
dp <- deepest %>%
  st_drop_geometry() %>%
  filter(!is.na(vg3_lcl)) %>%
  select(loc_wid, loc_code, hydroyear, filterdepth, filtertop_depth, vg3_lcl)
names(dp) <- paste0("deep.", names(dp))

ud <- undeep %>%
  filter(!is.na(vg3_lcl)) %>%
  select(
    loc_wid, loc_code, soilclass,
    hydroyear, filterdepth, filtertop_depth,
    vg3_lcl, deep_ref
  )

vg3_diffs <- ud %>%
  left_join(dp,
    by = join_by(deep_ref == deep.loc_wid, hydroyear == deep.hydroyear),
    relationship = "many-to-many"
  ) %>%
  filter(!is.na(deep.vg3_lcl)) %>%
  mutate(
    diff_vg3 = deep.vg3_lcl - vg3_lcl,
    diff_filterdepth = deep.filterdepth - filterdepth,
    diff_filtertop_depth = deep.filtertop_depth - filtertop_depth
  ) %>%
  group_by(loc_code, soilclass, diff_filterdepth, diff_filtertop_depth) %>%
  summarize(diff_vg3 = mean(diff_vg3))
```


```{r}
soilclass_counts <- vg3_diffs %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
soilclass_colors <- c("#284971", "#8eafd7", "#f0d70f", "#cccccc")
names(soilclass_colors) <- names(soilclass_counts)

p <- vg3_diffs %>%
  ggplot(aes(x = diff_filtertop_depth, y = diff_vg3)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  theme_bw() +
  geom_hline(yintercept = c(0)) +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  ggtitle(
    "obswell clusters: filtertop depth has limited influence on local VG3"
  ) +
  xlab("difference in filtertop depth (m)") +
  ylab("site-averaged difference in local VG3 (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.9, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )


p2 <- ggExtra::ggMarginal(p, type = "histogram", binwidth = 0.05)
p2

if (save_figures) {
  ggsave("figures/diff_vg3_soilclass.pdf", p2,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```


```{r}
p <- vg3_diffs %>%
  filter(soilclass != "unknown") %>%
  ggplot(aes(x = diff_filtertop_depth, y = diff_vg3)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  geom_hline(yintercept = c(0)) +
  facet_wrap(~soilclass, ncol = 2) +
  theme_bw() +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  ggtitle(
    "obswell clusters: on heavy soil, filtertop depth influences VG3"
  ) +
  xlab("difference in filtertop depth (m)") +
  ylab("site-averaged difference in local VG3 (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.9, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )

if (save_figures) {
  ggsave("figures/diff_vg3_soilclass_specific.pdf", p,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```




## conclusion
Sites with heavy soil show a systematic difference between deep and shallow wells, with deep wells over-reporting the water levels. 
The effect increases with filtertop depth, and is most apparent for HG3/VG3 with a filtertop difference of more than 2m.
As expected, this effect is not universal, but site-specific: many clusters aggregate around the zero line, but some deviate from it, linked most likely to infiltration and spring areas.


In the other soil classes, except for very big differences in filtertop_depth, 
there is little difference in HG3 (or VG3) values.


# Cluster Timelines
## strategy
As with the derived XG3 values above, 
it might be instructive to check variability within clusters 
for actual water level measurements.

Ingredients, from above:

- `obswell_clusters`
- `FactPeilMeting.Niveau`


This turned out to be inefficient in `R/dbplyr`: the table is huge, and even lazy queries collects it multiple times before joining. 
SQL is far more efficient than that.

```{r}
# R - SQL helper functions
query <- function(query_string) {
  rs <- DBI::dbGetQuery(watina_conn, query_string)
  return(rs)
}

query_difflevels <- function(focal_loc, reference_loc) {
  query_string <- paste0(
    " SELECT obs1.t, ",
    "        i1, ",
    "        i2, ",
    "        l1, ",
    "        l2 ",
    " FROM (",
    "     SELECT MeetpuntWID as i1, TijdWID AS t, Niveau AS l1 ",
    "     FROM FactPeilMeting ",
    "     WHERE MeetpuntWID = ", focal_loc,
    " ) obs1 ",
    " INNER JOIN (",
    "     SELECT MeetpuntWID as i2, TijdWID AS t, Niveau AS l2 ",
    "     FROM FactPeilMeting",
    "     WHERE MeetpuntWID = ", reference_loc,
    " ) obs2 ",
    "  ON (obs1.t = obs2.t) ",
    " ;"
  )
  query(query_string)
}
```


This can be used to query all items (in `lapply`).
```{r}
cluster_metadata <- obswell_clusters %>%
  st_drop_geometry()

if (FALSE) {
  # optional: save cluster metadata for an external script
  saveRDS(cluster_metadata, file = "cluster_metadata.RDS")
  glimpse(cluster_metadata)
}

cluster_pairs <- cluster_metadata %>%
  filter(!(loc_wid == deep_ref)) %>%
  select(loc_wid, deep_ref)

progress <- utils::txtProgressBar(min = 0, max = nrow(cluster_pairs), style = 3)

get_diff_level_per_pair <- function(cluster_id) {
  utils::setTxtProgressBar(progress, cluster_id)
  loc <- cluster_pairs[cluster_id, 1]
  ref <- cluster_pairs[cluster_id, 2]

  diff_levels <- query_difflevels(loc, ref)
  return(diff_levels)
}
```



<center>
**ATTENTION**
</center>

And here we attempt to `apply` the functions.

- takes a long time
- may kill the RAM
- but works on SQL, so brace yourself and go.


```{r}
difflevel_file <- "all_diff_levels.rds"
if (FALSE || exists(difflevel_file)) {
  all_diff_levels <- lapply(
    seq(1, nrow(cluster_pairs)),
    get_diff_level_per_pair
  ) %>%
    bind_rows()

  saveRDS(all_diff_levels, file = difflevel_file)
} else {
  all_diff_levels <- readRDS(difflevel_file)
}
```


overview plot:
```{r}
all_diff_levels %>%
  mutate(diff_level = l2 - l1) %>%
  ggplot(aes(x = diff_level)) +
  geom_density() +
  xlim(-1, 1) +
  geom_vline(xintercept = 0)
```


detail plot:
```{r}
try_correlation <- function(l1, l2) {
  if (length(l1) < 10) {
    return(invisible(NA))
  }
  return(cor.test(l1, l2)$estimate)
}

mean_diffs <- all_diff_levels %>%
  mutate(
    hydroyear = floor((t - 400) / 10000),
    diff_level = l2 - l1
  ) %>%
  group_by(i1, i2) %>% # , hydroyear
  summarize(
    dl = mean(diff_level),
    corr = try_correlation(l1, l2)
  )

mean_diffs %>%
  ggplot(aes(x = dl)) +
  geom_density() +
  xlim(-1, 1) +
  geom_vline(xintercept = 0)
```

Finally, join the `mean_diffs` to location info:

```{r}
refs <- mean_diffs %>%
  pull(i2) %>%
  unique()
ref_info <- cluster_metadata %>%
  filter(loc_wid %in% refs) %>%
  select(loc_wid, x, y, filterdepth, filtertop_depth, soilclass)
names(ref_info) <- paste0("ref.", names(ref_info))

euclid <- function(x1, y1, x2, y2) {
  sqrt((x2 - x1)^2 + (y2 - y1)^2)
}

informed_diffs <- mean_diffs %>%
  left_join(cluster_metadata,
            by = join_by(i1 == loc_wid, i2 == deep_ref)) %>%
  left_join(ref_info, by = join_by(i2 == ref.loc_wid)) %>%
  mutate(
    d_ftd = ref.filtertop_depth - filtertop_depth,
    d_fd = ref.filterdepth - filterdepth,
    d_dist = euclid(x, y, ref.x, ref.y)
  )
```

Analogous to the above: check how difference depends on depth difference

```{r}
soilclass_counts <- informed_diffs %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
soilclass_colors <- c("#284971", "#8eafd7", "#f0d70f", "#cccccc")
names(soilclass_colors) <- names(soilclass_counts)

p <- informed_diffs %>%
  ggplot(aes(x = d_ftd, y = dl)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  theme_bw() +
  geom_hline(yintercept = c(0)) +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  # ggtitle(
  #   "obswell clusters: filtertop depth has limited influence on water level"
  # ) +
  ylim(-2, 1) +
  xlab("difference in filtertop depth (m)") +
  ylab("cluster-averaged difference of water level (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.85, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )


p2 <- ggExtra::ggMarginal(p, type = "histogram", binwidth = 0.05, margins = "y")
p2

if (save_figures) {
  ggsave("figures/diff_level_soilclass.png", p2,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```


Again, a soilclass-specific analysis is justified.
However, in contrast to HG3, deeper wells do not seem to over-estimate actual levels.

```{r}

p <- informed_diffs %>%
  filter(soilclass != "unknown") %>%
  ggplot(aes(x = d_ftd, y = dl)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  geom_hline(yintercept = c(0)) +
  facet_wrap(~soilclass, ncol = 2) +
  theme_bw() +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  # ggtitle(
  #   "obswell clusters: filtertop depth has limited influence on water level"
  # ) +
  ylim(-2, 1) +
  xlab("difference in filtertop depth (m)") +
  ylab("cluster-averaged difference of water level (m)") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.85, 0.2),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    ),
    text = element_text(size = 20)
  )

if (save_figures) {
  ggsave("figures/diff_level_soilclass_specific.png", p,
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```

Besides, we can check (bluntly) how much correlation there is with clusters that
have >10 shared observation days:

```{r}
informed_diffs %>%
  ggplot(aes(x = corr)) +
  geom_density()
```


```{r}
soilclass_counts <- informed_diffs %>%
  filter(!is.na(corr)) %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
soilclass_colors <- c("#284971", "#8eafd7", "#f0d70f", "#cccccc")
names(soilclass_colors) <- names(soilclass_counts)

p <- informed_diffs %>%
  filter(!is.na(corr)) %>%
  ggplot(aes(x = d_ftd, y = corr)) +
  geom_point(aes(color = soilclass)) +
  geom_smooth(color = "black") +
  theme_bw() +
  geom_hline(yintercept = c(0)) +
  scale_color_manual(values = soilclass_colors, labels = soilclass_counts) +
  ggtitle("obswell clusters: clusters are highly correlated") +
  xlab("difference in filtertop depth (m)") +
  ylab("cluster-averaged correlation coefficient (arb. units)") +
  ylim(-1, 1)


p2 <- ggExtra::ggMarginal(p, type = "histogram", binwidth = 0.05)
p2
```

Observations:

- bias might be *a tiny bit positive*, i.e. deeper filters over-estimate the water level
- yet the effect is O(5cm)



# NOTES

+ [☇] IDEA: see `ZWIP017`, query exchanges with significant change in depth; plot before/after; HG3 discrepancy water level | pressure level
+ Why do some HG3 lie below VG3?
+ Why are HG3&LG3 only calculated intermittently, whereas VG3 is mostly there?
+ What is the difference between `soilsurf_ost` and `measuringref_ost`? Can it explain the HG3 discrepancy?
+ mathjax does not work out of the box (self-contained? js to server?)
+ [☇] there is a `vg3_lcl` > 50m at `STRP048` in hydroyear 2023
+ [TODO] if you use `watina::cluster_locs()` with obswells_db (non-sf), there is a "bad error message" (indeed, it is.)
+ [TODO] there is "PeilmetingCategorie == 'twijfelachtig'" in 6016... :/
+ [TODO] there are measurements from outside Flanders: check `area_code = VBS`

TODO: contact Frederic, Jo Loos for feedback on 

 - the discrepancy in HG3 values (*versus* `FactPeilMeting.Niveau`).
 - the skyscraper-high vg3 single observation


- TODO: Gaussian fit of diff_level, possibly multilevel (involving distance, soilclass)



```{r}
DBI::dbDisconnect(watina_conn)
```
