---
title: "Sensibility of a shallow top of a piezometer filter"
author: "Floris Vanderhaeghe"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(watina)
library(dplyr)
library(dbplot)
library(n2khab)
library(stringr)
```

# Exploring distribution of the filter top

## Fetching observation well properties

In order to explore distribution of filter tops, we look at observation wells with filter no deeper than 5 m.

```{r}
watina <- connect_watina()
obswells <- get_locs(
  watina, 
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P", 
  loc_validity = c("VLD", "ENT"),
  obswells = TRUE
)
```

Structure of the data:

```{r}
glimpse(obswells)
```

How many observation wells?

```{r}
count(obswells) |> print()
```

What is the distribution of `filterlength`?

```{r}
obswells |> pull(filterlength) |> quantile(seq(0, 1, 0.05))
```

## Exploring the availability of shallow filtertops

Derive `filtertop` from `filterdepth` and `filterlength`:

```{r}
obswells <- obswells |> 
  mutate(filtertop = filterdepth + filterlength/2)
```

Explore `filtertop` values:

```{r}
filtertop_vals <- obswells |> pull(filtertop)
length(filtertop_vals)
quantile(filtertop_vals, seq(0, 1, 0.05))
quantile(filtertop_vals, 0.01)
```

How many shallow filtertops?

```{r}
filtertop_vals[filtertop_vals <= 0.3] |> table()
```

# Sensibility of shallow filtertops: what about HG3 values?

Let's have a look at how frequently HG3 values are above soil surface in piezometers with filterdepth <= 3 m (the usual limit), and how this might be related to soil texture and to the `filterdepth` or `filtertop` value.

General distribution is like this:

```{r}
get_locs(
  watina, 
  filterdepth_range = c(0, 3),
  filterdepth_guess = FALSE,
  loc_type = "P", 
  loc_validity = c("VLD", "ENT"),
  obswells = FALSE
) |> 
  get_xg3(
    watina,
    startyear = 1900,
    truncated = FALSE
  ) |> 
  dbplot_histogram(hg3_lcl)
```

Simplified texture classes from the soilmap:

```{r}
soilmap <- 
  read_soilmap() |> 
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      TRUE ~ "unknown") |> 
      factor()
    )
```


```{r}
dbDisconnect(watina)
```

