---
title: "Sensibility of a shallow top of a piezometer filter"
author: "Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(watina)
library(dplyr)
library(dbplot)
library(n2khab)
library(stringr)
library(ggplot2)

save_figures <- TRUE # whether/not to save plots in utils::mkdir("./figures")

watina_conn <- connect_watina()
```

# Key Observations (or maybe not...)

- the majority of measurement locations are equipped with piezometers ([see here](#datajoin))
- there are non-piezo wells with short filters ([see here](#filterlengths))
- a cluster of the shallowest filter tops are actually non-piezo; some are above ground ([see here](#filtertops))
- apparrently, HG3 can be positive ([see here](#hg3_dist)): this would be the case with normal rain or floods, to which `Hg3` is extra sensitive; it affects both types of measurement setups ([see here](#hg3_shallow))
- the HG3 distribution and obove-surface HG3 does not depend much on soilclass ([see here](#soilclass))


# Data Collection

## Fetching observation well properties

In order to explore distribution of filter tops, we look at observation wells with filter no deeper than 5 m.

```{r}
obswells_db <- get_locs(
  watina_conn,
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P",
  loc_validity = c("VLD", "ENT"),
  obswells = TRUE
) %>% mutate(line_nr = row_number())
```

Structure of the data:

```{r}
glimpse(obswells_db)
count(obswells_db)
```

Duplicates are replaced sites (as read from `PeilpuntStatusCode, PeilpuntStatusNaam, PeilpuntToestandCode`):
```{r}
obswells_db %>%
  group_by(loc_wid) %>%
  count() %>%
  filter(n > 4) %>%
  arrange(desc(n))
```


## OBSOLETE: Query Measurement Technique
<center>
**WARNING**
</center>
The database columns `PeilpuntTypeNaam` and `PeilpuntTypeCode` used below are not reliable. 
Their content is flawed due to the similarity of the installation type names in Dutch, and the general similarity of observation wells and piezometer installations.
It is advised to instead use the filter top depth (less than 0.3m) as a distinguishing characteristic, or analyze with filter top depth as a continuous variable.


To find out what technique was used to get the water levels, we query extra info: `PeilpuntType*`.
```{r}
pptype <- tbl(watina_conn, "vwDimPeilpunt") %>%
  filter(
    PeilpuntStatusCode %in% c("VLD", "ENT", "CLD"),
    PeilpuntOpenbaarheidTypeCode == "PLME",
    PeilpuntOpenbaarheidCode == "UNKWN"
  ) %>%
  select(
    loc_wid = MeetpuntWID,
    pp_typecode = PeilpuntTypeCode,
    pp_type = PeilpuntTypeNaam,
    obswell_installdate = PeilpuntPlaatsing
  )
```


Are there any duplicates in the `pptype`? -> yes, but they should disappear on joining

```{r}
pptype %>%
  group_by(loc_wid, obswell_installdate) %>%
  count() %>%
  filter(n > 4) %>%
  arrange(desc(n))
```

```{r}
pptype <- pptype %>% distinct()
glimpse(pptype)
count(pptype)
```


```{r}
pptype %>%
  group_by(loc_wid) %>%
  count() %>%
  filter(n > 1) %>%
  arrange(desc(n))
```


## Joining Data {#datajoin}
Then join this to the obswells data:
```{r}
count(obswells_db) %>% print()
obswells <- obswells_db %>%
  left_join(pptype, join_by(loc_wid, obswell_installdate))
glimpse(obswells)
count(obswells) %>% print()
```


Finally, it is time to collect the lazy SQL table.
Collaction is necessary to use stringr/tidyr manipulations which do not translate to SQL.

```{r}
obswells <- obswells %>% collect()
obswells <- obswells %>% mutate(pp_type_grp = word(pp_type, 1))
obswells <- obswells %>%
  mutate_at(vars(pp_type_grp), ~ tidyr::replace_na(., "onbekend"))
glimpse(obswells)
```

And a quick check what we have in the pipe:
```{r}
ggplot(obswells, aes(x = "", y = pp_type_grp, fill = pp_type_grp)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0)

type_counts <- obswells %>%
  group_by(pp_type_grp) %>%
  summarize(n = n()) %>%
  mutate(freq = 100.0 * n / sum(n))
get_frq <- function(typ) {
  sprintf(
    "%.1f %%",
    type_counts[type_counts$pp_type_grp == typ, "freq"][[1]]
  )
}
```

Indeed, most measuring stations seem to be piezometers (`r get_frq("piëzometer")`).
About `r get_frq("peilbuis")` of measurements are wells, and `r get_frq("onbekend")` are unknown.


## Technical Documentation
some locations have been replaced over time; there can have been a switch from piezo to well or the other way round.

- example: `MeetpuntWID == 9930`
- the install date is relevant for joins, otherwise successors are cross-mixed

Some have `MeetpuntWID == -1`, this seems nonsense (but should be removed by the `left_join`).


# Data Inspection
## General
How many observation wells are there?

```{r}
obswells %>%
  group_by(pp_type_grp, pp_type) %>%
  count() %>%
  print()
count(obswells) |> print()
```

## Filter Lengths {#filterlengths}
What is the **absolute** distribution of `filterlength`?

```{r}
# Histogram
ggplot(
  obswells,
  aes(x = filterlength, fill = pp_type_grp, colour = pp_type_grp)
) +
  geom_histogram(aes(y = after_stat(count)),
    binwidth = 0.05,
    colour = "black",
  ) +
  xlim(-0.01, 2.01) +
  ggtitle("filterlength usage count per installation type")


filterlengths <- obswells |> pull(filterlength)
filterlengths |> quantile(seq(0, 1, 0.05))
```


There are some preferred lengths: 0.3, 0.5, 1.0 and 2.0 meters.


What is the **relative** distribution of `filterlength` within groups?

```{r}
# Densities; normalized per group
ggplot(
  obswells,
  aes(x = filterlength, fill = pp_type_grp, colour = pp_type_grp)
) +
  geom_density(alpha = 0.8, color = "black") +
  xlim(-0.2, 2.2) +
  ggtitle("group-normalized densities of filter length")
```


And in a different view: 

```{r}
obswells %>%
  mutate(age = lubridate::year(obswell_installdate) - 1979) %>%
  filter(age > 0) %>%
  ggplot(aes(y = filterlength, x = filterdepth,
             fill = pp_type_grp, color = pp_type_grp)) +
  geom_point(alpha = 0.67, aes(size = age^2)) +
  ylim(-0.01, 2.01)

ggsave("figures/filterlengths.pdf")
```




A considerable number of non-piezo wells has short filters.


Did the types (e.g. "unknown") change over time?
```{r}
obswells %>%
  mutate(year = lubridate::year(obswell_installdate)) %>%
  ggplot(aes(x = year, fill = pp_type_grp, colour = pp_type_grp)) +
  geom_histogram(binwidth = 1, color = "black") +
  ggtitle("new installation type per year")
```

2001 was a busy year...


## Shallow Filtertops {#filtertops}

To explore the availability of shallow filtertops, we need to derive `filtertop_depth` from `filterdepth` and `filterlength`.
(See `watina/R/get.R#get_locs()` for the original calculation.)

```{r}
obswells <- obswells |>
  mutate(filtertop_depth = filterdepth - filterlength / 2)
# filterdepth is positive downwards; thus filtertop_depth must be smaller
```

Explore `filtertop_depth` values:

```{r}
filtertop_depth_vals <- obswells |> pull(filtertop_depth)
length(filtertop_depth_vals)
quantile(filtertop_depth_vals, seq(0, 1, 0.05))

ggplot(obswells, aes(x = filtertop_depth, fill = pp_type_grp)) +
  geom_histogram(aes(y = after_stat(count)),
    binwidth = 0.05,
    colour = "black"
  ) +
  geom_vline(xintercept = c(0, 0.3)) +
  scale_x_reverse() +
  coord_flip() +
  xlab("filter top depth (m; positive: down)") +
  guides(fill = guide_legend(title = "Well Type")) +
  theme(text = element_text(size = 20))
# ggtitle(
# "distribution of filter top depths per installation type"
# )
if (save_figures) {
  ggsave("figures/filtertop_depth_distribution.svg",
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```

- Conventionally, 0.3m is the criterium for "shallow".
- From a distribution standpoint, I guess 49cm is still shallow; yet I prefer to exclude the bunch of 50cm's.
- I cannot make sense of above-ground filtertops, and will remove them.

How many shallow filtertops?

```{r}
depth_classes <- obswells %>%
  mutate(
    above_ground = filtertop_depth < 0,
    below_30cm = filtertop_depth < 0.3,
    below_50cm = filtertop_depth < 0.5
  ) %>%
  count(pp_type_grp, above_ground, below_30cm, below_50cm) %>%
  tidyr::pivot_wider(values_from = n, names_from = pp_type_grp)

knitr::kable(depth_classes)
```

Noted: We have quite a number of shallow filtertop locations.
- Some are unplausible (<=0), and should be excluded.
- A cluster of the shallowest ones are actually non-piezo.


```{r}
obswells %>%
  ggplot(aes(x = pp_type_grp, y = filtertop_depth, fill = pp_type_grp)) +
  geom_violin(draw_quantiles = c(0.05, 0.25, 0.50, 0.75, 0.95), alpha = 0.6) +
  geom_point(
    position = position_jitter(seed = 1, width = 0.2),
    alpha = 0.2, fill = "white"
  ) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = 0.3) +
  scale_y_continuous(trans = "reverse")
```

```{r}
obswells %>%
  ggplot(aes(x = pp_type_grp, y = filterlength, fill = pp_type_grp)) +
  geom_point(
    position = position_jitter(seed = 1, width = 0.40),
    alpha = 0.33, fill = "white"
  )
```



# Sensibility of shallow filtertops: what about HG3 values?
## Append Data

We first need to load the `xg3` data corresponding to the selected locations.
```{r}
xg3s_db <- obswells %>%
  # filter( (filtertop_depth > 0) & (filtertop_depth < 0.5) )  %>%
  get_xg3(
    watina_conn,
    startyear = 1900,
    truncated = FALSE
  )
glimpse(xg3s_db %>% arrange(desc(hydroyear), loc_code))

xg3s <- xg3s_db %>% collect()
```

*side note:* the `Precalc` table only seems to contain the latest data for our example (`MeetpuntWID == 9930`, `loc_code == "OPHP004"`).
```{r}
xg3s %>%
  arrange(desc(hydroyear), loc_code) %>%
  filter(loc_code == "OPHP004")
```

The tricky thing is to get the "pp_type" right; it can change during the year.
```{r}
latest_pp <- obswells %>%
  arrange(desc(obswell_installdate)) %>%
  summarize(
    latest_installment = dplyr::first(obswell_installdate),
    filterdepth = dplyr::first(filterdepth),
    filtertop_depth = dplyr::first(filtertop_depth),
    pp_type = dplyr::first(pp_type),
    pp_type_grp = dplyr::first(pp_type_grp),
    .by = c(loc_wid, loc_code)
  ) %>%
  mutate(shallow = (filtertop_depth > 0) & (filtertop_depth < 0.5)) %>%
  select(-latest_installment)

glimpse(latest_pp %>% filter(loc_code == "OPHP004"))
latest_pp %>% count()
```

## HG3 distribution {#hg3_dist}
Let's have a look at how frequently HG3 values are above soil surface in piezometers with filterdepth <= 3 m (the usual limit), and how this might be related to soil texture and to the `filterdepth` or `filtertop_depth` value.

General distribution is like this:
```{r}
xg3s %>%
  ggplot(aes(x = hg3_lcl)) +
  geom_histogram(binwidth = 0.1, color = "black", fill = "white") +
  geom_vline(xintercept = 0) +
  coord_flip() +
  ggtitle("bulk HG3 distribution (water level height: positive up)")
```

*reference frame:**

- hg3 is defined from soil surface, pointing up!
- filtertop_depth is defined from surface, pointing down

```{r}
xg3_obswells <- xg3s %>%
  left_join(latest_pp, join_by(loc_code))
```


```{r}
xg3_obswells %>%
  filter(pp_type_grp != "onbekend", !is.na(hg3_lcl)) %>%
  ggplot(aes(x = hg3_lcl, fill = shallow)) +
  theme_bw() +
  geom_histogram(binwidth = 0.1, color = "black") +
  geom_vline(xintercept = 0) +
  facet_grid(. ~ pp_type_grp, scales = "free_x") +
  scale_fill_manual(values = c("#356196", "#f0d70f")) +
  coord_flip() +
  xlab("local HG3 (m)") +
  ylab("observation count") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.95, 0.1),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    )
  ) +
  ggtitle("HG3 (local), direct measurement")

if (save_figures) {
  ggsave("figures/hg3_local_details.pdf",
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```


## shallow piezos {#hg3_shallow}
```{r}
xg3_shallowpiez <- xg3_obswells %>%
  filter(pp_type_grp == "piëzometer", shallow == TRUE) # , filtertop_depth > 0)
glimpse(xg3_shallowpiez)

xg3_shallowpiez %>%
  filter(!if_any(hg3_lcl, is.na)) %>%
  ggplot(aes(x = hg3_lcl)) +
  geom_histogram(binwidth = 0.1, color = "black", fill = "white") +
  geom_vline(xintercept = 0) +
  coord_flip() +
  ggtitle("HG3 (local), only shallow piezometers")
```

Apparrently, HG3 can be positive: 

- affects both types of measurement setups (piezometers and wells)
- could be the case with floods, maybe even just intense rain and tube filling, to which `HG3` is extra sensitive


## check coordinate reference systems

The CRS of the watina data seems to match `soilmap` coordinates (see below *Belgian Lambert 72*).
According to documentation,

> DimBodem	PeilpuntXCoordinaat	Description	X Coördinaat van het peilpunt in Lambert 72
> DimBodem	PeilpuntYCoordinaat	Description	Y Coördinaat van het peilpunt in Lambert 72

Unfortunately, I fail to reproduce the documentation query.
```{r}
query_table_names <- "SELECT name FROM sys.tables ORDER BY name ASC;"
table_names <- DBI::dbGetQuery(watina_conn, query_table_names)

query_columns <- "
  SELECT *,
       OBJECT_NAME(object_id) as table_name 
  FROM sys.columns;"

column_infos <- DBI::dbGetQuery(watina_conn, query_columns) %>%
  filter(table_name == "DimPeilpunt")

column_infos %>%
  filter((name %in% c("PeilpuntXCoordinaat", "PeilpuntYCoordinaat")))

if (FALSE) {
  # the documentation could not be queried from the database
  query_documentation <- "SELECT * FROM sys.documentation;"
  documentation <- DBI::dbGetQuery(watina_conn, query_documentation)

  documentation <- DBI::dbGetQuery(watina_conn, "SELECT * FROM vwDocumentatie;")
}
```


## case study extract

Select a suitable example case for the geospatial experience below:
```{r}
obswells[grep("zwin", obswells$area_name, ignore.case = TRUE), ] %>%
  saveRDS("geospatial_example_subset.rds")

obswells %>% saveRDS("geospatial_all_data.rds")
xg3s %>% saveRDS("ecohydrological_xg3s.rds")
```



# code break!
```{r}
dbDisconnect(watina_conn)
```


*Grab a coffee, restart your R session, and continue on the code below...*

The code chunks below should mostly be de-serialized from the above.

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("n2khab")
library("sf")
library("stringr")
library("ggplot2")
library("ggridges")
library("mapview")
```

# Soil

## data loading
```{r}
obswells_sample <- readRDS("geospatial_example_subset.rds")
print(obswells_sample)
```


Simplified texture classes from the soilmap:

```{r} 
soilmap_raw <-
  read_soilmap()

soilmap <- soilmap_raw[grep(
  "zwin",
  soilmap_raw$bsm_region,
  ignore.case = TRUE
), ]

soilmap <- soilmap %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  )

glimpse(soilmap)
```

- https://r-spatial.github.io/mapview/articles/mapview_02-advanced.html

Visual inspection of the soil class for focus location:
```{r}
soilmap %>%
  mapview(
    zcol = "soilclass",
    alpha.regions = 0.6,
    map.types = c("OpenStreetMap", "OpenTopoMap"),
    alpha = 0.4
  )
```


## CRS check
The CRS of the soilmap is **BD72 / Belgian Lambert 72** (ID: EPSG 31370)
```{r}
st_crs(soilmap)
```


## inspection
Let's take a closer look at what we have.
Multipolygons:
```{r}
soilmap_light <- soilmap %>%
  filter(soilclass == "heavy")
glimpse(soilmap_light)
```


... and points:
```{r}
obswells_sample[, c("x", "y")]
```

converting points (https://stackoverflow.com/a/45484314)
```{r}
sf_points <- st_as_sf(
  x = obswells_sample,
  coords = c("x", "y"),
  crs = st_crs(soilmap_light)
)
```

This makes them plottable:
```{r}
soilmap %>%
  mapview(
    zcol = "soilclass",
    alpha.regions = 0.6,
    map.types = c("OpenStreetMap", "OpenTopoMap"),
    alpha = 0.4
  ) +
  mapview(sf_points)
```


We can get intersects:

- https://gis.stackexchange.com/a/282769
```{r}
intersects <- st_intersects(soilmap, sf_points, sparse = FALSE)
print(paste(nrow(soilmap), nrow(sf_points), nrow(intersects), ncol(intersects)))
heatmap(1 * intersects, Colv = NA, Rowv = NA)

which(intersects, arr.ind = TRUE)
```
noted: I intersect a soilmap with points; the rows are the soilmap, the cols are the points.


... yet to transfer attributes, I might want to join:

- https://gis.stackexchange.com/a/466059
```{r}
joined_spatial <- st_join(sf_points, soilmap)
glimpse(joined_spatial %>% filter(!if_any(soilclass, is.na)))
```
... here we go.


Let's check whether there's more soil info for Zwin in the full dataset.
```{r}
joined_spatial <- st_join(sf_points, soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  ))
glimpse(joined_spatial %>% filter(!if_any(soilclass, is.na)))

mapview(joined_spatial,
  zcol = "soilclass",
  map.types = c("OpenStreetMap", "OpenTopoMap"),
  alpha = 0.6
)
```

Good. I can work with this.


# All Soil and Dirt {#soilclass}
With this data example in place, I will get the soil type info to all the `obswells`.

## prepare data
This requires `obswells` and `xg3s`, as used above.
```{r}
if (!exists("obswells")) {
  obswells <- readRDS("geospatial_all_data.rds")
}
if (!exists("xg3s")) {
  xg3s <- readRDS("ecohydrological_xg3s.rds")
}

obswells_placed <- obswells %>%
  filter(!if_any(c("x", "y"), is.na))
glimpse(obswells_placed)
print(paste("Loss of locations: ", nrow(obswells), nrow(obswells_placed)))
```

Additionally, the complete soilmap will be required, with soil class.
```{r}
soilmap <- soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      .default = "unknown"
    ) %>%
      factor()
  )
glimpse(soilmap)
```

Obswells can be converted to an `sf` object (based on `x, y` location of the wells).
```{r}
obswells_points <- st_as_sf(
  x = obswells_placed,
  coords = c("x", "y"),
  crs = st_crs(soilmap)
)
```


Then, the spatial join of well and soil data will append the soil class information.
```{r}
soiled_obswells <- st_join(obswells_points, soilmap) %>%
  filter(!if_any(soilclass, is.na))
glimpse(soiled_obswells)
```


As seen above, the xg3 seems to cover the latest installations, which are to be extracted here:
```{r}
latest_soiled <- soiled_obswells %>%
  arrange(desc(obswell_installdate)) %>%
  group_by(loc_wid, loc_code) %>%
  summarize(
    latest_installment = dplyr::first(obswell_installdate),
    filterdepth = dplyr::first(filterdepth),
    filtertop_depth = dplyr::first(filtertop_depth),
    soilclass = dplyr::first(soilclass),
  ) %>%
  ungroup() %>%
  mutate(shallow = (filtertop_depth > 0) & (filtertop_depth < 0.5)) %>%
  select(-latest_installment)
```

Finally, `xg3s` can be joined.
```{r}
soilclass_obswell_xg3s <- xg3s %>%
  left_join(latest_soiled, join_by(loc_code)) %>%
  filter(!if_any(hg3_lcl, is.na)) %>%
  mutate(soilclass = tidyr::replace_na(soilclass, "unknown"))

# sample size, for display in the plots below
soilclass_counts <- soilclass_obswell_xg3s %>%
  group_by(soilclass) %>%
  count() %>%
  ungroup() %>%
  mutate(label = paste0(soilclass, " (", n, ")"))
soilclass_counts <- setNames(soilclass_counts$label, soilclass_counts$soilclass)
```


## Visualize

The question is whether the occurrence of unexpectedly high `HG3` is associated with a soil type. 

Visually, this seems not to be the case:
```{r}
soilclass_obswell_xg3s %>%
  ggplot(aes(x = hg3_lcl, y = soilclass, fill = soilclass)) +
  theme_bw() +
  geom_density_ridges(
    stat = "binline",
    binwidth = 0.05,
    scale = 2.4,
    alpha = 0.6,
    color = "black"
  ) +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = c("#284971", "#8eafd7", "#f0d70f", "#b2b2b2")) +
  xlim(-2, 1) +
  scale_y_discrete(labels = soilclass_counts) +
  coord_flip() +
  xlab("local HG3 (m)") +
  ylab("observation count") +
  theme(
    legend.position = "inside", legend.position.inside = c(0.95, 0.1),
    legend.background = element_rect(
      fill = "white",
      linewidth = 0.5,
      linetype = "solid",
      colour = "lightgray"
    )
  ) +
  ggtitle("HG3 (local), direct measurement")

if (save_figures) {
  ggsave("figures/hg3_soilclass_details.pdf",
    dpi = 300, width = 16, height = 8, units = "in"
  )
}
```
## Statistical Tests
... do not make sense because

- data is not normally distributed
- sample size is unequal
- sample size is large (i.e. conventional tests will trigger upon miniscule differences).

```{r}
get_group <- function(grp) {
  x <- soilclass_obswell_xg3s %>%
    filter(soilclass == grp) %>%
    select("hg3_lcl")
  x[[1]]
}
ks.test(get_group("light"), get_group("heavy"))
ks.test(get_group("light"), get_group("peat"))

wilcox.test(get_group("light"), get_group("heavy"))
wilcox.test(get_group("light"), get_group("peat"))
```
↑ **ignore the p-values!**

What I would rather like to do is a model with an adequate distribution for HG3 which incorporates (and quantifies) an effect of `soilclass`.

My skill in R is (yet) insufficient to tackle this.
But I will learn.




