---
title: "Sensibility of a shallow top of a piezometer filter"
author: "Floris Vanderhaeghe, Falk Mielke"
date: "`r lubridate::today()`"
output:
  html_document:
    keep_md: TRUE
    number_sections: yes
    code_folding: show
    code_download: yes
    fig_width: 10
    fig_height: 5
    fig_caption: yes
    df_print: paged
    toc: TRUE
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
    includes:
        in_header: ../../header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(watina)
library(dplyr)
library(dbplot)
library(n2khab)
library(stringr)
library(ggplot2)


watina_conn <- connect_watina()
```

# Key Observations (or maybe not...)

- the majority of measurement locations are equipped with piezometers ([see here](#datajoin))
- there are non-piezo wells with short filters ([see here](#filterlengths))
- a cluster of the shallowest filter tops are actually non-piezo; some are above ground ([see here](#filtertops))
- apparrently, HG3 can be positive ([see here](#hg3_dist)): this would be the case with floods, to which `Hg3` is extra sensitive; it affects both types of measurement setups ([see here](#hg3_shallow))


# Data Collection

## Fetching observation well properties

In order to explore distribution of filter tops, we look at observation wells with filter no deeper than 5 m.

```{r}
obswells_db <- get_locs(
  watina_conn, 
  filterdepth_range = c(0, 5),
  filterdepth_guess = FALSE,
  loc_type = "P", 
  loc_validity = c("VLD", "ENT"),
  obswells = TRUE
) %>% mutate(line_nr = row_number())
```

Structure of the data:

```{r}
glimpse(obswells_db)
count(obswells_db) |> print()
obswells_db %>% group_by(loc_wid) %>% count %>% filter(n>4) %>% arrange(desc(n)) %>% print()
```
(duplicates are replaced sites, as read from `PeilpuntStatusCode, PeilpuntStatusNaam, PeilpuntToestandCode`)


## Extra: Query Measurement Technique
To find out what technique was used to get the water levels, we query extra info: `PeilpuntType*`.
```{r}
# pptype <- tbl(watina_conn, "vwDimPeilpunt") %>% collect
# write.csv2(pptype, "pptype.csv")
pptype <- tbl(watina_conn, "vwDimPeilpunt") %>%
          filter(PeilpuntStatusCode %in% c("VLD", "ENT", "CLD"),
                 PeilpuntOpenbaarheidTypeCode == "PLME",
                 PeilpuntOpenbaarheidCode == "UNKWN") %>%
          select(loc_wid = MeetpuntWID,
                 pp_typecode = PeilpuntTypeCode,
                 pp_type = PeilpuntTypeNaam,
                 obswell_installdate = PeilpuntPlaatsing
                 # fd2 = PeilbuisLengte - ReferentieNiveauMaaiveld - (FilterLengte / 2)
          ) 
    # collect %>%
    # mutate_if(is.character, as.factor)
# glimpse(pptype)
```

Are there any duplicates in the pptype? -> yes, but they should disappear on filtering
```{r}
pptype %>% group_by(loc_wid, obswell_installdate) %>% count %>% filter(n>4) %>% arrange(desc(n)) %>% print()
# pptype %>% filter(loc_wid == -1) %>% print()
# pptype %>% filter(loc_wid == 10385) %>% print()
pptype <- pptype %>% distinct()
glimpse(pptype)
count(pptype) %>% print()
pptype %>% group_by(loc_wid) %>% count %>% filter(n>1) %>% arrange(desc(n)) %>% print()
```


## Joining Data {#datajoin}
Then join this to the obswells data:
```{r}
count(obswells_db) %>% print()
obswells <- obswells_db %>% left_join(pptype, by = c("loc_wid", "obswell_installdate"))
glimpse(obswells)
count(obswells) %>% print()
# write.csv2(obswells, "obswells.csv")
```

Finally, it is time to collect the lazy SQL table.
Collaction is necessary to use stringr/tidyr manipulations which do not translate to SQL.

```{r}
obswells <- obswells %>% collect
obswells <- obswells %>% mutate(pp_type_grp = word(pp_type, 1))
obswells <- obswells %>% mutate_at(vars(pp_type_grp), ~tidyr::replace_na(., "onbekend"))
glimpse(obswells)
```

And a quick check what we have in the pipe:
```{r}
ggplot(obswells, aes(x="", y=pp_type_grp, fill=pp_type_grp)) + 
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0)

type_counts <- obswells %>% group_by(pp_type_grp) %>% summarize(n=n()) %>% mutate(freq = 100.0 * n / sum(n))
get_frq <- function (typ) sprintf("%.1f %%", type_counts[type_counts$pp_type_grp == typ, "freq"][[1]])
```

Indeed, most measuring stations seem to be piezometers (`r get_frq("piëzometer")`).
About `r get_frq("peilbuis")` of measurements are wells, and `r get_frq("onbekend")` are unknown.


## Technical Documentation
some locations have been replaced over time; there can have been a switch from piezo to well or the other way round.
- example: `MeetpuntWID == 9930`
- the install date is relevant for joins, otherwise successors are cross-mixed

Some have `MeetpuntWID == -1`, this seems nonsense (but should be removed by the `left_join`).


# Data Inspection
## General
How many observation wells are there?

```{r}
obswells %>% group_by(pp_type_grp, pp_type) %>% count %>% print()
count(obswells) |> print()
```

## Filter Lengths {#filterlengths}
What is the **absolute** distribution of `filterlength`?

```{r}
# Histogram of Densities
ggplot(obswells,
       aes(x=filterlength, fill = pp_type_grp, colour = pp_type_grp)) + 
    geom_histogram(aes(y=after_stat(count)),      # histogram with density instead of count on y-axis
                   binwidth = 0.05,
                   colour="black",
                   ) +
    xlim(-0.2, 2.2) +
    ggtitle("filterlength usage count per installation type")


filterlengths <- obswells |> pull(filterlength)
filterlengths |> quantile(seq(0, 1, 0.05))
```

There are some preferred lengths: 0.3, 0.5, 1.0 and 2.0 meters.


What is the **relative** distribution of `filterlength` within groups?

```{r}
# Densities; normalized per group
ggplot(obswells,
       aes(x=filterlength, fill = pp_type_grp, colour = pp_type_grp)) + 
    geom_density(alpha = 0.8, color = "black") +
    xlim(-0.2, 2.2) +
    ggtitle("group-normalized densities of filter length")
```


A considerable number of non-piezo wells has short filters.


Did the types (e.g. "unknown") change over time?
```{r}
obswells %>% mutate(year = lubridate::year(obswell_installdate)) %>%
    ggplot(aes(x=year, fill=pp_type_grp, colour=pp_type_grp)) +
    geom_histogram(binwidth = 1, color = "black") +
    ggtitle("new installation type per year")
```

2001 was a busy year...


## Shallow Filtertops {#filtertops}

To explore the availability of shallow filtertops, we need to derive `filtertop` from `filterdepth` and `filterlength`.
(See `watina/R/get.R#get_locs()` for the original calculation.)

```{r}
obswells <- obswells |> 
  mutate(filtertop = filterdepth - filterlength/2) # filterdepth is positive downwards; thus filtertop must be smaller
```

Explore `filtertop` values:

```{r}
filtertop_vals <- obswells |> pull(filtertop)
length(filtertop_vals)
quantile(filtertop_vals, seq(0, 1, 0.05))

ggplot(obswells, aes(x=filtertop, fill = pp_type_grp)) + 
    geom_histogram(aes(y=after_stat(count)),
                   binwidth = 0.05, 
                   colour="black") +
    geom_vline(xintercept = c(0, 0.5)) +
    scale_x_reverse() +
    coord_flip() +
    ggtitle("distribution of filter top depth (positive: down) per installation type")
```

- Conventionally, 0.3m is the criterium for "shallow".
- From a distribution standpoint, I guess 49cm is still shallow; yet I prefer to exclude the bunch of 50cm's.
- I cannot make sense of above-ground filtertops, and will remove them.

How many shallow filtertops?

```{r}
sum(filtertop_vals[(filtertop_vals < 0.5) & (filtertop_vals > 0.0)] |> table())
```

Noted: We have quite a number of shallow filtertop locations.
- Some are unplausible (<=0), and should be excluded.
- A cluster of the shallowest ones are actually non-piezo.


# Sensibility of shallow filtertops: what about HG3 values?
## Append Data

We first need to load the `xg3` data corresponding to the selected locations.
```{r}
xg3s_db <- obswells %>%
    # filter( (filtertop > 0) & (filtertop < 0.5) )  %>% 
  get_xg3(
    watina_conn,
    startyear = 1900,
    truncated = FALSE
  )
glimpse(xg3s_db %>% arrange(desc(hydroyear), loc_code))

xg3s <- xg3s_db %>% collect
```

*side note:* the `Precalc` table only seems to contain the latest data for our example (`MeetpuntWID == 9930`, `loc_code == "OPHP004"`).
```{r}
xg3s %>% arrange(desc(hydroyear), loc_code) %>%
    filter(loc_code == "OPHP004")
```

The tricky thing is to get the "pp_type" right; it can change during the year.
```{r}
latest_pp <- obswells %>% arrange(desc(obswell_installdate)) %>%
    group_by(loc_wid, loc_code) %>%
    summarize(latest_installment = dplyr::first(obswell_installdate),
              filterdepth = dplyr::first(filterdepth),
              filtertop = dplyr::first(filtertop),
              pp_type = dplyr::first(pp_type),
              pp_type_grp = dplyr::first(pp_type_grp)
              ) %>% ungroup %>%
    mutate(shallow = (filtertop > 0) & (filtertop < 0.5)) %>%
    select(-latest_installment)

glimpse(latest_pp %>% filter(loc_code == "OPHP004"))
latest_pp %>% count
```

## HG3 distribution {#hg3_dist}
Let's have a look at how frequently HG3 values are above soil surface in piezometers with filterdepth <= 3 m (the usual limit), and how this might be related to soil texture and to the `filterdepth` or `filtertop` value.

General distribution is like this:
```{r}
xg3s %>% 
    ggplot(aes(x = hg3_lcl)) +
    geom_histogram(binwidth = 0.1, color = "black", fill = "white") +
    geom_vline(xintercept = 0) +
    coord_flip() +
    ggtitle("bulk HG3 distribution (water level height: positive up)")
```

*reference frame:**
- hg3 is defined from soil surface, pointing up!
- filtertop is defined from surface, pointing down

```{r}
xg3_obswells <- xg3s %>%
    left_join(latest_pp, by = c(loc_code = "loc_code"))
```


```{r}
xg3_obswells %>%
    filter(!(pp_type_grp == "onbekend"), !if_any(hg3_lcl, is.na)) %>% 
    ggplot(aes(x = hg3_lcl, fill = shallow)) +
    theme_bw() +
    geom_histogram(binwidth = 0.1, color = "black") +
    geom_vline(xintercept = 0) +
    facet_grid(. ~ pp_type_grp, scales = "free_x") +
    scale_fill_manual(values=c("#356196", "#f0d70f")) +
    coord_flip() +
    xlab("local HG3 (m)") + ylab("observation count") +
    theme(legend.position = "inside", legend.position.inside = c(0.95, 0.1),
          legend.background = element_rect(fill="white", linewidth=0.5, linetype="solid", colour ="lightgray")
          ) +
    ggtitle("HG3 (local), direct measurement")

# ggsave("hg3_local_details.pdf", dpi = 300, width = 16, height = 8, units = "in")
```


## shallow piezos {#hg3_shallow}
```{r}
xg3_shallowpiez <- xg3_obswells %>%
    filter(pp_type_grp == "piëzometer", shallow == TRUE)#, filtertop > 0) 
glimpse(xg3_shallowpiez)

xg3_shallowpiez %>%
    filter(!if_any(hg3_lcl, is.na)) %>% 
    ggplot(aes(x = hg3_lcl)) +
    geom_histogram(binwidth = 0.1, color = "black", fill = "white") +
    geom_vline(xintercept = 0) +
    coord_flip() +
    ggtitle("HG3 (local), only shallow piezometers")
```

Apparrently, HG3 can be positive: 
- affects both types of measurement setups (piezometers and wells)
- could be the case with floods, maybe even just intense rain and tube filling, to which `HG3` is extra sensitive


## case study extract

Select a suitable example case for the geospatial experience below:
```{r}
obswells[grep("zwin", obswells$area_name, ignore.case = TRUE),] %>% saveRDS("geospatial_example_subset.rds")
```


```{r}
dbDisconnect(watina_conn)
```


# code break!
*Grab a coffee, restart your R session, and continue on the code below...*

The code chunks below should mostly be de-serialized from the above.

```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("dbplot")
library("n2khab")
library("sf")
library("stringr")
library("ggplot2")
library("mapview")

# options(browser = "librewolf")
```

# Soil

## data loading
```{r}
obswells_sample <- readRDS("geospatial_example_subset.rds")
print(obswells_sample)
```


Simplified texture classes from the soilmap:

```{r} 
soilmap_raw <- 
  read_soilmap()
soilmap <- soilmap_raw[grep("zwin", soilmap_raw$bsm_region, ignore.case = TRUE),] 
soilmap <- soilmap %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      TRUE ~ "unknown") |> 
      factor()
    )

glimpse(soilmap)
```

- https://r-spatial.github.io/mapview/articles/mapview_02-advanced.html
This does not work:
```{r}
soilmap %>% 
    mapview(
        zcol = "soilclass",
        alpha.region = 0.6,
        map.types = c("OpenStreetMap", "OpenTopoMap"),
        alpha = 0.4
    )
```

## inspection
Let's take a closer look at what we have.
Multipolygons:
```{r}
soilmap_light <- soilmap  %>%
    filter(soilclass == 'heavy')
glimpse(soilmap_light)
```


... and points:
```{r}
obswells_sample[, c('x','y')]
```

converting points (https://stackoverflow.com/a/45484314)
```{r}
sf_points <- st_as_sf(x = obswells_sample, 
                        coords = c("x", "y"), 
                      crs = st_crs(soilmap_light)
                      )
```

This makes them plottable:
```{r}
soilmap %>% 
    mapview(
        zcol = "soilclass",
        alpha.region = 0.6,
        map.types = c("OpenStreetMap", "OpenTopoMap"),
        alpha = 0.4
    ) +
    mapview(sf_points) 

```


We can get intersects:
- https://gis.stackexchange.com/a/282769
```{r}
intersects <- st_intersects(soilmap, sf_points, sparse = FALSE)
print(paste(nrow(soilmap), nrow(sf_points), nrow(intersects), ncol(intersects)))
heatmap(1*intersects, Colv = NA, Rowv = NA)

which(intersects, arr.ind = TRUE)

```
noted: I intersect a soilmap with points; the rows are the soilmap, the cols are the points.


... yet to transfer attributes, I might want to join:
- https://gis.stackexchange.com/a/466059
```{r}
joined_spatial <- st_join(sf_points, soilmap)
glimpse(joined_spatial %>% filter(!if_any(soilclass, is.na)))
```
... here we go.


Let's check whether there's more soil info for Zwin in the full dataset.
```{r}
joined_spatial <- st_join(sf_points, soilmap_raw %>%
  transmute(
    soilclass = case_when(
      str_detect(bsm_mo_tex, "V") ~ "peat",
      str_detect(bsm_mo_tex, "^U") ~ "heavy",
      str_detect(bsm_mo_tex, "[SZPLX]") ~ "light",
      str_detect(bsm_mo_tex, ".+") ~ "heavy",
      TRUE ~ "unknown") |> 
      factor()
    )
)
glimpse(joined_spatial %>% filter(!if_any(soilclass, is.na)))

mapview(joined_spatial, zcol = "soilclass",
        map.types = c("OpenStreetMap", "OpenTopoMap"),
        alpha = 0.6
        )
```
Good. I can work with this.
